### 2019-01-21 article.tex
Start

\documentclass{article}

% Bibliography
\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{;}

% Use 'It was found that A is B (Name 1234)' style
\setcitestyle{authoryear,open={},close={}}

% Affiliations
\usepackage{authblk}
\title{pirouette: the error BEAST2 makes in inferring a phylogeny}
\author[1]{Rich\`el J.C. Bilderbeek}
\author[1]{Giovanni Laudanno}
\author[1]{Rampal S. Etienne}
\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands}

% Use double spacing
\usepackage{setspace}
\doublespacing

\usepackage{listings}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage{pgf}
\usepackage{bm}

% Style of listings
% From http://r.789695.n4.nabble.com/How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
\usepackage{fancyvrb} 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolor}{rgb}{0.95,0.95,0.92}
\lstdefinestyle{mystyle}{
  language=R,% set programming language
  basicstyle=\ttfamily\small,% basic font style
  commentstyle=\color{gray},% comment style
  % numbers=left,% display line numbers on the left side
  numberstyle=\scriptsize,% use small line numbers
  numbersep=10pt,% space between line numbers and code
  tabsize=2,% sizes of tabs
  showstringspaces=false,% do not replace spaces in strings by a certain character
  captionpos=b,% positioning of the caption below
  breaklines=true,% automatic line breaking
  escapeinside={(*}{*)},% escaping to LaTeX
  fancyvrb=true,% verbatim code is typset by listings
  extendedchars=false,% prohibit extended chars (chars of codes 128--255)
  alsoletter={.<-},% becomes a letter
  alsoother={$},% becomes other
  otherkeywords={!=, ~, $, \&, \%/\%, \%*\%, \%\%, <-, <<-, /},% other keywords
  deletekeywords={c}% remove keywords 
}
\lstset{style=mystyle}

% Adds numbered lines
\usepackage{lineno}
\linenumbers

% Rename 'Abstract' to 'Summary 
\usepackage[english]{babel}
\addto{\captionsenglish}{\renewcommand{\abstractname}{Summary}}

\begin{document}

\maketitle

\begin{abstract}

  \textbf{1. }
    BEAST2 is a popular Bayesian phylogenetics software tool,
    that takes an alignment and inference model to create a
    posterior of jointly-estimated phylogenies and model parameter estimates.
    When a new macro-evolutionary speciation model is developed,
    a good first step is to measure the error BEAST2 makes on known
    phylogeny derived from the new speciation process, 
    with its current set of inference models. \\
  \textbf{2. }
    Here, we present a free, libre and open-source R package, \verb;pirouette;
    that assesses the inference error BEAST2 makes based on a known/true 
    phylogeny. \\
  \textbf{3. }
    We describe \verb;pirouette;'s usage and the biological scientific
    question it can answer, including full examples. \\
  \textbf{4. }
    As \verb;pirouette; is designed to be of high quality and extendable, 
    we conclude by describing the further development of the package. \\
\end{abstract}

{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, babette, R

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Phylogenies are commonly used to explore evolutionary hypotheses.
%Not only can phylogenies show us how species (or other
%evolutionary units) are related to each other, 
%but we can also estimate relevant parameters such as extinction and 
%speciation rates from them.
%There are many phylogenetics tools available to obtain an estimate 
%of the phylogeny of a given set of species. 
%BEAST2 (\cite{bouckaert2014beast}) is one of the most widely used ones.
%It uses a Bayesian statistical framework to estimate 
%the joint posterior distribution of phylogenies and model parameters, 
%from one or more DNA, RNA or amino acid alignments (see figure 1 
%for an overview of the workflow).

%Here, we present \verb;pirouette;:
%’BEAUti 2, BEAST2 and Tracer for R’, 
%which creates BEAST2 (v.2.4.7) configuration files,
%runs BEAST2, and analyzes its results,
%all from an R function call. This
%will save time, tedious mouse clicking and 
%reduces the chances of errors in such repetitive actions.
%The interface of \verb;babette; mimics the tools it
%is based on. This
%familiarity helps both beginner and experienced BEAST2 users 
%to make the step from those tools to \verb;babette;.
%\verb;babette; enables the creation of a single-script 
%pipeline from sequence alignments to posterior analysis in R.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Description}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\verb;babette; is written in the R programming language (\cite{R})
%and enables the full BEAST2 workflow from a single R function call,
%in a similar way to what subsequent usage of BEAUti, DensiTree and Tracer would produce.
%\verb;babette;'s main function is \verb;bbt_run;, which
%configures BEAST2, runs it and parses its output. 
%\verb;bbt_run; needs at least the name of a 
%FASTA file containing a DNA alignment. 
%The default settings for the other arguments of \verb;bbt_run; 
%are identical to BEAUti's and BEAST2's default settings.
%Per alignment, a site model, clock model and tree prior can be chosen.
%Multiple alignments can be used, each with its own (unlinked) site model, 
%clock model and tree prior.

%\verb;babette; currently has 108 exported functions to set up  
%a BEAST2 configuration file. 
%\verb;babette; can currently handle the majority of BEAUti use cases.
%Because of BEAUti's high number of plugins, 
%\verb;babette; uses a software architecture that is designed to be extended.
%Furthermore, \verb;babette; has 13 exported functions to run and help run BEAST2.
%One function is used to run BEAST2, another
%one installs BEAST2 to a default location.
%Finally, \verb;babette; has 21 exported function to parse the BEAST2 output
%files and analyze the created posterior. \verb;babette; gives the
%same ESSes and summary statistics as Tracer. The data is formatted
%such that it can easily be visualized using \verb;ggplot2; (for a trace,
%similar to Tracer) or \verb;phangorn; (\cite{phangorn}) (for 
%the phylogenies in a posterior, similar to DensiTree).

%Currently, \verb;babette; does not contain all functionality 
%in BEAUti, BEAST2 and their many plug-ins, 
%because these tools themselves also change in time.
%\verb;babette; currently works only on DNA data,
%because this is the most common use case. 
%Nevertheless, \verb;babette; provides the majority of default tree priors and supports the most important command-line arguments of BEAST2, provides
%the core Tracer analysis options, and has the most basic subset of 
%plotting options of DensiTree. Up till now, the \verb;babette; features 
%implemented are those requested by users. Further extension of \verb;babette; 
%will be based on future user requests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Installation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\verb;piouette; can be installed easily from CRAN:

\begin{lstlisting}[language=R, floatplacement=H]
install.packages("piouette")
\end{lstlisting}

For the most up-to-date version, 
one can download and install the package from \verb;pirouette;'s GitHub repository:

\begin{lstlisting}[language=R, floatplacement=H]
usthis::install_github("richelbilderbeek/pirouette")
\end{lstlisting}
To start using \verb;pirouette;, load its functions in the global namespace first:

\begin{lstlisting}[language=R, floatplacement=H]
library(pirouette)
\end{lstlisting}
Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
This can be done from within R, using:

\begin{lstlisting}[language=R, floatplacement=H]
install_beast2()
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

A first research question that

\begin{lstlisting}[language=R, floatplacement=H]
phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4):1);")
ape::plot.phylo(phylogeny, main = "The True Phylogeny")
\end{lstlisting}

This code will create a (temporary) BEAST2 configuration file,
from the FASTA file with name \verb;anthus_aco.fas; (which
is supplied with the package, from \cite{VanEls2018}), 
using the same default settings as BEAUti, which are, 
among others, a Jukes-Cantor site model, a strict clock, and a Yule birth tree prior.
\verb;babette; will then execute BEAST2 using that file, and
parses the output. The returned data structure, named \verb;out;, 
is a list of parameter estimates (called \verb;estimates;), posterior 
phylogenies (called \verb;anthus_aco_trees;, named after
the alignment's name) and MCMC operator performance (\verb;operators;).
An example of using a different site model, clock model 
and tree prior is:

\begin{lstlisting}[language=R, floatplacement=H]
out <- bbt_run(
  fasta_filenames = "anthus_aco.fas",
  site_models = create_hky_site_model(),
  clock_models = create_rln_clock_model(),
  tree_priors = create_bd_tree_prior()
)
\end{lstlisting}
This code uses an HKY site model, a relaxed log-normal clock model and a 
birth-death tree prior, each with their default settings in BEAUti.
Table \ref{tab:functions} shows an overview of all functions to 
create site models, clock models and tree priors.
Note that the arguments' names \verb;site_models;, \verb;clock_models; 
and \verb;tree_priors; are plural, as each of these
can be (a list of) one or more elements. Each of these arguments must 
have the same number of elements, so that each alignment has its
own site model, clock model and tree prior. 
An example of two alignments, each with its own site model, is:

\begin{lstlisting}[language=R, floatplacement=H]
out <- bbt_run(
  fasta_filenames = c(
    "anthus_aco.fas", 
    "anthus_nd2.fas"
  ),
  site_models = list(
    create_tn93_site_model(), 
    create_gtr_site_model()
  )
)
\end{lstlisting}
\verb;babette; also uses the same default prior distributions as BEAUti 
for each of the site models, clock models and tree priors. 
For example, by default, a Yule tree prior assumes that the birth rate 
follows a uniform distribution, 
from minus infinity to plus infinity. 
One may prefer a different ddistribution instead. 
Here is an example how to specify an exponential distribution for
the birth rate in a Yule tree prior in \verb;babette;:

\begin{lstlisting}[language=R, floatplacement=H]
out <- bbt_run(
  fasta_filenames = "anthus_aco.fas",
  tree_priors = create_yule_tree_prior(
    birth_rate_distr = create_exp_distr()    
  )
)
\end{lstlisting}
In this same example, one may specify
the initial shape parameters of the exponential distribution.
In BEAST2's implementation, an exponential distribution 
has one shape parameter: its mean, which can be set to any
value with \verb;BEAUti;. To set the 
mean value of the exponential distribution to a 
fixed (non-estimated) value, do:

\begin{lstlisting}[language=R, floatplacement=H]
out <- bbt_run(
  fasta_filenames = "anthus_aco.fas",
  tree_priors = create_yule_tree_prior(
    birth_rate_distr = create_exp_distr(
      mean = create_mean_param(
        value = 1.0, 
        estimate = FALSE
      )
    )    
  )
)
\end{lstlisting}
\verb;babette; also supports node dating. Like BEAUti, one
can specify Most Recent Common Ancestor ('MRCA') priors.
An MRCA prior allows to specify taxa having a common ancestor,
including a distribution for the date of that ancestor.
With \verb;babette;, this is achieved as follows:

\begin{lstlisting}[language=R, floatplacement=H]
out <- bbt_run(
  fasta_filenames = "anthus_aco.fas",
  mrca_priors = create_mrca_prior(
    taxa_names = sample(get_taxa_names("anthus_aco.fas"), size = 2),
    alignment_id = get_alignment_id("anthus_aco.fas"),
    is_monophyletic = TRUE,
    mrca_distr = create_normal_distr(
      mean = create_mean_param(value = 15.0, estimate = FALSE),
      sigma = create_sigma_param(value = 0.025, estimate = FALSE)
    )
  )
)
\end{lstlisting}
Instead of dating the ancestor of two random taxa, any subset of taxa can be selected,
and multiple sets are allowed.
\verb;babette; allows for the same core functionality as Tracer to show the values of the parameter estimates sampled
in the BEAST2 run. This is called the "trace" (hence the name).
The start of the trace, called the "burn-in", is usually discarded, as an MCMC 
algorithm (such as used by BEAST2) first has to converge to
its equilibrium and hence the parameter estimates are not 
representative. By default, Tracer discards the first 10\% of all 
the parameter estimates. 
To remove a 20\% burn-in from all parameter estimates 
in \verb;babette;, the following code can be used:

\begin{lstlisting}[language=R, floatplacement=H]
traces <- remove_burn_ins(
  traces = out$estimates, 
  burn_in_fraction = 0.2
)
\end{lstlisting}
Tracer shows the ESSes of each posterior's variables.
These ESSes are important to determine the strength of the
inference. As a rule of thumb, an ESS of 200 is acceptable 
for any parameter estimate.
To calculate the effective sample sizes (of all estimated variables) in \verb;babette;:

\begin{lstlisting}[language=R, floatplacement=H]
esses <- calc_esses(
  traces = traces, 
  sample_interval = 1000
)
\end{lstlisting}
Tracer displays multiple summary statistics for each
estimated variable: the mean and its standard error, standard deviation,
variance, median, mode, geometric mean, 95\% highest posterior density interval, 
auto-correlation time and effective sample size. It displays these statistics per
variable. In \verb;babette;, these summary statistics are collected for
all estimated parameters at once:

\begin{lstlisting}[language=R, floatplacement=H]
sum_stats <- calc_summary_stats(
  traces = traces, 
  sample_interval = 1000
)
\end{lstlisting}
\verb;babette; allows for the same functionality as DensiTree.
DensiTree displays the phylogenies in a posterior at the same
time scale, drawn one over one another, allowing to see the uncertainty in
topology and branch lengths. 
The posterior phylogenies are stored as \verb;anthus_aco_trees; in the object \verb;out;,
and can be plotted as follows:

\begin{lstlisting}[language=R, floatplacement=H]
plot_densitree(phylos = out$anthus_aco_trees)
\end{lstlisting}
Instead of running the full pipeline, 
\verb;babette; also allows to only create a BEAST2 configuration file.
To create a BEAST2 configuration file, with all settings to default, use:

\begin{lstlisting}[language=R, floatplacement=H]
create_beast2_input_file(
  input_filenames = babette::get_babette_path("anthus_aco.fas"),
  output_filename = "beast2.xml"
)
\end{lstlisting}
This file can then be loaded and edited by BEAUti, 
run by BEAST2, or run by \verb;babette;:

\begin{lstlisting}[language=R, floatplacement=H]
run_beast2(
  input_filename = "beast2.xml",
  output_log_filename = "run.log",
  output_trees_filenames = "posterior.trees",
  output_state_filename = "final.xml.state"
)
\end{lstlisting}
\verb;run_beast2; is a function that only runs BEAST2, 
and does not parse the output files (unlike \verb;bbt_run;). 
In the example above, 
we specify the names of the desired BEAST2 
output files explicitly, and these will be created in the
R working directory, after which they can be inspected 
with other tools, or used to continue a BEAST2 run.
When the names of these files are not specified, 
both \verb;bbt_run; and \verb;run_beast2;
put these files in the default temporary folder (as
obtained from \verb;temp.dir();) to keep
the working directory clean of intermediate files.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{babette resources}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\verb;babette; is free, libre and open source software available at 
\url{http://github.com/richelbilderbeek/babette}
and is licensed under the GNU General Public License v3.0.
\verb;babette; uses the Travis CI (\url{https://travis-ci.org})
continuous integration service, which is known to significantly 
increase the number of bugs exposed (\cite{vasilescu2015}) and increases
the speed at which new features are added (\cite{vasilescu2015}).
\verb;babette; has a 100\% code coverage, which correlates with 
code quality (\cite{horgan1994,del1995correlation}). 
\verb;babette; follows Hadley Wickham's style guide (\cite{style_guide}), 
which improves software quality (\cite{fang2001}).
\verb;babette; depends on multiple packages, which are 
\verb;ape; (\cite{APE}), 
\verb;beautier; (\cite{beautier}),
\verb;beastier; (\cite{beastier}),
\verb;devtools; (\cite{devtools}),
\verb;geiger; (\cite{GEIGER}),
\verb;ggplot2; (\cite{ggplot2}),
\verb;knitr; (\cite{knitr}),
\verb;phangorn; (\cite{phangorn}),
\verb;rmarkdown; (\cite{rmarkdown}),
\verb;seqinr; (\cite{seqinr}),
\verb;stringr; (\cite{stringr}),
\verb;testit; (\cite{testit}) and 
\verb;tracerer; (\cite{tracerer}).
We tested \verb;babette; to give a clean 
error message for incorrect input, by
calling \verb;babette; one million times
with random or random sensible inputs, 
using a high performance computer cluster. 
The test scripts are supplied with \verb;babette;.

\verb;babette;'s development takes place on GitHub,
\url{https://github.com/richelbilderbeek/babette}, 
which accommodates collaboration (\cite{perez2016ten}) 
and improves transparency (\cite{gorgolewski2016practical}).
\verb;babette;'s GitHub facilitates feature requests and 
has guidelines how to do so.

\verb;babette;'s documentation is extensive. All functions are documented
in the package's internal documentation. For quick use, 
each exported function shows a minimal example. 
For easy exploration, each exported function's documentation links to related functions.
Additionally, \verb;babette; has a vignette that demonstrates extensively how
to use it. There is documentation on the GitHub to get started, 
with a dozen examples of BEAUti screenshots with equivalent \verb;babette; code.
Finally, \verb;babette; has tutorial videos that can 
be downloaded or viewed on YouTube, \url{https://goo.gl/weKaaU}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Citation of babette}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Scientists using \verb;babette; in a published paper can cite this
article, and/or cite the \verb;babette; package 
directly. To obtain this citation from within an R script, use:

\begin{lstlisting}[language=R]
> citation("babette")
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Thanks to Yacine Ben Chehida and Paul van Els for supplying their 
BEAST2 use cases. Thanks again to Paul van Els for sharing his FASTA files 
for use by this package. Thanks to Leonel Herrera-Alsina, Raphael Scherrer 
and Giovanni Laudanno for their comments on this package and article.
Thanks to Huw Ogilvie, Michael Matschiner and one anonymous reviewer for reviewing this article.
Thanks to rOpenSci, and especially Noam Ross and Guangchuang Yu for reviewing the package's source code.
We would like to thank the Center for Information Technology of the University 
of Groningen for their support and for providing access to the Peregrine 
high performance computing cluster. We thank the Netherlands 
Organization for Scientific Research (NWO) for financial support 
through a VICI grant awarded to RSE.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Accessibility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

All code is archived at \url{http://github.com/richelbilderbeek/babette_article},
with DOI \url{https://doi.org/10.5281/zenodo.1251203}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Authors' contributions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

RJCB and RSE conceived the idea for the package. RJCB created and tested the package, and wrote the first draft of the manuscript. RSE contributed substantially to revisions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{mee}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
\begin{tabular}{ | l | l | }
\hline
\textbf{Name} & \textbf{Description} \\
\hline
\verb;bbt_run; & Run BEAST2 \\
\hline
\verb;create_gtr_site_model; & Create a GTR site model \\
\verb;create_hky_site_model; & Create an HKY site model \\
\verb;create_jc69_site_model; & Create a Jukes-Cantor site model \\
\verb;create_tn93_site_model; & Create a TN93 site model \\
\hline
\verb;create_rln_clock_model; & Create a relaxed log-normal clock model \\
\verb;create_strict_clock_model; & Create a strict clock model \\
\hline
\verb;create_bd_tree_prior; & Create a birth-death tree prior \\
\verb;create_cbs_tree_prior; & Create a coalescent Bayesian skyline tree prior \\
\verb;create_ccp_tree_prior; & Create a coalescent constant-population tree prior \\
\verb;create_cep_tree_prior; & Create a coalescent exponential-population tree prior \\
\verb;create_yule_tree_prior; & Create a Yule tree prior \\
\hline
\verb;create_beta_distr; & Create a beta distribution \\
\verb;create_exp_distr; & Create an exponential distribution \\
\verb;create_gamma_distr; & Create a gamma distribution \\
\verb;create_inv_gamma_distr; & Create an inverse gamma distribution \\
\verb;create_laplace_distr; & Create a Laplace distribution \\
\verb;create_log_normal_distr; & Create a log-normal distribution \\
\verb;create_normal_distr; & Create a normal distribution \\
\verb;create_one_div_x_distr; & Create a 1/X distribution \\
\verb;create_poisson_distr; & Create a Poisson distribution \\
\verb;create_uniform_distr; & Create a uniform distribution \\
\hline
\end{tabular}
\caption{babette's main functions}
\label{tab:functions}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
\grid
\grid
\grid

### 2019-01-21 article.tex
Progress

@@
 
-{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, babette, R
+{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, pirouette, R
 
@@
 
-%Phylogenies are commonly used to explore evolutionary hypotheses.
+Phylogenies are commonly used to explore evolutionary hypotheses.
 %Not only can phylogenies show us how species (or other
@@
 %reduces the chances of errors in such repetitive actions.
-%The interface of \verb;babette; mimics the tools it
+%The interface of \verb;pirouette; mimics the tools it
 %is based on. This
 %familiarity helps both beginner and experienced BEAST2 users 
-%to make the step from those tools to \verb;babette;.
-%\verb;babette; enables the creation of a single-script 
+%to make the step from those tools to \verb;pirouette;.
+%\verb;pirouette; enables the creation of a single-script 
 %pipeline from sequence alignments to posterior analysis in R.
@@
 
-%\verb;babette; is written in the R programming language (\cite{R})
-%and enables the full BEAST2 workflow from a single R function call,
+\verb;pirouette; is written in the R programming language (\cite{R}).
+
+\begin{figure}[h]
+  \includegraphics[width=\textwidth]{overview.png}
+\end{figure}
+
+and enables the full BEAST2 workflow from a single R function call,
 %in a similar way to what subsequent usage of BEAUti, DensiTree and Tracer would produce.
-%\verb;babette;'s main function is \verb;bbt_run;, which
+%\verb;pirouette;'s main function is \verb;bbt_run;, which
 %configures BEAST2, runs it and parses its output. 
@@
 
-%\verb;babette; currently has 108 exported functions to set up  
+%\verb;pirouette; currently has 108 exported functions to set up  
 %a BEAST2 configuration file. 
-%\verb;babette; can currently handle the majority of BEAUti use cases.
+%\verb;pirouette; can currently handle the majority of BEAUti use cases.
 %Because of BEAUti's high number of plugins, 
-%\verb;babette; uses a software architecture that is designed to be extended.
-%Furthermore, \verb;babette; has 13 exported functions to run and help run BEAST2.
+%\verb;pirouette; uses a software architecture that is designed to be extended.
+%Furthermore, \verb;pirouette; has 13 exported functions to run and help run BEAST2.
 %One function is used to run BEAST2, another
 %one installs BEAST2 to a default location.
-%Finally, \verb;babette; has 21 exported function to parse the BEAST2 output
-%files and analyze the created posterior. \verb;babette; gives the
+%Finally, \verb;pirouette; has 21 exported function to parse the BEAST2 output
+%files and analyze the created posterior. \verb;pirouette; gives the
 %same ESSes and summary statistics as Tracer. The data is formatted
@@
 
-%Currently, \verb;babette; does not contain all functionality 
+%Currently, \verb;pirouette; does not contain all functionality 
 %in BEAUti, BEAST2 and their many plug-ins, 
 %because these tools themselves also change in time.
-%\verb;babette; currently works only on DNA data,
+%\verb;pirouette; currently works only on DNA data,
 %because this is the most common use case. 
-%Nevertheless, \verb;babette; provides the majority of default tree priors and supports the most important command-line arguments of BEAST2, provides
+%Nevertheless, \verb;pirouette; provides the majority of default tree priors and supports the most important command-line arguments of BEAST2, provides
 %the core Tracer analysis options, and has the most basic subset of 
-%plotting options of DensiTree. Up till now, the \verb;babette; features 
-%implemented are those requested by users. Further extension of \verb;babette; 
+%plotting options of DensiTree. Up till now, the \verb;pirouette; features 
+%implemented are those requested by users. Further extension of \verb;pirouette; 
 %will be based on future user requests.
@@
 
-A first research question that
+A first research question that \verb;pirouette; answers, is:
+
+What is the error BEAST2 makes from a phylogeny using the same 
+speciation model as it was generated by?
+
+Say we have an idealised
 
@@
 phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4):1);")
-ape::plot.phylo(phylogeny, main = "The True Phylogeny")
 \end{lstlisting}
@@
 among others, a Jukes-Cantor site model, a strict clock, and a Yule birth tree prior.
-\verb;babette; will then execute BEAST2 using that file, and
+\verb;pirouette; will then execute BEAST2 using that file, and
 parses the output. The returned data structure, named \verb;out;, 
@@
 \end{lstlisting}
-\verb;babette; also uses the same default prior distributions as BEAUti 
+\verb;pirouette; also uses the same default prior distributions as BEAUti 
 for each of the site models, clock models and tree priors. 
@@
 Here is an example how to specify an exponential distribution for
-the birth rate in a Yule tree prior in \verb;babette;:
+the birth rate in a Yule tree prior in \verb;pirouette;:
 
@@
 \end{lstlisting}
-\verb;babette; also supports node dating. Like BEAUti, one
+\verb;pirouette; also supports node dating. Like BEAUti, one
 can specify Most Recent Common Ancestor ('MRCA') priors.
@@
 including a distribution for the date of that ancestor.
-With \verb;babette;, this is achieved as follows:
+With \verb;pirouette;, this is achieved as follows:
 
@@
 and multiple sets are allowed.
-\verb;babette; allows for the same core functionality as Tracer to show the values of the parameter estimates sampled
+\verb;pirouette; allows for the same core functionality as Tracer to show the values of the parameter estimates sampled
 in the BEAST2 run. This is called the "trace" (hence the name).
@@
 To remove a 20\% burn-in from all parameter estimates 
-in \verb;babette;, the following code can be used:
+in \verb;pirouette;, the following code can be used:
 
@@
 for any parameter estimate.
-To calculate the effective sample sizes (of all estimated variables) in \verb;babette;:
+To calculate the effective sample sizes (of all estimated variables) in \verb;pirouette;:
 
@@
 auto-correlation time and effective sample size. It displays these statistics per
-variable. In \verb;babette;, these summary statistics are collected for
+variable. In \verb;pirouette;, these summary statistics are collected for
 all estimated parameters at once:
@@
 \end{lstlisting}
-\verb;babette; allows for the same functionality as DensiTree.
+\verb;pirouette; allows for the same functionality as DensiTree.
 DensiTree displays the phylogenies in a posterior at the same
@@
 Instead of running the full pipeline, 
-\verb;babette; also allows to only create a BEAST2 configuration file.
+\verb;pirouette; also allows to only create a BEAST2 configuration file.
 To create a BEAST2 configuration file, with all settings to default, use:
@@
 create_beast2_input_file(
-  input_filenames = babette::get_babette_path("anthus_aco.fas"),
+  input_filenames = pirouette::get_pirouette_path("anthus_aco.fas"),
   output_filename = "beast2.xml"
@@
 This file can then be loaded and edited by BEAUti, 
-run by BEAST2, or run by \verb;babette;:
+run by BEAST2, or run by \verb;pirouette;:
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{babette resources}
+\section{pirouette resources}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\verb;babette; is free, libre and open source software available at 
-\url{http://github.com/richelbilderbeek/babette}
+\verb;pirouette; is free, libre and open source software available at 
+\url{http://github.com/richelbilderbeek/pirouette}
 and is licensed under the GNU General Public License v3.0.
-\verb;babette; uses the Travis CI (\url{https://travis-ci.org})
+\verb;pirouette; uses the Travis CI (\url{https://travis-ci.org})
 continuous integration service, which is known to significantly 
@@
 the speed at which new features are added (\cite{vasilescu2015}).
-\verb;babette; has a 100\% code coverage, which correlates with 
+\verb;pirouette; has a 100\% code coverage, which correlates with 
 code quality (\cite{horgan1994,del1995correlation}). 
-\verb;babette; follows Hadley Wickham's style guide (\cite{style_guide}), 
+\verb;pirouette; follows Hadley Wickham's style guide (\cite{style_guide}), 
 which improves software quality (\cite{fang2001}).
-\verb;babette; depends on multiple packages, which are 
+\verb;pirouette; depends on multiple packages, which are 
 \verb;ape; (\cite{APE}), 
-\verb;beautier; (\cite{beautier}),
-\verb;beastier; (\cite{beastier}),
-\verb;devtools; (\cite{devtools}),
-\verb;geiger; (\cite{GEIGER}),
+\verb;babette; (\cite{babette}),
 \verb;ggplot2; (\cite{ggplot2}),
 \verb;knitr; (\cite{knitr}),
+\verb;mcbette; (\cite{mcbette}),
 \verb;phangorn; (\cite{phangorn}),
 \verb;rmarkdown; (\cite{rmarkdown}),
-\verb;seqinr; (\cite{seqinr}),
 \verb;stringr; (\cite{stringr}),
 \verb;testit; (\cite{testit}) and 
-\verb;tracerer; (\cite{tracerer}).
-We tested \verb;babette; to give a clean 
-error message for incorrect input, by
-calling \verb;babette; one million times
-with random or random sensible inputs, 
-using a high performance computer cluster. 
-The test scripts are supplied with \verb;babette;.
+\verb;usethis; (\cite{usethis}).
 
-\verb;babette;'s development takes place on GitHub,
-\url{https://github.com/richelbilderbeek/babette}, 
+\verb;pirouette;'s development takes place on GitHub,
+\url{https://github.com/richelbilderbeek/pirouette}, 
 which accommodates collaboration (\cite{perez2016ten}) 
 and improves transparency (\cite{gorgolewski2016practical}).
-\verb;babette;'s GitHub facilitates feature requests and 
+\verb;pirouette;'s GitHub facilitates feature requests and 
 has guidelines how to do so.
 
-\verb;babette;'s documentation is extensive. All functions are documented
+\verb;pirouette;'s documentation is extensive. All functions are documented
 in the package's internal documentation. For quick use, 
@@
 For easy exploration, each exported function's documentation links to related functions.
-Additionally, \verb;babette; has a vignette that demonstrates extensively how
+Additionally, \verb;pirouette; has a vignette that demonstrates extensively how
 to use it. There is documentation on the GitHub to get started, 
-with a dozen examples of BEAUti screenshots with equivalent \verb;babette; code.
-Finally, \verb;babette; has tutorial videos that can 
+with a dozen examples of BEAUti screenshots with equivalent \verb;pirouette; code.
+Finally, \verb;pirouette; has tutorial videos that can 
 be downloaded or viewed on YouTube, \url{https://goo.gl/weKaaU}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Citation of babette}
+\section{Citation of pirouette}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Scientists using \verb;babette; in a published paper can cite this
-article, and/or cite the \verb;babette; package 
+Scientists using \verb;pirouette; in a published paper can cite this
+article, and/or cite the \verb;pirouette; package 
 directly. To obtain this citation from within an R script, use:
@@
 \begin{lstlisting}[language=R]
-> citation("babette")
+> citation("pirouette")
 \end{lstlisting}
@@
 
-Thanks to Yacine Ben Chehida and Paul van Els for supplying their 
-BEAST2 use cases. Thanks again to Paul van Els for sharing his FASTA files 
-for use by this package. Thanks to Leonel Herrera-Alsina, Raphael Scherrer 
-and Giovanni Laudanno for their comments on this package and article.
-Thanks to Huw Ogilvie, Michael Matschiner and one anonymous reviewer for reviewing this article.
-Thanks to rOpenSci, and especially Noam Ross and Guangchuang Yu for reviewing the package's source code.
 We would like to thank the Center for Information Technology of the University 
 of Groningen for their support and for providing access to the Peregrine 
-high performance computing cluster. We thank the Netherlands 
+high performance computing cluster. 
+We thank the Netherlands 
 Organization for Scientific Research (NWO) for financial support 
@@
 
-All code is archived at \url{http://github.com/richelbilderbeek/babette_article},
-with DOI \url{https://doi.org/10.5281/zenodo.1251203}.
+All code is archived at \url{http://github.com/richelbilderbeek/pirouette_article},
+with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
 
@@
 
-RJCB and RSE conceived the idea for the package. RJCB created and tested the package, and wrote the first draft of the manuscript. RSE contributed substantially to revisions.
+RJCB, GL and RSE conceived the idea for the package. 
+RJCB created and tested the package, and wrote the first draft of the manuscript.
+GL tested the package and contributed substantially to revisions.
+RSE contributed to revisions.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[h]
-\centering
-\begin{tabular}{ | l | l | }
-\hline
-\textbf{Name} & \textbf{Description} \\
-\hline
-\verb;bbt_run; & Run BEAST2 \\
-\hline
-\verb;create_gtr_site_model; & Create a GTR site model \\
-\verb;create_hky_site_model; & Create an HKY site model \\
-\verb;create_jc69_site_model; & Create a Jukes-Cantor site model \\
-\verb;create_tn93_site_model; & Create a TN93 site model \\
-\hline
-\verb;create_rln_clock_model; & Create a relaxed log-normal clock model \\
-\verb;create_strict_clock_model; & Create a strict clock model \\
-\hline
-\verb;create_bd_tree_prior; & Create a birth-death tree prior \\
-\verb;create_cbs_tree_prior; & Create a coalescent Bayesian skyline tree prior \\
-\verb;create_ccp_tree_prior; & Create a coalescent constant-population tree prior \\
-\verb;create_cep_tree_prior; & Create a coalescent exponential-population tree prior \\
-\verb;create_yule_tree_prior; & Create a Yule tree prior \\
-\hline
-\verb;create_beta_distr; & Create a beta distribution \\
-\verb;create_exp_distr; & Create an exponential distribution \\
-\verb;create_gamma_distr; & Create a gamma distribution \\
-\verb;create_inv_gamma_distr; & Create an inverse gamma distribution \\
-\verb;create_laplace_distr; & Create a Laplace distribution \\
-\verb;create_log_normal_distr; & Create a log-normal distribution \\
-\verb;create_normal_distr; & Create a normal distribution \\
-\verb;create_one_div_x_distr; & Create a 1/X distribution \\
-\verb;create_poisson_distr; & Create a Poisson distribution \\
-\verb;create_uniform_distr; & Create a uniform distribution \\
-\hline
-\end{tabular}
-\caption{babette's main functions}
-\label{tab:functions}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}
-\grid
-\grid
-\grid

### 2019-01-21 article.tex
First useable PDF

@@
 \usepackage{bm}
+
+% sidewaysfigure
+\usepackage{rotating}
 
@@
 Phylogenies are commonly used to explore evolutionary hypotheses.
+
 %Not only can phylogenies show us how species (or other
@@
 
-\begin{figure}[h]
+\begin{sidewaysfigure}
+  \centering
   \includegraphics[width=\textwidth]{overview.png}
-\end{figure}
+  \caption{pirouette pipeline}
+  \label{fig:pipeline}
+\end{sidewaysfigure}
 
@@
 
-\verb;piouette; can be installed easily from CRAN:
+\verb;pirouette; can be installed easily from CRAN:
 
@@
 
-This code will create a (temporary) BEAST2 configuration file,
-from the FASTA file with name \verb;anthus_aco.fas; (which
-is supplied with the package, from \cite{VanEls2018}), 
-using the same default settings as BEAUti, which are, 
-among others, a Jukes-Cantor site model, a strict clock, and a Yule birth tree prior.
-\verb;pirouette; will then execute BEAST2 using that file, and
-parses the output. The returned data structure, named \verb;out;, 
-is a list of parameter estimates (called \verb;estimates;), posterior 
-phylogenies (called \verb;anthus_aco_trees;, named after
-the alignment's name) and MCMC operator performance (\verb;operators;).
-An example of using a different site model, clock model 
-and tree prior is:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-out <- bbt_run(
-  fasta_filenames = "anthus_aco.fas",
-  site_models = create_hky_site_model(),
-  clock_models = create_rln_clock_model(),
-  tree_priors = create_bd_tree_prior()
-)
-\end{lstlisting}
-This code uses an HKY site model, a relaxed log-normal clock model and a 
-birth-death tree prior, each with their default settings in BEAUti.
-Table \ref{tab:functions} shows an overview of all functions to 
-create site models, clock models and tree priors.
-Note that the arguments' names \verb;site_models;, \verb;clock_models; 
-and \verb;tree_priors; are plural, as each of these
-can be (a list of) one or more elements. Each of these arguments must 
-have the same number of elements, so that each alignment has its
-own site model, clock model and tree prior. 
-An example of two alignments, each with its own site model, is:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-out <- bbt_run(
-  fasta_filenames = c(
-    "anthus_aco.fas", 
-    "anthus_nd2.fas"
-  ),
-  site_models = list(
-    create_tn93_site_model(), 
-    create_gtr_site_model()
-  )
-)
-\end{lstlisting}
-\verb;pirouette; also uses the same default prior distributions as BEAUti 
-for each of the site models, clock models and tree priors. 
-For example, by default, a Yule tree prior assumes that the birth rate 
-follows a uniform distribution, 
-from minus infinity to plus infinity. 
-One may prefer a different ddistribution instead. 
-Here is an example how to specify an exponential distribution for
-the birth rate in a Yule tree prior in \verb;pirouette;:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-out <- bbt_run(
-  fasta_filenames = "anthus_aco.fas",
-  tree_priors = create_yule_tree_prior(
-    birth_rate_distr = create_exp_distr()    
-  )
-)
-\end{lstlisting}
-In this same example, one may specify
-the initial shape parameters of the exponential distribution.
-In BEAST2's implementation, an exponential distribution 
-has one shape parameter: its mean, which can be set to any
-value with \verb;BEAUti;. To set the 
-mean value of the exponential distribution to a 
-fixed (non-estimated) value, do:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-out <- bbt_run(
-  fasta_filenames = "anthus_aco.fas",
-  tree_priors = create_yule_tree_prior(
-    birth_rate_distr = create_exp_distr(
-      mean = create_mean_param(
-        value = 1.0, 
-        estimate = FALSE
-      )
-    )    
-  )
-)
-\end{lstlisting}
-\verb;pirouette; also supports node dating. Like BEAUti, one
-can specify Most Recent Common Ancestor ('MRCA') priors.
-An MRCA prior allows to specify taxa having a common ancestor,
-including a distribution for the date of that ancestor.
-With \verb;pirouette;, this is achieved as follows:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-out <- bbt_run(
-  fasta_filenames = "anthus_aco.fas",
-  mrca_priors = create_mrca_prior(
-    taxa_names = sample(get_taxa_names("anthus_aco.fas"), size = 2),
-    alignment_id = get_alignment_id("anthus_aco.fas"),
-    is_monophyletic = TRUE,
-    mrca_distr = create_normal_distr(
-      mean = create_mean_param(value = 15.0, estimate = FALSE),
-      sigma = create_sigma_param(value = 0.025, estimate = FALSE)
-    )
-  )
-)
-\end{lstlisting}
-Instead of dating the ancestor of two random taxa, any subset of taxa can be selected,
-and multiple sets are allowed.
-\verb;pirouette; allows for the same core functionality as Tracer to show the values of the parameter estimates sampled
-in the BEAST2 run. This is called the "trace" (hence the name).
-The start of the trace, called the "burn-in", is usually discarded, as an MCMC 
-algorithm (such as used by BEAST2) first has to converge to
-its equilibrium and hence the parameter estimates are not 
-representative. By default, Tracer discards the first 10\% of all 
-the parameter estimates. 
-To remove a 20\% burn-in from all parameter estimates 
-in \verb;pirouette;, the following code can be used:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-traces <- remove_burn_ins(
-  traces = out$estimates, 
-  burn_in_fraction = 0.2
-)
-\end{lstlisting}
-Tracer shows the ESSes of each posterior's variables.
-These ESSes are important to determine the strength of the
-inference. As a rule of thumb, an ESS of 200 is acceptable 
-for any parameter estimate.
-To calculate the effective sample sizes (of all estimated variables) in \verb;pirouette;:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-esses <- calc_esses(
-  traces = traces, 
-  sample_interval = 1000
-)
-\end{lstlisting}
-Tracer displays multiple summary statistics for each
-estimated variable: the mean and its standard error, standard deviation,
-variance, median, mode, geometric mean, 95\% highest posterior density interval, 
-auto-correlation time and effective sample size. It displays these statistics per
-variable. In \verb;pirouette;, these summary statistics are collected for
-all estimated parameters at once:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-sum_stats <- calc_summary_stats(
-  traces = traces, 
-  sample_interval = 1000
-)
-\end{lstlisting}
-\verb;pirouette; allows for the same functionality as DensiTree.
-DensiTree displays the phylogenies in a posterior at the same
-time scale, drawn one over one another, allowing to see the uncertainty in
-topology and branch lengths. 
-The posterior phylogenies are stored as \verb;anthus_aco_trees; in the object \verb;out;,
-and can be plotted as follows:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-plot_densitree(phylos = out$anthus_aco_trees)
-\end{lstlisting}
-Instead of running the full pipeline, 
-\verb;pirouette; also allows to only create a BEAST2 configuration file.
-To create a BEAST2 configuration file, with all settings to default, use:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-create_beast2_input_file(
-  input_filenames = pirouette::get_pirouette_path("anthus_aco.fas"),
-  output_filename = "beast2.xml"
-)
-\end{lstlisting}
-This file can then be loaded and edited by BEAUti, 
-run by BEAST2, or run by \verb;pirouette;:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-run_beast2(
-  input_filename = "beast2.xml",
-  output_log_filename = "run.log",
-  output_trees_filenames = "posterior.trees",
-  output_state_filename = "final.xml.state"
-)
-\end{lstlisting}
-\verb;run_beast2; is a function that only runs BEAST2, 
-and does not parse the output files (unlike \verb;bbt_run;). 
-In the example above, 
-we specify the names of the desired BEAST2 
-output files explicitly, and these will be created in the
-R working directory, after which they can be inspected 
-with other tools, or used to continue a BEAST2 run.
-When the names of these files are not specified, 
-both \verb;bbt_run; and \verb;run_beast2;
-put these files in the default temporary folder (as
-obtained from \verb;temp.dir();) to keep
-the working directory clean of intermediate files.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \verb;ape; (\cite{APE}), 
-\verb;babette; (\cite{babette}),
+\verb;babette; (\cite{bilderbeek2018babette}),
 \verb;ggplot2; (\cite{ggplot2}),

### 2019-01-21 article.tex
Progress

@@
 \end{sidewaysfigure}
+
+The goal of \verb;pirouette; is to measure the inference error BEAST2
+makes from a given/known/true phylogeny. This involves multiple steps,
+described in detail below. 
+In brief: from a given phylogeny, an alignment is simulated, from which a BEAST2
+posterior is inferred, of which its phylogenies are compared with the given
+phylogeny. The more the posterior phylogenies match the given phylogeny,
+the lower the error. The is also the option to generate a 'twin tree',
+that goes through the same pipeline.
+
+The first step simulates a DNA alignment from the given phylogeny.
+One can specify a DNA sequence
+of any length at the root of the phylogeny, a DNA mutation rate, a
+site (also: nucleotide substitution) model, 
+a clock model, a random number generator (RNG) seed and a location
+where the alignment is saved to. This step is relatively fast, but longer
+DNA alignments will noticably slow down the inference step.
+
+The second step selects an inference model, using the alignment.
+The user can specify to use the generative model or select a best model
+from a set of inference models. When selecting the generative model,
+the site and clock model used in the alignment simulation are used
+in inference. Because the given phylogeny (on which the alignment is based)
+may have followed any tree prior (also: speciation model), the user needs
+to specify which tree prior is used in inference. When selecting a best
+model, the alignment is used to find the inference model that has the
+highest evidence (also: marginal likelihood)
+
+The third step infers a Bayesian posterior from the simulated alignment.
+From the
+DNA alignment a Bayesian posterior is generated, using a user-specified
+inference setup. The given tree is compared with all posterior trees, using
+a user-specified error measurement setup.
+
+A first and optional step is to generate a 'twin tree', that will be
+analyzed in the same way as the true tree.
 

### 2019-01-21 article.tex
@Giappo will love this

@@
 The user can specify to use the generative model or select a best model
-from a set of inference models. When selecting the generative model,
+from a set of inference models. 
+When selecting the generative model,
 the site and clock model used in the alignment simulation are used
@@
 may have followed any tree prior (also: speciation model), the user needs
-to specify which tree prior is used in inference. When selecting a best
+to specify which tree prior is used in inference. 
+When selecting a best
 model, the alignment is used to find the inference model that has the
-highest evidence (also: marginal likelihood)
+highest evidence (also: marginal likelihood) from a set of inference models.
+The evidence of an inference model is estimated using a nested-sampling
+approach, as decribed in Maturana et al., 2018. The nested sampling is
+performed by \verb;babette;, that call the 'NS' BEAST2 package. 
+Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac,
+which excludes Windows from this way of selecting an inference model.
 
-The third step infers a Bayesian posterior from the simulated alignment.
+The third step infers a Bayesian posterior from the simulated alignment,
+using the inference model(s) selected in the previous step. The user
+can specify the additional parameters needed for a BEAST2 run, which
+are the Markov-Chain Monte Carlo (MCMC) setup, 
+an optional Most Recent Common Ancestor (MRCA) prior and an RNG seed.
+The MCMC setup determines the number of posterior trees sampled.
+An MRCA prior allows the inferred phylogeny to have a dated crown age.
+
+The fourth step measures the inference error, using the phylogenies in the
+Bayesian posterior.
+
 From the

### 2019-01-21 article.tex
@Giappo will love this

@@
 
-%Not only can phylogenies show us how species (or other
-%evolutionary units) are related to each other, 
-%but we can also estimate relevant parameters such as extinction and 
-%speciation rates from them.
-%There are many phylogenetics tools available to obtain an estimate 
-%of the phylogeny of a given set of species. 
-%BEAST2 (\cite{bouckaert2014beast}) is one of the most widely used ones.
-%It uses a Bayesian statistical framework to estimate 
-%the joint posterior distribution of phylogenies and model parameters, 
-%from one or more DNA, RNA or amino acid alignments (see figure 1 
-%for an overview of the workflow).
-
-%Here, we present \verb;pirouette;:
-%’BEAUti 2, BEAST2 and Tracer for R’, 
-%which creates BEAST2 (v.2.4.7) configuration files,
-%runs BEAST2, and analyzes its results,
-%all from an R function call. This
-%will save time, tedious mouse clicking and 
-%reduces the chances of errors in such repetitive actions.
-%The interface of \verb;pirouette; mimics the tools it
-%is based on. This
-%familiarity helps both beginner and experienced BEAST2 users 
-%to make the step from those tools to \verb;pirouette;.
-%\verb;pirouette; enables the creation of a single-script 
-%pipeline from sequence alignments to posterior analysis in R.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The fourth step measures the inference error, using the phylogenies in the
-Bayesian posterior.
+Bayesian posterior. These phylogenies are compared to the given
+phylogeny using an error statistic, which is the nLTT 
+statistic [Janzen et al., ?2016]. Additionally, the user can specify the
+propertion of posterior phylogenies to discard (also called: the burn-in).
 
-From the
-DNA alignment a Bayesian posterior is generated, using a user-specified
-inference setup. The given tree is compared with all posterior trees, using
-a user-specified error measurement setup.
-
-A first and optional step is to generate a 'twin tree', that will be
+An optional step is to generate a 'twin tree', that will be
 analyzed in the same way as the true tree.
-
-and enables the full BEAST2 workflow from a single R function call,
-%in a similar way to what subsequent usage of BEAUti, DensiTree and Tracer would produce.
-%\verb;pirouette;'s main function is \verb;bbt_run;, which
-%configures BEAST2, runs it and parses its output. 
-%\verb;bbt_run; needs at least the name of a 
-%FASTA file containing a DNA alignment. 
-%The default settings for the other arguments of \verb;bbt_run; 
-%are identical to BEAUti's and BEAST2's default settings.
-%Per alignment, a site model, clock model and tree prior can be chosen.
-%Multiple alignments can be used, each with its own (unlinked) site model, 
-%clock model and tree prior.
-
-%\verb;pirouette; currently has 108 exported functions to set up  
-%a BEAST2 configuration file. 
-%\verb;pirouette; can currently handle the majority of BEAUti use cases.
-%Because of BEAUti's high number of plugins, 
-%\verb;pirouette; uses a software architecture that is designed to be extended.
-%Furthermore, \verb;pirouette; has 13 exported functions to run and help run BEAST2.
-%One function is used to run BEAST2, another
-%one installs BEAST2 to a default location.
-%Finally, \verb;pirouette; has 21 exported function to parse the BEAST2 output
-%files and analyze the created posterior. \verb;pirouette; gives the
-%same ESSes and summary statistics as Tracer. The data is formatted
-%such that it can easily be visualized using \verb;ggplot2; (for a trace,
-%similar to Tracer) or \verb;phangorn; (\cite{phangorn}) (for 
-%the phylogenies in a posterior, similar to DensiTree).
-
-%Currently, \verb;pirouette; does not contain all functionality 
-%in BEAUti, BEAST2 and their many plug-ins, 
-%because these tools themselves also change in time.
-%\verb;pirouette; currently works only on DNA data,
-%because this is the most common use case. 
-%Nevertheless, \verb;pirouette; provides the majority of default tree priors and supports the most important command-line arguments of BEAST2, provides
-%the core Tracer analysis options, and has the most basic subset of 
-%plotting options of DensiTree. Up till now, the \verb;pirouette; features 
-%implemented are those requested by users. Further extension of \verb;pirouette; 
-%will be based on future user requests.
 

### 2019-01-21 article.tex
Methods first draft done

@@
 An optional step is to generate a 'twin tree', that will be
-analyzed in the same way as the true tree.
+analyzed in the same way as the true tree. The twin tree has the same
+topology as the given phylogeny, yet its branch lengths follow the same
+distribution as a Yule (pure-birth) or (constant rate) birth-death tree prior.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage}
+\section{Usage: first research question}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A first research question that \verb;pirouette; answers, is:
-
 What is the error BEAST2 makes from a phylogeny using the same 
@@
 
-Say we have an idealised
+We start with an idealised birth-death tree of five taxa and a crown age of 10:
 
+% Need idealised birth-death tree
+% From https://github.com/richelbilderbeek/pirouette_article/issues/1
+% and https://github.com/richelbilderbeek/pirouette/issues/35
 \begin{lstlisting}[language=R, floatplacement=H]
@@
 
+The first step in \verb;pirouette; is to simulate a DNA alignment from the 
+given phylogeny. To do so, we must specify the DNA root sequence
+and a mutation rate. In this example, the DNA root sequence consists
+out of four block of 250 nucleotides each, where the per-nucleotide
+mutation rate is on average 0.1 mutations per time unit:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+alignment_params <- create_alignment_params(
+  root_sequence = create_blocked_dna(length = 1000),
+  mutation_rate = 0.1
+)
+\end{lstlisting}
+
+The second step is to selects how an inference model is picked.
+We just select the generative model, which are the site and clock
+model of the alignment. Additionally, we specify a birth-death tree prior
+to be the generative tree prior.
+
+% Note that this syntax does not work yet, see
+% https://github.com/richelbilderbeek/pirouette/issues/36
+\begin{lstlisting}[language=R, floatplacement=H]
+model_select_param <- create_gen_model_select_param(
+  alignment_params = alignment_params,
+  tree_prior = create_bd_tree_prior()
+)
+\end{lstlisting}
+
+The third and fourth step have sensible defaults, and we are not
+interested in using a twin tree. We can now measure the errors BEAST2
+makes when inferring the given phylogeny like this:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+errors <- pir_run(
+  phylogeny,
+  alignment_params = alignment_params,
+  model_select_params = model_select_param
+)
+\end{lstlisting}
+
+\verb;pirouette; has a nice plotting function:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+pir_plot(errors)
+\end{lstlisting}
+
+% TODO: write script to create figure 2, see
+% https://github.com/richelbilderbeek/pirouette/issues/37
+The resulting figure is shown in figure 2.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Usage: second research question}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A second research question that \verb;pirouette; answers, is:
+What is the error BEAST2 makes from a phylogeny 
+picking the best inference model?
+
+Here we start with an tree from an unknown speciation model,
+that has four taxa and a crown age of 5:
+
+% Need idealised birth-death tree
+% From https://github.com/richelbilderbeek/pirouette_article/issues/1
+% and https://github.com/richelbilderbeek/pirouette/issues/35
+\begin{lstlisting}[language=R, floatplacement=H]
+phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4):1);")
+\end{lstlisting}
+
+The first step in \verb;pirouette; is to simulate a DNA alignment from the 
+given phylogeny. We'll use the same parameters here:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+alignment_params <- create_alignment_params(
+  root_sequence = create_blocked_dna(length = 1000),
+  mutation_rate = 0.1
+)
+\end{lstlisting}
+
+The second step is to select how an inference model is picked.
+We will let the inference model that has the highest evidence be used
+in the Bayesian inference. By default, \verb;pirouette; will try
+all combinations of site models, clock models and tree priors.
+
+% Note that this syntax does not work yet, see
+% https://github.com/richelbilderbeek/pirouette/issues/36
+\begin{lstlisting}[language=R, floatplacement=H]
+model_select_param <- create_best_model_select_param()
+\end{lstlisting}
+
+Also here, the third and fourth step have sensible defaults, and we are not
+interested in using a twin tree. We can now measure the errors BEAST2
+makes when inferring the given phylogeny like this:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+errors <- pir_run(
+  phylogeny,
+  alignment_params = alignment_params,
+  model_select_params = model_select_param
+)
+\end{lstlisting}
+
+Again, showing the results:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+pir_plot(errors)
+\end{lstlisting}
+
+% TODO: write script to create figure 3, see
+% https://github.com/richelbilderbeek/pirouette/issues/38
+The resulting figure is shown in figure 3.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Usage: third research question}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A third research question that \verb;pirouette; answers, is:
+What is the error BEAST2 makes from a phylogeny 
+picking the best inference model, compared to the background noise?
+
+Here we again start with an tree from an unknown speciation model,
+that has four taxa and a crown age of 5:
+
+% Need idealised birth-death tree
+% From https://github.com/richelbilderbeek/pirouette_article/issues/1
+% and https://github.com/richelbilderbeek/pirouette/issues/35
+\begin{lstlisting}[language=R, floatplacement=H]
+phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4):1);")
+\end{lstlisting}
+
+The first step in \verb;pirouette; is to simulate a DNA alignment from the 
+given phylogeny. We'll use the same parameters here:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+alignment_params <- create_alignment_params(
+  root_sequence = create_blocked_dna(length = 1000),
+  mutation_rate = 0.1
+)
+\end{lstlisting}
+
+The second step is to select how an inference model is picked.
+We will let the inference model that has the highest evidence be used
+in the Bayesian inference. By default, \verb;pirouette; will try
+all combinations of site models, clock models and tree priors.
+
+% Note that this syntax does not work yet, see
+% https://github.com/richelbilderbeek/pirouette/issues/36
+\begin{lstlisting}[language=R, floatplacement=H]
+model_select_param <- create_best_model_select_param()
+\end{lstlisting}
+
+Also here, the third and fourth step have sensible defaults.
+
+This time, we are interested in creating a twin tree. A twin tree
+has the same topology as the given tree, yet its branch lengths follow
+a birth-death distribution. Creating a twinning parameter is easy,
+as it has sensible default settings:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+twinning_params <- create_twinning_params()
+\end{lstlisting}
+
+We can now measure the errors BEAST2
+makes when inferring the given phylogeny, compared
+to the background noise, like this:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+errors <- pir_run(
+  phylogeny,
+  twinning_params = twinning_params,
+  alignment_params = alignment_params,
+  model_select_params = model_select_param
+)
+\end{lstlisting}
+
+Again, showing the results:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+pir_plot(errors)
+\end{lstlisting}
+
+% TODO: write script to create figure 4, see
+% https://github.com/richelbilderbeek/pirouette/issues/39
+The resulting figure is shown in figure 4
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-01-23 article.tex
Changes @Giappo

@@
 
+%comments
+\newcommand{\giovanni}[1]{\textcolor{blue}{\textbf{[GL: #1]}}}
+
 \begin{document}
@@
     posterior of jointly-estimated phylogenies and model parameter estimates.
-    When a new macro-evolutionary speciation model is developed,
-    a good first step is to measure the error BEAST2 makes on known
-    phylogeny derived from the new speciation process, 
+    When a new macro-evolutionary speciation \giovanni{maybe "diversification" is a better word} model is developed,
+    a good first step is to measure the error BEAST2 makes on a known
+    phylogeny derived from the new speciation process \giovanni{I'd use "diversification mechanisms" in place of "speciation process"}, 
     with its current set of inference models. \\
@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given/known/true phylogeny. This involves multiple steps,
+makes from a given/known/true phylogeny \giovanni{I'd use: "given phylogeny of which full information is provided"}. This involves multiple steps,
 described in detail below.
+
+\giovanni{suggestion: expressing the same concepts of the "in brief" section in points could be an easier experience for the reader}
+\giovanni{start example}
+
+The pipeline is summarized by the following steps:
+\begin{enumerate}
+    \item from a given phylogeny an alignment is simulated;
+    \item the alignment is used as BEAST2 input to infer a posterior distribution of phylogenies;
+    \item posterior phylogenies are compared with the given phylogenies to estimate the error we make. The bigger the match, the lower the error;
+\end{enumerate}
+There is also the option to generate a 'twin tree',
+that goes through the same pipeline.
+\giovanni{end example}
+
 In brief: from a given phylogeny, an alignment is simulated, from which a BEAST2
@@
 phylogeny. The more the posterior phylogenies match the given phylogeny,
-the lower the error. The is also the option to generate a 'twin tree',
+the lower the error. There is also the option to generate a 'twin tree',
 that goes through the same pipeline.
@@
 of any length at the root of the phylogeny, a DNA mutation rate, a
-site (also: nucleotide substitution) model, 
+site (also: \giovanni{I'd use "i.e."} nucleotide substitution) model, 
 a clock model, a random number generator (RNG) seed and a location
 where the alignment is saved to. This step is relatively fast, but longer
-DNA alignments will noticably slow down the inference step.
+DNA alignments will noticeably slow down the inference step.
 
 The second step selects an inference model, using the alignment.
-The user can specify to use the generative model or select a best model
+The user can specify \giovanni{whether} to use the generative model or select a best model
 from a set of inference models. 
@@
 the site and clock model used in the alignment simulation are used
-in inference. Because the given phylogeny (on which the alignment is based)
-may have followed any tree prior (also: speciation model), the user needs
-to specify which tree prior is used in inference. 
-When selecting a best
+in \giovanni{the} inference. Because the given phylogeny (on which the alignment is based)
+may have followed any tree prior (also: \giovanni{I'd use "i.e."} speciation model), the user needs
+to specify which tree prior is used in \giovanni{the} inference. 
+When selecting a \giovanni{I'd use "the"} best
 model, the alignment is used to find the inference model that has the
-highest evidence (also: marginal likelihood) from a set of inference models.
-The evidence of an inference model is estimated using a nested-sampling
-approach, as decribed in Maturana et al., 2018. The nested sampling is
+highest evidence (also: \giovanni{I'd use "i.e."} marginal likelihood) from a set of inference models.
+The evidence of an inference model is estimated using a nested sampling (NS)
+approach, as described in \cite{maturana2018model}. The nested sampling is
 performed by \verb;babette;, that call the 'NS' BEAST2 package. 
-Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac,
-which excludes Windows from this way of selecting an inference model.
+Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; inference model selection is not available on Windows.
 
@@
 using the inference model(s) selected in the previous step. The user
-can specify the additional parameters needed for a BEAST2 run, which
+can specify the additional parameters needed for the BEAST2 run, which
 are the Markov-Chain Monte Carlo (MCMC) setup, 
@@
 phylogeny using an error statistic, which is the nLTT 
-statistic [Janzen et al., ?2016]. Additionally, the user can specify the
-propertion of posterior phylogenies to discard (also called: the burn-in).
+statistic \cite{janzen2015approximate}. Additionally, the user can specify the
+proportion of posterior phylogenies to discard (also called: the burn-in \giovanni{as far as I know the burn-in is to discard the start of a chain because it's still too far from the convergence, see http://www.beast2.org/glossary/}).
 
@@
 \begin{lstlisting}[language=R, floatplacement=H]
-install.packages("piouette")
+install.packages("pirouette")
 \end{lstlisting}
@@
 \end{lstlisting}
+\giovanni{is "usthis" supposed to be "devtools"?}
 To start using \verb;pirouette;, load its functions in the global namespace first:
@@
 
-A first research question that \verb;pirouette; answers, is:
+A first research question that \verb;pirouette; answers is:
 What is the error BEAST2 makes from a phylogeny using the same 
-speciation model as it was generated by?
+speciation \giovanni{I'd use: "diversification"} model as it was generated by?
 
@@
 out of four block of 250 nucleotides each, where the per-nucleotide
-mutation rate is on average 0.1 mutations per time unit:
+mutation rate is on average \giovanni{I don't think it's correct to say that's "on average".
+Usually when you have a rate $\lambda$ you suppose to use a Poisson process generated by this rate: $\frac{\lambda^k e^\lambda}{k!}$.
+$\lambda$ turns out to be also the expected number of occurrences} 0.1 mutations per unit time:
 
@@
 
-The second step is to selects how an inference model is picked.
-We just select the generative model, which are the site and clock
-model of the alignment. Additionally, we specify a birth-death tree prior
+The second step is to select how an inference model is picked.
+We just select the generative model, which consists in a collection of three elements: a site model, a clock
+model of the alignment and a birth-death tree prior
 to be the generative tree prior.
@@
 interested in using a twin tree. We can now measure the errors BEAST2
-makes when inferring the given phylogeny like this:
+makes when inferring the given phylogeny in this way:
 
@@
 A second research question that \verb;pirouette; answers, is:
-What is the error BEAST2 makes from a phylogeny 
+What is the error BEAST2 makes from a phylogeny when
 picking the best inference model?
 
-Here we start with an tree from an unknown speciation model,
+Here we start with a tree \giovanni{generated} from an unknown speciation \giovanni{I'd use: "diversification"} model,
 that has four taxa and a crown age of 5:
@@
 
-A third research question that \verb;pirouette; answers, is:
+\giovanni{I noticed that "phylogeny", "alignment params" and "model select param" are needed for all the three research questions.
+Some times, though, you repeat yourself.
+In an article you want to be as concise as possible so I'd suggest to avoid repetitions.
+One solution could be to summarize all the parameter settings in a summary scheme}
+
+A third research question that \verb;pirouette; answers is:
 What is the error BEAST2 makes from a phylogeny 
@@
 
-Here we again start with an tree from an unknown speciation model,
+Here we start again with a tree from an unknown speciation \giovanni{I'd use: "diversification"} model,
 that has four taxa and a crown age of 5:
@@
 \verb;pirouette;'s GitHub facilitates feature requests and 
-has guidelines how to do so.
+has guidelines on how to do so.
 
@@
 Finally, \verb;pirouette; has tutorial videos that can 
-be downloaded or viewed on YouTube, \url{https://goo.gl/weKaaU}.
+be downloaded or viewed on YouTube, \url{https://goo.gl/weKaaU} \giovanni{be sure that shortened link are not deprecated after a certain time}.
 

### 2019-01-24 article.tex
Revised the article, thanks @Giappo

@@
 \newcommand{\giovanni}[1]{\textcolor{blue}{\textbf{[GL: #1]}}}
+\newcommand{\richel}[1]{\textcolor{orange}{\textbf{[RB: #1]}}}
 
@@
     posterior of jointly-estimated phylogenies and model parameter estimates.
-    When a new macro-evolutionary speciation \giovanni{maybe "diversification" is a better word} model is developed,
+    When a new macro-evolutionary diversification model is developed,
     a good first step is to measure the error BEAST2 makes on a known
-    phylogeny derived from the new speciation process \giovanni{I'd use "diversification mechanisms" in place of "speciation process"}, 
+    phylogeny derived from a new diversification mechanism, 
     with its current set of inference models. \\
@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given/known/true phylogeny \giovanni{I'd use: "given phylogeny of which full information is provided"}. This involves multiple steps,
-described in detail below.
-
-\giovanni{suggestion: expressing the same concepts of the "in brief" section in points could be an easier experience for the reader}
-\giovanni{start example}
+makes from a given phylogeny of which full information is provided.
+This involves multiple steps, described in detail below.
 
@@
 that goes through the same pipeline.
-\giovanni{end example}
-
-In brief: from a given phylogeny, an alignment is simulated, from which a BEAST2
-posterior is inferred, of which its phylogenies are compared with the given
-phylogeny. The more the posterior phylogenies match the given phylogeny,
-the lower the error. There is also the option to generate a 'twin tree',
-that goes through the same pipeline.
 
@@
 of any length at the root of the phylogeny, a DNA mutation rate, a
-site (also: \giovanni{I'd use "i.e."} nucleotide substitution) model, 
+site (i.e. nucleotide substitution) model, 
 a clock model, a random number generator (RNG) seed and a location
@@
 The second step selects an inference model, using the alignment.
-The user can specify \giovanni{whether} to use the generative model or select a best model
+The user can specify whether to use the generative model or select a best model
 from a set of inference models. 
@@
 the site and clock model used in the alignment simulation are used
-in \giovanni{the} inference. Because the given phylogeny (on which the alignment is based)
-may have followed any tree prior (also: \giovanni{I'd use "i.e."} speciation model), the user needs
-to specify which tree prior is used in \giovanni{the} inference. 
-When selecting a \giovanni{I'd use "the"} best
+in the inference. Because the given phylogeny (on which the alignment is based)
+may have followed any tree prior (i.e. speciation model), the user needs
+to specify which tree prior is used in the inference. 
+When selecting for the best
 model, the alignment is used to find the inference model that has the
-highest evidence (also: \giovanni{I'd use "i.e."} marginal likelihood) from a set of inference models.
+highest evidence (i.e. marginal likelihood) from a set of inference models.
 The evidence of an inference model is estimated using a nested sampling (NS)
@@
 performed by \verb;babette;, that call the 'NS' BEAST2 package. 
-Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; inference model selection is not available on Windows.
+Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; 
+inference model selection is not available on Windows.
 
@@
 phylogeny using an error statistic, which is the nLTT 
-statistic \cite{janzen2015approximate}. Additionally, the user can specify the
-proportion of posterior phylogenies to discard (also called: the burn-in \giovanni{as far as I know the burn-in is to discard the start of a chain because it's still too far from the convergence, see http://www.beast2.org/glossary/}).
+statistic \cite{janzen2015approximate} by default,
+but also an custom error statistic is supplied, based
+on the absolute difference in the gamma statistic [Pybus & Harvy]. 
+Additionally, the user can specify the
+proportion of posterior phylogenies to 
+discard (i.e. the burn-in), throwing away the first 10%
+of all phylogenies by default. This burn-in is used to discard
+the part of the MCMC chain that has not yet converged to a
+representative part of state space.
 
@@
 distribution as a Yule (pure-birth) or (constant rate) birth-death tree prior.
+The use of such a twin tree, is to assess the minimum level of noise (i.e. 
+error) BEAST2 makes when the input (the twin phylogeny) has an ideal branch
+length distribution.
 
@@
 \begin{lstlisting}[language=R, floatplacement=H]
-usthis::install_github("richelbilderbeek/pirouette")
+usethis::install_github("richelbilderbeek/pirouette")
 \end{lstlisting}
-\giovanni{is "usthis" supposed to be "devtools"?}
+\giovanni{is "usethis" supposed to be "devtools"?}
+\richel{"usethis" is the next "devtools", "devtools" will be deprecated in the future}
 To start using \verb;pirouette;, load its functions in the global namespace first:
@@
 \end{lstlisting}
+For the option to select inference models,
+\verb;pirouette; uses the "NS" BEAST2 package [Maturana et al.].
+It can be installed from within R, using:
+
+\begin{lstlisting}[language=R, floatplacement=H]
+install_beast2_pkg("NS")
+\end{lstlisting}
 
@@
 What is the error BEAST2 makes from a phylogeny using the same 
-speciation \giovanni{I'd use: "diversification"} model as it was generated by?
+diversification model as it was generated by?
 
@@
 out of four block of 250 nucleotides each, where the per-nucleotide
-mutation rate is on average \giovanni{I don't think it's correct to say that's "on average".
-Usually when you have a rate $\lambda$ you suppose to use a Poisson process generated by this rate: $\frac{\lambda^k e^\lambda}{k!}$.
-$\lambda$ turns out to be also the expected number of occurrences} 0.1 mutations per unit time:
+mutation rate is 0.1 mutations per unit time.
 
@@
 The second step is to select how an inference model is picked.
-We just select the generative model, which consists in a collection of three elements: a site model, a clock
+We just select the generative model, which consists in a collection of 
+three elements: a site model, a clock
 model of the alignment and a birth-death tree prior
@@
 
-% Note that this syntax does not work yet, see
-% https://github.com/richelbilderbeek/pirouette/issues/36
 \begin{lstlisting}[language=R, floatplacement=H]
@@
 
-Here we start with a tree \giovanni{generated} from an unknown speciation \giovanni{I'd use: "diversification"} model,
-that has four taxa and a crown age of 5:
+Here we start with a tree generated from an unknown 
+diversification model, that has four taxa and a crown age of 5:
 
@@
 The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny. We'll use the same parameters here:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
-)
-\end{lstlisting}
+given phylogeny. We'll re-use the alignment parameters of the previous example.
 
@@
 
-\giovanni{I noticed that "phylogeny", "alignment params" and "model select param" are needed for all the three research questions.
-Some times, though, you repeat yourself.
-In an article you want to be as concise as possible so I'd suggest to avoid repetitions.
-One solution could be to summarize all the parameter settings in a summary scheme}
-
 A third research question that \verb;pirouette; answers is:
@@
 
-Here we start again with a tree from an unknown speciation \giovanni{I'd use: "diversification"} model,
-that has four taxa and a crown age of 5:
-
-% Need idealised birth-death tree
-% From https://github.com/richelbilderbeek/pirouette_article/issues/1
-% and https://github.com/richelbilderbeek/pirouette/issues/35
-\begin{lstlisting}[language=R, floatplacement=H]
-phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4):1);")
-\end{lstlisting}
-
-The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny. We'll use the same parameters here:
-
-\begin{lstlisting}[language=R, floatplacement=H]
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
-)
-\end{lstlisting}
-
-The second step is to select how an inference model is picked.
-We will let the inference model that has the highest evidence be used
-in the Bayesian inference. By default, \verb;pirouette; will try
-all combinations of site models, clock models and tree priors.
-
-% Note that this syntax does not work yet, see
-% https://github.com/richelbilderbeek/pirouette/issues/36
-\begin{lstlisting}[language=R, floatplacement=H]
-model_select_param <- create_best_model_select_param()
-\end{lstlisting}
-
-Also here, the third and fourth step have sensible defaults.
+We use all the same settings as the previous model:
+phylogeny, alignment parameters, inference model selection,
+shared inference parameters and error measuring parameters.
 
@@
 has the same topology as the given tree, yet its branch lengths follow
-a birth-death distribution. Creating a twinning parameter is easy,
+an idealized birth-death distribution. Creating a twinning parameter is easy,
 as it has sensible default settings:

### 2019-01-24 article.tex
Introduction

@@
 Phylogenies are commonly used to explore evolutionary hypotheses.
+The mode of diversification is one such hypothesis.
+Phylogenies constructed from DNA alignments are commonly fit to
+diversification models, or the other way around. For example,
+[Etienne & Rosindell 2012] fit four bird phylogenies to the protracted
+diversification model.
+
+There are plenty of diversification models: constant-birth [Yule, 19..],
+birth-death [Nee et al., 19..], time-dependent [?], diversity 
+dependent [Etienne & ?, approx 2012], protracted 
+birth-death [Rosindell & Etienne, 2011][Etienne & Rosindell, 2012],
+multiple-birth death [Laudanno, Bilderbeek & Etienne, in preparation],
+a rate shifting birth-death model [the SLS model. Laudanno & Etienne, in preparation]
+punctuated equilibrium aka early burst [?Harmon],
+dependent on a binary trait [BISSE model, ?Maddison], 
+continuous trait [HISSE model],
+concealed state [?MUSSE] or a combination of all [secSSE model, Leonel].
+
+Diversification models are getting more complex over time. There are many
+reasons for this. One reason is the increase in computing power,
+allowing this heightened complexity to be possible.
+Another reason is to explore a process that is not
+yet captured by the other diversification models. An example of this
+is the multiple-birth model [Laudanno, Bilderbeek & Etienne, in preparation], 
+that is the first model to allow speciation to co-occur.
 
@@
 Additionally, \verb;pirouette; has a vignette that demonstrates extensively how
-to use it. There is documentation on the GitHub to get started, 
-with a dozen examples of BEAUti screenshots with equivalent \verb;pirouette; code.
-Finally, \verb;pirouette; has tutorial videos that can 
-be downloaded or viewed on YouTube, \url{https://goo.gl/weKaaU} \giovanni{be sure that shortened link are not deprecated after a certain time}.
+to use it.
 

### 2019-01-24 article.tex
Finish first draft of introduction

@@
 reasons for this. One reason is the increase in computing power,
-allowing this heightened complexity to be possible.
+which allows this heightened complexity possible.
 Another reason is to explore a process that is not
@@
 that is the first model to allow speciation to co-occur.
+
+An open question for each new diversification model is if the increased
+complexity is worth the extra computational resources. There are many
+ways to answer this question. One possible experiment 
+is to use a phylogeny of the new diversification model,
+simulate its alignment and see how well our
+current inference models are at retrieving the phylogeny.
+If this difference is small enough, current models apparently
+suffice.
+
+\verb;pirouette; is an R package that performs such an experiment,
+and is built on \verb;babette; [Bilderbeek & Etienne, 2018], 
+which calls the popular Bayesian inference tool 
+BEAST2 [X and Bouckaert et al., ????]. With \verb;pirouette;, one
+can easily measure the error made by Bayesian inference in recovering
+any given phylogeny, helping us to valorize new diversification models.
 

### 2019-01-29 article.tex
Use '\&' instead of '&', added fig 2

@@
 diversification models, or the other way around. For example,
-[Etienne & Rosindell 2012] fit four bird phylogenies to the protracted
+[Etienne \& Rosindell 2012] fit four bird phylogenies to the protracted
 diversification model.
@@
 birth-death [Nee et al., 19..], time-dependent [?], diversity 
-dependent [Etienne & ?, approx 2012], protracted 
-birth-death [Rosindell & Etienne, 2011][Etienne & Rosindell, 2012],
-multiple-birth death [Laudanno, Bilderbeek & Etienne, in preparation],
-a rate shifting birth-death model [the SLS model. Laudanno & Etienne, in preparation]
+dependent [Etienne \& ?, approx 2012], protracted 
+birth-death [Rosindell \& Etienne, 2011][Etienne \& Rosindell, 2012],
+multiple-birth death [Laudanno, Bilderbeek \& Etienne, in preparation],
+a rate shifting birth-death model [the SLS model. Laudanno \& Etienne, in preparation]
 punctuated equilibrium aka early burst [?Harmon],
@@
 yet captured by the other diversification models. An example of this
-is the multiple-birth model [Laudanno, Bilderbeek & Etienne, in preparation], 
+is the multiple-birth model [Laudanno, Bilderbeek \& Etienne, in preparation], 
 that is the first model to allow speciation to co-occur.
@@
 \verb;pirouette; is an R package that performs such an experiment,
-and is built on \verb;babette; [Bilderbeek & Etienne, 2018], 
+and is built on \verb;babette; [Bilderbeek \& Etienne, 2018], 
 which calls the popular Bayesian inference tool 
@@
 but also an custom error statistic is supplied, based
-on the absolute difference in the gamma statistic [Pybus & Harvy]. 
+on the absolute difference in the gamma statistic [Pybus \& Harvy]. 
 Additionally, the user can specify the
@@
 The resulting figure is shown in figure 2.
+
+\begin{figure}[h]
+  \includegraphics[width=\textwidth]{figure_2.png}
+\end{figure}
 

### 2019-01-29 article.tex
Show fig 2
Add fig 3
Added figure 4

@@
 
-% TODO: write script to create figure 2, see
-% https://github.com/richelbilderbeek/pirouette/issues/37
 The resulting figure is shown in figure 2.
+
+\richel{Update figure 2 to use idealized tree}
 
@@
   \includegraphics[width=\textwidth]{figure_2.png}
+  \caption{
+    The error BEAST2 makes from a phylogeny using the same 
+    diversification model as it was generated by
+  }
 \end{figure}
@@
 
-% TODO: write script to create figure 3, see
-% https://github.com/richelbilderbeek/pirouette/issues/38
 The resulting figure is shown in figure 3.
+
+\begin{figure}[h]
+  \includegraphics[width=\textwidth]{figure_3.png}
+  \caption{
+    The error BEAST2 makes from a phylogeny when
+    picking the best inference model
+  }
+\end{figure}
 
@@
 
-% TODO: write script to create figure 4, see
-% https://github.com/richelbilderbeek/pirouette/issues/39
 The resulting figure is shown in figure 4
+
+\begin{figure}[h]
+  \includegraphics[width=\textwidth]{figure_4.png}
+  \caption{
+    The error BEAST2 makes from a phylogeny 
+    picking the best inference model, compared to the background noise?
+    Here, the twin column shows the error BEAST2 makes on an idealized
+    tree, to measure the noise, which is the minimal error.
+  }
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-01-29 article.tex
Updates from ShareLaTeX

--- 2019-01-24 article.tex
@@
 diversification models, or the other way around. For example,
-[Etienne & Rosindell 2012] fit four bird phylogenies to the protracted
+[Etienne $\&$ Rosindell 2012] fit four bird phylogenies to the protracted
 diversification model.
+\giovanni{ALTERNATIVE VERSION: The development of new powerful inference tools, such as BEAST [\cite{drummond2007beast}] or MrBayes [\cite{huelsenbeck2001mrbayes}], provided us the possibility to build phylogenetic trees from genetic material extracted from extant organisms.
+This has constituted a big step forward in the general understanding on how species evolved (this can be improved).
+Having access to this new kind of data has given rise to a high interest in understanding what are the main drivers and modes for diversification processes.
+}
 
 There are plenty of diversification models: constant-birth [Yule, 19..],
-birth-death [Nee et al., 19..], time-dependent [?], diversity 
-dependent [Etienne & ?, approx 2012], protracted 
-birth-death [Rosindell & Etienne, 2011][Etienne & Rosindell, 2012],
-multiple-birth death [Laudanno, Bilderbeek & Etienne, in preparation],
-a rate shifting birth-death model [the SLS model. Laudanno & Etienne, in preparation]
+birth-death [\cite{nee1994reconstructed}], time-dependent [?], diversity 
+dependent [\cite{etienne2011diversity}], protracted 
+birth-death [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
+multiple-birth death [Laudanno, Bilderbeek, Haegeman $\&$ Etienne, in preparation],
+a rate shifting birth-death model [the SLS model. Laudanno, Haegeman $\&$ Etienne, in preparation]
 punctuated equilibrium aka early burst [?Harmon],
-dependent on a binary trait [BISSE model, ?Maddison], 
-continuous trait [HISSE model],
-concealed state [?MUSSE] or a combination of all [secSSE model, Leonel].
-
+dependent on a binary trait [\cite{maddison2007estimating}], 
+multiple state trait [\cite{fitzjohn2012diversitree}],
+concealed state [\cite{beaulieu2016detecting}] or even a combination of multiple concealed and observable states [\cite{herrera2018detecting}].
+\giovanni{
+@Richel: I am wondering if it might be worth to give a brief explanation on the philosophy of the aforementioned models, to explain why we want to simulate the processes.
+In case, this could be a stub for it:
+Such models usually rely on the assumption that a particular diversification process can be explained mainly by specific mechanisms.
+Such mechanisms are built into likelihood functions, which provide the likelihood of the parameters, given the data.
+By maximizing the likelihood is possible to give the best inference of such parameters.
+It is therefore a standard practice, to validate such models, to test them versus simulated processes generated according to the hypothesised mechanisms.
+}
 Diversification models are getting more complex over time. There are many
@@
 yet captured by the other diversification models. An example of this
-is the multiple-birth model [Laudanno, Bilderbeek & Etienne, in preparation], 
+is the multiple-birth model [Laudanno, Bilderbeek, Haegeman $\&$ Etienne, in preparation], 
 that is the first model to allow speciation to co-occur.
 
-An open question for each new diversification model is if the increased
-complexity is worth the extra computational resources. There are many
+An open question for each new diversification model is whether the increased
+complexity is worth the extra computational resources \giovanni{I think here it is not only a matter of computational resources.
+I believe there two more factors: 1) developing a likelihood model takes time, human resources and therefore fundings; 2) in principle you can elaborate any model to be as complex as possibly requested, however there is a diminishing return on this.
+In fact the purpose of modelling for complex systems such as biology is to hypothesize the main driving factors of a process and trying to prove that they are indeed able to explain the main features of the phenomenon you want to explain.
+In my opinion the job of a modeller is to try to disentangle the intrinsic complexity of a system highlighting the main reasons of its behaviour.
+In such a light you don't want your model to be unnecessary complex.}.
+There are many
 ways to answer this question. One possible experiment 
@@
 \verb;pirouette; is an R package that performs such an experiment,
-and is built on \verb;babette; [Bilderbeek & Etienne, 2018], 
+and is built on \verb;babette; [Bilderbeek $\&$ Etienne, 2018], 
 which calls the popular Bayesian inference tool 
-BEAST2 [X and Bouckaert et al., ????]. With \verb;pirouette;, one
+BEAST2 [\cite{bouckaert2014beast}]. With \verb;pirouette;, one
 can easily measure the error made by Bayesian inference in recovering
@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given phylogeny of which full information is provided.
-This involves multiple steps, described in detail below.
+makes from a given phylogeny of which full information is provided \giovanni{is this true?
+do we also provide information about extinct species or we just use reconstructed tree (which always lack a lot of information)?}.
 
-The pipeline is summarized by the following steps:
+The pipeline is summarized by the following steps, which will be described in detail below:
 \begin{enumerate}
     \item from a given phylogeny an alignment is simulated;
+    \item from the alignment an inference model is chosen;
     \item the alignment is used as BEAST2 input to infer a posterior distribution of phylogenies;
@@
 phylogeny using an error statistic, which is the nLTT 
-statistic \cite{janzen2015approximate} by default,
-but also an custom error statistic is supplied, based
-on the absolute difference in the gamma statistic [Pybus & Harvy]. 
+statistic (\cite{janzen2015approximate}) by default,
+but also a custom error statistic is supplied, based
+on the absolute difference in the gamma statistic [Pybus $\&$ Harvy]. 
 Additionally, the user can specify the
 proportion of posterior phylogenies to 
-discard (i.e. the burn-in), throwing away the first 10%
+discard (i.e. the burn-in), throwing away the first $10\%$
 of all phylogenies by default. This burn-in is used to discard
@@
 \end{lstlisting}
-\giovanni{is "usethis" supposed to be "devtools"?}
-\richel{"usethis" is the next "devtools", "devtools" will be deprecated in the future}
+
 To start using \verb;pirouette;, load its functions in the global namespace first:
@@
 We just select the generative model, which consists in a collection of 
-three elements: a site model, a clock
+three elements: a site model \giovanni{do we have to explain what a site model is or what are the options?}, a clock
 model of the alignment and a birth-death tree prior
-to be the generative tree prior.
+to be the generative tree prior. \giovanni{would it be helpful to the reader to explicit that site and clock models are automatically taken from the generative model?}
 
@@
 The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny. We'll re-use the alignment parameters of the previous example.
+given phylogeny. We'll re-use the alignment parameters of the previous example \giovanni{it might be nice to add an explicit ref to the alignmentparams code}.
 
@@
 phylogeny, alignment parameters, inference model selection,
-shared inference parameters and error measuring parameters.
+shared inference parameters and error measuring parameters \giovanni{it might be nice to add an explicit ref to the lines of code we are referring to}.
 

### 2019-01-29 article.tex

@@
 
-% TODO: write script to create figure 2, see
-% https://github.com/richelbilderbeek/pirouette/issues/37
 The resulting figure is shown in figure 2.
+
+\richel{Update figure 2 to use idealized tree}
+
+\begin{figure}[h]
+  \includegraphics[width=\textwidth]{figure_2.png}
+  \caption{
+    The error BEAST2 makes from a phylogeny using the same 
+    diversification model as it was generated by
+  }
+\end{figure}
 

### 2019-01-29 article.tex
Use '\&' instead of '$\&$'

@@
 diversification models, or the other way around. For example,
-[Etienne $\&$ Rosindell 2012] fit four bird phylogenies to the protracted
+[Etienne \& Rosindell 2012] fit four bird phylogenies to the protracted
 diversification model.
@@
 birth-death [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
-multiple-birth death [Laudanno, Bilderbeek, Haegeman $\&$ Etienne, in preparation],
-a rate shifting birth-death model [the SLS model. Laudanno, Haegeman $\&$ Etienne, in preparation]
+multiple-birth death [Laudanno, Bilderbeek, Haegeman \& Etienne, in preparation],
+a rate shifting birth-death model [the SLS model. Laudanno, Haegeman \& Etienne, in preparation]
 punctuated equilibrium aka early burst [?Harmon],
@@
 yet captured by the other diversification models. An example of this
-is the multiple-birth model [Laudanno, Bilderbeek, Haegeman $\&$ Etienne, in preparation], 
+is the multiple-birth model [Laudanno, Bilderbeek, Haegeman \& Etienne, in preparation], 
 that is the first model to allow speciation to co-occur.
@@
 \verb;pirouette; is an R package that performs such an experiment,
-and is built on \verb;babette; [Bilderbeek $\&$ Etienne, 2018], 
+and is built on \verb;babette; [Bilderbeek \& Etienne, 2018], 
 which calls the popular Bayesian inference tool 
@@
 but also a custom error statistic is supplied, based
-on the absolute difference in the gamma statistic [Pybus $\&$ Harvy]. 
+on the absolute difference in the gamma statistic [Pybus \& Harvy]. 
 Additionally, the user can specify the

### 2019-01-29 article.tex
Use intro and other paragraph of @Giappo

@@
 
-Phylogenies are commonly used to explore evolutionary hypotheses.
-The mode of diversification is one such hypothesis.
-Phylogenies constructed from DNA alignments are commonly fit to
-diversification models, or the other way around. For example,
-[Etienne \& Rosindell 2012] fit four bird phylogenies to the protracted
-diversification model.
-\giovanni{ALTERNATIVE VERSION: The development of new powerful inference tools, such as BEAST [\cite{drummond2007beast}] or MrBayes [\cite{huelsenbeck2001mrbayes}], provided us the possibility to build phylogenetic trees from genetic material extracted from extant organisms.
-This has constituted a big step forward in the general understanding on how species evolved (this can be improved).
-Having access to this new kind of data has given rise to a high interest in understanding what are the main drivers and modes for diversification processes.
-}
+The development of new powerful inference tools, 
+such as BEAST [\cite{drummond2007beast}] or MrBayes [\cite{huelsenbeck2001mrbayes}], 
+provides us the possibility to build phylogenetic trees 
+from genetic material extracted from extant organisms.
+This has constituted a big step forward 
+in doing experiments to elucidate on how species 
+evolve.
+Having access to this new kind of tools has given rise 
+to a high interest in understanding what are the main drivers 
+and modes for diversification processes.
 
-There are plenty of diversification models: constant-birth [Yule, 19..],
-birth-death [\cite{nee1994reconstructed}], time-dependent [?], diversity 
-dependent [\cite{etienne2011diversity}], protracted 
-birth-death [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
+There are plenty of diversification models: 
+constant-birth [Yule, 19..],
+birth-death [\cite{nee1994reconstructed}], 
+time-dependent [Rabosky D., LASER package?], 
+diversity dependent [\cite{etienne2011diversity}], 
+protracted birth-death [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
 multiple-birth death [Laudanno, Bilderbeek, Haegeman \& Etienne, in preparation],
@@
 multiple state trait [\cite{fitzjohn2012diversitree}],
-concealed state [\cite{beaulieu2016detecting}] or even a combination of multiple concealed and observable states [\cite{herrera2018detecting}].
-\giovanni{
-@Richel: I am wondering if it might be worth to give a brief explanation on the philosophy of the aforementioned models, to explain why we want to simulate the processes.
-In case, this could be a stub for it:
-Such models usually rely on the assumption that a particular diversification process can be explained mainly by specific mechanisms.
-Such mechanisms are built into likelihood functions, which provide the likelihood of the parameters, given the data.
-By maximizing the likelihood is possible to give the best inference of such parameters.
-It is therefore a standard practice, to validate such models, to test them versus simulated processes generated according to the hypothesised mechanisms.
-}
+concealed state [\cite{beaulieu2016detecting}] 
+or even a combination of multiple concealed and observable states [\cite{herrera2018detecting}].
+
+Such models usually rely on the assumption 
+that a particular diversification process 
+can be explained mainly by a specific mechanism.
+Such a mechanism is built into a likelihood function, 
+which provides the likelihood of the parameters, given the data.
+By maximizing the likelihood, it is possible 
+to estimate the most likely model parameters. 
+It is a standard practice in assessing such models,
+how well they estimate their own parameters
+on simulated phylogenies in which the parameters are known.
+
 Diversification models are getting more complex over time. There are many

### 2019-01-29 article.tex
Use another paragraph of @Giappo

@@
 
-An open question for each new diversification model is whether the increased
-complexity is worth the extra computational resources \giovanni{I think here it is not only a matter of computational resources.
-I believe there two more factors: 1) developing a likelihood model takes time, human resources and therefore fundings; 2) in principle you can elaborate any model to be as complex as possibly requested, however there is a diminishing return on this.
-In fact the purpose of modelling for complex systems such as biology is to hypothesize the main driving factors of a process and trying to prove that they are indeed able to explain the main features of the phenomenon you want to explain.
-In my opinion the job of a modeller is to try to disentangle the intrinsic complexity of a system highlighting the main reasons of its behaviour.
-In such a light you don't want your model to be unnecessary complex.}.
-There are many
-ways to answer this question. One possible experiment 
+An open question for each new diversification model 
+is whether the increased complexity is worth it. 
+One of the purposes of modelling a complex systems
+is to hypothesize the main driving factors of a process 
+and trying to prove that they are indeed able 
+to explain the main features of the phenomenon you want to explain.
+Due to this, there are diminishing returns on adding complexity
+to a model: the job of a modeller is to try to 
+disentangle the intrinsic complexity of a system 
+highlighting the main reasons of its behaviour. 
+In such a light you don't want your model to be unnecessary complex.
+Additionally, developing a likelihood model takes time, 
+human resources, fundings and can strain computational resources.
+
+When assuming that a new diversification model is worth the
+(philosophical) effort, there are many ways to measure
+if the new model indeed gives a better explanation. 
+One possible experiment 
 is to use a phylogeny of the new diversification model,

### 2019-01-29 article.tex
Processed feedback, thanks @Giappo, fix https://github.com/richelbilderbeek/pirouette/issues/58

@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given phylogeny of which full information is provided \giovanni{is this true?
-do we also provide information about extinct species or we just use reconstructed tree (which always lack a lot of information)?}.
+makes from a given reconstructed phylogeny.
 
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 install.packages("pirouette")
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 usethis::install_github("richelbilderbeek/pirouette")
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 library(pirouette)
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 install_beast2()
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 install_beast2_pkg("NS")
@@
 % and https://github.com/richelbilderbeek/pirouette/issues/35
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4):1);")
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 alignment_params <- create_alignment_params(
@@
 
+By default, an alignment is created using a Jukes-Cantor (JC) site model
+and a strict clock model. A JC site model assumes that mutations
+between all nucloetides are constant at an equal rate. 
+A strict clock model assumes that
+the mutation rate of all species is equal and constant.
+Due to this, we state that the generative model for the alignment is
+a JC site model and a strict clock model.
+
 The second step is to select how an inference model is picked.
 We just select the generative model, which consists in a collection of 
-three elements: a site model \giovanni{do we have to explain what a site model is or what are the options?}, a clock
-model of the alignment and a birth-death tree prior
-to be the generative tree prior. \giovanni{would it be helpful to the reader to explicit that site and clock models are automatically taken from the generative model?}
+three elements: a site model, clock model and tree prior.
+If we know how the generative model of the alignment, as we did above,
+we can re-use that information. We do still need to specify 
+a generative tree prior, which is the tree prior underlying the phylogeny.
+In this case, we'll pick a (constant-rate) birth-death model:
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 model_select_param <- create_gen_model_select_param(
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 errors <- pir_run(
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 pir_plot(errors)
@@
 The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny. We'll re-use the alignment parameters of the previous example \giovanni{it might be nice to add an explicit ref to the alignmentparams code}.
+given phylogeny. We'll re-use the alignment parameters of the previous example 
+\richel{TODO: add ref to code}
 
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 model_select_param <- create_best_model_select_param()
@@
 phylogeny, alignment parameters, inference model selection,
-shared inference parameters and error measuring parameters \giovanni{it might be nice to add an explicit ref to the lines of code we are referring to}.
+shared inference parameters and error measuring parameters 
+\richel{TODO: add ref to code}
 
@@
 
-\begin{lstlisting}[language=R, floatplacement=H]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 twinning_params <- create_twinning_params()

### 2019-02-19 article.tex
Updates from Overleaf

@@
 from genetic material extracted from extant organisms.
-This has constituted a big step forward 
-in doing experiments to elucidate on how species 
+This has constituted an important step forward 
+in our understanding on how species 
 evolve.
-Having access to this new kind of tools has given rise 
-to a high interest in understanding what are the main drivers 
+Such tools have been increasingly exploited to hypothesize and test what are the main drivers
 and modes for diversification processes.
@@
 a rate shifting birth-death model [the SLS model. Laudanno, Haegeman \& Etienne, in preparation]
-punctuated equilibrium aka early burst [?Harmon],
+punctuated equilibrium (as known as "early burst") [?Harmon \giovanni{is it this one? https://onlinelibrary.wiley.com/doi/full/10.1111/j.1558-5646.2010.01025.x}],
 dependent on a binary trait [\cite{maddison2007estimating}], 
@@
 which provides the likelihood of the parameters, given the data.
-By maximizing the likelihood, it is possible 
-to estimate the most likely model parameters. 
+By maximizing it, it is possible 
+to estimate the what are the best values for model parameters. 
 It is a standard practice in assessing such models,
-how well they estimate their own parameters
-on simulated phylogenies in which the parameters are known.
+to verify the level of accuracy they can achieve
+on simulated phylogenies, of which full knowledge is given.
 
@@
 which allows this heightened complexity possible.
-Another reason is to explore a process that is not
-yet captured by the other diversification models. An example of this
-is the multiple-birth model [Laudanno, Bilderbeek, Haegeman \& Etienne, in preparation], 
-that is the first model to allow speciation to co-occur.
+Another reason is to explore a process that are not yet considered by currently existing diversification models. An example of this
+is the multiple-birth model [Laudanno et al., in preparation], 
+that is the first model to allow simultaneous speciations to co-occur.
 
-An open question for each new diversification model 
-is whether the increased complexity is worth it. 
+However, whenever a new model is proposed there is always a trade-off between the cost of increased complexity versus the benefits that it can provide.
+Therefore desirable models have to be able to disentangle the intrinsic complexity of a system, 
+highlight the main reasons of its behaviour and avoid unnecessary complexity. Furthermore developing models can cost time, human and computational resources.
+
+\iffalse
 One of the purposes of modelling a complex systems
@@
 human resources, fundings and can strain computational resources.
+\fi
 
+For this reason we believe it is important to have techniques to quantitatively assess the advancement provided by new models, in comparison to the existing ones.
+
+\iffalse
 When assuming that a new diversification model is worth the
@@
 if the new model indeed gives a better explanation. 
-One possible experiment 
-is to use a phylogeny of the new diversification model,
-simulate its alignment and see how well our
-current inference models are at retrieving the phylogeny.
+\fi
+
+To achieve this, one possible experiment
+is to simulate a phylogeny according to the mechanisms proposed by the new diversification model, then simulate its alignment and measure how well current inference models perform at retrieving the original phylogeny.
 If this difference is small enough, current models apparently
@@
 can easily measure the error made by Bayesian inference in recovering
-any given phylogeny, helping us to valorize new diversification models.
+any given phylogeny, helping us to validate new diversification models.
 
@@
     \item the alignment is used as BEAST2 input to infer a posterior distribution of phylogenies;
-    \item posterior phylogenies are compared with the given phylogenies to estimate the error we make. The bigger the match, the lower the error;
+    \item posterior phylogenies are compared with the given original phylogeny to estimate the error we make. The bigger the match, the lower the error;
 \end{enumerate}
+\giovanni{Here it could be useful to add a figure that explains the pipeline, a bit like the one I have on my desk}
 There is also the option to generate a 'twin tree',
@@
 model, the alignment is used to find the inference model that has the
-highest evidence (i.e. marginal likelihood) from a set of inference models.
+highest evidence (i.e. marginal likelihood) from a set of candidate inference models.
 The evidence of an inference model is estimated using a nested sampling (NS)
 approach, as described in \cite{maturana2018model}. The nested sampling is
-performed by \verb;babette;, that call the 'NS' BEAST2 package. 
+performed by \verb;babette;\giovanni{babette or mcbette?}, that calls the 'NS' BEAST2 package. 
 Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; 
@@
 Bayesian posterior. These phylogenies are compared to the given
-phylogeny using an error statistic, which is the nLTT 
+original phylogeny using an error statistic, which is the nLTT 
 statistic (\cite{janzen2015approximate}) by default,
@@
 For the option to select inference models,
-\verb;pirouette; uses the "NS" BEAST2 package [Maturana et al.].
+\verb;pirouette; uses the "NS" BEAST2 package [\cite{maturana2018model}].
 It can be installed from within R, using:
@@
 three elements: a site model, clock model and tree prior.
+\giovanni{Here you present the "alignment generative model" and the "inference model".
+These are general concepts, as they are used in all the experiments.
+I propose to define them from the start, for example in the "description" section; we can even refer to them in the figure I propose to add, so to make it clearer to the reader.
+It can be done in this way:
+A "model" is always defined as the collection of three elements: a site model, a clock model and a tree prior.
+We have two kind of models: generative and inference.
+The first is used to simulate alignments, the second is used to produce a posterior.
+Generative models always use a (constant-rate) birth-death tree prior.}
 If we know how the generative model of the alignment, as we did above,
@@
 \richel{Update figure 2 to use idealized tree}
+\giovanni{Check "pirouettepackage" repo. I uploaded a function to create the idealized tree that you might like.}
 
@@
     diversification model as it was generated by
+    \giovanni{If we define generative and inference model from the start we can rephrase this caption in a better way, referring to terms that the reader is already familiar to}
   }
@@
 The second step is to select how an inference model is picked.
-We will let the inference model that has the highest evidence be used
+We will let the inference model that has the highest evidence to be used
 in the Bayesian inference. By default, \verb;pirouette; will try
-all combinations of site models, clock models and tree priors.
+all combinations of site models, clock models and tree priors.\giovanni{again, if we give a name to this triplet, it could be good to call it by its name.
+We can still recall its composition in brackets, if we think it could be useful}
 
@@
 
-We use all the same settings as the previous model:
+The following settings are the same as in the previous section:
 phylogeny, alignment parameters, inference model selection,
@@
 has the same topology as the given tree, yet its branch lengths follow
-an idealized birth-death distribution. Creating a twinning parameter is easy,
+an idealized \giovanni{I don't think this is true.
+A twin tree is generated randomly.
+If we use the method "maxlikelihood" (see the function "createbdtree") the tree will be chosen in such a way it will maximizing the likelihood with the parameters we estimated from the original tree, but I would not call it "idealized".
+Furthermore I think it's not obvious to use the "maxlikelihood" method, as in general the "ideal" scenario should be established using a distribution of trees.
+A single tree is always the result of a stochastic process.} birth-death distribution.
+Creating a twinning parameter is easy,
 as it has sensible default settings:
@@
     Here, the twin column shows the error BEAST2 makes on an idealized
-    tree, to measure the noise, which is the minimal error.
+    tree, to measure the noise, which is the minimal error. \giovanni{again, I don't believe that using an "idealized" tree is the best thing to do. I would rather use more "true trees".}
   }
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+\giovanni{We need to discuss the results we got from the 3 experiments.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-02-20 article.tex

@@
 
-We start with an idealised birth-death tree of five taxa and a crown age of 10:
+We start with an idealised \giovanni{I still don't get why we need an "idealized" tree.} birth-death tree of five taxa and a crown age of 10:
 
@@
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]
-phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4):1);")
+phylogeny <- ape::read.tree(text = "((A:7.1,(B:2.2,C:2.2):4.9):2.9,(D:4.6,E:4.6):5.4):0;")
 \end{lstlisting}
+
+\begin{figure}[h]
+  \includegraphics[width=\textwidth]{figure_bd.png}
+  \caption{pippo}
+\end{figure}
 
@@
 A twin tree is generated randomly.
-If we use the method "maxlikelihood" (see the function "createbdtree") the tree will be chosen in such a way it will maximizing the likelihood with the parameters we estimated from the original tree, but I would not call it "idealized".
+If we use the method "maxlikelihood" (see the function "createbdtree") the tree will be chosen in such a way it will maximizing the likelihood given the parameters we estimated from the original tree, but I would not call it "idealized".
 Furthermore I think it's not obvious to use the "maxlikelihood" method, as in general the "ideal" scenario should be established using a distribution of trees.

### 2019-02-20 article.tex
Process feedback @Giappo

@@
 a rate shifting birth-death model [the SLS model. Laudanno, Haegeman \& Etienne, in preparation]
-punctuated equilibrium (as known as "early burst") [?Harmon \giovanni{is it this one? https://onlinelibrary.wiley.com/doi/full/10.1111/j.1558-5646.2010.01025.x}],
+punctuated equilibrium (as known as "early burst") [\cite{harmon2010early}],
 dependent on a binary trait [\cite{maddison2007estimating}], 
@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given reconstructed phylogeny.
+makes from a given reconstructed phylogeny. For this measurement,
+two models need to be specified. Both models are defined as a combination of 
+site model, clock model and a tree prior. The two kind of models are a 
+generative and inference model. The generative model is used to simulate 
+alignments, the inference model is used to produce a posterior. 
+Generative models always use a (constant-rate) birth-death tree prior.
 
@@
 \end{enumerate}
-\giovanni{Here it could be useful to add a figure that explains the pipeline, a bit like the one I have on my desk}
+\richel{figure to pipeline here}
 There is also the option to generate a 'twin tree',
@@
 approach, as described in \cite{maturana2018model}. The nested sampling is
-performed by \verb;babette;\giovanni{babette or mcbette?}, that calls the 'NS' BEAST2 package. 
+performed by \verb;mcbette; \cite{mcbette}, that calls the 'NS' BEAST2 package. 
 Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; 
@@
 
-We start with an idealised \giovanni{I still don't get why we need an "idealized" tree.} birth-death tree of five taxa and a crown age of 10:
+We start with an idealised 
+birth-death tree of five taxa and a crown age of 10.
+By idealized, we mean a phylogeny that is easiest to recover
+by the Bayesian inference. A phylogeny that is easy to recover has
+the most common branch length distribution imaginable. We
+use an idealised tree, as we want to measure the baseline noise; that is,
+the lowest error that BEAST2 can possibly make.
 
-% Need idealised birth-death tree
-% From https://github.com/richelbilderbeek/pirouette_article/issues/1
-% and https://github.com/richelbilderbeek/pirouette/issues/35
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]
@@
 three elements: a site model, clock model and tree prior. 
-\giovanni{Here you present the "alignment generative model" and the "inference model".
-These are general concepts, as they are used in all the experiments.
-I propose to define them from the start, for example in the "description" section; we can even refer to them in the figure I propose to add, so to make it clearer to the reader.
-It can be done in this way:
-A "model" is always defined as the collection of three elements: a site model, a clock model and a tree prior.
-We have two kind of models: generative and inference.
-The first is used to simulate alignments, the second is used to produce a posterior.
-Generative models always use a (constant-rate) birth-death tree prior.}
 If we know how the generative model of the alignment, as we did above,
@@
 \richel{Update figure 2 to use idealized tree}
-\giovanni{Check "pirouettepackage" repo. I uploaded a function to create the idealized tree that you might like.}
+\giovanni{
+  Check "pirouettepackage" repo. 
+  I uploaded a function to create the idealized tree that you might like.
+}
 
@@
   \caption{
-    The error BEAST2 makes from a phylogeny using the same 
-    diversification model as it was generated by
-    \giovanni{If we define generative and inference model from the start we can rephrase this caption in a better way, referring to terms that the reader is already familiar to}
+    The error BEAST2 makes from a phylogeny 
+    when the generative and inference model are the same
   }
@@
 We will let the inference model that has the highest evidence to be used
-in the Bayesian inference. By default, \verb;pirouette; will try
-all combinations of site models, clock models and tree priors.\giovanni{again, if we give a name to this triplet, it could be good to call it by its name.
-We can still recall its composition in brackets, if we think it could be useful}
+in the Bayesian inference. By default, \verb;pirouette; will use
+its full arsenal of 40 inference models, which are all combinations of 4 site 
+models, 2 clock models and 5 tree priors.
 
@@
 has the same topology as the given tree, yet its branch lengths follow
-an idealized \giovanni{I don't think this is true.
-A twin tree is generated randomly.
-If we use the method "maxlikelihood" (see the function "createbdtree") the tree will be chosen in such a way it will maximizing the likelihood given the parameters we estimated from the original tree, but I would not call it "idealized".
-Furthermore I think it's not obvious to use the "maxlikelihood" method, as in general the "ideal" scenario should be established using a distribution of trees.
-A single tree is always the result of a stochastic process.} birth-death distribution.
-Creating a twinning parameter is easy,
+an idealized 
+\giovanni{
+  I don't think this is true. A twin tree is generated randomly. 
+  If we use the method "maxlikelihood" (see the function "createbdtree") 
+  the tree will be chosen in such a way it will maximizing the likelihood 
+  given the parameters we estimated from the original tree, 
+  but I would not call it "idealized". Furthermore I think it's not obvious to 
+  use the "maxlikelihood" method, as in general the "ideal" scenario should be 
+  established using a distribution of trees. A single tree is always the result 
+  of a stochastic process.
+}
+\richel{we'll discuss this today} 
+birth-death distribution. Creating a twinning parameter is easy,
 as it has sensible default settings:
@@
     Here, the twin column shows the error BEAST2 makes on an idealized
-    tree, to measure the noise, which is the minimal error. \giovanni{again, I don't believe that using an "idealized" tree is the best thing to do. I would rather use more "true trees".}
+    tree, to measure the noise, which is the minimal error. 
+    \giovanni{again, I don't believe that using an "idealized" tree is the best 
+      thing to do. I would rather use more "true trees".}
+    \richel{we'll discuss this today}
   }
@@
 \section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \giovanni{We need to discuss the results we got from the 3 experiments.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\richel{Only iff the journal requires this}
 

### 2019-02-20 article.tex
Reword model definitions

@@
 makes from a given reconstructed phylogeny. For this measurement,
-two models need to be specified. Both models are defined as a combination of 
-site model, clock model and a tree prior. The two kind of models are a 
-generative and inference model. The generative model is used to simulate 
-alignments, the inference model is used to produce a posterior. 
-Generative models always use a (constant-rate) birth-death tree prior.
+two models need to be specified: a generative and an inference model.
+A generative model is a combination of site model, clock model and tree prior,
+used to simulate alignments. An inference model is a combination of 
+site model, clock model, tree prior, an optional node calibration prior
+and setup of the Markov chain Monte Carlo (MCMC) algorithm, 
+used to produce a posterior. Node calibration priors, called 'most
+recent common ancestor' (MRCA) priors within BEAST2 allow, among others,
+to specify a crown age.
 

### 2019-02-20 article.tex
Reword MCRA prior

@@
 used to produce a posterior. Node calibration priors, called 'most
-recent common ancestor' (MRCA) priors within BEAST2 allow, among others,
-to specify a crown age.
+recent common ancestor' (MRCA) priors within BEAST2 allow to create a 
+time-calibrated phylogeny, by specifying a normal distribution around
+an expected time when two or more taxa's ancestors diversified.
 

### 2019-02-21 article.tex
Show experiment setup of first research question

@@
 
-The second step is to select how an inference model is picked.
-We just select the generative model, which consists in a collection of 
-three elements: a site model, clock model and tree prior. 
+In the second step we state our experiments. In this context, we
+define an experiment as a combination of an inference model
+and the conditions under which a Bayesian inference is created.
+
+\begin{table}
+  \begin{tabular}{ | c | c | c | l | }
+    \hline
+    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & \textbf{inference model} \\ 
+    \hline
+    generative & always & no & generative model \\
+    \hline
+  \end{tabular}
+  \caption{Experimental setup to answer the first research question}
+\end{table}
+
 If we know how the generative model of the alignment, as we did above,

### 2019-02-21 article.tex
Example 2 finished writing, still need to create figure for it

@@
 
-We start with an idealised 
-birth-death tree of five taxa and a crown age of 10.
+We start with an idealised Yule (pure-birth) tree 
+with five taxa and a crown age of ten.
 By idealized, we mean a phylogeny that is easiest to recover
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-phylogeny <- ape::read.tree(text = "((A:7.1,(B:2.2,C:2.2):4.9):2.9,(D:4.6,E:4.6):5.4):0;")
+\begin{lstlisting}[
+    language=R, floatplacement=H, frame=single, 
+    label = {lst:create_ideal_tree}, 
+    caption = {Create an ideal tree}
+  ]
+phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
 \end{lstlisting}
@@
   \includegraphics[width=\textwidth]{figure_bd.png}
-  \caption{pippo}
+  \caption{The ideal tree, as created by listing \ref{lst:create_ideal_tree}}
 \end{figure}
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[
+    language=R, floatplacement=H, 
+    frame=single
+    label = {lst:create_alignment}, 
+    caption = {Create an alignment}
+  ]
 alignment_params <- create_alignment_params(
@@
 and a strict clock model. A JC site model assumes that mutations
-between all nucloetides are constant at an equal rate. 
+between all nucleotides are constant at an equal rate. 
 A strict clock model assumes that
@@
 define an experiment as a combination of an inference model
-and the conditions under which a Bayesian inference is created.
+and the conditions under which a Bayesian inference is executed.
+Within this example, we specify that our experiment uses the
+generative model, will always be run and we ignore the evidence for our model,
+for the inference model that is also the generative model (using a
+JC site model, strict clock model and Yule tree prior).
 
@@
     \hline
-    generative & always & no & generative model \\
+    generative & always & no & JC, strict, Yule \\
     \hline
   \end{tabular}
-  \caption{Experimental setup to answer the first research question}
+  \caption{
+    Experimental setup to answer the first research question.
+    JC: Jukes-Cantor site model.
+    strict: strict clock model.
+    Yule: Yule (pure-birth) tree prior
+  }
 \end{table}
 
-If we know how the generative model of the alignment, as we did above,
-we can re-use that information. We do still need to specify 
-a generative tree prior, which is the tree prior underlying the phylogeny.
-In this case, we'll pick a (constant-rate) birth-death model:
+Due to sensible defaults, specifying this
+results in:
 
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]
-model_select_param <- create_gen_model_select_param(
+experiments <- list(create_experiment())
+\end{lstlisting}
+
+All the \verb;pirouette; arguments are bundled
+and checked by \verb;create_pir_params;:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+pir_params <- create_pir_params(
   alignment_params = alignment_params,
-  tree_prior = create_bd_tree_prior()
+  experiments = experiments
 )
@@
 
-The third and fourth step have sensible defaults, and we are not
-interested in using a twin tree. We can now measure the errors BEAST2
-makes when inferring the given phylogeny in this way:
+Running the experiment:
 
@@
   phylogeny,
-  alignment_params = alignment_params,
-  model_select_params = model_select_param
+  pir_params = pir_params
 )
@@
 
-The resulting figure is shown in figure 2.
-
-\richel{Update figure 2 to use idealized tree}
-\giovanni{
-  Check "pirouettepackage" repo. 
-  I uploaded a function to create the idealized tree that you might like.
-}
+The resulting figure is shown in figure \ref{fig:example_1}.
 
@@
   }
+  \label{fig:example_1}
 \end{figure}
@@
 Here we start with a tree generated from an unknown 
-diversification model, that has four taxa and a crown age of 5:
+diversification model, that has four taxa and a crown age of five:
 
-% Need idealised birth-death tree
-% From https://github.com/richelbilderbeek/pirouette_article/issues/1
-% and https://github.com/richelbilderbeek/pirouette/issues/35
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[
+  language=R, 
+  floatplacement=H, 
+  frame=single, 
+  label = {lst:unknown_phylogeny},
+  caption = "A phylogeny with an unknown diversification model"
+]
 phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4):1);")
@@
 given phylogeny. We'll re-use the alignment parameters of the previous example
-\richel{TODO: add ref to code}
+as shown in listing \ref{lst:create_alignment}.
 
-The second step is to select how an inference model is picked.
-We will let the inference model that has the highest evidence to be used
-in the Bayesian inference. By default, \verb;pirouette; will use
-its full arsenal of 40 inference models, which are all combinations of 4 site 
+\begin{table}
+  \begin{tabular}{ | c | c | c | l | }
+    \hline
+    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & \textbf{inference model} \\ 
+    \hline
+    candidate & best candidate & yes & JC, strict, Yule \\
+    candidate & best candidate & yes & JC, strict, BD \\
+    ...       & ...            & ... & ... \\
+    candidate & best candidate & yes & GTR, RLN, CCP \\
+    candidate & best candidate & yes & GTR, RLN, CEP \\
+    \hline
+  \end{tabular}
+  \caption{
+    Experimental setup to answer the second research question.
+    JC: Jukes-Cantor site model.
+    strict: strict clock model.
+    Yule: Yule (pure-birth) tree prior.
+    BD: birth-death tree prior.
+    GTR: GTR site model.
+    RLN: relaxed log-normal clock model.
+    CCP: coalescent constant-population tree prior.
+    CEP: coalescent exponential-population tree prior.
+  }
+\end{table}
+
+In the second step we state our experiments. 
+Within this example, all our experiments are candidates,
+run only if it is the best candidate, the evidence is measured (ignoring
+this will give a helpful error) and we use the full arsenal of 
+40 inference models, which are all combinations of 4 site 
 models, 2 clock models and 5 tree priors.
@@
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]
-model_select_param <- create_best_model_select_param()
+experiment <- create_all_experiments()
 \end{lstlisting}
 
-Also here, the third and fourth step have sensible defaults, and we are not
-interested in using a twin tree. We can now measure the errors BEAST2
-makes when inferring the given phylogeny like this:
+Also here, the third (the BEAST2 inference) and fourth (measuring the error)
+step have sensible defaults, and we are not
+interested in using a twin tree. We can create the complete
+\verb;pirouette; parameter set (again) like this:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments
+)
+\end{lstlisting}
+
+Running the experiment:
 
@@
 The following settings are the same as in the previous section:
-phylogeny, alignment parameters, inference model selection,
+phylogeny (listing \ref{lst:unknown_phylogeny}), alignment parameters, inference model selection,
 shared inference parameters and error measuring parameters 

### 2019-02-21 scripts/create_fig_2.R

@@ -1,45 +1,1 @@
-#' Create figure 2 for pirouette article
-#' @inheritParams default_params_doc
-#' @return a ggplot2 plot
-#' @author Richel J.C. Bilderbeek, Giovanni Laudanno
-#' @export
-create_fig_2 <- function() {
-  phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4) :1);")
-  alignment_params <- create_alignment_params(
-    root_sequence = create_blocked_dna(length = 1000),
-    mutation_rate = 0.1
-  )
-  model_select_param <- create_gen_model_select_param(
-    alignment_params = alignment_params,
-    tree_prior = create_bd_tree_prior()
-  )
-  errors <- pir_run(
-    phylogeny = phylogeny,
-    alignment_params = alignment_params,
-    model_select_params = model_select_param
-  )
-  pir_plot(errors) # nolint pirouette function
-}
-
-#' Create figure 2 file
-#' @param filename name of the file to save figure 2 to
-#' @return the name of the file the plot is saved to
-#' @author Richel J.C. Bilderbeek, Giovanni Laudanno
-#' @export
-create_fig_2_file <- function(
-  fig_2_filename = "figure_2.png"
-) {
-  # create figure
-  fig_2 <- create_fig_2() # nolint internal function
-
-  # save output
-  ggplot2::ggsave(
-    filename = fig_2_filename,
-    plot = fig_2
-  )
-  fig_2_filename
-}
-
-# Run
-library(pirouette)
-create_fig_2_file()
+# See example_1

### 2019-02-21 scripts/example_1.R

@@ -1,0 +1,49 @@
+# Code of example 1
+#
+# Works under Windows
+library(pirouette)
+
+#phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
+# STUB
+phylogeny <- create_stunning_tree(n_taxa = 5, t_0 = 10)
+
+alignment_params <- create_alignment_params(
+  root_sequence = create_blocked_dna(length = 1000),
+  mutation_rate = 0.1
+)
+
+experiments <- list(create_experiment())
+
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments
+)
+
+errors <- pir_run(
+  phylogeny,
+  pir_params = pir_params
+)
+
+
+library(ggplot2)
+
+pir_plot(errors) + theme(
+  panel.background = element_rect(
+    fill = "white",
+    colour = "black"
+  ),
+  panel.grid.major = element_line(colour = "black"),
+  panel.grid.minor = element_line(color = "black")
+
+)
+
+library(ggthemes)
+pir_plot(errors) + theme_wsj() + ggsave("/home/richel/GitHubs/pirouette_article/figure_2.png")
+
+pir_plot(errors) + theme_tufte()
+pir_plot(errors) + theme_economist()
+pir_plot(errors) + theme_economist_white()
+pir_plot(errors) + theme_excel() # The irony!
+pir_plot(errors) + theme_few() # Like
+pir_plot(errors) + theme_fivethirtyeight()
+

### 2019-02-21 scripts/example_2.R

@@ -1,0 +1,47 @@
+# Code of example 2
+#
+# Works under Linux and MacOS only
+library(pirouette)
+
+phylogeny  <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4) :1);")
+
+alignment_params <- create_alignment_params(
+  root_sequence = create_blocked_dna(length = 1000),
+  mutation_rate = 0.1
+)
+
+experiments <- create_all_experiments()
+
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments
+)
+
+errors <- pir_run(
+  phylogeny,
+  pir_params = pir_params
+)
+
+
+library(ggplot2)
+
+pir_plot(errors) + theme(
+  panel.background = element_rect(
+    fill = "white",
+    colour = "black"
+  ),
+  panel.grid.major = element_line(colour = "black"),
+  panel.grid.minor = element_line(color = "black")
+
+)
+
+library(ggthemes)
+pir_plot(errors) + theme_wsj() + ggsave("/home/richel/GitHubs/pirouette_article/figure_2.png")
+
+pir_plot(errors) + theme_tufte()
+pir_plot(errors) + theme_economist()
+pir_plot(errors) + theme_economist_white()
+pir_plot(errors) + theme_excel() # The irony!
+pir_plot(errors) + theme_few() # Like
+pir_plot(errors) + theme_fivethirtyeight()
+

### 2019-02-21 article.tex
Add experiment setup, progress https://github.com/richelbilderbeek/pirouette/issues/58

@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[
+  language=R, 
+  floatplacement=H, frame=single,
+  label = {lst:create_generative_experiment},
+  caption = {
+    Create a default experiment. This default experiment is
+    for a generative model, that is always run, without its evidence measured,
+    assuming a JC, strict, Yule inference model
+  }
+]
 experiments <- list(create_experiment())
@@
 A third research question that \verb;pirouette; answers is:
-What is the error BEAST2 makes from a phylogeny 
-picking the best inference model, compared to the background noise?
+What is the error BEAST2 makes from a phylogeny, 
+when hand-picking an inference model, compared to the background noise?
 
 The following settings are the same as in the previous section:
-phylogeny (listing \ref{lst:unknown_phylogeny}), alignment parameters, inference model selection,
-shared inference parameters and error measuring parameters 
-\richel{TODO: add ref to code}
+phylogeny (listing \ref{lst:unknown_phylogeny}), 
+alignment parameters (listing \ref{lst:create_alignment}), 
+experiments (listing \ref{lst:create_generative_experiment})
+BEAST2 inference and error measuring parameters.
 
@@
 has the same topology as the given tree, yet its branch lengths follow
-an idealized 
-\giovanni{
-  I don't think this is true. A twin tree is generated randomly. 
-  If we use the method "maxlikelihood" (see the function "createbdtree") 
-  the tree will be chosen in such a way it will maximizing the likelihood 
-  given the parameters we estimated from the original tree, 
-  but I would not call it "idealized". Furthermore I think it's not obvious to 
-  use the "maxlikelihood" method, as in general the "ideal" scenario should be 
-  established using a distribution of trees. A single tree is always the result 
-  of a stochastic process.
-}
-\richel{we'll discuss this today} 
-birth-death distribution. Creating a twinning parameter is easy,
+a distribution as such that the BEAST2 error will be 
+close to \richel{'close to' is a bit weasly, but currently it is true}
+its minimal error.
+Creating a twinning parameter is easy,
 as it has sensible default settings:
@@
 
+All the \verb;pirouette; arguments are bundled
+and checked by \verb;create_pir_params;:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments,
+  twinning_params = twinning_params
+)
+\end{lstlisting}
+
+Running:
+
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]
@@
   phylogeny,
-  twinning_params = twinning_params,
-  alignment_params = alignment_params,
-  model_select_params = model_select_param
+  pir_params = pir_params
 )
@@
     tree, to measure the noise, which is the minimal error. 
-    \giovanni{again, I don't believe that using an "idealized" tree is the best 
-      thing to do. I would rather use more "true trees".}
-    \richel{we'll discuss this today}
   }
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Other examples}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+There are many other examples that can be found in the \verb;pirouette; 
+documentation.
 

### 2019-02-21 scripts/create_fig_3.R

@@ -1,42 +1,1 @@
-#' Create figure 3 for pirouette article
-#' @inheritParams default_params_doc
-#' @return a ggplot2 plot
-#' @author Richel J.C. Bilderbeek, Giovanni Laudanno
-#' @export
-create_fig_3 <- function() {
-  phylogeny  <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4) :1);")
-  alignment_params <- create_alignment_params(
-    root_sequence = create_blocked_dna(length = 1000),
-    mutation_rate = 0.1
-  )
-  model_select_param <- create_best_model_select_param() # nolint pirouette function
-  errors <- pir_run( # nolint pirouette function
-    phylogeny = phylogeny,
-    alignment_params = alignment_params,
-    model_select_params = model_select_param
-  )
-  pir_plot(errors) # nolint pirouette function
-}
-
-#' Create figure 3 file
-#' @inheritParams default_params_doc
-#' @return the name of the file the plot is saved to
-#' @author Richel J.C. Bilderbeek, Giovanni Laudanno
-#' @export
-create_fig_3_file <- function(
-  fig_3_filename = "figure_3.png"
-) {
-  # create figure
-  fig_3 <- create_fig_3() # nolint internal function
-
-  # save output
-  ggplot2::ggsave(
-    filename = fig_3_filename,
-    plot = fig_3
-  )
-  fig_3_filename
-}
-
-# Run
-library(pirouette)
-create_fig_3_file()
+# See example_2.R

### 2019-02-21 scripts/example_3.R

@@ -1,20 +1,25 @@
-# Code of example 2
+# Code of example 3
 #
-# Works under Linux and MacOS only
+# Works under Windows
 library(pirouette)
 
-phylogeny  <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4) :1);")
+#phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
+# STUB
+phylogeny <- create_stunning_tree(n_taxa = 5, t_0 = 10)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),
   mutation_rate = 0.1
 )
 
-experiments <- create_all_experiments()
+experiments <- list(create_experiment())
+
+twinning_params <- create_twinning_params()
 
 pir_params <- create_pir_params(
   alignment_params = alignment_params,
-  experiments = experiments
+  experiments = experiments,
+  twinning_params = twinning_params
 )
 
 errors <- pir_run(
@@ -24,19 +29,8 @@ errors <- pir_run(
 
 
 library(ggplot2)
-
-pir_plot(errors) + theme(
-  panel.background = element_rect(
-    fill = "white",
-    colour = "black"
-  ),
-  panel.grid.major = element_line(colour = "black"),
-  panel.grid.minor = element_line(color = "black")
-
-)
-
 library(ggthemes)
-pir_plot(errors) + theme_wsj() + ggsave("/home/richel/GitHubs/pirouette_article/figure_3.png")
+pir_plot(errors) + theme_wsj() + ggsave("/home/richel/GitHubs/pirouette_article/figure_4.png")
 
 pir_plot(errors) + theme_tufte()
 pir_plot(errors) + theme_economist()

### 2019-02-25 article.tex
Updates from Overleaf

@@
 such as BEAST [\cite{drummond2007beast}] or MrBayes [\cite{huelsenbeck2001mrbayes}], 
-provides us the possibility to build phylogenetic trees 
+allowed to build phylogenetic trees 
 from genetic material extracted from extant organisms.
@@
 evolve.
-Such tools have been increasingly exploited to hypothesize and test what are the main drivers
-and modes for diversification processes.
+Such tools have been increasingly exploited to hypothesize and test what are the main drivers and modes for diversification processes.
 
@@
 concealed state [\cite{beaulieu2016detecting}] 
-or even a combination of multiple concealed and observable states [\cite{herrera2018detecting}].
+or even a combination of multiple concealed and observable states [\cite{herrera2018detecting}]. \giovanni{we might need to polish this list a little bit}
 
@@
 Such a mechanism is built into a likelihood function, 
-which provides the likelihood of the parameters, given the data.
+which provides the probability to realize a process given the parameters.
 By maximizing it, it is possible 
-to estimate the what are the best values for model parameters. 
+to estimate what are the best values for model parameters. 
 It is a standard practice in assessing such models,
@@
 
-Diversification models are getting more complex over time. There are many
-reasons for this. One reason is the increase in computing power,
-which allows this heightened complexity possible.
-Another reason is to explore a process that are not yet considered by currently existing diversification models. An example of this
-is the multiple-birth model [Laudanno et al., in preparation], 
+Diversification models are getting more complex over time. This is facilitated by the increase in computing power due to the advancement of technology.
+Increased complexity made possible the exploration of new models encompassing mechanisms that are not yet considered by previous diversification models.
+An example of this is the multiple-birth model [Laudanno et al., in preparation],
 that is the first model to allow simultaneous speciations to co-occur.
@@
 However, whenever a new model is proposed there is always a trade-off between the cost of increased complexity versus the benefits that it can provide.
-Therefore desirable models have to be able to disentangle the intrinsic complexity of a system, 
-highlight the main reasons of its behaviour and avoid unnecessary complexity. Furthermore developing models can cost time, human and computational resources.
+In fact developing models can cost time, human and computational resources.
+Therefore desirable models have to be able to disentangle the intrinsic complexity of a system, highlight the main reasons of its behaviour and avoid unnecessary complexity.
 
@@
 
-For this reason we believe it is important to have techniques to quantitatively assess the advancement provided by new models, in comparison to the existing ones.
+For this reason we believe it is important to have techniques to quantitatively assess the advancement provided by the introduction of new models, in comparison to the existing ones.
 
@@
 
-To achieve this, one possible experiment
-is to simulate a phylogeny according to the mechanisms proposed by the new diversification model, then simulate its alignment and measure how well current inference models perform at retrieving the original phylogeny.
+To achieve this, one possible experiment is to simulate a phylogeny according to the mechanisms proposed by the new diversification model, then simulate its alignment and measure how well current inference models perform at retrieving the original phylogeny.
 If this difference is small enough, current models apparently
@@
 any given phylogeny, helping us to validate new diversification models.
+\giovanni{I think we should mention the fact that BEAST2 allows to implement customized tree priors more complicated than the simple yule and bd.
+These tree priors, in the end, are what we call likelihood in our models, so it's pivotal in proposing our argument, I think.}
 
@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given reconstructed phylogeny. For this measurement,
-two models need to be specified: a generative and an inference model.
+makes from a given reconstructed phylogeny \giovanni{We should mention here that the phylogenies that are meant to be submitted to pirouette are simulated phylogenies generated according to the model we want to test}.
+For this measurement, two models need to be specified: a generative and an inference model.
 A generative model is a combination of site model, clock model and tree prior,
@@
 time-calibrated phylogeny, by specifying a normal distribution around
-an expected time when two or more taxa's ancestors diversified.
+an expected time when two or more taxa's \giovanni{I have the feeling that taxa should be plural here. Should we mention that the group must be monophyletic?} ancestors diversified.
 
+\subsection{pirouette's pipeline}
 The pipeline is summarized by the following steps, which will be described in detail below:
@@
     \item the alignment is used as BEAST2 input to infer a posterior distribution of phylogenies;
-    \item posterior phylogenies are compared with the given original phylogeny to estimate the error we make. The bigger the match, the lower the error;
+    \item posterior phylogenies are compared with the given original phylogeny to estimate the error we make. The higher the match, the lower the error;
 \end{enumerate}
@@
 The second step selects an inference model, using the alignment.
-The user can specify whether to use the generative model or select a best model
+The user can specify whether to use the generative model or select the best model
 from a set of inference models. 
@@
 approach, as described in \cite{maturana2018model}. The nested sampling is
-performed by \verb;mcbette; \cite{mcbette}, that calls the 'NS' BEAST2 package. 
+performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 package. 
 Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; 
@@
 but also a custom error statistic is supplied, based
-on the absolute difference in the gamma statistic [Pybus \& Harvy]. 
+on the absolute difference in the gamma statistic [\cite{pybus2000testing}]. 
 Additionally, the user can specify the
@@
 \begin{lstlisting}[
-    language=R, floatplacement=H, 
-    frame=single
+    language=R,
+    floatplacement=H, frame=single,
     label = {lst:create_alignment}, 
@@
 By default, an alignment is created using a Jukes-Cantor (JC) site model
-and a strict clock model. A JC site model assumes that mutations
-between all nucleotides are constant at an equal rate. 
-A strict clock model assumes that
-the mutation rate of all species is equal and constant.
+and a strict clock model.
+A JC site model assumes that mutation rates from any nucleotide to any other are equal and constant.
+A strict clock model assumes that the mutation rates of all species are equal and constant.
 Due to this, we state that the generative model for the alignment is
@@
     strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior
+    Yule: Yule (pure-birth) tree prior.
   }
@@
     for a generative model, that is always run, without its evidence measured,
-    assuming a JC, strict, Yule inference model
+    assuming a JC site model, strict clock model, Yule inference model.
   }
@@
     The error BEAST2 makes from a phylogeny 
-    when the generative and inference model are the same
+    when the generative and inference model are the same.
   }
@@
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]
-experiment <- create_all_experiments()
+experiments <- create_all_experiments()
 \end{lstlisting}
@@
 Also here, the third (the BEAST2 inference) and fourth (measuring the error)
-step have sensible defaults, and we are not
+steps have sensible defaults, and we are not
 interested in using a twin tree. We can create the complete
@@
 alignment parameters (listing \ref{lst:create_alignment}), 
-experiments (listing \ref{lst:create_generative_experiment})
+experiments (listing \ref{lst:create_generative_experiment}),
 BEAST2 inference and error measuring parameters.
@@
     The error BEAST2 makes from a phylogeny 
-    picking the best inference model, compared to the background noise?
+    picking the best inference model, compared to the background noise.
     Here, the twin column shows the error BEAST2 makes on an idealized
@@
 documentation.
+\giovanni{it could be nice to add a listing with simple commands that refer to the most useful examples}
 
@@
 \giovanni{We need to discuss the results we got from the 3 experiments.}
-\richel{Only iff the journal requires this}
+\richel{Only if the journal requires this}
 

### 2019-02-25 article.tex
EOD

@@
 concealed state [\cite{beaulieu2016detecting}] 
-or even a combination of multiple concealed and observable states [\cite{herrera2018detecting}]. \giovanni{we might need to polish this list a little bit}
+or even a combination of multiple concealed and 
+observable states [\cite{herrera2018detecting}]. 
+\giovanni{we might need to polish this list a little bit}
 
@@
 which calls the popular Bayesian inference tool 
-BEAST2 [\cite{bouckaert2014beast}]. With \verb;pirouette;, one
+BEAST2 [\cite{bouckaert2014beast}]. 
+With \verb;pirouette;, one
 can easily measure the error made by Bayesian inference in recovering
 any given phylogeny, helping us to validate new diversification models.
-\giovanni{I think we should mention the fact that BEAST2 allows to implement customized tree priors more complicated than the simple yule and bd.
-These tree priors, in the end, are what we call likelihood in our models, so it's pivotal in proposing our argument, I think.}
+Such a new diversification model can then be added to BEAST2,
+which already supplies multiple speciation models (among others, 
+Yule and birth-death) and also allows to add 
+new diversification models via the BEAST2 package system.
+\richel{rework}
 
@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given reconstructed phylogeny \giovanni{We should mention here that the phylogenies that are meant to be submitted to pirouette are simulated phylogenies generated according to the model we want to test}.
-For this measurement, two models need to be specified: a generative and an inference model.
-A generative model is a combination of site model, clock model and tree prior,
+makes from a given reconstructed phylogeny 
+\giovanni{
+  We should mention here that the phylogenies that are meant to be 
+  submitted to pirouette are simulated phylogenies generated according 
+  to the model we want to test
+}. 
+To be able to explain how we do this, we first give definitions
+for generative model and inference model: 
+a generative model is a combination of site model, clock model and tree prior,
 used to simulate alignments. An inference model is a combination of 
@@
 time-calibrated phylogeny, by specifying a normal distribution around
-an expected time when two or more taxa's \giovanni{I have the feeling that taxa should be plural here. Should we mention that the group must be monophyletic?} ancestors diversified.
+an expected time when two or more taxa's 
+\giovanni{
+  I have the feeling that taxa should be plural here. 
+  Should we mention that the group must be monophyletic?
+} ancestors diversified.
 
@@
 documentation.
-\giovanni{it could be nice to add a listing with simple commands that refer to the most useful examples}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[h]
+\centering
+\begin{tabular}{ | l | l | }
+\hline
+\textbf{Name} & \textbf{Description} \\
+\hline
+\verb;pir_run; & Run \verb;pirouette; \\
+\verb;create_pir_params; & Create the \verb;pirouette; parameters \\
+\hline
+\verb;create_alignment_params; & Create the alignment parameters \\
+\verb;create_twinning_params; & Create the twinning parameters \\
+\verb;create_experiment; & Create one experiment \\
+\verb;create_error_measure_params; & Create the error measurement parameters \\
+\hline
+\end{tabular}
+\caption{pirouette's main functions}
+\label{tab:functions}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-02-26 article.tex
Update scripts

@@
 There are plenty of diversification models: 
-constant-birth [Yule, 19..],
-birth-death [\cite{nee1994reconstructed}], 
-time-dependent [Rabosky D., LASER package?], 
-diversity dependent [\cite{etienne2011diversity}], 
-protracted birth-death [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
-multiple-birth death [Laudanno, Bilderbeek, Haegeman \& Etienne, in preparation],
-a rate shifting birth-death model [the SLS model. Laudanno, Haegeman \& Etienne, in preparation]
-punctuated equilibrium (as known as "early burst") [\cite{harmon2010early}],
-dependent on a binary trait [\cite{maddison2007estimating}], 
-multiple state trait [\cite{fitzjohn2012diversitree}],
-concealed state [\cite{beaulieu2016detecting}] 
-or even a combination of multiple concealed and 
-observable states [\cite{herrera2018detecting}]. 
-\giovanni{we might need to polish this list a little bit}
+birth-death models that have rates that are 
+constant [\cite{nee1994reconstructed}], 
+time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
+diversity-dependent [\cite{etienne2011diversity}],
+or can shift [the SLS model. Laudanno, Haegeman \& Etienne, in preparation].
+There are birth death models in which speciation is a process that 
+takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
+can co-occur [Laudanno, Bilderbeek, Haegeman \& Etienne, in preparation],
+dependent on a trait that is binary [\cite{maddison2007estimating}], 
+has multiple states [\cite{fitzjohn2012diversitree}],
+has a concealed state [\cite{beaulieu2016detecting}] 
+or has multiple observed and concealed states [\cite{herrera2018detecting}].
 
@@
 
-Diversification models are getting more complex over time. This is facilitated by the increase in computing power due to the advancement of technology.
-Increased complexity made possible the exploration of new models encompassing mechanisms that are not yet considered by previous diversification models.
+Diversification models are getting more complex over time. 
+This is facilitated by the increase in computing power due to the advancement of technology.
+Increased complexity made possible the exploration of new models 
+encompassing mechanisms that are not yet considered by previous diversification models. 
 An example of this is the multiple-birth model [Laudanno et al., in preparation],
@@
 
-However, whenever a new model is proposed there is always a trade-off between the cost of increased complexity versus the benefits that it can provide.
+However, whenever a new model is proposed there is always a trade-off 
+between the cost of increased complexity versus the benefits that it can provide.
 In fact developing models can cost time, human and computational resources.
-Therefore desirable models have to be able to disentangle the intrinsic complexity of a system, highlight the main reasons of its behaviour and avoid unnecessary complexity.
+Therefore desirable models have to be able to disentangle 
+the intrinsic complexity of a system, 
+highlight the main reasons of its behaviour and avoid unnecessary complexity.
 
-\iffalse
-One of the purposes of modelling a complex systems
-is to hypothesize the main driving factors of a process 
-and trying to prove that they are indeed able 
-to explain the main features of the phenomenon you want to explain.
-Due to this, there are diminishing returns on adding complexity
-to a model: the job of a modeller is to try to
-disentangle the intrinsic complexity of a system 
-highlighting the main reasons of its behaviour. 
-In such a light you don't want your model to be unnecessary complex.
-Additionally, developing a likelihood model takes time, 
-human resources, fundings and can strain computational resources.
-\fi
+For this reason we believe it is important to have techniques 
+to quantitatively assess the advancement provided by the introduction of new models, 
+in comparison to the existing ones.
 
-For this reason we believe it is important to have techniques to quantitatively assess the advancement provided by the introduction of new models, in comparison to the existing ones.
-
-\iffalse
-When assuming that a new diversification model is worth the
-(philosophical) effort, there are many ways to measure
-if the new model indeed gives a better explanation. 
-\fi
-
-To achieve this, one possible experiment is to simulate a phylogeny according to the mechanisms proposed by the new diversification model, then simulate its alignment and measure how well current inference models perform at retrieving the original phylogeny.
+To achieve this, one possible experiment is to simulate a phylogeny 
+according to the mechanisms proposed by the new diversification model, 
+then simulate its alignment and measure how well current inference models perform 
+at retrieving the original phylogeny.
 If this difference is small enough, current models apparently
@@
 can easily measure the error made by Bayesian inference in recovering
-any given phylogeny, helping us to validate new diversification models.
-Such a new diversification model can then be added to BEAST2,
-which already supplies multiple speciation models (among others, 
-Yule and birth-death) and also allows to add 
-new diversification models via the BEAST2 package system.
-\richel{rework}
+any given phylogeny, helping us to judge the necessity of a diversification 
+model.
+Upon finding a diversification model is warranted, 
+it can then be added to BEAST2's existing set of (birth-death and 
+coalescent) models via the BEAST2 package system.
 
@@
   \centering
-  \includegraphics[width=\textwidth]{overview.png}
+  \includegraphics[width=\textwidth]{workflow.png}
   \caption{pirouette pipeline}
@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given reconstructed phylogeny 
-\giovanni{
-  We should mention here that the phylogenies that are meant to be 
-  submitted to pirouette are simulated phylogenies generated according 
-  to the model we want to test
-}. 
-To be able to explain how we do this, we first give definitions
-for generative model and inference model: 
+makes from a given reconstructed phylogeny, within the context of
+a simulation study on a (possibly new) diversification model. Such
+a study will typically construct phylogenies for different combinations of
+the diversification model's parameters, to assess under which scenario's
+the error BEAST2 makes cannot be ignored anymore. Within such a study,
+there need be many replicates.
+
+To be able to explain how \verb;pirouette; measures the inference error
+BEAST2 makes, we first give two definitions of what
+we mean by a generative model and by an inference model: 
 a generative model is a combination of site model, clock model and tree prior,
@@
 time-calibrated phylogeny, by specifying a normal distribution around
-an expected time when two or more taxa's 
-\giovanni{
-  I have the feeling that taxa should be plural here. 
-  Should we mention that the group must be monophyletic?
-} ancestors diversified.
+an expected time when the ancestors of two or more taxa diversified.
 
 \subsection{pirouette's pipeline}
-The pipeline is summarized by the following steps, which will be described in detail below:
+The pipeline is summarized by the following steps,
+which will be described in detail below:
 \begin{enumerate}
@@
 \end{enumerate}
-\richel{figure to pipeline here}
+The pipeline is visualized in \ref{fig:pipeline}.
 There is also the option to generate a 'twin tree',
@@
 
-\giovanni{We need to discuss the results we got from the 3 experiments.}
-\richel{Only if the journal requires this}
+In figure ...
 

### 2019-02-26 scripts/create_fig_4.R

@@ -1,44 +1,1 @@
-#' Create figure 4 for pirouette article
-#' @inheritParams default_params_doc
-#' @return a ggplot2 plot
-#' @author Richel J.C. Bilderbeek, Giovanni Laudanno
-#' @export
-create_fig_4 <- function() {
-  phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4) :1);")
-  alignment_params <- create_alignment_params(
-    root_sequence = create_blocked_dna(length = 1000),
-    mutation_rate = 0.1
-  )
-  model_select_param <- create_best_model_select_param() # nolint pirouette function
-  twinning_params <- create_twinning_params() # nolint pirouette function
-  errors  <- pir_run( # nolint pirouette function
-    phylogeny = phylogeny,
-    alignment_params = alignment_params,
-    model_select_params = model_select_param,
-    twinning_params = twinning_params
-  )
-  pir_plot(errors) # nolint pirouette function
-}
-
-#' Create figure 4 file
-#' @inheritParams default_params_doc
-#' @return the name of the file the plot is saved to
-#' @author Richel J.C. Bilderbeek, Giovanni Laudanno
-#' @export
-create_fig_4_file <- function(
-  fig_4_filename = "figure_4.png"
-) {
-  # create figure
-  fig_4 <- create_fig_4() # nolint internal function
-
-  # save output
-  ggplot2::ggsave(
-    filename = fig_4_filename,
-    plot = fig_4
-  )
-  fig_4_filename
-}
-
-# Run
-library(pirouette)
-create_fig_4_file()
+# Moved to example_3.R

### 2019-02-26 scripts/create_fig_bd.R

@@ -1,101 +1,1 @@
-#' Estimates best lambda for a given n_taxa and n_0 and then
-#'   calculates likelihood of a tree, given its branching times
-#' @inheritParams default_params_doc
-#' @author Giovanni Laudanno
-#' @return the likelihood
-#' @noRd
-calc_likelihood <- function(
-    brts,
-    t_0 = 10,
-    n_taxa = 5,
-    n_0 = 2
-  ) {
-  cond_1 <- (length(brts) != length(unique(brts)))
-  cond_2 <- (length(brts) != (n_taxa - n_0))
-  cond_3 <- (any(brts < 0))
-  cond_4 <- (any(brts > t_0))
-  if (cond_1 | cond_2 | cond_3 | cond_4) {
-    loglik <- -Inf
-  } else {
-    lambda <- (log(n_taxa) - log(n_0) ) / t_0
-    mu <- 0
-    r <- -(lambda - mu)
-    lik <- (1 - exp(2 * r * (t_0 - brts[1]))) *
-      (1 - exp(3 * r * (brts[1] - brts[2]))) *
-      (1 - exp(4 * r * (brts[2] - brts[3]))) *
-      (1 - exp(5 * r * brts[3]))
-    loglik <- log(lik)
-  }
-  out <- -loglik
-  return(out)
-}
-
-#' Create figure bd for pirouette article
-#' @inheritParams default_params_doc
-#' @return a list containing a tree in phylo format, the same tree in
-#' newick format and the plot of the tree
-#' @author Giovanni Laudanno
-#' @export
-create_fig_bd <- function(
-  t_0 = 10,
-  n_taxa = 5,
-  n_0 = 2
-) {
-  start_brts <- seq(from = 0, to = t_0, by = t_0 / (n_taxa - n_0 + 1))
-  start_brts <- sort(start_brts[2:(length(start_brts) - 1)], decreasing = TRUE)
-  x <- subplex::subplex(
-    par = start_brts,
-    fn = calc_likelihood,
-    t_0 = t_0,
-    n_taxa = n_taxa,
-    n_0 = n_0
-  )
-  brts <- sort(c(rep(t_0, n_0), x$par), decreasing = TRUE)
-  brts <- signif(brts, digits = 2)
-  l_table <- matrix(NA, nrow = 5, ncol = 4)
-  l_table[, 4] <- -1
-  l_table[, 3] <- c(1, -2, 3, -4, 5)
-  l_table[, 2] <- c(0, 1, 1, -2, 3)
[…]

### 2019-02-26 scripts/create_fig_2.R
Cleaned up scripts

@@ -1,1 +1,0 @@
-# See example_1

### 2019-02-26 scripts/create_fig_3.R

@@ -1,1 +1,0 @@
-# See example_2.R

### 2019-02-26 scripts/create_fig_4.R

@@ -1,1 +1,0 @@
-# Moved to example_3.R

### 2019-02-26 scripts/create_fig_bd.R

@@ -1,1 +1,0 @@
-# Moved to create_stunning_tree

### 2019-02-26 article.tex
Simplify scripts

@@
   \textbf{4. }
-    As \verb;pirouette; is designed to be of high quality and extendable, 
-    we conclude by describing the further development of the package. \\
+    Last, we'll discuss the results obtained by the examples. \\
 \end{abstract}
@@
 The development of new powerful inference tools, 
-such as BEAST [\cite{drummond2007beast}] or MrBayes [\cite{huelsenbeck2001mrbayes}], 
+such as BEAST [\cite{drummond2007beast}], MrBayes [\cite{huelsenbeck2001mrbayes}]
+or RevBayes [\cite{hohna2016revbayes}], 
 allowed to build phylogenetic trees 

### 2019-02-26 scripts/example_1.R

@@ -2,6 +2,8 @@
 #
 # Works under Windows
 library(pirouette)
+library(ggplot2)
+library(ggthemes)
 
 phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
 
@@ -22,26 +24,4 @@ errors <- pir_run(
   pir_params = pir_params
 )
 
-
-library(ggplot2)
-
-pir_plot(errors) + theme(
-  panel.background = element_rect(
-    fill = "white",
-    colour = "black"
-  ),
-  panel.grid.major = element_line(colour = "black"),
-  panel.grid.minor = element_line(color = "black")
-
-)
-
-library(ggthemes)
 pir_plot(errors) + theme_wsj() + ggsave("figure_example_1.png")
-
-pir_plot(errors) + theme_tufte()
-pir_plot(errors) + theme_economist()
-pir_plot(errors) + theme_economist_white()
-pir_plot(errors) + theme_excel() # The irony!
-pir_plot(errors) + theme_few() # Like
-pir_plot(errors) + theme_fivethirtyeight()
-

### 2019-02-26 scripts/example_2.R

@@ -2,6 +2,8 @@
 #
 # Works under Linux and MacOS only
 library(pirouette)
+library(ggplot2)
+library(ggthemes)
 
 phylogeny  <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4) :1);")
 
@@ -22,9 +24,6 @@ errors <- pir_run(
   pir_params = pir_params
 )
 
-
-library(ggplot2)
-
 pir_plot(errors) + theme(
   panel.background = element_rect(
     fill = "white",
@@ -35,13 +34,4 @@ pir_plot(errors) + theme(
 
 )
 
-library(ggthemes)
 pir_plot(errors) + theme_wsj() + ggsave("figure_example_2.png")
-
-pir_plot(errors) + theme_tufte()
-pir_plot(errors) + theme_economist()
-pir_plot(errors) + theme_economist_white()
-pir_plot(errors) + theme_excel() # The irony!
-pir_plot(errors) + theme_few() # Like
-pir_plot(errors) + theme_fivethirtyeight()
-

### 2019-02-26 scripts/example_3.R

@@ -2,6 +2,8 @@
 #
 # Works under Windows
 library(pirouette)
+library(ggplot2)
+library(ggthemes)
 
 #phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
 # STUB
@@ -28,14 +30,4 @@ errors <- pir_run(
 )
 
 
-library(ggplot2)
-library(ggthemes)
 pir_plot(errors) + theme_wsj() + ggsave("figure_example_3")
-
-pir_plot(errors) + theme_tufte()
-pir_plot(errors) + theme_economist()
-pir_plot(errors) + theme_economist_white()
-pir_plot(errors) + theme_excel() # The irony!
-pir_plot(errors) + theme_few() # Like
-pir_plot(errors) + theme_fivethirtyeight()
-

### 2019-02-26 article.tex
Reviewed article, progress #58, fix #151

@@
 or RevBayes [\cite{hohna2016revbayes}], 
-allowed to build phylogenetic trees 
+allows to build phylogenetic trees 
 from genetic material extracted from extant organisms.
 This has constituted an important step forward 
-in our understanding on how species 
-evolve.
-Such tools have been increasingly exploited to hypothesize and test what are the main drivers and modes for diversification processes.
+in our understanding on how species evolve.
+Such tools have been increasingly exploited to hypothesize 
+and test what the main drivers and modes for diversification processes are.
 
@@
 Such a mechanism is built into a likelihood function, 
-which provides the probability to realize a process given the parameters.
-By maximizing it, it is possible 
-to estimate what are the best values for model parameters. 
-It is a standard practice in assessing such models,
-to verify the level of accuracy they can achieve
-on simulated phylogenies, of which full knowledge is given.
+which provides the probability that a phylogeny is realized 
+from a given set of parameters.
+Starting from a phylogeny, this likelihood is maximized
+to find the parameters that are most likely.
+For such diversification models, it is a standard practice 
+to measure how well parameters can be recovered in this way,
+by simulating 'true' phylogenies from known parameters, then
+measuring the parameters deduced by maximum likelihood. 
+It is a sign of quality of a diversification model 
+if its parameters can be recovered well by this method.
+\richel{
+  We should bridge this method with how pirouette does things differently;
+  why the maximum likelihood does not apply in this context  
+}
 
@@
 \begin{figure}[h]
-  \includegraphics[width=\textwidth]{figure_2.png}
+  \includegraphics[width=\textwidth]{figure_example_1.png}
   \caption{
@@
 
-The resulting figure is shown in figure 3.
+The resulting figure is shown in figure \ref{fig:example_2}.
 
 \begin{figure}[h]
-  \includegraphics[width=\textwidth]{figure_3.png}
+  \includegraphics[width=\textwidth]{figure_example_2.png}
   \caption{
@@
   }
+  \label{fig:example_2}
 \end{figure}
@@
 
-The resulting figure is shown in figure 4
+The resulting figure is shown in figure \ref{fig:example_3}
 
 \begin{figure}[h]
-  \includegraphics[width=\textwidth]{figure_4.png}
+  \includegraphics[width=\textwidth]{figure_example_3.png}
   \caption{
@@
   }
+  \label{fig:example_3}
 \end{figure}
@@
 
-In figure ...
+Here we discuss how to interpret the results of \verb;pirouette;.
+Figure \ref{fig:example_1} shows that the generative model 
+gives an error distribution with a highest density at 0.02,
+where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0.
+The range of errors goes from 0.0 to 0.27. As we know the generative model,
+this error distribution is the minimum noise that BEAST2 produces for this
+generative model.
+\richel{Show effective sample size}
+
+Figure \ref{fig:example_2} shows that the best candidate model 
+gives an error distribution with a highest density at 0.02,
+where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0.
+The range of errors goes from 0.0 to 0.08. 
+\richel{Describe and show evidences for the candidate model}
+\richel{Describe and show the best candidate model}
+This means that it is rational to use this inference model to 
+infer a posterior phylogeny distribution, 
+as is done by \verb;pirouette;.
+\richel{Show effective sample size}
+\richel{Show posterior phylogenies as densitree}
+
+Figure \ref{fig:example_3} shows that the hand-picked model 
+gives an error distribution with a highest density at 0.018,
+with a range from 0.0 to 0.1.
+The twin tree has an error distribution with a highest density at 0.019,
+with a range from 0.0 to 0.11, which is the noise caused by the tree prior.
+As there is little difference between the true and twin tree,
+the (unknown) diversification model has little impact on the error
+BEAST2 makes, hinting that the constant-rate birth-death model is
+sufficiently good enough to explain the true phylogeny. 
+The diversification model apparently has minor impact and need not to be
+added to BEAST2.
 
@@
 and is licensed under the GNU General Public License v3.0.
-\verb;pirouette; uses the Travis CI (\url{https://travis-ci.org})
-continuous integration service, which is known to significantly 
-increase the number of bugs exposed (\cite{vasilescu2015}) and increases
-the speed at which new features are added (\cite{vasilescu2015}).
-\verb;pirouette; has a 100\% code coverage, which correlates with 
-code quality (\cite{horgan1994,del1995correlation}). 
-\verb;pirouette; follows Hadley Wickham's style guide (\cite{style_guide}), 
-which improves software quality (\cite{fang2001}).
+\verb;pirouette; follows all practices as recommended by the
+literature: continuous integration, a code coverage of 100\%
+and a style guide (\cite{style_guide}).
 \verb;pirouette; depends on multiple packages, which are 
@@
 \url{https://github.com/richelbilderbeek/pirouette}, 
-which accommodates collaboration (\cite{perez2016ten}) 
-and improves transparency (\cite{gorgolewski2016practical}).
-\verb;pirouette;'s GitHub facilitates feature requests and 
+which facilitates feature requests and 
 has guidelines on how to do so.

### 2019-02-26 scripts/example_1.R

@@ -24,4 +24,7 @@ errors <- pir_run(
   pir_params = pir_params
 )
 
-pir_plot(errors) + theme_wsj() + ggsave("figure_example_1.png")
+pir_plot(errors) +
+  scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
+  theme_wsj() +
+  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_1.png")

### 2019-02-26 scripts/example_2.R

@@ -24,14 +24,7 @@ errors <- pir_run(
   pir_params = pir_params
 )
 
-pir_plot(errors) + theme(
-  panel.background = element_rect(
-    fill = "white",
-    colour = "black"
-  ),
-  panel.grid.major = element_line(colour = "black"),
-  panel.grid.minor = element_line(color = "black")
-
-)
-
-pir_plot(errors) + theme_wsj() + ggsave("figure_example_2.png")
+pir_plot(errors) +
+  scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
+  theme_wsj() +
+  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_2.png")

### 2019-02-26 scripts/example_3.R

@@ -29,5 +29,7 @@ errors <- pir_run(
   pir_params = pir_params
 )
 
-
-pir_plot(errors) + theme_wsj() + ggsave("figure_example_3")
+pir_plot(errors) +
+  scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
+  theme_wsj() +
+  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_3.png")

### 2019-02-26 article.tex
Updates from Overleaf

--- 2019-02-26 article.tex
@@
 There are plenty of diversification models: 
-birth-death models that have rates that are 
-constant [\cite{nee1994reconstructed}], 
-time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
-diversity-dependent [\cite{etienne2011diversity}],
-or can shift [the SLS model. Laudanno, Haegeman \& Etienne, in preparation].
-There are birth death models in which speciation is a process that 
-takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
-can co-occur [Laudanno, Bilderbeek, Haegeman \& Etienne, in preparation],
-dependent on a trait that is binary [\cite{maddison2007estimating}], 
-has multiple states [\cite{fitzjohn2012diversitree}],
-has a concealed state [\cite{beaulieu2016detecting}] 
-or has multiple observed and concealed states [\cite{herrera2018detecting}].
+constant-birth [Yule, 19..],
+birth-death [\cite{nee1994reconstructed}], 
+time-dependent [Rabosky D., LASER package?], 
+diversity dependent [\cite{etienne2011diversity}], 
+protracted birth-death [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
+multiple-birth death [Laudanno, Bilderbeek, Haegeman \& Etienne, in preparation],
+a rate shifting birth-death model [the SLS model. Laudanno, Haegeman \& Etienne, in preparation]
+punctuated equilibrium (as known as "early burst") [\cite{harmon2010early}],
+dependent on a binary trait [\cite{maddison2007estimating}], 
+multiple state trait [\cite{fitzjohn2012diversitree}],
+concealed state [\cite{beaulieu2016detecting}] 
+or even a combination of multiple concealed and observable states [\cite{herrera2018detecting}]. \giovanni{we might need to polish this list a little bit}
 
@@
 
-Diversification models are getting more complex over time. 
-This is facilitated by the increase in computing power due to the advancement of technology.
-Increased complexity made possible the exploration of new models 
-encompassing mechanisms that are not yet considered by previous diversification models. 
+Diversification models are getting more complex over time. This is facilitated by the increase in computing power due to the advancement of technology.
+Increased complexity made possible the exploration of new models encompassing mechanisms that are not yet considered by previous diversification models.
 An example of this is the multiple-birth model [Laudanno et al., in preparation],
@@
 
-However, whenever a new model is proposed there is always a trade-off 
-between the cost of increased complexity versus the benefits that it can provide.
+However, whenever a new model is proposed there is always a trade-off between the cost of increased complexity versus the benefits that it can provide.
 In fact developing models can cost time, human and computational resources.
-Therefore desirable models have to be able to disentangle 
-the intrinsic complexity of a system, 
-highlight the main reasons of its behaviour and avoid unnecessary complexity.
+Therefore desirable models have to be able to disentangle the intrinsic complexity of a system, highlight the main reasons of its behaviour and avoid unnecessary complexity.
 
-For this reason we believe it is important to have techniques 
-to quantitatively assess the advancement provided by the introduction of new models, 
-in comparison to the existing ones.
+\iffalse
+One of the purposes of modelling a complex systems
+is to hypothesize the main driving factors of a process 
+and trying to prove that they are indeed able 
+to explain the main features of the phenomenon you want to explain.
+Due to this, there are diminishing returns on adding complexity
+to a model: the job of a modeller is to try to
+disentangle the intrinsic complexity of a system 
+highlighting the main reasons of its behaviour. 
+In such a light you don't want your model to be unnecessary complex.
+Additionally, developing a likelihood model takes time, 
+human resources, fundings and can strain computational resources.
+\fi
 
-To achieve this, one possible experiment is to simulate a phylogeny 
-according to the mechanisms proposed by the new diversification model, 
-then simulate its alignment and measure how well current inference models perform 
-at retrieving the original phylogeny.
+For this reason we believe it is important to have techniques to quantitatively assess the advancement provided by the introduction of new models, in comparison to the existing ones.
+
+\iffalse
+When assuming that a new diversification model is worth the
+(philosophical) effort, there are many ways to measure
+if the new model indeed gives a better explanation. 
+\fi
+
+To achieve this, one possible experiment is to simulate a phylogeny according to the mechanisms proposed by the new diversification model, then simulate its alignment and measure how well current inference models perform at retrieving the original phylogeny.
 If this difference is small enough, current models apparently
@@
 which calls the popular Bayesian inference tool 
-BEAST2 [\cite{bouckaert2014beast}]. 
-With \verb;pirouette;, one
+BEAST2 [\cite{bouckaert2014beast}]. With \verb;pirouette;, one
 can easily measure the error made by Bayesian inference in recovering
-any given phylogeny, helping us to judge the necessity of a diversification 
-model.
-Upon finding a diversification model is warranted, 
-it can then be added to BEAST2's existing set of (birth-death and 
-coalescent) models via the BEAST2 package system.
+any given phylogeny, helping us to validate new diversification models.
+\giovanni{I think we should mention the fact that BEAST2 allows to implement customized tree priors more complicated than the simple yule and bd.
+These tree priors, in the end, are what we call likelihood in our models, so it's pivotal in proposing our argument, I think.}
 
@@
   \centering
-  \includegraphics[width=\textwidth]{workflow.png}
+  \includegraphics[width=\textwidth]{overview.png}
   \caption{pirouette pipeline}
@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given reconstructed phylogeny, within the context of
-a simulation study on a (possibly new) diversification model. Such
-a study will typically construct phylogenies for different combinations of
-the diversification model's parameters, to assess under which scenario's
-the error BEAST2 makes cannot be ignored anymore. Within such a study,
-there need be many replicates.
-
-To be able to explain how \verb;pirouette; measures the inference error
-BEAST2 makes, we first give two definitions of what
-we mean by a generative model and by an inference model: 
-a generative model is a combination of site model, clock model and tree prior,
+makes from a given reconstructed phylogeny \giovanni{We should mention here that the phylogenies that are meant to be submitted to pirouette are simulated phylogenies generated according to the model we want to test}.
+For this measurement, two models need to be specified: a generative and an inference model.
+A generative model is a combination of site model, clock model and tree prior,
 used to simulate alignments. An inference model is a combination of 
@@
 time-calibrated phylogeny, by specifying a normal distribution around
-an expected time when the ancestors of two or more taxa diversified.
+an expected time when two or more taxa's \giovanni{I have the feeling that taxa should be plural here. Should we mention that the group must be monophyletic?} ancestors diversified.
 
 \subsection{pirouette's pipeline}
-The pipeline is summarized by the following steps,
-which will be described in detail below:
+The pipeline is summarized by the following steps, which will be described in detail below:
 \begin{enumerate}
@@
 \end{enumerate}
-The pipeline is visualized in \ref{fig:pipeline}.
+\richel{figure to pipeline here}
 There is also the option to generate a 'twin tree',
@@
 topology as the given phylogeny, yet its branch lengths follow the same
-distribution as a Yule (pure-birth) or (constant rate) birth-death tree prior.
+distribution as a Yule (pure-birth) or (constant rate) birth-death tree prior \giovanni{let's revise this after we meet Rampal and define how a twin tree should be created}.
 The use of such a twin tree, is to assess the minimum level of noise (i.e. 
@@
 
-We start with an idealised Yule (pure-birth) tree 
+We start with an idealised \giovanni{the word idealized cannot be used unless we specify what it means.} Yule (pure-birth) tree 
 with five taxa and a crown age of ten.
@@
 documentation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[h]
-\centering
-\begin{tabular}{ | l | l | }
-\hline
-\textbf{Name} & \textbf{Description} \\
-\hline
-\verb;pir_run; & Run \verb;pirouette; \\
-\verb;create_pir_params; & Create the \verb;pirouette; parameters \\
-\hline
-\verb;create_alignment_params; & Create the alignment parameters \\
-\verb;create_twinning_params; & Create the twinning parameters \\
-\verb;create_experiment; & Create one experiment \\
-\verb;create_error_measure_params; & Create the error measurement parameters \\
-\hline
-\end{tabular}
-\caption{pirouette's main functions}
-\label{tab:functions}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\giovanni{it could be nice to add a listing with simple commands that refer to the most useful examples}
 
@@
 
-In figure ...
+\giovanni{We need to discuss the results we got from the 3 experiments.}
+\richel{Only if the journal requires this}
 

### 2019-02-28 article.tex

@@
     BEAST2 is a popular Bayesian phylogenetics software tool,
-    that takes an alignment and inference model to create a
+    that takes an alignment and an inference model to create a
     posterior of jointly-estimated phylogenies and model parameter estimates.
     When a new macro-evolutionary diversification model is developed,
-    a good first step is to measure the error BEAST2 makes on a known
-    phylogeny derived from a new diversification mechanism, 
-    with its current set of inference models. \\
+    a good first step is to measure the error BEAST2 makes, with its current set of inference models, on a known
+    phylogeny derived from a new diversification mechanism. \\
   \textbf{2. }
-    Here, we present a free, libre and open-source R package, \verb;pirouette;
-    that assesses the inference error BEAST2 makes based on a known/true 
+    Here we present \verb;pirouette;, a free, libre and open-source R package that assesses the inference error BEAST2 makes based on a known/true 
     phylogeny. \\
@@
   \textbf{4. }
-    Last, we'll discuss the results obtained by the examples. \\
+    Last, we will discuss the results obtained by the examples. \\
 \end{abstract}
@@
 Such tools have been increasingly exploited to hypothesize 
-and test what the main drivers and modes for diversification processes are.
+and test what the main drivers and modes for diversification are.
 
-There are plenty of diversification models: 
-birth-death models that have rates that are 
+Since BEAST performs a bayesian analysis, it needs not only genetic data but also tree priors, describing how the diversification occur, to return posterior phylogenies.
+BEAST works using standard tree priors like Yule or birth-death \giovanni{integrate this if necessary}.
+The recent development of BEAST2 [\cite{bouckaert2014beast}] provides, unlike its predecessor, the possibility for third-party users to integrate its prior libraries with new ones.
+Many of these have been developed, in the form of likelihood diversification models: birth-death models that have rates that are
 constant [\cite{nee1994reconstructed}], 
@@
 diversity-dependent [\cite{etienne2011diversity}],
-or can shift [the SLS model. Laudanno, Haegeman \& Etienne, in preparation].
+or can shift [the SLS model. Laudanno et al., in preparation].
 There are birth death models in which speciation is a process that 
 takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
-can co-occur [Laudanno, Bilderbeek, Haegeman \& Etienne, in preparation],
+can co-occur [Laudanno et al., in preparation],
 dependent on a trait that is binary [\cite{maddison2007estimating}], 
@@
 
+\iffalse
 Such models usually rely on the assumption 
@@
 Starting from a phylogeny, this likelihood is maximized
-to find the parameters that are most likely.
+to infer the best estimates for the parameter values.
 For such diversification models, it is a standard practice 
@@
 measuring the parameters deduced by maximum likelihood. 
-It is a sign of quality of a diversification model 
-if its parameters can be recovered well by this method.
+\fi
+
+From a given phylogeny, the likelihood is maximized
+to infer the best estimates for the parameter values.
+For such  models, it is a standard practice to assess their performance on a controlled environment: this is done by simulating phylogenies from known parameters, then comparing the estimations obtained by maximum likelihood with the original parameters.
+
 \richel{
@@
 }
+\giovanni{
+I will try writing a first draft of this
+}
+ If the likelihood proves to be effective to describe the process, it can be implemented into BEAST2 as a new prior.
+Despite that it is still possible that already existing BEAST2 priors will perform well enough and, therefore, that the introduction of a new model is not necessary.
 
@@
 and is built on \verb;babette; [Bilderbeek \& Etienne, 2018], 
-which calls the popular Bayesian inference tool 
-BEAST2 [\cite{bouckaert2014beast}]. 
+which calls BEAST2 [\cite{bouckaert2014beast}]. 
 With \verb;pirouette;, one
 can easily measure the error made by Bayesian inference in recovering
-any given phylogeny, helping us to judge the necessity of a diversification 
-model.
+any given phylogeny, helping us to evaluate the necessity of a new diversification model.
+Once verified that the introduction of a new model is indeed beneficial,
+\iffalse
 Upon finding a diversification model is warranted, 
+\fi
 it can then be added to BEAST2's existing set of (birth-death and 
@@
 
-\begin{sidewaysfigure}
-  \centering
-  \includegraphics[width=\textwidth]{workflow.png}
-  \caption{pirouette pipeline}
-  \label{fig:pipeline}
-\end{sidewaysfigure}
-
+\iffalse
 The goal of \verb;pirouette; is to measure the inference error BEAST2
@@
 a simulation study on a (possibly new) diversification model. Such
-a study will typically construct phylogenies for different combinations of
-the diversification model's parameters, to assess under which scenario's
-the error BEAST2 makes cannot be ignored anymore. Within such a study,
-there need be many replicates.
+a study will typically construct phylogenies for different combinations of the diversification model's parameters, to assess under which scenarios the error made by BEAST2 cannot be ignored anymore. 
+Such a study needs to be done on, there need be many replicates.
+\fi
+
+The goal of \verb;pirouette; is to measure the inference error made by BEAST2 from a given reconstructed phylogeny, which has been simulated according to a (possibly new) diversification model.
+To do that the full procedure would be to simulate a great number of phylogenies for different combinations of the diversification model's parameters, to assess under which scenarios the error made by BEAST2 cannot be ignored anymore.
 
@@
 used to produce a posterior. Node calibration priors, called 'most
-recent common ancestor' (MRCA) priors within BEAST2 allow to create a 
-time-calibrated phylogeny, by specifying a normal distribution around
-an expected time when the ancestors of two or more taxa diversified.
+recent common ancestor' (MRCA) priors within BEAST2, allow to create a time-calibrated phylogeny, by specifying a normal distribution around an expected time when the ancestors of two or more taxa diversified.
 
 \subsection{pirouette's pipeline}
-The pipeline is summarized by the following steps,
-which will be described in detail below:
+
+\begin{sidewaysfigure}
+  \centering
+  \includegraphics[width=\textwidth]{workflow.png}
+  \caption{
+  pirouette pipeline
+  \giovanni{
+  see \texttt{https://github.com/richelbilderbeek/pirouette\_article/issues/9}
+  }
+  }
+  \label{fig:pipeline}
+\end{sidewaysfigure}
+
+The pipeline is summarized by the following steps, which will be described in detail below:
 \begin{enumerate}
@@
 \end{enumerate}
-The pipeline is visualized in \ref{fig:pipeline}.
+The pipeline is visualized in Fig.~\ref{fig:pipeline}.
 There is also the option to generate a 'twin tree',
@@
 are the Markov-Chain Monte Carlo (MCMC) setup, 
-an optional Most Recent Common Ancestor (MRCA) prior and an RNG seed.
+an optional Most Recent Common Ancestor (MRCA) \giovanni{MRCA has been already been defined. It is therefore possible to use the acronym only, if you want} prior and an RNG seed.
 The MCMC setup determines the number of posterior trees sampled.
@@
 the part of the MCMC chain that has not yet converged to a
-representative part of state space.
+representative part of the state space.
+
+\subsection{Twin tree}
+
+\iffalse
+An optional step is to generate a 'twin tree', that will be
+analyzed in the same way as the true tree.
+The twin tree has the same topology as the given phylogeny, yet its branch lengths follow the same distribution as a Yule (pure-birth) or (constant rate) birth-death tree prior \giovanni{let's revise this after we meet Rampal and define how a twin tree should be created}.
+The use of such a twin tree, is to assess the minimum level of noise (i.e. error) BEAST2 makes when the input (the twin phylogeny) has an ideal branch length distribution.
+\fi
 
 An optional step is to generate a 'twin tree', that will be
-analyzed in the same way as the true tree. The twin tree has the same
-topology as the given phylogeny, yet its branch lengths follow the same
-distribution as a Yule (pure-birth) or (constant rate) birth-death tree prior \giovanni{let's revise this after we meet Rampal and define how a twin tree should be created}.
-The use of such a twin tree, is to assess the minimum level of noise (i.e. 
-error) BEAST2 makes when the input (the twin phylogeny) has an ideal branch
-length distribution.
+analyzed in the same way as the true tree.
+The twin tree has the same topology as the given phylogeny, yet its branching times are simulated according to the tree prior used by BEAST2 to produce a posterior.
+The goal of this procedure is to generate a tree to use as a control for the main experiment.
+The procedure to generate a twin tree is the following:
+
+From the given phylogeny we maximize a birth-death likelihood [\cite{nee1994reconstructed}] to infer the standard birth-death rates, namely speciation rate $\lambda$ and extinction rate $\mu$.
+
+We use $\lambda$ and $\mu$ to simulate a new birth-death tree, conditioned on having the same number of tips as the original tree. We store its branching times.
+
+We then combine such branching times with the original tree's topology to obtain the twin tree.
 
@@
 A first research question that \verb;pirouette; answers is:
-What is the error BEAST2 makes from a phylogeny using the same 
-diversification model as it was generated by?
+"What is the error BEAST2 makes \giovanni{I would prefer "the error made by BEAST2" in place of "the error BEAST2 makes"} from a phylogeny using the same diversification model as it was generated by?"
 
+\iffalse
 We start with an idealised \giovanni{the word idealized cannot be used unless we specify what it means.} Yule (pure-birth) tree 
@@
 the most common branch length distribution imaginable. We
-use an idealised tree, as we want to measure the baseline noise; that is,
-the lowest error that BEAST2 can possibly make.
+use an idealised tree, as we want to measure the baseline noise; that is, the lowest error that BEAST2 can possibly make.
+\fi
+
+We start with a Yule (pure-birth) tree with five taxa and a crown age of ten.
 
@@
 
-The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny. To do so, we must specify the DNA root sequence
-and a mutation rate. In this example, the DNA root sequence consists
-out of four block of 250 nucleotides each, where the per-nucleotide
-mutation rate is 0.1 mutations per unit time.
+The first step in \verb;pirouette; is to simulate a DNA alignment from the given phylogeny.
+To do so, we must specify the DNA root sequence and a mutation rate.
+In this example, the DNA root sequence consists out of four block of 250 nucleotides each, where the per-nucleotide mutation rate is 0.1 mutations per unit time.
 
@@
 
+\iffalse
 In the second step we state our experiments. In this context, we
@@
 JC site model, strict clock model and Yule tree prior).
+\fi
+
+In the second step we state our experiments. In this context, we
+define an experiment as a combination of an inference model
+and the conditions under which a Bayesian inference is executed.
+Within this example, we specify that our experiment uses the
+generative model (which means that the inference model will be a combination of JC site model, strict clock model and Yule tree prior), will always be run and we don't need to measure the evidence for our model.
 
@@
   }
+  \label{tbl:RQ1}
 \end{table}
@@
   caption = {
-    Create a default experiment. This default experiment is
-    for a generative model, that is always run, without its evidence measured,
-    assuming a JC site model, strict clock model, Yule inference model.
+    Create a default experiment, as described in Table~\ref{tbl:RQ1}.
   }
@@
 A second research question that \verb;pirouette; answers, is:
-What is the error BEAST2 makes from a phylogeny when
-picking the best inference model?
+"What is the error BEAST2 makes from a phylogeny when
+picking the best inference model?"
 
@@
   label = {lst:unknown_phylogeny},
-  caption = "A phylogeny with an unknown diversification model"
+  caption = A phylogeny generated by an unknown diversification model
 ]
@@
 
-The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny. We'll re-use the alignment parameters of the previous example
-as shown in listing \ref{lst:create_alignment}.
+The first step in \verb;pirouette; is to simulate a DNA alignment from the given phylogeny.
+We will re-use the alignment parameters of the previous example as shown in listing \ref{lst:create_alignment}.
 
@@
 
+\giovanni{After having explained that we are going to use as candidates all the possible inference models, it comes natural to wonder what is the best inference model selected by marginal likelihoods.
+It could be nice maybe, adding the name in the lengend on top-right, in place of the generic "candidate" word.
+It could be also nice to add it in the caption and/or in the main text.}
+
 \begin{figure}[h]
@@
 A third research question that \verb;pirouette; answers is:
-What is the error BEAST2 makes from a phylogeny, 
-when hand-picking an inference model, compared to the background noise?
+"What is the error BEAST2 makes from a phylogeny, 
+when hand-picking an inference model, compared to the background noise?"
 
@@
 
+\iffalse
 This time, we are interested in creating a twin tree. A twin tree
@@
 its minimal error.
-Creating a twinning parameter is easy,
-as it has sensible default settings:
+\fi
+
+This time, we are interested in creating a twin tree. A twin tree
+has the same topology as the given tree, yet its branching times are simulated according to the tree prior used by BEAST2 to produce a posterior.
+Since there is coherence between the inference model and the generative model, we expect BEAST2 to produce a posterior of phylogenies clearly more similar to the original one.
+We intend to use this as a control experiment.
+
+Creating a twinning parameter is easy, as it has sensible default settings:
 
@@
 
-We can now measure the errors BEAST2
-makes when inferring the given phylogeny, compared
-to the background noise, like this:
+We can now measure the errors made by BEAST2 when inferring the given phylogeny versus the error it makes when the same procedure is applied to the twin tree.
 
-All the \verb;pirouette; arguments are bundled
-and checked by \verb;create_pir_params;:
+All the \verb;pirouette; arguments are bundled and checked by \verb;create_pir_params;:
 
@@
   \caption{
+    The error BEAST2 makes from a phylogeny picking the best inference model versus the control.
+    Here, the twin column shows the error BEAST2 makes starting from the original tree versus the twin. 
+  }
+  \label{fig:example_3}
+\end{figure}
+
+\iffalse
 The error BEAST2 makes from a phylogeny 
@@
 tree, to measure the noise, which is the minimal error. 
-  }
-  \label{fig:example_3}
-\end{figure}
+\fi
 
@@
 Figure \ref{fig:example_1} shows that the generative model 
-gives an error distribution with a highest density at 0.02,
-where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0.
-The range of errors goes from 0.0 to 0.27. As we know the generative model,
-this error distribution is the minimum noise that BEAST2 produces for this
-generative model.
+gives an error distribution with a highest density \giovanni{probably "median" is a better word} at 0.02 \giovanni{I would increase the precision to the second significative digit, e.g. 0.017},
+where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0 \giovanni{providing standard deviation could be a better choice}.
+The range of errors goes from 0.0 to 0.27 \giovanni{this is contrast with the last sentence}.
+As we know the generative model, this error distribution is the minimum noise that BEAST2 produces for this generative model.
 \richel{Show effective sample size}
 
-Figure \ref{fig:example_2} shows that the best candidate model 
-gives an error distribution with a highest density at 0.02,
-where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0.
+Figure \ref{fig:example_2} \giovanni{same as RQ1} shows that the best candidate model gives an error distribution with a highest density at 0.02, where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0.
 The range of errors goes from 0.0 to 0.08.
@@
 
-Figure \ref{fig:example_3} shows that the hand-picked model 
-gives an error distribution with a highest density at 0.018,
-with a range from 0.0 to 0.1.
+Figure \ref{fig:example_3} \giovanni{same as RQ1} shows that the hand-picked model gives an error distribution with a highest density at 0.018, with a range from 0.0 to 0.1.
 The twin tree has an error distribution with a highest density at 0.019,
 with a range from 0.0 to 0.11, which is the noise caused by the tree prior.
+\giovanni{I am still wondering if "noise" is the right term to use here.
+I still don't have a precise opinion on that.
+The problem is that for me it's a bit hard to interpret this without a distribution...}
 As there is little difference between the true and twin tree,
@@
 added to BEAST2.
+
+\giovanni{since you refer, for every RQ, to the same values, it could be better to show them as a table}
 
@@
 
-\verb;pirouette;'s documentation is extensive. All functions are documented
-in the package's internal documentation. For quick use, 
+\verb;pirouette;'s documentation is extensive. All functions are documented in the package's internal documentation. For quick use, 
 each exported function shows a minimal example. 
 For easy exploration, each exported function's documentation links to related functions.
-Additionally, \verb;pirouette; has a vignette that demonstrates extensively how
-to use it.
+Additionally, \verb;pirouette; has a vignette that demonstrates extensively how to use it.
 

### 2019-02-28 article.tex
Improved pipeline figure, fix https://github.com/richelbilderbeek/pirouette_article/issues/9

@@
 
-\begin{sidewaysfigure}
+\begin{figure}
   \centering
@@
     pirouette pipeline
-  \giovanni{
-  see \texttt{https://github.com/richelbilderbeek/pirouette\_article/issues/9}
-  }
   }
   \label{fig:pipeline}
-\end{sidewaysfigure}
+\end{figure}
 
@@
 are the Markov-Chain Monte Carlo (MCMC) setup, 
-an optional Most Recent Common Ancestor (MRCA) \giovanni{MRCA has been already been defined. It is therefore possible to use the acronym only, if you want} prior and an RNG seed.
+an optional MRCA prior and an RNG seed.
 The MCMC setup determines the number of posterior trees sampled.
@@
 analyzed in the same way as the true tree.
-The twin tree has the same topology as the given phylogeny, yet its branch lengths follow the same distribution as a Yule (pure-birth) or (constant rate) birth-death tree prior \giovanni{let's revise this after we meet Rampal and define how a twin tree should be created}.
+The twin tree has the same topology as the given phylogeny, yet its branch lengths follow the same distribution as a Yule (pure-birth) or (constant rate) birth-death tree prior 
+\giovanni{let's revise this after we meet Rampal and define how a twin tree should be created}.
 The use of such a twin tree, is to assess the minimum level of noise (i.e. error) BEAST2 makes when the input (the twin phylogeny) has an ideal branch length distribution.
@@
 A first research question that \verb;pirouette; answers is:
-"What is the error BEAST2 makes \giovanni{I would prefer "the error made by BEAST2" in place of "the error BEAST2 makes"} from a phylogeny using the same diversification model as it was generated by?"
-
-\iffalse
-We start with an idealised \giovanni{the word idealized cannot be used unless we specify what it means.} Yule (pure-birth) tree 
-with five taxa and a crown age of ten.
-By idealized, we mean a phylogeny that is easiest to recover
-by the Bayesian inference. A phylogeny that is easy to recover has
-the most common branch length distribution imaginable. We
-use an idealised tree, as we want to measure the baseline noise; that is, the lowest error that BEAST2 can possibly make.
-\fi
+"What is the error made by BEAST2 from a phylogeny using the same diversification model as it was generated by?"
 
@@
 Figure \ref{fig:example_1} shows that the generative model 
-gives an error distribution with a highest density \giovanni{probably "median" is a better word} at 0.02 \giovanni{I would increase the precision to the second significative digit, e.g. 0.017},
-where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0 \giovanni{providing standard deviation could be a better choice}.
-The range of errors goes from 0.0 to 0.27 \giovanni{this is contrast with the last sentence}.
+gives an error distribution with a highest density 
+\giovanni{probably "median" is a better word}
+at 0.02 
+\giovanni{I would increase the precision to the second significative digit, e.g. 0.017},
+where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0 
+\giovanni{providing standard deviation could be a better choice}.
+The range of errors goes from 0.0 to 0.27 
+\giovanni{this is contrast with the last sentence}. 
 As we know the generative model, this error distribution is the minimum noise that BEAST2 produces for this generative model.

### 2019-02-28 article.tex
Re-create figure 1, fix https://github.com/richelbilderbeek/pirouette_article/issues/10

@@
   \caption{
-    pirouette pipeline
+    \texttt{pirouette} pipeline. 
+    The pipeline starts from a true phylogeny (1a), 
+    that is converted to an alignment (2a), which is used
+    to create a BEAST2 posterior of phylogenies (3a),
+    which is compared to the true phylogeny (1a) using
+    the nLTT statistic (4a), resulting in an error distribution (5a).
+    Optionally, a twin phylogeny (1b) can be generated from a true
+    phylogeny (1a), which then results in an error distribution (5b) with
+    the similar intermediate steps 
   }

### 2019-02-28 article.tex
Better caption for fig 1

--- 2019-02-28 article.tex
@@
   \caption{
-    pirouette pipeline
+    \texttt{pirouette} pipeline. 
+    The pipeline starts from a true phylogeny (1a), 
+    that is converted to an alignment (2a), which is used
+    to create a BEAST2 posterior of phylogenies (3a),
+    which is compared to the true phylogeny (1a) using
+    the nLTT statistic. (4a) shows the nLTT of the true tree in red,
+    the nLTT of all individual posterior trees in grey, and the average
+    nLTT statistic of all posterior trees in black.
+    Taking the differences between the true tree's nLTT and each posterior's
+    nLTT results in an error distribution, here displayed as a histogram
+    of error values (5a).
+    Optionally, a twin phylogeny (1b) can be generated from a true
+    phylogeny (1a), which then results in an error distribution (5b) with
+    the similar intermediate steps 
   }

### 2019-02-28 article.tex
Add example 4 with figure

@@
 \usepackage{bm}
+\usepackage{multirow}
 
@@
   \textbf{2. }
-    Here we present \verb;pirouette;, a free, libre and open-source R package that assesses the inference error BEAST2 makes based on a known/true 
+    Here we present \verb;pirouette;, 
+    a free, libre and open-source R package that assesses 
+    the inference error BEAST2 makes based on a known/true 
     phylogeny. \\
@@
 The development of new powerful inference tools, 
-such as BEAST [\cite{drummond2007beast}], MrBayes [\cite{huelsenbeck2001mrbayes}]
+such as BEAST [\cite{drummond2007beast}], 
+MrBayes [\cite{huelsenbeck2001mrbayes}]
 or RevBayes [\cite{hohna2016revbayes}], 
@@
 
-Since BEAST performs a bayesian analysis, it needs not only genetic data but also tree priors, describing how the diversification occur, to return posterior phylogenies.
-BEAST works using standard tree priors like Yule or birth-death \giovanni{integrate this if necessary}.
-The recent development of BEAST2 [\cite{bouckaert2014beast}] provides, unlike its predecessor, the possibility for third-party users to integrate its prior libraries with new ones.
-Many of these have been developed, in the form of likelihood diversification models: birth-death models that have rates that are
+Since BEAST performs a Bayesian analysis, 
+it needs not only genetic data but also tree priors, 
+describing how the diversification occur, to return posterior phylogenies.
+BEAST works using standard tree priors like Yule or birth-death 
+\giovanni{integrate this if necessary}.
+The recent development of BEAST2 [\cite{bouckaert2014beast}] provides, 
+unlike its predecessor, the possibility for third-party users 
+to integrate its prior libraries with new ones. 
+Many of these have been developed, 
+in the form of likelihood diversification models: 
+birth-death models that have rates that are 
 constant [\cite{nee1994reconstructed}], 
@@
 to infer the best estimates for the parameter values.
-For such  models, it is a standard practice to assess their performance on a controlled environment: this is done by simulating phylogenies from known parameters, then comparing the estimations obtained by maximum likelihood with the original parameters.
-
-\richel{
-  We should bridge this method with how pirouette does things differently;
-  why the maximum likelihood does not apply in this context  
-}
-\giovanni{
-I will try writing a first draft of this
-}
- If the likelihood proves to be effective to describe the process, it can be implemented into BEAST2 as a new prior.
-Despite that it is still possible that already existing BEAST2 priors will perform well enough and, therefore, that the introduction of a new model is not necessary.
+For such  models, it is a standard practice to assess their performance 
+on a controlled environment: this is done by simulating phylogenies from known parameters, 
+then comparing the estimations obtained by maximum likelihood with the original parameters.
+If the likelihood proves to be effective to describe the process, 
+it can be implemented into BEAST2 as a new prior. 
+Despite that it is still possible that already existing BEAST2 priors 
+will perform well enough and, therefore, 
+that the introduction of a new model is not necessary.
 
@@
 Once verified that the introduction of a new model is indeed beneficial,
-\iffalse
-Upon finding a diversification model is warranted, 
-\fi
 it can then be added to BEAST2's existing set of (birth-death and 
@@
 
-The first step simulates a DNA alignment from the given phylogeny.
+The first step simulates a DNA alignment from the given 
+phylogeny (Fig.~\ref{fig:pipeline} 1a $\rightarrow$ 2a).
 One can specify a DNA sequence
@@
 
-The third step infers a Bayesian posterior from the simulated alignment,
+The third step infers a Bayesian posterior from the simulated 
+alignment (Fig.~\ref{fig:pipeline} 2a $\rightarrow$ 3a),
 using the inference model(s) selected in the previous step. The user
@@
 
-The fourth step measures the inference error, using the phylogenies in the
-Bayesian posterior. These phylogenies are compared to the given
-original phylogeny using an error statistic, which is the nLTT 
-statistic (\cite{janzen2015approximate}) by default,
-but also a custom error statistic is supplied, based
-on the absolute difference in the gamma statistic [\cite{pybus2000testing}]. 
+The fourth step compares the true tree to each posterior tree
+error using
+the nLTT statistic (\cite{janzen2015approximate}) 
+(Fig.~\ref{fig:pipeline} 3a $\rightarrow$ 4a), resulting
+in an error distribution (Fig.~\ref{fig:pipeline} 4a $\rightarrow$ 5a).
+\richel{I would not mind to remove 4a and 4b from the figure}
+The nLTT statistic is used by default,
+but also a user-defined error statistic can be used. As an example,
+\verb;pirouette; suplies such a custom error statistic,
+that uses an absolute difference in the gamma statistic [\cite{pybus2000testing}].
 Additionally, the user can specify the
@@
 
-\iffalse
-An optional step is to generate a 'twin tree', that will be
-analyzed in the same way as the true tree.
-The twin tree has the same topology as the given phylogeny, yet its branch lengths follow the same distribution as a Yule (pure-birth) or (constant rate) birth-death tree prior 
-\giovanni{let's revise this after we meet Rampal and define how a twin tree should be created}.
-The use of such a twin tree, is to assess the minimum level of noise (i.e. error) BEAST2 makes when the input (the twin phylogeny) has an ideal branch length distribution.
-\fi
-
-An optional step is to generate a 'twin tree', that will be
-analyzed in the same way as the true tree.
-The twin tree has the same topology as the given phylogeny, yet its branching times are simulated according to the tree prior used by BEAST2 to produce a posterior.
-The goal of this procedure is to generate a tree to use as a control for the main experiment.
+An optional step is to generate a 'twin tree' (Fig.~\ref{fig:pipeline} 1a $\rightarrow$ 1b), 
+that will be analyzed in the same way as the true tree.
+The twin tree has the same topology as the given phylogeny, 
+yet its branching times are simulated according to 
+the tree prior used by BEAST2 to produce a posterior.
+The goal of this procedure is to generate a tree 
+to use as a control for the main experiment.
 The procedure to generate a twin tree is the following:
 
-From the given phylogeny we maximize a birth-death likelihood [\cite{nee1994reconstructed}] to infer the standard birth-death rates, namely speciation rate $\lambda$ and extinction rate $\mu$.
-
-We use $\lambda$ and $\mu$ to simulate a new birth-death tree, conditioned on having the same number of tips as the original tree. We store its branching times.
-
-We then combine such branching times with the original tree's topology to obtain the twin tree.
+From the given (true) phylogeny, we maximize a birth-death 
+likelihood [\cite{nee1994reconstructed}] to infer the standard 
+birth-death rates, namely speciation rate $\lambda$ and extinction rate $\mu$.
+We use $\lambda$ and $\mu$ to simulate a new birth-death tree, 
+conditioned on having the same number of tips as the original tree. 
+We store its branching times. 
+We then combine such branching times with the original tree's 
+topology to obtain the twin tree.
 
@@
     language=R, floatplacement=H, frame=single, 
-    label = {lst:create_ideal_tree}, 
-    caption = {Create an ideal tree}
+    label = {lst:create_yule_tree}, 
+    caption = {Create a Yule tree}
   ]
@@
   \includegraphics[width=\textwidth]{figure_bd.png}
-  \caption{The ideal tree, as created by listing \ref{lst:create_ideal_tree}}
+  \caption{The Yule tree, as created by listing \ref{lst:create_yule_tree}}
 \end{figure}
@@
 To do so, we must specify the DNA root sequence and a mutation rate. 
-In this example, the DNA root sequence consists out of four block of 250 nucleotides each, where the per-nucleotide mutation rate is 0.1 mutations per unit time.
+In this example, the DNA root sequence consists out of four block of 250 nucleotides each, 
+where the per-nucleotide mutation rate is 0.1 mutations per unit time.
 
@@
   \caption{
-    The error BEAST2 makes from a phylogeny 
+    Example 1: the error BEAST2 makes from a phylogeny 
     when the generative and inference model are the same.
+    The x-axis specifies the tree type (which can be 'true' or 'twin'),
+    the y-axis shows the error density
   }
@@
 
-\giovanni{After having explained that we are going to use as candidates all the possible inference models, it comes natural to wonder what is the best inference model selected by marginal likelihoods.
-It could be nice maybe, adding the name in the lengend on top-right, in place of the generic "candidate" word.
-It could be also nice to add it in the caption and/or in the main text.}
+\giovanni{describe winner}
 
@@
   \caption{
-    The error BEAST2 makes from a phylogeny when
-    picking the best inference model
+    Example 2: the error BEAST2 makes from a phylogeny when
+    picking the best inference model.
+    The x-axis specifies the tree type (which can be 'true' or 'twin'),
+    the y-axis shows the error density
   }
@@
 This time, we are interested in creating a twin tree. A twin tree
-has the same topology as the given tree, yet its branching times are simulated according to the tree prior used by BEAST2 to produce a posterior.
-Since there is coherence between the inference model and the generative model, we expect BEAST2 to produce a posterior of phylogenies clearly more similar to the original one.
+has the same topology as the given tree, 
+yet its branching times are simulated according 
+to the tree prior used by BEAST2 to produce a posterior. 
+Since there is coherence between the inference model and the generative model, 
+we expect BEAST2 to produce a posterior of phylogenies clearly 
+more similar to the original one. 
 We intend to use this as a control experiment.
@@
   \caption{
-    The error BEAST2 makes from a phylogeny picking the best inference model versus the control.
+    Example 3: the error BEAST2 makes from a phylogeny picking the best inference model versus the control.
     Here, the twin column shows the error BEAST2 makes starting from the original tree versus the twin. 
+    The x-axis specifies the tree type (which can be 'true' or 'twin'),
+    the y-axis shows the error density
   }
@@
 \fi
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Usage: fourth research question}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A fourth research question that \verb;pirouette; answers, is:
+"What is the error BEAST2 makes from a phylogeny for a known
+generative model, compared to candidates with similar inference models,
+in relationship to the background error?"
+
+Here, same as example 2, we start with a tree generated by the Yule process,
+by using the code shown in listing \ref{lst:create_yule_tree}.
+The first step in \verb;pirouette; is to simulate a DNA alignment 
+from the given phylogeny, for which we'll use the same code 
+as shown in listing \ref{lst:create_alignment}.
+
+\begin{table}
+  \begin{tabular}{ | c | c | c | l | }
+    \hline
+    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & \textbf{inference model} \\ 
+    \hline
+    generative & always         & yes & JC, strict, Yule \\
+    candidate  & best candidate & yes & JC, strict, BD \\
+    candidate  & best candidate & yes & JC, strict, CCP \\
+    candidate  & best candidate & yes & JC, strict, CEP \\
+    \hline
+  \end{tabular}
+  \caption{
+    Experimental setup to answer the fourth research question.
+    JC: Jukes-Cantor site model.
+    strict: strict clock model.
+    Yule: Yule (pure-birth) tree prior.
+    BD: birth-death tree prior.
+    CCP: coalescent constant-population tree prior.
+    CEP: coalescent exponential-population tree prior.
+  }
+  \label{tab:experiment_4}
+\end{table}
+
+In the second step we state our experiments. 
+Within this example, we must both specify the generative and candidate 
+experiments. Here we'll specify the generative model:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+generative_experiment <- create_experiment(
+  model_type = "generative",
+  run_if = "always",
+  do_measure_evidence = TRUE
+)
+\end{lstlisting}
+
+Specifying all candidates can be done in a factorial fashion using 
+\verb;create_all_experiments;. In this case, all candidates will use the
+same site and clock model, and only differ in their tree priors:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+candidate_experiments <- create_all_experiments(
+  site_models = list(create_jc69_site_model()),
+  clock_models = list(create_strict_clock_model()),
+  tree_priors = list(
+    create_bd_tree_prior(), 
+    create_ccp_tree_prior(), 
+    create_cep_tree_prior()
+  )
+)
+\end{lstlisting}
+
+We need to combine these experiments into one set:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+experiments <- c(list(generative_experiment), candidate_experiments)
+\end{lstlisting}
+
+Also here, the third (the BEAST2 inference) and fourth (measuring the error)
+steps have sensible defaults. To get an idea of a baseline error made by
+BEAST2, we also make use of the twinning option:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments,
+  twinning_params = create_twinning_params()
+)
+\end{lstlisting}
+
+Running the experiment:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+errors <- pir_run(
+  phylogeny,
+  pir_params = pir_params
+)
+\end{lstlisting}
+
+Again, showing the results:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+pir_plot(errors)
+\end{lstlisting}
+
+The resulting figure is shown in figure \ref{fig:example_4}.
+
+\giovanni{describe winner}
+
+\begin{figure}[h]
+  \includegraphics[width=\textwidth]{figure_example_4.png}
+  \caption{
+    Example 4: the error BEAST2 makes from a phylogeny for a known
+    generative model, compared to candidates with similar inference models,
+    in relationship to the background error.
+    The x-axis specifies the tree type,
+    the y-axis shows the error density
+  }
+  \label{fig:example_4}
+\end{figure}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[h]
+\centering
+\begin{tabular}{ | r | l | l | l | l | }
+\hline
+\multirow{2}{*}{\textbf{Example}} & \multicolumn{2}{c|}{\textbf{True tree}} 
+                                  & \multicolumn{2}{c|}{\textbf{Twin tree}} \\
+\cline{2-5}
+                                  & \textbf{HPD} & \textbf{range} & \textbf{HPD} & \textbf{range} \\
+\hline
+1  & $0.20$  & $0.0 - 0.27$  &         &               \\
+2  & $0.02$  & $0.0 - 0.08$  &         &               \\
+3  & $0.018$ & $0.0 - 0.1$   & $0.019$ & $0.0 - 0.11$  \\
+4C & $0.22$  & $0.0 - 0.088$ & $0.009$ & $0.0 - 0.056$ \\
+4G & $0.020$ & $0.0 - 0.075$ & $0.013$ & $0.0 - 0.052$ \\
+\hline
+\end{tabular}
+\caption{Results of the examples. Errors are shows with the 
+  highest posterior density first, and the range of the errors in
+  round brackets second. Experiment 4 is split in two rows
+  by either the candidate ('C') or generative model ('G') 
+}
+\label{tab:results}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Here we discuss how to interpret the results of \verb;pirouette;.
 Figure \ref{fig:example_1} shows that the generative model 
-gives an error distribution with a highest density 
+gives an error distribution with a highest posterior density (HPD) 
 \giovanni{probably "median" is a better word}
+\richel{HPC is a valid Bayesian term I learned from the BEAST book}
 at 0.02
 \giovanni{I would increase the precision to the second significative digit, e.g. 0.017},
-where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0 
-\giovanni{providing standard deviation could be a better choice}.
 The range of errors goes from 0.0 to 0.27 
-\giovanni{this is contrast with the last sentence}. 
-As we know the generative model, this error distribution is the minimum noise that BEAST2 produces for this generative model.
+As we know the generative model, 
+this error distribution is the minimum noise that BEAST2 produces 
+for that generative model.
 \richel{Show effective sample size}
 
-Figure \ref{fig:example_2} \giovanni{same as RQ1} shows that the best candidate model gives an error distribution with a highest density at 0.02, where the range of errors (in this case, the nLTT statistic) is from 0.0 to 1.0.
-The range of errors goes from 0.0 to 0.08.
+Figure \ref{fig:example_2} shows that the best candidate model gives an error distribution 
+with an HPD at 0.02. The range of errors goes from 0.0 to 0.08. 
 \richel{Describe and show evidences for the candidate model}
@@
 
-Figure \ref{fig:example_3} \giovanni{same as RQ1} shows that the hand-picked model gives an error distribution with a highest density at 0.018, with a range from 0.0 to 0.1.
+Figure \ref{fig:example_3} shows that the hand-picked model 
+gives an error distribution with an HPD at 0.018, 
+with a range from 0.0 to 0.1.
 The twin tree has an error distribution with a highest density at 0.019,
-with a range from 0.0 to 0.11, which is the noise caused by the tree prior.
-\giovanni{I am still wondering if "noise" is the right term to use here.
-I still don't have a precise opinion on that.
-The problem is that for me it's a bit hard to interpret this without a distribution...}
+with a range from 0.0 to 0.11, which is an indication of the
+error made by BEAST on that tree prior.
 As there is little difference between the true and twin tree,
@@
 
-\giovanni{since you refer, for every RQ, to the same values, it could be better to show them as a table}
+\giovanni{
+  since you refer, for every RQ, to the same values, 
+  it could be better to show them as a table
+}
 

### 2019-02-28 scripts/example_4.R

@@ -1,0 +1,50 @@
+# Code of example 4
+#
+# Works under Linux and MacOS only
+library(pirouette)
+library(ggplot2)
+library(ggthemes)
+
+set.seed(314)
+
+phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
+
+alignment_params <- create_alignment_params(
+  root_sequence = create_blocked_dna(length = 1000),
+  mutation_rate = 0.1
+)
+
+# JC69, strict, Yule
+generative_experiment <- create_experiment(
+  model_type = "generative",
+  run_if = "always",
+  do_measure_evidence = TRUE
+)
+check_experiment(generative_experiment)
+
+# All non-Yule tree priors
+candidate_experiments <- create_all_experiments(
+  site_models = list(create_jc69_site_model()),
+  clock_models = list(create_strict_clock_model()),
+  tree_priors = list(create_bd_tree_prior(), create_ccp_tree_prior(), create_cep_tree_prior())
+)
+check_experiments(candidate_experiments)
+
+experiments <- c(list(generative_experiment), candidate_experiments)
+check_experiments(experiments)
+
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments,
+  twinning_params = create_twinning_params()
+)
+
+errors <- pir_run(
+  phylogeny,
+  pir_params = pir_params
+)
+
+pir_plot(errors) +
+  scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
+  theme_wsj() +
+  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_4.png")

### 2019-02-28 article.tex
Show first ESSes

@@
 \end{lstlisting}
+
+An overview of \verb;pirouette;'s main functions is shown in 
+table \ref{tab:functions}, which are all demonstrated in the examples
+below.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[h]
+\centering
+\begin{tabular}{ | l | l | }
+\hline
+\textbf{Name} & \textbf{Description} \\
+\hline
+\verb;pir_run; & Run \verb;pirouette; \\
+\verb;create_pir_params; & Create the \verb;pirouette; parameters \\
+\hline
+\verb;create_alignment_params; & Create the alignment parameters \\
+\verb;create_twinning_params; & Create the twinning parameters \\
+\verb;create_experiment; & Create one experiment \\
+\verb;create_error_measure_params; & Create the error measurement parameters \\
+\hline
+\end{tabular}
+\caption{pirouette's main functions}
+\label{tab:functions}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Other examples}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-There are many other examples that can be found in the \verb;pirouette; 
-documentation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[h]
-\centering
-\begin{tabular}{ | l | l | }
-\hline
-\textbf{Name} & \textbf{Description} \\
-\hline
-\verb;pir_run; & Run \verb;pirouette; \\
-\verb;create_pir_params; & Create the \verb;pirouette; parameters \\
-\hline
-\verb;create_alignment_params; & Create the alignment parameters \\
-\verb;create_twinning_params; & Create the twinning parameters \\
-\verb;create_experiment; & Create one experiment \\
-\verb;create_error_measure_params; & Create the error measurement parameters \\
-\hline
-\end{tabular}
-\caption{pirouette's main functions}
-\label{tab:functions}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
@@
 \end{tabular}
-\caption{Results of the examples. Errors are shows with the 
-  highest posterior density first, and the range of the errors in
-  round brackets second. Experiment 4 is split in two rows
+\caption{Results of the examples. For the true and twin tree,
+  the highest posterior density (HPC) and range of the errors 
+  is shown. Experiment 4 is split in two rows
   by either the candidate ('C') or generative model ('G') 
@@
 \richel{HPC is a valid Bayesian term I learned from the BEAST book}
-at 0.02
-\giovanni{I would increase the precision to the second significative digit, e.g. 0.017},
-The range of errors goes from 0.0 to 0.27 
+at 0.015
+The range of errors goes from 0.0 to 0.097 
 As we know the generative model, 
@@
 for that generative model.
-\richel{Show effective sample size}
+We measure the effective samples as well, and all are well above
+the recommended value of 200 \ref{drummond2015bayesian}, 
+as shown in table \ref{tab:esses_example_1}.
+
+\input{example_1_esses.latex}
+% has label tab:esses_example_1
 
@@
 added to BEAST2.
-
-\giovanni{
-  since you refer, for every RQ, to the same values, 
-  it could be better to show them as a table
-}
 

### 2019-02-28 scripts/example_1.R

@@ -5,6 +5,8 @@ library(pirouette)
 library(ggplot2)
 library(ggthemes)
 
+set.seed(314)
+
 phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
 
 alignment_params <- create_alignment_params(
@@ -26,10 +28,25 @@ errors <- pir_run(
 
 pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
-  theme_excel()
-
+  theme_wsj() +
+  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_1.png")
 
 pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
-  theme_wsj() +
-  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_1.png")
+  theme_excel()
+
+
+testit::assert(pir_params$experiments[[1]]$inference_model$mcmc$store_every != -1)
+
+
+esses <- tracerer::calc_esses(
+  traces = tracerer::parse_beast_log(pir_params$experiments[[1]]$beast2_options$output_log_filename),
+  sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
+)
+sink("/home/richel/GitHubs/pirouette_article/example_1_esses.latex")
+xtable::xtable(esses, caption = "ESSes of example 1", label = "tab:esses_example_1", digits = 0)
+sink()
+
+
+
+#  geom_text(stat='count', aes(label=..count..), vjust=-1)

### 2019-02-28 article.tex
Example 1 has ESSes

@@
 for that generative model.
-We measure the effective samples as well, and all are well above
-the recommended value of 200 \ref{drummond2015bayesian}, 
+We can measure the effective samples as well (not shown), 
+and all ESSes are above
+the recommended value of 200 \cite{drummond2015bayesian}, 
 as shown in table \ref{tab:esses_example_1}.
-
-\input{example_1_esses.latex}
-% has label tab:esses_example_1
 
@@
 
+\appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results in detail}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\input{example_1_esses.latex}
+% has label tab:esses_example_1
+
 \end{document}

### 2019-02-28 article.tex
Example 2 worked out better

@@
 with an HPD at 0.02. The range of errors goes from 0.0 to 0.08. 
-\richel{Describe and show evidences for the candidate model}
-\richel{Describe and show the best candidate model}
+The inference model with the highest evidence follows a JC site model, 
+RLN clock model and BD tree prior, with a weight of $0.675$. 
+This is (unexpectedly) better than the generative
+model with a weight of $0.166$. 
+The complete list of all evidences is shown in table \ref{tab:evidences_example_2}.
 This means that it is rational to use this inference model to 
@@
 as is done by \verb;pirouette;.
-\richel{Show effective sample size}
-\richel{Show posterior phylogenies as densitree}
+Also here, the ESSes are well above 200, 
+as shown in table \ref{tab:esses_example_2}.
 
@@
 
+\input{example_2_esses.latex}
+% has label tab:esses_example_2
+
+\input{example_2_evidences.latex}
+% has label tab:evidences_example_2
+
 \end{document}

### 2019-02-28 scripts/example_2.R

@@ -5,6 +5,8 @@ library(pirouette)
 library(ggplot2)
 library(ggthemes)
 
+set.seed(314)
+
 phylogeny  <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4) :1);")
 
 alignment_params <- create_alignment_params(
@@ -12,7 +14,14 @@ alignment_params <- create_alignment_params(
   mutation_rate = 0.1
 )
 
-experiments <- create_all_experiments()
+experiments <- create_all_experiments(
+  tree_priors = list(
+    create_bd_tree_prior(),
+    create_ccp_tree_prior(),
+    create_cep_tree_prior(),
+    create_yule_tree_prior()
+  )
+)
 
 pir_params <- create_pir_params(
   alignment_params = alignment_params,
@@ -28,3 +37,36 @@ pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
   theme_wsj() +
   ggsave("/home/richel/GitHubs/pirouette_article/figure_example_2.png")
+
+
+testit::assert(pir_params$experiments[[1]]$inference_model$mcmc$store_every != -1)
+esses <- tracerer::calc_esses(
+  traces = tracerer::parse_beast_log(pir_params$experiments[[1]]$beast2_options$output_log_filename),
+  sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
+)
+sink("/home/richel/GitHubs/pirouette_article/example_2_esses.latex")
+xtable::xtable(esses, caption = "ESSes of example 2", label = "tab:esses_example_2", digits = 0)
+sink()
+
+
+df_evidences <- utils::read.csv(pir_params$evidence_filename)[, c(-1, -6)]
+df_evidences$site_model_name <- plyr::revalue(df_evidences$site_model_name, c("JC69" = "JC", "TN93" = "TN"))
+df_evidences$clock_model_name <- plyr::revalue(
+  df_evidences$clock_model_name,
+  c("strict" = "Strict", "relaxed_log_normal" = "RLN")
+)
+df_evidences$tree_prior_name <- plyr::revalue(
+  df_evidences$tree_prior_name,
+  c("yule" = "Yule", "birth_death" = "BD", "coalescent_constant_population" = "CCP", "coalescent_exp_population" = "CEP")
+)
[…]

### 2019-02-28 article.tex
Create all ESSes and evidences and figures

@@
 \cline{2-5}
-                                  & \textbf{HPD} & \textbf{range} & \textbf{HPD} & \textbf{range} \\
-\hline
-1  & $0.20$  & $0.0 - 0.27$  &         &               \\
+                                  & \textbf{median} & \textbf{range} & \textbf{median} & \textbf{range} \\
+\hline
+1  & $0.02$  & $0.0 - 0.27$  &         &               \\
 2  & $0.02$  & $0.0 - 0.08$  &         &               \\
@@
 \end{tabular}
-\caption{Results of the examples. For the true and twin tree,
-  the highest posterior density (HPC) and range of the errors 
+\caption{Results of the examples. 
+  For the true and twin tree,
+  the median and range of the errors 
   is shown. Experiment 4 is split in two rows
@@
 
+\richel{check numbers}
+
 Here we discuss how to interpret the results of \verb;pirouette;.
 Figure \ref{fig:example_1} shows that the generative model 
-gives an error distribution with a highest posterior density (HPD) 
-\giovanni{probably "median" is a better word}
-\richel{HPC is a valid Bayesian term I learned from the BEAST book}
-at 0.015
-The range of errors goes from 0.0 to 0.097 
+gives an error distribution with a median of \richel{?}
+The range of errors goes from \richel{?} to \richel{?}.
 As we know the generative model, 
@@
 Figure \ref{fig:example_2} shows that the best candidate model gives an error distribution 
-with an HPD at 0.02. The range of errors goes from 0.0 to 0.08. 
+with a median of \richel{?}. 
+The range of errors goes from \richel{?} to \richel{?}. 
 The inference model with the highest evidence follows a JC site model, 
@@
 Figure \ref{fig:example_3} shows that the hand-picked model 
-gives an error distribution with an HPD at 0.018, 
-with a range from 0.0 to 0.1.
-The twin tree has an error distribution with a highest density at 0.019,
-with a range from 0.0 to 0.11, which is an indication of the
+gives an error distribution with a median of \richel{?}, 
+The twin tree has an error distribution with a median of \richel{?},
+with a range from 0.0 to \richel{?}, which is an indication of the
 error made by BEAST on that tree prior.
@@
 added to BEAST2.
+We can draw this conclusion with confidence, as the ESSes are well above 200, 
+as shown in table \ref{tab:esses_example_3}.
+
+Figure \ref{fig:example_4} shows that \richel{?}
+and 
+gives an error distribution with a median of \richel{?}, 
+The twin tree has an error distribution with a median of \richel{?},
+with a range from 0.0 to \richel{?}, which is an indication of the
+error made by BEAST on that tree prior.
+The inference model with the highest evidence follows a JC site model, 
+RLN clock model and BD tree prior, with a weight of ?. 
+The complete list of all evidences is shown in table \ref{tab:evidences_example_4}.
+As there is little difference between the true and twin tree,
+the (unknown) diversification model has little impact on the error
+BEAST2 makes, hinting that the constant-rate birth-death model is
+sufficiently good enough to explain the true phylogeny. 
+The diversification model apparently has minor impact and need not to be
+added to BEAST2.
+We can draw this conclusion with confidence, as the ESSes are well above 200, 
+as shown in table \ref{tab:esses_example_4}.
 
@@
 
+\richel{
+  I'd enjoy to put all intermediate output here: alignments, twin trees,
+  posteriors, nLTTs, errors. Now we only show the ones relevant in the text
+}
+
 \input{example_1_esses.latex}
@@
 
+\input{example_3_esses.latex}
+% has label tab:esses_example_3
+
+\input{example_4_esses.latex}
+% has label tab:esses_example_4
+
+\input{example_4_evidences.latex}
+% has label tab:evidences_example_4
+
 \end{document}

### 2019-02-28 scripts/example_3.R

@@ -5,8 +5,8 @@ library(pirouette)
 library(ggplot2)
 library(ggthemes)
 
-#phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
-# STUB
+set.seed(314)
+
 phylogeny <- create_stunning_tree(n_taxa = 5, t_0 = 10)
 
 alignment_params <- create_alignment_params(
@@ -33,3 +33,13 @@ pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
   theme_wsj() +
   ggsave("/home/richel/GitHubs/pirouette_article/figure_example_3.png")
+
+testit::assert(pir_params$experiments[[1]]$inference_model$mcmc$store_every != -1)
+esses <- tracerer::calc_esses(
+  traces = tracerer::parse_beast_log(pir_params$experiments[[1]]$beast2_options$output_log_filename),
+  sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
+)
+
+sink("/home/richel/GitHubs/pirouette_article/example_3_esses.latex")
+xtable::xtable(esses, caption = "ESSes of example 3", label = "tab:esses_example_3", digits = 0)
+sink()

### 2019-02-28 scripts/example_4.R

@@ -48,3 +48,39 @@ pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
   theme_wsj() +
   ggsave("/home/richel/GitHubs/pirouette_article/figure_example_4.png")
+
+
+
+
+
+testit::assert(pir_params$experiments[[1]]$inference_model$mcmc$store_every != -1)
+esses <- tracerer::calc_esses(
+  traces = tracerer::parse_beast_log(pir_params$experiments[[1]]$beast2_options$output_log_filename),
+  sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
+)
+sink("/home/richel/GitHubs/pirouette_article/example_4_esses.latex")
+xtable::xtable(esses, caption = "ESSes of example 4", label = "tab:esses_example_4", digits = 0)
+sink()
+
+
+df_evidences <- utils::read.csv(pir_params$evidence_filename)[, c(-1, -6)]
+df_evidences$site_model_name <- plyr::revalue(df_evidences$site_model_name, c("JC69" = "JC", "TN93" = "TN"))
+df_evidences$clock_model_name <- plyr::revalue(
+  df_evidences$clock_model_name,
+  c("strict" = "Strict", "relaxed_log_normal" = "RLN")
+)
+df_evidences$tree_prior_name <- plyr::revalue(
+  df_evidences$tree_prior_name,
+  c("yule" = "Yule", "birth_death" = "BD", "coalescent_constant_population" = "CCP", "coalescent_exp_population" = "CEP")
+)
+names(df_evidences) <- c("Site model", "Clock model", "Tree prior", "log(evidence)", "Weight")
+
+
+sink("/home/richel/GitHubs/pirouette_article/example_4_evidences.latex")
+xtable::xtable(
+  df_evidences,
+  caption = "Evidences of example 4", label = "tab:evidences_example_4", digits = 3
+)
+sink()
+
+

### 2019-02-28 article.tex
Updates from Overleaf

@@
 Many of these have been developed, 
-in the form of likelihood diversification models: 
+in the form of likelihood diversification models \giovanni{Not of them are already available for BEAST2 yet, or maybe they will never be}: 
 birth-death models that have rates that are 
@@
 diversity-dependent [\cite{etienne2011diversity}],
-or can shift [the SLS model. Laudanno et al., in preparation].
-There are birth death models in which speciation is a process that 
-takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}],
-can co-occur [Laudanno et al., in preparation],
+or can shift [\cite{etienne2012conceptual}, the SLS model. Laudanno et al., in preparation].
+There are birth death models in which speciation is a process that takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}], can co-occur [Laudanno et al., in preparation],
 dependent on a trait that is binary [\cite{maddison2007estimating}], 
@@
 or has multiple observed and concealed states [\cite{herrera2018detecting}].
-
-\iffalse
-Such models usually rely on the assumption 
-that a particular diversification process 
-can be explained mainly by a specific mechanism.
-Such a mechanism is built into a likelihood function, 
-which provides the probability that a phylogeny is realized 
-from a given set of parameters.
-Starting from a phylogeny, this likelihood is maximized
-to infer the best estimates for the parameter values.
-For such diversification models, it is a standard practice 
-to measure how well parameters can be recovered in this way,
-by simulating 'true' phylogenies from known parameters, then
-measuring the parameters deduced by maximum likelihood. 
-\fi
 
@@
 If the likelihood proves to be effective to describe the process, 
-it can be implemented into BEAST2 as a new prior. 
-Despite that it is still possible that already existing BEAST2 priors 
+it can potentially be implemented into BEAST2 as a new prior. 
+Despite that, it is still possible that already existing BEAST2 priors 
 will perform well enough and, therefore, 
@@
     to create a BEAST2 posterior of phylogenies (3a),
-    which is compared to the true phylogeny (1a) using
+    which is compared to the true phylogeny (1a) \giovanni{this makes me think that we might need an arrow going from 1a to 4a, maybe (and i say maybe) the arrow could be in red to highlight that the "true" nltt is in red too.
+BTW, the red line is not supervisible.} using
     the nLTT statistic. (4a) shows the nLTT of the true tree in red,
@@
 The first step simulates a DNA alignment from the given 
-phylogeny (Fig.~\ref{fig:pipeline} 1a $\rightarrow$ 2a).
+phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
 One can specify a DNA sequence
@@
 
-The second step selects an inference model, using the alignment.
-The user can specify whether to use the generative model or select the best model
-from a set of inference models. 
-When selecting the generative model,
-the site and clock model used in the alignment simulation are used
-in the inference. Because the given phylogeny (on which the alignment is based)
+The second step selects an inference model, using the alignment \giovanni{Is it worth to add it in the figure, maybe on the side of the vertical pipeline?
+It could be something like an arrow moving horizontally to the box of inference model.
+Then the box could have another arrow pointing at the arrow connecting 2 and 3.
+If you think this is too complicated, and not easy to follow for the reader you can discard it.}.
+The user can specify whether to use the generative model or select the best model from a set of inference models. 
+When selecting the generative model, the site and clock model used in the alignment simulation are used in the inference. Because the given phylogeny (on which the alignment is based)
 may have followed any tree prior (i.e. speciation model), the user needs
@@
 The third step infers a Bayesian posterior from the simulated 
-alignment (Fig.~\ref{fig:pipeline} 2a $\rightarrow$ 3a),
+alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a),
 using the inference model(s) selected in the previous step. The user
@@
 are the Markov-Chain Monte Carlo (MCMC) setup, 
-an optional MRCA prior and an RNG seed.
+an optional MRCA prior and an RNG seed \giovanni{This has been already said when defining the inference model.
+I suggest to define it only once in detail (like here) and just mention it the first time (something like: "An inference model is a combination of site model, clock model, tree prior as well as additional BEAST2 settings." in the description section, around line 98), which will probably look a bit more neat, as it's a bit clearer the parallel between alignment model and inference model.}.
 The MCMC setup determines the number of posterior trees sampled.
@@
 The fourth step compares the true tree to each posterior tree
-error using
-the nLTT statistic (\cite{janzen2015approximate}) 
-(Fig.~\ref{fig:pipeline} 3a $\rightarrow$ 4a), resulting
-in an error distribution (Fig.~\ref{fig:pipeline} 4a $\rightarrow$ 5a).
+error using the nLTT statistic (\cite{janzen2015approximate}) 
+(Fig.~\ref{fig:pipeline}, 3a $\rightarrow$ 4a), resulting
+in an error distribution (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
 \richel{I would not mind to remove 4a and 4b from the figure}
-The nLTT statistic is used by default,
-but also a user-defined error statistic can be used. As an example,
-\verb;pirouette; suplies such a custom error statistic,
+The nLTT statistic is used by default, but also a user-defined error statistic can be used. As an example,
+\verb;pirouette; supplies one custom error statistic,
 that uses an absolute difference in the gamma statistic [\cite{pybus2000testing}].
@@
 
-An optional step is to generate a 'twin tree' (Fig.~\ref{fig:pipeline} 1a $\rightarrow$ 1b), 
+An optional step is to generate a 'twin tree' (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
 that will be analyzed in the same way as the true tree.
@@
 yet its branching times are simulated according to 
-the tree prior used by BEAST2 to produce a posterior.
+the tree prior used by BEAST2 to produce a posterior \giovanni{is it true? is it Yule or BD? Please check}.
 The goal of this procedure is to generate a tree 
 to use as a control for the main experiment.
-The procedure to generate a twin tree is the following:
+We produce the twin tree according to the following procedure.
 
@@
 An overview of \verb;pirouette;'s main functions is shown in 
-table \ref{tab:functions}, which are all demonstrated in the examples
-below.
+table \ref{tab:functions}. We will exploit these functions to answer our research questions in the next sections.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: first research question}
+All \verb;pirouette;'s functions are documented. Useful examples and sensible defaults, which we will refer to in the next sections, can be accessed from the documentation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Usage: First research question}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A first research question that \verb;pirouette; answers is:
+
 "What is the error made by BEAST2 from a phylogeny using the same diversification model as it was generated by?"
@@
 \end{lstlisting}
+
+\giovanni{do we really want to use a function called "create\_ideal\_tree"?}
 
@@
     when the generative and inference model are the same.
-    The x-axis specifies the tree type (which can be 'true' or 'twin'),
-    the y-axis shows the error density
+    The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error.
+\giovanni{This holds true for all the figures: if we want to add these numbers we need to specify what they represent in the caption.}
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: second research question}
+\section{Usage: Second research question}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A second research question that \verb;pirouette; answers, is:
-"What is the error BEAST2 makes from a phylogeny when
+
+"What is the error made by BEAST2 from a phylogeny when
 picking the best inference model?"
@@
 run only if it is the best candidate, the evidence is measured (ignoring
-this will give a helpful error) and we use the full arsenal of 
+this will give a helpful error) and we use the full set of 
 40 inference models, which are all combinations of 4 site 
@@
     picking the best inference model.
-    The x-axis specifies the tree type (which can be 'true' or 'twin'),
-    the y-axis shows the error density
+    The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error.
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: third research question}
+\section{Usage: Third research question}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A third research question that \verb;pirouette; answers is:
-"What is the error BEAST2 makes from a phylogeny, 
+
+"What is the error made by BEAST2 from a phylogeny, 
 when hand-picking an inference model, compared to the background noise?"
@@
 experiments (listing \ref{lst:create_generative_experiment}),
-BEAST2 inference and error measuring parameters.
-
-\iffalse
-This time, we are interested in creating a twin tree. A twin tree
-has the same topology as the given tree, yet its branch lengths follow
-a distribution as such that the BEAST2 error will be 
-close to \richel{'close to' is a bit weasly, but currently it is true}
-its minimal error.
-\fi
+BEAST2 inference and error measuring parameters \giovanni{can we add a reference to the "BEAST2 inference and error measuring parameters" we are using?}.
 
@@
 to the tree prior used by BEAST2 to produce a posterior. 
-Since there is coherence between the inference model and the generative model, 
-we expect BEAST2 to produce a posterior of phylogenies clearly 
-more similar to the original one. 
+Since there is coherence between the inference model and the generative model \giovanni{Do we need to specify what is a generative model?
+Are we using BD or Yule?
+In case we are using Yule are we sure there is actually coherence when inference model is bd?}, we expect BEAST2 to produce a posterior of phylogenies clearly more similar to the original one.
 We intend to use this as a control experiment.
@@
   \caption{
-    Example 3: the error BEAST2 makes from a phylogeny picking the best inference model versus the control.
+    Example 3: the error made by BEAST2 from a phylogeny, picking the best inference model versus the control.
     Here, the twin column shows the error BEAST2 makes starting from the original tree versus the twin. 
-    The x-axis specifies the tree type (which can be 'true' or 'twin'),
-    the y-axis shows the error density
+    The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error.
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: fourth research question}
+\section{Usage: Fourth research question}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A fourth research question that \verb;pirouette; answers, is:
-"What is the error BEAST2 makes from a phylogeny for a known
-generative model, compared to candidates with similar inference models,
-in relationship to the background error?"
-
-Here, same as example 2, we start with a tree generated by the Yule process,
-by using the code shown in listing \ref{lst:create_yule_tree}.
+
+"What is the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error?"
+\giovanni{Did we define already the "generative" model? If not we can maybe refer to it when we define alignment and inference model}
+
+Here, same as example 2, we start with a tree generated by the Yule process, by using the code shown in listing \ref{lst:create_yule_tree}.
 The first step in \verb;pirouette; is to simulate a DNA alignment 
-from the given phylogeny, for which we'll use the same code 
+from the given phylogeny, for which we will use the same code 
 as shown in listing \ref{lst:create_alignment}.
@@
 In the second step we state our experiments. 
-Within this example, we must both specify the generative and candidate 
-experiments. Here we'll specify the generative model:
+In this example, we must both specify the generative and candidate experiments. Here we specify the generative model:
 
@@
 
-Specifying all candidates can be done in a factorial fashion using 
-\verb;create_all_experiments;. In this case, all candidates will use the
-same site and clock model, and only differ in their tree priors:
+Specifying all candidates can be done simply by calling \verb;create_all_experiments;. In this case, all candidates will use the same site and clock model, and only differ in their tree priors:
 
@@
 
-Also here, the third (the BEAST2 inference) and fourth (measuring the error)
-steps have sensible defaults. To get an idea of a baseline error made by
-BEAST2, we also make use of the twinning option:
+We choose again to use sensible defaults for the third (the BEAST2 inference) and fourth (measuring the error) steps, as we previously did to answer the second research question.
+To get an idea of a baseline error made by BEAST2, we also make use of the twinning option:
 
@@
   \caption{
-    Example 4: the error BEAST2 makes from a phylogeny for a known
-    generative model, compared to candidates with similar inference models,
-    in relationship to the background error.
-    The x-axis specifies the tree type,
-    the y-axis shows the error density
+    Example 4: the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error.
+    The x-axis specifies the tree type, the y-axis shows the error.
   }
@@
   is shown. Experiment 4 is split in two rows
-  by either the candidate ('C') or generative model ('G') 
+  by either the candidate ('C') or generative model ('G'). \giovanni{How do you define range? Are we using the standard error? \url{https://en.wikipedia.org/wiki/Standard_error}}
 }
@@
 \richel{check numbers}
+\giovanni{I think that the current version of the discussion is too didascalic.
+Most of what is said can be effectively summarized by table 5.
+We can even add ESSes in it.
+Another point is that we should understand what is the message we want to deliver in this discussion.
+Since we are only using one tree we are not in the position to claim anything, because a single tree is not an effective sample.
+I saw that in some occasions you refer to the RQ as examples.
+This might be the right way to proceed, as our purpose is just to illustrate how to run all the possible experiments on a full tree distribution.
+I think we should make it clearer throughout the entire manuscript.
+I am also interested in hearing Rampal's opinion on this.}
 
@@
 The range of errors goes from \richel{?} to \richel{?}.
-As we know the generative model, 
-this error distribution is the minimum noise that BEAST2 produces 
-for that generative model.
+As we know the generative model, this error distribution is the minimum noise that BEAST2 produces for that generative model.
 We can measure the effective samples as well (not shown), 
-and all ESSes are above
-the recommended value of 200 \cite{drummond2015bayesian}, 
+and all the Estimated Sample Sizes (ESSes) are above
+the recommended value of 200 [\cite{drummond2015bayesian}], 
 as shown in table \ref{tab:esses_example_1}.
 
-Figure \ref{fig:example_2} shows that the best candidate model gives an error distribution 
-with a median of \richel{?}. 
+Figure \ref{fig:example_2} shows that the best candidate model gives an error distribution with a median of \richel{?}. 
 The range of errors goes from \richel{?} to \richel{?}. 
-The inference model with the highest evidence follows a JC site model, 
-RLN clock model and BD tree prior, with a weight of $0.675$. 
-This is (unexpectedly) better than the generative
+The inference model with the highest evidence follows a JC site model, RLN clock model and BD tree prior, with a weight of $0.675$. 
+This is (unexpectedly) \giovanni{We are not using distributions, so we are totally subject to stochasticity. Nothing can be expected from just one tree.} better than the generative
 model with a weight of $0.166$. 
@@
 This means that it is rational to use this inference model to 
-infer a posterior phylogeny distribution, 
-as is done by \verb;pirouette;.
+infer a posterior phylogeny distribution, as is done by \verb;pirouette;.
 Also here, the ESSes are well above 200, 
@@
 added to BEAST2.
-We can draw this conclusion with confidence, as the ESSes are well above 200, 
+We can draw this conclusion with confidence \giovanni{no, we cannot: one tree is not enough}, as the ESSes are well above 200, 
 as shown in table \ref{tab:esses_example_4}.

### 2019-02-28 article.tex
First draft ready for feedback by @rsetienne

@@
     The pipeline starts from a true phylogeny (1a), 
-    that is converted to an alignment (2a), which is used
-    to create a BEAST2 posterior of phylogenies (3a),
-    which is compared to the true phylogeny (1a) \giovanni{this makes me think that we might need an arrow going from 1a to 4a, maybe (and i say maybe) the arrow could be in red to highlight that the "true" nltt is in red too.
-BTW, the red line is not supervisible.} using
-    the nLTT statistic. (4a) shows the nLTT of the true tree in red,
+    that is converted to an alignment (2a).
+    The experiments are filtered (3a), possibly based on the
+    alignment (2a), to determine which experiments are actually run.
+    The experiment that are actually performed, use the 
+    alignment (2a) to create a BEAST2 posterior of phylogenies (4a),
+    which is compared to the true phylogeny (1a) 
+    using the nLTT statistic. (5a) shows the nLTT of the true tree in red,
     the nLTT of all individual posterior trees in grey, and the average
@@
     nLTT results in an error distribution, here displayed as a histogram
-    of error values (5a).
+    of error values (6a).
     Optionally, a twin phylogeny (1b) can be generated from a true
-    phylogeny (1a), which then results in an error distribution (5b) with
+    phylogeny (1a), which then results in an error distribution (6b) with
     the similar intermediate steps 
@@
 
-The second step selects an inference model, using the alignment \giovanni{Is it worth to add it in the figure, maybe on the side of the vertical pipeline?
-It could be something like an arrow moving horizontally to the box of inference model.
-Then the box could have another arrow pointing at the arrow connecting 2 and 3.
-If you think this is too complicated, and not easy to follow for the reader you can discard it.}.
+The second step selects experiments to be actually run, 
+based on the experiments and the alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a) 
 The user can specify whether to use the generative model or select the best model from a set of inference models. 
@@
 The third step infers a Bayesian posterior from the simulated 
-alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a),
+alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a $\rightarrow$ 4a),
 using the inference model(s) selected in the previous step. The user
@@
 error using the nLTT statistic (\cite{janzen2015approximate}) 
-(Fig.~\ref{fig:pipeline}, 3a $\rightarrow$ 4a), resulting
-in an error distribution (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
-\richel{I would not mind to remove 4a and 4b from the figure}
+(Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a), resulting
+in an error distribution (Fig.~\ref{fig:pipeline}, 5a $\rightarrow$ 6a).
 The nLTT statistic is used by default, but also a user-defined error statistic can be used. As an example,
@@
 yet its branching times are simulated according to 
-the tree prior used by BEAST2 to produce a posterior \giovanni{is it true? is it Yule or BD? Please check}.
+the tree prior used by BEAST2 to produce a posterior 
+\giovanni{is it true? is it Yule or BD? Please check}
+\richel{It is Yule}.
 The goal of this procedure is to generate a tree 
@@
   ]
-phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
-\end{lstlisting}
-
-\giovanni{do we really want to use a function called "create\_ideal\_tree"?}
+phylogeny <- create_yule_tree(n_taxa = 5, crown_age = 10)
+\end{lstlisting}
 
@@
     when the generative and inference model are the same.
-    The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error.
-\giovanni{This holds true for all the figures: if we want to add these numbers we need to specify what they represent in the caption.}
+    The x-axis specifies the tree type (which can be 'true' or 'twin'), 
+    the y-axis shows the error. 
+    The boxplot shows the minimum, first quartile, median, third 
+    quartile and maximum error, of which the values are displayed 
+    in the adjacent label
   }
@@
     The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error.
+    The boxplot shows the minimum, first quartile, median, third 
+    quartile and maximum error, of which the values are displayed 
+    in the adjacent label
   }
@@
     The x-axis specifies the tree type, the y-axis shows the error.
+    The boxplot shows the minimum, first quartile, median, third 
+    quartile and maximum error, of which the values are displayed 
+    in the adjacent label
   }
@@
   For the true and twin tree,
-  the median and range of the errors 
+  the median and range (from minimum to maximum value) of the errors 
   is shown. Experiment 4 is split in two rows
-  by either the candidate ('C') or generative model ('G'). \giovanni{How do you define range? Are we using the standard error? \url{https://en.wikipedia.org/wiki/Standard_error}}
+  by either the candidate ('C') or generative model ('G')
 }

### 2019-03-04 article.tex
Changes @rsetienne

@@
 \usepackage{authblk}
-\title{pirouette: the error BEAST2 makes in inferring a phylogeny}
+\title{pirouette: Determining the error BEAST2 makes in inferring a phylogeny}
 \author[1]{Rich\`el J.C. Bilderbeek}
@@
 \newcommand{\richel}[1]{\textcolor{orange}{\textbf{[RB: #1]}}}
+\newcommand{\rampal}[1]{\textcolor{green}{\textbf{[RSE: #1]}}}
 
@@
     BEAST2 is a popular Bayesian phylogenetics software tool,
-    that takes an alignment and an inference model to create a
+    that takes an alignment of characters (usually nucleotides) and an inference model to create a
     posterior of jointly-estimated phylogenies and model parameter estimates.
-    When a new macro-evolutionary diversification model is developed,
-    a good first step is to measure the error BEAST2 makes, with its current set of inference models, on a known
-    phylogeny derived from a new diversification mechanism. \\
+An important ingredient of BEAST2 is the model for the species tree prior, and the tool can in principle be used for inference of macroevolutionary processes underlying the species tree.
+However, only models that allow for fast computation of the prior probability are possible in practice.
+This begs the question how accurate the tree estimation is when the real macroevolutionary processes are different from those assumed in BEAST2.
   \textbf{2. }
@@
     the inference error BEAST2 makes based on a known/true 
-    phylogeny. \\
+    phylogeny, for example generated by a new macroevolutionary diversification model. \\
   \textbf{3. }
@@
   \textbf{4. }
-    Last, we will discuss the results obtained by the examples. \\
+    Last, we discuss the results obtained by the examples. \\
 \end{abstract}
@@
 allows to build phylogenetic trees 
-from genetic material extracted from extant organisms.
+from character data (usually nucleotide sequences) extracted from extant organisms.
 This has constituted an important step forward 
-in our understanding on how species evolve.
+in our understanding of how species evolve.
 Such tools have been increasingly exploited to hypothesize 
@@
 
-Since BEAST performs a Bayesian analysis, 
-it needs not only genetic data but also tree priors, 
-describing how the diversification occur, to return posterior phylogenies.
-BEAST works using standard tree priors like Yule or birth-death 
+Because BEAST performs a Bayesian analysis, 
+it does not only need genetic data but also tree priors, 
+describing how the diversification occurs, to return posterior phylogenies.
+BEAST uses standard tree priors such as Yule or constant-rate birth-death models
 \giovanni{integrate this if necessary}.
@@
 unlike its predecessor, the possibility for third-party users 
-to integrate its prior libraries with new ones. 
-Many of these have been developed, 
-in the form of likelihood diversification models \giovanni{Not of them are already available for BEAST2 yet, or maybe they will never be}: 
+to use their own preferred diversification model if an algorithm is provided to calculate the probability of a tree given this model. 
+Many such models and their associated probability algorithms have been developed, 
+ \giovanni{Not of them are already available for BEAST2 yet, or maybe they will never be}: 
 birth-death models that have rates that are 
@@
 time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
-diversity-dependent [\cite{etienne2011diversity}],
-or can shift [\cite{etienne2012conceptual}, the SLS model. Laudanno et al., in preparation].
-There are birth death models in which speciation is a process that takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}], can co-occur [Laudanno et al., in preparation],
-dependent on a trait that is binary [\cite{maddison2007estimating}], 
-has multiple states [\cite{fitzjohn2012diversitree}],
-has a concealed state [\cite{beaulieu2016detecting}] 
-or has multiple observed and concealed states [\cite{herrera2018detecting}].
+or diversity-dependent [\cite{etienne2011diversity}],
+or that can shift [\cite{etienne2012conceptual}, \cite{rabosky2014automatic}, \cite{alfaro2009nine}, the SLS model. Laudanno et al., in preparation].
+Other birth-death models treat speciation as a process that takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}], or that happens in bursts of simultaneous events [Laudanno et al., in preparation], or that
+depends on a trait that has two [\cite{maddison2007estimating}], 
+or more [\cite{fitzjohn2012diversitree}] states,
+or even allows concealed states [\cite{beaulieu2016detecting}] 
+or a combination of all these [\cite{herrera2018detecting}].
 
-From a given phylogeny, the likelihood is maximized
-to infer the best estimates for the parameter values.
+For a given phylogeny the parameter values of the diversification model are usually estimated using maximum likelihood.
 For such  models, it is a standard practice to assess their performance 
-on a controlled environment: this is done by simulating phylogenies from known parameters, 
-then comparing the estimations obtained by maximum likelihood with the original parameters.
+on a controlled environment: phylogenies are simulated with known parameters, and then the estimated parameters are compared with the original parameters.
 If the likelihood proves to be effective to describe the process, 
 it can potentially be implemented into BEAST2 as a new prior.
-Despite that, it is still possible that already existing BEAST2 priors 
-will perform well enough and, therefore, 
-that the introduction of a new model is not necessary.
-
-Diversification models are getting more complex over time. 
-This is facilitated by the increase in computing power due to the advancement of technology.
-Increased complexity made possible the exploration of new models 
-encompassing mechanisms that are not yet considered by previous diversification models. 
-An example of this is the multiple-birth model [Laudanno et al., in preparation],
-that is the first model to allow simultaneous speciations to co-occur.
-
-However, whenever a new model is proposed there is always a trade-off 
-between the cost of increased complexity versus the benefits that it can provide.
-In fact developing models can cost time, human and computational resources.
-Therefore desirable models have to be able to disentangle 
-the intrinsic complexity of a system, 
-highlight the main reasons of its behaviour and avoid unnecessary complexity.
-
-For this reason we believe it is important to have techniques 
-to quantitatively assess the advancement provided by the introduction of new models, 
-in comparison to the existing ones.
-
-To achieve this, one possible experiment is to simulate a phylogeny 
+However, in many cases the computation of this likelihood (the tree prior) is computationally demanding and it is unfeasible to use it in a Bayesian inference framework where it needs to be computed millions to billions of times.
+However, if the data are very informative, the influence of the prior is often relatively small, and hence under this assumption it has become common practice to use already existing BEAST2 priors.
+However, it is unclear whether this assumption holds.
+Here we introduce a pipeline to assess this assumption for any given diversification model. Briefly, it consists of the following steps: first simulate a phylogeny 
 according to the mechanisms proposed by the new diversification model, 
-then simulate its alignment and measure how well current inference models perform 
-at retrieving the original phylogeny.
-If this difference is small enough, current models apparently
-suffice.
-
-\verb;pirouette; is an R package that performs such an experiment,
-and is built on \verb;babette; [Bilderbeek \& Etienne, 2018], 
+then simulate a nucleotide alignment and use current inference tools (such as BEAST2) to estimate the phylogeny.
+If the estimated tree differs only slightly from the original phylogeny, the original assumption is justified and the simple diversification models in the inference tools suffice.
+If the trees differ markedly, it will be worth the effort and computational burden to develop a species tree module to be used in these tools.
+Our pipeline is programmed as an R package and is called 
+\verb;pirouette; it is built on \verb;babette; [Bilderbeek \& Etienne, 2018], 
 which calls BEAST2 [\cite{bouckaert2014beast}]. 
@@
 any given phylogeny, helping us to evaluate the necessity of a new diversification model.
-Once verified that the introduction of a new model is indeed beneficial,
-it can then be added to BEAST2's existing set of (birth-death and 
-coalescent) models via the BEAST2 package system.
 
@@
 
-\iffalse
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes from a given reconstructed phylogeny, within the context of
+makes for a given reconstructed phylogeny, within the context of
 a simulation study on a (possibly new) diversification model. Such
-a study will typically construct phylogenies for different combinations of the diversification model's parameters, to assess under which scenarios the error made by BEAST2 cannot be ignored anymore. 
-Such a study needs to be done on, there need be many replicates.
-\fi
+a study will typically construct phylogenies for different combinations of the diversification model's parameters, to assess under which scenarios the error made by BEAST2 cannot be neglected. 
+This usually requires many replicates.
 
-The goal of \verb;pirouette; is to measure the inference error made by BEAST2 from a given reconstructed phylogeny, which has been simulated according to a (possibly new) diversification model.
-To do that the full procedure would be to simulate a great number of phylogenies for different combinations of the diversification model's parameters, to assess under which scenarios the error made by BEAST2 cannot be ignored anymore.
-
-To be able to explain how \verb;pirouette; measures the inference error
+To explain how \verb;pirouette; measures the inference error
 BEAST2 makes, we first give two definitions of what
@@
 a generative model is a combination of site model, clock model and tree prior,
-used to simulate alignments. An inference model is a combination of 
+used to simulate DNA alignments. An inference model is a combination of 
 site model, clock model, tree prior, an optional node calibration prior
 and setup of the Markov chain Monte Carlo (MCMC) algorithm, 
-used to produce a posterior. Node calibration priors, called 'most
+used to produce a posterior distribution of trees and model parameters. Node calibration priors, called 'most
 recent common ancestor' (MRCA) priors within BEAST2, allow to create a time-calibrated phylogeny, by specifying a normal distribution around an expected time when the ancestors of two or more taxa diversified.
@@
     The pipeline starts from a true phylogeny (1a), 
-    that is converted to an alignment (2a).
+    that is converted to an alignment (2a). \rampal{this is unclear; the reader does not know what an experiment is, why there are more, what filtering means, etc}
     The experiments are filtered (3a), possibly based on the
@@
     which is compared to the true phylogeny (1a) 
-    using the nLTT statistic. (5a) shows the nLTT of the true tree in red,
+    using the nLTT statistic [\cite{janzen2015approximate}]. (5a) shows the nLTT of the true tree in red,
     the nLTT of all individual posterior trees in grey, and the average
@@
     phylogeny (1a), which then results in an error distribution (6b) with
-    the similar intermediate steps 
+    similar intermediate steps 
   }
@@
 \begin{enumerate}
-    \item from a given phylogeny an alignment is simulated;
-    \item from the alignment an inference model is chosen;
-    \item the alignment is used as BEAST2 input to infer a posterior distribution of phylogenies;
-    \item posterior phylogenies are compared with the given original phylogeny to estimate the error we make. The higher the match, the lower the error;
+    \item for a given phylogeny an alignment is simulated for a known 'alignment' model;
+    \item for this alignment, model selection is performed to choose the best-fitting alignment model in inference (which may be different from the generative alignment model);
+    \item the alignment is then used as BEAST2 input to infer a posterior distribution of phylogenies;
+    \item posterior phylogenies are compared with the given original phylogeny to estimate the error made;
 \end{enumerate}
@@
 There is also the option to generate a 'twin tree',
-that goes through the same pipeline.
+that goes through the same pipeline. The utility of this twin tree will be explained below.
 
-The first step simulates a DNA alignment from the given 
+The first step simulates a DNA alignment from a given 
 phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
@@
 
-The second step selects experiments to be actually run, 
-based on the experiments and the alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a) 
-The user can specify whether to use the generative model or select the best model from a set of inference models. 
-When selecting the generative model, the site and clock model used in the alignment simulation are used in the inference. Because the given phylogeny (on which the alignment is based)
+In the second step selects (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a), the user can specify whether to use the generative model as the inference model or to select the best model from a set of inference models.
+When selecting the generative model, the site and clock model used in the alignment simulation are used in the inference. Furthermore, because the given phylogeny (on which the alignment is based)
 may have followed any tree prior (i.e. speciation model), the user needs
@@
 Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; 
-inference model selection is not available on Windows.
+inference model selection is not available on Windows.\rampal{clarify this. One can use mcbette on a browser, right?}
 
-The third step infers a Bayesian posterior from the simulated 
+The third step infers the actual Bayesian posterior from the simulated 
 alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a $\rightarrow$ 4a),
@@
 \richel{It is Yule}.
+\rampal{I think it's better to use BD for more generality. Yule is too limited in the trees it can produce.}
 The goal of this procedure is to generate a tree 
@@
 We use $\lambda$ and $\mu$ to simulate a new birth-death tree, 
-conditioned on having the same number of tips as the original tree. 
-We store its branching times. 
-We then combine such branching times with the original tree's 
+conditioned on having the same number of tips as the original tree, we store its branching times, and then  combine the branching times with the original tree's 
 topology to obtain the twin tree.
@@
 \verb;pirouette; can be installed easily from CRAN:
-
+\rampal{be careful here; it is not on CRAN yet. Better to say it can be installed from GitHub and will be made available on CRAN)}
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]

### 2019-03-04 article.tex
Process feedback rsetienne

@@
     The pipeline starts from a true phylogeny (1a), 
-    that is converted to an alignment (2a). \rampal{this is unclear; the reader does not know what an experiment is, why there are more, what filtering means, etc}
-    The experiments are filtered (3a), possibly based on the
-    alignment (2a), to determine which experiments are actually run.
+    that is converted to an alignment (2a).
+    \rampal{this is unclear; the reader does not know what an experiment is, why there are more, what filtering means, etc}
+    The use supplies one or more experiments, which are a combination of 
+    an inference model and the conditions in which it is actually used in
+    inference.
+    It is determined which experiments will be actually run (3a), 
+    possibly based on the
+    alignment (2a).
     The experiment that are actually performed, use the 
@@
 performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 package. 
-Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac; 
-inference model selection is not available on Windows.\rampal{clarify this. One can use mcbette on a browser, right?}
+Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac,
+therefore inference model selection is not available on Windows. Note that
+there is the option to do model selection from the browser, 
+using \verb;mcbette;. The computation for this model selection is 
+performed on the servers of the Travis CI continuous integration service,
+that restricts computation time to one hour for free users.
 
@@
 the tree prior used by BEAST2 to produce a posterior 
-\giovanni{is it true? is it Yule or BD? Please check}
-\richel{It is Yule}.
-\rampal{I think it's better to use BD for more generality. Yule is too limited in the trees it can produce.}
+\richel{Should be switched to BD, pirouette Issue 161}.
 The goal of this procedure is to generate a tree 
@@
 \rampal{be careful here; it is not on CRAN yet. Better to say it can be installed from GitHub and will be made available on CRAN)}
+\richel{We adopted this policy (to say it's on CRAN in advance) with babette as well. Is this a hint to get pirouette on CRAN before the publication of this article :-) ?}
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]

### 2019-03-04 article.tex
Processed feedback, thanks @rsetienne

@@
     BEAST2 is a popular Bayesian phylogenetics software tool,
-    that takes an alignment of characters (usually nucleotides) and an inference model to create a
-    posterior of jointly-estimated phylogenies and model parameter estimates.
-An important ingredient of BEAST2 is the model for the species tree prior, and the tool can in principle be used for inference of macroevolutionary processes underlying the species tree.
-However, only models that allow for fast computation of the prior probability are possible in practice.
-This begs the question how accurate the tree estimation is when the real macroevolutionary processes are different from those assumed in BEAST2.
+    that takes an alignment of characters (usually nucleotides) 
+    and an inference model to create a
+    posterior of jointly-estimated phylogenies and model parameter 
+    estimates. An important ingredient of BEAST2 is 
+    the model for the species tree prior, 
+    and the tool can in principle be used for inference of 
+    macroevolutionary processes underlying the species tree. 
+    However, only models that allow for fast computation of 
+    the prior probability are possible in practice. 
+    This begs the question how accurate the tree estimation is 
+    when the real macroevolutionary processes are different 
+    from those assumed in BEAST2. \\
   \textbf{2. }
@@
     the inference error BEAST2 makes based on a known/true 
-    phylogeny, for example generated by a new macroevolutionary diversification model. \\
+    phylogeny, for example generated by a new 
+    macroevolutionary diversification model. \\
   \textbf{3. }
@@
 allows to build phylogenetic trees 
-from character data (usually nucleotide sequences) extracted from extant organisms.
+from character data (usually nucleotide sequences) extracted 
+from extant (but also extinct) organisms.
 This has constituted an important step forward 
@@
 Because BEAST performs a Bayesian analysis, 
-it does not only need genetic data but also tree priors, 
+it does not only need character data but also tree priors, 
 describing how the diversification occurs, to return posterior phylogenies.
-BEAST uses standard tree priors such as Yule or constant-rate birth-death models
-\giovanni{integrate this if necessary}.
+BEAST uses standard tree priors such as Yule or 
+constant-rate birth-death [\cite{nee1994reconstructed}] models.
+These models are most commonly picked in biological research,
+as these are of a desired minimum biological complexity, 
+yet computationally fast enough.
 The recent development of BEAST2 [\cite{bouckaert2014beast}] provides, 
 unlike its predecessor, the possibility for third-party users 
-to use their own preferred diversification model if an algorithm is provided to calculate the probability of a tree given this model. 
-Many such models and their associated probability algorithms have been developed, 
- \giovanni{Not of them are already available for BEAST2 yet, or maybe they will never be}: 
-birth-death models that have rates that are 
-constant [\cite{nee1994reconstructed}], 
+to use their own new or preferred diversification model, 
+if an algorithm is provided to calculate the probability 
+of a tree given such a diversification model.
+There have been plenty of these models (and their associated probability algorithms) 
+developed, among others, those in which diversification is 
 time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
@@
 or a combination of all these [\cite{herrera2018detecting}].
-
-For a given phylogeny the parameter values of the diversification model are usually estimated using maximum likelihood.
-For such  models, it is a standard practice to assess their performance 
-on a controlled environment: phylogenies are simulated with known parameters, and then the estimated parameters are compared with the original parameters.
+Note all of these diversification models are available in BEAST2, 
+but the option is available.
+
+For a given phylogeny, the parameter values of the diversification model 
+are usually estimated using maximum likelihood.
+For such  models, 
+it is a standard practice to assess their performance 
+on a controlled environment: 
+phylogenies are simulated with known parameters, 
+and then the estimated parameters are compared with the original parameters.
 If the likelihood proves to be effective to describe the process, 
 it can potentially be implemented into BEAST2 as a new prior. 
-However, in many cases the computation of this likelihood (the tree prior) is computationally demanding and it is unfeasible to use it in a Bayesian inference framework where it needs to be computed millions to billions of times.
-However, if the data are very informative, the influence of the prior is often relatively small, and hence under this assumption it has become common practice to use already existing BEAST2 priors.
+However, in many cases the computation of this likelihood (the tree prior) 
+is computationally demanding and it is unfeasible to use it 
+in a Bayesian inference framework where it needs to be 
+computed millions to billions of times.
+However, if the data are very informative, the influence of the prior 
+is often relatively small, and hence under this assumption 
+it has become common practice to use already existing BEAST2 priors. 
 However, it is unclear whether this assumption holds.
-Here we introduce a pipeline to assess this assumption for any given diversification model. Briefly, it consists of the following steps: first simulate a phylogeny 
+Here we introduce a pipeline to assess this assumption 
+for any given diversification model. 
+Briefly, it consists of the following steps: 
+first simulate a phylogeny 
 according to the mechanisms proposed by the new diversification model, 
-then simulate a nucleotide alignment and use current inference tools (such as BEAST2) to estimate the phylogeny.
-If the estimated tree differs only slightly from the original phylogeny, the original assumption is justified and the simple diversification models in the inference tools suffice.
-If the trees differ markedly, it will be worth the effort and computational burden to develop a species tree module to be used in these tools.
+then simulate a nucleotide alignment 
+and use current inference tools (such as BEAST2) to estimate the phylogeny. 
+If the estimated tree differs only slightly from the original phylogeny, 
+the original assumption is justified and the simple diversification models 
+in the inference tools suffice. 
+If the trees differ markedly, it will be worth the effort and computational burden 
+to develop a species tree module to be used in these tools.
 Our pipeline is programmed as an R package and is called 

### 2019-03-04 article.tex
Improved 'Introduction' and 'Description' sections

@@
 
+\begin{table}
+  \begin{tabular}{ | l | l | }
+    \hline
+    \textbf{Term} & \textbf{Definition} \\ 
+    \hline
+    Clock model & Model for the DNA mutation rates \\
+    Experiment & Combination of an inference model and the conditions to actually use it in inference \\
+    Generative model & Combination of site model, clock model and tree prior, used to simulate a DNA alignment \\
+    Inference model & Combination of site model, clock model, tree prior, MRCA prior and MCMC, used to infer a BEAST2 posterior \\
+    MCMC & Setup of the Markov Chain Monte Carlo algorithm \\
+    MRCA prior & Prior knowledge of a most recent common ancestor \\
+    Site model & Model for the DNA nucleotide substitions \\
+    Tree prior & Model for the macroevolutionary speciation process \\
+    \hline
+  \end{tabular}
+  \caption{
+    Definitions of the terms used in this article
+  }
+  \label{tab:definitions}
+\end{table}
+
 \verb;pirouette; is written in the R programming language (\cite{R}).
-
 The goal of \verb;pirouette; is to measure the inference error BEAST2
@@
 a simulation study on a (possibly new) diversification model. Such
-a study will typically construct phylogenies for different combinations of the diversification model's parameters, to assess under which scenarios the error made by BEAST2 cannot be neglected. 
-This usually requires many replicates.
-
+a study will typically construct phylogenies for different 
+combinations of the diversification model's parameters, 
+to assess under which scenarios the error made by BEAST2 cannot be neglected.
+Although a well-conducted theoretical study uses many replicates,
+this article only shows examples with one replicate, 
+for the sake of readability.
+
+\richel{
+  I do think this is a clumsy paragraph. I hope you can merge it
+  in the text
+}
 To explain how \verb;pirouette; measures the inference error
-BEAST2 makes, we first give two definitions of what
-we mean by a generative model and by an inference model: 
-a generative model is a combination of site model, clock model and tree prior,
-used to simulate DNA alignments. An inference model is a combination of 
-site model, clock model, tree prior, an optional node calibration prior
-and setup of the Markov chain Monte Carlo (MCMC) algorithm, 
-used to produce a posterior distribution of trees and model parameters. Node calibration priors, called 'most
-recent common ancestor' (MRCA) priors within BEAST2, allow to create a time-calibrated phylogeny, by specifying a normal distribution around an expected time when the ancestors of two or more taxa diversified.
+BEAST2 makes, we first give two 
+definitions (see table \ref{tab:definitions} for an overview of definitions):
+(1) a generative model is a combination of a known tree prior (that generated
+a given phylogeny) and a site and clock model (that generate an alignment 
+from the given tree), (2) an inference model is a combination of 
+site model, clock model, tree prior, an optional most recent 
+common-ancestor (MRCA) prior (to create a time-calibrated phylogeny)
+and a Markov chain Monte Carlo (MCMC) setup.
+A generative model is a known/true model underlying a phylogeny 
+and is used to simulate DNA alignments.
+An inference model
+is used to produce a posterior distribution of trees and model parameters.
 
@@
     that is converted to an alignment (2a).
-    \rampal{this is unclear; the reader does not know what an experiment is, why there are more, what filtering means, etc}
-    The use supplies one or more experiments, which are a combination of 
+    The use supplies one or more experiments, in which an experiment is
+    defined as a combination of 
     an inference model and the conditions in which it is actually used in
     inference.
-    It is determined which experiments will be actually run (3a), 
-    possibly based on the
-    alignment (2a).
-    The experiment that are actually performed, use the 
-    alignment (2a) to create a BEAST2 posterior of phylogenies (4a),
-    which is compared to the true phylogeny (1a) 
-    using the nLTT statistic [\cite{janzen2015approximate}]. (5a) shows the nLTT of the true tree in red,
+    Based on the experiments and possibly the alignment (2a),
+    experiments are selected to be actually used used in the next step (3a).
+    Those experiments that are selected, use the   
+    alignment (2a) and their inference model 
+    to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a).
+    The posterior trees are compared to the true phylogeny (1a) 
+    using the nLTT statistic [\cite{janzen2015approximate}]. 
+    (5a) shows the nLTT of the true tree in red,
     the nLTT of all individual posterior trees in grey, and the average
@@
 There is also the option to generate a 'twin tree',
-that goes through the same pipeline. The utility of this twin tree will be explained below.
+that goes through the same pipeline. 
+The utility of this twin tree will be explained below.
 
@@
 phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
-One can specify a DNA sequence
-of any length at the root of the phylogeny, a DNA mutation rate, a
-site (i.e. nucleotide substitution) model, 
-a clock model, a random number generator (RNG) seed and a location
-where the alignment is saved to. This step is relatively fast, but longer
+One can specify, most importantly, a root DNA sequence, 
+a DNA mutation rate, a site model, and a clock model
+\richel{
+  as the DNA mutation rate is the same over all branches, 
+  only the strict clock model makes sense. I suggest to
+  remove the option to specify a clock model
+}.
+This step is relatively fast, but longer
 DNA alignments will noticeably slow down the inference step.
 
-In the second step selects (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a), the user can specify whether to use the generative model as the inference model or to select the best model from a set of inference models.
-When selecting the generative model, the site and clock model used in the alignment simulation are used in the inference. Furthermore, because the given phylogeny (on which the alignment is based)
-may have followed any tree prior (i.e. speciation model), the user needs
-to specify which tree prior is used in the inference. 
-When selecting for the best
-model, the alignment is used to find the inference model that has the
-highest evidence (i.e. marginal likelihood) from a set of candidate inference models.
+The second step (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a)
+starts from experiments as set up by the user. 
+We define an experiment as the combination of an inference model 
+and the conditions to actually use it in inference.
+There a two types of inference models, which are
+generative and candidate inference models.
+An experiment uses a generative model if the inference model is known.
+An experiment uses a candidate model if either the inference model is unknown,
+or there is an interest in seeing how well non-generative models compete with
+the generative model.
+Typically, a generative model and/or a best candidate model are used 
+in the actual inference.
+When using an experiment with a generative model,
+it must use the same site and clock model used in the alignment simulation,
+as well as the tree prior that underlied the creation of the given phylogeny. 
+When using experiments with candidate models,
+\verb;pirouette; selects the best candidate based
+on the evidence (i.e. marginal likelihood) for the model given the alignment.
 The evidence of an inference model is estimated using a nested sampling (NS)
@@
 Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac,
-therefore inference model selection is not available on Windows. Note that
-there is the option to do model selection from the browser, 
-using \verb;mcbette;. The computation for this model selection is 
-performed on the servers of the Travis CI continuous integration service,
-that restricts computation time to one hour for free users.
-
-The third step infers the actual Bayesian posterior from the simulated 
+therefore using the selection of candidate models is not available on Windows. 
+Note that there is the option to do just the model selection on Windows 
+from the browser using \verb;mcbette;, yet the computation for this model selection is
+restricted to one hour for free users of the host (Travis CI) continuous
+integration service.
+
+The third step infers the actual Bayesian posteriors from the simulated 
 alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a $\rightarrow$ 4a),
-using the inference model(s) selected in the previous step. The user
-can specify the additional parameters needed for the BEAST2 run, which
-are the Markov-Chain Monte Carlo (MCMC) setup, 
-an optional MRCA prior and an RNG seed \giovanni{This has been already said when defining the inference model.
-I suggest to define it only once in detail (like here) and just mention it the first time (something like: "An inference model is a combination of site model, clock model, tree prior as well as additional BEAST2 settings." in the description section, around line 98), which will probably look a bit more neat, as it's a bit clearer the parallel between alignment model and inference model.}.
-The MCMC setup determines the number of posterior trees sampled.
-An MRCA prior allows the inferred phylogeny to have a dated crown age.
+using the inference models from the experiments 
+selected in the previous step. Per selected experiment, 
+one BEAST2 posterior is inferred, 
+using the \verb;babette; [\cite{bilderbeek2018babette}] R package.
 
@@
 in an error distribution (Fig.~\ref{fig:pipeline}, 5a $\rightarrow$ 6a).
-The nLTT statistic is used by default, but also a user-defined error statistic can be used. As an example,
+The nLTT statistic is used by default, 
+but also a user-defined error statistic can be used. 
+As an example,
 \verb;pirouette; supplies one custom error statistic,
@@
 
-An optional step is to generate a 'twin tree' (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
+An optional step is to generate a 'twin tree' 
+(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
 that will be analyzed in the same way as the true tree.
@@
 yet its branching times are simulated according to 
-the tree prior used by BEAST2 to produce a posterior 
+a birth-death tree model.
 \richel{Should be switched to BD, pirouette Issue 161}.
 The goal of this procedure is to generate a tree
-to use as a control for the main experiment.
-We produce the twin tree according to the following procedure.
-
-From the given (true) phylogeny, we maximize a birth-death 
-likelihood [\cite{nee1994reconstructed}] to infer the standard 
-birth-death rates, namely speciation rate $\lambda$ and extinction rate $\mu$.
-We use $\lambda$ and $\mu$ to simulate a new birth-death tree, 
-conditioned on having the same number of tips as the original tree, we store its branching times, and then  combine the branching times with the original tree's 
-topology to obtain the twin tree.
+with a tree prior present in BEAST2, 
+that is used as a control for the main experiment.
+To produce such a birth-death tree, we maximize a birth-death 
+likelihood [\cite{nee1994reconstructed}], 
+conditioned on having the same number of tips as the original tree,
+to infer its parameters.
+We use these parameters,  
+which are a speciation rate and extinction rate,
+to simulate the branching times for the 
+twin tree, which we combinate with the original tree's 
+topology.
 
@@
 An overview of \verb;pirouette;'s main functions is shown in 
-table \ref{tab:functions}. We will exploit these functions to answer our research questions in the next sections.
+table \ref{tab:functions}. 
+We will exploit these functions to answer our research questions in the next sections.
 
@@
 
-All \verb;pirouette;'s functions are documented. Useful examples and sensible defaults, which we will refer to in the next sections, can be accessed from the documentation.
+All \verb;pirouette;'s functions are documented. 
+Useful examples and sensible defaults, 
+which we will refer to in the next sections, 
+can be accessed from the documentation.
 
@@
 
+\richel{Suggest: use BD tree (again)}
 We start with a Yule (pure-birth) tree with five taxa and a crown age of ten.
@@
 and a strict clock model. 
-A JC site model assumes that mutation rates from any nucleotide to any other are equal and constant.
+A JC site model assumes that mutation rates from any nucleotide to any other 
+are equal and constant. 
 A strict clock model assumes that the mutation rates of all species are equal and constant.
@@
 
-\iffalse
 In the second step we state our experiments. In this context, we
@@
 Within this example, we specify that our experiment uses the
-generative model, will always be run and we ignore the evidence for our model,
-for the inference model that is also the generative model (using a
-JC site model, strict clock model and Yule tree prior).
-\fi
-
-In the second step we state our experiments. In this context, we
-define an experiment as a combination of an inference model
-and the conditions under which a Bayesian inference is executed.
-Within this example, we specify that our experiment uses the
-generative model (which means that the inference model will be a combination of JC site model, strict clock model and Yule tree prior), will always be run and we don't need to measure the evidence for our model.
+generative model (which means that the inference model will be 
+a combination of JC site model, strict clock model and Yule tree prior), 
+will always be run and we skip to measure the evidence for our model.
 
@@
 
-\verb;pirouette; has a nice plotting function:
+\verb;pirouette; has a plotting function used for convenience:
 
@@
 The first step in \verb;pirouette; is to simulate a DNA alignment from the given phylogeny. 
-We will re-use the alignment parameters of the previous example as shown in listing \ref{lst:create_alignment}.
+We will re-use the alignment parameters of the previous example 
+as shown in listing \ref{lst:create_alignment}.
 

### 2019-03-05 article.tex
Changes @Giappo

@@
 \usepackage{multirow}
+\usepackage{amsfonts}
+\usepackage{array}
+\usepackage{array}
+\usepackage{booktabs}
+\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
+\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
+\usepackage{longtable}
 
@@
 
+\verb;pirouette; is written in the R programming language (\cite{R}).
+The goal of \verb;pirouette; is to measure the inference error BEAST2
+makes for a given reconstructed phylogeny, simulated under the mechanisms of a (possibly new) diversification model, which we will refer to as the 'generative prior'.
+Such a study will typically construct phylogenies for different combinations of the diversification model's parameters, to assess under which scenarios the error made by BEAST2 cannot be neglected.
+Although a well-conducted theoretical study uses many replicates, this article only shows examples with one replicate, for the sake of clarity.
+
+\verb;pirouette; is very flexible and allows the user to specify a big variety of custom settings.
+These settings can be grouped in macro-sections, according to how they operate in the pipeline.
+We summarize them in Table~\ref{tab:definitions}.
+
+Although many possible tests can be performed, we present here only four example research questions (ERQ) that \verb;pirouette; can answer.
+For these ERQs we will estimate the error made by BEAST2 when using birth-death tree prior, in recovering an original tree generated under a Yule generative prior.
+
 \begin{table}
-  \begin{tabular}{ | l | l | }
-    \hline
-    \textbf{Term} & \textbf{Definition} \\ 
-    \hline
-    Clock model & Model for the DNA mutation rates \\
-    Experiment & Combination of an inference model and the conditions to actually use it in inference \\
-    Generative model & Combination of site model, clock model and tree prior, used to simulate a DNA alignment \\
-    Inference model & Combination of site model, clock model, tree prior, MRCA prior and MCMC, used to infer a BEAST2 posterior \\
+\centering
+  \begin{tabular}{|@{}c|@{}l|p{6cm}@{}|}
+    \hline
+    \textbf{Name} & \textbf{Meaning} & \textbf{Definition} \\ 
+    \hline
+    $\mathit{t}$ & tree prior & Used to simulate a DNA alignment\\
+    $\mathit{c}$ & clock model & Specifies the clock for the DNA mutation rates\\
+    $\mathit{s}$ & site model & Specifies the nucleotide substitution model\\
+    $\mathit{\mu}$ & mutation rate & Defines the pace at which mutation occurs\\
+    $\mathit{r}$ & root sequence & Specifies DNA sequence at the root of the tree\\
+    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent common ancestor\\
+    $\mathit{v}$ & MCMC settings & Setup of the Markov Chain Monte Carlo algorithm\\
+    \hline
+    $\mathit{P}$ & Generative prior & Specifies the diversification model according to which the 'true' tree is simulated\\
+    $\mathit{A}$ & Alignment model & Model for the alignment simulation. It's a choice of
+    $(\mathit{r},\mathit{\mu},\mathit{c},\mathit{s})$\\
+    $\mathit{G}$ & Generative model & A combination of the chosen $(\mathit{P},\mathit{A})$\\
+    $\mathit{I}$ & Inference model & Specifies how to execute the inference. It's a choice of
+    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
+    $\mathit{C}$ & Inference conditions & Specifies under which conditions an inference model is used\\
+    $\mathit{X}$ & Experiment & A combination of the chosen $(\mathit{I},\mathit{C})$\\
+    $\mathit{E}$ & Error parameters & Sets how to measure errors\\
+    $\mathit{T}$ & Twinning parameters & Specifies how to twin the 'true' tree\\
+    \hline
+  \end{tabular}
+  \caption{
+    Definitions of terms and relative symbols used in this article.
+  }
+  \label{tab:definitions}
+  \giovanni{It would be nice to specify (in the first part of the table) the basic elements composing all the elements of the second part of the table.
+This could be probably done for "Alignment parameters", "Inference conditions" and "Twinning parameters".}
+  \giovanni{Shall we add another column with the possible option for each lower case setting? e.g. "clock model" = c("strict", "rln"); "tree prior" = c("yule", "bd")}
+  \giovanni{How do we specify what constitutes 'twinning parameters'?}
+\end{table}
+
+\iffalse
+Experiment & Combination of an inference model and the \\ 
+    & conditions to actually use it in inference \\
+    Generative model & Combination of site model, clock model and \\
+    & tree prior, used to simulate a DNA alignment \\
+    Inference model & Combination of site model, clock model, tree prior,\\
+    & MRCA prior and MCMC, used to infer \\
+    & a BEAST2 posterior \\
     MCMC & Setup of the Markov Chain Monte Carlo algorithm \\
@@
     Tree prior & Model for the macroevolutionary speciation process \\
-    \hline
-  \end{tabular}
-  \caption{
-    Definitions of the terms used in this article
-  }
-  \label{tab:definitions}
-\end{table}
-
-\verb;pirouette; is written in the R programming language (\cite{R}).
-The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes for a given reconstructed phylogeny, within the context of
-a simulation study on a (possibly new) diversification model. Such
-a study will typically construct phylogenies for different 
-combinations of the diversification model's parameters, 
-to assess under which scenarios the error made by BEAST2 cannot be neglected.
-Although a well-conducted theoretical study uses many replicates,
-this article only shows examples with one replicate, 
-for the sake of readability.
-
+\fi
+
+\iffalse
+$\mathit{t}$ & tree prior & Used to simulate a DNA alignment\\
+    $\mathit{c}$ & clock model & Specifies the clock for the DNA \\
+    & & mutation rates\\
+    $\mathit{s}$ & site model & Specifies how nucleotides change\\
+    & & in one another\\
+    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent\\
+    & & common ancestor\\
+    $\mathit{v}$ & MCMC settings & Setup of the Markov Chain\\
+    & &  Monte Carlo algorithm\\
+
+    $\mathit{G}$ & Generative model & Specifies how to create the 'true'\\
+    & &  tree. It's a choice of $(\mathit{c},\mathit{s},\mathit{t})$\\
+    $\mathit{T}$ & Twinning parameters & Specifies how to twin the 'true' tree\\
+    $\mathit{A}$ & Alignment parameters & Parameters for the alignment\\
+    & & simulation\\
+    $\mathit{I}$ & Inference model & Specifies how to execute the inference. \\
+    & & It's a choice of
+    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
+    $\mathit{C}$ & Inference conditions & Specifies under which conditions an\\
+    & & inference model is used\\
+    $\mathit{X}$ & Experiment & A combination of the chosen $(\mathit{I},\mathit{C})$\\
+    $\mathit{E}$ & Error parameters & Sets how to measure errors\\
+\fi
+
+\iffalse
 \richel{
-  I do think this is a clumsy paragraph. I hope you can merge it
-  in the text
-}
-To explain how \verb;pirouette; measures the inference error
-BEAST2 makes, we first give two 
-definitions (see table \ref{tab:definitions} for an overview of definitions):
-(1) a generative model is a combination of a known tree prior (that generated
-a given phylogeny) and a site and clock model (that generate an alignment 
-from the given tree), (2) an inference model is a combination of 
-site model, clock model, tree prior, an optional most recent 
-common-ancestor (MRCA) prior (to create a time-calibrated phylogeny)
-and a Markov chain Monte Carlo (MCMC) setup.
-A generative model is a known/true model underlying a phylogeny 
-and is used to simulate DNA alignments.
-An inference model
-is used to produce a posterior distribution of trees and model parameters.
+  I do think this is a clumsy paragraph. I hope you can merge it in the text
+}
+To explain how \verb;pirouette; measures the inference error BEAST2 makes, we first give two definitions (see Table~\ref{tab:definitions} for an overview of definitions): (1) a generative model is a combination of a known tree prior (that generated a given phylogeny) and a site and clock model (that generate an alignment from the given tree), (2) an inference model is a combination of site model, clock model, tree prior, an optional most recent common-ancestor (MRCA) prior (to create a time-calibrated phylogeny) and a Markov chain Monte Carlo (MCMC) setup.
+A generative model is a known/true model underlying a phylogeny and is used to simulate DNA alignments. An inference model is used to produce a posterior distribution of trees and model parameters.
+\fi
 
@@
     \texttt{pirouette} pipeline. 
-    The pipeline starts from a true phylogeny (1a), 
-    that is converted to an alignment (2a).
-    The use supplies one or more experiments, in which an experiment is
-    defined as a combination of 
-    an inference model and the conditions in which it is actually used in
-    inference.
-    Based on the experiments and possibly the alignment (2a),
-    experiments are selected to be actually used used in the next step (3a).
-    Those experiments that are selected, use the   
-    alignment (2a) and their inference model 
+    The pipeline starts from a true phylogeny (1a), that is converted to an alignment (2a). The user supplies one or more experiments. Every experiment is
+    defined as a combination of an inference model and the conditions in which it is actually used in the inference.
+Based on the experiments and possibly the alignment (2a), experiments are selected to be actually used used in the next step (3a).
+    Selected experiments use the alignment (2a) and their inference model 
     to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a).
-    The posterior trees are compared to the true phylogeny (1a) 
-    using the nLTT statistic [\cite{janzen2015approximate}]. 
+The posterior trees are compared to the true phylogeny (1a) using the nLTT statistic [\cite{janzen2015approximate}].
 (5a) shows the nLTT of the true tree in red,
-    the nLTT of all individual posterior trees in grey, and the average
-    nLTT statistic of all posterior trees in black.
-    Taking the differences between the true tree's nLTT and each posterior's
-    nLTT results in an error distribution, here displayed as a histogram
-    of error values (6a).
-    Optionally, a twin phylogeny (1b) can be generated from a true
-    phylogeny (1a), which then results in an error distribution (6b) with
-    similar intermediate steps 
+    the nLTT of all individual posterior trees in grey, and the average nLTT statistic of all posterior trees in black.
+Taking the differences between the true tree's nLTT and each posterior's nLTT results in an error distribution, here displayed as a histogram
+    of error values (6a). Optionally, a twin phylogeny (1b) can be generated from a true
+    phylogeny (1a), which then results in an error distribution (6b) with similar intermediate steps.
   }
   \label{fig:pipeline}
+    \giovanni{The black line for the average of all posterior trees is barely visible. Please, make it thicker.}
+    \giovanni{Modify the figure to include elements in table~\ref{tab:definitions}.
+It should be specified clearly where elements corresponding to capital symbols play their roles in the pipeline.
+Then we should include also these elements in the caption.}
 \end{figure}
+
+\iffalse
+\caption{
+    \texttt{pirouette} pipeline. 
+    The pipeline starts from a true phylogeny (1a), that is converted to an alignment (2a). The user supplies one or more experiments. Every experiment is
+    defined as a combination of an inference model and the conditions in which it is actually used in the inference.
+Based on the experiments and possibly the alignment (2a), experiments are selected to be actually used used in the next step (3a).
+    Selected experiments use the alignment (2a) and their inference model 
+    to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a).
+The posterior trees are compared to the true phylogeny (1a) using the nLTT statistic [\cite{janzen2015approximate}].
+(5a) shows the nLTT of the true tree in red,
+    the nLTT of all individual posterior trees in grey, and the average nLTT statistic of all posterior trees in black.
+Taking the differences between the true tree's nLTT and each posterior's nLTT results in an error distribution, here displayed as a histogram
+    of error values (6a). Optionally, a twin phylogeny (1b) can be generated from a true
+    phylogeny (1a), which then results in an error distribution (6b) with similar intermediate steps.
+  }
+\fi
 
@@
 \begin{enumerate}
-    \item for a given phylogeny an alignment is simulated for a known 'alignment' model;
-    \item for this alignment, model selection is performed to choose the best-fitting alignment model in inference (which may be different from the generative alignment model);
+    \item for a given phylogeny an alignment is simulated under a known alignment model;
+    \item for this alignment, according to the specified inference conditions, an inference model is picked (which may be different from the generative alignment model);
     \item the alignment is then used as BEAST2 input to infer a posterior distribution of phylogenies;
-    \item posterior phylogenies are compared with the given original phylogeny to estimate the error made;
+    \item posterior phylogenies are compared with the given original phylogeny to estimate the error made, according to the error parameters specified by the user;
 \end{enumerate}
-The pipeline is visualized in Fig.~\ref{fig:pipeline}.
-There is also the option to generate a 'twin tree',
-that goes through the same pipeline. 
+The pipeline is visualized in Fig.~\ref{fig:pipeline}. There is also the option to generate a 'twin tree', that goes through the same pipeline. 
 The utility of this twin tree will be explained below.
@@
 phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
-One can specify, most importantly, a root DNA sequence, 
-a DNA mutation rate, a site model, and a clock model
-\richel{
-  as the DNA mutation rate is the same over all branches, 
-  only the strict clock model makes sense. I suggest to
-  remove the option to specify a clock model
-}.
-This step is relatively fast, but longer
-DNA alignments will noticeably slow down the inference step.
+This operation is performed according to the alignment model specified by the user, which consists in a root DNA sequence, a DNA mutation rate, a site model and a clock model.
+Since all the branches must evolve at the same pace, the clock model is always set to strict for this step.
+\giovanni{I think we need to specify the clock model, even though the user cannot change it.
+This is necessary because we refer to it when specifying what it means to set the inference model equal to the generative model.
+Please check if my description of why the clock can only be strict is accurate.}
+This step is relatively fast, but longer DNA alignments will noticeably slow down the inference step.
 
@@
 We define an experiment as the combination of an inference model 
-and the conditions to actually use it in inference.
-There a two types of inference models, which are
-generative and candidate inference models.
-An experiment uses a generative model if the inference model is known.
-An experiment uses a candidate model if either the inference model is unknown,
+and the conditions to actually use it in the inference step.
+The inference model can be selected in two ways, either using the
+generative model or selecting one or more candidates out of a list of candidate inference models.
+An experiment can use the generative model as the inference model, if the generative model is known.
+An experiment uses a candidate model if either the generative model is unknown,
 or there is an interest in seeing how well non-generative models compete with
 the generative model.
-Typically, a generative model and/or a best candidate model are used 
-in the actual inference.
+Typically, a generative model and/or a best candidate model are used in the actual inference.
 When using an experiment with a generative model,
 it must use the same site and clock model used in the alignment simulation,
-as well as the tree prior that underlied the creation of the given phylogeny. 
+as well as the tree prior that underlied the creation of the given phylogeny \giovanni{the tree prior appears not to be part of the alignment model.
+This means that should be part of the generative prior.
+Is it true?
+In case we can refer explicitly to it, according to our definitions in table \ref{tab:definitions}}.
 When using experiments with candidate models,
@@
 Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac,
-therefore using the selection of candidate models is not available on Windows. 
-Note that there is the option to do just the model selection on Windows 
-from the browser using \verb;mcbette;, yet the computation for this model selection is
+therefore the selection of candidate models is not available directly on Windows. 
+However, even on Windows systems there is the option to perform the model selection from browser, using \verb;mcbette;, yet the computation for this model selection is
 restricted to one hour for free users of the host (Travis CI) continuous
@@
 alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a $\rightarrow$ 4a),
-using the inference models from the experiments 
-selected in the previous step. Per selected experiment, 
-one BEAST2 posterior is inferred, 
-using the \verb;babette; [\cite{bilderbeek2018babette}] R package.
-
-The fourth step compares the true tree to each posterior tree
+using the inference models from the experiments selected in the previous step.
+For each selected experiment, one BEAST2 posterior is inferred, using the \verb;babette; [\cite{bilderbeek2018babette}] R package.
+
+The fourth step compares the true tree to each posterior tree, calculating the
 error using the nLTT statistic (\cite{janzen2015approximate}) 
@@
 in an error distribution (Fig.~\ref{fig:pipeline}, 5a $\rightarrow$ 6a).
-The nLTT statistic is used by default, 
-but also a user-defined error statistic can be used. 
-As an example,
-\verb;pirouette; supplies one custom error statistic,
+The nLTT statistic is used by default, but also a user-defined error statistic can be used. 
+As an example, \verb;pirouette; supplies one custom error statistic,
 that uses an absolute difference in the gamma statistic [\cite{pybus2000testing}].
-Additionally, the user can specify the
-proportion of posterior phylogenies to 
+Additionally, the user can specify the proportion of posterior phylogenies to 
 discard (i.e. the burn-in), throwing away the first $10\%$
@@
 The goal of this procedure is to generate a tree
-with a tree prior present in BEAST2, 
+accordingly to the tree prior used in the generative model, 
 that is used as a control for the main experiment.
-To produce such a birth-death tree, we maximize a birth-death 
-likelihood [\cite{nee1994reconstructed}], 
-conditioned on having the same number of tips as the original tree,
-to infer its parameters.
-We use these parameters,  
-which are a speciation rate and extinction rate,
-to simulate the branching times for the 
-twin tree, which we combinate with the original tree's 
-topology.
+To produce such a birth-death tree, we maximize a birth-death likelihood [\cite{nee1994reconstructed}], conditioned on having the same number of tips as the original tree, to infer its parameters.
+We use these parameters, which are the per-species speciation and extinction rates, to simulate the branching times for the twin tree, which we combine with the original tree's topology.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: First research question}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A first research question that \verb;pirouette; answers is:
+\section{Usage: Example Research Question 1}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A first research question that \verb;pirouette; can answer is:
 
@@
 
-\richel{Suggest: use BD tree (again)}
-We start with a Yule (pure-birth) tree with five taxa and a crown age of ten.
+\richel{Suggest: use BD tree (again)} \giovanni{I think in this case we should indeed use a Yule tree prior, as we want to use a generative prior that is not birth-death (which is what BEAST2 is using for the inference).
+What do you think?}
+We start from a tree simulated under a Yule (pure-birth) generative prior. We choose to use a tree with five taxa and a crown age of ten.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Second research question}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A second research question that \verb;pirouette; answers, is:
+\section{Usage: Example Research Question 2}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A second research question that \verb;pirouette; can answer, is:
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Third research question}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A third research question that \verb;pirouette; answers is:
+\section{Usage: Example Research Question 3}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A third research question that \verb;pirouette; can answer is:
 
@@
 to the tree prior used by BEAST2 to produce a posterior. 
-Since there is coherence between the inference model and the generative model \giovanni{Do we need to specify what is a generative model?
-Are we using BD or Yule?
-In case we are using Yule are we sure there is actually coherence when inference model is bd?}, we expect BEAST2 to produce a posterior of phylogenies clearly more similar to the original one.
+Since there is coherence between the inference model and the generative model, we expect BEAST2 to produce a posterior of phylogenies clearly more similar to the original one. 
 We intend to use this as a control experiment.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Fourth research question}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A fourth research question that \verb;pirouette; answers, is:
+\section{Usage: Example Research Question 4}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A fourth research question that \verb;pirouette; can answer, is:
 
 "What is the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error?"
-\giovanni{Did we define already the "generative" model? If not we can maybe refer to it when we define alignment and inference model}
 
@@
 
+\iffalse
 Here we discuss how to interpret the results of \verb;pirouette;.
@@
 as shown in table \ref{tab:esses_example_4}.
+\fi
+
+\giovanni{This is a first draft on how I think the discussion should be setup. In this context there is no much need for the results in the table. Let me know what you think about it.}
+
+In the four example research questions we showed how to use \verb;pirouette; to assess if BEAST2 is able, with a birth-death prior, to recover the original tree simulated by a Yule process.
+In principle any other, possibly much more complex, generative prior can be tested.
+All tests were performed on a single original tree, hence not providing us enough statistical power to evaluate the level of accuracy of the inference.
+However if the same procedure were repeated and performed on a distribution with a sufficient number of generative trees, it could actually constitute a quantitative assessment of the quality of the inference, as shown in figures \ref{fig:example_1}, \ref{fig:example_2}, \ref{fig:example_3}, \ref{fig:example_4}.
+Importantly, \verb;pirouette; can directly show the comparison between the results obtained on a new generative prior as compared to the error BEAST2 would make when operating on a tree generated under the right tree prior (the twin tree).
+From such results an user can estimate whether or not a new tree prior, tailored on the generative process, is needed.
+If this would indeed be the case, one could implement their own tree prior as a new module in BEAST2.
+\giovanni{Maybe we can add a bit more on the nested sampling.}
 

### 2019-03-05 article.tex
Getting better

@@
 The goal of \verb;pirouette; is to measure the inference error BEAST2
-makes for a given reconstructed phylogeny, simulated under the mechanisms of a (possibly new) diversification model, which we will refer to as the 'generative prior'.
-Such a study will typically construct phylogenies for different combinations of the diversification model's parameters, to assess under which scenarios the error made by BEAST2 cannot be neglected.
-Although a well-conducted theoretical study uses many replicates, this article only shows examples with one replicate, for the sake of clarity.
-
-\verb;pirouette; is very flexible and allows the user to specify a big variety of custom settings.
-These settings can be grouped in macro-sections, according to how they operate in the pipeline.
+makes for a given reconstructed phylogeny, 
+simulated under the mechanisms of a (possibly new) diversification model, 
+which we will refer to as the 'generative prior'. 
+Such a study will typically construct phylogenies 
+for different combinations of the diversification model's parameters, 
+to assess under which scenarios the error made by BEAST2 cannot be neglected. 
+Although a well-conducted theoretical study uses many replicates, 
+this article only shows examples with one replicate, for the sake of clarity.
+
+\verb;pirouette; is very flexible and allows the user 
+to specify a big variety of custom settings. 
+These settings can be grouped in macro-sections, 
+according to how they operate in the pipeline. 
 We summarize them in Table~\ref{tab:definitions}.
 
-Although many possible tests can be performed, we present here only four example research questions (ERQ) that \verb;pirouette; can answer.
-For these ERQs we will estimate the error made by BEAST2 when using birth-death tree prior, in recovering an original tree generated under a Yule generative prior.
+Although many possible tests can be performed, 
+we present here only four example research questions (ERQ) that \verb;pirouette; can answer. 
+For these ERQs we will estimate the error made by BEAST2 when 
+using a birth-death tree prior, in recovering an original tree 
+generated under a Yule generative prior.
 
@@
 \centering
-  \begin{tabular}{|@{}c|@{}l|p{6cm}@{}|}
-    \hline
-    \textbf{Name} & \textbf{Meaning} & \textbf{Definition} \\ 
-    \hline
-    $\mathit{t}$ & tree prior & Used to simulate a DNA alignment\\
-    $\mathit{c}$ & clock model & Specifies the clock for the DNA mutation rates\\
-    $\mathit{s}$ & site model & Specifies the nucleotide substitution model\\
-    $\mathit{\mu}$ & mutation rate & Defines the pace at which mutation occurs\\
-    $\mathit{r}$ & root sequence & Specifies DNA sequence at the root of the tree\\
+  \begin{tabular}{|@{}c|p{3cm}|@{}l|p{6cm}@{}|}
+    \hline
+    \textbf{Symbol} & \textbf{Name} & \textbf{Definition} \\ 
+    \hline
+    $\mathit{t}$ & Tree prior & Macroevolutionary speciation model \\
+    $\mathit{c}$ & Clock model & Clock for the DNA mutation rates\\
+    $\mathit{s}$ & Site model & Nucleotide substitution model\\
+    $\mathit{\mu}$ & Mutation rate & Pace at which mutation occurs, $\mu \in \mathit{s}$ \\
+    $\mathit{r}$ & Root sequence & DNA sequence at the root of the tree\\
     $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent common ancestor\\
-    $\mathit{v}$ & MCMC settings & Setup of the Markov Chain Monte Carlo algorithm\\
-    \hline
-    $\mathit{P}$ & Generative prior & Specifies the diversification model according to which the 'true' tree is simulated\\
-    $\mathit{A}$ & Alignment model & Model for the alignment simulation. It's a choice of
-    $(\mathit{r},\mathit{\mu},\mathit{c},\mathit{s})$\\
-    $\mathit{G}$ & Generative model & A combination of the chosen $(\mathit{P},\mathit{A})$\\
-    $\mathit{I}$ & Inference model & Specifies how to execute the inference. It's a choice of
+    $\mathit{v}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup \\
+    \hline
+    $\mathit{P}$ & Generative prior & Diversification model the 'true' tree is simulated with \\
+    $\mathit{A}$ & Alignment model & Alignment simulation model, $\mathit{A} \ni \{ \mathit{r}, \mathit{\mu}, \mathit{c}, \mathit{s} \}$, in which $\mathit{c}$ = strict \\
+    $\mathit{G}$ & Generative model & Combination of a known $\mathit{P}$ and chosen (thus known) $\mathit{A}$ \\
+    $\mathit{I}$ & Inference model & Bayesian phylogenetic inference model, a combination of
     $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
-    $\mathit{C}$ & Inference conditions & Specifies under which conditions an inference model is used\\
-    $\mathit{X}$ & Experiment & A combination of the chosen $(\mathit{I},\mathit{C})$\\
-    $\mathit{E}$ & Error parameters & Sets how to measure errors\\
-    $\mathit{T}$ & Twinning parameters & Specifies how to twin the 'true' tree\\
+    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in inference \\
+    $\mathit{X}$ & Experiment & Combination of $(\mathit{I},\mathit{C})$\\
+    $\mathit{E}$ & Error measurement parameters & Errors measurement setup \\
+    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure \\
     \hline 
@@
   \label{tab:definitions}
-  \giovanni{It would be nice to specify (in the first part of the table) the basic elements composing all the elements of the second part of the table.
-This could be probably done for "Alignment parameters", "Inference conditions" and "Twinning parameters".}
-  \giovanni{Shall we add another column with the possible option for each lower case setting? e.g. "clock model" = c("strict", "rln"); "tree prior" = c("yule", "bd")}
-  \giovanni{How do we specify what constitutes 'twinning parameters'?}
+  \richel{I've some set notations in some spots, as I am unsure if you like it}
+  \giovanni{It would be nice to specify (in the first part of the table) 
+    the basic elements composing all the elements of the second part of the table. 
+    This could be probably done for "Alignment parameters", "Inference conditions" 
+    and "Twinning parameters".
+  }
+  \richel{
+    I feel the next table (which was also your suggestion), does so
+    at the right amount of detail
+  }
+  \giovanni{
+    How do we specify what constitutes 'twinning parameters'?
+  }
+  \richel{Perhaps you'd enjoy my definition?}
 \end{table}
 
-\iffalse
-Experiment & Combination of an inference model and the \\ 
-    & conditions to actually use it in inference \\
-    Generative model & Combination of site model, clock model and \\
-    & tree prior, used to simulate a DNA alignment \\
-    Inference model & Combination of site model, clock model, tree prior,\\
-    & MRCA prior and MCMC, used to infer \\
-    & a BEAST2 posterior \\
-    MCMC & Setup of the Markov Chain Monte Carlo algorithm \\
-    MRCA prior & Prior knowledge of a most recent common ancestor \\
-    Site model & Model for the DNA nucleotide substitions \\
-    Tree prior & Model for the macroevolutionary speciation process \\
-\fi
-
-\iffalse
-$\mathit{t}$ & tree prior & Used to simulate a DNA alignment\\
-    $\mathit{c}$ & clock model & Specifies the clock for the DNA \\
-    & & mutation rates\\
-    $\mathit{s}$ & site model & Specifies how nucleotides change\\
-    & & in one another\\
-    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent\\
-    & & common ancestor\\
-    $\mathit{v}$ & MCMC settings & Setup of the Markov Chain\\
-    & &  Monte Carlo algorithm\\
-
-    $\mathit{G}$ & Generative model & Specifies how to create the 'true'\\
-    & &  tree. It's a choice of $(\mathit{c},\mathit{s},\mathit{t})$\\
-    $\mathit{T}$ & Twinning parameters & Specifies how to twin the 'true' tree\\
-    $\mathit{A}$ & Alignment parameters & Parameters for the alignment\\
-    & & simulation\\
-    $\mathit{I}$ & Inference model & Specifies how to execute the inference. \\
-    & & It's a choice of
-    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
-    $\mathit{C}$ & Inference conditions & Specifies under which conditions an\\
-    & & inference model is used\\
-    $\mathit{X}$ & Experiment & A combination of the chosen $(\mathit{I},\mathit{C})$\\
-    $\mathit{E}$ & Error parameters & Sets how to measure errors\\
-\fi
-
-\iffalse
-\richel{
-  I do think this is a clumsy paragraph. I hope you can merge it in the text
-}
-To explain how \verb;pirouette; measures the inference error BEAST2 makes, we first give two definitions (see Table~\ref{tab:definitions} for an overview of definitions): (1) a generative model is a combination of a known tree prior (that generated a given phylogeny) and a site and clock model (that generate an alignment from the given tree), (2) an inference model is a combination of site model, clock model, tree prior, an optional most recent common-ancestor (MRCA) prior (to create a time-calibrated phylogeny) and a Markov chain Monte Carlo (MCMC) setup.
-A generative model is a known/true model underlying a phylogeny and is used to simulate DNA alignments. An inference model is used to produce a posterior distribution of trees and model parameters.
-\fi
+\begin{table}
+\centering
+  \begin{tabular}{| c | c | l |}
+    \hline
+    \textbf{Symbol} & \textbf{Subparameter} & \textbf{Possible values} \\ 
+    \hline
+    $\mathit{t}$ & name & Yule, BD, CBS, CCP, CEP \\
+    $\mathit{c}$ & clock model & strict, relaxed log-normal \\
+    $\mathit{s}$ & site model & JC, HKY, TN, GTR \\
+    $\mathit{C}$ & model type & Generative, candidate \\
+    $\mathit{C}$ & run if & Always, best candidate \\
+    $\mathit{E}$ & function & nLTT, $|\gamma|$ \\
+    $\mathit{T}$ & model & Yule, BD \\
+    \hline
+  \end{tabular}
+  \caption{
+    Most important parameter options. BD = birth death, CBS = coalescent
+    Bayesian Skyline, CCP = coalescent constant-population, CEP = coalescent
+    exponential-population, JC = Jukes-Cantor 1969, HKY = Hasegawa, Kishino and 
+    Yano 1985, TN = Tamura 1992, GTR = Generalised time-reversible 
+    model (Tavaré 1986)
+  }
+  \label{tab:options}
+\end{table}
 
@@
   \label{fig:pipeline}
-    \giovanni{The black line for the average of all posterior trees is barely visible. Please, make it thicker.}
-    \giovanni{Modify the figure to include elements in table~\ref{tab:definitions}.
-It should be specified clearly where elements corresponding to capital symbols play their roles in the pipeline.
-Then we should include also these elements in the caption.}
 \end{figure}
-
-\iffalse
-\caption{
-    \texttt{pirouette} pipeline. 
-    The pipeline starts from a true phylogeny (1a), that is converted to an alignment (2a). The user supplies one or more experiments. Every experiment is
-    defined as a combination of an inference model and the conditions in which it is actually used in the inference.
-Based on the experiments and possibly the alignment (2a), experiments are selected to be actually used used in the next step (3a).
-    Selected experiments use the alignment (2a) and their inference model 
-    to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a).
-The posterior trees are compared to the true phylogeny (1a) using the nLTT statistic [\cite{janzen2015approximate}].
-(5a) shows the nLTT of the true tree in red,
-    the nLTT of all individual posterior trees in grey, and the average nLTT statistic of all posterior trees in black.
-Taking the differences between the true tree's nLTT and each posterior's nLTT results in an error distribution, here displayed as a histogram
-    of error values (6a). Optionally, a twin phylogeny (1b) can be generated from a true
-    phylogeny (1a), which then results in an error distribution (6b) with similar intermediate steps.
-  }
-\fi
 
@@
 \end{enumerate}
-The pipeline is visualized in Fig.~\ref{fig:pipeline}. There is also the option to generate a 'twin tree', that goes through the same pipeline. 
+The pipeline is visualized in Fig.~\ref{fig:pipeline}. 
+There is also the option to generate a 'twin tree', 
+that goes through the same pipeline. 
 The utility of this twin tree will be explained below.
@@
 phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
-This operation is performed according to the alignment model specified by the user, which consists in a root DNA sequence, a DNA mutation rate, a site model and a clock model.
-Since all the branches must evolve at the same pace, the clock model is always set to strict for this step.
-\giovanni{I think we need to specify the clock model, even though the user cannot change it.
-This is necessary because we refer to it when specifying what it means to set the inference model equal to the generative model.
-Please check if my description of why the clock can only be strict is accurate.}
-This step is relatively fast, but longer DNA alignments will noticeably slow down the inference step.
+This operation is performed according to the alignment model 
+specified by the user, which consists in a root DNA sequence, 
+a DNA mutation rate and a site model.
+A strict clock model is used as the only clock model
+type, as the currently used alignment simulation algorithm
+only supports a mutation rate that is equal across all the branches.
+This step is relatively fast, 
+but longer DNA alignments will noticeably slow down the inference step.
 
@@
 it must use the same site and clock model used in the alignment simulation,
-as well as the tree prior that underlied the creation of the given phylogeny \giovanni{the tree prior appears not to be part of the alignment model.
+as well as the tree prior that underlied the creation of the given phylogeny 
+\giovanni{
+  the tree prior appears not to be part of the alignment model. 
   This means that should be part of the generative prior. 
-Is it true?
-In case we can refer explicitly to it, according to our definitions in table \ref{tab:definitions}}.
+  Is it true? In case we can refer explicitly to it, 
+  according to our definitions in table \ref{tab:definitions}
+}
+\richel{
+  Indeed, a tree prior is not part of the alignment model.
+  Go ahead and show how you'd like to refer explicitly to the
+  table!
+}
+. 
 When using experiments with candidate models,
@@
 a birth-death tree model.
-\richel{Should be switched to BD, pirouette Issue 161}.
 The goal of this procedure is to generate a tree
@@
 \verb;pirouette; can be installed easily from CRAN:
-\rampal{be careful here; it is not on CRAN yet. Better to say it can be installed from GitHub and will be made available on CRAN)}
-\richel{We adopted this policy (to say it's on CRAN in advance) with babette as well. Is this a hint to get pirouette on CRAN before the publication of this article :-) ?}
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]
@@
 
-\richel{Suggest: use BD tree (again)} \giovanni{I think in this case we should indeed use a Yule tree prior, as we want to use a generative prior that is not birth-death (which is what BEAST2 is using for the inference).
-What do you think?}
-We start from a tree simulated under a Yule (pure-birth) generative prior. We choose to use a tree with five taxa and a crown age of ten.
+\giovanni{
+  I think in this case we should indeed use a Yule tree prior [...]. What do you think?}
+\richel{
+  Agrees. I was wrong, possibly due to carnaval. Yule is the way to go!
+}
+We start from a tree simulated under a Yule (pure-birth) generative prior. 
+We choose to use a tree with five taxa and a crown age of ten.
 
@@
 
-The following settings are the same as in the previous section:
-phylogeny (listing \ref{lst:unknown_phylogeny}), 
-alignment parameters (listing \ref{lst:create_alignment}), 
-experiments (listing \ref{lst:create_generative_experiment}),
-BEAST2 inference and error measuring parameters \giovanni{can we add a reference to the "BEAST2 inference and error measuring parameters" we are using?}.
-
-This time, we are interested in creating a twin tree. A twin tree
+Most of the settings for this example research question are
+the same as in the previous section, which are
+the phylogeny (listing \ref{lst:unknown_phylogeny}), 
+the alignment parameters (listing \ref{lst:create_alignment}) 
+and the list of experiments (listing \ref{lst:create_generative_experiment}).
+
+This time, however, we are interested in creating a twin tree,
+as such a twin tree allows us to measure a baseline error produced
+by the BEAST2 run.
+A twin tree
 has the same topology as the given tree, 
 yet its branching times are simulated according 
-to the tree prior used by BEAST2 to produce a posterior. 
-Since there is coherence between the inference model and the generative model, we expect BEAST2 to produce a posterior of phylogenies clearly more similar to the original one. 
-We intend to use this as a control experiment.
+to a tree prior already implemented by BEAST2 to produce a posterior. 
+Because for a twin tree we know the tree prior 
+and because we use that tree prior in inference, 
+we expect BEAST2 to produce a posterior of phylogenies
+that is close to the twin tree.
+As the inference model for the twin tree is (more) completely
+specified than the true tree, 
+we expect that a twin tree and its posterior trees are more similar
+than a true tree and its posterior trees.
+Due to this, the twinning is used a control for the experiment.
 
@@
 
-We can now measure the errors made by BEAST2 when inferring the given phylogeny versus the error it makes when the same procedure is applied to the twin tree.
+We can now measure the errors made by BEAST2 
+when inferring the given phylogeny 
+versus the error it makes when the same procedure is applied to the twin tree.
 
@@
 \giovanni{describe winner}
+\richel{indeed, after the figures also show it :-) }
 
@@
 \richel{check numbers}
-\giovanni{I think that the current version of the discussion is too didascalic.
+\giovanni{
+  I think that the current version of the discussion is too didascalic. 
   Most of what is said can be effectively summarized by table 5. 
   We can even add ESSes in it. 
-Another point is that we should understand what is the message we want to deliver in this discussion.
-Since we are only using one tree we are not in the position to claim anything, because a single tree is not an effective sample.
+}
+\giovanni{
+  Another point is that we should understand what is 
+  the message we want to deliver in this discussion. 
+  Since we are only using one tree we are not in the position to claim anything, 
+  because a single tree is not an effective sample. 
   I saw that in some occasions you refer to the RQ as examples. 
-This might be the right way to proceed, as our purpose is just to illustrate how to run all the possible experiments on a full tree distribution.
+  This might be the right way to proceed, 
+  as our purpose is just to illustrate how to run all the possible experiments 
+  on a full tree distribution. 
   I think we should make it clearer throughout the entire manuscript. 
-I am also interested in hearing Rampal's opinion on this.}
+  I am also interested in hearing Rampal's opinion on this.
+}
 
@@
 
-\giovanni{This is a first draft on how I think the discussion should be setup. In this context there is no much need for the results in the table. Let me know what you think about it.}
+\giovanni{
+  This is a first draft on how I think the discussion should be setup. 
+  In this context there is no much need for the results in the table. 
+  Let me know what you think about it.
+}
 

### 2019-03-05 article.tex
Add symbols to workflow figure

@@
 \giovanni{describe winner}
+\richel{agreed, after the figures also show it :-) }
 
@@
 \giovanni{describe winner}
-\richel{indeed, after the figures also show it :-) }
+\richel{agreed, after the figures also show it :-) }
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[h]
-\centering
-\begin{tabular}{ | r | l | l | l | l | }
-\hline
-\multirow{2}{*}{\textbf{Example}} & \multicolumn{2}{c|}{\textbf{True tree}} 
-                                  & \multicolumn{2}{c|}{\textbf{Twin tree}} \\
-\cline{2-5}
-                                  & \textbf{median} & \textbf{range} & \textbf{median} & \textbf{range} \\
-\hline
-1  & $0.02$  & $0.0 - 0.27$  &         &               \\
-2  & $0.02$  & $0.0 - 0.08$  &         &               \\
-3  & $0.018$ & $0.0 - 0.1$   & $0.019$ & $0.0 - 0.11$  \\
-4C & $0.22$  & $0.0 - 0.088$ & $0.009$ & $0.0 - 0.056$ \\
-4G & $0.020$ & $0.0 - 0.075$ & $0.013$ & $0.0 - 0.052$ \\
-\hline
-\end{tabular}
-\caption{Results of the examples. 
-  For the true and twin tree,
-  the median and range (from minimum to maximum value) of the errors 
-  is shown. Experiment 4 is split in two rows
-  by either the candidate ('C') or generative model ('G')
-}
-\label{tab:results}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \richel{check numbers}
@@
   We can even add ESSes in it. 
+}
+\richel{
+  Agreed, I disliked it as well
 }
@@
 
-\iffalse
-Here we discuss how to interpret the results of \verb;pirouette;.
-Figure \ref{fig:example_1} shows that the generative model 
-gives an error distribution with a median of \richel{?}
-The range of errors goes from \richel{?} to \richel{?}.
-As we know the generative model, this error distribution is the minimum noise that BEAST2 produces for that generative model.
-We can measure the effective samples as well (not shown), 
-and all the Estimated Sample Sizes (ESSes) are above
-the recommended value of 200 [\cite{drummond2015bayesian}], 
-as shown in table \ref{tab:esses_example_1}.
-
-Figure \ref{fig:example_2} shows that the best candidate model gives an error distribution with a median of \richel{?}. 
-The range of errors goes from \richel{?} to \richel{?}. 
-The inference model with the highest evidence follows a JC site model, RLN clock model and BD tree prior, with a weight of $0.675$. 
-This is (unexpectedly) \giovanni{We are not using distributions, so we are totally subject to stochasticity. Nothing can be expected from just one tree.} better than the generative
-model with a weight of $0.166$. 
-The complete list of all evidences is shown in table \ref{tab:evidences_example_2}.
-This means that it is rational to use this inference model to 
-infer a posterior phylogeny distribution, as is done by \verb;pirouette;.
-Also here, the ESSes are well above 200, 
-as shown in table \ref{tab:esses_example_2}.
-
-Figure \ref{fig:example_3} shows that the hand-picked model 
-gives an error distribution with a median of \richel{?}, 
-The twin tree has an error distribution with a median of \richel{?},
-with a range from 0.0 to \richel{?}, which is an indication of the
-error made by BEAST on that tree prior.
-As there is little difference between the true and twin tree,
-the (unknown) diversification model has little impact on the error
-BEAST2 makes, hinting that the constant-rate birth-death model is
-sufficiently good enough to explain the true phylogeny. 
-The diversification model apparently has minor impact and need not to be
-added to BEAST2.
-We can draw this conclusion with confidence, as the ESSes are well above 200, 
-as shown in table \ref{tab:esses_example_3}.
-
-Figure \ref{fig:example_4} shows that \richel{?}
-and 
-gives an error distribution with a median of \richel{?}, 
-The twin tree has an error distribution with a median of \richel{?},
-with a range from 0.0 to \richel{?}, which is an indication of the
-error made by BEAST on that tree prior.
-The inference model with the highest evidence follows a JC site model, 
-RLN clock model and BD tree prior, with a weight of ?. 
-The complete list of all evidences is shown in table \ref{tab:evidences_example_4}.
-As there is little difference between the true and twin tree,
-the (unknown) diversification model has little impact on the error
-BEAST2 makes, hinting that the constant-rate birth-death model is
-sufficiently good enough to explain the true phylogeny. 
-The diversification model apparently has minor impact and need not to be
-added to BEAST2.
-We can draw this conclusion with confidence \giovanni{no, we cannot: one tree is not enough}, as the ESSes are well above 200, 
-as shown in table \ref{tab:esses_example_4}.
-\fi
-
 \giovanni{
@@
 }
-
-In the four example research questions we showed how to use \verb;pirouette; to assess if BEAST2 is able, with a birth-death prior, to recover the original tree simulated by a Yule process.
-In principle any other, possibly much more complex, generative prior can be tested.
-All tests were performed on a single original tree, hence not providing us enough statistical power to evaluate the level of accuracy of the inference.
-However if the same procedure were repeated and performed on a distribution with a sufficient number of generative trees, it could actually constitute a quantitative assessment of the quality of the inference, as shown in figures \ref{fig:example_1}, \ref{fig:example_2}, \ref{fig:example_3}, \ref{fig:example_4}.
-Importantly, \verb;pirouette; can directly show the comparison between the results obtained on a new generative prior as compared to the error BEAST2 would make when operating on a tree generated under the right tree prior (the twin tree).
-From such results an user can estimate whether or not a new tree prior, tailored on the generative process, is needed.
-If this would indeed be the case, one could implement their own tree prior as a new module in BEAST2.
-\giovanni{Maybe we can add a bit more on the nested sampling.}
+\richel{
+  I think that was the right way forward. 
+  I do think the reader would like help interpreting the
+  results of the example research questions as if these were
+  proper experiments. I added that
+}
+
+In the four example research questions we showed how to use \verb;pirouette; 
+to assess if BEAST2 is able, with a birth-death prior, 
+to recover the original tree simulated by a Yule process. 
+In principle any other, possibly much more complex, 
+generative prior can be tested.
+All tests were performed on a single original tree, 
+hence not providing us enough statistical power to 
+evaluate the level of accuracy of the inference.
+
+However if the same procedure were repeated and performed on a distribution 
+with a sufficient number of generative trees, 
+it could actually constitute a quantitative assessment 
+of the quality of the inference.
+To give an idea of the insights \verb;pirouette; can give,
+we pretend the results of the example research questions
+to be obtained from thousands of replicates.
+In that case, figure \ref{fig:example_1} shows 
+the minimum error BEAST2 when the
+generative model is known, which can serve as 
+a minimum noise level assessment in a bigger experiment.  
+Figure \ref{fig:example_2} shows that 
+\richel{
+  reword after new results with 6 taxa tree
+} 
+the best candidate model gives an error distribution similar to the unknown tree prior,
+hinting that the candidate model would be just as good in inference.
+Figure \ref{fig:example_3} shows that 
+\richel{
+  reword after new results with 6 taxa tree
+} 
+the twin model gives an error distribution similar to the unknown tree prior,
+hinting that the shape of the phylogeny with unknown tree prior
+does not effect the error made by BEAST2.
+Figure \ref{fig:example_4} shows that 
+\richel{
+  reword after new results with 6 taxa tree
+} 
+the twin model and best candidate model give 
+an error distribution similar to the unknown tree prior,
+hinting that the shape of the phylogeny with unknown tree prior
+does not effect the error made by BEAST2, 
+and that the candidate model would be just as good in inference.
+
+In conclusion, \verb;pirouette; can directly show 
+the comparison between the results obtained 
+on a new generative prior as compared to the error BEAST2 
+would make when operating on a tree generated 
+under a known, right and existing tree prior (the twin tree).
+From such results an user can estimate whether or not a new tree prior, 
+tailored on the generative process, is needed. 
+If this would indeed be the case, 
+one could implement their own tree prior as a new module in BEAST2.
 
@@
 
-\verb;pirouette;'s documentation is extensive. All functions are documented in the package's internal documentation. For quick use, 
+\verb;pirouette;'s documentation is extensive. 
+All functions are documented in the package's internal documentation. 
+For quick use, 
 each exported function shows a minimal example. 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[h]
+\centering
+\begin{tabular}{ | r | l | l | l | l | }
+\hline
+\multirow{2}{*}{\textbf{Example}} & \multicolumn{2}{c|}{\textbf{True tree}} 
+                                  & \multicolumn{2}{c|}{\textbf{Twin tree}} \\
+\cline{2-5}
+                                  & \textbf{median} & \textbf{range} & \textbf{median} & \textbf{range} \\
+\hline
+1  & $0.02$  & $0.0 - 0.27$  &         &               \\
+2  & $0.02$  & $0.0 - 0.08$  &         &               \\
+3  & $0.018$ & $0.0 - 0.1$   & $0.019$ & $0.0 - 0.11$  \\
+4C & $0.22$  & $0.0 - 0.088$ & $0.009$ & $0.0 - 0.056$ \\
+4G & $0.020$ & $0.0 - 0.075$ & $0.013$ & $0.0 - 0.052$ \\
+\hline
+\end{tabular}
+\caption{
+  \richel{redo}
+  Results of the examples. 
+  For the true and twin tree,
+  the median and range (1st and 3rd quartile) of the errors 
+  is shown. Experiment 4 is split in two rows
+  by either the candidate ('C') or generative model ('G')
+}
+\label{tab:results}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \richel{

### 2019-03-07 article.tex
Updates from Overleaf

@@
 according to how they operate in the pipeline. 
-We summarize them in Table~\ref{tab:definitions}.
+We summarize them in Tables~\ref{tab:definitions} and \ref{tab:options}.
 
@@
 
+\begin{sidewaystable}
+%\begin{table}
+\centering
+  %\begin{tabular}{| c | c | l | l |}
+  \begin{tabular}{|@{}c|p{5.5cm}|p{2.85cm}@{}|}
+    \hline
+    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Possible values} \\ 
+    \hline
+    $\mathit{tp}$ & 'tree prior', i.e a macroevolutionary speciation model & Yule, BD, CBS, CCP, CEP \\
+    $\mathit{cm}$ & 'clock model' for the DNA mutation rates & strict, relaxed log-normal \\
+    $\mathit{sm}$ & 'site model' for nucleotide substitution & JC, HKY, TN, GTR \\
+    $\mathit{mr}$ & 'mutation rate' & \\
+    $\mathit{rs}$ & 'root sequence' of nucleotides at the root of the tree & \\
+    $\mathit{mp}$ & 'MRCA prior' knowledge of a most recent common ancestor & \\
+    $\mathit{ms}$ & 'MCMC settings' i.e. Markov Chain Monte Carlo algorithm setup & \\
+    $\mathit{et}$ & 'experiment type' & Generative, candidate \\
+    $\mathit{ri}$ & 'run if' & Always, Best candidate\\
+    $\mathit{ef}$ & 'error function' that specifies how to measure the error & nLTT, $|\gamma|$ \\
+    $\mathit{tm}$ & 'twin model' & Yule, BD \\
+    \hline
+  \end{tabular}
+  \caption{
+    Most important parameter options. BD = birth death, CBS = coalescent
+    Bayesian Skyline, CCP = coalescent constant-population, CEP = coalescent
+    exponential-population, JC = Jukes-Cantor 1969, HKY = Hasegawa, Kishino and 
+    Yano 1985, TN = Tamura 1992, GTR = Generalised time-reversible 
+    model (Tavaré 1986)
+  }
+  \label{tab:options}
+%\end{table}
+\end{sidewaystable}
+
+\begin{sidewaystable}
+%\begin{table}
+\centering
+  \begin{tabular}{|@{}c|p{6cm}|p{2.85cm}@{}|}
+    \hline
+    \textbf{Symbol} & \textbf{Macro-argument} & \textbf{Composed by}\\ 
+    \hline
+    $\mathit{P}$ & 'Generative prior' under which the 'true' tree is simulated & $\mathit{tp}$ \\
+    $\mathit{A}$ & 'Alignment model' used for simulation & $\mathit{cm}, \mathit{sm}, \mathit{mr}, \mathit{rs}$ \\
+    $\mathit{G}$ & 'Generative model' & $\mathit{P},\mathit{A}$ \\
+    $\mathit{I}$ & 'Inference model' to run BEAST2 & $\mathit{cm},\mathit{sm},\mathit{tp},\mathit{mp},\mathit{ms}$ \\
+    $\mathit{C}$ & 'Inference conditions' under which $\mathit{I}$ is used in the inference & $\mathit{et}, \mathit{ri}$\\
+    $\mathit{X}$ & 'Experiment' & $\mathit{I},\mathit{C}$ \\
+    $\mathit{E}$ & 'Error measurement parameters' & $\mathit{ef}$\\
+    $\mathit{T}$ & 'Twinning parameters' to set the twin creation procedure & $\mathit{tm}$\\
+    \hline 
+  \end{tabular}
+  \caption{
+    Definitions of terms and relative symbols used in this article.
+  }
+  \label{tab:definitions}
+%\end{table}
+\end{sidewaystable}
+
+\iffalse
 \begin{table}
 \centering
-  \begin{tabular}{|@{}c|p{3cm}|@{}l|p{6cm}@{}|}
+  \begin{tabular}{|@{}c|p{2.8cm}|p{6cm}@{}|}
+    \hline
+    \textbf{Symbol} & \textbf{Name} & \textbf{Definition} \\ 
+    \hline
+    $\mathit{t}$ & Tree prior & Macroevolutionary speciation model \\
+    $\mathit{c}$ & Clock model & Clock for the DNA mutation rates\\
+    $\mathit{s}$ & Site model & Nucleotide substitution model\\
+    $\mathit{\mu}$ & Mutation rate & Pace at which mutation occurs, $\mu \in \mathit{s}$ \\
+    $\mathit{r}$ & Root sequence & DNA sequence at the root of the tree\\
+    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent common ancestor \\
+    $\mathit{v}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup \\
+    \hline
+    $\mathit{P}$ & Generative prior & Diversification model under which the 'true' tree is simulated \\
+    $\mathit{A}$ & Alignment model & Alignment simulation model. It is a combination of $(\mathit{r}, \mathit{\mu}, \mathit{c}, \mathit{s})$, where $\mathit{c}$ = strict \\
+    $\mathit{G}$ & Generative model & Combination of a known $\mathit{P}$ and chosen (thus known) $\mathit{A}$ \\
+    $\mathit{I}$ & Inference model & Bayesian phylogenetic inference model, a combination of
+    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
+    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in inference \\
+    $\mathit{X}$ & Experiment & Combination of $(\mathit{I},\mathit{C})$\\
+    $\mathit{E}$ & Error measurement parameters & Errors measurement setup \\
+    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure \\
+    \hline 
+  \end{tabular}
+  \caption{
+    Definitions of terms and relative symbols used in this article.
+  }
+  \label{tab:definitions}
+  \richel{I've some set notations in some spots, as I am unsure if you like it}
+  \giovanni{It would be nice to specify (in the first part of the table) 
+    the basic elements composing all the elements of the second part of the table. 
+    This could be probably done for "Alignment parameters", "Inference conditions" 
+    and "Twinning parameters".
+  }
+  \richel{
+    I feel the next table (which was also your suggestion), does so
+    at the right amount of detail
+  }
+  \giovanni{
+    How do we specify what constitutes 'twinning parameters'?
+  }
+  \richel{Perhaps you'd enjoy my definition?}
+\end{table}
+\fi
+
+\iffalse
+% previous version
+\begin{table}
+\centering
+  \begin{tabular}{|@{}c|p{2.8cm}|@{}l|p{5.5cm}@{}|}
     \hline
@@
 \end{table}
-
+\fi
+
+\iffalse
+% previous version
 \begin{table}
@@
 \end{table}
+\fi
 
@@
     \texttt{pirouette} pipeline. 
-    The pipeline starts from a true phylogeny (1a), that is converted to an alignment (2a). The user supplies one or more experiments. Every experiment is
-    defined as a combination of an inference model and the conditions in which it is actually used in the inference.
-Based on the experiments and possibly the alignment (2a), experiments are selected to be actually used used in the next step (3a).
-    Selected experiments use the alignment (2a) and their inference model 
-    to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a).
-The posterior trees are compared to the true phylogeny (1a) using the nLTT statistic [\cite{janzen2015approximate}].
-(5a) shows the nLTT of the true tree in red,
-    the nLTT of all individual posterior trees in grey, and the average nLTT statistic of all posterior trees in black.
+    The pipeline starts from a true phylogeny (1a) simulated under the selected generative prior $\mathit{P}$, that is converted to an alignment (2a) according to the alignment model $\mathit{A}$.
+The user supplies one or more experiments.
+Every experiment $\mathit{X}$ is defined as a combination of an inference model $\mathit{I}$ and the conditions $\mathit{C}$ under which it is actually used in the inference.
+Based on the experiments and possibly the alignment (2a), experiments are selected to be actually used in the next step (3a).
+    Selected experiments use the alignment (2a) and their inference model $\mathit{I}$ to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a).
+The posterior trees are compared to the true phylogeny (1a) using the selected error measure $\mathit{E}$, in our case the nLTT statistic [\cite{janzen2015approximate}].
 Taking the differences between the true tree's nLTT and each posterior's nLTT results in an error distribution, here displayed as a histogram
-    of error values (6a). Optionally, a twin phylogeny (1b) can be generated from a true
-    phylogeny (1a), which then results in an error distribution (6b) with similar intermediate steps.
+    of error values (5a). Optionally, a twin phylogeny (1b) can be generated from the true phylogeny (1a), which then results in an error distribution (5b) with similar intermediate steps.
   }
   \label{fig:pipeline}
+  \giovanni{add reference to the generative prior in the figure}
 \end{figure}
@@
     \item the alignment is then used as BEAST2 input to infer a posterior distribution of phylogenies;
-    \item posterior phylogenies are compared with the given original phylogeny to estimate the error made, according to the error parameters specified by the user;
+    \item posterior phylogenies are compared with the given original phylogeny to estimate the error made, according to the error measure specified by the user;
 \end{enumerate}
@@
 
-The fourth step compares the true tree to each posterior tree, calculating the
-error using the nLTT statistic (\cite{janzen2015approximate}) 
-(Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a), resulting
-in an error distribution (Fig.~\ref{fig:pipeline}, 5a $\rightarrow$ 6a).
-The nLTT statistic is used by default, but also a user-defined error statistic can be used. 
+The fourth step compares the true tree to each posterior tree, resulting in an error distribution (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
+The nLTT statistic (\cite{janzen2015approximate}) is used by default, but also a user-defined error statistic can be used. 
 As an example, \verb;pirouette; supplies one custom error statistic,
@@
 
-\giovanni{
-  I think in this case we should indeed use a Yule tree prior [...]. What do you think?}
-\richel{
-  Agrees. I was wrong, possibly due to carnaval. Yule is the way to go!
-}
 We start from a tree simulated under a Yule (pure-birth) generative prior. 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{check numbers}
-\giovanni{
-  I think that the current version of the discussion is too didascalic. 
-  Most of what is said can be effectively summarized by table 5. 
-  We can even add ESSes in it. 
-}
-\richel{
-  Agreed, I disliked it as well
-}
-\giovanni{
-  Another point is that we should understand what is 
-  the message we want to deliver in this discussion. 
-  Since we are only using one tree we are not in the position to claim anything, 
-  because a single tree is not an effective sample. 
-  I saw that in some occasions you refer to the RQ as examples. 
-  This might be the right way to proceed, 
-  as our purpose is just to illustrate how to run all the possible experiments 
-  on a full tree distribution. 
-  I think we should make it clearer throughout the entire manuscript. 
-  I am also interested in hearing Rampal's opinion on this.
-}
-
-\giovanni{
-  This is a first draft on how I think the discussion should be setup. 
-  In this context there is no much need for the results in the table. 
-  Let me know what you think about it.
-}
-\richel{
-  I think that was the right way forward. 
-  I do think the reader would like help interpreting the
-  results of the example research questions as if these were
-  proper experiments. I added that
-}
 

### 2019-03-07 article.tex
Describe twinning better

@@
 that will be analyzed in the same way as the true tree.
-The twin tree has the same topology as the given phylogeny, 
+The twin tree will be the true tree 
+converted to follow the tree prior of the generative model. 
+We do so, by finding the parameters 
+(e.g. speciation and extinction rate for a birth-death model) 
+with the maximum likelihood, 
+conditioned on having the same number of tips as the original tree.
+We use these parameters to simulate the branching times of the twin tree, 
+which we combine with the original tree's topology.
+
+Using a twin tree serves as a control: both the given/true tree
+and twin tree will produce an error distribution.
+For the true tree, this error distribution is caused by
+the shape of the phylogeny (which is different from the tree priors used), 
+the stochastic alignment simulation and stochastic Bayesian inference.  
+The twin tree is created from the given phylogeny,
+as it has the same topology, 
 yet its branching times are simulated according to 
-a birth-death tree model.
-The goal of this procedure is to generate a tree
-accordingly to the tree prior used in the generative model, 
-that is used as a control for the main experiment.
-To produce such a birth-death tree, we maximize a birth-death likelihood [\cite{nee1994reconstructed}], conditioned on having the same number of tips as the original tree, to infer its parameters.
-We use these parameters, which are the per-species speciation and extinction rates, to simulate the branching times for the twin tree, which we combine with the original tree's topology.
+the (implemented) generative tree prior.
+Due to this, this error distribution for the twin tree is caused 
+only by the stochastic alignment simulation and stochastic Bayesian inference.
+The difference between these two error distribution is the error caused
+by the new/unimplemented tree prior.
 

### 2019-03-07 article.tex
Use 'ht' float placement, as indicated by Overleaf

@@
 \verb;pirouette; can be installed easily from CRAN:
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 install.packages("pirouette")
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 usethis::install_github("richelbilderbeek/pirouette")
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 library(pirouette)
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 install_beast2()
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 install_beast2_pkg("NS")
@@
 We start from a tree simulated under a Yule (pure-birth) generative prior. 
-We choose to use a tree with five taxa and a crown age of ten.
+We choose to use a tree with six taxa and a crown age of ten.
 
 \begin{lstlisting}[
-    language=R, floatplacement=H, frame=single, 
+    language=R, floatplacement=ht, frame=single, 
     label = {lst:create_yule_tree}, 
@@
   ]
-phylogeny <- create_yule_tree(n_taxa = 5, crown_age = 10)
+phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
 \end{lstlisting}
@@
     language=R,
-    floatplacement=H, frame=single,
+    floatplacement=ht, frame=single,
     label = {lst:create_alignment}, 
@@
   language=R, 
-  floatplacement=H, frame=single,
+  floatplacement=ht, frame=single,
   label = {lst:create_generative_experiment},
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 pir_params <- create_pir_params(
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 errors <- pir_run(
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 pir_plot(errors)
@@
 Here we start with a tree generated from an unknown 
-diversification model, that has four taxa and a crown age of five:
+diversification model, that has six taxa and a crown age of ten:
 
@@
   language=R, 
-  floatplacement=H, 
+  floatplacement=ht, 
   frame=single, 
@@
 ]
-phylogeny <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4):1);")
+phylogeny  <- ape::read.tree(
+  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
+)
 \end{lstlisting}
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 experiments <- create_all_experiments()
@@
     picking the best inference model.
-    The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error.
-    The boxplot shows the minimum, first quartile, median, third 
-    quartile and maximum error, of which the values are displayed 
-    in the adjacent label
+    The x-axis specifies the tree type (which can be 'true' or 'twin'), 
+    the y-axis shows the error
   }
@@
 the same as in the previous section, which are
-the phylogeny (listing \ref{lst:unknown_phylogeny}), 
+the phylogeny from an unknown speciation model (listing \ref{lst:unknown_phylogeny}), 
 the alignment parameters (listing \ref{lst:create_alignment}) 
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 twinning_params <- create_twinning_params()
@@
 
-"What is the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error?"
-
-Here, same as example 2, we start with a tree generated by the Yule process, by using the code shown in listing \ref{lst:create_yule_tree}.
+"What is the error made by BEAST2 from a phylogeny, 
+when the generative model is known, 
+compared to candidates with similar inference models, 
+in relationship to the background error?"
+\richel{
+  I think this is a nonsense research question and example:
+  we should use a tree of unknown speciation model instead.
+  Only then does twinning make sense
+}
+
+Here, same as example 2, we start with a tree generated by the Yule process, 
+by using the code shown in listing \ref{lst:create_yule_tree}.
 The first step in \verb;pirouette; is to simulate a DNA alignment 
@@
 In the second step we state our experiments. 
-In this example, we must both specify the generative and candidate experiments. Here we specify the generative model:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+In this example, we must both specify the generative and candidate experiments. 
+Here we specify the generative model:
+
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 generative_experiment <- create_experiment(
@@
 
-Specifying all candidates can be done simply by calling \verb;create_all_experiments;. In this case, all candidates will use the same site and clock model, and only differ in their tree priors:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+Specifying all candidates can be done simply by 
+calling \verb;create_all_experiments;. 
+In this case, all candidates will use the same site and clock model, 
+and only differ in their tree priors:
+
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 candidate_experiments <- create_all_experiments(
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 experiments <- c(list(generative_experiment), candidate_experiments)

### 2019-03-07 article.tex
Use 'ht' float placement, as indicated by Overleaf
Added 4th example, intermediate between 3rd and 5th

@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \includegraphics[width=\textwidth]{figure_bd.png}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \includegraphics[width=\textwidth]{figure_example_1.png}
@@
 \end{lstlisting}
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{tree_unknown.png}
+  \caption{The tree derived from an unknown diversification process, 
+    as created by listing \ref{lst:unknown_phylogeny}
+  }
+\end{figure}
 
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[
+  language=R, 
+  floatplacement=ht, 
+  frame=single, 
+  label = {lst:create_all_experiments},
+  caption = Create all 40 candidate experiments
+]
 experiments <- create_all_experiments()
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \includegraphics[width=\textwidth]{figure_example_2.png}
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[
+  language=R, 
+  floatplacement=ht, 
+  frame=single,
+  label = {lst:create_twinning_params},
+  caption = Create the default twinning parameters
+]
 twinning_params <- create_twinning_params()
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \includegraphics[width=\textwidth]{figure_example_3.png}
@@
 
+"What is the error made by BEAST2 from a phylogeny 
+derived from an unknown speciation model,
+when selecting the best inference model, 
+in relationship to the background error?"
+\richel{
+  I inserted this example research question, 
+  as think this is natural intermediate between the 
+  previous and next/final example
+}
+
+Here, same as example 2 and 3, we start with the phylogeny 
+from an unknown speciation model (listing \ref{lst:unknown_phylogeny}), 
+The first step in \verb;pirouette; is to simulate a DNA alignment 
+from the given phylogeny, for which we will use the same code 
+as shown in listing \ref{lst:create_alignment}.
+
+In the second step we specify we want all 
+40 inference models to compete, by 
+calling \verb;create_all_experiments;,
+as done in listing \ref{lst:create_all_experiments}.
+
+We choose again to use sensible defaults for the third (the BEAST2 inference) 
+and fourth (measuring the error) steps, 
+as we previously did to answer the second research question.
+
+To get an idea of a baseline error made by BEAST2, 
+we also make use of the twinning option by using the
+default created twinning parameters, as shown in 
+listing \ref{lst:create_twinning_params}.
+
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments,
+  twinning_params = twinning_params
+)
+\end{lstlisting}
+
+Running the experiment:
+
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+errors <- pir_run(
+  phylogeny,
+  pir_params = pir_params
+)
+\end{lstlisting}
+
+Again, showing the results:
+
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+pir_plot(errors)
+\end{lstlisting}
+
+The resulting figure is shown in figure \ref{fig:example_4}.
+
+\giovanni{describe winner}
+\richel{agreed, after the figures also show it :-) }
+
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{figure_example_4.png}
+  \caption{
+    Example 4: the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error.
+    The x-axis specifies the tree type, the y-axis shows the error.
+    The boxplot shows the minimum, first quartile, median, third 
+    quartile and maximum error, of which the values are displayed 
+    in the adjacent label
+  }
+  \label{fig:example_4}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Usage: Example Research Question 5}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A fifth research question that \verb;pirouette; can answer, is:
+
 "What is the error made by BEAST2 from a phylogeny, 
@@
 in relationship to the background error?"
-\richel{
-  I think this is a nonsense research question and example:
-  we should use a tree of unknown speciation model instead.
-  Only then does twinning make sense
-}
-
-Here, same as example 2, we start with a tree generated by the Yule process, 
+
+In this example, we use a tree generated by the Yule process.
+Instead of using the common Yule tree as in example 2, we'll use
+a Yule tree of low (actually: zero) likelihood: 
+the one as used in examples 3 and 4.
 by using the code shown in listing \ref{lst:create_yule_tree}.
@@
   }
-  \label{tab:experiment_4}
+  \label{tab:experiment_5}
 \end{table}
@@
 In the second step we state our experiments. 
-In this example, we must both specify the generative and candidate experiments. 
-Here we specify the generative model:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+In this example, we must both specify the generative and candidate experiments. Here we specify the generative model:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 generative_experiment <- create_experiment(
@@
 
-Specifying all candidates can be done simply by 
-calling \verb;create_all_experiments;. 
-In this case, all candidates will use the same site and clock model, 
-and only differ in their tree priors:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+Specifying all candidates can be done simply by calling \verb;create_all_experiments;. In this case, all candidates will use the same site and clock model, and only differ in their tree priors:
+
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 candidate_experiments <- create_all_experiments(
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 experiments <- c(list(generative_experiment), candidate_experiments)
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 pir_params <- create_pir_params(
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 errors <- pir_run(
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[language=R, floatplacement=H, frame=single]
 pir_plot(errors)
@@
 
-The resulting figure is shown in figure \ref{fig:example_4}.
+The resulting figure is shown in figure \ref{fig:example_5}.
 
@@
 \begin{figure}[h]
-  \includegraphics[width=\textwidth]{figure_example_4.png}
-  \caption{
-    Example 4: the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error.
+  \includegraphics[width=\textwidth]{figure_example_5.png}
+  \caption{
+    Example 5: the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error.
     The x-axis specifies the tree type, the y-axis shows the error.
@@
   }
-  \label{fig:example_4}
+  \label{fig:example_5}
 \end{figure}

### 2019-03-07 scripts/example_1.R

@@ -7,7 +7,7 @@ library(ggthemes)
 
 set.seed(314)
 
-phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
+phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),

### 2019-03-07 scripts/example_2.R

@@ -4,24 +4,20 @@
 library(pirouette)
 library(ggplot2)
 library(ggthemes)
+library(ggtree)
 
 set.seed(314)
 
-phylogeny  <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4) :1);")
+phylogeny  <- ape::read.tree(
+  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
+)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),
   mutation_rate = 0.1
 )
 
-experiments <- create_all_experiments(
-  tree_priors = list(
-    create_bd_tree_prior(),
-    create_ccp_tree_prior(),
-    create_cep_tree_prior(),
-    create_yule_tree_prior()
-  )
-)
+experiments <- create_all_experiments()
 
 pir_params <- create_pir_params(
   alignment_params = alignment_params,
@@ -69,4 +65,5 @@ xtable::xtable(
 )
 sink()
 
-
+ggtree::ggtree(phylogeny) + theme_tree2() + geom_tiplab() +
+  ggsave("/home/richel/GitHubs/pirouette_article/tree_unknown.png")

### 2019-03-07 scripts/example_3.R

@@ -7,7 +7,7 @@ library(ggthemes)
 
 set.seed(314)
 
-phylogeny <- create_stunning_tree(n_taxa = 5, t_0 = 10)
+phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),

### 2019-03-07 scripts/example_4.R

@@ -7,7 +7,8 @@ library(ggthemes)
 
 set.seed(314)
 
-phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
+phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
+# phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),

### 2019-03-07 article.tex
Example research question 1 peeks ahead at twinnning

@@
 
-"What is the error made by BEAST2 from a phylogeny using the same diversification model as it was generated by?"
+"What is the error made by BEAST2 from a phylogeny 
+using the same diversification model as it was generated by?"
 
@@
     label = {lst:create_yule_tree}, 
-    caption = {Create a Yule tree}
+    caption = {
+      Create a Yule tree. 
+      The resulting tree is shown in figure \ref{fig:yule_tree}
+    }
   ]
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{figure_bd.png}
+  \includegraphics[width=\textwidth]{tree_yule.png}
   \caption{The Yule tree, as created by listing \ref{lst:create_yule_tree}}
+  \label{fig:yule_tree} 
 \end{figure}
@@
 \end{figure}
+
+The error distribution obtained in this way can serve as a control:
+imagine one develops a novel tree prior related to the Yule model.
+One can predict that BEAST2 will be worse in recovering the phylogenies
+from that novel tree prior, as only the existing tree priors are used
+in inference. 
+To determine how much worse the recovery is, one needs to know the 
+inference error made by existing tree priors. 
+As this control is important, it is an inherent part of \verb;pirouette;:
+it is called twinning, which is shown in example research question 3.
 

### 2019-03-07 scripts/example_1.R

@@ -4,10 +4,16 @@
 library(pirouette)
 library(ggplot2)
 library(ggthemes)
+library(ggtree)
 
 set.seed(314)
 
-phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
+# Should run one day, https://github.com/richelbilderbeek/pirouette/issues/192
+#phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
+phylogeny  <- ape::read.tree(
+  text = "(((((A:2, B:2):2, C:4):2, D:6):2, E:8):2, F:10);"
+)
+
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),
@@ -26,9 +32,9 @@ errors <- pir_run(
   pir_params = pir_params
 )
 
+
 pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
-  theme_wsj() +
   ggsave("/home/richel/GitHubs/pirouette_article/figure_example_1.png")
 
 testit::assert(pir_params$experiments[[1]]$inference_model$mcmc$store_every != -1)
@@ -39,3 +45,6 @@ esses <- tracerer::calc_esses(
 sink("/home/richel/GitHubs/pirouette_article/example_1_esses.latex")
 xtable::xtable(esses, caption = "ESSes of example 1", label = "tab:esses_example_1", digits = 0)
 sink()
+
+ggtree::ggtree(phylogeny) + theme_tree2() + geom_tiplab() + ggtitle("Faked Yule tree, #192") +
+  ggsave("/home/richel/GitHubs/pirouette_article/tree_yule.png")

### 2019-03-07 scripts/example_3.R

@@ -4,6 +4,7 @@
 library(pirouette)
 library(ggplot2)
 library(ggthemes)
+library(ggtree)
 
 set.seed(314)
 

### 2019-03-07 article.tex
Add TikZ picture, check if Overleaf accepts that

@@
 \usepackage{longtable}
+
+\usepackage{tkz-graph}
+\usetikzlibrary{arrows,automata}
+\usetikzlibrary{calc}
+\usetikzlibrary{arrows.meta}
 
@@
     when the generative and inference model are the same.
-    The x-axis specifies the tree type (which can be 'true' or 'twin'), 
-    the y-axis shows the error. 
-    The boxplot shows the minimum, first quartile, median, third 
-    quartile and maximum error, of which the values are displayed 
-    in the adjacent label
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Errors}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}[h]
@@
 
-\richel{
-  I'd enjoy to put all intermediate output here: alignments, twin trees,
-  posteriors, nLTTs, errors. Now we only show the ones relevant in the text
-}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Example 1}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h]
+  \centering
+  \resizebox {0.8\columnwidth} {!} {
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm, semithick]   
+    \tikzstyle{every state}=[]
+    \node[state] (A) [rectangle] {
+      \includegraphics[width=0.2\textwidth]{tree_yule.png}
+    };   
+    \node[state] (B) [below of = A, rectangle] {
+      \includegraphics[width=0.2\textwidth]{tree_yule.png}
+    };   
+    \node[state] (C) [below of = B, rectangle] {
+      \includegraphics[width=0.2\textwidth]{tree_yule.png}
+    };   
+    \path 
+      (A) edge [anchor = south] node {} (B)
+      (B) edge [anchor = south] node {} (C)
+    ; 
+    \end{tikzpicture}
+  }
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Example 2}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \input{example_2_esses.latex}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Example 3}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \input{example_3_esses.latex}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Example 4}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \input{example_4_esses.latex}

### 2019-03-07 article.tex
Example 1 complete

@@
   \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm, semithick]   
+    \begin{tikzpicture}[
+      ->,>=stealth',shorten >=1pt,auto,
+      node distance=0.25\textheight, 
+      semithick
+    ]   
     \tikzstyle{every state}=[]
-    \node[state] (A) [rectangle] {
-      \includegraphics[width=0.2\textwidth]{tree_yule.png}
+    \node[state, draw=none] (O) [] {
+    };   
+    \node[state] (A) [right of = O, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_1_true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[width=0.2\textwidth]{tree_yule.png}
+      \includegraphics[height=0.2\textheight]{example_1_true_alignment.png}
     };   
     \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[width=0.2\textwidth]{tree_yule.png}
+      \includegraphics[height=0.2\textheight]{example_1_true_posterior_gen.png}
+    };   
+    \node[state] (D) [below of = C, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_1_error_violin.png}
     };   
     \path 
+      (O) edge [anchor = south] node {} (A)
       (A) edge [anchor = south] node {} (B)
       (B) edge [anchor = south] node {} (C)
+      (C) edge [anchor = south] node {} (D)
     ; 

### 2019-03-07 scripts/example_1.R

@@ -6,6 +6,8 @@ library(ggplot2)
 library(ggthemes)
 library(ggtree)
 
+root_folder <- path.expand("~/GitHubs/pirouette_article/")
+example_no <- 1
 set.seed(314)
 
 # Should run one day, https://github.com/richelbilderbeek/pirouette/issues/192
@@ -13,7 +15,6 @@ set.seed(314)
 phylogeny  <- ape::read.tree(
   text = "(((((A:2, B:2):2, C:4):2, D:6):2, E:8):2, F:10);"
 )
-
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),
@@ -35,7 +36,7 @@ errors <- pir_run(
 
 pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
-  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_1.png")
+  ggsave(file.path(root_folder, paste0("example_", example_no, "_errors.png")))
 
 testit::assert(pir_params$experiments[[1]]$inference_model$mcmc$store_every != -1)
 esses <- tracerer::calc_esses(
@@ -48,3 +49,128 @@ sink()
 
 ggtree::ggtree(phylogeny) + theme_tree2() + geom_tiplab() + ggtitle("Faked Yule tree, #192") +
   ggsave("/home/richel/GitHubs/pirouette_article/tree_yule.png")
+
+# Appendix figure
+################################################################################
+# trees
+################################################################################
+
+# True tree
+ggtree::ggtree(phylogeny) + theme_tree2() + geom_tiplab() +
+  ggtitle(paste0("Example ", example_no, "'s faked Yule tree, #192")) +
+  ggsave(file.path(root_folder, paste0("example_", example_no, "_true_tree.png")))
+
+# Twin tree
+if (!is.na(pir_params$twinning_params)) {
+  ggtree::ggtree(ape::read.tree(pir_params$twinning_params$twin_tree_filename)) + theme_tree2() + geom_tiplab() +
+    ggtitle(paste0("Example ", example_no, "'s faked Yule tree, #192")) +
+    ggsave(file.path(root_folder, paste0("example_", example_no, "_twin_tree.png")))
+}
+
+################################################################################
+# alignment
+################################################################################
+
+
+png(
[…]

### 2019-03-07 scripts/example_2.R

@@ -6,6 +6,8 @@ library(ggplot2)
 library(ggthemes)
 library(ggtree)
 
+root_folder <- path.expand("~/GitHubs/pirouette_article/")
+example_no <- 2
 set.seed(314)
 
 phylogeny  <- ape::read.tree(

### 2019-03-07 scripts/example_3.R

@@ -6,6 +6,8 @@ library(ggplot2)
 library(ggthemes)
 library(ggtree)
 
+root_folder <- path.expand("~/GitHubs/pirouette_article/")
+example_no <- 3
 set.seed(314)
 
 phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")

### 2019-03-07 scripts/example_4.R

@@ -6,6 +6,8 @@ library(ggplot2)
 library(ggthemes)
 library(ggtree)
 
+root_folder <- path.expand("~/GitHubs/pirouette_article/")
+example_no <- 1
 set.seed(314)
 
 phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")

### 2019-03-07 scripts/example_2.R
Create all example intermediate figures in the same way

@@ -69,3 +69,128 @@ sink()
 
 ggtree::ggtree(phylogeny) + theme_tree2() + geom_tiplab() +
   ggsave("/home/richel/GitHubs/pirouette_article/tree_unknown.png")
+
+# Appendix figure
+################################################################################
+# trees
+################################################################################
+
+# True tree
+ggtree::ggtree(phylogeny) + theme_tree2() + geom_tiplab() +
+  ggtitle(paste0("Example ", example_no, "'s faked Yule tree, #192")) +
+  ggsave(file.path(root_folder, paste0("example_", example_no, "_true_tree.png")))
+
+# Twin tree
+if (!is.na(pir_params$twinning_params)) {
+  ggtree::ggtree(ape::read.tree(pir_params$twinning_params$twin_tree_filename)) + theme_tree2() + geom_tiplab() +
+    ggtitle(paste0("Example ", example_no, "'s faked Yule tree, #192")) +
+    ggsave(file.path(root_folder, paste0("example_", example_no, "_twin_tree.png")))
+}
+
+################################################################################
+# alignment
+################################################################################
+
+
+png(
+  filename = file.path(root_folder, paste0("example_", example_no, "_true_alignment.png")),
+  width = 800,
+  height = 300
+)
+ape::image.DNAbin(
+  ape::read.FASTA(file = pir_params$alignment_params$fasta_filename),
+  grid = TRUE,
+  show.bases = FALSE,
+  legend = FALSE,
+  cex.lab = 2.0,
+  cex.axis = 2.0
+)
+dev.off()
+
+if (!is.na(pir_params$twinning_params)) {
+  png(
+    filename = file.path(root_folder, paste0("example_", example_no, "_twin_alignment.png")),
+    width = 800,
+    height = 300
+  )
+  ape::image.DNAbin(
+    ape::read.FASTA(file = pir_params$twinning_params$twin_alignment_filename),
+    grid = TRUE,
+    show.bases = FALSE,
+    legend = FALSE,
+    cex.lab = 2.0,
+    cex.axis = 2.0
+  )
+  dev.off()
+}
+################################################################################
+# posteriors
+################################################################################
+
+png(
+  filename = file.path(root_folder, paste0("example_", example_no, "_true_posterior_gen.png")),
[…]

### 2019-03-07 scripts/example_1.R
Can run example 1 on Peregrine

@@ -36,6 +36,7 @@ pir_params$experiments[[1]]$beast2_options$input_filename <- file.path(root_fold
 pir_params$experiments[[1]]$beast2_options$output_log_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.log"))
 pir_params$experiments[[1]]$beast2_options$output_trees_filenames <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.trees"))
 pir_params$experiments[[1]]$beast2_options$output_state_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.xml.state"))
+pir_params$experiments[[1]]$errors_filename <- file.path(root_folder, paste0("example_", example_no, "_error.csv"))
 
 errors <- pir_run(
   phylogeny,

### 2019-03-07 scripts/example_2.R

@@ -33,7 +33,7 @@ errors <- pir_run(
 
 pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
-  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_2.png")
+  ggsave(file.path(root_folder, paste0("example_", example_no, "_errors.png")))
 
 
 testit::assert(pir_params$experiments[[1]]$inference_model$mcmc$store_every != -1)

### 2019-03-07 scripts/example_3.R

@@ -34,7 +34,7 @@ errors <- pir_run(
 
 pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
-  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_3.png")
+  ggsave(file.path(root_folder, paste0("example_", example_no, "_errors.png")))
 
 testit::assert(pir_params$experiments[[1]]$inference_model$mcmc$store_every != -1)
 esses <- tracerer::calc_esses(

### 2019-03-07 scripts/example_4.R

@@ -50,7 +50,7 @@ errors <- pir_run(
 
 pir_plot(errors) +
   scale_y_continuous(breaks = seq(0.0, 0.11, by = 0.01), limits = c(0, 0.11)) +
-  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_4.png")
+  ggsave(file.path(root_folder, paste0("example_", example_no, "_errors.png")))
 
 
 

### 2019-03-07 scripts/example_2.R
Prepare for Peregrine

@@ -26,6 +26,19 @@ pir_params <- create_pir_params(
   experiments = experiments
 )
 
+################################################################################
+# Settings to run on Peregrine cluster
+################################################################################
+pir_params$alignment_params$fasta_filename <- file.path(root_folder, paste0("example_", example_no, "_true.fasta"))
+for (experiment in pir_params$experiments) {
+  experiment$beast2_options$input_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_input.xml"))
+  experiment$beast2_options$output_log_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.log"))
+  experiment$beast2_options$output_trees_filenames <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.trees"))
+  experiment$beast2_options$output_state_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.xml.state"))
+  experiment$errors_filename <- file.path(root_folder, paste0("example_", example_no, "_error.csv"))
+}
+################################################################################
+
 errors <- pir_run(
   phylogeny,
   pir_params = pir_params

### 2019-03-11 article.tex
Can run example 1 on Peregrine
Use new figures
Changes made by @Giappo

@@
+\RequirePackage{rotating}
 \documentclass{article}
@@
 \usepackage{longtable}
+\usepackage{rotating}
+\usepackage{mathtools}
 
@@
     the prior probability are possible in practice. 
-    This begs the question how accurate the tree estimation is 
+    This begs the question of how accurate the tree estimation is 
     when the real macroevolutionary processes are different 
@@
 or a combination of all these [\cite{herrera2018detecting}].
-Note all of these diversification models are available in BEAST2, 
+Not all of these diversification models are available in BEAST2, 
 but the option is available.
@@
 according to how they operate in the pipeline. 
-We summarize them in Tables~\ref{tab:definitions} and \ref{tab:options}.
+We summarize them in Tables~\ref{tab:options} and \ref{tab:definitions}.
 
 Although many possible tests can be performed, 
-we present here only four example research questions (ERQ) that \verb;pirouette; can answer. 
+we present here only four \giovanni{five?} example research questions (ERQ) that \verb;pirouette; can answer. 
 For these ERQs we will estimate the error made by BEAST2 when 
@@
 \begin{sidewaystable}
-%\begin{table}
 \centering
-  %\begin{tabular}{| c | c | l | l |}
-  \begin{tabular}{|@{}c|p{5.5cm}|p{2.85cm}@{}|}
-    \hline
-    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Possible values} \\ 
-    \hline
-    $\mathit{tp}$ & 'tree prior', i.e a macroevolutionary speciation model & Yule, BD, CBS, CCP, CEP \\
-    $\mathit{cm}$ & 'clock model' for the DNA mutation rates & strict, relaxed log-normal \\
-    $\mathit{sm}$ & 'site model' for nucleotide substitution & JC, HKY, TN, GTR \\
-    $\mathit{mr}$ & 'mutation rate' & \\
-    $\mathit{rs}$ & 'root sequence' of nucleotides at the root of the tree & \\
-    $\mathit{mp}$ & 'MRCA prior' knowledge of a most recent common ancestor & \\
-    $\mathit{ms}$ & 'MCMC settings' i.e. Markov Chain Monte Carlo algorithm setup & \\
-    $\mathit{et}$ & 'experiment type' & Generative, candidate \\
-    $\mathit{ri}$ & 'run if' & Always, Best candidate\\
-    $\mathit{ef}$ & 'error function' that specifies how to measure the error & nLTT, $|\gamma|$ \\
-    $\mathit{tm}$ & 'twin model' & Yule, BD \\
+  \begin{tabular}{|@{}c|p{2.5cm}|p{9cm}|p{4.5cm}@{}|}
+    \hline
+    \centering
+    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} & \textbf{Possible values} \\ 
+    \hline
+    $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model & Yule, BD, CBS, CCP, CEP \\
+    $\mathit{c_{m}}$ & clock model & Clock for the DNA mutation rates & strict, relaxed log-normal \\
+    $\mathit{s_{m}}$ & site model & Nucleotide substitution model & JC, HKY, TN93, GTR \\
+    $\mathit{m_{r}}$ & mutation rate & Pace at which mutation occurs & $m_{r} \in \mathbb{R}_{>0}$\\
+    $\mathit{r_{s}}$ & root sequence & DNA sequence at the root of the tree & any combination of a, c, g, t \\
+    $\mathit{m_{p}}$ & MRCA prior & Prior knowledge of a most recent common ancestor & see documentation \\
+    $\mathit{m_{s}}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup & see documentation \\
+    $\mathit{m_{t}}$ & model type & Criterion to select an inference model & Generative, Candidate \\
+    $\mathit{r_{i}}$ & run if & Condition under which an inference model is used & Always, Best candidate\\
+    $\mathit{e_{f}}$ & error function & Specifies how to measure the error & nLTT, $|\gamma|$ \\
+    $\mathit{t_{m}}$ & twin model & Sets the tree prior for the twin tree & Yule, BD \\
     \hline
@@
     Bayesian Skyline, CCP = coalescent constant-population, CEP = coalescent
-    exponential-population, JC = Jukes-Cantor 1969, HKY = Hasegawa, Kishino and 
-    Yano 1985, TN = Tamura 1992, GTR = Generalised time-reversible 
-    model (Tavaré 1986)
+    exponential-population, JC = Jukes-Cantor 1969 (\cite{jukes1969evolution}), HKY = Hasegawa, Kishino and 
+    Yano (\cite{hasegawa1985dating}), TN93 = \cite{tamura1993estimation}, GTR = Generalised time-reversible 
+    model (\cite{tavare1986some}).
   }
   \label{tab:options}
-%\end{table}
-\end{sidewaystable}
-
-\begin{sidewaystable}
-%\begin{table}
+\bigskip
+
+  \begin{tabular}{|@{}c|p{4cm}|p{9cm}|p{3cm}@{}|}
+    \hline
     \centering
-  \begin{tabular}{|@{}c|p{6cm}|p{2.85cm}@{}|}
-    \hline
-    \textbf{Symbol} & \textbf{Macro-argument} & \textbf{Composed by}\\ 
-    \hline
-    $\mathit{P}$ & 'Generative prior' under which the 'true' tree is simulated & $\mathit{tp}$ \\
-    $\mathit{A}$ & 'Alignment model' used for simulation & $\mathit{cm}, \mathit{sm}, \mathit{mr}, \mathit{rs}$ \\
-    $\mathit{G}$ & 'Generative model' & $\mathit{P},\mathit{A}$ \\
-    $\mathit{I}$ & 'Inference model' to run BEAST2 & $\mathit{cm},\mathit{sm},\mathit{tp},\mathit{mp},\mathit{ms}$ \\
-    $\mathit{C}$ & 'Inference conditions' under which $\mathit{I}$ is used in the inference & $\mathit{et}, \mathit{ri}$\\
-    $\mathit{X}$ & 'Experiment' & $\mathit{I},\mathit{C}$ \\
-    $\mathit{E}$ & 'Error measurement parameters' & $\mathit{ef}$\\
-    $\mathit{T}$ & 'Twinning parameters' to set the twin creation procedure & $\mathit{tm}$\\
+    \textbf{Symbol} & \textbf{Macro-argument} & \textbf{Description} & \textbf{Includes} \\
+    \hline
+    $\mathit{G}$ & Generative prior & Diversification model to generate the 'true' tree & $\mathit{t_{p}}$ \\
+    $\mathit{A}$ & Alignment model & Model used to simulate the alignment ($\mathit{c_{m}}$ must be strict) & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{m_{r}}, \mathit{r_{s}}$ \\
+    $\mathit{I}$ & Inference model & Phylogenetic inference model to run BEAST2 & $\mathit{c_{m}},\mathit{s_{m}},\mathit{t_{p}},\mathit{m_{p}},\mathit{m_{s}}$ \\
+    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference & $\mathit{m_{t}}, \mathit{r_{i}}$\\
+    $\mathit{X}$ & Experiment & Full setting for the BEAST2 run & $\mathit{I},\mathit{C}$ \\
+    $\mathit{E}$ & Error measure parameters & Errors measurement setup & $\mathit{e_{f}}$\\
+    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & $\mathit{t_{m}}$\\
     \hline 
@@
   \caption{
-    Definitions of terms and relative symbols used in this article.
+    Definitions of terms and relative symbols used in the main text and in Fig~\ref{fig:pipeline}.
   }
   \label{tab:definitions}
-%\end{table}
+  \giovanni{Should we rename "Twinning Parameters" to "Twinning model" and "twin model" to "twin prior"?}
 \end{sidewaystable}
@@
     $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
-    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in inference \\
+    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference \\
     $\mathit{X}$ & Experiment & Combination of $(\mathit{I},\mathit{C})$\\
@@
     \hline
-    $\mathit{P}$ & Generative prior & Diversification model the 'true' tree is simulated with \\
+    $\mathit{G}$ & Generative prior & Diversification model the 'true' tree is simulated with \\
     $\mathit{A}$ & Alignment model & Alignment simulation model, $\mathit{A} \ni \{ \mathit{r}, \mathit{\mu}, \mathit{c}, \mathit{s} \}$, in which $\mathit{c}$ = strict \\
@@
   \label{fig:pipeline}
-  \giovanni{add reference to the generative prior in the figure}
+  \giovanni{add reference to the generative prior P (or G) in the figure}
+  \giovanni{I think that a list of possible models should replace the "experiment" cake in the middle, then C should be drawn on the arrow connecting the figure in the middle to 3a.
+"I" should replace the circled x of passage 3a.
+An additional C can be drawn alongside the arrow 2a -> 3a.
+After that a rectangle can be drawn around the whole middle section (including 3a, 3b and the middle C) to denote that this is phase "X".}
+  \giovanni{We might also need an arrow going from the central figure to the top arrow (the one with "T") to specify that the twin is generated according to the chosen inference model.}
 \end{figure}
@@
 it must use the same site and clock model used in the alignment simulation,
-as well as the tree prior that underlied the creation of the given phylogeny 
-\giovanni{
-  the tree prior appears not to be part of the alignment model. 
-  This means that should be part of the generative prior. 
-  Is it true? In case we can refer explicitly to it, 
-  according to our definitions in table \ref{tab:definitions}
-}
-\richel{
-  Indeed, a tree prior is not part of the alignment model.
-  Go ahead and show how you'd like to refer explicitly to the
-  table!
-}
-. 
+as well as the tree prior that underlied the creation of the given phylogeny. 
 When using experiments with candidate models,
@@
 
+\iffalse
+%version 1
 An optional step is to generate a 'twin tree' 
@@
 The twin tree will be the true tree 
-converted to follow the tree prior of the generative model. 
-We do so, by finding the parameters 
-(e.g. speciation and extinction rate for a birth-death model) 
-with the maximum likelihood, 
-conditioned on having the same number of tips as the original tree.
+converted to follow the tree prior of the chosen inference model.
+
+To do so we exploit the likelihood function associated with the chosen inference model.
+We find the parameters (e.g.
+speciation and extinction rate for a birth-death model) that maximize such likelihood applied to the true tree, conditioned on its number of tips.
+
 We use these parameters to simulate the branching times of the twin tree, 
@@
 by the new/unimplemented tree prior.
-
+\fi
+
+\iffalse
+%version 2 with no formulae
+An optional step is to generate a 'twin tree' 
+(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
+that will be analyzed in the same way as the true tree.
+We want to produce such tree 'converting' the original tree, created according to the generative prior $\mathit{G}$, into the chosen inference model $\mathit{I}$.
+
+To do so we exploit the likelihood function associated with the chosen inference model $\mathit{I}$.
+We find the parameters (e.g.
+speciation and extinction rate for a birth-death model) that maximize such likelihood applied to the true tree, conditioned on its number of tips.
+
+We use these parameters to simulate an equal amount of branching times for the twin tree simulated under the chosen inference model's process.
+We then combine these branching times with the original tree's topology.
+
+Using a twin tree serves as a control: both the given/true tree and twin tree will produce an error distribution.
+We assume that these errors are a natural product of the input tree and the intermediate steps of the pipeline.
+All the factors that can cause such errors are: the branching times of the tree (as a consequence of the tree prior), the topology of the tree, the stochasticity induced by simulating the alignment (step 1a $\rightarrow$ 2a in Fig.~\ref{fig:pipeline}) and the stochasticity induced by the Bayesian inference (step 2a $\rightarrow$ 4a in Fig.~\ref{fig:pipeline}).
+By imposing the same topology and replicating the pipeline a sufficient number of times we can sort out all the possible causes that are not a direct consequence of the choice of the tree prior.
+
+Potential differences in shapes of error distributions will therefore be consequences of different tree priors for the two parallel pipelines: the use of prior $\mathit{G}$ for the original pipeline and of prior $\mathit{I}$ for the twin pipeline.
+In the second case, though, such error distribution is not influenced by the mismatch between how the tree is generated and how it is analyzed by BEAST2.
+\fi
+
+\giovanni{This is a slightly more formal version.
+I personally find it clearer.
+If you don't like it we can use version 2 (see above in the source).
+In case you really like it, instead, I can try to make the rest a bit more similar to this.}
+
+%version 3, more formal
+An optional step is to generate a 'twin tree' $\tau_{I}$
+(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
+that will be analyzed in the same way as the true tree, following a parallel pipeline.
+We want to produce such tree 'converting' the original tree $\tau_{G}$, created according to the generative prior $\mathit{G}$, to be in accordance with the chosen inference model $\mathit{I}$:
+\begin{align}
+    \tau_{G} = (\Vec{t}_{G}, \psi_{G}) \xrightarrow[]{\mathit{T}} \tau_{I} = (\Vec{t}_{I}, \psi_{I}),
+\end{align}
+where $\Vec{t}$ are the tree branching times and $\psi$ is tree topology, which provides information on where new lineages branch out at each branching time.
+
+To execute the twinning process we exploit the likelihood function $L_{I}$ associated with the chosen inference model $\mathit{I}$.
+We find the parameters $\theta^{*}_{I}$ (e.g.
+speciation and extinction rates, in case of a birth-death model) that maximize $L_{I}$ applied to the true tree, conditioned on its number of tips $n_{G}$,
+\begin{align}
+    max[L_{I}(\theta_{I}|\tau_{G}, n_{G})] \rightarrow \theta^{*}_{I}.
+\end{align}
+We use $\theta^{*}_{I}$ to simulate an equal amount $n_{I} = n_{G}$ of branching times $\Vec{t}_{I}$ for the twin tree $\tau_{I}$, under the process $I$.
+We then impose that $\tau_{G}$ and $\tau_{I}$ must respect the same topology
+\begin{align}
+    \psi_{I} = \psi_{G}.
+\end{align}
+Using a twin tree serves as a control: both $\tau_{G}$ and $\tau_{I}$ will produce an error distribution.
+We assume that these errors are a natural product of the input tree and the intermediate steps of the pipeline.
+All the factors that can cause such errors are: the branching times $\vec{t}$ of the tree (as a consequence of the tree prior), its topology $\psi$, the stochasticity induced by simulating the alignment (step 1a $\rightarrow$ 2a in Fig.~\ref{fig:pipeline}) and the stochasticity induced by the Bayesian inference (step 2a $\rightarrow$ 4a in Fig.~\ref{fig:pipeline}).
+By imposing the same topology and replicating the pipelines a sufficient number of times we can sort out all the possible causes that are not a direct consequence of the choice of the tree prior.
+
+Potential differences in shape of error distributions will therefore be consequences of different tree priors for the two parallel pipelines: the use of prior $\mathit{G}$ for the original pipeline and of prior $\mathit{I}$ for the twin pipeline.
+In the second case, though, such error distribution is not influenced by the mismatch between how the tree is generated and how it is analyzed by BEAST2.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{tabular}
-\caption{pirouette's main functions}
+\caption{pirouette's main functions.}
 \label{tab:functions}
@@
   \includegraphics[width=\textwidth]{tree_yule.png}
-  \caption{The Yule tree, as created by listing \ref{lst:create_yule_tree}}
+  \caption{The Yule tree, as created by listing \ref{lst:create_yule_tree}.}
   \label{fig:yule_tree}
+  \giovanni{Remove the text in top-left corner.}
+  \giovanni{I'd use a sexier topology, purely for esthetical reasons.}
 \end{figure}
@@
 To do so, we must specify the DNA root sequence and a mutation rate. 
-In this example, the DNA root sequence consists out of four block of 250 nucleotides each, 
+In this example, the DNA root sequence consists in four blocks of 250 nucleotides each, 
 where the per-nucleotide mutation rate is 0.1 mutations per unit time.
@@
     label = {lst:create_alignment}, 
-    caption = {Create an alignment}
+    caption = {Create an alignment.}
   ]
@@
 
-By default, an alignment is created using a Jukes-Cantor (JC) site model
+By default, an alignment is created using a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
 and a strict clock model. 
@@
 a combination of JC site model, strict clock model and Yule tree prior), 
-will always be run and we skip to measure the evidence for our model.
+it will always be run and we skip the evidence measurement for our model.
 
@@
 results in:
+\giovanni{Would it be clearer not to use sensible defaults where we call the functions with arguments we refer to in the main text? For example here it could be clearer for the reader to see:
+create\_experiment(site\_model = create\_jc69\_site\_model(), clock\_model = create\_strict\_clock\_model(), tree\_prior = create\_yule\_tree\_prior())
+}
 
@@
   label = {lst:unknown_phylogeny},
-  caption = A phylogeny generated by an unknown diversification model
+  caption = A phylogeny generated by an unknown diversification model.
 ]
@@
   \caption{The tree derived from an unknown diversification process, 
-    as created by listing \ref{lst:unknown_phylogeny}
+    as created by listing \ref{lst:unknown_phylogeny}.
   }
@@
     quartile and maximum error, of which the values are displayed 
-    in the adjacent label
+    in the adjacent label.
   }
   \label{fig:example_3}
+  \giovanni{remove boxplots}
 \end{figure}
@@
     quartile and maximum error, of which the values are displayed 
-    in the adjacent label
+    in the adjacent label.
   }
   \label{fig:example_4}
+  \giovanni{Remove boxplots.}
 \end{figure}

### 2019-03-11 article.tex

@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{figure_example_1.png}
+  \includegraphics[width=\textwidth]{example_1_errors.png}
   \caption{
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{figure_example_2.png}
+  \includegraphics[width=\textwidth]{example_2_errors.png}
   \caption{
@@
     picking the best inference model.
-    The x-axis specifies the tree type (which can be 'true' or 'twin'), 
-    the y-axis shows the error
   }
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{figure_example_3.png}
+  \includegraphics[width=\textwidth]{example_3_errors.png}
   \caption{
@@
     Here, the twin column shows the error BEAST2 makes starting from the original tree versus the twin. 
-    The x-axis specifies the tree type (which can be 'true' or 'twin'), the y-axis shows the error.
-    The boxplot shows the minimum, first quartile, median, third 
-    quartile and maximum error, of which the values are displayed 
-    in the adjacent label.
   }
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{figure_example_4.png}
+  \includegraphics[width=\textwidth]{example_4_errors.png}
   \caption{
     Example 4: the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error.
-    The x-axis specifies the tree type, the y-axis shows the error.
-    The boxplot shows the minimum, first quartile, median, third 
-    quartile and maximum error, of which the values are displayed 
-    in the adjacent label.
   }
@@
 \begin{figure}[h]
-  \includegraphics[width=\textwidth]{figure_example_5.png}
+  \includegraphics[width=\textwidth]{example_5_errors.png}
   \caption{
@@
   }
+  \label{fig:example_1_full_pipeline}
+  \caption{Example 1: full pipeline}
 \end{figure}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h]
+  \centering
+  \resizebox {0.8\columnwidth} {!} {
+    \begin{tikzpicture}[
+      ->,>=stealth',shorten >=1pt,auto,
+      node distance=0.25\textheight, 
+      semithick
+    ]   
+    \tikzstyle{every state}=[]
+    \node[state, draw=none] (O) [] {
+    };   
+    \node[state] (A) [right of = O, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_2_true_tree.png}
+    };   
+    \node[state] (B) [below of = A, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_2_true_alignment.png}
+    };   
+    \node[state] (C) [below of = B, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_2_true_posterior_gen.png}
+    };   
+    \node[state] (D) [below of = C, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_2_error_violin.png}
+    };   
+    \path 
+      (O) edge [anchor = south] node {} (A)
+      (A) edge [anchor = south] node {} (B)
+      (B) edge [anchor = south] node {} (C)
+      (C) edge [anchor = south] node {} (D)
+    ; 
+    \end{tikzpicture}
+  }
+  \label{fig:example_2_full_pipeline}
+  \caption{Example 2: full pipeline}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \input{example_2_esses.latex}

### 2019-03-11 article.tex
Add first half of example 3's pipeline

@@
 Although many possible tests can be performed, 
-we present here only four \giovanni{five?} example research questions (ERQ) that \verb;pirouette; can answer. 
+we present here only five example research questions (ERQ) that \verb;pirouette; can answer. 
 For these ERQs we will estimate the error made by BEAST2 when 
@@
   \label{tab:definitions}
-  \giovanni{Should we rename "Twinning Parameters" to "Twinning model" and "twin model" to "twin prior"?}
+  \giovanni{
+    Should we rename "Twinning Parameters" to "Twinning model"?
+  }
+  \richel{
+    As the twinning parameters contain more than only the twin model's tree prior,
+    I suggest to keep it 'twinning parameters'
+  }
+  \giovanni{
+    Should we rename "twin model" to "twin prior"?
+  }
+  \richel{
+    I think that's a great idea! I would prefer the name 'twin tree prior' even
+    better
+  }
 \end{sidewaystable}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[h]
+  \centering
+  \resizebox {0.8\columnwidth} {!} {
+    \begin{tikzpicture}[
+      ->,>=stealth',shorten >=1pt,auto,
+      node distance=0.25\textheight, 
+      semithick
+    ]   
+    \tikzstyle{every state}=[]
+    \node[state, draw=none] (O) [] {
+    };   
+    \node[state] (A) [right of = O, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_3_true_tree.png}
+    };   
+    \node[state] (B) [below of = A, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_3_true_alignment.png}
+    };   
+    \node[state] (C) [below of = B, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_3_true_posterior_gen.png}
+    };   
+    \node[state] (D) [below of = C, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_3_error_violin.png}
+    };   
+    \path 
+      (O) edge [anchor = south] node {} (A)
+      (A) edge [anchor = south] node {} (B)
+      (B) edge [anchor = south] node {} (C)
+      (C) edge [anchor = south] node {} (D)
+    ; 
+    \end{tikzpicture}
+  }
+  \label{fig:example_3_full_pipeline}
+  \caption{Example 3: full pipeline}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \input{example_3_esses.latex}

### 2019-03-11 article.tex
Pipeline example 3 done

@@
       ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.25\textheight, 
+      node distance=0.4\textheight, 
       semithick
@@
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_true_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_3_true_alignment.png}
     };   
@@
       \includegraphics[height=0.2\textheight]{example_3_error_violin.png}
+    };   
+    \node[state] (AT) [right of = A, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_3_twin_tree.png}
+    };   
+    \node[state] (BT) [below of = AT, rectangle] {
+      \includegraphics[height=0.13\textheight]{example_3_twin_alignment.png}
+    };   
+    \node[state] (CT) [below of = BT, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_3_twin_posterior_gen.png}
+    };   
+    \node[state] (DT) [below of = CT, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_3_twin_error_violin.png}
     };   
@@
       (C) edge [anchor = south] node {} (D)
+      (A) edge [anchor = east] node {} (AT)
+      (AT) edge [anchor = south] node {} (BT)
+      (BT) edge [anchor = south] node {} (CT)
+      (CT) edge [anchor = south] node {} (DT)
     ; 

### 2019-03-11 article.tex
Remove table with results

@@
   \label{tab:definitions}
-  \richel{I've some set notations in some spots, as I am unsure if you like it}
-  \giovanni{It would be nice to specify (in the first part of the table) 
-    the basic elements composing all the elements of the second part of the table. 
-    This could be probably done for "Alignment parameters", "Inference conditions" 
-    and "Twinning parameters".
-  }
-  \richel{
-    I feel the next table (which was also your suggestion), does so
-    at the right amount of detail
-  }
-  \giovanni{
-    How do we specify what constitutes 'twinning parameters'?
-  }
-  \richel{Perhaps you'd enjoy my definition?}
 \end{table}
@@
   \label{fig:pipeline}
-  \giovanni{add reference to the generative prior P (or G) in the figure}
-  \giovanni{I think that a list of possible models should replace the "experiment" cake in the middle, then C should be drawn on the arrow connecting the figure in the middle to 3a.
+  \giovanni{
+    add reference to the generative prior P (or G) in the figure
+  }
+  \giovanni{
+    I think that a list of possible models should replace the "experiment" 
+    cake in the middle, then C should be drawn on the arrow 
+    connecting the figure in the middle to 3a. 
     "I" should replace the circled x of passage 3a. 
     An additional C can be drawn alongside the arrow 2a -> 3a. 
-After that a rectangle can be drawn around the whole middle section (including 3a, 3b and the middle C) to denote that this is phase "X".}
+    After that a rectangle can be drawn around the whole middle section 
+    (including 3a, 3b and the middle C) to denote that this is phase "X".}
   \giovanni{We might also need an arrow going from the central figure to the top arrow (the one with "T") to specify that the twin is generated according to the chosen inference model.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[h]
-\centering
-\begin{tabular}{ | r | l | l | l | l | }
-\hline
-\multirow{2}{*}{\textbf{Example}} & \multicolumn{2}{c|}{\textbf{True tree}} 
-                                  & \multicolumn{2}{c|}{\textbf{Twin tree}} \\
-\cline{2-5}
-                                  & \textbf{median} & \textbf{range} & \textbf{median} & \textbf{range} \\
-\hline
-1  & $0.02$  & $0.0 - 0.27$  &         &               \\
-2  & $0.02$  & $0.0 - 0.08$  &         &               \\
-3  & $0.018$ & $0.0 - 0.1$   & $0.019$ & $0.0 - 0.11$  \\
-4C & $0.22$  & $0.0 - 0.088$ & $0.009$ & $0.0 - 0.056$ \\
-4G & $0.020$ & $0.0 - 0.075$ & $0.013$ & $0.0 - 0.052$ \\
-\hline
-\end{tabular}
-\caption{
-  \richel{redo}
-  Results of the examples. 
-  For the true and twin tree,
-  the median and range (1st and 3rd quartile) of the errors 
-  is shown. Experiment 4 is split in two rows
-  by either the candidate ('C') or generative model ('G')
-}
-\label{tab:results}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Example 1}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\iffalse
+\begin{figure}[h]
+  \centering
+  \resizebox {0.8\columnwidth} {!} {
+    \begin{tikzpicture}[
+      ->,>=stealth',shorten >=1pt,auto,
+      node distance=0.4\textheight, 
+      semithick
+    ]   
+    \tikzstyle{every state}=[]
+    \node[state, draw=none] (O) [] {
+    };   
+    \node[state] (A) [right of = O, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_true_tree.png}
+    };   
+    \node[state] (B) [below of = A, rectangle] {
+      \includegraphics[height=0.13\textheight]{example_4_true_alignment.png}
+    };   
+    \node[state] (CG) [below of = B, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_true_posterior_gen.png}
+    };   
+    \node[state] (DG) [below of = CG, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_error_violin_gen.png}
+    };   
+    \node[state] (CB) [below right of = B, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_true_posterior_best.png}
+    };   
+    \node[state] (DB) [below of = CB, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_error_violin_best.png}
+    };   
+    \node[state] (AT) [right of = A, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_twin_tree.png}
+    };   
+    \node[state] (BT) [below of = AT, rectangle] {
+      \includegraphics[height=0.13\textheight]{example_4_twin_alignment.png}
+    };   
+    \node[state] (CTG) [below of = BT, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_gen.png}
+    };   
+    \node[state] (DTG) [below of = CTG, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_gen.png}
+    };   
+    \node[state] (CTB) [below right of = BT, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_best.png}
+    };   
+    \node[state] (DTB) [below of = CTB, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_best.png}
+    };   
+    \path 
+      (O) edge [anchor = south] node {} (A)
+      (A) edge [anchor = south] node {} (B)
+      (B) edge [anchor = south] node {} (CG)
+      (CG) edge [anchor = south] node {} (DG)
+      (B) edge [anchor = southeast] node {} (CB)
+      (CB) edge [anchor = south] node {} (DB)
+      (A) edge [anchor = east] node {} (AT)
+      (AT) edge [anchor = south] node {} (BT)
+      (BT) edge [anchor = southeast] node {} (CTG)
+      (CTG) edge [anchor = south] node {} (DTG)
+      (BT) edge [anchor = south] node {} (CTB)
+      (CTB) edge [anchor = south] node {} (DTB)
+    ; 
+    \end{tikzpicture}
+  }
+  \label{fig:example_4_full_pipeline}
+  \caption{Example 4: full pipeline}
+\end{figure}
+\fi
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \input{example_4_esses.latex}

### 2019-03-11 article.tex
Process feedback

@@
   \label{fig:pipeline}
-  \giovanni{
-    add reference to the generative prior P (or G) in the figure
-  }
-  \giovanni{
-    I think that a list of possible models should replace the "experiment" 
-    cake in the middle, then C should be drawn on the arrow 
-    connecting the figure in the middle to 3a. 
-    "I" should replace the circled x of passage 3a. 
-    An additional C can be drawn alongside the arrow 2a -> 3a. 
-    After that a rectangle can be drawn around the whole middle section 
-    (including 3a, 3b and the middle C) to denote that this is phase "X".}
-  \giovanni{We might also need an arrow going from the central figure to the top arrow (the one with "T") to specify that the twin is generated according to the chosen inference model.}
 \end{figure}
@@
 
-\giovanni{This is a slightly more formal version.
+\giovanni{
+  This is a slightly more formal version. 
   I personally find it clearer. 
   If you don't like it we can use version 2 (see above in the source). 
-In case you really like it, instead, I can try to make the rest a bit more similar to this.}
+  In case you really like it, instead, I can try to make the rest a bit more similar to this.
+}
 
@@
   \caption{
-    Example 5: the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error.
-    The x-axis specifies the tree type, the y-axis shows the error.
-    The boxplot shows the minimum, first quartile, median, third 
-    quartile and maximum error, of which the values are displayed 
-    in the adjacent label
+    Example 5: the error made by BEAST2 from a phylogeny, 
+    when the generative model is known, 
+    compared to candidates with similar inference models, 
+    in relationship to the background error.
   }

### 2019-03-12 article.tex
Example 4 pipeline figure pretty

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\iffalse
 \begin{figure}[h]
@@
     };   
-    \node[state] (CB) [below right of = B, rectangle] {
+    \node[state] (CB) [right of = C, rectangle] {
       \includegraphics[height=0.2\textheight]{example_4_true_posterior_best.png}
@@
     };   
-    \node[state] (AT) [right of = A, rectangle] {
+    \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
       \includegraphics[height=0.2\textheight]{example_4_twin_tree.png}
@@
     };   
-    \node[state] (CTB) [below right of = BT, rectangle] {
+    \node[state] (CTB) [right of = CTG, rectangle] {
       \includegraphics[height=0.2\textheight]{example_4_twin_posterior_best.png}
@@
       (CG) edge [anchor = south] node {} (DG)
-      (B) edge [anchor = southeast] node {} (CB)
+      (B) edge [anchor = south east] node {} (CB)
       (CB) edge [anchor = south] node {} (DB)
@@
       (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = southeast] node {} (CTG)
+      (BT) edge [anchor = south east] node {} (CTG)
       (CTG) edge [anchor = south] node {} (DTG)
@@
 \end{figure}
-\fi
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-03-12 scripts/example_4.R

@@ -182,6 +182,20 @@ babette::plot_densitree(
 )
 dev.off()
 
+png(
+  filename = file.path(root_folder, paste0("example_", example_no, "_true_posterior_best.png")),
+  width = 1000, height = 800
+)
+babette::plot_densitree(
+  phylos = tracerer::parse_beast_trees(pir_params$experiments[[2]]$beast2_options$output_trees_filenames),
+  alpha = 0.01,
+  consensus = rev(LETTERS[1:6]),
+  cex = 2.0,
+  scaleX = TRUE,
+  scale.bar = FALSE
+)
+dev.off()
+
 if (!is_one_na(pir_params$twinning_params)) {
   png(
     filename = file.path(root_folder, paste0("example_", example_no, "_twin_posterior_gen.png")),
@@ -196,43 +210,94 @@ if (!is_one_na(pir_params$twinning_params)) {
     scale.bar = FALSE
   )
   dev.off()
+
+  png(
+    filename = file.path(root_folder, paste0("example_", example_no, "_twin_posterior_best.png")),
+    width = 1000, height = 800
+  )
+  babette::plot_densitree(
+    phylos = tracerer::parse_beast_trees(to_twin_filename(pir_params$experiments[[2]]$beast2_options$output_trees_filenames)),
+    alpha = 0.01,
+    consensus = rev(LETTERS[1:6]),
+    cex = 2.0,
+    scaleX = TRUE,
+    scale.bar = FALSE
+  )
+  dev.off()
 }
 
 ################################################################################
 # histogram of errors
 ################################################################################
 
-df_errors <- data.frame(error = read.csv(pir_params$experiments[[1]]$errors_filename)$x)
+# True, gen
+df_errors_gen <- data.frame(error = read.csv(pir_params$experiments[[1]]$errors_filename)$x)
 
 ggplot2::ggplot(
-  df_errors,
+  df_errors_gen,
   aes(x = error)
 ) + geom_histogram(binwidth = 0.01) +
-  ggsave(file.path(root_folder, paste0("example_", example_no, "_error_histogram.png")))
+  ggsave(file.path(root_folder, paste0("example_", example_no, "_error_histogram_gen.png")))
 
 ggplot2::ggplot(
-  df_errors,
+  df_errors_gen,
   aes(x = "", y = error)
 ) + geom_violin() +
   xlab("") +
[…]

### 2019-03-12 article.tex
ESSes and evidences for example 5

@@
 
+%\input{example_4_esses.latex}
+% has label tab:esses_example_4
+
+%\input{example_4_evidences.latex}
+% has label tab:evidences_example_4
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Example 5}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_4_true_tree.png}
+      \includegraphics[height=0.2\textheight]{example_5_true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_4_true_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_5_true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_4_true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_5_true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_4_error_violin_gen.png}
+      \includegraphics[height=0.2\textheight]{example_5_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_4_true_posterior_best.png}
+      \includegraphics[height=0.2\textheight]{example_5_true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_4_error_violin_best.png}
+      \includegraphics[height=0.2\textheight]{example_5_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.2\textheight]{example_4_twin_tree.png}
+      \includegraphics[height=0.2\textheight]{example_5_twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_4_twin_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_5_twin_alignment.png}
     };   
     \node[state] (CTG) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_gen.png}
     };   
     \node[state] (DTG) [below of = CTG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_gen.png}
+      \includegraphics[height=0.2\textheight]{example_5_twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_best.png}
+      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_best.png}
     };   
     \node[state] (DTB) [below of = CTB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_best.png}
+      \includegraphics[height=0.2\textheight]{example_5_twin_error_violin_best.png}
     };   
@@
   }
-  \label{fig:example_4_full_pipeline}
-  \caption{Example 4: full pipeline}
+  \label{fig:example_5_full_pipeline}
+  \caption{Example 5: full pipeline}
 \end{figure}
@@
 
-\input{example_4_esses.latex}
-% has label tab:esses_example_4
-
-\input{example_4_evidences.latex}
-% has label tab:evidences_example_4
+\input{example_5_esses_gen.latex}
+% has label tab:esses_example_5_gen
+
+\input{example_5_esses_best.latex}
+% has label tab:esses_example_5_best
+
+\input{example_5_esses_twin_gen.latex}
+% has label tab:esses_example_5_twin_gen
+
+\input{example_5_esses_twin_best.latex}
+% has label tab:esses_example_5_twin_best
+
+\input{example_5_evidence_true.latex}
+% has label tab:evidences_example_5
+
+\input{example_5_evidence_twin.latex}
+% has label tab:evidences_example_5
 

### 2019-03-12 scripts/example_5.R

@@ -1,0 +1,366 @@
+# Code of example 5
+#
+# Works under Linux and MacOS only
+library(pirouette)
+library(ggplot2)
+library(ggthemes)
+library(ggtree)
+
+root_folder <- path.expand("~/GitHubs/pirouette_article/")
+example_no <- 5
+set.seed(314)
+
+phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
+# phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
+
+alignment_params <- create_alignment_params(
+  root_sequence = create_blocked_dna(length = 1000),
+  mutation_rate = 0.1
+)
+
+# JC69, strict, Yule
+generative_experiment <- create_experiment(
+  inference_conditions = create_inference_conditions(
+    model_type = "generative",
+    run_if = "always",
+    do_measure_evidence = TRUE
+  )
+)
+check_experiment(generative_experiment)
+
+# All non-Yule tree priors
+candidate_experiments <- create_all_experiments(
+  site_models = list(create_jc69_site_model()),
+  clock_models = list(create_strict_clock_model()),
+  tree_priors = list(create_bd_tree_prior(), create_ccp_tree_prior(), create_cep_tree_prior())
+)
+check_experiments(candidate_experiments)
+
+experiments <- c(list(generative_experiment), candidate_experiments)
+check_experiments(experiments)
+
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments,
+  twinning_params = create_twinning_params()
+)
+
+################################################################################
+# Settings to run on Peregrine cluster
+################################################################################
+pir_params$alignment_params$fasta_filename <- file.path(root_folder, paste0("example_", example_no, "_true.fasta"))
+for (i in seq_along(pir_params$experiments)) {
+  pir_params$experiments[[i]]$beast2_options$input_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_input_best.xml"))
+  pir_params$experiments[[i]]$beast2_options$output_log_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output_best.log"))
[…]

### 2019-03-12 article.tex
Create figures example 4

@@
 "What is the error made by BEAST2 from a phylogeny, 
-when the generative model is known, 
+when the generative model is hand-picked, 
 compared to candidates with similar inference models, 
@@
 In this example, we use a tree generated by the Yule process.
-Instead of using the common Yule tree as in example 2, we'll use
+Instead of using the common Yule tree as in example 1, we'll use
 a Yule tree of low (actually: zero) likelihood: 
-the one as used in examples 3 and 4.
-by using the code shown in listing \ref{lst:create_yule_tree}.
+the one as used in examples 2-4.
+by using the code shown in listing \ref{lst:unknown_phylogeny}.
 The first step in \verb;pirouette; is to simulate a DNA alignment 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \includegraphics[width=\textwidth]{example_5_errors.png}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering

### 2019-03-12 scripts/example_2.R

@@ -60,11 +60,11 @@ esses <- tracerer::calc_esses(
   traces = tracerer::parse_beast_log(pir_params$experiments[[1]]$beast2_options$output_log_filename),
   sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
 )
-sink(file.path(root_folder, paste0("example_", example_no, "_esses.latex"))
+sink(file.path(root_folder, paste0("example_", example_no, "_esses.latex")))
 xtable::xtable(
-  esses, 
-  caption = paste0("ESSes of example ", example_no), 
-  label = paste0("tab:esses_example_", example_no), 
+  esses,
+  caption = paste0("ESSes of example ", example_no),
+  label = paste0("tab:esses_example_", example_no),
   digits = 0
 )
 sink()
@@ -78,7 +78,13 @@ df_evidences$clock_model_name <- plyr::revalue(
 )
 df_evidences$tree_prior_name <- plyr::revalue(
   df_evidences$tree_prior_name,
-  c("yule" = "Yule", "birth_death" = "BD", "coalescent_constant_population" = "CCP", "coalescent_exp_population" = "CEP")
+  c(
+    "yule" = "Yule",
+    "birth_death" = "BD",
+    "coalescent_bayesian_skyline" = "CBS",
+    "coalescent_constant_population" = "CCP",
+    "coalescent_exp_population" = "CEP"
+  )
 )
 names(df_evidences) <- c("Site model", "Clock model", "Tree prior", "log(evidence)", "Weight")
 

### 2019-03-12 scripts/example_4.R

@@ -3,7 +3,6 @@
 # Works under Linux and MacOS only
 library(pirouette)
 library(ggplot2)
-library(ggthemes)
 library(ggtree)
 
 root_folder <- path.expand("~/GitHubs/pirouette_article/")
@@ -11,32 +10,13 @@ example_no <- 4
 set.seed(314)
 
 phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
-# phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),
   mutation_rate = 0.1
 )
 
-# JC69, strict, Yule
-generative_experiment <- create_experiment(
-  inference_conditions = create_inference_conditions(
-    model_type = "generative",
-    run_if = "always",
-    do_measure_evidence = TRUE
-  )
-)
-check_experiment(generative_experiment)
-
-# All non-Yule tree priors
-candidate_experiments <- create_all_experiments(
-  site_models = list(create_jc69_site_model()),
-  clock_models = list(create_strict_clock_model()),
-  tree_priors = list(create_bd_tree_prior(), create_ccp_tree_prior(), create_cep_tree_prior())
-)
-check_experiments(candidate_experiments)
-
-experiments <- c(list(generative_experiment), candidate_experiments)
+experiments <- create_all_experiments()
 check_experiments(experiments)
 
 pir_params <- create_pir_params(
@@ -56,11 +36,6 @@ for (i in seq_along(pir_params$experiments)) {
   pir_params$experiments[[i]]$beast2_options$output_state_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output_best.xml.state"))
   pir_params$experiments[[i]]$errors_filename <- file.path(root_folder, paste0("example_", example_no, "_error_best.csv"))
 }
-pir_params$experiments[[1]]$beast2_options$input_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_input_gen.xml"))
-pir_params$experiments[[1]]$beast2_options$output_log_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output_gen.log"))
[…]

### 2019-03-12 scripts/example_5.R

@@ -10,8 +10,8 @@ root_folder <- path.expand("~/GitHubs/pirouette_article/")
 example_no <- 5
 set.seed(314)
 
+# An infintely rare Yule tree
 phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
-# phylogeny <- create_ideal_tree(n_taxa = 5, crown_age = 10)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),

### 2019-03-12 article.tex
Nice figure on ERQ 4

@@
 
-%\input{example_4_esses.latex}
-% has label tab:esses_example_4
-
-%\input{example_4_evidences.latex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[ht]
+  \centering
+  \resizebox {1.0\columnwidth} {!} {
+    \begin{tikzpicture}[
+      ->,>=stealth',shorten >=1pt,auto,
+      node distance=0.5\textheight, 
+      semithick
+    ]   
+    \tikzstyle{every state}=[]
+    \node[state, draw=none] (O) [] {
+    };   
+    \node[state] (A) [right of = O, rectangle] {
+      \includegraphics[height=0.4\textheight]{example_4_true_tree.png}
+    };   
+    \node[state] (B) [below of = A, rectangle] {
+      \includegraphics[height=0.45\textheight]{example_4_true_alignment.png}
+    };   
+    \node[state] (CB) [below of = B, rectangle, node distance=0.8\textheight] {
+      \includegraphics[height=0.95\textheight]{example_4_true_posterior_best.png}
+    };   
+    \node[state] (DB) [below of = CB, rectangle, node distance=1.0\textheight] {
+      \includegraphics[height=0.65\textheight]{example_4_error_violin_best.png}
+    };   
+    \node[state] (AT) [right of = A, rectangle, node distance=2.0\textwidth] {
+      \includegraphics[height=0.4\textheight]{example_4_twin_tree.png}
+    };   
+    \node[state] (BT) [below of = AT, rectangle] {
+      \includegraphics[height=0.45\textheight]{example_4_twin_alignment.png}
+    };   
+    \node[state] (CTB) [below of = BT, rectangle, node distance=0.8\textheight] {
+      \includegraphics[height=0.95\textheight]{example_4_twin_posterior_best.png}
+    };   
+    \node[state] (DTB) [below of = CTB, rectangle, node distance=1.0\textheight] {
+      \includegraphics[height=0.65\textheight]{example_4_twin_error_violin_best.png}
+    };   
+    \path 
+      (O) edge [anchor = south] node {} (A)
+      (A) edge [anchor = south] node {} (B)
+      (B) edge [anchor = south east] node {} (CB)
+      (CB) edge [anchor = south] node {} (DB)
+      (A) edge [anchor = east] node {} (AT)
+      (AT) edge [anchor = south] node {} (BT)
+      (BT) edge [anchor = south] node {} (CTB)
+      (CTB) edge [anchor = south] node {} (DTB)
+    ; 
+    \end{tikzpicture}
+  }
+  \label{fig:example_4_full_pipeline}
+  \caption{Example 4: full pipeline}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\input{example_4_esses_best.latex}
+% has label tab:esses_example_4_best
+
+\input{example_4_esses_twin_best.latex}
+% has label tab:esses_example_4_twin_best
+
+\input{example_4_evidence_true.latex}
+% has label tab:evidences_example_4
+
+\input{example_4_evidence_twin.latex}
 % has label tab:evidences_example_4

### 2019-03-12 scripts/example_4.R

@@ -250,3 +250,4 @@ if (!is_one_na(pir_params$twinning_params)) {
     ggsave(file.path(root_folder, paste0("example_", example_no, "_twin_error_violin_best.png")))
 
 }
+

### 2019-03-12 article.tex
Updates from Overleaf

--- 2019-03-12 article.tex
@@
     $\mathit{e_{f}}$ & error function & Specifies how to measure the error & nLTT, $|\gamma|$ \\
-    $\mathit{t_{m}}$ & twin model & Sets the tree prior for the twin tree & Yule, BD \\
+    $\mathit{t_{m}}$ & twin tree prior & Sets the tree prior for the twin tree & Yule, BD \\
     \hline
@@
     \hline
-    $\mathit{G}$ & Generative prior & Diversification model to generate the 'true' tree & $\mathit{t_{p}}$ \\
+    $\mathit{P}$ & Generative tree prior & Diversification model to generate the 'true' tree & $\mathit{t_{p}}$ \\
     $\mathit{A}$ & Alignment model & Model used to simulate the alignment ($\mathit{c_{m}}$ must be strict) & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{m_{r}}, \mathit{r_{s}}$ \\
+    $\mathit{G}$ & Generative model & The full setting to produce BEAST2 input data & $\mathit{P}$, $\mathit{A}$ \\
     $\mathit{I}$ & Inference model & Phylogenetic inference model to run BEAST2 & $\mathit{c_{m}},\mathit{s_{m}},\mathit{t_{p}},\mathit{m_{p}},\mathit{m_{s}}$ \\
@@
 
-%\input{example_4_esses.latex}
-% has label tab:esses_example_4
-
-%\input{example_4_evidences.latex}
-% has label tab:evidences_example_4
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 5}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\iffalse
 \begin{figure}[h]
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_true_tree.png}
+      \includegraphics[height=0.2\textheight]{example_4_true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5_true_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_4_true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_4_true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_error_violin_gen.png}
-    };   
-    \node[state] (CB) [right of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_true_posterior_best.png}
+      \includegraphics[height=0.2\textheight]{example_4_error_violin_gen.png}
+    };   
+    \node[state] (CB) [below right of = B, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_error_violin_best.png}
-    };   
-    \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_tree.png}
+      \includegraphics[height=0.2\textheight]{example_4_error_violin_best.png}
+    };   
+    \node[state] (AT) [right of = A, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5_twin_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_4_twin_alignment.png}
     };   
     \node[state] (CTG) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_gen.png}
     };   
     \node[state] (DTG) [below of = CTG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_error_violin_gen.png}
-    };   
-    \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_best.png}
+      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_gen.png}
+    };   
+    \node[state] (CTB) [below right of = BT, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_4_twin_posterior_best.png}
     };   
     \node[state] (DTB) [below of = CTB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_error_violin_best.png}
+      \includegraphics[height=0.2\textheight]{example_4_twin_error_violin_best.png}
     };   
@@
       (CG) edge [anchor = south] node {} (DG)
-      (B) edge [anchor = south east] node {} (CB)
+      (B) edge [anchor = southeast] node {} (CB)
       (CB) edge [anchor = south] node {} (DB)
@@
       (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = south east] node {} (CTG)
+      (BT) edge [anchor = southeast] node {} (CTG)
       (CTG) edge [anchor = south] node {} (DTG)
@@
   }
-  \label{fig:example_5_full_pipeline}
-  \caption{Example 5: full pipeline}
+  \label{fig:example_4_full_pipeline}
+  \caption{Example 4: full pipeline}
 \end{figure}
+\fi
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\input{example_5_esses_gen.latex}
-% has label tab:esses_example_5_gen
-
-\input{example_5_esses_best.latex}
-% has label tab:esses_example_5_best
-
-\input{example_5_esses_twin_gen.latex}
-% has label tab:esses_example_5_twin_gen
-
-\input{example_5_esses_twin_best.latex}
-% has label tab:esses_example_5_twin_best
-
-\input{example_5_evidence_true.latex}
-% has label tab:evidences_example_5
-
-\input{example_5_evidence_twin.latex}
-% has label tab:evidences_example_5
+\input{example_4_esses.latex}
+% has label tab:esses_example_4
+
+\input{example_4_evidences.latex}
+% has label tab:evidences_example_4
 

### 2019-03-13 article.tex
Improve introduction

@@
     estimates. An important ingredient of BEAST2 is 
-    the model for the species tree prior, 
-    and the tool can in principle be used for inference of 
+    the species tree prior, 
+    due to which the tool can be used for inference of 
     macroevolutionary processes underlying the species tree. 
@@
 constant-rate birth-death [\cite{nee1994reconstructed}] models.
-These models are most commonly picked in biological research,
-as these are of a desired minimum biological complexity, 
-yet computationally fast enough.
+These models are the most commonly picked in biological research,
+as these embody a reasonable level of biological complexity, 
+yet are computationally fast enough.
 The recent development of BEAST2 [\cite{bouckaert2014beast}] provides, 
@@
 or diversity-dependent [\cite{etienne2011diversity}],
-or that can shift [\cite{etienne2012conceptual}, \cite{rabosky2014automatic}, \cite{alfaro2009nine}, the SLS model. Laudanno et al., in preparation].
-Other birth-death models treat speciation as a process that takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}], or that happens in bursts of simultaneous events [Laudanno et al., in preparation], or that
+or that can shift [\cite{etienne2012conceptual}, 
+\cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}].
+Other birth-death models treat speciation as a process that takes 
+time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}], 
+or that happens in bursts of simultaneous 
+events [\cite{laudanno2018mbd}], or that
 depends on a trait that has two [\cite{maddison2007estimating}], 
@@
 
-For a given phylogeny, the parameter values of the diversification model 
-are usually estimated using maximum likelihood.
-For such  models, 
-it is a standard practice to assess their performance 
-on a controlled environment: 
-phylogenies are simulated with known parameters, 
-and then the estimated parameters are compared with the original parameters.
-If the likelihood proves to be effective to describe the process, 
-it can potentially be implemented into BEAST2 as a new prior. 
+When investigating a diversification process,
+starting from a given phylogeny, 
+the parameter values of the diversification model 
+are commonly estimated using maximum likelihood.
+To test the effectiveness of the diversification model,
+it is standard practice to simulate new phylogenies
+from the (maximum likely) parameters and use maximum
+likelihood to again estimate the parameters.
+An effective diversification model has an
+initial parameter estimate that is similar to the control parameter estimates.
+
 However, in many cases the computation of this likelihood (the tree prior) 

### 2019-03-13 article.tex
Improve introduction

@@
 
-However, in many cases the computation of this likelihood (the tree prior) 
-is computationally demanding and it is unfeasible to use it 
-in a Bayesian inference framework where it needs to be 
-computed millions to billions of times.
-However, if the data are very informative, the influence of the prior 
-is often relatively small, and hence under this assumption 
-it has become common practice to use already existing BEAST2 priors. 
-However, it is unclear whether this assumption holds.
-Here we introduce a pipeline to assess this assumption 
-for any given diversification model. 
+Not all diversification models that pass the procedure described above
+are used in Bayesian inference.
+In Bayesian inference, the likelihood of a diversification 
+model, also called tree prior, is computed millions of time.
+Due to this, effective but computationally expensive tree priors 
+are unused in Bayesian inference.
+Also, there may be no need to add such a tree prior to a Bayesian
+inference framework: the more informative the data, 
+the lower the influence of the tree prior.
+Due to this, existing and easy to compute tree priors are most commonly used.
+However, it is unknown if it is valid to assume that tree prior choice is 
+of low importance.
+Here we introduce a pipeline to assess this assumption and quantify the
+effect of the tree prior of any given diversification model. 
 Briefly, it consists of the following steps: 

### 2019-03-13 article.tex
Improved introduction

@@
 of low importance.
-Here we introduce a pipeline to assess this assumption and quantify the
-effect of the tree prior of any given diversification model. 
-Briefly, it consists of the following steps: 
-first simulate a phylogeny 
-according to the mechanisms proposed by the new diversification model, 
-then simulate a nucleotide alignment 
-and use current inference tools (such as BEAST2) to estimate the phylogeny. 
-If the estimated tree differs only slightly from the original phylogeny, 
-the original assumption is justified and the simple diversification models 
-in the inference tools suffice. 
-If the trees differ markedly, it will be worth the effort and computational burden 
-to develop a species tree module to be used in these tools.
-Our pipeline is programmed as an R package and is called 
-\verb;pirouette; it is built on \verb;babette; [Bilderbeek \& Etienne, 2018], 
+
+Here we introduce a method to test the importance of a tree prior.
+The method starts with a phylogeny generated by the process of 
+a (usually novel) tree prior.
+Using existing tree priors, a Bayesian posterior with phylogenies is inferred.
+The more simular the posterior phylogenies 
+and the original phylogeny are,
+the less important the novel tree prior has been.
+Only tree priors that are measured to be influential
+will be worth the effort and computational burden 
+to use in a Bayesian framework.
+
+This method is programmed as an R package called \verb;pirouette;.
+\verb;pirouette; is built on \verb;babette; [\cite{bilderbeek2018babette}], 
 which calls BEAST2 [\cite{bouckaert2014beast}]. 
-With \verb;pirouette;, one
-can easily measure the error made by Bayesian inference in recovering
-any given phylogeny, helping us to evaluate the necessity of a new diversification model.
+With \verb;pirouette;, one has a method to quantify the importance 
+of a novel diversification model.
 

### 2019-03-13 article.tex
Use '~\ref' instead of ' \ref', as the tilde marks a non-breaking space. Thanks @Giappo

@@
 \verb;pirouette; is written in the R programming language (\cite{R}).
-The goal of \verb;pirouette; is to measure the inference error BEAST2
+The goal of \verb;pirouette; is to quantify the importance of a tree
+prior. It does so by measuring the inference error BEAST2
 makes for a given reconstructed phylogeny, 
-simulated under the mechanisms of a (possibly new) diversification model, 
-which we will refer to as the 'generative prior'. 
+simulated under the mechanisms of a (usually novel) diversification model, 
+which we will refer to as the 'generative tree prior', $\mathit{P}$. 
 Such a study will typically construct phylogenies 
@@
 according to how they operate in the pipeline. 
-We summarize them in Tables~\ref{tab:options} and \ref{tab:definitions}.
+We summarize them in tables~\ref{tab:options} and~\ref{tab:definitions}.
 
@@
 An overview of \verb;pirouette;'s main functions is shown in 
-table \ref{tab:functions}. 
+table~\ref{tab:functions}. 
 We will exploit these functions to answer our research questions in the next sections.
@@
       Create a Yule tree. 
-      The resulting tree is shown in figure \ref{fig:yule_tree}
+      The resulting tree is shown in figure~\ref{fig:yule_tree}
     }
@@
   \includegraphics[width=\textwidth]{tree_yule.png}
-  \caption{The Yule tree, as created by listing \ref{lst:create_yule_tree}.}
+  \caption{The Yule tree, as created by listing~\ref{lst:create_yule_tree}.}
   \label{fig:yule_tree}
@@
 
-The resulting figure is shown in figure \ref{fig:example_1}.
+The resulting figure is shown in figure~\ref{fig:example_1}.
 
@@
   \caption{The tree derived from an unknown diversification process, 
-    as created by listing \ref{lst:unknown_phylogeny}.
+    as created by listing~\ref{lst:unknown_phylogeny}.
   }
@@
 We will re-use the alignment parameters of the previous example 
-as shown in listing \ref{lst:create_alignment}.
+as shown in listing~\ref{lst:create_alignment}.
 
@@
 
-The resulting figure is shown in figure \ref{fig:example_2}.
+The resulting figure is shown in figure~\ref{fig:example_2}.
 
@@
 the same as in the previous section, which are
-the phylogeny from an unknown speciation model (listing \ref{lst:unknown_phylogeny}), 
-the alignment parameters (listing \ref{lst:create_alignment}) 
-and the list of experiments (listing \ref{lst:create_generative_experiment}).
+the phylogeny from an unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
+the alignment parameters (listing~\ref{lst:create_alignment}) 
+and the list of experiments (listing~\ref{lst:create_generative_experiment}).
 
@@
 
-The resulting figure is shown in figure \ref{fig:example_3}
+The resulting figure is shown in figure~\ref{fig:example_3}
 
@@
 Here, same as example 2 and 3, we start with the phylogeny 
-from an unknown speciation model (listing \ref{lst:unknown_phylogeny}), 
+from an unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
 The first step in \verb;pirouette; is to simulate a DNA alignment 
 from the given phylogeny, for which we will use the same code 
-as shown in listing \ref{lst:create_alignment}.
+as shown in listing~\ref{lst:create_alignment}.
 
@@
 calling \verb;create_all_experiments;,
-as done in listing \ref{lst:create_all_experiments}.
+as done in listing~\ref{lst:create_all_experiments}.
 
@@
 default created twinning parameters, as shown in 
-listing \ref{lst:create_twinning_params}.
+listing~\ref{lst:create_twinning_params}.
 
@@
 
-The resulting figure is shown in figure \ref{fig:example_4}.
+The resulting figure is shown in figure~\ref{fig:example_4}.
 
@@
 the one as used in examples 2-4.
-by using the code shown in listing \ref{lst:unknown_phylogeny}.
+by using the code shown in listing~\ref{lst:unknown_phylogeny}.
 The first step in \verb;pirouette; is to simulate a DNA alignment 
 from the given phylogeny, for which we will use the same code 
-as shown in listing \ref{lst:create_alignment}.
+as shown in listing~\ref{lst:create_alignment}.
 
@@
 
-The resulting figure is shown in figure \ref{fig:example_5}.
+The resulting figure is shown in figure~\ref{fig:example_5}.
 
@@
 to be obtained from thousands of replicates.
-In that case, figure \ref{fig:example_1} shows 
+In that case, figure~\ref{fig:example_1} shows 
 the minimum error BEAST2 when the
@@
 a minimum noise level assessment in a bigger experiment.  
-Figure \ref{fig:example_2} shows that 
+Figure~\ref{fig:example_2} shows that 
 \richel{
@@
 hinting that the candidate model would be just as good in inference.
-Figure \ref{fig:example_3} shows that 
+Figure~\ref{fig:example_3} shows that 
 \richel{
@@
 does not effect the error made by BEAST2.
-Figure \ref{fig:example_4} shows that 
+Figure~\ref{fig:example_4} shows that 
 \richel{

### 2019-03-13 article.tex
To and including '2 Description'

@@
 \usepackage{authblk}
-\title{pirouette: Determining the error BEAST2 makes in inferring a phylogeny}
+
+% RJCB: I think this title is better: 
+% (1) tree prior importantance is the goal, the error is just the means
+% (2) no need to specify BEAST2, it's also just a means
+\title{pirouette: Quantifying the importance of a tree prior in Bayesian phylogenetics}
+% \title{pirouette: Determining the error BEAST2 makes in inferring a phylogeny}
+
 \author[1]{Rich\`el J.C. Bilderbeek}
@@
 Although many possible tests can be performed, 
-we present here only five example research questions (ERQ) that \verb;pirouette; can answer. 
-For these ERQs we will estimate the error made by BEAST2 when 
-using a birth-death tree prior, in recovering an original tree 
-generated under a Yule generative prior.
+we present here what we think are the five most important 
+example research questions (ERQ) that \verb;pirouette; can answer. 
+For these ERQs we will quantify the importance of a tree prior.
 
@@
   \caption{
-    Most important parameter options. BD = birth death, CBS = coalescent
-    Bayesian Skyline, CCP = coalescent constant-population, CEP = coalescent
-    exponential-population, JC = Jukes-Cantor 1969 (\cite{jukes1969evolution}), HKY = Hasegawa, Kishino and 
-    Yano (\cite{hasegawa1985dating}), TN93 = \cite{tamura1993estimation}, GTR = Generalised time-reversible 
-    model (\cite{tavare1986some}).
+    Most important parameter options. 
+    BD = birth death (\cite{nee1994reconstructed}), 
+    CBS = coalescent Bayesian Skyline (\cite{drummond2005bayesian}), 
+    CCP = coalescent constant-population, 
+    CEP = coalescent exponential-population,
+    JC = Jukes-Cantor 1969 (\cite{jukes1969evolution}), 
+    HKY = Hasegawa, Kishino and  Yano (\cite{hasegawa1985dating}), 
+    TN93 = \cite{tamura1993estimation}, 
+    GTR = Generalised time-reversible model (\cite{tavare1986some}).
   }
@@
     \texttt{pirouette} pipeline. 
-    The pipeline starts from a true phylogeny (1a) simulated under the selected generative prior $\mathit{P}$, that is converted to an alignment (2a) according to the alignment model $\mathit{A}$.
+    The pipeline starts from a phylogeny (1a) simulated by generative prior $\mathit{P}$.
+    The phylogeny is converted to an alignment (2a) using alignment model $\mathit{A}$. 
     The user supplies one or more experiments.
-Every experiment $\mathit{X}$ is defined as a combination of an inference model $\mathit{I}$ and the conditions $\mathit{C}$ under which it is actually used in the inference.
-Based on the experiments and possibly the alignment (2a), experiments are selected to be actually used in the next step (3a).
-    Selected experiments use the alignment (2a) and their inference model $\mathit{I}$ to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a).
-The posterior trees are compared to the true phylogeny (1a) using the selected error measure $\mathit{E}$, in our case the nLTT statistic [\cite{janzen2015approximate}].
-Taking the differences between the true tree's nLTT and each posterior's nLTT results in an error distribution, here displayed as a histogram
-    of error values (5a). Optionally, a twin phylogeny (1b) can be generated from the true phylogeny (1a), which then results in an error distribution (5b) with similar intermediate steps.
+    For each experiment $\mathit{X}$ (a tuple of inference model $\mathit{I}$ and condition $\mathit{C}$),
+    the inference model $\mathit{I}$ is used in the next step (3a),
+    if the condition $\mathit{C}$ are satisfied.
+    The condition $\mathit{C}$ can depend on the alignment (2a).
+    The allowed inference models use the alignment (2a) 
+    to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a). 
+    Each of the posteriors' trees is compared to the true phylogeny (1a) 
+    using error measure $\mathit{E}$, 
+    resulting in an error distribution (5a). 
+    Optionally, a twin phylogeny (1b) can be generated from the original 
+    phylogeny (1a), which then results in an error distribution (5b) 
+    with similar intermediate steps.
   }
@@
 
-The pipeline is summarized by the following steps, which will be described in detail below:
+The pipeline is summarized by the following steps:
 \begin{enumerate}

### 2019-03-13 article.tex
I enjoy ERQ 1 so much better now

@@
 \end{lstlisting}
-For the option to select inference models,
-\verb;pirouette; uses the "NS" BEAST2 package [\cite{maturana2018model}].
+For the option to select a best candidate model,
+\verb;pirouette; needs the "NS" BEAST2 package [\cite{maturana2018model}].
 It can be installed from within R, using:
@@
 table~\ref{tab:functions}. 
-We will exploit these functions to answer our research questions in the next sections.
+These functions are demonstrated when in the example code below.
+All \verb;pirouette;'s functions are documented,
+have a useful example, and have sensible defaults.
 
@@
 
-All \verb;pirouette;'s functions are documented. 
-Useful examples and sensible defaults, 
-which we will refer to in the next sections, 
-can be accessed from the documentation.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\verb;pirouette; quantifies the influence of a tree prior,
+by measuring the difference between a given tree and 
+phylogenies in a posterior. Due to stochasticity, all posterior trees
+will differ from the given phylogeny, even when the generative tree
+prior G and alignment model A are the same as 
+those in the inference model.
+Measuring this difference results allows us to know the baseline error
+of the \verb;pirouette; pipeline.
+Because only standard tree priors are used in inference, 
+and G must be used such a tree prior, it follows that
+G must also be a standard tree prior.
+
 A first research question that \verb;pirouette; can answer is:
 
-"What is the error made by BEAST2 from a phylogeny 
-using the same diversification model as it was generated by?"
-
-We start from a tree simulated under a Yule (pure-birth) generative prior. 
-We choose to use a tree with six taxa and a crown age of ten.
+"What is the inference error made on phylogenies
+created by a standard diversification model?"
+
+In this example we use the Yule (pure-birth) tree prior 
+as the generative tree prior G,
+because it is the simplest standard tree prior. 
+Table~\ref{tab:options} shows an overview of the standard
+tree priors.
+We choose to use a small tree with six taxa to keep
+the calculations short and the figure more readable.
+We pick a crown age of ten time units. This value is 
+completely arbitrary, but does tie in with the mutation rate 
+used in simulating an alignment in the next step.
 
@@
   \giovanni{Remove the text in top-left corner.}
+  \richel{Agreed, iff the text matches the code}
   \giovanni{I'd use a sexier topology, purely for esthetical reasons.}
+  \richel{Agreed, will pick a sexy RNG seed after create\_yule\_tree is there}
 \end{figure}
@@
 The first step in \verb;pirouette; is to simulate a DNA alignment from the given phylogeny. 
-To do so, we must specify the DNA root sequence and a mutation rate. 
+To do so, we must specify the DNA root sequence, DNA mutation rate,
+site model (in this context: a nucleotide substitution model) and a clock model. 
 In this example, the DNA root sequence consists in four blocks of 250 nucleotides each, 
 where the per-nucleotide mutation rate is 0.1 mutations per unit time.
+Because the crown age of the phylogeny is 10 time units, each nucleotide will
+have one (possibly-silent) mutation on average.
+We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
+and a strict clock model as these are the simplest.
+A Jukes-Cantor (JC, \cite{jukes1969evolution}) site model,
+assumes that mutation rates between nucleotides are equal and constant. 
+A strict clock model assumes that the mutation rates 
+of all species are equal and constant.
 
@@
   root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
+  mutation_rate = 0.1,
+  site_model = create_jc69_site_model(),
+  clock_model = create_strict_clock_model()
 )
@@
 
-By default, an alignment is created using a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
-and a strict clock model. 
-A JC site model assumes that mutation rates from any nucleotide to any other 
-are equal and constant. 
-A strict clock model assumes that the mutation rates of all species are equal and constant.
-Due to this, we state that the generative model for the alignment is
-a JC site model and a strict clock model.
-
-In the second step we state our experiments. In this context, we
-define an experiment as a combination of an inference model
-and the conditions under which a Bayesian inference is executed.
-Within this example, we specify that our experiment uses the
-generative model (which means that the inference model will be 
-a combination of JC site model, strict clock model and Yule tree prior), 
-it will always be run and we skip the evidence measurement for our model.
+As the site and clock model used here are also the defaults, 
+these function arguments can be safely omitted: we just show these
+explicitly for clarity.
+
+In the second step we state our experiments, which will be just one.
+We define an experiment X as a combination of an inference model I
+and conditions, C.
+In this example we pick I to be the same inference model as the generative one,
+which is Yule tree prior, JC site model and strict clock model.
+We specify C that the experiment will always be run.
 
@@
 results in:
-\giovanni{Would it be clearer not to use sensible defaults where we call the functions with arguments we refer to in the main text? For example here it could be clearer for the reader to see:
-create\_experiment(site\_model = create\_jc69\_site\_model(), clock\_model = create\_strict\_clock\_model(), tree\_prior = create\_yule\_tree\_prior())
-}
+
+\begin{lstlisting}[
+  language=R, 
+  floatplacement=ht, frame=single,
+  label = {lst:create_generative_experiment_explicit},
+  caption = {
+    Create the experiments to answer research question 1, 
+    using explicit arguments
+  }
+]
+experiments <- list(
+  create_experiment(
+    inference_conditions = create_inference_conditions(
+      model_type = "generative", 
+      run_if = "always"
+    ), 
+   inference_model = create_inference_model(
+     site_model = create_jc69_site_model(), 
+     clock_model = create_strict_clock_model(), 
+     tree_prior = create_yule_tree_prior()
+    )
+  )
+)
+\end{lstlisting}
+
+Due to sensible defaults, the same can be achieved by:
 
@@
   caption = {
-    Create a default experiment, as described in Table~\ref{tbl:RQ1}.
+    Create the experiments to answer research question 1, 
+    using the default arguments
   }
@@
 
-All the \verb;pirouette; arguments are bundled
+The objects we just created need to be bundled
+in a \verb;pirouette; parameters structure,
+together with error measurement parameters E and
+twinning parameters T. We use the default error measurement
+parameters, which uses a burn-in of 10\% and the nLTT statistic to
+measure the difference between phylogenies. We will not use twinning
+here yet.  
+All these arguments are bundled
 and checked by \verb;create_pir_params;:
@@
   alignment_params = alignment_params,
-  experiments = experiments
+  experiments = experiments,
+  error_measure_params = create_error_measure_params(),
+  twinning_params = NA
 )
@@
 
-Running the experiment:
+Now we can use our Yule tree and \verb;pir_params; to measure 
+the inference error made on phylogenies
+created by a standard diversification model:
 
@@
 errors <- pir_run(
-  phylogeny,
+  phylogeny = phylogeny,
   pir_params = pir_params
@@
 
-\verb;pirouette; has a plotting function used for convenience:
+One the hardware we used, it will take approximately 10 minutes
+to obtain the error distribution returned by \verb;pir_run;.
+This error distribution is a data frame in the non-tidy short form.
+Instead of converting it to tidy data and plotting it,
+\verb;pirouette; has a plotting function that does exactly that:
 
@@
 \end{lstlisting}
-
-The resulting figure is shown in figure~\ref{fig:example_1}.
 
@@
 
-The error distribution obtained in this way can serve as a control:
-imagine one develops a novel tree prior related to the Yule model.
-One can predict that BEAST2 will be worse in recovering the phylogenies
-from that novel tree prior, as only the existing tree priors are used
-in inference. 
-To determine how much worse the recovery is, one needs to know the 
-inference error made by existing tree priors. 
-As this control is important, it is an inherent part of \verb;pirouette;:
-it is called twinning, which is shown in example research question 3.
+The error distribution obtained is shown as a violin plot
+in figure~\ref{fig:example_1}.
+
+This error distribution has little relevance on its own. 
+It only gives a baseline error made by
+the Bayesian inference. We use 'a', because we have used only one Yule
+tree. To obtain the baseline error, one should simulate multiple Yule
+trees and merge their error distributions.
+
+For this error distribution to have more relevance it needs to be compared
+to an error distribution derived from a phylogeny 
+derived from a non-standard speciation process.
+In that way, the error distribution we have now serves as a control,
+as it is the baseline error caused by stochasticity.
+In this example, that non-standard speciation process must 
+be closest related to a Yule process for the comparison to make sense.
+Would the two error distributions be similar, the non-standard speciation 
+process is of little influence.
+Because measuring the baseline error is so important,
+we made it an inherent part of \verb;pirouette;. 
+We call it twinning and we use it in example research question 3.
 

### 2019-03-13 article.tex
Doing nicely reviewing ERQ 2

@@
 
-Due to sensible defaults, specifying this
-results in:
+Listing~\ref{lst:create_generative_experiment_explicit} shows how to
+set up this experiment:
 
@@
 
+There are more ways to pick a standard inference model. In the
+previous example, we picked the known generative inference model.
+A novel tree prior, by definition, is non-standard, thus cannot be
+part of a standard inference model. 
+One option is to choose to hand-pick a related 
+standard tree prior, as done in ERCs 3 and 5, and combine it with 
+a known generative alignment model.
+An alternative is to let multiple candidate inference models compete 
+and pick the best one, based on the highest evidence (aka marginal likelihood):
+when an alignment is simulated from the given tree, one can estimate the
+evidence for each inference model given that 
+alignment \richel{or the other way around?}.
+The benefit of this approach is that this is a commonly used procedure,
+which picks an inference model that is
+best at explaining the alignment, where more complex inference models are
+penalized for having more parameters.
+
 A second research question that \verb;pirouette; can answer, is:
 
-"What is the error made by BEAST2 from a phylogeny when
+"What is the inference error made on a novel phylogeny when
 picking the best inference model?"
@@
 
-In the second step we state our experiments. 
-Within this example, all our experiments are candidates,
-run only if it is the best candidate, the evidence is measured (ignoring
-this will give a helpful error) and we use the full set of 
-40 inference models, which are all combinations of 4 site 
+In the second step we set up our experiments. 
+Within this ERQ, each of our experiments contain a different
+inference model $I$, but has condition $C$ that will only let $I$
+be run in inference if that inference model $I$ is the 
+best candidate inference model, that is, 
+it is the inference model with the highest evidence.
+Setting up experiments like this is thought to be so common,
+that \verb;pirouette; has a simple function to do:
+\verb;create_all_experiments; will create 40 experiments,
+containing inference models of all combinations of 4 site 
 models, 2 clock models and 5 tree priors.
@@
 
-Also here, the third (the BEAST2 inference) and fourth (measuring the error)
-steps have sensible defaults, and we are not
-interested in using a twin tree. We can create the complete
-\verb;pirouette; parameter set (again) like this:
+For all other steps we use the default setups. 
+We can create the complete
+\verb;pirouette; parameter set like this:
 

### 2019-03-14 article.tex
Improved worksflow figure

@@
     The user supplies one or more experiments.
-    For each experiment $\mathit{X}$ (a tuple of inference model $\mathit{I}$ and condition $\mathit{C}$),
-    the inference model $\mathit{I}$ is used in the next step (3a),
-    if the condition $\mathit{C}$ are satisfied.
-    The condition $\mathit{C}$ can depend on the alignment (2a).
+    For each experiment $\mathit{X}$ 
+    (a tuple of inference model $\mathit{I}$ and condition $\mathit{C}$),
+    if its condition $\mathit{C}$ is 
+    satisfied (which can depend on the alignment), 
+    its inference model $\mathit{I}$ is used in the next step (3a).
     The allowed inference models use the alignment (2a) 
@@
     Optionally, a twin phylogeny (1b) can be generated from the original 
-    phylogeny (1a), which then results in an error distribution (5b) 
-    with similar intermediate steps.
+    phylogeny (1a) using twinning parameters $T$, where $T$ may depend on
+    the experiments \richel{improve}.
+    From a twin phylogeny, the same intermediate steps are taken,
+    resulting in a twin error distribution (5b)
   }

### 2019-03-14 article.tex
Give listings labels

@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[
+  language=R, floatplacement=ht, frame=single,
+  label = {lst:create_pir_params},
+  caption = Calling \texttt{create\_pir\_params}
+]
 pir_params <- create_pir_params(
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[
+  language=R, floatplacement=ht, frame=single,
+  label = {lst:pir_run},
+  caption = Calling \texttt{pir\_run}
+]
 errors <- pir_run(
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[
+  language=R, floatplacement=ht, frame=single,
+  label = {lst:pir_plot},
+  caption = Calling \texttt{pir\_plot}
+]
 pir_plot(errors)
@@
 \verb;create_all_experiments; will create 40 experiments,
-containing inference models of all combinations of 4 site 
+containing the inference models of all combinations of 4 site 
 models, 2 clock models and 5 tree priors.
@@
 We can create the complete
-\verb;pirouette; parameter set like this:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\verb;pirouette; parameter set like this (which is the
+same as listing~\ref{lst:create_pir_params}, but using the defaults):
+
+\begin{lstlisting}[
+  language=R, floatplacement=ht, frame=single,
+  label = {lts:create_pir_params_defaults},
+  caption = Create a \texttt{pir\_params} with many defaults
+]
 pir_params <- create_pir_params(
@@
 
-Running the experiment:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-\end{lstlisting}
-
-Again, showing the results:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-pir_plot(errors)
-\end{lstlisting}
-
+We can run the \verb;pirouette; method with listing~\ref{lst:pir_run}
+and plot the results with listing~\ref{lst:pir_plot}.
 The resulting figure is shown in figure~\ref{fig:example_2}.

### 2019-03-15 article.tex
Changes @Giappo

@@
 
-\iffalse
-\begin{table}
-\centering
-  \begin{tabular}{|@{}c|p{2.8cm}|p{6cm}@{}|}
-    \hline
-    \textbf{Symbol} & \textbf{Name} & \textbf{Definition} \\ 
-    \hline
-    $\mathit{t}$ & Tree prior & Macroevolutionary speciation model \\
-    $\mathit{c}$ & Clock model & Clock for the DNA mutation rates\\
-    $\mathit{s}$ & Site model & Nucleotide substitution model\\
-    $\mathit{\mu}$ & Mutation rate & Pace at which mutation occurs, $\mu \in \mathit{s}$ \\
-    $\mathit{r}$ & Root sequence & DNA sequence at the root of the tree\\
-    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent common ancestor \\
-    $\mathit{v}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup \\
-    \hline
-    $\mathit{P}$ & Generative prior & Diversification model under which the 'true' tree is simulated \\
-    $\mathit{A}$ & Alignment model & Alignment simulation model. It is a combination of $(\mathit{r}, \mathit{\mu}, \mathit{c}, \mathit{s})$, where $\mathit{c}$ = strict \\
-    $\mathit{G}$ & Generative model & Combination of a known $\mathit{P}$ and chosen (thus known) $\mathit{A}$ \\
-    $\mathit{I}$ & Inference model & Bayesian phylogenetic inference model, a combination of
-    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
-    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference \\
-    $\mathit{X}$ & Experiment & Combination of $(\mathit{I},\mathit{C})$\\
-    $\mathit{E}$ & Error measurement parameters & Errors measurement setup \\
-    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure \\
-    \hline 
-  \end{tabular}
-  \caption{
-    Definitions of terms and relative symbols used in this article.
-  }
-  \label{tab:definitions}
-\end{table}
-\fi
-
-\iffalse
-% previous version
-\begin{table}
-\centering
-  \begin{tabular}{|@{}c|p{2.8cm}|@{}l|p{5.5cm}@{}|}
-    \hline
-    \textbf{Symbol} & \textbf{Name} & \textbf{Definition} \\ 
-    \hline
-    $\mathit{t}$ & Tree prior & Macroevolutionary speciation model \\
-    $\mathit{c}$ & Clock model & Clock for the DNA mutation rates\\
-    $\mathit{s}$ & Site model & Nucleotide substitution model\\
-    $\mathit{\mu}$ & Mutation rate & Pace at which mutation occurs, $\mu \in \mathit{s}$ \\
-    $\mathit{r}$ & Root sequence & DNA sequence at the root of the tree\\
-    $\mathit{m}$ & MRCA prior & Prior knowledge of a most recent common ancestor\\
-    $\mathit{v}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup \\
-    \hline
-    $\mathit{G}$ & Generative prior & Diversification model the 'true' tree is simulated with \\
-    $\mathit{A}$ & Alignment model & Alignment simulation model, $\mathit{A} \ni \{ \mathit{r}, \mathit{\mu}, \mathit{c}, \mathit{s} \}$, in which $\mathit{c}$ = strict \\
-    $\mathit{G}$ & Generative model & Combination of a known $\mathit{P}$ and chosen (thus known) $\mathit{A}$ \\
-    $\mathit{I}$ & Inference model & Bayesian phylogenetic inference model, a combination of
-    $(\mathit{c},\mathit{s},\mathit{t},\mathit{m},\mathit{v})$\\
-    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in inference \\
-    $\mathit{X}$ & Experiment & Combination of $(\mathit{I},\mathit{C})$\\
-    $\mathit{E}$ & Error measurement parameters & Errors measurement setup \\
-    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure \\
-    \hline 
-  \end{tabular}
-  \caption{
-    Definitions of terms and relative symbols used in this article.
-  }
-  \label{tab:definitions}
-\end{table}
-\fi
-
-\iffalse
-% previous version
-\begin{table}
-\centering
-  \begin{tabular}{| c | c | l |}
-    \hline
-    \textbf{Symbol} & \textbf{Subparameter} & \textbf{Possible values} \\ 
-    \hline
-    $\mathit{t}$ & name & Yule, BD, CBS, CCP, CEP \\
-    $\mathit{c}$ & clock model & strict, relaxed log-normal \\
-    $\mathit{s}$ & site model & JC, HKY, TN, GTR \\
-    $\mathit{C}$ & model type & Generative, candidate \\
-    $\mathit{C}$ & run if & Always, best candidate \\
-    $\mathit{E}$ & function & nLTT, $|\gamma|$ \\
-    $\mathit{T}$ & model & Yule, BD \\
-    \hline
-  \end{tabular}
-  \caption{
-    Most important parameter options. BD = birth death, CBS = coalescent
-    Bayesian Skyline, CCP = coalescent constant-population, CEP = coalescent
-    exponential-population, JC = Jukes-Cantor 1969, HKY = Hasegawa, Kishino and 
-    Yano 1985, TN = Tamura 1992, GTR = Generalised time-reversible 
-    model (Tavaré 1986)
-  }
-  \label{tab:options}
-\end{table}
-\fi
-
 \subsection{pirouette's pipeline}
@@
 that will be analyzed in the same way as the true tree, following a parallel pipeline.
-We want to produce such tree 'converting' the original tree $\tau_{G}$, created according to the generative prior $\mathit{G}$, to be in accordance with the chosen inference model $\mathit{I}$:
+We want to produce such tree 'converting' the original tree $\tau_{\mathit{P}}$, created according to the generative prior $\mathit{P}$, to be in accordance with the chosen inference model $\mathit{I}$:
 \begin{align}
-    \tau_{G} = (\Vec{t}_{G}, \psi_{G}) \xrightarrow[]{\mathit{T}} \tau_{I} = (\Vec{t}_{I}, \psi_{I}),
+    \tau_{\mathit{P}} = (\Vec{t}_{\mathit{P}}, \psi_{\mathit{P}}) \xrightarrow[]{\mathit{T}} \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, \psi_{\mathit{I}}),
 \end{align}
@@
 
-To execute the twinning process we exploit the likelihood function $L_{I}$ associated with the chosen inference model $\mathit{I}$.
-We find the parameters $\theta^{*}_{I}$ (e.g.
-speciation and extinction rates, in case of a birth-death model) that maximize $L_{I}$ applied to the true tree, conditioned on its number of tips $n_{G}$,
+To execute the twinning process we exploit the likelihood function $L_{\mathit{I}}$ associated with the chosen inference model $\mathit{I}$.
+We find the parameters $\theta^{*}_{\mathit{I}}$ (e.g.
+speciation and extinction rates, in case of a birth-death model) that maximize $L_{\mathit{I}}$ applied to the true tree, conditioned on its number of tips $n_{\mathit{P}}$,
 \begin{align}
-    max[L_{I}(\theta_{I}|\tau_{G}, n_{G})] \rightarrow \theta^{*}_{I}.
+    max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{P}}, n_{\mathit{P}})] \rightarrow \theta^{*}_{\mathit{I}}.
 \end{align}
-We use $\theta^{*}_{I}$ to simulate an equal amount $n_{I} = n_{G}$ of branching times $\Vec{t}_{I}$ for the twin tree $\tau_{I}$, under the process $I$.
-We then impose that $\tau_{G}$ and $\tau_{I}$ must respect the same topology
+We use $\theta^{*}_{\mathit{I}}$ to simulate an equal amount $n_{\mathit{I}} = n_{\mathit{P}}$ of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree $\tau_{\mathit{I}}$, under the process $\mathit{I}$.
+We then impose that $\tau_{\mathit{P}}$ and $\tau_{\mathit{I}}$ must respect the same topology
 \begin{align}
-    \psi_{I} = \psi_{G}.
+    \psi_{\mathit{I}} = \psi_{\mathit{P}}.
 \end{align}
-Using a twin tree serves as a control: both $\tau_{G}$ and $\tau_{I}$ will produce an error distribution.
+Using a twin tree serves as a control: both $\tau_{\mathit{P}}$ and $\tau_{\mathit{I}}$ will produce an error distribution.
 We assume that these errors are a natural product of the input tree and the intermediate steps of the pipeline.
@@
 
-Potential differences in shape of error distributions will therefore be consequences of different tree priors for the two parallel pipelines: the use of prior $\mathit{G}$ for the original pipeline and of prior $\mathit{I}$ for the twin pipeline.
+Potential differences in shape of error distributions will therefore be consequences of different tree priors for the two parallel pipelines: the use of prior $\mathit{P}$ for the original pipeline and of prior $\mathit{I}$ for the twin pipeline.
 In the second case, though, such error distribution is not influenced by the mismatch between how the tree is generated and how it is analyzed by BEAST2.
@@
 
-\verb;pirouette; quantifies the influence of a tree prior,
+\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's inference,
 by measuring the difference between a given tree and 
@@
 will differ from the given phylogeny, even when the generative tree
-prior G and alignment model A are the same as 
-those in the inference model.
-Measuring this difference results allows us to know the baseline error
-of the \verb;pirouette; pipeline.
-Because only standard tree priors are used in inference, 
-and G must be used such a tree prior, it follows that
-G must also be a standard tree prior.
-
-A first research question that \verb;pirouette; can answer is:
+prior $\mathit{P}$ and alignment model $\mathit{A}$ are the same as 
+those in the inference model $\mathit{I}$.
+Measuring this difference allows us to know the baseline error
+of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' all the tree priors that are available to be used within an inference model (see Table~\ref{tab:options}).
+
+Now we can formulate the first example research question that \verb;pirouette; can answer:
 
@@
 
-In this example we use the Yule (pure-birth) tree prior 
-as the generative tree prior G,
-because it is the simplest standard tree prior. 
-Table~\ref{tab:options} shows an overview of the standard
-tree priors.
+In this example we use a standard generative tree prior $\mathit{P_{0}}$, namely the Yule (pure-birth) tree prior. 
 We choose to use a small tree with six taxa to keep
@@
 The first step in \verb;pirouette; is to simulate a DNA alignment from the given phylogeny. 
-To do so, we must specify the DNA root sequence, DNA mutation rate,
-site model (in this context: a nucleotide substitution model) and a clock model. 
+To do so, we must specify an alignment model $\mathit{A}$ which consists of: a DNA root sequence $\mathit{r_{s}}$, the DNA mutation rate $\mathit{m_{r}}$, a
+site model $\mathit{s_{m}}$ (in this context: a nucleotide substitution model) and a clock model $\mathit{c_{m}}$. 
 In this example, the DNA root sequence consists in four blocks of 250 nucleotides each, 
 where the per-nucleotide mutation rate is 0.1 mutations per unit time.
+\iffalse
+\giovanni{this looks unnecessary to me}
 Because the crown age of the phylogeny is 10 time units, each nucleotide will
 have one (possibly-silent) mutation on average.
+\fi
 We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
 and a strict clock model as these are the simplest.
-A Jukes-Cantor (JC, \cite{jukes1969evolution}) site model,
+A Jukes-Cantor site model,
 assumes that mutation rates between nucleotides are equal and constant. 
@@
 
-As the site and clock model used here are also the defaults, 
+As the site and clock models used here are also the defaults, 
 these function arguments can be safely omitted: we just show these
@@
 
-In the second step we state our experiments, which will be just one.
-We define an experiment X as a combination of an inference model I
-and conditions, C.
-In this example we pick I to be the same inference model as the generative one,
-which is Yule tree prior, JC site model and strict clock model.
-We specify C that the experiment will always be run.
+In the second step we state our experiment.
+We define an experiment $\mathit{X}$ as a combination of an inference model $\mathit{I}$
+and conditions $\mathit{C}$.
+In this example we pick $\mathit{I}$ to be the same inference model as the generative one,
+which is the chosen Yule tree prior $\mathit{P_{0}}$, as well as site and clock models defined in $\mathit{A}$, respectively JC and strict.
+We specify in $\mathit{C}$ that the experiment will always be run.
 
@@
 ]
-experiments <- list(
-  create_experiment(
+experiment <- create_experiment(
   inference_conditions = create_inference_conditions(
@@
 )
-)
-\end{lstlisting}
-
-Due to sensible defaults, the same can be achieved by:
+\end{lstlisting}
+
+Experiments must be bundled in a list to work, even if only one is provided, as in this case:
 
@@
   caption = {
-    Create the experiments to answer research question 1, 
-    using the default arguments
+    Create the experiments to answer research question 1.
   }
 ]
-experiments <- list(create_experiment())
+experiments <- list(experiment)
 \end{lstlisting}
@@
 in a \verb;pirouette; parameters structure,
-together with error measurement parameters E and
-twinning parameters T. We use the default error measurement
+together with error measurement parameters $\mathit{E}$ and
+twinning parameters $\mathit{T}$. We use the default error measurement
 parameters, which uses a burn-in of 10\% and the nLTT statistic to
@@
 
+\iffalse
+\giovanni{I think we are providing too many information. Let's try to keep the main text as dry as possible.}
 One the hardware we used, it will take approximately 10 minutes
@@
 \verb;pirouette; has a plotting function that does exactly that:
+\fi
+
+The error distribution returned by \verb;pir_run; is stored as a data frame. It can be directly plotted using \verb;pir_plot;:
 
@@
 
+\iffalse
 This error distribution has little relevance on its own.
+\giovanni{maybe it's better to explain later the multiple tree argument}
 It only gives a baseline error made by
@@
 trees and merge their error distributions.
-
+\fi
+
+\iffalse
 For this error distribution to have more relevance it needs to be compared
@@
 derived from a non-standard speciation process.
-In that way, the error distribution we have now serves as a control,
-as it is the baseline error caused by stochasticity.
+\fi
+
+This error distribution has little relevance on its own, as we constructed the experiment in such a way $\mathit{I}$ matches with the generative model given by $\mathit{A}$ and $\mathit{P_{0}}$.
+However it will become clear when compared to the errors obtained from a phylogeny 
+derived from a non-standard diversification process $\mathit{P}$.
+
+\iffalse
+By doing so, this error distribution can serve as a control,
+as it can be seen as the baseline error caused by stochasticity.
 In this example, that non-standard speciation process must 
-be closest related to a Yule process for the comparison to make sense.
+be closest related to a Yule process for the comparison to make sense \giovanni{why?}.
 Would the two error distributions be similar, the non-standard speciation 
@@
 We call it twinning and we use it in example research question 3.
+\fi
+
+By doing so, this error distribution will serve as a control,
+as it can be seen as the baseline error caused by stochasticity.
+If the two error distributions obtained from $\mathit{P_{0}}$ and $\mathit{P}$ will appear similar, the error induced by the choice of a non-standard speciation process will be of little influence.
+Given the necessity of such comparison,
+we made it an inherent part of \verb;pirouette;. 
+We call it 'twinning' and we use it in ERQ 3.
 
@@
 
+\iffalse
 There are more ways to pick a standard inference model. In the
 previous example, we picked the known generative inference model.
-A novel tree prior, by definition, is non-standard, thus cannot be
-part of a standard inference model. 
+\fi
+In the previous example, we used a known generative tree prior to match with the inference model.
+However, a novel tree prior, by definition, is non-standard, thus cannot be part of a standard inference model (i.e. an inference model using a standard tree prior). 
 One option is to choose to hand-pick a related 

### 2019-03-15 article.tex
Process feedback @Giappo, thanks @Giappo

@@
 
-\giovanni{
-  This is a slightly more formal version. 
-  I personally find it clearer. 
-  If you don't like it we can use version 2 (see above in the source). 
-  In case you really like it, instead, I can try to make the rest a bit more similar to this.
-}
+\giovanni{This is a slightly more formal version}
+\richel{As agreed, go ahead and expend this formalization}
 
@@
 where the per-nucleotide mutation rate is 0.1 mutations per unit time.
-\iffalse
-\giovanni{this looks unnecessary to me}
-Because the crown age of the phylogeny is 10 time units, each nucleotide will
-have one (possibly-silent) mutation on average.
-\fi
 We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
@@
 
-\iffalse
-\giovanni{I think we are providing too many information. Let's try to keep the main text as dry as possible.}
-One the hardware we used, it will take approximately 10 minutes
-to obtain the error distribution returned by \verb;pir_run;.
-This error distribution is a data frame in the non-tidy short form.
-Instead of converting it to tidy data and plotting it,
-\verb;pirouette; has a plotting function that does exactly that:
-\fi
+\giovanni{I think we are providing too many information}
+\richel{agreed, thanks for pushing back}
 
@@
 
-\iffalse
-This error distribution has little relevance on its own.
-\giovanni{maybe it's better to explain later the multiple tree argument}
-It only gives a baseline error made by
-the Bayesian inference. We use 'a', because we have used only one Yule
-tree. To obtain the baseline error, one should simulate multiple Yule
-trees and merge their error distributions.
-\fi
-
-\iffalse
-For this error distribution to have more relevance it needs to be compared
-to an error distribution derived from a phylogeny 
-derived from a non-standard speciation process.
-\fi
-
-This error distribution has little relevance on its own, as we constructed the experiment in such a way $\mathit{I}$ matches with the generative model given by $\mathit{A}$ and $\mathit{P_{0}}$.
+This error distribution has little relevance on its own, 
+as we constructed the experiment in such a way $\mathit{I}$ 
+matches with the generative model given by $\mathit{A}$ and $\mathit{P_{0}}$.
 However it will become clear when compared to the errors obtained from a phylogeny 
 derived from a non-standard diversification process $\mathit{P}$.
-
-\iffalse
-By doing so, this error distribution can serve as a control,
-as it can be seen as the baseline error caused by stochasticity.
-In this example, that non-standard speciation process must 
-be closest related to a Yule process for the comparison to make sense \giovanni{why?}.
-Would the two error distributions be similar, the non-standard speciation 
-process is of little influence.
-Because measuring the baseline error is so important,
-we made it an inherent part of \verb;pirouette;. 
-We call it twinning and we use it in example research question 3.
-\fi
 
@@
 
-\iffalse
-There are more ways to pick a standard inference model. In the
-previous example, we picked the known generative inference model.
-\fi
 In the previous example, we used a known generative tree prior to match with the inference model.
@@
   \label{fig:example_3}
-  \giovanni{remove boxplots}
 \end{figure}
@@
   \label{fig:example_4}
-  \giovanni{Remove boxplots.}
 \end{figure}

### 2019-03-15 article.tex
Review ERQ 3, progress #218

@@
 
-\subsection{Twin tree}
+\subsection{Twin tree}\label{subsec:twinning}
 
@@
 
+In ERQ 1 we measured the baseline error of the \verb;pirouette; method, 
+by using a standard tree prior in both the generation of the tree 
+and in our inference.
+The goal of \verb;pirouette; is to quantify the 
+influence of a new (non-standard) tree prior.
+To do so, we need to compare the inference error 
+from the new tree prior to a baseline error,
+by use of a twin tree (see subsection~\ref{subsec:twinning}).
+In this ERQ, we assume the new tree prior so closely related to
+a standard tree prior, that we hand-pick a standard tree prior to
+serve as a stand-in generative tree prior.
+
 A third research question that \verb;pirouette; can answer is:
@@
 "What is the error made by BEAST2 from a phylogeny, 
-when hand-picking an inference model, compared to the background noise?"
+when hand-picking a tree prior, compared to the background noise?"
 
@@
 the alignment parameters (listing~\ref{lst:create_alignment})
-and the list of experiments (listing~\ref{lst:create_generative_experiment}).
-
-This time, however, we are interested in creating a twin tree,
-as such a twin tree allows us to measure a baseline error produced
-by the BEAST2 run.
-A twin tree
-has the same topology as the given tree, 
-yet its branching times are simulated according 
-to a tree prior already implemented by BEAST2 to produce a posterior. 
-Because for a twin tree we know the tree prior 
-and because we use that tree prior in inference, 
-we expect BEAST2 to produce a posterior of phylogenies
-that is close to the twin tree.
-As the inference model for the twin tree is (more) completely
-specified than the true tree, 
-we expect that a twin tree and its posterior trees are more similar
-than a true tree and its posterior trees.
-Due to this, the twinning is used a control for the experiment.
-
-Creating a twinning parameter is easy, as it has sensible default settings:
+and the list of experiments (listing~\ref{lst:create_generative_experiment}),
+the latter stating that the generative tree prior is a Yule tree prior.
+
+This time, however, we are interested measuring both the inference error
+and a baseline error, which is achieved by creating a twinning parameter
+set. Doing so is easy, as it has sensible default settings:
 
@@
 
-We can now measure the errors made by BEAST2 
-when inferring the given phylogeny 
-versus the error it makes when the same procedure is applied to the twin tree.
-
-All the \verb;pirouette; arguments are bundled and checked by \verb;create_pir_params;:
+Bundling all the \verb;pirouette; arguments by \verb;create_pir_params;:
 
@@
 
-Running:
+Running and plotting the results:
 
@@
 )
-\end{lstlisting}
-
-Again, showing the results:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
 pir_plot(errors)
@@
   \caption{
-    Example 3: the error made by BEAST2 from a phylogeny, picking the best inference model versus the control.
-    Here, the twin column shows the error BEAST2 makes starting from the original tree versus the twin. 
+    Example 3: the error made by BEAST2 from a phylogeny, 
+    when hand-picking a generative tree prior versus the control.
+    Here, the 'twin' column shows the baseline inference error,
+    as measured on the twin tree
   }
@@
 \end{figure}
-
-\iffalse
-The error BEAST2 makes from a phylogeny 
-picking the best inference model, compared to the background noise.
-Here, the twin column shows the error BEAST2 makes on an idealized
-tree, to measure the noise, which is the minimal error. 
-\fi
 

### 2019-03-18 article.tex
Updates from Overleaf

--- 2019-03-15 article.tex
@@
 
+\subsection{Question}
+
 \verb;pirouette; quantifies the influence of a new tree prior on BEAST2's inference,
@@
 
+\subsection{Answer}
+
 In this example we use a standard generative tree prior $\mathit{P_{0}}$, namely the Yule (pure-birth) tree prior. 
@@
       Create a Yule tree. 
-      The resulting tree is shown in figure~\ref{fig:yule_tree}
+      The resulting tree is shown in figure~\ref{fig:yule_tree}.
     }
@@
   \caption{
-    Experimental setup to answer the first research question.
+    Inference conditions and inference model for ERQ 1.
     JC: Jukes-Cantor site model.
@@
   }
-  \label{tbl:RQ1}
+  \label{tab:RQ1}
 \end{table}
@@
 
+\subsection{Question}
+
+\iffalse
 In the previous example, we used a known generative tree prior to match with the inference model.
@@
 One option is to choose to hand-pick a related 
-standard tree prior, as done in ERCs 3 and 5, and combine it with 
+standard tree prior, as done in ERQs 3 and 5, and combine it with 
 a known generative alignment model.
@@
 penalized for having more parameters.
-
+\fi
+
+In the previous example we selected the inference model $\mathit{I}$ to match with a known generative tree prior $\mathit{P_{0}}$.
+However, a novel tree prior $\mathit{P}$, by definition, is non-standard, thus cannot be part of a standard inference model (i.e. an inference model using a standard tree prior).
+In this case the user has to specify the the tree prior $\mathit{t_{p}}$ to be used in the inference from a list of standard tree priors.
+In general, having no information on $\mathit{P}$, this choice is arbitrary.
+\verb;pirouette; provides two ways to deal with this problem. The choice can be made specifying the inference conditions $\mathit{C}$ to be used.
+
+If the user has sufficient reasons to believe that one standard tree prior could be more suitable than others for the inference, they can set it as part of the inference model and to adopt site and clock models used to simulate the alignment.
+This option can be specified selecting "model\_type" to be "generative" as part of the inference conditions (as seen in ERQ 1, table~\ref{tab:RQ1}).
+
+If, instead, there is no preference for a specific tree prior, the user can let \verb;pirouette; select the best inference model from a list of candidate models.
+The model selection is performed automatically based on the criterion of the highest evidence (aka marginal likelihood).
+This option can be specified providing any number of experiments, selecting "model\_type" to be "candidate" for each of them, and "run\_if" to be "best\_candidate" (table~\ref{tab:RQ2} will summarize this for ERQ 2).
+
+\iffalse
+One option is to choose to hand-pick a related 
+standard tree prior, as done in ERQs 3 and 5, and combine it with 
+a known generative alignment model.
+An alternative is to let multiple candidate inference models compete 
+and pick the best one, based on the highest evidence (aka marginal likelihood):
+when an alignment is simulated from the given tree, one can estimate the
+evidence for each inference model given that 
+alignment \richel{or the other way around?}.
+The benefit of this approach is that this is a commonly used procedure,
+which picks an inference model that is
+best at explaining the alignment, where more complex inference models are
+penalized for having more parameters.
+\fi
+\iffalse
 A second research question that \verb;pirouette; can answer, is:
+\fi
+
+We show the procedure using \verb;pirouette; to answer the second research question:
 
@@
 picking the best inference model?"
+
+\subsection{Answer}
 
@@
   \caption{
-    Experimental setup to answer the second research question.
+    Inference conditions and inference model for ERQ 2.
     JC: Jukes-Cantor site model.
@@
   }
+  \label{tab:RQ2}
 \end{table}
 
+\iffalse
 In the second step we set up our experiments. 
@@
 inference model $I$, but has condition $C$ that will only let $I$
-be run in inference if that inference model $I$ is the 
+be run in inference only if that inference model $I$ is the 
 best candidate inference model, that is, 
@@
 models, 2 clock models and 5 tree priors.
+\fi
+
+In the second step we set up our experiments.
+Unlike ERQ 1, here we provide a set of experiments $\{\mathit{X_{i}}\}_{i=1}^{40}$.
+All the experiments have the same inference conditions $\mathit{C_{i}} = \mathit{C}$ but different candidate inference models $\mathit{I_{i}}$.
+Yet, only one of them will be used to run the actual inference.
+According to the specified $\mathit{C}$, \verb;pirouette; will measure the evidence for each of them and select  only the one with the highest evidence as the best candidate.
+
+Setting up experiments like this is thought to be so common,
+that \verb;pirouette; has a simple function to do:
+\verb;create_all_experiments; will create the entire set of 40 experiments,
+containing the inference models of all combinations of 4 site 
+models, 2 clock models and 5 tree priors.
 
@@
   label = {lst:create_all_experiments},
-  caption = Create all 40 candidate experiments
+  caption = Create all 40 candidate experiments.
 ]
@@
   label = {lts:create_pir_params_defaults},
-  caption = Create a \texttt{pir\_params} with many defaults
+  caption = Create a \texttt{pir\_params} with many defaults.
 ]
@@
 
-A third research question that \verb;pirouette; can answer is:
+\subsection{Question}
+
+In ERQ 1 we built the twin pipeline.
+At the end of the section we discussed how it has not much relevance if considered as a stand-alone process.
+Here, instead, we will use it as a control for the main pipeline, where the generative tree prior $\mathit{P}$ is unknown, but the site and clock models used for the inference are the ones used to simulate the alignment.
+This leads to the third ERQ:
 
@@
 when hand-picking an inference model, compared to the background noise?"
+
+\subsection{Answer}
 
@@
 
+\iffalse
 This time, however, we are interested in creating a twin tree,
@@
 Due to this, the twinning is used a control for the experiment.
+\fi
+
+This time, however, we are interested in creating a twin tree. A twin tree
+has the same topology as the given tree, yet its branching times are simulated according to the standard tree prior used by BEAST2 to produce the posterior for the main pipeline. 
+Since there is coherence between the inference model and the generative model, we expect BEAST2 to produce a posterior of phylogenies more similar to the original one, respect to the outcome of the main pipeline.
+Due to this, the twinning process is used to provide a control for the main experiment.
 
@@
 
-A fourth research question that \verb;pirouette; can answer, is:
+\subsection{Question}
+
+To answer ERQ 2 we used a evidence-based model selection procedure to obtain an error distribution.
+However, it is unclear to what extent this error is due to the mismatch between the tree priors (from generative and inference models) or due to stochasticity.
+As mentioned in ERQ 1, \verb;pirouette; provides a tool to evaluate this, using the twinning process.
+We now apply it to the same pipeline followed in ERQ 2.
+This leads to our fourth example research question:
 
@@
 
+\subsection{Answer}
+
 Here, same as example 2 and 3, we start with the phylogeny 
@@
 
-A fifth research question that \verb;pirouette; can answer, is:
+\subsection{Question}
+
+In the previous research question we have seen how to use \verb;pirouette; to  run the analysis on main pipeline, either using a 'generative' or a 'best\_candidate' inference model.
+We also saw that to exclude the baseline error due to stochasticity it is possible to run a parallel twin pipeline.
+In this way two error distributions are obtained.
+Comparing them it is possible to estimate the error made by BEAST2 when inferring a posterior distribution starting from a phylogeny generated under an unknown generative tree prior.
+We can sum up all \verb;pirouette;'s features presented so far in the fifth research question:
 
@@
 in relationship to the background error?"
+
+\subsection{Answer}
 

### 2019-03-18 article.tex

@@
 
+In ERQ 1 we measured the baseline error of the \verb;pirouette; method, 
+by using a standard tree prior in both the generation of the tree 
+and in our inference.
+The goal of \verb;pirouette; is to quantify the 
+influence of a new (non-standard) tree prior.
+To do so, we need to compare the inference error 
+from the new tree prior to a baseline error,
+by use of a twin tree (see subsection~\ref{subsec:twinning}).
+In this ERQ, we assume the new tree prior so closely related to
+a standard tree prior, that we hand-pick a standard tree prior to
+serve as a stand-in generative tree prior.
+
 \subsection{Question}
@@
 the alignment parameters (listing~\ref{lst:create_alignment})
-and the list of experiments (listing~\ref{lst:create_generative_experiment}).
-
-\iffalse
-This time, however, we are interested in creating a twin tree,
-as such a twin tree allows us to measure a baseline error produced
-by the BEAST2 run.
-A twin tree
-has the same topology as the given tree, 
-yet its branching times are simulated according 
-to a tree prior already implemented by BEAST2 to produce a posterior. 
-Because for a twin tree we know the tree prior 
-and because we use that tree prior in inference, 
-we expect BEAST2 to produce a posterior of phylogenies
-that is close to the twin tree.
-As the inference model for the twin tree is (more) completely
-specified than the true tree, 
-we expect that a twin tree and its posterior trees are more similar
-than a true tree and its posterior trees.
-Due to this, the twinning is used a control for the experiment.
-\fi
-
+and the list of experiments (listing~\ref{lst:create_generative_experiment}),
+the latter stating that the generative tree prior is a Yule tree prior.
+
+This time, however, we are interested measuring both the inference error
+and a baseline error, which is achieved by creating a twinning parameter
+set. 
 This time, however, we are interested in creating a twin tree. A twin tree

### 2019-03-18 article.tex
Merge introduction of ERQ 3

@@
 by use of a twin tree (see subsection~\ref{subsec:twinning}).
-In this ERQ, we assume the new tree prior so closely related to
-a standard tree prior, that we hand-pick a standard tree prior to
-serve as a stand-in generative tree prior.
+To do so, we use the twin tree as a control for the main pipeline.
+In this example, we do not know the generative tree prior $\mathit{P}$,
+but we do know the alignment model, $\mathit{A}$.
+In this ERQ, we assume the new tree prior is so closely related to
+a standard tree prior, that it is reasonable to hand-pick a standard tree prior 
+to serve as a stand-in generative tree prior.
 
 \subsection{Question}
-
-In ERQ 1 we built the twin pipeline.
-At the end of the section we discussed how it has not much relevance if considered as a stand-alone process.
-Here, instead, we will use it as a control for the main pipeline, where the generative tree prior $\mathit{P}$ is unknown, but the site and clock models used for the inference are the ones used to simulate the alignment.
-This leads to the third ERQ:
 

### 2019-03-18 article.tex
Remove all iffalse sections

@@
 
-\iffalse
-%version 1
-An optional step is to generate a 'twin tree' 
-(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
-that will be analyzed in the same way as the true tree.
-The twin tree will be the true tree 
-converted to follow the tree prior of the chosen inference model.
-
-To do so we exploit the likelihood function associated with the chosen inference model.
-We find the parameters (e.g.
-speciation and extinction rate for a birth-death model) that maximize such likelihood applied to the true tree, conditioned on its number of tips.
-
-We use these parameters to simulate the branching times of the twin tree, 
-which we combine with the original tree's topology.
-
-Using a twin tree serves as a control: both the given/true tree
-and twin tree will produce an error distribution.
-For the true tree, this error distribution is caused by
-the shape of the phylogeny (which is different from the tree priors used), 
-the stochastic alignment simulation and stochastic Bayesian inference.  
-The twin tree is created from the given phylogeny,
-as it has the same topology, 
-yet its branching times are simulated according to 
-the (implemented) generative tree prior.
-Due to this, this error distribution for the twin tree is caused 
-only by the stochastic alignment simulation and stochastic Bayesian inference.
-The difference between these two error distribution is the error caused
-by the new/unimplemented tree prior.
-\fi
-
-\iffalse
-%version 2 with no formulae
-An optional step is to generate a 'twin tree' 
-(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
-that will be analyzed in the same way as the true tree.
-We want to produce such tree 'converting' the original tree, created according to the generative prior $\mathit{G}$, into the chosen inference model $\mathit{I}$.
-
-To do so we exploit the likelihood function associated with the chosen inference model $\mathit{I}$.
-We find the parameters (e.g.
-speciation and extinction rate for a birth-death model) that maximize such likelihood applied to the true tree, conditioned on its number of tips.
-
-We use these parameters to simulate an equal amount of branching times for the twin tree simulated under the chosen inference model's process.
-We then combine these branching times with the original tree's topology.
-
-Using a twin tree serves as a control: both the given/true tree and twin tree will produce an error distribution.
-We assume that these errors are a natural product of the input tree and the intermediate steps of the pipeline.
-All the factors that can cause such errors are: the branching times of the tree (as a consequence of the tree prior), the topology of the tree, the stochasticity induced by simulating the alignment (step 1a $\rightarrow$ 2a in Fig.~\ref{fig:pipeline}) and the stochasticity induced by the Bayesian inference (step 2a $\rightarrow$ 4a in Fig.~\ref{fig:pipeline}).
-By imposing the same topology and replicating the pipeline a sufficient number of times we can sort out all the possible causes that are not a direct consequence of the choice of the tree prior.
-
-Potential differences in shapes of error distributions will therefore be consequences of different tree priors for the two parallel pipelines: the use of prior $\mathit{G}$ for the original pipeline and of prior $\mathit{I}$ for the twin pipeline.
-In the second case, though, such error distribution is not influenced by the mismatch between how the tree is generated and how it is analyzed by BEAST2.
-\fi
-
-\giovanni{This is a slightly more formal version}
-\richel{As agreed, go ahead and expend this formalization}
-
-%version 3, more formal
 An optional step is to generate a 'twin tree' $\tau_{I}$
@@
 
-\iffalse
-In the previous example, we used a known generative tree prior to match with the inference model.
-However, a novel tree prior, by definition, is non-standard, thus cannot be part of a standard inference model (i.e. an inference model using a standard tree prior).
-One option is to choose to hand-pick a related 
-standard tree prior, as done in ERQs 3 and 5, and combine it with 
-a known generative alignment model.
-An alternative is to let multiple candidate inference models compete 
-and pick the best one, based on the highest evidence (aka marginal likelihood):
-when an alignment is simulated from the given tree, one can estimate the
-evidence for each inference model given that 
-alignment \richel{or the other way around?}.
-The benefit of this approach is that this is a commonly used procedure,
-which picks an inference model that is
-best at explaining the alignment, where more complex inference models are
-penalized for having more parameters.
-\fi
-
 In the previous example we selected the inference model $\mathit{I}$ to match with a known generative tree prior $\mathit{P_{0}}$.
@@
 This option can be specified providing any number of experiments, selecting "model\_type" to be "candidate" for each of them, and "run\_if" to be "best\_candidate" (table~\ref{tab:RQ2} will summarize this for ERQ 2).
-
-\iffalse
-One option is to choose to hand-pick a related 
-standard tree prior, as done in ERQs 3 and 5, and combine it with 
-a known generative alignment model.
-An alternative is to let multiple candidate inference models compete 
-and pick the best one, based on the highest evidence (aka marginal likelihood):
-when an alignment is simulated from the given tree, one can estimate the
-evidence for each inference model given that 
-alignment \richel{or the other way around?}.
-The benefit of this approach is that this is a commonly used procedure,
-which picks an inference model that is
-best at explaining the alignment, where more complex inference models are
-penalized for having more parameters.
-\fi
-\iffalse
-A second research question that \verb;pirouette; can answer, is:
-\fi
 
@@
 \end{table}
-
-\iffalse
-In the second step we set up our experiments. 
-Within this ERQ, each of our experiments contain a different
-inference model $I$, but has condition $C$ that will only let $I$
-be run in inference only if that inference model $I$ is the 
-best candidate inference model, that is, 
-it is the inference model with the highest evidence.
-Setting up experiments like this is thought to be so common,
-that \verb;pirouette; has a simple function to do:
-\verb;create_all_experiments; will create 40 experiments,
-containing the inference models of all combinations of 4 site 
-models, 2 clock models and 5 tree priors.
-\fi
 

### 2019-03-18 article.tex
80 chars per line max

@@
 % (2) no need to specify BEAST2, it's also just a means
-\title{pirouette: Quantifying the importance of a tree prior in Bayesian phylogenetics}
+\title{pirouette: Quantifying the importance of a tree prior in Bayesian 
+phylogenetics}
 % \title{pirouette: Determining the error BEAST2 makes in inferring a phylogeny}
@@
 \author[1]{Rampal S. Etienne}
-\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands}
+\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of 
+Groningen, Groningen, The Netherlands}
 
@@
 % Style of listings
-% From http://r.789695.n4.nabble.com/How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
+% From http://r.789695.n4.nabble.com/
+%   How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
 \usepackage{fancyvrb} 
@@
   tabsize=2,% sizes of tabs
-  showstringspaces=false,% do not replace spaces in strings by a certain character
+  showstringspaces=false,
   captionpos=b,% positioning of the caption below
@@
 
-{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, pirouette, R
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, 
+pirouette, R
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 of a tree given such a diversification model.
-There have been plenty of these models (and their associated probability algorithms) 
+There have been plenty of these models (and their associated probability 
+algorithms) 
 developed, among others, those in which diversification is 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Description}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
     \centering
-    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} & \textbf{Possible values} \\ 
-    \hline
-    $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model & Yule, BD, CBS, CCP, CEP \\
-    $\mathit{c_{m}}$ & clock model & Clock for the DNA mutation rates & strict, relaxed log-normal \\
-    $\mathit{s_{m}}$ & site model & Nucleotide substitution model & JC, HKY, TN93, GTR \\
-    $\mathit{m_{r}}$ & mutation rate & Pace at which mutation occurs & $m_{r} \in \mathbb{R}_{>0}$\\
-    $\mathit{r_{s}}$ & root sequence & DNA sequence at the root of the tree & any combination of a, c, g, t \\
-    $\mathit{m_{p}}$ & MRCA prior & Prior knowledge of a most recent common ancestor & see documentation \\
-    $\mathit{m_{s}}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup & see documentation \\
-    $\mathit{m_{t}}$ & model type & Criterion to select an inference model & Generative, Candidate \\
-    $\mathit{r_{i}}$ & run if & Condition under which an inference model is used & Always, Best candidate\\
-    $\mathit{e_{f}}$ & error function & Specifies how to measure the error & nLTT, $|\gamma|$ \\
-    $\mathit{t_{m}}$ & twin tree prior & Sets the tree prior for the twin tree & Yule, BD \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} & 
+\textbf{Possible values} \\ 
+    \hline
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model 
+      & Yule, BD, CBS, CCP, CEP \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{c_{m}}$ & clock model & Clock for the DNA mutation rates 
+      & strict, relaxed log-normal \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{s_{m}}$ & site model & Nucleotide substitution model 
+      & JC, HKY, TN93, GTR \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{m_{r}}$ & mutation rate & Pace at which mutation occurs 
+      & $m_{r} \in \mathbb{R}_{>0}$\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{r_{s}}$ & root sequence & DNA sequence at the root of the tree 
+      & any combination of a, c, g, t \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{m_{p}}$ & MRCA prior & Prior knowledge of a most recent common 
+      ancestor & see documentation \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{m_{s}}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup 
+      & see documentation \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{m_{t}}$ & model type & Criterion to select an inference model 
+      & Generative, Candidate \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{r_{i}}$ & run if & Condition under which an inference model is 
+      used & Always, Best candidate\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{e_{f}}$ & error function & Specifies how to measure the error 
+      & nLTT, $|\gamma|$ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{t_{m}}$ & twin tree prior & Sets the tree prior for the twin tree 
+      & Yule, BD \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \hline
@@
     \centering
-    \textbf{Symbol} & \textbf{Macro-argument} & \textbf{Description} & \textbf{Includes} \\
-    \hline
-    $\mathit{P}$ & Generative tree prior & Diversification model to generate the 'true' tree & $\mathit{t_{p}}$ \\
-    $\mathit{A}$ & Alignment model & Model used to simulate the alignment ($\mathit{c_{m}}$ must be strict) & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{m_{r}}, \mathit{r_{s}}$ \\
-    $\mathit{G}$ & Generative model & The full setting to produce BEAST2 input data & $\mathit{P}$, $\mathit{A}$ \\
-    $\mathit{I}$ & Inference model & Phylogenetic inference model to run BEAST2 & $\mathit{c_{m}},\mathit{s_{m}},\mathit{t_{p}},\mathit{m_{p}},\mathit{m_{s}}$ \\
-    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference & $\mathit{m_{t}}, \mathit{r_{i}}$\\
-    $\mathit{X}$ & Experiment & Full setting for the BEAST2 run & $\mathit{I},\mathit{C}$ \\
-    $\mathit{E}$ & Error measure parameters & Errors measurement setup & $\mathit{e_{f}}$\\
-    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & $\mathit{t_{m}}$\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{Symbol} & \textbf{Macro-argument} & \textbf{Description} 
+      & \textbf{Includes} \\
+    \hline
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{P}$ & Generative tree prior & Diversification model to generate 
+      the 'true' tree & $\mathit{t_{p}}$ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{A}$ & Alignment model & Model used to simulate the 
+      alignment ($\mathit{c_{m}}$ must be strict) 
+      & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{m_{r}}, \mathit{r_{s}}$ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{G}$ & Generative model & The full setting to produce BEAST2 input 
+      data & $\mathit{P}$, $\mathit{A}$ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{I}$ & Inference model & Phylogenetic inference model to run BEAST2 
+      & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{t_{p}}, 
+      \mathit{m_{p}}, \mathit{m_{s}}$ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ 
+      is used in the inference & $\mathit{m_{t}}, \mathit{r_{i}}$\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{X}$ & Experiment & Full setting for the BEAST2 run & 
+      $\mathit{I},\mathit{C}$ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{E}$ & Error measure parameters & Errors measurement setup & 
+      $\mathit{e_{f}}$\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & 
+      $\mathit{t_{m}}$\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \hline 
@@
   \caption{
-    Definitions of terms and relative symbols used in the main text and in Fig~\ref{fig:pipeline}.
+    Definitions of terms and relative symbols used in the main text and in 
+    Fig~\ref{fig:pipeline}.
   }
@@
   \richel{
-    As the twinning parameters contain more than only the twin model's tree prior,
-    I suggest to keep it 'twinning parameters'
+    As the twinning parameters contain more than only the twin model's tree 
+    prior, I suggest to keep it 'twinning parameters'
   }
@@
     \texttt{pirouette} pipeline. 
-    The pipeline starts from a phylogeny (1a) simulated by generative prior $\mathit{P}$.
-    The phylogeny is converted to an alignment (2a) using alignment model $\mathit{A}$. 
+    The pipeline starts from a phylogeny (1a) simulated by generative prior 
+$\mathit{P}$.
+    The phylogeny is converted to an alignment (2a) using alignment model 
+$\mathit{A}$. 
     The user supplies one or more experiments.
@@
 \begin{enumerate}
-    \item for a given phylogeny an alignment is simulated under a known alignment model;
-    \item for this alignment, according to the specified inference conditions, an inference model is picked (which may be different from the generative alignment model);
-    \item the alignment is then used as BEAST2 input to infer a posterior distribution of phylogenies;
-    \item posterior phylogenies are compared with the given original phylogeny to estimate the error made, according to the error measure specified by the user;
+    \item for a given phylogeny an alignment is simulated under a known 
+alignment model;
+    \item for this alignment, according to the specified inference conditions, 
+an inference model is picked (which may be different from the generative 
+alignment model);
+    \item the alignment is then used as BEAST2 input to infer a posterior 
+distribution of phylogenies;
+    \item posterior phylogenies are compared with the given original phylogeny 
+to estimate the error made, according to the error measure specified by the 
+user;
 \end{enumerate}
@@
 The inference model can be selected in two ways, either using the
-generative model or selecting one or more candidates out of a list of candidate inference models.
-An experiment can use the generative model as the inference model, if the generative model is known.
+generative model or selecting one or more candidates out of a list of candidate 
+inference models.
+An experiment can use the generative model as the inference model, if the 
+generative model is known.
 An experiment uses a candidate model if either the generative model is unknown,
@@
 the generative model.
-Typically, a generative model and/or a best candidate model are used in the actual inference.
+Typically, a generative model and/or a best candidate model are used in the 
+actual inference.
 When using an experiment with a generative model,
@@
 approach, as described in \cite{maturana2018model}. The nested sampling is
-performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 package. 
+performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 
+package. 
 Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac,
-therefore the selection of candidate models is not available directly on Windows. 
-However, even on Windows systems there is the option to perform the model selection from browser, using \verb;mcbette;, yet the computation for this model selection is
+therefore the selection of candidate models is not available directly on 
+Windows. 
+However, even on Windows systems there is the option to perform the model 
+selection from browser, using \verb;mcbette;, yet the computation for this 
+model selection is
 restricted to one hour for free users of the host (Travis CI) continuous
@@
 using the inference models from the experiments selected in the previous step. 
-For each selected experiment, one BEAST2 posterior is inferred, using the \verb;babette; [\cite{bilderbeek2018babette}] R package.
-
-The fourth step compares the true tree to each posterior tree, resulting in an error distribution (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
-The nLTT statistic (\cite{janzen2015approximate}) is used by default, but also a user-defined error statistic can be used. 
+For each selected experiment, one BEAST2 posterior is inferred, using the 
+\verb;babette; [\cite{bilderbeek2018babette}] R package.
+
+The fourth step compares the true tree to each posterior tree, resulting in an 
+error distribution (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
+The nLTT statistic (\cite{janzen2015approximate}) is used by default, but also 
+a user-defined error statistic can be used. 
 As an example, \verb;pirouette; supplies one custom error statistic,
-that uses an absolute difference in the gamma statistic [\cite{pybus2000testing}].
+that uses an absolute difference in the gamma statistic 
+[\cite{pybus2000testing}].
 Additionally, the user can specify the proportion of posterior phylogenies to 
@@
 (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
-that will be analyzed in the same way as the true tree, following a parallel pipeline.
-We want to produce such tree 'converting' the original tree $\tau_{\mathit{P}}$, created according to the generative prior $\mathit{P}$, to be in accordance with the chosen inference model $\mathit{I}$:
+that will be analyzed in the same way as the true tree, following a parallel 
+pipeline. We want to produce such tree 'converting' the original tree 
+$\tau_{\mathit{P}}$, created according to the generative prior $\mathit{P}$, to 
+be in accordance with the chosen inference model $\mathit{I}$:
 \begin{align}
-    \tau_{\mathit{P}} = (\Vec{t}_{\mathit{P}}, \psi_{\mathit{P}}) \xrightarrow[]{\mathit{T}} \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, \psi_{\mathit{I}}),
+    \tau_{\mathit{P}} = (\Vec{t}_{\mathit{P}}, \psi_{\mathit{P}}) 
+\xrightarrow[]{\mathit{T}} \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, 
+\psi_{\mathit{I}}),
 \end{align}
-where $\Vec{t}$ are the tree branching times and $\psi$ is tree topology, which provides information on where new lineages branch out at each branching time.
-
-To execute the twinning process we exploit the likelihood function $L_{\mathit{I}}$ associated with the chosen inference model $\mathit{I}$.
-We find the parameters $\theta^{*}_{\mathit{I}}$ (e.g.
-speciation and extinction rates, in case of a birth-death model) that maximize $L_{\mathit{I}}$ applied to the true tree, conditioned on its number of tips $n_{\mathit{P}}$,
+where $\Vec{t}$ are the tree branching times and $\psi$ is tree topology, which 
+provides information on where new lineages branch out at each branching time.
+
+To execute the twinning process we exploit the likelihood function 
+$L_{\mathit{I}}$ associated with the chosen inference model $\mathit{I}$. We 
+find the parameters $\theta^{*}_{\mathit{I}}$ (e.g. speciation and extinction 
+rates, in case of a birth-death model) that maximize $L_{\mathit{I}}$ applied 
+to the true tree, conditioned on its number of tips $n_{\mathit{P}}$,
 \begin{align}
-    max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{P}}, n_{\mathit{P}})] \rightarrow \theta^{*}_{\mathit{I}}.
+    max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{P}}, n_{\mathit{P}})] 
+\rightarrow \theta^{*}_{\mathit{I}}.
 \end{align}
-We use $\theta^{*}_{\mathit{I}}$ to simulate an equal amount $n_{\mathit{I}} = n_{\mathit{P}}$ of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree $\tau_{\mathit{I}}$, under the process $\mathit{I}$.
-We then impose that $\tau_{\mathit{P}}$ and $\tau_{\mathit{I}}$ must respect the same topology
+We use $\theta^{*}_{\mathit{I}}$ to simulate an equal amount $n_{\mathit{I}} = 
+n_{\mathit{P}}$ of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree 
+$\tau_{\mathit{I}}$, under the process $\mathit{I}$. We then impose that 
+$\tau_{\mathit{P}}$ and $\tau_{\mathit{I}}$ must respect the same topology
 \begin{align}
@@
 \end{align}
-Using a twin tree serves as a control: both $\tau_{\mathit{P}}$ and $\tau_{\mathit{I}}$ will produce an error distribution.
-We assume that these errors are a natural product of the input tree and the intermediate steps of the pipeline.
-All the factors that can cause such errors are: the branching times $\vec{t}$ of the tree (as a consequence of the tree prior), its topology $\psi$, the stochasticity induced by simulating the alignment (step 1a $\rightarrow$ 2a in Fig.~\ref{fig:pipeline}) and the stochasticity induced by the Bayesian inference (step 2a $\rightarrow$ 4a in Fig.~\ref{fig:pipeline}).
-By imposing the same topology and replicating the pipelines a sufficient number of times we can sort out all the possible causes that are not a direct consequence of the choice of the tree prior.
-
-Potential differences in shape of error distributions will therefore be consequences of different tree priors for the two parallel pipelines: the use of prior $\mathit{P}$ for the original pipeline and of prior $\mathit{I}$ for the twin pipeline.
-In the second case, though, such error distribution is not influenced by the mismatch between how the tree is generated and how it is analyzed by BEAST2.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Using a twin tree serves as a control: both $\tau_{\mathit{P}}$ and 
+$\tau_{\mathit{I}}$ will produce an error distribution. We assume that these 
+errors are a natural product of the input tree and the intermediate steps of 
+the pipeline. All the factors that can cause such errors are: the branching 
+times $\vec{t}$ of the tree (as a consequence of the tree prior), its topology 
+$\psi$, the stochasticity induced by simulating the alignment (step 1a 
+$\rightarrow$ 2a in Fig.~\ref{fig:pipeline}) and the stochasticity induced by 
+the Bayesian inference (step 2a $\rightarrow$ 4a in Fig.~\ref{fig:pipeline}). 
+By imposing the same topology and replicating the pipelines a sufficient number 
+of times we can sort out all the possible causes that are not a direct 
+consequence of the choice of the tree prior.
+
+Potential differences in shape of error distributions will therefore be 
+consequences of different tree priors for the two parallel pipelines: the use 
+of prior $\mathit{P}$ for the original pipeline and of prior $\mathit{I}$ for 
+the twin pipeline. In the second case, though, such error distribution is not 
+influenced by the mismatch between how the tree is generated and how it is 
+analyzed by BEAST2.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Installation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 For the most up-to-date version, 
-one can download and install the package from \verb;pirouette;'s GitHub repository:
+one can download and install the package from \verb;pirouette;'s GitHub 
+repository:
 
@@
 
-To start using \verb;pirouette;, load its functions in the global namespace first:
+To start using \verb;pirouette;, load its functions in the global namespace 
+first:
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{table}[h]
@@
 \end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Usage: Example Research Question 1}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's inference,
+\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's 
+inference,
 by measuring the difference between a given tree and 
@@
 Measuring this difference allows us to know the baseline error
-of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' all the tree priors that are available to be used within an inference model (see Table~\ref{tab:options}).
-
-Now we can formulate the first example research question that \verb;pirouette; can answer:
+of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' 
+all the tree priors that are available to be used within an inference model 
+(see Table~\ref{tab:options}).
+
+Now we can formulate the first example research question that \verb;pirouette; 
+can answer:
 
@@
 
-In this example we use a standard generative tree prior $\mathit{P_{0}}$, namely the Yule (pure-birth) tree prior. 
+In this example we use a standard generative tree prior $\mathit{P_{0}}$, 
+namely the Yule (pure-birth) tree prior. 
 We choose to use a small tree with six taxa to keep
@@
 
-The first step in \verb;pirouette; is to simulate a DNA alignment from the given phylogeny. 
-To do so, we must specify an alignment model $\mathit{A}$ which consists of: a DNA root sequence $\mathit{r_{s}}$, the DNA mutation rate $\mathit{m_{r}}$, a
-site model $\mathit{s_{m}}$ (in this context: a nucleotide substitution model) and a clock model $\mathit{c_{m}}$. 
-In this example, the DNA root sequence consists in four blocks of 250 nucleotides each, 
+The first step in \verb;pirouette; is to simulate a DNA alignment from the 
+given phylogeny. 
+To do so, we must specify an alignment model $\mathit{A}$ which consists of: a 
+DNA root sequence $\mathit{r_{s}}$, the DNA mutation rate $\mathit{m_{r}}$, a
+site model $\mathit{s_{m}}$ (in this context: a nucleotide substitution model) 
+and a clock model $\mathit{c_{m}}$. 
+In this example, the DNA root sequence consists in four blocks of 250 
+nucleotides each, 
 where the per-nucleotide mutation rate is 0.1 mutations per unit time.
@@
 In the second step we state our experiment.
-We define an experiment $\mathit{X}$ as a combination of an inference model $\mathit{I}$
+We define an experiment $\mathit{X}$ as a combination of an inference model 
+$\mathit{I}$
 and conditions $\mathit{C}$.
-In this example we pick $\mathit{I}$ to be the same inference model as the generative one,
-which is the chosen Yule tree prior $\mathit{P_{0}}$, as well as site and clock models defined in $\mathit{A}$, respectively JC and strict.
+In this example we pick $\mathit{I}$ to be the same inference model as the 
+generative one,
+which is the chosen Yule tree prior $\mathit{P_{0}}$, as well as site and clock 
+models defined in $\mathit{A}$, respectively JC and strict.
 We specify in $\mathit{C}$ that the experiment will always be run.
@@
     \hline
-    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & \textbf{inference model} \\ 
+    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & 
+\textbf{inference model} \\ 
     \hline
@@
 
-Experiments must be bundled in a list to work, even if only one is provided, as in this case:
+Experiments must be bundled in a list to work, even if only one is provided, as 
+in this case:
 
@@
 
-The error distribution returned by \verb;pir_run; is stored as a data frame. It can be directly plotted using \verb;pir_plot;:
+The error distribution returned by \verb;pir_run; is stored as a data frame. It 
+can be directly plotted using \verb;pir_plot;:
 
@@
 matches with the generative model given by $\mathit{A}$ and $\mathit{P_{0}}$.
-However it will become clear when compared to the errors obtained from a phylogeny 
+However it will become clear when compared to the errors obtained from a 
+phylogeny 
 derived from a non-standard diversification process $\mathit{P}$.
@@
 as it can be seen as the baseline error caused by stochasticity.
-If the two error distributions obtained from $\mathit{P_{0}}$ and $\mathit{P}$ will appear similar, the error induced by the choice of a non-standard speciation process will be of little influence.
+If the two error distributions obtained from $\mathit{P_{0}}$ and $\mathit{P}$ 
+will appear similar, the error induced by the choice of a non-standard 
+speciation process will be of little influence.
 Given the necessity of such comparison,
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Usage: Example Research Question 2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-In the previous example we selected the inference model $\mathit{I}$ to match with a known generative tree prior $\mathit{P_{0}}$.
-However, a novel tree prior $\mathit{P}$, by definition, is non-standard, thus cannot be part of a standard inference model (i.e. an inference model using a standard tree prior).
-In this case the user has to specify the the tree prior $\mathit{t_{p}}$ to be used in the inference from a list of standard tree priors.
-In general, having no information on $\mathit{P}$, this choice is arbitrary.
-\verb;pirouette; provides two ways to deal with this problem. The choice can be made specifying the inference conditions $\mathit{C}$ to be used.
-
-If the user has sufficient reasons to believe that one standard tree prior could be more suitable than others for the inference, they can set it as part of the inference model and to adopt site and clock models used to simulate the alignment.
-This option can be specified selecting "model\_type" to be "generative" as part of the inference conditions (as seen in ERQ 1, table~\ref{tab:RQ1}).
-
-If, instead, there is no preference for a specific tree prior, the user can let \verb;pirouette; select the best inference model from a list of candidate models.
-The model selection is performed automatically based on the criterion of the highest evidence (aka marginal likelihood).
-This option can be specified providing any number of experiments, selecting "model\_type" to be "candidate" for each of them, and "run\_if" to be "best\_candidate" (table~\ref{tab:RQ2} will summarize this for ERQ 2).
-
-We show the procedure using \verb;pirouette; to answer the second research question:
+In the previous example we selected the inference model $\mathit{I}$ to match 
+with a known generative tree prior $\mathit{P_{0}}$.
+However, a novel tree prior $\mathit{P}$, by definition, is non-standard, thus 
+cannot be part of a standard inference model (i.e. an inference model using a 
+standard tree prior).
+In this case the user has to specify the the tree prior $\mathit{t_{p}}$ to be 
+used in the inference from a list of standard tree priors. In general, having 
+no information on $\mathit{P}$, this choice is arbitrary.
+\verb;pirouette; provides two ways to deal with this problem. The choice can be 
+made specifying the inference conditions $\mathit{C}$ to be used.
+
+If the user has sufficient reasons to believe that one standard tree prior 
+could be more suitable than others for the inference, they can set it as part 
+of the inference model and to adopt site and clock models used to simulate the 
+alignment. This option can be specified selecting "model\_type" to be 
+"generative" as part of the inference conditions (as seen in ERQ 1, 
+table~\ref{tab:RQ1}).
+
+If, instead, there is no preference for a specific tree prior, the user can let 
+\verb;pirouette; select the best inference model from a list of candidate 
+models. The model selection is performed automatically based on the criterion 
+of the highest evidence (aka marginal likelihood). This option can be specified 
+providing any number of experiments, selecting "model\_type" to be "candidate" 
+for each of them, and "run\_if" to be "best\_candidate" (table~\ref{tab:RQ2} 
+will summarize this for ERQ 2).
+
+We show the procedure using \verb;pirouette; to answer the second research 
+question:
 
@@
 
-The first step in \verb;pirouette; is to simulate a DNA alignment from the given phylogeny. 
+The first step in \verb;pirouette; is to simulate a DNA alignment from the 
+given phylogeny. 
 We will re-use the alignment parameters of the previous example 
@@
 
-In the second step we set up our experiments.
-Unlike ERQ 1, here we provide a set of experiments $\{\mathit{X_{i}}\}_{i=1}^{40}$.
-All the experiments have the same inference conditions $\mathit{C_{i}} = \mathit{C}$ but different candidate inference models $\mathit{I_{i}}$.
-Yet, only one of them will be used to run the actual inference.
-According to the specified $\mathit{C}$, \verb;pirouette; will measure the evidence for each of them and select  only the one with the highest evidence as the best candidate.
+In the second step we set up our experiments. Unlike ERQ 1, here we provide a 
+set of experiments $\{\mathit{X_{i}}\}_{i=1}^{40}$. All the experiments have 
+the same inference conditions $\mathit{C_{i}} = \mathit{C}$ but different 
+candidate inference models $\mathit{I_{i}}$. Yet, only one of them will be used 
+to run the actual inference. According to the specified $\mathit{C}$, 
+\verb;pirouette; will measure the evidence for each of them and select  only 
+the one with the highest evidence as the best candidate.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Usage: Example Research Question 3}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 the same as in the previous section, which are
-the phylogeny from an unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
+the phylogeny from an unknown speciation model 
+(listing~\ref{lst:unknown_phylogeny}), 
 the alignment parameters (listing~\ref{lst:create_alignment})
@@
 This time, however, we are interested in creating a twin tree. A twin tree
-has the same topology as the given tree, yet its branching times are simulated according to the standard tree prior used by BEAST2 to produce the posterior for the main pipeline. 
-Since there is coherence between the inference model and the generative model, we expect BEAST2 to produce a posterior of phylogenies more similar to the original one, respect to the outcome of the main pipeline.
-Due to this, the twinning process is used to provide a control for the main experiment.
+has the same topology as the given tree, yet its branching times are simulated 
+according to the standard tree prior used by BEAST2 to produce the posterior 
+for the main pipeline. 
+Since there is coherence between the inference model and the generative model, 
+we expect BEAST2 to produce a posterior of phylogenies more similar to the 
+original one, respect to the outcome of the main pipeline. 
+Due to this, the twinning process is used to provide a control for the main 
+experiment.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Usage: Example Research Question 4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-To answer ERQ 2 we used a evidence-based model selection procedure to obtain an error distribution.
-However, it is unclear to what extent this error is due to the mismatch between the tree priors (from generative and inference models) or due to stochasticity.
-As mentioned in ERQ 1, \verb;pirouette; provides a tool to evaluate this, using the twinning process.
-We now apply it to the same pipeline followed in ERQ 2.
+To answer ERQ 2 we used a evidence-based model selection procedure to obtain an 
+error distribution.
+However, it is unclear to what extent this error is due to the mismatch between 
+the tree priors (from generative and inference models) or due to stochasticity. 
+As mentioned in ERQ 1, \verb;pirouette; provides a tool to evaluate this, using 
+the twinning process. We now apply it to the same pipeline followed in ERQ 2.
 This leads to our fourth example research question:
@@
   \caption{
-    Example 4: the error made by BEAST2 from a phylogeny, when the generative model is known, compared to candidates with similar inference models, in relationship to the background error.
+    Example 4: the error made by BEAST2 from a phylogeny, when the generative 
+model is known, compared to candidates with similar inference models, in 
+relationship to the background error.
   }
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Usage: Example Research Question 5}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-In the previous research question we have seen how to use \verb;pirouette; to  run the analysis on main pipeline, either using a 'generative' or a 'best\_candidate' inference model.
-We also saw that to exclude the baseline error due to stochasticity it is possible to run a parallel twin pipeline.
-In this way two error distributions are obtained.
-Comparing them it is possible to estimate the error made by BEAST2 when inferring a posterior distribution starting from a phylogeny generated under an unknown generative tree prior.
-We can sum up all \verb;pirouette;'s features presented so far in the fifth research question:
+In the previous research question we have seen how to use \verb;pirouette; to  
+run the analysis on main pipeline, either using a 'generative' or a 
+'best\_candidate' inference model. We also saw that to exclude the baseline 
+error due to stochasticity it is possible to run a parallel twin pipeline. In 
+this way two error distributions are obtained. Comparing them it is possible to 
+estimate the error made by BEAST2 when inferring a posterior distribution 
+starting from a phylogeny generated under an unknown generative tree prior.
+We can sum up all \verb;pirouette;'s features presented so far in the fifth 
+research question:
 
@@
 In the second step we state our experiments. 
-In this example, we must both specify the generative and candidate experiments. Here we specify the generative model:
+In this example, we must both specify the generative and candidate experiments. 
+Here we specify the generative model:
 
@@
 
-Specifying all candidates can be done simply by calling \verb;create_all_experiments;. In this case, all candidates will use the same site and clock model, and only differ in their tree priors:
+Specifying all candidates can be done simply by calling 
+\verb;create_all_experiments;. In this case, all candidates will use the same 
+site and clock model, and only differ in their tree priors:
 
@@
 
-We choose again to use sensible defaults for the third (the BEAST2 inference) and fourth (measuring the error) steps, as we previously did to answer the second research question.
-To get an idea of a baseline error made by BEAST2, we also make use of the twinning option:
+We choose again to use sensible defaults for the third (the BEAST2 inference) 
+and fourth (measuring the error) steps, as we previously did to answer the 
+second research question. To get an idea of a baseline error made by BEAST2, we 
+also make use of the twinning option:
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 } 
-the best candidate model gives an error distribution similar to the unknown tree prior,
+the best candidate model gives an error distribution similar to the unknown 
+tree prior,
 hinting that the candidate model would be just as good in inference.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{pirouette resources}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 each exported function shows a minimal example. 
-For easy exploration, each exported function's documentation links to related functions.
-Additionally, \verb;pirouette; has a vignette that demonstrates extensively how to use it.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+For easy exploration, each exported function's documentation links to related 
+functions.
+Additionally, \verb;pirouette; has a vignette that demonstrates extensively how 
+to use it.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Citation of pirouette}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Acknowledgements}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%
 \section{Data Accessibility}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-All code is archived at \url{http://github.com/richelbilderbeek/pirouette_article},
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%
+
+All code is archived at 
+\url{http://github.com/richelbilderbeek/pirouette_article},
 with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Authors' contributions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 RJCB, GL and RSE conceived the idea for the package. 
-RJCB created and tested the package, and wrote the first draft of the manuscript.
+RJCB created and tested the package, and wrote the first draft of the 
+manuscript.
 GL tested the package and contributed substantially to revisions.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Bibliography
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % MEE style
@@
 \bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results in detail}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Errors}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Example 1}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[ht]
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Example 2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[ht]
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Example 3}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[ht]
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Example 4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[ht]
@@
     \node[state] (CB) [below of = B, rectangle, node distance=0.8\textheight] {
+
 \includegraphics[height=0.95\textheight]{example_4_true_posterior_best.png}
@@
     };   
-    \node[state] (CTB) [below of = BT, rectangle, node distance=0.8\textheight] {
+    \node[state] (CTB) [below of = BT, rectangle, node distance=0.8\textheight] 
+{
+
 \includegraphics[height=0.95\textheight]{example_4_twin_posterior_best.png}
     };   
-    \node[state] (DTB) [below of = CTB, rectangle, node distance=1.0\textheight] {
+    \node[state] (DTB) [below of = CTB, rectangle, node 
+distance=1.0\textheight] {
+
 \includegraphics[height=0.65\textheight]{example_4_twin_error_violin_best.png}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Example 5}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[ht]
@@
     \node[state] (DTG) [below of = CTG, rectangle] {
+
 \includegraphics[height=0.2\textheight]{example_5_twin_error_violin_gen.png}
@@
     \node[state] (DTB) [below of = CTB, rectangle] {
+
 \includegraphics[height=0.2\textheight]{example_5_twin_error_violin_best.png}

### 2019-03-18 article.tex
Minor processsing of feedback

@@
 and a strict clock model as these are the simplest.
-A Jukes-Cantor site model,
+A Jukes-Cantor site model
 assumes that mutation rates between nucleotides are equal and constant. 
@@
 
-\giovanni{I think we are providing too many information}
-\richel{agreed, thanks for pushing back}
-
-The error distribution returned by \verb;pir_run; is stored as a data frame. It 
-can be directly plotted using \verb;pir_plot;:
+The error distribution returned by \verb;pir_run; is stored as a data frame. 
+\richel{
+  I suggest \texttt{pir\_run} to return a list of run experiments,
+  as described in Issue 225, 
+  \url{https://github.com/richelbilderbeek/pirouette/issues/225}
+}
+It can be directly plotted using \verb;pir_plot;:
 
@@
 The error distribution obtained is shown as a violin plot
-in figure~\ref{fig:example_1}.
+in figure~\ref{fig:example_1}
+\richel{
+  I suggest to use a histogram (or density plot) instead:
+  I like how the error distribution in figure 1 looks.
+}
+.
 

### 2019-03-18 article.tex
Code and text of ERQ 3 must match

@@
 together with error measurement parameters $\mathit{E}$ and
-twinning parameters $\mathit{T}$. We use the default error measurement
+twinning parameters $\mathit{T}$.
+
+We use the default error measurement
 parameters, which uses a burn-in of 10\% and the nLTT statistic to
-measure the difference between phylogenies. We will not use twinning
+measure the difference between phylogenies. For clarity,
+we create this setup explicitly here:
+
+\begin{lstlisting}[
+  language=R, floatplacement=ht, frame=single,
+  label = {lst:create_error_measure_params},
+  caption = Calling \texttt{create\_error\_measure\_params}
+]
+error_measure_params <- create_error_measure_params(
+  burn_in_fraction = 0.1, 
+  error_function = get_nltt_error_function()
+)
+\end{lstlisting}
+
+We will not use twinning
 here yet.  
@@
   experiments = experiments,
-  error_measure_params = create_error_measure_params(),
+  error_measure_params = error_measure_params,
   twinning_params = NA
@@
 
-Creating a twinning parameter is easy, as it has sensible default settings:
+Creating a twinning parameter is easy, as it has sensible default settings.
+For clarity, we explicitly show the default values of the most
+important arguments:
 
@@
 ]
-twinning_params <- create_twinning_params()
-\end{lstlisting}
+twinning_params <- create_twinning_params(
+  twin_model = "bd", 
+  method = "random_tree"
+)
+\end{lstlisting}
+
+\richel{
+  I would enjoy being able to write ...
+}
+\begin{lstlisting}
+  twinning_params <- create_twinning_params(
+    tree_prior = create_bd_tree_prior(),
+    ...
+  )
+)
+\end{lstlisting}
+\richel{
+  ... as that would connect the tree priors more
+}
 
@@
 in relationship to the background error?"
-\richel{
-  I inserted this example research question, 
-  as think this is natural intermediate between the 
-  previous and next/final example
-}
 
@@
 
-In the second step we specify we want all 
-40 inference models to compete, by 
-calling \verb;create_all_experiments;,
+We agin specify we want all 40 inference models to compete, 
+by calling \verb;create_all_experiments;,
 as done in listing~\ref{lst:create_all_experiments}.
 
-We choose again to use sensible defaults for the third (the BEAST2 inference) 
-and fourth (measuring the error) steps, 
-as we previously did to answer the second research question.
+We again pick to use the default setup to measure the error,
+as shown in listing~\ref{lst:create_error_measure_params}.
 

### 2019-03-18 scripts/example_3.R

@@ -19,7 +19,10 @@ alignment_params <- create_alignment_params(
 
 experiments <- list(create_experiment())
 
-twinning_params <- create_twinning_params()
+twinning_params <- create_twinning_params(
+  twin_model = "bd", 
+  method = "random_tree"
+)
 
 pir_params <- create_pir_params(
   alignment_params = alignment_params,

### 2019-03-18 scripts/example_4.R
Fix https://github.com/richelbilderbeek/pirouette/issues/226

@@ -65,7 +65,13 @@ df_evidences$clock_model_name <- plyr::revalue(
 )
 df_evidences$tree_prior_name <- plyr::revalue(
   df_evidences$tree_prior_name,
-  c("yule" = "Yule", "birth_death" = "BD", "coalescent_constant_population" = "CCP", "coalescent_exp_population" = "CEP")
+  c(
+    "yule" = "Yule",
+    "birth_death" = "BD",
+    "coalescent_bayesian_skyline" = "CBS",
+    "coalescent_constant_population" = "CCP",
+    "coalescent_exp_population" = "CEP"
+  )
 )
 names(df_evidences) <- c("Site model", "Clock model", "Tree prior", "log(evidence)", "Weight")
 
@@ -83,9 +89,16 @@ df_evidences$clock_model_name <- plyr::revalue(
   df_evidences$clock_model_name,
   c("strict" = "Strict", "relaxed_log_normal" = "RLN")
 )
+
 df_evidences$tree_prior_name <- plyr::revalue(
   df_evidences$tree_prior_name,
-  c("yule" = "Yule", "birth_death" = "BD", "coalescent_constant_population" = "CCP", "coalescent_exp_population" = "CEP")
+  c(
+    "yule" = "Yule",
+    "birth_death" = "BD",
+    "coalescent_bayesian_skyline" = "CBS",
+    "coalescent_constant_population" = "CCP",
+    "coalescent_exp_population" = "CEP"
+  )
 )
 names(df_evidences) <- c("Site model", "Clock model", "Tree prior", "log(evidence)", "Weight")
 

### 2019-03-18 article.tex
Review ERQ4, progress https://github.com/richelbilderbeek/pirouette/issues/219

@@
 
-We agin specify we want all 40 inference models to compete, 
+We again specify we want all 40 inference models to compete, 
 by calling \verb;create_all_experiments;,
@@
   experiments = experiments,
+  error_measure_params = error_measure_params,
   twinning_params = twinning_params
@@
 \end{lstlisting}
+\richel{
+  Refer to earlier listing for this call to \texttt{create\_pir\_params}
+}
 
@@
 \end{lstlisting}
+\richel{
+  Refer to earlier listing for this call to \texttt{pir\_run}
+}
 

### 2019-03-18 article.tex
Pretend that create_all_experiments can exlude one inference model as shown in https://github.com/richelbilderbeek/pirouette/issues/227

@@
 
-In the previous research question we have seen how to use \verb;pirouette; to  
-run the analysis on main pipeline, either using a 'generative' or a 
-'best\_candidate' inference model. We also saw that to exclude the baseline 
-error due to stochasticity it is possible to run a parallel twin pipeline. In 
-this way two error distributions are obtained. Comparing them it is possible to 
-estimate the error made by BEAST2 when inferring a posterior distribution 
-starting from a phylogeny generated under an unknown generative tree prior.
+In ERQs 1 and 2 we used \verb;pirouette; to  
+run the analysis on the main pipeline, either using a 'generative' or a 
+'best\_candidate' inference model, resulting in one error distribution,
+that does not mean much on its own.
+In ERQs 3 and 4 we measure the baseline error caused by stochasticity,
+by running a parallel twin pipeline, resulting in two error distributions:
+the baseline error and the complete inference error.
 We can sum up all \verb;pirouette;'s features presented so far in the fifth 
@@
 a Yule tree of low (actually: zero) likelihood: 
-the one as used in examples 2-4.
-by using the code shown in listing~\ref{lst:unknown_phylogeny}.
+the one as used in ERQs 2-4, by using the code shown 
+in listing~\ref{lst:unknown_phylogeny}.
 The first step in \verb;pirouette; is to simulate a DNA alignment 
@@
 generative_experiment <- create_experiment(
+  inference_conditions = create_inference_conditions(
     model_type = "generative",
@@
   )
+  inference_model = create_inference_model(
+    site_model = create_jc69_site_model(),
+    clock_model = create_strict_clock_model(),
+    tree_prior = create_yule_tree_prior()
+  )
+)
 \end{lstlisting}
@@
 Specifying all candidates can be done simply by calling 
-\verb;create_all_experiments;. In this case, all candidates will use the same 
-site and clock model, and only differ in their tree priors:
+\verb;create_all_experiments; and excluding the generative inference model,
+resulting in 39 candidate experiments:
 
@@
 candidate_experiments <- create_all_experiments(
-  site_models = list(create_jc69_site_model()),
-  clock_models = list(create_strict_clock_model()),
-  tree_priors = list(
-    create_bd_tree_prior(), 
-    create_ccp_tree_prior(), 
-    create_cep_tree_prior()
-  )
+  exclude_model = generative_experiment$inference_model
 )

### 2019-03-18 scripts/example_5.R

@@ -25,14 +25,17 @@ generative_experiment <- create_experiment(
     run_if = "always",
     do_measure_evidence = TRUE
   )
+  inference_model = create_inference_model(
+    site_model = create_jc69_site_model(),
+    clock_model = create_strict_clock_model(),
+    tree_prior = create_yule_tree_prior()
+  )
 )
 check_experiment(generative_experiment)
 
 # All non-Yule tree priors
 candidate_experiments <- create_all_experiments(
-  site_models = list(create_jc69_site_model()),
-  clock_models = list(create_strict_clock_model()),
-  tree_priors = list(create_bd_tree_prior(), create_ccp_tree_prior(), create_cep_tree_prior())
+  exclude_model = generative_experiment$inference_model
 )
 check_experiments(candidate_experiments)
 

### 2019-03-18 article.tex
Review ERQ 5, progress https://github.com/richelbilderbeek/pirouette/issues/220

@@
 
-We need to combine these experiments into one set:
+We need to combine these experiments into one set of experiments:
 
@@
 
-We choose again to use sensible defaults for the third (the BEAST2 inference) 
-and fourth (measuring the error) steps, as we previously did to answer the 
-second research question. To get an idea of a baseline error made by BEAST2, we 
-also make use of the twinning option:
+We choose again to use the default error measurement parameters
+as created in listing~\ref{lst:create_error_measure_params}
+and combine this with the same twinning parameters 
+as used in listing~\ref{lst:create_twinning_params}:
 
@@
   experiments = experiments,
-  twinning_params = create_twinning_params()
-)
-\end{lstlisting}
-
-Running the experiment:
+  error_measure_params = error_measure_params,
+  twinning_params = twinning_params
+)
+\end{lstlisting}
+
+Running the experiment in the usual fashion:
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%
 \section{Data Accessibility}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%
 

### 2019-03-18 article.tex
Reviewed discussion and beyond

@@
 % (2) no need to specify BEAST2, it's also just a means
-\title{pirouette: Quantifying the importance of a tree prior in Bayesian 
-phylogenetics}
+\title{
+  pirouette:
+  Quantifying the importance of a tree prior in Bayesian phylogenetics
+}
 % \title{pirouette: Determining the error BEAST2 makes in inferring a phylogeny}
@@
 
-In the four example research questions we showed how to use \verb;pirouette; 
-to assess if BEAST2 is able, with a birth-death prior, 
-to recover the original tree simulated by a Yule process. 
-In principle any other, possibly much more complex, 
-generative prior can be tested.
+In the five example research questions, 
+we showed how to use \verb;pirouette; 
+to quantify the importance of a tree prior in Bayesian phylogenetics.
+In all ERQs we used and assumed the most simple tree prior possible,
+which is the Yule tree prior.
+In principle any other (more complex) generative tree prior 
+can be tested, but we chose to provide the 
+simplest (and fastest to run) examples.
 All tests were performed on a single original tree, 
@@
 
-However if the same procedure were repeated and performed on a distribution 
+However, if the same procedure were repeated and performed on a distribution 
 with a sufficient number of generative trees, 
@@
 Figure~\ref{fig:example_2} shows that 
-\richel{
-  reword after new results with 6 taxa tree
-} 
 the best candidate model gives an error distribution similar to the unknown 
@@
 Figure~\ref{fig:example_3} shows that 
-\richel{
-  reword after new results with 6 taxa tree
-} 
 the twin model gives an error distribution similar to the unknown tree prior,
@@
 Figure~\ref{fig:example_4} shows that 
-\richel{
-  reword after new results with 6 taxa tree
-} 
 the twin model and best candidate model give 
@@
 \verb;pirouette; is free, libre and open source software available at 
-\url{http://github.com/richelbilderbeek/pirouette}
-and is licensed under the GNU General Public License v3.0.
-\verb;pirouette; follows all practices as recommended by the
-literature: continuous integration, a code coverage of 100\%
-and a style guide (\cite{style_guide}).
-\verb;pirouette; depends on multiple packages, which are 
+\url{http://github.com/richelbilderbeek/pirouette},
+licensed under the GNU General Public License version 3.
+\verb;pirouette; depends on multiple packages, which are:
+\richel{I think we can get rid of some dependencies, we just need to check}
 \verb;ape; (\cite{APE}), 
 \verb;babette; (\cite{bilderbeek2018babette}),
+\verb;becosys; (\cite{becosys}),
+\verb;DDD; (\cite{DDD}),
+\verb;devtools; (\cite{devtools}),
+\verb;geiger; (\cite{geiger}),
 \verb;ggplot2; (\cite{ggplot2}),
 \verb;knitr; (\cite{knitr}),
+\verb;lintr; (\cite{lintr}),
 \verb;mcbette; (\cite{mcbette}),
+\verb;nLTT; (\cite{nLTT}),
+\verb;PBD; (\cite{PBD}),
 \verb;phangorn; (\cite{phangorn}),
+\verb;phytools; (\cite{phytools}),
+\verb;rappdirs; (\cite{rappdirs}),
 \verb;rmarkdown; (\cite{rmarkdown}),
+\verb;Rmpfr; (\cite{Rmpfr}),
 \verb;stringr; (\cite{stringr}),
-\verb;testit; (\cite{testit}) and 
-\verb;usethis; (\cite{usethis}).
+\verb;subplex; (\cite{subplex}),
+\verb;TESS; (\cite{TESS}),
+\verb;testit; (\cite{testit}), 
+\verb;testthat; (\cite{testthat}) and
+\verb;tidyr; (\cite{tidyr}).
 
@@
 \url{https://github.com/richelbilderbeek/pirouette},
-which facilitates feature requests and 
-has guidelines on how to do so.
-
-\verb;pirouette;'s documentation is extensive. 
-All functions are documented in the package's internal documentation. 
-For quick use, 
-each exported function shows a minimal example. 
-For easy exploration, each exported function's documentation links to related 
-functions.
-Additionally, \verb;pirouette; has a vignette that demonstrates extensively how 
-to use it.
+which allows submitting bug reports, requesting features, 
+or adding code. To ensure a high quality, \verb;pirouette; 
+uses a continuous integration service, has a code coverage of 100\%
+and enforces the most commonly used R style guide (\cite{style_guide}).
+
+\verb;pirouette;'s is extensively documentated on its website,
+its documentation and its vignettes.
+The \verb;pirouette; website is a good starting point to learn
+how to use \verb;pirouette;, as it links to tutorials and videos.
+The \verb;pirouette; package documentation describes
+all functions and liberally links to related functions.
+All exported functions show a minimal example as part of their documentation.
+The \verb;pirouette; vignette demonstrates extensively how 
+to use \verb;pirouette; in a more lightly written way.
 
@@
 
-Scientists using \verb;pirouette; in a published paper can cite this
-article, and/or cite the \verb;pirouette; package 
-directly. To obtain this citation from within an R script, use:
+Professionals using \verb;pirouette; can cite this
+article. To obtain the citation from within R, use:
 

### 2019-03-18 article.tex
Use X' for a selected eXperiment, formalize its usage

@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{X}$ & Experiment & Full setting for the BEAST2 run & 
-      $\mathit{I},\mathit{C}$ \\
+    $\mathit{X}$ & Experiment (candidate) 
+      & Full setting for a Bayesian inference 
+      & $\mathit{I},\mathit{C}$ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{X'}$ & Experiment (selected) & As above & 
+      $\mathit{I},\mathit{C}$, $Pr(D)$, $\omega$ \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Definitions of terms and relative symbols used in the main text and in 
-    Fig~\ref{fig:pipeline}.
+    Fig~\ref{fig:pipeline}. $Pr(D)$ denotes the evidence, $\omega$ denotes
+    a model's weight 
   }
@@
     The user supplies one or more experiments.
-    For each experiment $\mathit{X}$ 
+    For each candidate experiment $\mathit{X}$ 
     (a tuple of inference model $\mathit{I}$ and condition $\mathit{C}$),
@@
     satisfied (which can depend on the alignment), 
-    its inference model $\mathit{I}$ is used in the next step (3a).
-    The allowed inference models use the alignment (2a) 
-    to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a). 
+    it becomes a selected experiment $\mathit{X'}$, 
+    to be used in the next step (3a).
+    The inference models of the selected experimenst use the alignment (2a) 
+    to each create a Bayesian posterior of (parameter estimates and) 
+    phylogenies (4a). 
     Each of the posteriors' trees is compared to the true phylogeny (1a) 

### 2019-03-19 article.tex
Updates from Overleaf

@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{t_{m}}$ & twin tree prior & Sets the tree prior for the twin tree 
+    $\mathit{\tau_{p}}$ & twin tree prior & Sets the tree prior for the twin tree 
       & Yule, BD \\
@@
     $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & 
-      $\mathit{t_{m}}$\\
+      $\mathit{\tau_{p}}$\\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   }
-  \giovanni{
-    Should we rename "twin model" to "twin prior"?
-  }
-  \richel{
-    I think that's a great idea! I would prefer the name 'twin tree prior' even
-    better
-  }
+  \giovanni{I don't completely get the $X'$ description. What are $Pr(D)$ and $\omega$? Is it something that we really need or it brings more complications than benefits?}
 \end{sidewaystable}
@@
     resulting in a twin error distribution (5b)
+    \giovanni{Ok, that's my take on this: 1) all the capital letters should go in a ball (or a rhombus, as you already did for C) which stands in the middle of the arrows (as with C now); 2) The box in the middle should contain all the inference models $I_{i}$; 3) We need another C at the intersection of the arrows that point to T and A; 4) In between of the arrow 2a->4a there should be another rhombus (or circle) with an I (or I' or I*, according on how we want to call the selected inference model), which is pointed by the rhombus (of circle) containing C; 5) The first arrow should start from a rhombus (or circle) with P inside.}
   }
@@
 
-For all other steps we use the default setups. 
+For all the other steps we use the default setups. 
 We can create the complete
@@
 
-To answer ERQ 2 we used a evidence-based model selection procedure to obtain an 
+To answer ERQ 2 we used an evidence-based model selection procedure to obtain an 
 error distribution.

### 2019-03-25 article.tex
To 80 chars
EOD

--- 2019-03-18 article.tex
@@
 
-% RJCB: I think this title is better: 
-% (1) tree prior importantance is the goal, the error is just the means
-% (2) no need to specify BEAST2, it's also just a means
 \title{
-  pirouette:
-  Quantifying the importance of a tree prior in Bayesian phylogenetics
+  Indirect evidence white is actually a funny color
 }
-% \title{pirouette: Determining the error BEAST2 makes in inferring a phylogeny}
-
 \author[1]{Rich\`el J.C. Bilderbeek}
-\author[1]{Giovanni Laudanno}
-\author[1]{Rampal S. Etienne}
 \affil[1]{Groningen Institute for Evolutionary Life Sciences, University of 
@@
 
-\begin{abstract}
-
-  \textbf{1. }
-    BEAST2 is a popular Bayesian phylogenetics software tool,
-    that takes an alignment of characters (usually nucleotides) 
-    and an inference model to create a
-    posterior of jointly-estimated phylogenies and model parameter 
-    estimates. An important ingredient of BEAST2 is 
-    the species tree prior, 
-    due to which the tool can be used for inference of 
-    macroevolutionary processes underlying the species tree. 
-    However, only models that allow for fast computation of 
-    the prior probability are possible in practice. 
-    This begs the question of how accurate the tree estimation is 
-    when the real macroevolutionary processes are different 
-    from those assumed in BEAST2. \\
-  \textbf{2. }
-    Here we present \verb;pirouette;, 
-    a free, libre and open-source R package that assesses 
-    the inference error BEAST2 makes based on a known/true 
-    phylogeny, for example generated by a new 
-    macroevolutionary diversification model. \\
-  \textbf{3. }
-    We describe \verb;pirouette;'s usage and the biological scientific
-    question it can answer, including full examples. \\
-  \textbf{4. }
-    Last, we discuss the results obtained by the examples. \\
-\end{abstract}
-
-{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, 
-pirouette, R
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The development of new powerful inference tools, 
-such as BEAST [\cite{drummond2007beast}], 
-MrBayes [\cite{huelsenbeck2001mrbayes}]
-or RevBayes [\cite{hohna2016revbayes}], 
-allows to build phylogenetic trees 
-from character data (usually nucleotide sequences) extracted 
-from extant (but also extinct) organisms.
-This has constituted an important step forward 
-in our understanding of how species evolve.
-Such tools have been increasingly exploited to hypothesize 
-and test what the main drivers and modes for diversification are.
-
-Because BEAST performs a Bayesian analysis, 
-it does not only need character data but also tree priors, 
-describing how the diversification occurs, to return posterior phylogenies.
-BEAST uses standard tree priors such as Yule or 
-constant-rate birth-death [\cite{nee1994reconstructed}] models.
-These models are the most commonly picked in biological research,
-as these embody a reasonable level of biological complexity, 
-yet are computationally fast enough.
-The recent development of BEAST2 [\cite{bouckaert2014beast}] provides, 
-unlike its predecessor, the possibility for third-party users 
-to use their own new or preferred diversification model, 
-if an algorithm is provided to calculate the probability 
-of a tree given such a diversification model.
-There have been plenty of these models (and their associated probability 
-algorithms) 
-developed, among others, those in which diversification is 
-time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
-or diversity-dependent [\cite{etienne2011diversity}],
-or that can shift [\cite{etienne2012conceptual}, 
-\cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}].
-Other birth-death models treat speciation as a process that takes 
-time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}], 
-or that happens in bursts of simultaneous 
-events [\cite{laudanno2018mbd}], or that
-depends on a trait that has two [\cite{maddison2007estimating}], 
-or more [\cite{fitzjohn2012diversitree}] states,
-or even allows concealed states [\cite{beaulieu2016detecting}] 
-or a combination of all these [\cite{herrera2018detecting}].
-Not all of these diversification models are available in BEAST2, 
-but the option is available.
-
-When investigating a diversification process,
-starting from a given phylogeny, 
-the parameter values of the diversification model 
-are commonly estimated using maximum likelihood.
-To test the effectiveness of the diversification model,
-it is standard practice to simulate new phylogenies
-from the (maximum likely) parameters and use maximum
-likelihood to again estimate the parameters.
-An effective diversification model has an
-initial parameter estimate that is similar to the control parameter estimates.
-
-Not all diversification models that pass the procedure described above
-are used in Bayesian inference.
-In Bayesian inference, the likelihood of a diversification 
-model, also called tree prior, is computed millions of time.
-Due to this, effective but computationally expensive tree priors 
-are unused in Bayesian inference.
-Also, there may be no need to add such a tree prior to a Bayesian
-inference framework: the more informative the data, 
-the lower the influence of the tree prior.
-Due to this, existing and easy to compute tree priors are most commonly used.
-However, it is unknown if it is valid to assume that tree prior choice is 
-of low importance.
-
-Here we introduce a method to test the importance of a tree prior.
-The method starts with a phylogeny generated by the process of 
-a (usually novel) tree prior.
-Using existing tree priors, a Bayesian posterior with phylogenies is inferred.
-The more simular the posterior phylogenies 
-and the original phylogeny are,
-the less important the novel tree prior has been.
-Only tree priors that are measured to be influential
-will be worth the effort and computational burden 
-to use in a Bayesian framework.
-
-This method is programmed as an R package called \verb;pirouette;.
-\verb;pirouette; is built on \verb;babette; [\cite{bilderbeek2018babette}], 
-which calls BEAST2 [\cite{bouckaert2014beast}]. 
-With \verb;pirouette;, one has a method to quantify the importance 
-of a novel diversification model.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Description}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\verb;pirouette; is written in the R programming language (\cite{R}).
-The goal of \verb;pirouette; is to quantify the importance of a tree
-prior. It does so by measuring the inference error BEAST2
-makes for a given reconstructed phylogeny, 
-simulated under the mechanisms of a (usually novel) diversification model, 
-which we will refer to as the 'generative tree prior', $\mathit{P}$. 
-Such a study will typically construct phylogenies 
-for different combinations of the diversification model's parameters, 
-to assess under which scenarios the error made by BEAST2 cannot be neglected. 
-Although a well-conducted theoretical study uses many replicates, 
-this article only shows examples with one replicate, for the sake of clarity.
-
-\verb;pirouette; is very flexible and allows the user 
-to specify a big variety of custom settings. 
-These settings can be grouped in macro-sections, 
-according to how they operate in the pipeline. 
-We summarize them in tables~\ref{tab:options} and~\ref{tab:definitions}.
-
-Although many possible tests can be performed, 
-we present here what we think are the five most important 
-example research questions (ERQ) that \verb;pirouette; can answer. 
-For these ERQs we will quantify the importance of a tree prior.
-
-\begin{sidewaystable}
-\centering
-  \begin{tabular}{|@{}c|p{2.5cm}|p{9cm}|p{4.5cm}@{}|}
-    \hline
-    \centering
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} & 
-\textbf{Possible values} \\ 
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model 
-      & Yule, BD, CBS, CCP, CEP \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{c_{m}}$ & clock model & Clock for the DNA mutation rates 
-      & strict, relaxed log-normal \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{s_{m}}$ & site model & Nucleotide substitution model 
-      & JC, HKY, TN93, GTR \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{m_{r}}$ & mutation rate & Pace at which mutation occurs 
-      & $m_{r} \in \mathbb{R}_{>0}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{r_{s}}$ & root sequence & DNA sequence at the root of the tree 
-      & any combination of a, c, g, t \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{m_{p}}$ & MRCA prior & Prior knowledge of a most recent common 
-      ancestor & see documentation \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{m_{s}}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup 
-      & see documentation \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{m_{t}}$ & model type & Criterion to select an inference model 
-      & Generative, Candidate \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{r_{i}}$ & run if & Condition under which an inference model is 
-      used & Always, Best candidate\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{e_{f}}$ & error function & Specifies how to measure the error 
-      & nLTT, $|\gamma|$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{t_{m}}$ & twin tree prior & Sets the tree prior for the twin tree 
-      & Yule, BD \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-  \end{tabular}
-  \caption{
-    Most important parameter options. 
-    BD = birth death (\cite{nee1994reconstructed}), 
-    CBS = coalescent Bayesian Skyline (\cite{drummond2005bayesian}), 
-    CCP = coalescent constant-population, 
-    CEP = coalescent exponential-population,
-    JC = Jukes-Cantor 1969 (\cite{jukes1969evolution}), 
-    HKY = Hasegawa, Kishino and  Yano (\cite{hasegawa1985dating}), 
-    TN93 = \cite{tamura1993estimation}, 
-    GTR = Generalised time-reversible model (\cite{tavare1986some}).
-  }
-  \label{tab:options}
-\bigskip
-
-  \begin{tabular}{|@{}c|p{4cm}|p{9cm}|p{3cm}@{}|}
-    \hline
-    \centering
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{Symbol} & \textbf{Macro-argument} & \textbf{Description} 
-      & \textbf{Includes} \\
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{P}$ & Generative tree prior & Diversification model to generate 
-      the 'true' tree & $\mathit{t_{p}}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{A}$ & Alignment model & Model used to simulate the 
-      alignment ($\mathit{c_{m}}$ must be strict) 
-      & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{m_{r}}, \mathit{r_{s}}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{G}$ & Generative model & The full setting to produce BEAST2 input 
-      data & $\mathit{P}$, $\mathit{A}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{I}$ & Inference model & Phylogenetic inference model to run BEAST2 
-      & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{t_{p}}, 
-      \mathit{m_{p}}, \mathit{m_{s}}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ 
-      is used in the inference & $\mathit{m_{t}}, \mathit{r_{i}}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{X}$ & Experiment & Full setting for the BEAST2 run & 
-      $\mathit{I},\mathit{C}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{E}$ & Error measure parameters & Errors measurement setup & 
-      $\mathit{e_{f}}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & 
-      $\mathit{t_{m}}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline 
-  \end{tabular}
-  \caption{
-    Definitions of terms and relative symbols used in the main text and in 
-    Fig~\ref{fig:pipeline}.
-  }
-  \label{tab:definitions}
-  \giovanni{
-    Should we rename "Twinning Parameters" to "Twinning model"?
-  }
-  \richel{
-    As the twinning parameters contain more than only the twin model's tree 
-    prior, I suggest to keep it 'twinning parameters'
-  }
-  \giovanni{
-    Should we rename "twin model" to "twin prior"?
-  }
-  \richel{
-    I think that's a great idea! I would prefer the name 'twin tree prior' even
-    better
-  }
-\end{sidewaystable}
-
-\subsection{pirouette's pipeline}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{workflow.png}
-  \caption{
-    \texttt{pirouette} pipeline. 
-    The pipeline starts from a phylogeny (1a) simulated by generative prior 
-$\mathit{P}$.
-    The phylogeny is converted to an alignment (2a) using alignment model 
-$\mathit{A}$. 
-    The user supplies one or more experiments.
-    For each experiment $\mathit{X}$ 
-    (a tuple of inference model $\mathit{I}$ and condition $\mathit{C}$),
-    if its condition $\mathit{C}$ is 
-    satisfied (which can depend on the alignment), 
-    its inference model $\mathit{I}$ is used in the next step (3a).
-    The allowed inference models use the alignment (2a) 
-    to create a BEAST2 posterior of (parameter estimates and) phylogenies (4a). 
-    Each of the posteriors' trees is compared to the true phylogeny (1a) 
-    using error measure $\mathit{E}$, 
-    resulting in an error distribution (5a). 
-    Optionally, a twin phylogeny (1b) can be generated from the original 
-    phylogeny (1a) using twinning parameters $T$, where $T$ may depend on
-    the experiments \richel{improve}.
-    From a twin phylogeny, the same intermediate steps are taken,
-    resulting in a twin error distribution (5b)
-  }
-  \label{fig:pipeline}
-\end{figure}
-
-The pipeline is summarized by the following steps:
-\begin{enumerate}
-    \item for a given phylogeny an alignment is simulated under a known 
-alignment model;
-    \item for this alignment, according to the specified inference conditions, 
-an inference model is picked (which may be different from the generative 
-alignment model);
-    \item the alignment is then used as BEAST2 input to infer a posterior 
-distribution of phylogenies;
-    \item posterior phylogenies are compared with the given original phylogeny 
-to estimate the error made, according to the error measure specified by the 
-user;
-\end{enumerate}
-The pipeline is visualized in Fig.~\ref{fig:pipeline}. 
-There is also the option to generate a 'twin tree', 
-that goes through the same pipeline. 
-The utility of this twin tree will be explained below.
-
-The first step simulates a DNA alignment from a given 
-phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
-This operation is performed according to the alignment model 
-specified by the user, which consists in a root DNA sequence, 
-a DNA mutation rate and a site model.
-A strict clock model is used as the only clock model
-type, as the currently used alignment simulation algorithm
-only supports a mutation rate that is equal across all the branches.
-This step is relatively fast, 
-but longer DNA alignments will noticeably slow down the inference step.
-
-The second step (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a)
-starts from experiments as set up by the user. 
-We define an experiment as the combination of an inference model 
-and the conditions to actually use it in the inference step.
-The inference model can be selected in two ways, either using the
-generative model or selecting one or more candidates out of a list of candidate 
-inference models.
-An experiment can use the generative model as the inference model, if the 
-generative model is known.
-An experiment uses a candidate model if either the generative model is unknown,
-or there is an interest in seeing how well non-generative models compete with
-the generative model.
-Typically, a generative model and/or a best candidate model are used in the 
-actual inference.
-When using an experiment with a generative model,
-it must use the same site and clock model used in the alignment simulation,
-as well as the tree prior that underlied the creation of the given phylogeny. 
-When using experiments with candidate models,
-\verb;pirouette; selects the best candidate based
-on the evidence (i.e. marginal likelihood) for the model given the alignment.
-The evidence of an inference model is estimated using a nested sampling (NS)
-approach, as described in \cite{maturana2018model}. The nested sampling is
-performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 
-package. 
-Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac,
-therefore the selection of candidate models is not available directly on 
-Windows. 
-However, even on Windows systems there is the option to perform the model 
-selection from browser, using \verb;mcbette;, yet the computation for this 
-model selection is
-restricted to one hour for free users of the host (Travis CI) continuous
-integration service.
-
-The third step infers the actual Bayesian posteriors from the simulated 
-alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a $\rightarrow$ 4a),
-using the inference models from the experiments selected in the previous step. 
-For each selected experiment, one BEAST2 posterior is inferred, using the 
-\verb;babette; [\cite{bilderbeek2018babette}] R package.
-
-The fourth step compares the true tree to each posterior tree, resulting in an 
-error distribution (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
-The nLTT statistic (\cite{janzen2015approximate}) is used by default, but also 
-a user-defined error statistic can be used. 
-As an example, \verb;pirouette; supplies one custom error statistic,
-that uses an absolute difference in the gamma statistic 
-[\cite{pybus2000testing}].
-Additionally, the user can specify the proportion of posterior phylogenies to 
-discard (i.e. the burn-in), throwing away the first $10\%$
-of all phylogenies by default. This burn-in is used to discard
-the part of the MCMC chain that has not yet converged to a
-representative part of the state space.
-
-\subsection{Twin tree}\label{subsec:twinning}
-
-An optional step is to generate a 'twin tree' $\tau_{I}$
-(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
-that will be analyzed in the same way as the true tree, following a parallel 
-pipeline. We want to produce such tree 'converting' the original tree 
-$\tau_{\mathit{P}}$, created according to the generative prior $\mathit{P}$, to 
-be in accordance with the chosen inference model $\mathit{I}$:
-\begin{align}
-    \tau_{\mathit{P}} = (\Vec{t}_{\mathit{P}}, \psi_{\mathit{P}}) 
-\xrightarrow[]{\mathit{T}} \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, 
-\psi_{\mathit{I}}),
-\end{align}
-where $\Vec{t}$ are the tree branching times and $\psi$ is tree topology, which 
-provides information on where new lineages branch out at each branching time.
-
-To execute the twinning process we exploit the likelihood function 
-$L_{\mathit{I}}$ associated with the chosen inference model $\mathit{I}$. We 
-find the parameters $\theta^{*}_{\mathit{I}}$ (e.g. speciation and extinction 
-rates, in case of a birth-death model) that maximize $L_{\mathit{I}}$ applied 
-to the true tree, conditioned on its number of tips $n_{\mathit{P}}$,
-\begin{align}
-    max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{P}}, n_{\mathit{P}})] 
-\rightarrow \theta^{*}_{\mathit{I}}.
-\end{align}
-We use $\theta^{*}_{\mathit{I}}$ to simulate an equal amount $n_{\mathit{I}} = 
-n_{\mathit{P}}$ of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree 
-$\tau_{\mathit{I}}$, under the process $\mathit{I}$. We then impose that 
-$\tau_{\mathit{P}}$ and $\tau_{\mathit{I}}$ must respect the same topology
-\begin{align}
-    \psi_{\mathit{I}} = \psi_{\mathit{P}}.
-\end{align}
-Using a twin tree serves as a control: both $\tau_{\mathit{P}}$ and 
-$\tau_{\mathit{I}}$ will produce an error distribution. We assume that these 
-errors are a natural product of the input tree and the intermediate steps of 
-the pipeline. All the factors that can cause such errors are: the branching 
-times $\vec{t}$ of the tree (as a consequence of the tree prior), its topology 
-$\psi$, the stochasticity induced by simulating the alignment (step 1a 
-$\rightarrow$ 2a in Fig.~\ref{fig:pipeline}) and the stochasticity induced by 
-the Bayesian inference (step 2a $\rightarrow$ 4a in Fig.~\ref{fig:pipeline}). 
-By imposing the same topology and replicating the pipelines a sufficient number 
-of times we can sort out all the possible causes that are not a direct 
-consequence of the choice of the tree prior.
-
-Potential differences in shape of error distributions will therefore be 
-consequences of different tree priors for the two parallel pipelines: the use 
-of prior $\mathit{P}$ for the original pipeline and of prior $\mathit{I}$ for 
-the twin pipeline. In the second case, though, such error distribution is not 
-influenced by the mismatch between how the tree is generated and how it is 
-analyzed by BEAST2.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Installation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\verb;pirouette; can be installed easily from CRAN:
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install.packages("pirouette")
-\end{lstlisting}
-
-For the most up-to-date version, 
-one can download and install the package from \verb;pirouette;'s GitHub 
-repository:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-usethis::install_github("richelbilderbeek/pirouette")
-\end{lstlisting}
-
-To start using \verb;pirouette;, load its functions in the global namespace 
-first:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-library(pirouette)
-\end{lstlisting}
-Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
-This can be done from within R, using:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install_beast2()
-\end{lstlisting}
-For the option to select a best candidate model,
-\verb;pirouette; needs the "NS" BEAST2 package [\cite{maturana2018model}].
-It can be installed from within R, using:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install_beast2_pkg("NS")
-\end{lstlisting}
-
-An overview of \verb;pirouette;'s main functions is shown in 
-table~\ref{tab:functions}. 
-These functions are demonstrated when in the example code below.
-All \verb;pirouette;'s functions are documented,
-have a useful example, and have sensible defaults.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[h]
-\centering
-\begin{tabular}{ | l | l | }
-\hline
-\textbf{Name} & \textbf{Description} \\
-\hline
-\verb;pir_run; & Run \verb;pirouette; \\
-\verb;create_pir_params; & Create the \verb;pirouette; parameters \\
-\hline
-\verb;create_alignment_params; & Create the alignment parameters \\
-\verb;create_twinning_params; & Create the twinning parameters \\
-\verb;create_experiment; & Create one experiment \\
-\verb;create_error_measure_params; & Create the error measurement parameters \\
-\hline
-\end{tabular}
-\caption{pirouette's main functions.}
-\label{tab:functions}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 1}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Question}
-
-\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's 
-inference,
-by measuring the difference between a given tree and 
-phylogenies in a posterior. Due to stochasticity, all posterior trees
-will differ from the given phylogeny, even when the generative tree
-prior $\mathit{P}$ and alignment model $\mathit{A}$ are the same as 
-those in the inference model $\mathit{I}$.
-Measuring this difference allows us to know the baseline error
-of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' 
-all the tree priors that are available to be used within an inference model 
-(see Table~\ref{tab:options}).
-
-Now we can formulate the first example research question that \verb;pirouette; 
-can answer:
-
-"What is the inference error made on phylogenies
-created by a standard diversification model?"
-
-\subsection{Answer}
-
-In this example we use a standard generative tree prior $\mathit{P_{0}}$, 
-namely the Yule (pure-birth) tree prior. 
-We choose to use a small tree with six taxa to keep
-the calculations short and the figure more readable.
-We pick a crown age of ten time units. This value is 
-completely arbitrary, but does tie in with the mutation rate 
-used in simulating an alignment in the next step.
-
-\begin{lstlisting}[
-    language=R, floatplacement=ht, frame=single, 
-    label = {lst:create_yule_tree}, 
-    caption = {
-      Create a Yule tree. 
-      The resulting tree is shown in figure~\ref{fig:yule_tree}.
-    }
-  ]
-phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
-\end{lstlisting}
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{tree_yule.png}
-  \caption{The Yule tree, as created by listing~\ref{lst:create_yule_tree}.}
-  \label{fig:yule_tree}
-  \giovanni{Remove the text in top-left corner.}
-  \richel{Agreed, iff the text matches the code}
-  \giovanni{I'd use a sexier topology, purely for esthetical reasons.}
-  \richel{Agreed, will pick a sexy RNG seed after create\_yule\_tree is there}
-\end{figure}
-
-The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny. 
-To do so, we must specify an alignment model $\mathit{A}$ which consists of: a 
-DNA root sequence $\mathit{r_{s}}$, the DNA mutation rate $\mathit{m_{r}}$, a
-site model $\mathit{s_{m}}$ (in this context: a nucleotide substitution model) 
-and a clock model $\mathit{c_{m}}$. 
-In this example, the DNA root sequence consists in four blocks of 250 
-nucleotides each, 
-where the per-nucleotide mutation rate is 0.1 mutations per unit time.
-We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
-and a strict clock model as these are the simplest.
-A Jukes-Cantor site model
-assumes that mutation rates between nucleotides are equal and constant. 
-A strict clock model assumes that the mutation rates 
-of all species are equal and constant.
-
-\begin{lstlisting}[
-    language=R,
-    floatplacement=ht, frame=single,
-    label = {lst:create_alignment}, 
-    caption = {Create an alignment.}
-  ]
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1,
-  site_model = create_jc69_site_model(),
-  clock_model = create_strict_clock_model()
-)
-\end{lstlisting}
-
-As the site and clock models used here are also the defaults, 
-these function arguments can be safely omitted: we just show these
-explicitly for clarity.
-
-In the second step we state our experiment.
-We define an experiment $\mathit{X}$ as a combination of an inference model 
-$\mathit{I}$
-and conditions $\mathit{C}$.
-In this example we pick $\mathit{I}$ to be the same inference model as the 
-generative one,
-which is the chosen Yule tree prior $\mathit{P_{0}}$, as well as site and clock 
-models defined in $\mathit{A}$, respectively JC and strict.
-We specify in $\mathit{C}$ that the experiment will always be run.
-
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & 
-\textbf{inference model} \\ 
-    \hline
-    generative & always & no & JC, strict, Yule \\
-    \hline
-  \end{tabular}
-  \caption{
-    Inference conditions and inference model for ERQ 1.
-    JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-  }
-  \label{tab:RQ1}
-\end{table}
-
-Listing~\ref{lst:create_generative_experiment_explicit} shows how to
-set up this experiment:
-
-\begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, frame=single,
-  label = {lst:create_generative_experiment_explicit},
-  caption = {
-    Create the experiments to answer research question 1, 
-    using explicit arguments
-  }
-]
-experiment <- create_experiment(
-  inference_conditions = create_inference_conditions(
-    model_type = "generative", 
-    run_if = "always"
-  ), 
-  inference_model = create_inference_model(
-   site_model = create_jc69_site_model(), 
-   clock_model = create_strict_clock_model(), 
-   tree_prior = create_yule_tree_prior()
-  )
-)
-\end{lstlisting}
-
-Experiments must be bundled in a list to work, even if only one is provided, as 
-in this case:
-
-\begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, frame=single,
-  label = {lst:create_generative_experiment},
-  caption = {
-    Create the experiments to answer research question 1.
-  }
-]
-experiments <- list(experiment)
-\end{lstlisting}
-
-The objects we just created need to be bundled
-in a \verb;pirouette; parameters structure,
-together with error measurement parameters $\mathit{E}$ and
-twinning parameters $\mathit{T}$.
-
-We use the default error measurement
-parameters, which uses a burn-in of 10\% and the nLTT statistic to
-measure the difference between phylogenies. For clarity,
-we create this setup explicitly here:
-
-\begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
-  label = {lst:create_error_measure_params},
-  caption = Calling \texttt{create\_error\_measure\_params}
-]
-error_measure_params <- create_error_measure_params(
-  burn_in_fraction = 0.1, 
-  error_function = get_nltt_error_function()
-)
-\end{lstlisting}
-
-We will not use twinning
-here yet.  
-All these arguments are bundled
-and checked by \verb;create_pir_params;:
-
-\begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
-  label = {lst:create_pir_params},
-  caption = Calling \texttt{create\_pir\_params}
-]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  error_measure_params = error_measure_params,
-  twinning_params = NA
-)
-\end{lstlisting}
-
-Now we can use our Yule tree and \verb;pir_params; to measure 
-the inference error made on phylogenies
-created by a standard diversification model:
-
-\begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
-  label = {lst:pir_run},
-  caption = Calling \texttt{pir\_run}
-]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-\end{lstlisting}
-
-The error distribution returned by \verb;pir_run; is stored as a data frame. 
-\richel{
-  I suggest \texttt{pir\_run} to return a list of run experiments,
-  as described in Issue 225, 
-  \url{https://github.com/richelbilderbeek/pirouette/issues/225}
-}
-It can be directly plotted using \verb;pir_plot;:
-
-\begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
-  label = {lst:pir_plot},
-  caption = Calling \texttt{pir\_plot}
-]
-pir_plot(errors)
-\end{lstlisting}
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_1_errors.png}
-  \caption{
-    Example 1: the error BEAST2 makes from a phylogeny 
-    when the generative and inference model are the same.
-  }
-  \label{fig:example_1}
-\end{figure}
-
-The error distribution obtained is shown as a violin plot
-in figure~\ref{fig:example_1}
-\richel{
-  I suggest to use a histogram (or density plot) instead:
-  I like how the error distribution in figure 1 looks.
-}
-.
-
-This error distribution has little relevance on its own, 
-as we constructed the experiment in such a way $\mathit{I}$ 
-matches with the generative model given by $\mathit{A}$ and $\mathit{P_{0}}$.
-However it will become clear when compared to the errors obtained from a 
-phylogeny 
-derived from a non-standard diversification process $\mathit{P}$.
-
-By doing so, this error distribution will serve as a control,
-as it can be seen as the baseline error caused by stochasticity.
-If the two error distributions obtained from $\mathit{P_{0}}$ and $\mathit{P}$ 
-will appear similar, the error induced by the choice of a non-standard 
-speciation process will be of little influence.
-Given the necessity of such comparison,
-we made it an inherent part of \verb;pirouette;. 
-We call it 'twinning' and we use it in ERQ 3.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Question}
-
-In the previous example we selected the inference model $\mathit{I}$ to match 
-with a known generative tree prior $\mathit{P_{0}}$.
-However, a novel tree prior $\mathit{P}$, by definition, is non-standard, thus 
-cannot be part of a standard inference model (i.e. an inference model using a 
-standard tree prior).
-In this case the user has to specify the the tree prior $\mathit{t_{p}}$ to be 
-used in the inference from a list of standard tree priors. In general, having 
-no information on $\mathit{P}$, this choice is arbitrary.
-\verb;pirouette; provides two ways to deal with this problem. The choice can be 
-made specifying the inference conditions $\mathit{C}$ to be used.
-
-If the user has sufficient reasons to believe that one standard tree prior 
-could be more suitable than others for the inference, they can set it as part 
-of the inference model and to adopt site and clock models used to simulate the 
-alignment. This option can be specified selecting "model\_type" to be 
-"generative" as part of the inference conditions (as seen in ERQ 1, 
-table~\ref{tab:RQ1}).
-
-If, instead, there is no preference for a specific tree prior, the user can let 
-\verb;pirouette; select the best inference model from a list of candidate 
-models. The model selection is performed automatically based on the criterion 
-of the highest evidence (aka marginal likelihood). This option can be specified 
-providing any number of experiments, selecting "model\_type" to be "candidate" 
-for each of them, and "run\_if" to be "best\_candidate" (table~\ref{tab:RQ2} 
-will summarize this for ERQ 2).
-
-We show the procedure using \verb;pirouette; to answer the second research 
-question:
-
-"What is the inference error made on a novel phylogeny when
-picking the best inference model?"
-
-\subsection{Answer}
-
-Here we start with a tree generated from an unknown 
-diversification model, that has six taxa and a crown age of ten:
-
-\begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, 
-  frame=single, 
-  label = {lst:unknown_phylogeny},
-  caption = A phylogeny generated by an unknown diversification model.
-]
-phylogeny  <- ape::read.tree(
-  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
-)
-\end{lstlisting}
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{tree_unknown.png}
-  \caption{The tree derived from an unknown diversification process, 
-    as created by listing~\ref{lst:unknown_phylogeny}.
-  }
-\end{figure}
-
-The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny. 
-We will re-use the alignment parameters of the previous example 
-as shown in listing~\ref{lst:create_alignment}.
-
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & 
-\textbf{inference model} \\ 
-    \hline
-    candidate & best candidate & yes & JC, strict, Yule \\
-    candidate & best candidate & yes & JC, strict, BD   \\
-    ...       & ...            & ... & ...              \\
-    candidate & best candidate & yes & GTR, RLN, CCP    \\
-    candidate & best candidate & yes & GTR, RLN, CEP    \\
-    \hline
-  \end{tabular}
-  \caption{
-    Inference conditions and inference model for ERQ 2.
-    JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-    BD: birth-death tree prior.
-    GTR: GTR site model.
-    RLN: relaxed log-normal clock model.
-    CCP: coalescent constant-population tree prior.
-    CEP: coalescent exponential-population tree prior.
-  }
-  \label{tab:RQ2}
-\end{table}
-
-In the second step we set up our experiments. Unlike ERQ 1, here we provide a 
-set of experiments $\{\mathit{X_{i}}\}_{i=1}^{40}$. All the experiments have 
-the same inference conditions $\mathit{C_{i}} = \mathit{C}$ but different 
-candidate inference models $\mathit{I_{i}}$. Yet, only one of them will be used 
-to run the actual inference. According to the specified $\mathit{C}$, 
-\verb;pirouette; will measure the evidence for each of them and select  only 
-the one with the highest evidence as the best candidate.
-
-Setting up experiments like this is thought to be so common,
-that \verb;pirouette; has a simple function to do:
-\verb;create_all_experiments; will create the entire set of 40 experiments,
-containing the inference models of all combinations of 4 site 
-models, 2 clock models and 5 tree priors.
-
-\begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, 
-  frame=single, 
-  label = {lst:create_all_experiments},
-  caption = Create all 40 candidate experiments.
-]
-experiments <- create_all_experiments()
-\end{lstlisting}
-
-For all other steps we use the default setups. 
-We can create the complete
-\verb;pirouette; parameter set like this (which is the
-same as listing~\ref{lst:create_pir_params}, but using the defaults):
-
-\begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
-  label = {lts:create_pir_params_defaults},
-  caption = Create a \texttt{pir\_params} with many defaults.
-]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments
-)
-\end{lstlisting}
-
-We can run the \verb;pirouette; method with listing~\ref{lst:pir_run}
-and plot the results with listing~\ref{lst:pir_plot}.
-The resulting figure is shown in figure~\ref{fig:example_2}.
-
-\giovanni{describe winner}
-\richel{agreed, after the figures also show it :-) }
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_2_errors.png}
-  \caption{
-    Example 2: the error BEAST2 makes from a phylogeny when
-    picking the best inference model.
-  }
-  \label{fig:example_2}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 3}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In ERQ 1 we measured the baseline error of the \verb;pirouette; method, 
-by using a standard tree prior in both the generation of the tree 
-and in our inference.
-The goal of \verb;pirouette; is to quantify the 
-influence of a new (non-standard) tree prior.
-To do so, we need to compare the inference error 
-from the new tree prior to a baseline error,
-by use of a twin tree (see subsection~\ref{subsec:twinning}).
-To do so, we use the twin tree as a control for the main pipeline.
-In this example, we do not know the generative tree prior $\mathit{P}$,
-but we do know the alignment model, $\mathit{A}$.
-In this ERQ, we assume the new tree prior is so closely related to
-a standard tree prior, that it is reasonable to hand-pick a standard tree prior 
-to serve as a stand-in generative tree prior.
-
-\subsection{Question}
-
-"What is the error made by BEAST2 from a phylogeny, 
-when hand-picking a tree prior, compared to the background noise?"
-
-\subsection{Answer}
-
-Most of the settings for this example research question are
-the same as in the previous section, which are
-the phylogeny from an unknown speciation model 
-(listing~\ref{lst:unknown_phylogeny}), 
-the alignment parameters (listing~\ref{lst:create_alignment})
-and the list of experiments (listing~\ref{lst:create_generative_experiment}),
-the latter stating that the generative tree prior is a Yule tree prior.
-
-This time, however, we are interested measuring both the inference error
-and a baseline error, which is achieved by creating a twinning parameter
-set. 
-This time, however, we are interested in creating a twin tree. A twin tree
-has the same topology as the given tree, yet its branching times are simulated 
-according to the standard tree prior used by BEAST2 to produce the posterior 
-for the main pipeline. 
-Since there is coherence between the inference model and the generative model, 
-we expect BEAST2 to produce a posterior of phylogenies more similar to the 
-original one, respect to the outcome of the main pipeline. 
-Due to this, the twinning process is used to provide a control for the main 
-experiment.
-
-Creating a twinning parameter is easy, as it has sensible default settings.
-For clarity, we explicitly show the default values of the most
-important arguments:
-
-\begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, 
-  frame=single,
-  label = {lst:create_twinning_params},
-  caption = Create the default twinning parameters
-]
-twinning_params <- create_twinning_params(
-  twin_model = "bd", 
-  method = "random_tree"
-)
-\end{lstlisting}
-
-\richel{
-  I would enjoy being able to write ...
-}
-\begin{lstlisting}
-  twinning_params <- create_twinning_params(
-    tree_prior = create_bd_tree_prior(),
-    ...
-  )
-)
-\end{lstlisting}
-\richel{
-  ... as that would connect the tree priors more
-}
-
-Bundling all the \verb;pirouette; arguments by \verb;create_pir_params;:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  twinning_params = twinning_params
-)
-\end{lstlisting}
-
-Running and plotting the results:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-pir_plot(errors)
-\end{lstlisting}
-
-The resulting figure is shown in figure~\ref{fig:example_3}
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_3_errors.png}
-  \caption{
-    Example 3: the error made by BEAST2 from a phylogeny, 
-    when hand-picking a generative tree prior versus the control.
-    Here, the 'twin' column shows the baseline inference error,
-    as measured on the twin tree
-  }
-  \label{fig:example_3}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Question}
-
-To answer ERQ 2 we used a evidence-based model selection procedure to obtain an 
-error distribution.
-However, it is unclear to what extent this error is due to the mismatch between 
-the tree priors (from generative and inference models) or due to stochasticity. 
-As mentioned in ERQ 1, \verb;pirouette; provides a tool to evaluate this, using 
-the twinning process. We now apply it to the same pipeline followed in ERQ 2.
-This leads to our fourth example research question:
-
-"What is the error made by BEAST2 from a phylogeny 
-derived from an unknown speciation model,
-when selecting the best inference model, 
-in relationship to the background error?"
-
-\subsection{Answer}
-
-Here, same as example 2 and 3, we start with the phylogeny 
-from an unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
-The first step in \verb;pirouette; is to simulate a DNA alignment 
-from the given phylogeny, for which we will use the same code 
-as shown in listing~\ref{lst:create_alignment}.
-
-We again specify we want all 40 inference models to compete, 
-by calling \verb;create_all_experiments;,
-as done in listing~\ref{lst:create_all_experiments}.
-
-We again pick to use the default setup to measure the error,
-as shown in listing~\ref{lst:create_error_measure_params}.
-
-To get an idea of a baseline error made by BEAST2, 
-we also make use of the twinning option by using the
-default created twinning parameters, as shown in 
-listing~\ref{lst:create_twinning_params}.
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  error_measure_params = error_measure_params,
-  twinning_params = twinning_params
-)
-\end{lstlisting}
-\richel{
-  Refer to earlier listing for this call to \texttt{create\_pir\_params}
-}
-
-Running the experiment:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-\end{lstlisting}
-\richel{
-  Refer to earlier listing for this call to \texttt{pir\_run}
-}
-
-Again, showing the results:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-pir_plot(errors)
-\end{lstlisting}
-\richel{
-  Refer to earlier listing for this call to \texttt{pir\_run}
-}
-
-The resulting figure is shown in figure~\ref{fig:example_4}.
-
-\giovanni{describe winner}
-\richel{agreed, after the figures also show it :-) }
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_4_errors.png}
-  \caption{
-    Example 4: the error made by BEAST2 from a phylogeny, when the generative 
-    model is known, compared to candidates with similar inference models, in 
-    relationship to the background error.
-  }
-  \label{fig:example_4}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 5}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Question}
-
-In ERQs 1 and 2 we used \verb;pirouette; to  
-run the analysis on the main pipeline, either using a 'generative' or a 
-'best\_candidate' inference model, resulting in one error distribution,
-that does not mean much on its own.
-In ERQs 3 and 4 we measure the baseline error caused by stochasticity,
-by running a parallel twin pipeline, resulting in two error distributions:
-the baseline error and the complete inference error.
-We can sum up all \verb;pirouette;'s features presented so far in the fifth 
-research question:
-
-"What is the error made by BEAST2 from a phylogeny, 
-when the generative model is hand-picked, 
-compared to candidates with similar inference models, 
-in relationship to the background error?"
-
-\subsection{Answer}
-
-In this example, we use a tree generated by the Yule process.
-Instead of using the common Yule tree as in example 1, we'll use
-a Yule tree of low (actually: zero) likelihood: 
-the one as used in ERQs 2-4, by using the code shown 
-in listing~\ref{lst:unknown_phylogeny}.
-The first step in \verb;pirouette; is to simulate a DNA alignment 
-from the given phylogeny, for which we will use the same code 
-as shown in listing~\ref{lst:create_alignment}.
-
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & 
-\textbf{inference model} \\ 
-    \hline
-    generative & always         & yes & JC, strict, Yule \\
-    candidate  & best candidate & yes & JC, strict, BD \\
-    candidate  & best candidate & yes & JC, strict, CCP \\
-    candidate  & best candidate & yes & JC, strict, CEP \\
-    \hline
-  \end{tabular}
-  \caption{
-    Experimental setup to answer the fourth research question.
-    JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-    BD: birth-death tree prior.
-    CCP: coalescent constant-population tree prior.
-    CEP: coalescent exponential-population tree prior.
-  }
-  \label{tab:experiment_5}
-\end{table}
-
-In the second step we state our experiments. 
-In this example, we must both specify the generative and candidate experiments. 
-Here we specify the generative model:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-generative_experiment <- create_experiment(
-  inference_conditions = create_inference_conditions(
-    model_type = "generative",
-    run_if = "always",
-    do_measure_evidence = TRUE
-  )
-  inference_model = create_inference_model(
-    site_model = create_jc69_site_model(),
-    clock_model = create_strict_clock_model(),
-    tree_prior = create_yule_tree_prior()
-  )
-)
-\end{lstlisting}
-
-Specifying all candidates can be done simply by calling 
-\verb;create_all_experiments; and excluding the generative inference model,
-resulting in 39 candidate experiments:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-candidate_experiments <- create_all_experiments(
-  exclude_model = generative_experiment$inference_model
-)
-\end{lstlisting}
-
-We need to combine these experiments into one set of experiments:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-experiments <- c(list(generative_experiment), candidate_experiments)
-\end{lstlisting}
-
-We choose again to use the default error measurement parameters
-as created in listing~\ref{lst:create_error_measure_params}
-and combine this with the same twinning parameters 
-as used in listing~\ref{lst:create_twinning_params}:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  error_measure_params = error_measure_params,
-  twinning_params = twinning_params
-)
-\end{lstlisting}
-
-Running the experiment in the usual fashion:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-\end{lstlisting}
-
-Again, showing the results:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-pir_plot(errors)
-\end{lstlisting}
-
-The resulting figure is shown in figure~\ref{fig:example_5}.
-
-\giovanni{describe winner}
-\richel{agreed, after the figures also show it :-) }
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_5_errors.png}
-  \caption{
-    Example 5: the error made by BEAST2 from a phylogeny, 
-    when the generative model is known, 
-    compared to candidates with similar inference models, 
-    in relationship to the background error.
-  }
-  \label{fig:example_5}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In the five example research questions, 
-we showed how to use \verb;pirouette; 
-to quantify the importance of a tree prior in Bayesian phylogenetics.
-In all ERQs we used and assumed the most simple tree prior possible,
-which is the Yule tree prior.
-In principle any other (more complex) generative tree prior 
-can be tested, but we chose to provide the 
-simplest (and fastest to run) examples.
-All tests were performed on a single original tree, 
-hence not providing us enough statistical power to 
-evaluate the level of accuracy of the inference.
-
-However, if the same procedure were repeated and performed on a distribution 
-with a sufficient number of generative trees, 
-it could actually constitute a quantitative assessment 
-of the quality of the inference.
-To give an idea of the insights \verb;pirouette; can give,
-we pretend the results of the example research questions
-to be obtained from thousands of replicates.
-In that case, figure~\ref{fig:example_1} shows 
-the minimum error BEAST2 when the
-generative model is known, which can serve as 
-a minimum noise level assessment in a bigger experiment.  
-Figure~\ref{fig:example_2} shows that 
-the best candidate model gives an error distribution similar to the unknown 
-tree prior,
-hinting that the candidate model would be just as good in inference.
-Figure~\ref{fig:example_3} shows that 
-the twin model gives an error distribution similar to the unknown tree prior,
-hinting that the shape of the phylogeny with unknown tree prior
-does not effect the error made by BEAST2.
-Figure~\ref{fig:example_4} shows that 
-the twin model and best candidate model give 
-an error distribution similar to the unknown tree prior,
-hinting that the shape of the phylogeny with unknown tree prior
-does not effect the error made by BEAST2, 
-and that the candidate model would be just as good in inference.
-
-In conclusion, \verb;pirouette; can directly show 
-the comparison between the results obtained 
-on a new generative prior as compared to the error BEAST2 
-would make when operating on a tree generated 
-under a known, right and existing tree prior (the twin tree).
-From such results an user can estimate whether or not a new tree prior, 
-tailored on the generative process, is needed. 
-If this would indeed be the case, 
-one could implement their own tree prior as a new module in BEAST2.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{pirouette resources}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\verb;pirouette; is free, libre and open source software available at 
-\url{http://github.com/richelbilderbeek/pirouette},
-licensed under the GNU General Public License version 3.
-\verb;pirouette; depends on multiple packages, which are:
-\richel{I think we can get rid of some dependencies, we just need to check}
-\verb;ape; (\cite{APE}), 
-\verb;babette; (\cite{bilderbeek2018babette}),
-\verb;becosys; (\cite{becosys}),
-\verb;DDD; (\cite{DDD}),
-\verb;devtools; (\cite{devtools}),
-\verb;geiger; (\cite{geiger}),
-\verb;ggplot2; (\cite{ggplot2}),
-\verb;knitr; (\cite{knitr}),
-\verb;lintr; (\cite{lintr}),
-\verb;mcbette; (\cite{mcbette}),
-\verb;nLTT; (\cite{nLTT}),
-\verb;PBD; (\cite{PBD}),
-\verb;phangorn; (\cite{phangorn}),
-\verb;phytools; (\cite{phytools}),
-\verb;rappdirs; (\cite{rappdirs}),
-\verb;rmarkdown; (\cite{rmarkdown}),
-\verb;Rmpfr; (\cite{Rmpfr}),
-\verb;stringr; (\cite{stringr}),
-\verb;subplex; (\cite{subplex}),
-\verb;TESS; (\cite{TESS}),
-\verb;testit; (\cite{testit}), 
-\verb;testthat; (\cite{testthat}) and
-\verb;tidyr; (\cite{tidyr}).
-
-\verb;pirouette;'s development takes place on GitHub,
-\url{https://github.com/richelbilderbeek/pirouette},
-which allows submitting bug reports, requesting features, 
-or adding code. To ensure a high quality, \verb;pirouette; 
-uses a continuous integration service, has a code coverage of 100\%
-and enforces the most commonly used R style guide (\cite{style_guide}).
-
-\verb;pirouette;'s is extensively documentated on its website,
-its documentation and its vignettes.
-The \verb;pirouette; website is a good starting point to learn
-how to use \verb;pirouette;, as it links to tutorials and videos.
-The \verb;pirouette; package documentation describes
-all functions and liberally links to related functions.
-All exported functions show a minimal example as part of their documentation.
-The \verb;pirouette; vignette demonstrates extensively how 
-to use \verb;pirouette; in a more lightly written way.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Citation of pirouette}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Professionals using \verb;pirouette; can cite this
-article. To obtain the citation from within R, use:
-
-\begin{lstlisting}[language=R]
-> citation("pirouette")
-\end{lstlisting}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgements}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We would like to thank the Center for Information Technology of the University 
-of Groningen for their support and for providing access to the Peregrine 
-high performance computing cluster. 
-We thank the Netherlands 
-Organization for Scientific Research (NWO) for financial support 
-through a VICI grant awarded to RSE.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Data Accessibility}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-All code is archived at 
-\url{http://github.com/richelbilderbeek/pirouette_article},
-with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Authors' contributions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-RJCB, GL and RSE conceived the idea for the package. 
-RJCB created and tested the package, and wrote the first draft of the 
-manuscript.
-GL tested the package and contributed substantially to revisions.
-RSE contributed to revisions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% MEE style
-\bibliographystyle{mee}
-\bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results in detail}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Errors}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 1}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.25\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1_true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1_true_alignment.png}
-    };   
-    \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1_true_posterior_gen.png}
-    };   
-    \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1_error_violin.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (C)
-      (C) edge [anchor = south] node {} (D)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_1_full_pipeline}
-  \caption{Example 1: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_1_esses.latex}
-% has label tab:esses_example_1
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.25\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2_true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2_true_alignment.png}
-    };   
-    \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2_true_posterior_gen.png}
-    };   
-    \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2_error_violin.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (C)
-      (C) edge [anchor = south] node {} (D)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_2_full_pipeline}
-  \caption{Example 2: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_2_esses.latex}
-% has label tab:esses_example_2
-
-\input{example_2_evidences.latex}
-% has label tab:evidences_example_2
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 3}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.4\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3_true_alignment.png}
-    };   
-    \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_true_posterior_gen.png}
-    };   
-    \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_error_violin.png}
-    };   
-    \node[state] (AT) [right of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_twin_tree.png}
-    };   
-    \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3_twin_alignment.png}
-    };   
-    \node[state] (CT) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_twin_posterior_gen.png}
-    };   
-    \node[state] (DT) [below of = CT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_twin_error_violin.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (C)
-      (C) edge [anchor = south] node {} (D)
-      (A) edge [anchor = east] node {} (AT)
-      (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = south] node {} (CT)
-      (CT) edge [anchor = south] node {} (DT)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_3_full_pipeline}
-  \caption{Example 3: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_3_esses.latex}
-% has label tab:esses_example_3
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {1.0\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.5\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{example_4_true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_4_true_alignment.png}
-    };   
-    \node[state] (CB) [below of = B, rectangle, node distance=0.8\textheight] {
-
-\includegraphics[height=0.95\textheight]{example_4_true_posterior_best.png}
-    };   
-    \node[state] (DB) [below of = CB, rectangle, node distance=1.0\textheight] {
-      \includegraphics[height=0.65\textheight]{example_4_error_violin_best.png}
-    };   
-    \node[state] (AT) [right of = A, rectangle, node distance=2.0\textwidth] {
-      \includegraphics[height=0.4\textheight]{example_4_twin_tree.png}
-    };   
-    \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_4_twin_alignment.png}
-    };   
-    \node[state] (CTB) [below of = BT, rectangle, node distance=0.8\textheight] 
-{
-
-\includegraphics[height=0.95\textheight]{example_4_twin_posterior_best.png}
-    };   
-    \node[state] (DTB) [below of = CTB, rectangle, node 
-distance=1.0\textheight] {
-
-\includegraphics[height=0.65\textheight]{example_4_twin_error_violin_best.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south east] node {} (CB)
-      (CB) edge [anchor = south] node {} (DB)
-      (A) edge [anchor = east] node {} (AT)
-      (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = south] node {} (CTB)
-      (CTB) edge [anchor = south] node {} (DTB)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_4_full_pipeline}
-  \caption{Example 4: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_4_esses_best.latex}
-% has label tab:esses_example_4_best
-
-\input{example_4_esses_twin_best.latex}
-% has label tab:esses_example_4_twin_best
-
-\input{example_4_evidence_true.latex}
-% has label tab:evidences_example_4
-
-\input{example_4_evidence_twin.latex}
-% has label tab:evidences_example_4
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 5}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.4\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5_true_alignment.png}
-    };   
-    \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_true_posterior_gen.png}
-    };   
-    \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_error_violin_gen.png}
-    };   
-    \node[state] (CB) [right of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_true_posterior_best.png}
-    };   
-    \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_error_violin_best.png}
-    };   
-    \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_tree.png}
-    };   
-    \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5_twin_alignment.png}
-    };   
-    \node[state] (CTG) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_gen.png}
-    };   
-    \node[state] (DTG) [below of = CTG, rectangle] {
-
-\includegraphics[height=0.2\textheight]{example_5_twin_error_violin_gen.png}
-    };   
-    \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_best.png}
-    };   
-    \node[state] (DTB) [below of = CTB, rectangle] {
-
-\includegraphics[height=0.2\textheight]{example_5_twin_error_violin_best.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (CG)
-      (CG) edge [anchor = south] node {} (DG)
-      (B) edge [anchor = south east] node {} (CB)
-      (CB) edge [anchor = south] node {} (DB)
-      (A) edge [anchor = east] node {} (AT)
-      (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = south east] node {} (CTG)
-      (CTG) edge [anchor = south] node {} (DTG)
-      (BT) edge [anchor = south] node {} (CTB)
-      (CTB) edge [anchor = south] node {} (DTB)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_5_full_pipeline}
-  \caption{Example 5: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_5_esses_gen.latex}
-% has label tab:esses_example_5_gen
-
-\input{example_5_esses_best.latex}
-% has label tab:esses_example_5_best
-
-\input{example_5_esses_twin_gen.latex}
-% has label tab:esses_example_5_twin_gen
-
-\input{example_5_esses_twin_best.latex}
-% has label tab:esses_example_5_twin_best
-
-\input{example_5_evidence_true.latex}
-% has label tab:evidences_example_5
-
-\input{example_5_evidence_twin.latex}
-% has label tab:evidences_example_5
-
 \end{document}

### 2019-03-25 scripts/example_1.R

@@ -1,197 +1,0 @@
-# Code of example 1
-#
-# Works under Windows
-library(pirouette)
-library(ggplot2)
-library(ggthemes)
-library(ggtree)
-
-root_folder <- path.expand("~/GitHubs/pirouette_article/")
-example_no <- 1
-set.seed(314)
-
-# Should run one day, https://github.com/richelbilderbeek/pirouette/issues/192
-#phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
-phylogeny  <- ape::read.tree(
-  text = "(((((A:2, B:2):2, C:4):2, D:6):2, E:8):2, F:10);"
-)
-
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
-)
-
-experiments <- list(create_experiment())
-
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments
-)
-
-################################################################################
-# Settings to run on Peregrine cluster
-################################################################################
-pir_params$alignment_params$fasta_filename <- file.path(root_folder, paste0("example_", example_no, "_true.fasta"))
-for (i in seq_along(pir_params$experiments)) {
-  pir_params$experiments[[i]]$beast2_options$input_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_input.xml"))
-  pir_params$experiments[[i]]$beast2_options$output_log_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.log"))
-  pir_params$experiments[[i]]$beast2_options$output_trees_filenames <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.trees"))
-  pir_params$experiments[[i]]$beast2_options$output_state_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.xml.state"))
-  pir_params$experiments[[i]]$errors_filename <- file.path(root_folder, paste0("example_", example_no, "_error.csv"))
-}
-pir_params$evidence_filename <- file.path(root_folder, paste0("example_", example_no, "_evidence_true.csv"))
-if (!is_one_na(pir_params$twinning_params)) {
-  twinning_params$twin_tree_filename <- file.path(root_folder, paste0("example_", example_no, "_twin.tree"))
[…]

### 2019-03-25 scripts/example_2.R

@@ -1,225 +1,0 @@
-# Code of example 2
-#
-# Works under Linux and MacOS only
-library(pirouette)
-library(ggplot2)
-library(ggthemes)
-library(ggtree)
-
-root_folder <- path.expand("~/GitHubs/pirouette_article/")
-example_no <- 2
-set.seed(314)
-
-phylogeny  <- ape::read.tree(
-  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
-)
-
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
-)
-
-experiments <- create_all_experiments()
-
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments
-)
-
-################################################################################
-# Settings to run on Peregrine cluster
-################################################################################
-pir_params$alignment_params$fasta_filename <- file.path(root_folder, paste0("example_", example_no, "_true.fasta"))
-for (i in seq_along(pir_params$experiments)) {
-  pir_params$experiments[[i]]$beast2_options$input_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_input.xml"))
-  pir_params$experiments[[i]]$beast2_options$output_log_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.log"))
-  pir_params$experiments[[i]]$beast2_options$output_trees_filenames <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.trees"))
-  pir_params$experiments[[i]]$beast2_options$output_state_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.xml.state"))
-  pir_params$experiments[[i]]$errors_filename <- file.path(root_folder, paste0("example_", example_no, "_error.csv"))
-}
-pir_params$evidence_filename <- file.path(root_folder, paste0("example_", example_no, "_evidence_true.csv"))
-if (!is_one_na(pir_params$twinning_params)) {
-  twinning_params$twin_tree_filename <- file.path(root_folder, paste0("example_", example_no, "_twin.tree"))
-  twinning_params$twin_alignment_filename <- file.path(root_folder, paste0("example_", example_no, "_twin.fasta"))
[…]

### 2019-03-25 scripts/example_3.R

@@ -1,206 +1,0 @@
-# Code of example 3
-#
-# Works under Windows
-library(pirouette)
-library(ggplot2)
-library(ggtree)
-
-root_folder <- path.expand("~/GitHubs/pirouette_article/")
-example_no <- 3
-set.seed(314)
-
-testit::assert(is_beast2_installed())
-
-phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
-
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
-)
-
-experiments <- list(create_experiment())
-
-twinning_params <- create_twinning_params(
-  twin_model = "bd",
-  method = "random_tree"
-)
-
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  twinning_params = twinning_params
-)
-
-print("#######################################################################")
-print("Settings to run on Peregrine cluster")
-print("#######################################################################")
-pir_params$alignment_params$fasta_filename <- file.path(root_folder, paste0("example_", example_no, "_true.fasta"))
-for (i in seq_along(pir_params$experiments)) {
-  pir_params$experiments[[i]]$beast2_options$input_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_input.xml"))
-  pir_params$experiments[[i]]$beast2_options$output_log_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.log"))
-  pir_params$experiments[[i]]$beast2_options$output_trees_filenames <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.trees"))
-  pir_params$experiments[[i]]$beast2_options$output_state_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output.xml.state"))
-  pir_params$experiments[[i]]$errors_filename <- file.path(root_folder, paste0("example_", example_no, "_error.csv"))
-}
-pir_params$evidence_filename <- file.path(root_folder, paste0("example_", example_no, "_evidence_true.csv"))
-if (!is_one_na(pir_params$twinning_params)) {
-  twinning_params$twin_tree_filename <- file.path(root_folder, paste0("example_", example_no, "_twin.tree"))
[…]

### 2019-03-25 scripts/example_4.R

@@ -1,283 +1,0 @@
-# Code of example 4
-#
-# Works under Linux and MacOS only
-library(pirouette)
-library(ggplot2)
-library(ggtree)
-
-root_folder <- path.expand("~/GitHubs/pirouette_article/")
-example_no <- 4
-set.seed(314)
-
-phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
-
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
-)
-
-experiments <- create_all_experiments()
-check_experiments(experiments)
-
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  twinning_params = create_twinning_params()
-)
-
-print("#######################################################################")
-print("Settings to run on Peregrine cluster")
-print("#######################################################################")
-pir_params$alignment_params$fasta_filename <- file.path(root_folder, paste0("example_", example_no, "_true.fasta"))
-for (i in seq_along(pir_params$experiments)) {
-  pir_params$experiments[[i]]$beast2_options$input_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_input_best.xml"))
-  pir_params$experiments[[i]]$beast2_options$output_log_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output_best.log"))
-  pir_params$experiments[[i]]$beast2_options$output_trees_filenames <- file.path(root_folder, paste0("example_", example_no, "_beast2_output_best.trees"))
-  pir_params$experiments[[i]]$beast2_options$output_state_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output_best.xml.state"))
-  pir_params$experiments[[i]]$errors_filename <- file.path(root_folder, paste0("example_", example_no, "_error_best.csv"))
-}
-pir_params$evidence_filename <- file.path(root_folder, paste0("example_", example_no, "_evidence_true.csv"))
-if (!is_one_na(pir_params$twinning_params)) {
-  pir_params$twinning_params$twin_tree_filename <- file.path(root_folder, paste0("example_", example_no, "_twin.tree"))
[…]

### 2019-03-25 scripts/example_5.R

@@ -1,373 +1,0 @@
-# Code of example 5
-#
-# Works under Linux and MacOS only
-library(pirouette)
-library(ggplot2)
-library(ggtree)
-
-root_folder <- path.expand("~/GitHubs/pirouette_article/")
-example_no <- 5
-set.seed(314)
-
-# An infintely rare Yule tree
-phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
-
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
-)
-
-# JC69, strict, Yule
-generative_experiment <- create_experiment(
-  inference_conditions = create_inference_conditions(
-    model_type = "generative",
-    run_if = "always",
-    do_measure_evidence = TRUE
-  ),
-  inference_model = create_inference_model(
-    site_model = create_jc69_site_model(),
-    clock_model = create_strict_clock_model(),
-    tree_prior = create_yule_tree_prior()
-  )
-)
-check_experiment(generative_experiment)
-
-# All non-Yule tree priors
-candidate_experiments <- create_all_experiments(
-  exclude_model = generative_experiment$inference_model
-)
-check_experiments(candidate_experiments)
-
-experiments <- c(list(generative_experiment), candidate_experiments)
-check_experiments(experiments)
-
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  twinning_params = create_twinning_params()
-)
-
-print("#######################################################################")
-print("Settings to run on Peregrine cluster")
-print("#######################################################################")
-pir_params$alignment_params$fasta_filename <- file.path(root_folder, paste0("example_", example_no, "_true.fasta"))
-for (i in seq_along(pir_params$experiments)) {
-  pir_params$experiments[[i]]$beast2_options$input_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_input_best.xml"))
-  pir_params$experiments[[i]]$beast2_options$output_log_filename <- file.path(root_folder, paste0("example_", example_no, "_beast2_output_best.log"))
[…]

### 2019-03-25 scripts/example_all.R
Scripts are back

@@ -1,0 +1,53 @@
+# Code of example 2, yet now all experiments are run
+#
+# Works under Linux and MacOS only
+library(pirouette)
+library(ggplot2)
+library(ggthemes)
+
+phylogeny  <- ape::read.tree(text = "((A:4, B:4):1, (C:4, D:4) :1);")
+
+alignment_params <- create_alignment_params(
+  root_sequence = create_blocked_dna(length = 1000),
+  mutation_rate = 0.1
+)
+
+experiments <- create_all_experiments()
+
+if (1 == 2) {
+  # Run all experiments
+  for (i in seq_along(experiments)) {
+    experiments[[i]]$run_if <- "always"
+    experiments[[1]]$errors_filename <- paste0("errors_", i, ".csv")
+  }
+} else {
+  experiments[[1]]$run_if <- "always"
+  experiments[[2]]$run_if <- "always"
+  experiments[[3]]$run_if <- "always"
+  experiments[[1]]$errors_filename <- paste0("errors_1.csv")
+  experiments[[2]]$errors_filename <- paste0("errors_2.csv")
+  experiments[[3]]$errors_filename <- paste0("errors_3.csv")
+}
+
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments
+)
+
+errors <- pir_run(
+  phylogeny,
+  pir_params = pir_params
+)
+
+pir_plot(errors) + theme(
+  panel.background = element_rect(
+    fill = "white",
+    colour = "black"
+  ),
+  panel.grid.major = element_line(colour = "black"),
+  panel.grid.minor = element_line(color = "black")
+
+)
+
+pir_plot(errors) + theme_wsj() +
+  ggsave("/home/richel/GitHubs/pirouette_article/figure_example_all.png")

### 2019-03-25 article.tex
Even the text is back :-)

--- 2019-03-19 article.tex
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{\tau_{p}}$ & twin tree prior & Sets the tree prior for the twin tree 
+    $\mathit{\tau_{p}}$ & twin tree prior & Sets the tree prior for the twin 
+tree 
       & Yule, BD \\
@@
   }
-  \giovanni{I don't completely get the $X'$ description. What are $Pr(D)$ and $\omega$? Is it something that we really need or it brings more complications than benefits?}
+  \giovanni{I don't completely get the $X'$ description. What are $Pr(D)$ and 
+$\omega$? Is it something that we really need or it brings more complications 
+than benefits?}
 \end{sidewaystable}
@@
     resulting in a twin error distribution (5b)
-    \giovanni{Ok, that's my take on this: 1) all the capital letters should go in a ball (or a rhombus, as you already did for C) which stands in the middle of the arrows (as with C now); 2) The box in the middle should contain all the inference models $I_{i}$; 3) We need another C at the intersection of the arrows that point to T and A; 4) In between of the arrow 2a->4a there should be another rhombus (or circle) with an I (or I' or I*, according on how we want to call the selected inference model), which is pointed by the rhombus (of circle) containing C; 5) The first arrow should start from a rhombus (or circle) with P inside.}
+    \giovanni{Ok, that's my take on this: 1) all the capital letters should go 
+in a ball (or a rhombus, as you already did for C) which stands in the middle 
+of the arrows (as with C now); 2) The box in the middle should contain all the 
+inference models $I_{i}$; 3) We need another C at the intersection of the 
+arrows that point to T and A; 4) In between of the arrow 2a->4a there should be 
+another rhombus (or circle) with an I (or I' or I*, according on how we want to 
+call the selected inference model), which is pointed by the rhombus (of circle) 
+containing C; 5) The first arrow should start from a rhombus (or circle) with P 
+inside.}
   }
@@
 
-To answer ERQ 2 we used an evidence-based model selection procedure to obtain an 
+To answer ERQ 2 we used an evidence-based model selection procedure to obtain 
+an 
 error distribution.
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Interpretation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{
+  I think we can put the interpretation in a table as well.
+  Question is: do you agree with the interpretations I gave?
+  Please correct me.
+}
+
+\begin{table}
+  \begin{tabular}{ | r | c | c | c | }
+    \hline
+    \textbf{Compare with} & \textbf{E2} & \textbf{E3} & \textbf{E4} \\ 
+    \hline
+    \textbf{E1} & Y & Y & N \\
+    \textbf{E2} & . & N & Y \\
+    \textbf{E3} & . & . & Y \\
+    \textbf{E4} & . & . & . \\
+    \hline
+  \end{tabular}
+  \caption{
+    Error distributions that are relevant to compare.
+    E1: true tree, generative model.
+    E2: true tree, best candidate model.
+    E3: true tree, generative model.
+    E4: true tree, best candidate model.
+  }
+  \label{tab:relevant_comparisions}
+\end{table}
+
+\begin{table}
+  \begin{tabular}{ | r | c | p{9cm} | }
+    \hline
+    \textbf{Condition} & \textbf{Expectation} & \textbf{Interpretation} \\ 
+    \hline
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E1 > E2$}       & Unexpected & Novel tree prior is more related to 
+      best candate than hand-picked tree prior \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E1 \approx E2$} & Possible   & Hand-picked tree prior is just as 
+      suitable as the best candidate tree prior \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E1 < E2$}       & Expected   & Hand-picked tree prior is the most 
+      related tree prior \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \hline
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E1 > E3$}       & Unknown    & Novel tree prior important \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E1 \approx E3$} & Unknown    & Novel tree prior unimportant \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E1 < E3$}       & Unexpected & Twinning procedure increases 
+      inference errors when using hand-picked tree prior \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \hline
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E2 > E4$}       & Expected   & Impact of novel tree prior cannot 
+      be compensated for by model selection: twin tree with low likelihood? \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E2 \approx E4$} & Possible   & Best candidate tree priors perform 
+      equally well in true and twin tree: true and twin tree similar? \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E2 < E4$}       & Unexpected & Twinning procedure increases 
+      inference errors  when using best tree prior candidate \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \hline
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E3 > E4$}       & Unexpected & Hand-picked tree prior (that equals 
+      the twinning tree prior!) worse than best candidate tree prior \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E3 \approx E4$} & Possible   & Twin tree fits equally well to the 
+      hand-picked and best candidate tree prior  \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$E3 < E4$}       & Expected   & Hand-pick tree prior (that equals 
+      the twinning tree prior) performs as expected \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \hline
+  \end{tabular}
+  \caption{
+    Interpretation of the different error distributions.
+    This assumes that the operators ($<$, $\approx$ and $>$) to compare
+    distributions are defined.
+    E1: true tree, generative model.
+    E2: true tree, best candidate model.
+    E3: true tree, generative model.
+    E4: true tree, best candidate model.
+  }
+  \label{tab:interpretation}
+\end{table}
 

### 2019-04-02 article.tex
Updates from Overleaf

@@
     to be used in the next step (3a).
-    The inference models of the selected experimenst use the alignment (2a) 
+    The inference models of the selected experiment use the alignment (2a) 
     to each create a Bayesian posterior of (parameter estimates and) 
@@
 models. The model selection is performed automatically based on the criterion 
-of the highest evidence (aka marginal likelihood). This option can be specified 
+of the highest evidence (also known as marginal likelihood). This option can be specified 
 providing any number of experiments, selecting "model\_type" to be "candidate" 
@@
 
-Setting up experiments like this is thought to be so common,
-that \verb;pirouette; has a simple function to do:
+As model selection is commonly performed on the full list of available candidate models,
+\verb;pirouette; has a dedicated function to provide it:
 \verb;create_all_experiments; will create the entire set of 40 experiments,
@@
 models, 2 clock models and 5 tree priors.
+\giovanni{Do we still want to use CBS, CCP and CEP?}
 
@@
 
-We can run the \verb;pirouette; method with listing~\ref{lst:pir_run}
+We can now run \verb;pirouette;  with listing~\ref{lst:pir_run}
 and plot the results with listing~\ref{lst:pir_plot}.
@@
 
-In ERQ 1 we measured the baseline error of the \verb;pirouette; method, 
-by using a standard tree prior in both the generation of the tree 
-and in our inference.
+In ERQ 1 we measured the baseline error of the \verb;pirouette; method. We did so by using a standard tree prior to generate the original tree as well as in the inference.
 The goal of \verb;pirouette; is to quantify the 
@@
 from the new tree prior to a baseline error,
-by use of a twin tree (see subsection~\ref{subsec:twinning}).
+by using a twin tree (see subsection~\ref{subsec:twinning}).
 To do so, we use the twin tree as a control for the main pipeline.
-In this example, we do not know the generative tree prior $\mathit{P}$,
-but we do know the alignment model, $\mathit{A}$.
-In this ERQ, we assume the new tree prior is so closely related to
-a standard tree prior, that it is reasonable to hand-pick a standard tree prior 
-to serve as a stand-in generative tree prior.
+In this example, we have no knowledge of the generative tree prior $\mathit{P}$
+but we know the alignment model $\mathit{A}$.
+In this ERQ it is reasonable to hand-pick a standard tree prior 
+to serve as a stand-in generative tree prior as we assume that they are closely related.
 
@@
 
-Most of the settings for this example research question are
-the same as in the previous section, which are
+Most of the settings for this ERQ are
+the same as in the previous section. They are
 the phylogeny from an unknown speciation model 
@@
 
-This time, however, we are interested measuring both the inference error
-and a baseline error, which is achieved by creating a twinning parameter
-set. 
-This time, however, we are interested in creating a twin tree. A twin tree
+This time, however, we are interested in measuring both the inference error
+and a baseline error. We do so by setting the twinning parameters $\mathit{T}$ to create a twin tree. A twin tree
 has the same topology as the given tree, yet its branching times are simulated 
@@
 Since there is coherence between the inference model and the generative model, 
-we expect BEAST2 to produce a posterior of phylogenies more similar to the 
+we expect BEAST2 to produce a posterior of phylogenies that are more similar to the 
 original one, respect to the outcome of the main pipeline. 
@@
 
-Creating a twinning parameter is easy, as it has sensible default settings.
-For clarity, we explicitly show the default values of the most
+Creating the twinning parameters is easy, as it has sensible default settings.
+For the sake of clarity, we explicitly show the default values of the most
 important arguments:
@@
 
-Bundling all the \verb;pirouette; arguments by \verb;create_pir_params;:
+We bundle again all the \verb;pirouette; arguments using \verb;create_pir_params;:
 
@@
 
-Running and plotting the results:
+Finally, we run and plot the results:
 
@@
 
-The resulting figure is shown in figure~\ref{fig:example_3}
+The output is shown in figure~\ref{fig:example_3}
 
@@
 
-\subsection{Question}
-
 To answer ERQ 2 we used an evidence-based model selection procedure to obtain 
@@
 the twinning process. We now apply it to the same pipeline followed in ERQ 2.
-This leads to our fourth example research question:
+
+\subsection{Question}
 
@@
 when selecting the best inference model, 
-in relationship to the background error?"
+in relation to the background error?"
 
@@
 
-Here, same as example 2 and 3, we start with the phylogeny 
-from an unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
+Here, as in examples 2 and 3, we start with the phylogeny 
+generated by an unknown speciation model (listing~\ref{lst:unknown_phylogeny}). 
 The first step in \verb;pirouette; is to simulate a DNA alignment 
@@
 
-We again specify we want all 40 inference models to compete, 
+We specify that we want all 40 inference models to compete 
 by calling \verb;create_all_experiments;,
-as done in listing~\ref{lst:create_all_experiments}.
-
-We again pick to use the default setup to measure the error,
+as we previously did in listing~\ref{lst:create_all_experiments}.
+
+We choose again to use the default error measurement setup,
 as shown in listing~\ref{lst:create_error_measure_params}.
@@
 we also make use of the twinning option by using the
-default created twinning parameters, as shown in 
+default twinning parameters, as shown in 
 listing~\ref{lst:create_twinning_params}.
@@
 
-Running the experiment:
+We run the experiment:
 
@@
 
-Again, showing the results:
+and show the results:
 
@@
 
-The resulting figure is shown in figure~\ref{fig:example_4}.
+The output is shown in figure~\ref{fig:example_4}.
 
@@
 
-\subsection{Question}
-
 In ERQs 1 and 2 we used \verb;pirouette; to  
@@
 that does not mean much on its own.
-In ERQs 3 and 4 we measure the baseline error caused by stochasticity,
-by running a parallel twin pipeline, resulting in two error distributions:
+In ERQs 3 and 4 we measured the baseline error caused by stochasticity,
+by running a parallel twin pipeline. This yielded two error distributions:
 the baseline error and the complete inference error.
-We can sum up all \verb;pirouette;'s features presented so far in the fifth 
-research question:
+
+\subsection{Question}
 
@@
 compared to candidates with similar inference models, 
-in relationship to the background error?"
+in relation to the background error?"
 
@@
 In this example, we use a tree generated by the Yule process.
-Instead of using the common Yule tree as in example 1, we'll use
+Instead of using the common Yule tree as in example 1, we will use
 a Yule tree of low (actually: zero) likelihood: 
@@
 
-Specifying all candidates can be done simply by calling 
+We can specify all the candidates simply by calling 
 \verb;create_all_experiments; and excluding the generative inference model,
@@
 \end{lstlisting}
+\giovanni{Does this code actually work?}
 
@@
 
-Running the experiment in the usual fashion:
+We run the experiment in the usual fashion:
 
@@
 
-Again, showing the results:
+We plot the results:
 
@@
 
-The resulting figure is shown in figure~\ref{fig:example_5}.
+The output is shown in figure~\ref{fig:example_5}.
 
@@
     E4: true tree, best candidate model.
+    \giovanni{What if instead we use two letters, such as "TG", "TB", "WG", "WB"?}
+    \giovanni{There is a typo: true tree is everywhere and twin is not present.}
   }
@@
     \textbf{$E1 > E2$}       & Unexpected & Novel tree prior is more related to 
-      best candate than hand-picked tree prior \\
+      best candidate than hand-picked tree prior \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-04-03 article.tex
Put all pictures in subfolders

@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{tree_yule.png}
+  \includegraphics[width=\textwidth]{example_1/true_tree.png}
   \caption{The Yule tree, as created by listing~\ref{lst:create_yule_tree}.}
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_1_errors.png}
+  \includegraphics[width=\textwidth]{example_1/errors.png}
   \caption{
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{tree_unknown.png}
+  \includegraphics[width=\textwidth]{example_2/tree_unknown.png}
   \caption{The tree derived from an unknown diversification process, 
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_2_errors.png}
+  \includegraphics[width=\textwidth]{example_2/errors.png}
   \caption{
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_3_errors.png}
+  \includegraphics[width=\textwidth]{example_3/errors.png}
   \caption{
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_4_errors.png}
+  \includegraphics[width=\textwidth]{example_4/errors.png}
   \caption{
@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_5_errors.png}
+  \includegraphics[width=\textwidth]{example_5/errors.png}
   \caption{
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1_true_tree.png}
+      \includegraphics[height=0.2\textheight]{example_1/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1_true_alignment.png}
+      \includegraphics[height=0.2\textheight]{example_1/true_alignment.png}
     };   
     \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1_true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_1/true_posterior_gen.png}
     };   
     \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1_error_violin.png}
+      \includegraphics[height=0.2\textheight]{example_1/true_error_violin_gen.png}
     };   
@@
 
-\input{example_1_esses.latex}
+\input{example_1/esses.latex}
 % has label tab:esses_example_1
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2_true_tree.png}
+      \includegraphics[height=0.2\textheight]{example_2/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2_true_alignment.png}
+      \includegraphics[height=0.2\textheight]{example_2/true_alignment.png}
     };   
     \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2_true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_2/true_posterior_best.png}
     };   
     \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2_error_violin.png}
+      \includegraphics[height=0.2\textheight]{example_2/true_error_violin_best.png}
     };   
@@
 
-\input{example_2_esses.latex}
+\input{example_2/esses.latex}
 % has label tab:esses_example_2
 
-\input{example_2_evidences.latex}
+\input{example_2/evidences.latex}
 % has label tab:evidences_example_2
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_true_tree.png}
+      \includegraphics[height=0.2\textheight]{example_3/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3_true_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_3/true_alignment.png}
     };   
     \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_3/true_posterior_gen.png}
     };   
     \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_error_violin.png}
+      \includegraphics[height=0.2\textheight]{example_3/true_error_violin_gen.png}
     };   
     \node[state] (AT) [right of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_twin_tree.png}
+      \includegraphics[height=0.2\textheight]{example_3/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3_twin_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_3/twin_alignment.png}
     };   
     \node[state] (CT) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_twin_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_3/twin_posterior_gen.png}
     };   
     \node[state] (DT) [below of = CT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3_twin_error_violin.png}
+      \includegraphics[height=0.2\textheight]{example_3/twin_error_violin_gen.png}
     };   
@@
 
-\input{example_3_esses.latex}
+\input{example_3/esses.latex}
 % has label tab:esses_example_3
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{example_4_true_tree.png}
+      \includegraphics[height=0.4\textheight]{example_4/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_4_true_alignment.png}
+      \includegraphics[height=0.45\textheight]{example_4/true_alignment.png}
     };   
@@
 
-\includegraphics[height=0.95\textheight]{example_4_true_posterior_best.png}
+\includegraphics[height=0.95\textheight]{example_4/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle, node distance=1.0\textheight] {
-      \includegraphics[height=0.65\textheight]{example_4_error_violin_best.png}
+      \includegraphics[height=0.65\textheight]{example_4/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=2.0\textwidth] {
-      \includegraphics[height=0.4\textheight]{example_4_twin_tree.png}
+      \includegraphics[height=0.4\textheight]{example_4/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_4_twin_alignment.png}
+      \includegraphics[height=0.45\textheight]{example_4/twin_alignment.png}
     };   
@@
 
-\includegraphics[height=0.95\textheight]{example_4_twin_posterior_best.png}
+\includegraphics[height=0.95\textheight]{example_4/twin_posterior_best.png}
     };   
@@
 
-\includegraphics[height=0.65\textheight]{example_4_twin_error_violin_best.png}
+\includegraphics[height=0.65\textheight]{example_4/twin_error_violin_best.png}
     };   
@@
 
-\input{example_4_esses_best.latex}
+\input{example_4/esses_best.latex}
 % has label tab:esses_example_4_best
 
-\input{example_4_esses_twin_best.latex}
+\input{example_4/esses_twin_best.latex}
 % has label tab:esses_example_4_twin_best
 
-\input{example_4_evidence_true.latex}
+\input{example_4/evidence_true.latex}
 % has label tab:evidences_example_4
 
-\input{example_4_evidence_twin.latex}
+\input{example_4/evidence_twin.latex}
 % has label tab:evidences_example_4
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_true_tree.png}
+      \includegraphics[height=0.2\textheight]{example_5/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5_true_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_5/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_5/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_error_violin_gen.png}
+      \includegraphics[height=0.2\textheight]{example_5/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_true_posterior_best.png}
+      \includegraphics[height=0.2\textheight]{example_5/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_error_violin_best.png}
+      \includegraphics[height=0.2\textheight]{example_5/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_tree.png}
+      \includegraphics[height=0.2\textheight]{example_5/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5_twin_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_5/twin_alignment.png}
     };   
     \node[state] (CTG) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_5/twin_posterior_gen.png}
     };   
@@
 
-\includegraphics[height=0.2\textheight]{example_5_twin_error_violin_gen.png}
+\includegraphics[height=0.2\textheight]{example_5/twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5_twin_posterior_best.png}
+      \includegraphics[height=0.2\textheight]{example_5/twin_posterior_best.png}
     };   
@@
 
-\includegraphics[height=0.2\textheight]{example_5_twin_error_violin_best.png}
+\includegraphics[height=0.2\textheight]{example_5/twin_error_violin_best.png}
     };   
@@
 
-\input{example_5_esses_gen.latex}
+\input{example_5/esses_gen.latex}
 % has label tab:esses_example_5_gen
 
-\input{example_5_esses_best.latex}
+\input{example_5/esses_best.latex}
 % has label tab:esses_example_5_best
 
-\input{example_5_esses_twin_gen.latex}
+\input{example_5/esses_twin_gen.latex}
 % has label tab:esses_example_5_twin_gen
 
-\input{example_5_esses_twin_best.latex}
+\input{example_5/esses_twin_best.latex}
 % has label tab:esses_example_5_twin_best
 
-\input{example_5_evidence_true.latex}
+\input{example_5/evidence_true.latex}
 % has label tab:evidences_example_5
 
-\input{example_5_evidence_twin.latex}
+\input{example_5/evidence_twin.latex}
 % has label tab:evidences_example_5

### 2019-04-03 scripts/example_2.R

@@ -94,7 +94,7 @@ df_evidences$tree_prior_name <- plyr::revalue(
 names(df_evidences) <- c("Site model", "Clock model", "Tree prior", "log(evidence)", "Weight")
 
 
-sink(file.path(example_folder, "example_2_evidences.latex"))
+sink(file.path(example_folder, "evidences.latex"))
 xtable::print.xtable(
   xtable::xtable(
     df_evidences,

### 2019-04-03 scripts/example_5.R
Fix script

@@ -100,7 +100,13 @@ df_evidences$clock_model_name <- plyr::revalue(
 )
 df_evidences$tree_prior_name <- plyr::revalue(
   df_evidences$tree_prior_name,
-  c("yule" = "Yule", "birth_death" = "BD", "coalescent_constant_population" = "CCP", "coalescent_exp_population" = "CEP")
+  c(
+    "yule" = "Yule", 
+    "birth_death" = "BD", 
+    "coalescent_bayesian_skyline" = "CBS", 
+    "coalescent_constant_population" = "CCP", 
+    "coalescent_exp_population" = "CEP"
+  )
 )
 names(df_evidences) <- c("Site model", "Clock model", "Tree prior", "log(evidence)", "Weight")
 
@@ -123,7 +129,13 @@ df_evidences$clock_model_name <- plyr::revalue(
 )
 df_evidences$tree_prior_name <- plyr::revalue(
   df_evidences$tree_prior_name,
-  c("yule" = "Yule", "birth_death" = "BD", "coalescent_constant_population" = "CCP", "coalescent_exp_population" = "CEP")
+  c(
+    "yule" = "Yule", 
+    "birth_death" = "BD", 
+    "coalescent_bayesian_skyline" = "CBS", 
+    "coalescent_constant_population" = "CCP", 
+    "coalescent_exp_population" = "CEP"
+  )
 )
 names(df_evidences) <- c("Site model", "Clock model", "Tree prior", "log(evidence)", "Weight")
 

### 2019-04-03 article.tex
Reviewed abstract

@@
     and an inference model to create a
-    posterior of jointly-estimated phylogenies and model parameter 
-    estimates. An important ingredient of BEAST2 is 
-    the species tree prior, 
+    posterior of jointly-estimated phylogenies and model parameters. 
+    An important ingredient of BEAST2 is the species tree prior, 
     due to which the tool can be used for inference of 
@@
     a free, libre and open-source R package that assesses 
-    the inference error BEAST2 makes based on a known/true 
-    phylogeny, for example generated by a new 
-    macroevolutionary diversification model. \\
+    the inference error BEAST2 makes based on a 
+    phylogeny from any macroevolutionary diversification model. \\
   \textbf{3. }
@@
 
-{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, 
-pirouette, R
+{\bf Keywords:} computational biology, evolution, phylogenetics, tree prior,
+  Bayesian model selection, BEAST2, babette, pirouette, R
 

### 2019-04-03 article.tex
Reviewed and improved Introduction

@@
 
-Because BEAST performs a Bayesian analysis, 
-it does not only need character data but also tree priors, 
-describing how the diversification occurs, to return posterior phylogenies.
-BEAST uses standard tree priors such as Yule or 
-constant-rate birth-death [\cite{nee1994reconstructed}] models.
-These models are the most commonly picked in biological research,
-as these embody a reasonable level of biological complexity, 
-yet are computationally fast enough.
-The recent development of BEAST2 [\cite{bouckaert2014beast}] provides, 
-unlike its predecessor, the possibility for third-party users 
-to use their own new or preferred diversification model, 
-if an algorithm is provided to calculate the probability 
-of a tree given such a diversification model.
+BEAST [\cite{drummond2007beast}] is a Bayesian phylogenetics tool, 
+that needs both data and priors to infer a posterior.
+Of these priors, the tree prior describes how diversification occurs.
+BEAST has built-in tree priors such as the Yule [\cite{yule}] and 
+(constant-rate) birth-death [\cite{nee1994reconstructed}] model.
+These simple tree priors are most commonly used, as these 
+have sufficient biological complexity, while being computationally fast.
+BEAST's successor, BEAST2 [\cite{bouckaert2014beast}],
+has a package manager, that allows third-party users to extend existing 
+functionality.
+By writing a BEAST2 package, novel diversification models can be added.
+The key ingredient of such a package would be the likelihood equation
+of a phylogeny for the novel diversification model.
 There have been plenty of these models (and their associated probability 
@@
 or a combination of all these [\cite{herrera2018detecting}].
-Not all of these diversification models are available in BEAST2, 
-but the option is available.
-
-When investigating a diversification process,
-starting from a given phylogeny, 
-the parameter values of the diversification model 
-are commonly estimated using maximum likelihood.
-To test the effectiveness of the diversification model,
-it is standard practice to simulate new phylogenies
-from the (maximum likely) parameters and use maximum
-likelihood to again estimate the parameters.
-An effective diversification model has an
-initial parameter estimate that is similar to the control parameter estimates.
-
-Not all diversification models that pass the procedure described above
-are used in Bayesian inference.
+Only few of these diversification models are available as a BEAST2 package.
+
+When a novel diversification model is introduced,
+its performance is commonly tested as well.
+Part of a model's performance is its parameter 
+recovery (e.g. [\cite{etienne2014estimating}]),
+in which phylogenies are simulated from known parameters, 
+after which the parameters are estimated from the phylogenies.
+Ideally, the estimated parameter values closely match the known/true values.
+
+Even when a diversification model passes the procedure described above,
+it is not necessarily used in Bayesian inference.
 In Bayesian inference, the likelihood of a diversification 
-model, also called tree prior, is computed millions of time.
-Due to this, effective but computationally expensive tree priors 
-are unused in Bayesian inference.
+model, is computed millions of times.
+Due to this, effective but computationally expensive tree priors are inviable.
 Also, there may be no need to add such a tree prior to a Bayesian
 inference framework: the more informative the data, 
-the lower the influence of the tree prior.
-Due to this, existing and easy to compute tree priors are most commonly used.
+the lower the influence the tree prior has.
+Due to this, computionally faster tree priors are more commonly used.
 However, it is unknown if it is valid to assume that tree prior choice is 
@@
 
-Here we introduce a method to test the importance of a tree prior.
+\richel{
+  I think this is a usefull paragraph, adding some depth to this paper,
+  but it could use some rewording
+}
+There have been multiple attemps to investigate the importance of tree
+prior choice. For example, recently Sarver et al., [\cite{sarver2019choice}] 
+show that the choice of tree prior does not 
+substantially affect phylogenetic inferences of diversification rates.
+Also recently, Duchene et al. [\cite{duchene2018phylodynamic}] released
+a BEAST2 package to assess how well posterior predictive simulations
+recover a given tree when using the standard diversification models.
+These investigations show how current diversification models compare
+to one another, but these do not help assess the importance of a 
+new tree prio.
+
+Here we introduce a method to quantify the importance of a novel tree prior.
 The method starts with a phylogeny generated by the process of 
 a (usually novel) tree prior.
-Using existing tree priors, a Bayesian posterior with phylogenies is inferred.
-The more simular the posterior phylogenies 
-and the original phylogeny are,
+Using the tree priors built-in into BEAST2, 
+a Bayesian posterior is inferred.
+The more similar the phylogenies in the posterior are to the original phylogeny,
 the less important the novel tree prior has been.
-Only tree priors that are measured to be influential
-will be worth the effort and computational burden 
+Only impactful tree priors will be worth the effort and computational burden 
 to use in a Bayesian framework.
@@
 which calls BEAST2 [\cite{bouckaert2014beast}].
-With \verb;pirouette;, one has a method to quantify the importance 
-of a novel diversification model.
 

### 2019-04-04 article.tex
Processed feedback @Giappo

@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{X'}$ & Experiment (selected) & As above & 
-      $\mathit{I},\mathit{C}$, $Pr(D)$, $\omega$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     $\mathit{E}$ & Error measure parameters & Errors measurement setup & 
@@
   \label{tab:definitions}
-  \giovanni{
-    Should we rename "Twinning Parameters" to "Twinning model"?
-  }
-  \richel{
-    As the twinning parameters contain more than only the twin model's tree 
-    prior, I suggest to keep it 'twinning parameters'
-  }
-  \giovanni{I don't completely get the $X'$ description. What are $Pr(D)$ and 
-$\omega$? Is it something that we really need or it brings more complications 
-than benefits?}
 \end{sidewaystable}
@@
   \label{fig:yule_tree}
-  \giovanni{Remove the text in top-left corner.}
-  \richel{Agreed, iff the text matches the code}
-  \giovanni{I'd use a sexier topology, purely for esthetical reasons.}
-  \richel{Agreed, will pick a sexy RNG seed after create\_yule\_tree is there}
 \end{figure}
@@
 \giovanni{Do we still want to use CBS, CCP and CEP?}
+\richel{Yes, as we don't know anything about the process underlying the tree}
 
@@
 
-We can now run \verb;pirouette;  with listing~\ref{lst:pir_run}
-and plot the results with listing~\ref{lst:pir_plot}.
+We run \verb;pirouette (listing~\ref{lst:pir_run}) 
+and plot the results (listing~\ref{lst:pir_plot}).
 The resulting figure is shown in figure~\ref{fig:example_2}.
-
-\giovanni{describe winner}
-\richel{agreed, after the figures also show it :-) }
+In this case, the inference model with a JC69 site model, 
+strict clock model and Yule tree prior has the highest likelihood
+amongst all candidates (table \ref{tab:evidences_example_2} shows
+the evidences).
 
@@
 
-The output is shown in figure~\ref{fig:example_4}.
-
-\giovanni{describe winner}
-\richel{agreed, after the figures also show it :-) }
+The output is shown in figure~\ref{fig:example_4}. The alignment
+generated for the true tree had most evidence for the JC site model,
+strict clock model and Yule tree prior, 
+with a weight of 0.981 (as 
+shown in table \ref{tab:evidences_example_4}). 
+The alignment
+generated for the twin tree had most evidence for the JC site model,
+RLN clock model and CBS tree prior,
+with a weight of 0.707 (as 
+shown in table \ref{tab:evidences_example_4}).
 
@@
 
-\giovanni{describe winner}
-\richel{agreed, after the figures also show it :-) }
+The alignment
+generated for the true tree had most evidence for the JC site model,
+strict clock model and Yule tree prior, 
+with a weight of 0.981 \richel{again} (as 
+shown in table \ref{tab:evidences_example_5}). 
+The alignment
+generated for the twin tree had most evidence for the JC site model,
+RLN clock model and CBS tree prior,
+with a weight of 0.707 \richel{again} (as 
+shown in table \ref{tab:evidences_example_5_twin}).
 
@@
     \hline
-    \textbf{Compare with} & \textbf{E2} & \textbf{E3} & \textbf{E4} \\ 
-    \hline
-    \textbf{E1} & Y & Y & N \\
-    \textbf{E2} & . & N & Y \\
-    \textbf{E3} & . & . & Y \\
-    \textbf{E4} & . & . & . \\
+    \textbf{Compare with} & \textbf{TB} & \textbf{WG} & \textbf{WB} \\ 
+    \hline
+    \textbf{TG} & Y & Y & N \\
+    \textbf{TB} & . & N & Y \\
+    \textbf{WG} & . & . & Y \\
+    \textbf{WB} & . & . & . \\
     \hline
@@
     Error distributions that are relevant to compare.
-    E1: true tree, generative model.
-    E2: true tree, best candidate model.
-    E3: true tree, generative model.
-    E4: true tree, best candidate model.
-    \giovanni{What if instead we use two letters, such as "TG", "TB", "WG", "WB"?}
-    \giovanni{There is a typo: true tree is everywhere and twin is not present.}
+    TG: true tree, generative model.
+    TB: true tree, best candidate model.
+    WG: twin tree, generative model.
+    WB: twin tree, best candidate model.
   }
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E1 > E2$}       & Unexpected & Novel tree prior is more related to 
+    \textbf{$TG > TB$}       & Unexpected & Novel tree prior is more related to 
       best candidate than hand-picked tree prior \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E1 \approx E2$} & Possible   & Hand-picked tree prior is just as 
+    \textbf{$TG \approx TB$} & Possible   & Hand-picked tree prior is just as 
       suitable as the best candidate tree prior \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E1 < E2$}       & Expected   & Hand-picked tree prior is the most 
+    \textbf{$TG < TB$}       & Expected   & Hand-picked tree prior is the most 
       related tree prior \\
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E1 > E3$}       & Unknown    & Novel tree prior important \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E1 \approx E3$} & Unknown    & Novel tree prior unimportant \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E1 < E3$}       & Unexpected & Twinning procedure increases 
+    \textbf{$TG > WG$}       & Unknown    & Novel tree prior important \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$TG \approx WG$} & Unknown    & Novel tree prior unimportant \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{$TG < WG$}       & Unexpected & Twinning procedure increases 
       inference errors when using hand-picked tree prior \\
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E2 > E4$}       & Expected   & Impact of novel tree prior cannot 
+    \textbf{$TB > WB$}       & Expected   & Impact of novel tree prior cannot 
       be compensated for by model selection: twin tree with low likelihood? \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E2 \approx E4$} & Possible   & Best candidate tree priors perform 
+    \textbf{$TB \approx WB$} & Possible   & Best candidate tree priors perform 
       equally well in true and twin tree: true and twin tree similar? \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E2 < E4$}       & Unexpected & Twinning procedure increases 
+    \textbf{$TB < WB$}       & Unexpected & Twinning procedure increases 
       inference errors  when using best tree prior candidate \\
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E3 > E4$}       & Unexpected & Hand-picked tree prior (that equals 
+    \textbf{$WG > WB$}       & Unexpected & Hand-picked tree prior (that equals 
       the twinning tree prior!) worse than best candidate tree prior \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E3 \approx E4$} & Possible   & Twin tree fits equally well to the 
+    \textbf{$WG \approx WB$} & Possible   & Twin tree fits equally well to the 
       hand-picked and best candidate tree prior  \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$E3 < E4$}       & Expected   & Hand-pick tree prior (that equals 
+    \textbf{$WG < WB$}       & Expected   & Hand-pick tree prior (that equals 
       the twinning tree prior) performs as expected \\
@@
     distributions are defined.
-    E1: true tree, generative model.
-    E2: true tree, best candidate model.
-    E3: true tree, generative model.
-    E4: true tree, best candidate model.
+    TG: true tree, generative model.
+    TB: true tree, best candidate model.
+    WG: true tree, generative model.
+    WB: true tree, best candidate model.
   }
@@
 
+\begin{table}
+  \begin{tabular}{ | r | c | }
+    \hline
+    \textbf{ERQ} & \textbf{run-time} \\ 
+    \hline
+      1 & 00:10:12 \\
+      2 & 00:34:10 \\
+      3 & 00:24:53 \\
+      4 & 01:04:50 \\
+      5 & 01:29:40 \\
+    \hline
+  \end{tabular}
+  \caption{
+    Time for an example research question to finish.
+    Time is in in hh:mm::ss.
+  }
+  \label{tab:runtime}
+\end{table}
 \end{document}

### 2019-04-04 scripts/example_4.R

@@ -86,7 +86,9 @@ sink(file.path(example_folder, "evidence_true.latex"))
 xtable::print.xtable(
   xtable::xtable(
     df_evidences,
-    caption = "Evidences of example 5", label = "tab:evidences_example_5", digits = 3
+    caption = paste0("Evidences of example ", example_no, " for the true tree"), 
+    label = paste0("tab:evidences_example_", example_no, "_true"), 
+    digits = 3
   ),
   include.rownames = FALSE
 )
@@ -116,7 +118,9 @@ sink(file.path(example_folder, "evidence_twin.latex"))
 xtable::print.xtable(
   xtable::xtable(
     df_evidences,
-    caption = "Evidences of example 5, twin tree", label = "tab:evidences_example_5_twin", digits = 3
+    caption = paste0("Evidences of example ", example_no, " for the twin tree"), 
+    label = paste0("tab:evidences_example_", example_no, "_twin"), 
+    digits = 3
   ),
   include.rownames = FALSE
 )

### 2019-04-04 article.tex
Fix

@@
 % RJCB: I think this title is better: 
-% (1) tree prior importantance is the goal, the error is just the means
+% (1) tree prior importance is the goal, the error is just the means
 % (2) no need to specify BEAST2, it's also just a means
 \title{
-  pirouette:
   Quantifying the importance of a tree prior in Bayesian phylogenetics
@@
 
-We run \verb;pirouette (listing~\ref{lst:pir_run}) 
+We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
 and plot the results (listing~\ref{lst:pir_plot}).

### 2019-04-04 article.tex
Partially reviewed 'Description'

@@
 The goal of \verb;pirouette; is to quantify the importance of a tree
-prior. It does so by measuring the inference error BEAST2
-makes for a given reconstructed phylogeny, 
-simulated under the mechanisms of a (usually novel) diversification model, 
-which we will refer to as the 'generative tree prior', $\mathit{P}$. 
-Such a study will typically construct phylogenies 
+prior.
+It does so by measuring the inference error
+made for a given reconstructed phylogeny, 
+simulated under the mechanisms of a (usually novel) diversification model.
+We refer to the true prior that generated the given tree
+as the 'generative tree prior', $\mathit{P}$.
+Many tree priors have a parameter setting in which these are equivalent
+to a standard tree prior. For example, a protracted birth-death model
+becomes equivalent to a birth-death when speciation rate approaches infinity.
+When benchmarking a novel tree prior, 
+such a study will typically construct phylogenies 
 for different combinations of the diversification model's parameters, 
-to assess under which scenarios the error made by BEAST2 cannot be neglected. 
+to assess under which scenarios the inference error cannot be neglected. 
 Although a well-conducted theoretical study uses many replicates, 
@@
 \verb;pirouette; is very flexible and allows the user 
-to specify a big variety of custom settings. 
+to specify a wide variety of custom settings. 
 These settings can be grouped in macro-sections, 
@@
 example research questions (ERQ) that \verb;pirouette; can answer. 
-For these ERQs we will quantify the importance of a tree prior.
+These ERQs gradually build up in complexity.
+For each of these ERQs we will quantify the importance of a tree prior.
 
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} & 
-\textbf{Possible values} \\ 
+    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} 
+      & \textbf{Possible values} \\ 
     \hline
@@
     $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model 
-      & Yule, BD, CBS, CCP, CEP \\
+      & Yule [\cite{yule}], BD, CBS, CCP, CEP \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{\tau_{p}}$ & twin tree prior & Sets the tree prior for the twin 
-tree 
+    $\mathit{\tau_{p}}$ & twin tree prior 
+      & Sets the tree prior for the twin tree 
       & Yule, BD \\
@@
     BD = birth death (\cite{nee1994reconstructed}), 
-    CBS = coalescent Bayesian Skyline (\cite{drummond2005bayesian}), 
+    CBS = coalescent Bayesian skyline (\cite{drummond2005bayesian}), 
     CCP = coalescent constant-population, 
     CEP = coalescent exponential-population,
-    JC = Jukes-Cantor 1969 (\cite{jukes1969evolution}), 
+    JC = Jukes and Cantor (\cite{jukes1969evolution}), 
     HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
-    TN93 = \cite{tamura1993estimation}, 
+    TN93 = Tamura and Nei \cite{tamura1993estimation}, 
     GTR = Generalised time-reversible model (\cite{tavare1986some}).
@@
 \begin{enumerate}
-    \item for a given phylogeny an alignment is simulated under a known 
-alignment model;
-    \item for this alignment, according to the specified inference conditions, 
-an inference model is picked (which may be different from the generative 
-alignment model);
-    \item the alignment is then used as BEAST2 input to infer a posterior 
-distribution of phylogenies;
-    \item posterior phylogenies are compared with the given original phylogeny 
-to estimate the error made, according to the error measure specified by the 
-user;
+  \item from a given phylogeny an alignment is simulated 
+    under a known alignment model;
+  \item from this alignment, according to the specified inference conditions, 
+    an inference model is picked (which may differ from the 
+    generative alignment model);
+  \item the alignment and inference model are used 
+    to infer a posterior distribution;
+  \item the phylogenies in the posterior are compared with the given phylogeny 
+    to estimate the error made, according to the error measure specified 
+    by the user;
 \end{enumerate}
@@
 The first step simulates a DNA alignment from a given 
-phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
-This operation is performed according to the alignment model 
-specified by the user, which consists in a root DNA sequence, 
-a DNA mutation rate and a site model.
-A strict clock model is used as the only clock model
-type, as the currently used alignment simulation algorithm
-only supports a mutation rate that is equal across all the branches.
-This step is relatively fast, 
-but longer DNA alignments will noticeably slow down the inference step.
+phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a)
+using the DNA alignment parameters.
+The DNA alignment parameters consist of a DNA root sequence, a DNA mutation rate,
+a site model and a (strict) clock model.
+The DNA root sequence is the DNA sequence the shared common ancestor has,
+and is set to four equally-sized mononucleotide blocks by default, as this
+helps interpreting the resulting alignment.
+In the context of this research, a 'site model' is a nucleotide substitution
+model. Supported site models are JC, HKY, TN and GTR. Only the strict
+clock model is supported.
 
 The second step (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a)
-starts from experiments as set up by the user. 
-We define an experiment as the combination of an inference model 
-and the conditions to actually use it in the inference step.
-The inference model can be selected in two ways, either using the
-generative model or selecting one or more candidates out of a list of candidate 
-inference models.
-An experiment can use the generative model as the inference model, if the 
-generative model is known.
-An experiment uses a candidate model if either the generative model is unknown,
-or there is an interest in seeing how well non-generative models compete with
-the generative model.
-Typically, a generative model and/or a best candidate model are used in the 
-actual inference.
-When using an experiment with a generative model,
-it must use the same site and clock model used in the alignment simulation,
-as well as the tree prior that underlied the creation of the given phylogeny. 
-When using experiments with candidate models,
-\verb;pirouette; selects the best candidate based
-on the evidence (i.e. marginal likelihood) for the model given the alignment.
-The evidence of an inference model is estimated using a nested sampling (NS)
-approach, as described in \cite{maturana2018model}. The nested sampling is
-performed by \verb;mcbette; [\cite{mcbette}], that calls the 'NS' BEAST2 
-package. 
-Using BEAST2 packages (in a scripted way) can only be done under Linux and Mac,
-therefore the selection of candidate models is not available directly on 
-Windows. 
-However, even on Windows systems there is the option to perform the model 
-selection from browser, using \verb;mcbette;, yet the computation for this 
-model selection is
-restricted to one hour for free users of the host (Travis CI) continuous
-integration service.
+selects inference models from a set of experiments. 
+We define an experiment $X$ as the combination of an inference model $I$
+and the conditions, $C$ to actually use it in the inference step.
+There are two main setups of condition $C$: either the attached
+inference model $I$ is the generative model and should always be selected,
+or the attached inference model is a candidate model, that will be
+be selected if it is the best model.
+When setting $C$ to be attached to the generative model,
+it is implied that attached inference model $I$ uses the 
+same site and clock model used in the alignment simulation,
+and $I$ uses the same tree prior as used in the creation of the given phylogeny. 
+A 'best' model is defined as the inference model with
+the highest evidence (a.k.a. marginal likelihood) for that inference model, 
+given the alignment simulated in the previous step.
+The evidence of an inference model is estimated by nested 
+sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package.
+Scripted use of BEAST2 packages is only possible under under Linux and Mac.
+Windows systems can do the model comparison for shorter DNA sequences
+using the web interface of \verb;mcbette; [\cite{mcbette}].
+
+\richel{Up until here}
 

### 2019-04-09 article.tex

@@
 \usepackage{mathtools}
+% \usepackage{float}
 
@@
     An important ingredient of BEAST2 is the species tree prior, 
-    due to which the tool can be used for inference of 
+    due to which the tool can be used for the inference of 
     macroevolutionary processes underlying the species tree. 
@@
 has a package manager, that allows third-party users to extend existing 
-functionality.
+functionalities.
 By writing a BEAST2 package, novel diversification models can be added.
@@
 of a phylogeny for the novel diversification model.
+
+\iffalse
 There have been plenty of these models (and their associated probability 
@@
 \cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}].
+\fi
+
+Many of such models (and their associated probability 
+algorithms) have been
+developed to include diverse features: those in which diversification rates are 
+time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
+or diversity-dependent [\cite{etienne2011diversity}],
+or that can shift [\cite{etienne2012conceptual}, 
+\cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}].
+
 Other birth-death models treat speciation as a process that takes 
@@
 
-\subsection{Question}
-
 \verb;pirouette; quantifies the influence of a new tree prior on BEAST2's 
@@
 Now we can formulate the first example research question that \verb;pirouette; 
-can answer:
+can answer.
+
+\subsection{Question}
 
@@
 We pick a crown age of ten time units. This value is 
-completely arbitrary, but does tie in with the mutation rate 
+completely arbitrary, but it ties in with the mutation rate 
 used in simulating an alignment in the next step.
@@
   label = {lst:create_error_measure_params},
-  caption = Calling \texttt{create\_error\_measure\_params}
+  caption = Calling \texttt{create\_error\_measure\_params}.
 ]
@@
   label = {lst:create_pir_params},
-  caption = Calling \texttt{create\_pir\_params}
+  caption = Calling \texttt{create\_pir\_params}.
 ]
@@
   label = {lst:pir_run},
-  caption = Calling \texttt{pir\_run}
+  caption = Calling \texttt{pir\_run}.
 ]
@@
   label = {lst:pir_plot},
-  caption = Calling \texttt{pir\_plot}
+  caption = Calling \texttt{pir\_plot}.
 ]
@@
 }
+\giovanni{Can we show everything at the same time with standard histograms?}
 .
@@
 we made it an inherent part of \verb;pirouette;. 
-We call it 'twinning' and we use it in ERQ 3.
+We call it 'twinning' and we use it from ERQ 3.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Question}
 
@@
 could be more suitable than others for the inference, they can set it as part 
-of the inference model and to adopt site and clock models used to simulate the 
+of the inference model along with site and clock models used to simulate the 
 alignment. This option can be specified selecting "model\_type" to be 
@@
 We show the procedure using \verb;pirouette; to answer the second research 
-question:
+question.
+
+\subsection{Question}
 
@@
 
-Here we start with a tree generated from an unknown 
+This time we start with a tree generated from an unknown 
 diversification model, that has six taxa and a crown age of ten:
@@
   \label{fig:example_2}
+  \giovanni{There is a weird NA at the top of the figure.}
 \end{figure}
@@
 for the main pipeline. 
-Since there is coherence between the inference model and the generative model, 
+Since in the twin pipeline there is coherence between the inference model and the generative model, 
 we expect BEAST2 to produce a posterior of phylogenies that are more similar to the 
@@
     Here, the 'twin' column shows the baseline inference error,
-    as measured on the twin tree
+    as measured on the twin tree.
   }
@@
 the tree priors (from generative and inference models) or due to stochasticity. 
-As mentioned in ERQ 1, \verb;pirouette; provides a tool to evaluate this, using 
+As mentioned in ERQs 1 and 3, \verb;pirouette; provides a tool to evaluate this, using 
 the twinning process. We now apply it to the same pipeline followed in ERQ 2.
@@
 
-Here, as in examples 2 and 3, we start with the phylogeny 
+Here, as in ERQs 2 and 3, we start with a phylogeny 
 generated by an unknown speciation model (listing~\ref{lst:unknown_phylogeny}). 
@@
 
-The output is shown in figure~\ref{fig:example_4}. The alignment
+The output is shown in figure~\ref{fig:example_4}\giovanni{figure~\ref{fig:example_4} needs to be closer}. The alignment
 generated for the true tree had most evidence for the JC site model,
@@
   \label{fig:example_4}
+  \giovanni{NA on the top.}
 \end{figure}
@@
 by running a parallel twin pipeline. This yielded two error distributions:
-the baseline error and the complete inference error.
+the baseline error and the complete inference error. Here we want to combine all the instances of the previous ERQs at once.
 
@@
 
-In this example, we use a tree generated by the Yule process.
+In this example we use a tree generated by the Yule process.
 Instead of using the common Yule tree as in example 1, we will use
-a Yule tree of low (actually: zero) likelihood: 
-the one as used in ERQs 2-4, by using the code shown 
-in listing~\ref{lst:unknown_phylogeny}.
+a Yule tree of low (actually: zero) likelihood \giovanni{what do you mean by zero likelihood?}, i.e. the one already used in ERQs 2-4, by using the code shown in listing~\ref{lst:unknown_phylogeny}.
 The first step in \verb;pirouette; is to simulate a DNA alignment 
@@
 In the second step we state our experiments. 
-In this example, we must both specify the generative and candidate experiments. 
+In this example we must specify both the generative and candidate experiments. 
 Here we specify the generative model:
@@
 We choose again to use the default error measurement parameters
-as created in listing~\ref{lst:create_error_measure_params}
+created in listing~\ref{lst:create_error_measure_params}
 and combine this with the same twinning parameters 
-as used in listing~\ref{lst:create_twinning_params}:
+used in listing~\ref{lst:create_twinning_params}:
 
@@
 The output is shown in figure~\ref{fig:example_5}.
-
+\giovanni{figure~\ref{fig:example_5} needs to be closer}
 The alignment
-generated for the true tree had most evidence for the JC site model,
+generated from the true tree had most evidence for the JC site model,
 strict clock model and Yule tree prior, 
@@
   }
+  \giovanni{na on the top-right area. labels on x superimpose each other. We should change the names to their shortened versions: e.g. relaxed\_log\_normal $\to$ rln}
   \label{fig:example_5}

### 2019-04-09 article.tex
Improve description part

@@
 
-\richel{Up until here}
-
-The third step infers the actual Bayesian posteriors from the simulated 
-alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a $\rightarrow$ 4a),
-using the inference models from the experiments selected in the previous step. 
-For each selected experiment, one BEAST2 posterior is inferred, using the 
+The third step infers the posteriors,
+using the simulated alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a $\rightarrow$ 4a),
+and those inference models that were selected in the previous step. 
+Per selected experiment a posterior is inferred, using the 
 \verb;babette; [\cite{bilderbeek2018babette}] R package. 
-
-The fourth step compares the true tree to each posterior tree, resulting in an 
-error distribution (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
-The nLTT statistic (\cite{janzen2015approximate}) is used by default, but also 
-a user-defined error statistic can be used. 
-As an example, \verb;pirouette; supplies one custom error statistic,
-that uses an absolute difference in the gamma statistic 
-[\cite{pybus2000testing}].
-Additionally, the user can specify the proportion of posterior phylogenies to 
-discard (i.e. the burn-in), throwing away the first $10\%$
-of all phylogenies by default. This burn-in is used to discard
-the part of the MCMC chain that has not yet converged to a
-representative part of the state space.
+This step usually takes up most of the full pipeline's computation time.
+
+The fourth step quantifies the inference error made.
+It creates an error distribution, by measuring the difference
+between the one true tree and each of the posterior 
+trees (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
+The default way to quantify the difference between two phylogenies
+is the nLTT statistic (\cite{janzen2015approximate}), but any 
+user-defined error statistic can be used.
+An MCMC needs to 'warm up', as initially it will sample
+an unrepresentative part of state space. This phase, called 
+the burn-in, is commonly set up to be a small part of the entire run, 
+after which it is discarded. By default, \verb;pirouette; 
+uses a cut-off value of 10\%.
 
@@
 (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
-that will be analyzed in the same way as the true tree, following a parallel 
-pipeline. We want to produce such tree 'converting' the original tree 
-$\tau_{\mathit{P}}$, created according to the generative prior $\mathit{P}$, to 
-be in accordance with the chosen inference model $\mathit{I}$:
+that will be analyzed in the same way as the true tree.
+Any phylogeny $\tau$ is the combination of 
+branching times $\Vec{t}$ and topology $\psi$.
+Starting from a non-standard generative prior $t_p = \mathit{P}$,
+we can observe a given tree $\tau_{\mathit{P}}$,
+with branching times $\Vec{t}_{\mathit{P}}$ and 
+topology $\psi_{\mathit{P}}$, 
+but we cannot use the tree prior in Bayesian inference.
+The twinning process, T, in accordance with a chosen inference 
+model $t_p = \mathit{I}$, creates a tree $\tau_{\mathit{I}}$
+with branching times $\Vec{t}_{\mathit{I}}$ while preserving 
+topology $\psi_{\mathit{P}}$:
+
 \begin{align}
   \tau_{\mathit{P}} = (\Vec{t}_{\mathit{P}}, \psi_{\mathit{P}}) 
-\xrightarrow[]{\mathit{T}} \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, 
-\psi_{\mathit{I}}),
+  \xrightarrow[]{\mathit{T}} 
+  \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, \psi_{\mathit{P}})
 \end{align}
-where $\Vec{t}$ are the tree branching times and $\psi$ is tree topology, which 
-provides information on where new lineages branch out at each branching time.
-
-To execute the twinning process we exploit the likelihood function 
-$L_{\mathit{I}}$ associated with the chosen inference model $\mathit{I}$. We 
-find the parameters $\theta^{*}_{\mathit{I}}$ (e.g. speciation and extinction 
-rates, in case of a birth-death model) that maximize $L_{\mathit{I}}$ applied 
+
+To achieve this, we find the parameters $\theta^{*}_{\mathit{I}}$ 
+(e.g. speciation and extinction rates, in case of a birth-death model) 
+that maximize likelihood $L_{\mathit{I}}$ applied 
 to the true tree, conditioned on its number of tips $n_{\mathit{P}}$,
+where $L_{\mathit{I}}$ is  the likelihood function  
+associated with the chosen inference model $\mathit{I}$:
+
 \begin{align}
@@
 \end{align}
-We use $\theta^{*}_{\mathit{I}}$ to simulate an equal amount $n_{\mathit{I}} = 
-n_{\mathit{P}}$ of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree 
-$\tau_{\mathit{I}}$, under the process $\mathit{I}$. We then impose that 
-$\tau_{\mathit{P}}$ and $\tau_{\mathit{I}}$ must respect the same topology
-\begin{align}
-    \psi_{\mathit{I}} = \psi_{\mathit{P}}.
-\end{align}
+
+We use $\theta^{*}_{\mathit{I}}$ to simulate an equal amount 
+$n_{\mathit{I}} = n_{\mathit{P}}$ 
+of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree 
+$\tau_{\mathit{I}}$, under the process $\mathit{I}$, 
+while preserving the same topology.
+
 Using a twin tree serves as a control: both $\tau_{\mathit{P}}$ and 
-$\tau_{\mathit{I}}$ will produce an error distribution. We assume that these 
-errors are a natural product of the input tree and the intermediate steps of 
-the pipeline. All the factors that can cause such errors are: the branching 
+$\tau_{\mathit{I}}$ will produce an error distribution. 
+These errors are unavoidably accumulated in the \verb;pirouette; pipeline. 
+All the factors that can cause such errors are: the branching 
 times $\vec{t}$ of the tree (as a consequence of the tree prior), its topology 

### 2019-04-10 article.tex
Updates from Overleaf

--- 2019-04-09 article.tex
@@
 model, is computed millions of times.
-Due to this, effective but computationally expensive tree priors are inviable.
+Due to this, effective but computationally expensive tree priors are not viable.
 Also, there may be no need to add such a tree prior to a Bayesian
@@
 \richel{
-  I think this is a usefull paragraph, adding some depth to this paper,
+  I think this is a useful paragraph, adding some depth to this paper,
   but it could use some rewording
 }
-There have been multiple attemps to investigate the importance of tree
+There have been multiple attempts to investigate the importance of tree
 prior choice. For example, recently Sarver et al., [\cite{sarver2019choice}] 

### 2019-04-10 article.tex
Reviewed description

--- 2019-04-09 article.tex
@@
 
-\richel{
-  I think this is a usefull paragraph, adding some depth to this paper,
-  but it could use some rewording
-}
-There have been multiple attemps to investigate the importance of tree
+There have been multiple attempts to investigate the importance of tree
 prior choice. For example, recently Sarver et al., [\cite{sarver2019choice}] 
@@
 $\tau_{\mathit{I}}$ will produce an error distribution. 
-These errors are unavoidably accumulated in the \verb;pirouette; pipeline. 
-All the factors that can cause such errors are: the branching 
-times $\vec{t}$ of the tree (as a consequence of the tree prior), its topology 
-$\psi$, the stochasticity induced by simulating the alignment (step 1a 
-$\rightarrow$ 2a in Fig.~\ref{fig:pipeline}) and the stochasticity induced by 
-the Bayesian inference (step 2a $\rightarrow$ 4a in Fig.~\ref{fig:pipeline}). 
-By imposing the same topology and replicating the pipelines a sufficient number 
-of times we can sort out all the possible causes that are not a direct 
-consequence of the choice of the tree prior.
-
-Potential differences in shape of error distributions will therefore be 
-consequences of different tree priors for the two parallel pipelines: the use 
-of prior $\mathit{P}$ for the original pipeline and of prior $\mathit{I}$ for 
-the twin pipeline. In the second case, though, such error distribution is not 
-influenced by the mismatch between how the tree is generated and how it is 
-analyzed by BEAST2.
+These errors unavoidably accumulate in the \verb;pirouette; pipeline,
+in both the true and twin tree: either tree may be an unlikely representative
+of the underlying tree prior, the alignment simulation may give different
+numbers of mutations for a same evolutionary time, 
+and the MCMC will sample different states.
+These errors can be equalized by imposing a same tree topology and by
+replicating the pipelines a sufficient number.
+
+The left-over differences in error distributions will be 
+consequences of different phylogeny and tree priors for the two parallel 
+pipelines: the main pipeline uses tree $\tau_{\mathit{P}}$ and
+possibly mismatched prior $\mathit{P}$, where the twin pipeline
+uses tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
+So, for the twin route, the resulting error distribution cannot be 
+attributed to a mismatch between true tree prior and tree prior used in
+inference.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \verb;pir_run; & Run \verb;pirouette; \\
+    \verb;pir_plot; & Show the \verb;pirouette; results as a plot \\
     \verb;create_pir_params; & Create the \verb;pirouette; parameters \\
@@
 The error distribution returned by \verb;pir_run; is stored as a data frame. 
-\richel{
-  I suggest \texttt{pir\_run} to return a list of run experiments,
-  as described in Issue 225, 
-  \url{https://github.com/richelbilderbeek/pirouette/issues/225}
-}
 It can be directly plotted using \verb;pir_plot;:

### 2019-04-10 article.tex
Reviewing ERQ 1

@@
 phylogenies in a posterior. Due to stochasticity, all posterior trees
-will differ from the given phylogeny, even when the generative tree
-prior $\mathit{P}$ and alignment model $\mathit{A}$ are the same as 
+will differ from the given phylogeny $\tau_{\mathit{P}}$, 
+even when the generative tree prior $\mathit{P}$ 
+and alignment model $\mathit{A}$ are the same as 
 those in the inference model $\mathit{I}$.
@@
 given phylogeny. 
-To do so, we must specify an alignment model $\mathit{A}$ which consists of: a 
-DNA root sequence $\mathit{r_{s}}$, the DNA mutation rate $\mathit{m_{r}}$, a
+To do so, we must specify an alignment model $\mathit{A}$ which consists of a 
+DNA root sequence $\mathit{r_{s}}$, a DNA mutation rate $\mathit{m_{r}}$, a
 site model $\mathit{s_{m}}$ (in this context: a nucleotide substitution model) 
@@
 In this example, the DNA root sequence consists in four blocks of 250 
-nucleotides each, 
+mononucleotides each, 
 where the per-nucleotide mutation rate is 0.1 mutations per unit time.
@@
 and a strict clock model as these are the simplest.
-A Jukes-Cantor site model
+A JC site model
 assumes that mutation rates between nucleotides are equal and constant. 
@@
 
-The error distribution returned by \verb;pir_run; is stored as a data frame. 
-It can be directly plotted using \verb;pir_plot;:
+The error distribution returned by \verb;pir_run; is a data frame,
+that can be plotted directly using \verb;pir_plot;:
 
@@
   \caption{
-    Example 1: the error BEAST2 makes from a phylogeny 
+    Example 1: the inference error made from a phylogeny 
     when the generative and inference model are the same.

### 2019-04-10 article.tex
Reviewed ERQs 1 and 2, part of 3

@@
 
-The objects we just created need to be bundled
+The objects we just created must be bundled
 in a \verb;pirouette; parameters structure,
@@
 
-We will not use twinning
-here yet.  
+We will not use twinning here yet.  
 All these arguments are bundled
@@
 
-The error distribution returned by \verb;pir_run; is a data frame,
-that can be plotted directly using \verb;pir_plot;:
+The error distribution can be plotted directly using \verb;pir_plot;:
 
@@
   \caption{
-    Example 1: the inference error made from a phylogeny 
+    ERQ 1: the inference error made from a phylogeny 
     when the generative and inference model are the same.
+    The vertical dashed line indicates the median error
   }
@@
 
-The error distribution obtained is shown as a violin plot
-in figure~\ref{fig:example_1}
-\richel{
-  I suggest to use a histogram (or density plot) instead:
-  I like how the error distribution in figure 1 looks.
-}
-\giovanni{Can we show everything at the same time with standard histograms?}
-.
-
-This error distribution has little relevance on its own, 
-as we constructed the experiment in such a way $\mathit{I}$ 
-matches with the generative model given by $\mathit{A}$ and $\mathit{P_{0}}$.
-However it will become clear when compared to the errors obtained from a 
-phylogeny 
-derived from a non-standard diversification process $\mathit{P}$.
-
-By doing so, this error distribution will serve as a control,
-as it can be seen as the baseline error caused by stochasticity.
-If the two error distributions obtained from $\mathit{P_{0}}$ and $\mathit{P}$ 
-will appear similar, the error induced by the choice of a non-standard 
-speciation process will be of little influence.
-Given the necessity of such comparison,
-we made it an inherent part of \verb;pirouette;. 
+The resulting error distribution, as shown figure~\ref{fig:example_1}),
+shows the inference error 
+when $\mathit{I}$ matches with the generative model given 
+by $\mathit{A}$ and $\mathit{P_{0}}$.
+This error distribution can serve as a control,
+as it is obtained from a tree of which the tree prior is standard and known.
+Obtaining this control is an inherent part of \verb;pirouette;. 
 We call it 'twinning' and we use it from ERQ 3.
@@
 with a known generative tree prior $\mathit{P_{0}}$.
-However, a novel tree prior $\mathit{P}$, by definition, is non-standard, thus 
-cannot be part of a standard inference model (i.e. an inference model using a 
-standard tree prior).
-In this case the user has to specify the the tree prior $\mathit{t_{p}}$ to be 
-used in the inference from a list of standard tree priors. In general, having 
-no information on $\mathit{P}$, this choice is arbitrary.
-\verb;pirouette; provides two ways to deal with this problem. The choice can be 
-made specifying the inference conditions $\mathit{C}$ to be used.
-
-If the user has sufficient reasons to believe that one standard tree prior 
-could be more suitable than others for the inference, they can set it as part 
-of the inference model along with site and clock models used to simulate the 
-alignment. This option can be specified selecting "model\_type" to be 
-"generative" as part of the inference conditions (as seen in ERQ 1, 
-table~\ref{tab:RQ1}).
-
-If, instead, there is no preference for a specific tree prior, the user can let 
-\verb;pirouette; select the best inference model from a list of candidate 
-models. The model selection is performed automatically based on the criterion 
-of the highest evidence (also known as marginal likelihood). This option can be specified 
-providing any number of experiments, selecting "model\_type" to be "candidate" 
-for each of them, and "run\_if" to be "best\_candidate" (table~\ref{tab:RQ2} 
-will summarize this for ERQ 2).
+However, a novel tree prior $\mathit{P}$, by our definition, is non-standard, 
+thus cannot be part of a standard inference 
+model (i.e. an inference model using a standard tree prior).
+In such a case, the user has to somehow select 
+a standard tree prior $\mathit{t_{p}}$ to be 
+used in the inference.
+There are two ways to deal with this problem: hand-pick a tree
+prior, or let the best of a set of models be picked. 
+Hand-picking a tree prior can be rational if one standard tree prior 
+is more suitable than others for the inference, for example, when
+the novel tree prior is most similar to a certain tree prior.
+Instead of hand-picking, the best inference model from a list of candidate 
+models can be used in inference.
+\verb;pirouette; selects the 'best' model, that is, 
+the model with the highest evidence (also known as marginal likelihood).
 
@@
 "What is the inference error made on a novel phylogeny when
-picking the best inference model?"
+using the best inference model?"
 
@@
 
-The first step in \verb;pirouette; is to simulate a DNA alignment from the 
+First step is to simulate a DNA alignment from the 
 given phylogeny. 
@@
 the same inference conditions $\mathit{C_{i}} = \mathit{C}$ but different 
-candidate inference models $\mathit{I_{i}}$. Yet, only one of them will be used 
+candidate inference models $\mathit{I_{i}}$. Only one of them will be used 
 to run the actual inference. According to the specified $\mathit{C}$, 
@@
 models, 2 clock models and 5 tree priors.
-\giovanni{Do we still want to use CBS, CCP and CEP?}
-\richel{Yes, as we don't know anything about the process underlying the tree}
 
@@
 We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
-and plot the results (listing~\ref{lst:pir_plot}).
-The resulting figure is shown in figure~\ref{fig:example_2}.
+and plot the results (listing~\ref{lst:pir_plot}),
+as shown in figure~\ref{fig:example_2}.
 In this case, the inference model with a JC69 site model, 
@@
 
-In ERQ 1 we measured the baseline error of the \verb;pirouette; method. We did so by using a standard tree prior to generate the original tree as well as in the inference.
-The goal of \verb;pirouette; is to quantify the 
-influence of a new (non-standard) tree prior.
-To do so, we need to compare the inference error 
-from the new tree prior to a baseline error,
+In ERQ 1 we measured the baseline error of the \verb;pirouette; method,
+by using the same standard tree prior in both the generation of 
+the original tree, as well as in the inference.
+This baseline error distribution serves as a control,
+to quantify the impact of a new (non-standard) tree prior.
+To measure that impact, we compare the inference error 
+from the new tree prior to the baseline error,
 by using a twin tree (see subsection~\ref{subsec:twinning}).
-To do so, we use the twin tree as a control for the main pipeline.
-In this example, we have no knowledge of the generative tree prior $\mathit{P}$
-but we know the alignment model $\mathit{A}$.
-In this ERQ it is reasonable to hand-pick a standard tree prior 
-to serve as a stand-in generative tree prior as we assume that they are closely related.
+In this ERQ, we hand-pick a standard tree prior $\mathit{P}$.
 
@@
 
-"What is the error made by BEAST2 from a phylogeny, 
+"What is the inference error made from a phylogeny, 
 when hand-picking a tree prior, compared to the background noise?"
@@
 Most of the settings for this ERQ are
-the same as in the previous section. They are
-the phylogeny from an unknown speciation model 
-(listing~\ref{lst:unknown_phylogeny}), 
+the same as in ERQ 2: we'll re-use the phylogeny from an 
+unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
 the alignment parameters (listing~\ref{lst:create_alignment})
@@
 
-This time, however, we are interested in measuring both the inference error
-and a baseline error. We do so by setting the twinning parameters $\mathit{T}$ to create a twin tree. A twin tree
-has the same topology as the given tree, yet its branching times are simulated 
-according to the standard tree prior used by BEAST2 to produce the posterior 
-for the main pipeline. 
-Since in the twin pipeline there is coherence between the inference model and the generative model, 
-we expect BEAST2 to produce a posterior of phylogenies that are more similar to the 
-original one, respect to the outcome of the main pipeline. 
-Due to this, the twinning process is used to provide a control for the main 
-experiment.
-
-Creating the twinning parameters is easy, as it has sensible default settings.
-For the sake of clarity, we explicitly show the default values of the most
+This time, however, we measure both the baseline and complete inference error.
+We set the twinning parameters $\mathit{T}$ to create a twin tree,
+as described in the 'Twinning' section.
+
+Creating the twinning parameters is trivial with the default settings.
+For clarity, however, we explicitly show the most
 important arguments:
+
+\richel{hiero!}
 

### 2019-04-12 article.tex
Re-run

@@
 \usepackage{mathtools}
-% \usepackage{float}
+\usepackage{float}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \includegraphics[width=\textwidth]{example_1/errors.png}
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \includegraphics[width=\textwidth]{example_2/errors.png}
   \caption{
-    Example 2: the error BEAST2 makes from a phylogeny when
-    picking the best inference model.
+    ERQ 2: the inference error from
+    when selecting the best inference model.
   }
   \label{fig:example_2}
-  \giovanni{There is a weird NA at the top of the figure.}
 \end{figure}
@@
 important arguments:
-
-\richel{hiero!}
 
@@
 
-\richel{
-  I would enjoy being able to write ...
-}
-\begin{lstlisting}
-  twinning_params <- create_twinning_params(
-    tree_prior = create_bd_tree_prior(),
-    ...
-  )
-)
-\end{lstlisting}
-\richel{
-  ... as that would connect the tree priors more
-}
-
-We bundle again all the \verb;pirouette; arguments using \verb;create_pir_params;:
+We combine all the parameters using \verb;create_pir_params;:
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \includegraphics[width=\textwidth]{example_3/errors.png}
   \caption{
-    Example 3: the error made by BEAST2 from a phylogeny, 
+    ERQ 3: the inference error made 
     when hand-picking a generative tree prior versus the control.
@@
 
-To answer ERQ 2 we used an evidence-based model selection procedure to obtain 
-an 
-error distribution.
-However, it is unclear to what extent this error is due to the mismatch between 
-the tree priors (from generative and inference models) or due to stochasticity. 
-As mentioned in ERQs 1 and 3, \verb;pirouette; provides a tool to evaluate this, using 
-the twinning process. We now apply it to the same pipeline followed in ERQ 2.
+When no inference model can be hand-picked, 
+one can select a best model from a set,
+as shown in ERQ2.
+The obtained error distribution contains the error accumulated in the pipeline,
+without quantifying to what extent this error is due to the mismatch between 
+the generative tree prior $\mathit{P}$ and the selected inference model 
+$\mathit{I}$.
+As twinning (see section 'Twinning') shows us the baseline error,
+we can quantify the contribution of the tree prior to the error.
 
@@
 
-"What is the error made by BEAST2 from a phylogeny 
+"What is the inference error from a phylogeny 
 derived from an unknown speciation model,
 when selecting the best inference model, 
-in relation to the background error?"
+in relation to the baseline error?"
 
@@
 generated by an unknown speciation model (listing~\ref{lst:unknown_phylogeny}). 
-The first step in \verb;pirouette; is to simulate a DNA alignment 
+The first step is to simulate a DNA alignment 
 from the given phylogeny, for which we will use the same code 
@@
 \end{lstlisting}
-\richel{
-  Refer to earlier listing for this call to \texttt{create\_pir\_params}
-}
 
@@
 \end{lstlisting}
-\richel{
-  Refer to earlier listing for this call to \texttt{pir\_run}
-}
 
@@
 \end{lstlisting}
-\richel{
-  Refer to earlier listing for this call to \texttt{pir\_run}
-}
-
-The output is shown in figure~\ref{fig:example_4}\giovanni{figure~\ref{fig:example_4} needs to be closer}. The alignment
+
+The output is shown in figure~\ref{fig:example_4}. The alignment
 generated for the true tree had most evidence for the JC site model,
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \includegraphics[width=\textwidth]{example_4/errors.png}
@@
   \label{fig:example_4}
-  \giovanni{NA on the top.}
 \end{figure}
@@
 The output is shown in figure~\ref{fig:example_5}.
-\giovanni{figure~\ref{fig:example_5} needs to be closer}
 The alignment
@@
   }
-  \giovanni{na on the top-right area. labels on x superimpose each other. We should change the names to their shortened versions: e.g. relaxed\_log\_normal $\to$ rln}
   \label{fig:example_5}

### 2019-04-12 article.tex
Changes @Giappo

--- 2019-04-10 article.tex
@@
   \item from a given phylogeny an alignment is simulated 
-    under a known alignment model;
-  \item from this alignment, according to the specified inference conditions, 
-    an inference model is picked (which may differ from the 
+    under a known alignment model $\mathit{A}$;
+  \item from this alignment, according to the specified inference conditions $\mathit{C}$, 
+    an inference model $\mathit{I}$ is picked (which may differ from the 
     generative alignment model);
@@
   \item the phylogenies in the posterior are compared with the given phylogeny 
-    to estimate the error made, according to the error measure specified 
+    to estimate the error made, according to the error measure $\mathit{E}$ specified 
     by the user;

### 2019-04-15 article.tex
Changes @Giappo

@@
 Other birth-death models treat speciation as a process that takes 
-time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}], 
+time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
 or that happens in bursts of simultaneous 
@@
 in which phylogenies are simulated from known parameters, 
-after which the parameters are estimated from the phylogenies.
+and then the model is used to infer back the parameters from the phylogenies.
 Ideally, the estimated parameter values closely match the known/true values.
@@
 model, is computed millions of times.
-Due to this, effective but computationally expensive tree priors are inviable.
+Due to this, effective but computationally expensive tree priors are non-viable.
 Also, there may be no need to add such a tree prior to a Bayesian
@@
 These investigations show how current diversification models compare
-to one another, but these do not help assess the importance of a 
-new tree prio.
+to one another, but these do not help to assess the importance of a 
+new tree prior.
 
@@
 to a standard tree prior. For example, a protracted birth-death model
-becomes equivalent to a birth-death when speciation rate approaches infinity.
+becomes equivalent to a standard birth-death model when speciation-completion rate approaches infinity.
 When benchmarking a novel tree prior, 
@@
     generative alignment model);
-  \item the alignment and inference model are used 
-    to infer a posterior distribution;
+  \item the inference model and the alignment are used 
+    to infer a posterior distribution of phylogenies;
   \item the phylogenies in the posterior are compared with the given phylogeny 
@@
 a site model and a (strict) clock model.
-The DNA root sequence is the DNA sequence the shared common ancestor has,
+The DNA root sequence is the DNA sequence of the shared common ancestor,
 and is set to four equally-sized mononucleotide blocks by default, as this
@@
 model. Supported site models are JC, HKY, TN and GTR. Only the strict
-clock model is supported.
+clock model is supported in this step.
 
@@
 We define an experiment $X$ as the combination of an inference model $I$
-and the conditions, $C$ to actually use it in the inference step.
+and the conditions $C$ to actually use it in the inference step.
 There are two main setups of condition $C$: either the attached
@@
 or the attached inference model is a candidate model, that will be
-be selected if it is the best model.
+be selected if it is the best model \giovanni{then why do you need the "run\_if" option?}.
 When setting $C$ to be attached to the generative model,
@@
 and those inference models that were selected in the previous step. 
-Per selected experiment a posterior is inferred, using the 
+For each selected experiment a posterior is inferred, using the 
 \verb;babette; [\cite{bilderbeek2018babette}] R package. 
@@
 It creates an error distribution, by measuring the difference
-between the one true tree and each of the posterior 
+between the true tree and each of the posterior 
 trees (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
@@
 user-defined error statistic can be used.
-An MCMC needs to 'warm up', as initially it will sample
+\giovanni{How is the MCMC topic connected to what previously said?} An MCMC needs to 'warm up', as initially it will sample
 an unrepresentative part of state space. This phase, called 
@@
 The twinning process, T, in accordance with a chosen inference 
-model $t_p = \mathit{I}$, creates a tree $\tau_{\mathit{I}}$
+model $t_p \in \mathit{I}$, creates a tree $\tau_{\mathit{I}}$
 with branching times $\Vec{t}_{\mathit{I}}$ while preserving 
@@
 (e.g. speciation and extinction rates, in case of a birth-death model) 
-that maximize likelihood $L_{\mathit{I}}$ applied 
+that maximize the likelihood $L_{\mathit{I}}$ applied 
 to the true tree, conditioned on its number of tips $n_{\mathit{P}}$,
@@
 table~\ref{tab:functions}. 
-These functions are demonstrated when in the example code below.
+Their usage is demonstrated in the example code below.
 All \verb;pirouette;'s functions are documented,
@@
 DNA root sequence $\mathit{r_{s}}$, a DNA mutation rate $\mathit{m_{r}}$, a
-site model $\mathit{s_{m}}$ (in this context: a nucleotide substitution model) 
+site model $\mathit{s_{m}}$ (in this context: a nucleotide substitution model \giovanni{We already specified this in 2.1, when explaining the first step. Do we need to repeat it?}) 
 and a clock model $\mathit{c_{m}}$. 
@@
     Create the experiments to answer research question 1, 
-    using explicit arguments
+    using explicit arguments.
   }
@@
     when the generative and inference model are the same.
-    The vertical dashed line indicates the median error
+    The vertical dashed line indicates the median error.
   }
@@
 Most of the settings for this ERQ are
-the same as in ERQ 2: we'll re-use the phylogeny from an 
+the same as in ERQ 2: we will re-use the phylogeny from an 
 unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
@@
 We set the twinning parameters $\mathit{T}$ to create a twin tree,
-as described in the 'Twinning' section.
+as described in the 'Twinning' section \ref{subsec:twinning}.
 
@@
   label = {lst:create_twinning_params},
-  caption = Create the default twinning parameters
+  caption = Create the default twinning parameters.
 ]
@@
 When no inference model can be hand-picked, 
-one can select a best model from a set,
+one can select the best model from a set,
 as shown in ERQ2.
@@
 $\mathit{I}$.
-As twinning (see section 'Twinning') shows us the baseline error,
+As twinning (see section \ref{subsec:twinning}: 'Twinning') shows us the baseline error,
 we can quantify the contribution of the tree prior to the error.
@@
 Instead of using the common Yule tree as in example 1, we will use
-a Yule tree of low (actually: zero) likelihood \giovanni{what do you mean by zero likelihood?}, i.e. the one already used in ERQs 2-4, by using the code shown in listing~\ref{lst:unknown_phylogeny}.
+a Yule tree of low (actually: zero) likelihood \giovanni{What do you mean by zero likelihood?
+This word might have a heavy meaning int his context...}, i.e.
+the one already used in ERQs 2-4, by using the code shown in listing~\ref{lst:unknown_phylogeny}.
 The first step in \verb;pirouette; is to simulate a DNA alignment 

### 2019-04-15 scripts/example_1.R
Use 95% of the error values

@@ -64,6 +64,11 @@ errors <- pir_run(
 )
 
 if (1 == 2) {
+  errors <- utils::read.csv(
+    file = file.path(example_folder, "errors.csv")
+  )
+  check_pir_out(errors)
+
   pir_plot(errors)
 }
 

### 2019-04-15 scripts/example_2.R

@@ -73,6 +73,10 @@ errors <- pir_run(
 Sys.time()
 
 if (1 == 2) {
+  errors <- utils::read.csv(
+    file = file.path(example_folder, "errors.csv")
+  )
+  check_pir_out(errors)
   pir_plot(pir_out = errors)
 }
 

### 2019-04-15 scripts/example_3.R

@@ -75,6 +75,10 @@ errors <- pir_run(
 Sys.time()
 
 if (1 == 2) {
+  errors <- utils::read.csv(
+    file = file.path(example_folder, "errors.csv")
+  )
+  check_pir_out(errors)
   pir_plot(pir_out = errors)
 }
 

### 2019-04-15 scripts/example_4.R

@@ -71,6 +71,10 @@ errors <- pir_run(
 Sys.time()
 
 if (1 == 2) {
+  errors <- utils::read.csv(
+    file = file.path(example_folder, "errors.csv")
+  )
+  check_pir_out(errors)
   pir_plot(pir_out = errors)
 }
 

### 2019-04-15 scripts/example_5.R

@@ -99,6 +99,10 @@ errors <- pir_run(
 Sys.time()
 
 if (1 == 2) {
+  errors <- utils::read.csv(
+    file = file.path(example_folder, "errors.csv")
+  )
+  check_pir_out(errors)
   pir_plot(pir_out = errors)
 }
 

### 2019-04-15 article.tex
Put ERQ5 figure here
Removed duplicate listings

@@
   \textbf{4. }
-    Last, we discuss the results obtained by the examples. \\
+    Last, we discuss the results obtained by the examples,
+    their interpretation. \\
 \end{abstract}
@@
 unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
-the alignment parameters (listing~\ref{lst:create_alignment})
-and the list of experiments (listing~\ref{lst:create_generative_experiment}),
-the latter stating that the generative tree prior is a Yule tree prior.
+the alignment parameters (listing~\ref{lst:create_alignment}).
+
+We will create a list of experiments as in 
+listing~\ref{lst:create_generative_experiment},
+but with all defaults:
+
+\begin{lstlisting}[
+  language=R, 
+  floatplacement=ht, 
+  frame=single,
+  label = {lst:create_gen_experiment_defaults},
+  caption = Create a list of one default generative experiment.
+]
+experiments <- list(create_gen_experiment())
+\end{lstlisting}
 
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[
+  language=R, 
+  floatplacement=ht, 
+  frame=single,
+  label = {lst:create_pir_params_with_twinning},
+  caption = Create the default twinning parameters.
+]
 pir_params <- create_pir_params(
@@
 
-Finally, we run and plot the results:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-pir_plot(errors)
-\end{lstlisting}
-
+We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
+and plot the results (listing~\ref{lst:pir_plot}),
 The output is shown in figure~\ref{fig:example_3}
@@
 To get an idea of a baseline error made by BEAST2, 
-we also make use of the twinning option by using the
-default twinning parameters, as shown in 
-listing~\ref{lst:create_twinning_params}.
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  error_measure_params = error_measure_params,
-  twinning_params = twinning_params
-)
-\end{lstlisting}
-
-We run the experiment:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-\end{lstlisting}
-
-and show the results:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-pir_plot(errors)
-\end{lstlisting}
-
-The output is shown in figure~\ref{fig:example_4}. The alignment
+we also make use of the twinning option by using
+the same pir\_params XXXXXXXXXXXXXXXXXXXXXXXXXXXx as shown in 
+listing~\ref{lst:create_pir_params_with_twinning}.
+We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
+and plot the results (listing~\ref{lst:pir_plot}).
+The output is shown in figure~\ref{fig:example_4}.
+
+The alignment
 generated for the true tree had most evidence for the JC site model,
@@
 
-In ERQs 1 and 2 we used \verb;pirouette; to  
-run the analysis on the main pipeline, either using a 'generative' or a 
-'best\_candidate' inference model, resulting in one error distribution,
-that does not mean much on its own.
-In ERQs 3 and 4 we measured the baseline error caused by stochasticity,
-by running a parallel twin pipeline. This yielded two error distributions:
-the baseline error and the complete inference error. Here we want to combine all the instances of the previous ERQs at once.
+When you are moderately sure about hand-picking an inference model,
+but would also like to check how well your hand-picked model compares
+to others, we can do both. Of course, you are interested in the
+baseline error for both the hand-picked and best candidate model,
+so you'll enable twinning as well. This example is a combination
+of the previos two ERQs.
 
@@
 Instead of using the common Yule tree as in example 1, we will use
-a Yule tree of low (actually: zero) likelihood \giovanni{What do you mean by zero likelihood?
-This word might have a heavy meaning int his context...}, i.e.
-the one already used in ERQs 2-4, by using the code shown in listing~\ref{lst:unknown_phylogeny}.
-The first step in \verb;pirouette; is to simulate a DNA alignment 
+a Yule tree of low likelihood, 
+i.e. the one already used in ERQs 2-4, 
+by using the code shown in listing~\ref{lst:unknown_phylogeny}.
+The first step is to simulate a DNA alignment 
 from the given phylogeny, for which we will use the same code 
@@
 \begin{lstlisting}[language=R, floatplacement=H, frame=single]
-generative_experiment <- create_experiment(
-  inference_conditions = create_inference_conditions(
-    model_type = "generative",
-    run_if = "always",
-    do_measure_evidence = TRUE
-  )
-  inference_model = create_inference_model(
-    site_model = create_jc69_site_model(),
-    clock_model = create_strict_clock_model(),
-    tree_prior = create_yule_tree_prior()
-  )
-)
+generative_experiment <- create_gen_experiment()
 \end{lstlisting}
@@
 \end{lstlisting}
-\giovanni{Does this code actually work?}
 
 We choose again to use the default error measurement parameters
-created in listing~\ref{lst:create_error_measure_params}
-and combine this with the same twinning parameters 
-used in listing~\ref{lst:create_twinning_params}:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  error_measure_params = error_measure_params,
-  twinning_params = twinning_params
-)
-\end{lstlisting}
-
-We run the experiment in the usual fashion:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-\end{lstlisting}
-
-We plot the results:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-pir_plot(errors)
-\end{lstlisting}
-
+created in listing~\ref{lst:create_error_measure_params},
+with twinning, as alsready shown in 
+listing~\ref{lst:create_pir_params_with_twinning}. XXXXXXXXXXXXXXXXXXXXX
+
+We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
+and plot the results (listing~\ref{lst:pir_plot}).
 The output is shown in figure~\ref{fig:example_5}.
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
   \includegraphics[width=\textwidth]{example_5/errors.png}
   \caption{
-    Example 5: the error made by BEAST2 from a phylogeny, 
-    when the generative model is known, 
-    compared to candidates with similar inference models, 
-    in relationship to the background error.
+    Example 5: the inference error made, 
+    when the generative model is hand-picked, 
+    compared to the best of 39 candidate models, 
+    each in relationship to the background error.
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{
-  I think we can put the interpretation in a table as well.
-  Question is: do you agree with the interpretations I gave?
-  Please correct me.
-}
 

### 2019-04-15 scripts/example_3.R

@@ -21,7 +21,7 @@ alignment_params <- create_alignment_params(
   mutation_rate = 0.1
 )
 
-experiments <- list(create_experiment())
+experiments <- list(create_gen_experiment())
 
 # Testing
 if (1 == 2) {

### 2019-04-15 scripts/example_5.R

@@ -36,6 +36,7 @@ generative_experiment <- create_experiment(
     mcmc = create_mcmc(chain_length = 1e+07, store_every = 1000)
   )
 )
+generative_experiment <- create_gen_experiment()
 check_experiment(generative_experiment)
 
 # All non-Yule tree priors

### 2019-04-15 article.tex
Reference to better superior listings

@@
 
-Listing~\ref{lst:create_generative_experiment_explicit} shows how to
+Listing~\ref{lst:create_gen_experiment_explicit} shows how to
 set up this experiment:
@@
   floatplacement=ht, frame=single,
-  label = {lst:create_generative_experiment_explicit},
+  label = {lst:create_gen_experiment_explicit},
   caption = {
@@
   floatplacement=ht, frame=single,
-  label = {lst:create_generative_experiment},
+  label = {lst:create_gen_experiment},
   caption = {
@@
 
-We will create a list of experiments as in 
-listing~\ref{lst:create_generative_experiment},
+We will create an experiment as in 
+listing~\ref{lst:create_gen_experiment},
 but with all defaults:
@@
 ]
-experiments <- list(create_gen_experiment())
-\end{lstlisting}
+experiment <- create_gen_experiment()
+\end{lstlisting}
+
+Creating the complete set of experiments:
+
+YYYYYYYYYYYYYYYYYYYYYYYYYYYYYYYY
+experiments <- list(experiment)
 
@@
 We choose again to use the default error measurement setup,
-as shown in listing~\ref{lst:create_error_measure_params}.
+as shown in listing~\ref{lst:create_error_measure_params},
+which is the default in \verb;create_pir_params;, so we'll
+leave it out for now.
 

### 2019-04-15 scripts/example_1.R
Transpose ESSes

@@ -22,7 +22,7 @@ alignment_params <- create_alignment_params(
   mutation_rate = 0.1
 )
 
-experiment <- create_experiment()
+experiment <- create_gen_experiment()
 experiments <- list(experiment)
 
 # Testing
@@ -88,9 +88,11 @@ esses <- tracerer::calc_esses(
   sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
 )
 
+df_esses <- data.frame(parameter = colnames(esses), ESS = as.character(esses))
+
 sink(file.path(example_folder, "esses.latex"))
 xtable::print.xtable(
-  xtable::xtable(esses, caption = "ESSes of example 1", label = "tab:esses_example_1", digits = 0),
+  xtable::xtable(df_esses, caption = "ESSes of example 1", label = "tab:esses_example_1", digits = 0),
   include.rownames = FALSE
 )
 sink()

### 2019-04-15 scripts/example_2.R

@@ -95,9 +95,12 @@ esses <- tracerer::calc_esses(
   traces = tracerer::parse_beast_log(pir_params$experiments[[1]]$beast2_options$output_log_filename),
   sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
 )
+
+df_esses <- data.frame(parameter = colnames(esses), ESS = as.character(esses))
+
 sink(file.path(example_folder, "esses.latex"))
 xtable::xtable(
-  esses,
+  df_esses,
   caption = paste0("ESSes of example ", example_no),
   label = paste0("tab:esses_example_", example_no),
   digits = 0

### 2019-04-15 scripts/example_3.R

@@ -96,6 +96,7 @@ esses <- tracerer::calc_esses(
   traces = tracerer::parse_beast_log(pir_params$experiments[[1]]$beast2_options$output_log_filename),
   sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
 )
+df_esses <- data.frame(parameter = colnames(esses), ESS = as.character(esses))
 
 print("#######################################################################")
 print("ESSes")
@@ -103,7 +104,7 @@ print("#######################################################################")
 sink(file.path(example_folder, "esses.latex"))
 xtable::print.xtable(
   xtable::xtable(
-    esses,
+    df_esses,
     caption = paste0("ESSes of example ", example_no),
     label = paste0("tab:esses_example_", example_no),
     digits = 0

### 2019-04-15 scripts/example_4.R

@@ -166,11 +166,13 @@ esses_twin_best <- tracerer::calc_esses(
   traces = tracerer::parse_beast_log(to_twin_filename(pir_params$experiments[[2]]$beast2_options$output_log_filename)),
   sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
 )
+df_esses_best <- data.frame(parameter = colnames(esses_best), ESS = as.character(esses_best))
+df_esses_twin_best <- data.frame(parameter = colnames(esses_twin_best), ESS = as.character(esses_twin_best))
 
 sink(file.path(example_folder, "esses_best.latex"))
 xtable::print.xtable(
   xtable::xtable(
-    esses_best,
+    df_esses_best,
     caption = paste0("ESSes of example ", example_no, " for best candidate model"),
     label = paste0("tab:esses_example_", example_no, "_best"),
     digits = 0
@@ -183,7 +185,7 @@ sink()
 sink(file.path(example_folder, "esses_twin_best.latex"))
 xtable::print.xtable(
   xtable::xtable(
-    esses_twin_best,
+    df_esses_twin_best,
     caption = paste0("ESSes of example ", example_no, " for best candidate model, twin tree"),
     label = paste0("tab:esses_example_", example_no, "_twin__best"),
     digits = 0

### 2019-04-15 scripts/example_5.R

@@ -199,10 +199,15 @@ esses_twin_best <- tracerer::calc_esses(
   sample_interval = pir_params$experiments[[1]]$inference_model$mcmc$store_every
 )
 
+df_esses_gen <- data.frame(parameter = colnames(esses_gen), ESS = as.character(esses_gen))
+df_esses_best <- data.frame(parameter = colnames(esses_best), ESS = as.character(esses_best))
+df_esses_twin_gen <- data.frame(parameter = colnames(esses_twin_gen), ESS = as.character(esses_twin_gen))
+df_esses_twin_best <- data.frame(parameter = colnames(esses_twin_best), ESS = as.character(esses_twin_best))
+
 sink(file.path(example_folder, "esses_gen.latex"))
 xtable::print.xtable(
   xtable::xtable(
-    esses_gen,
+    df_esses_gen,
     caption = paste0("ESSes of example ", example_no, " for generative model"),
     label = paste0("tab:esses_example_", example_no, "_gen"),
     digits = 0
@@ -214,7 +219,7 @@ sink()
 sink(file.path(example_folder, "esses_best.latex"))
 xtable::print.xtable(
   xtable::xtable(
-    esses_best,
+    df_esses_best,
     caption = paste0("ESSes of example ", example_no, " for best candidate model"),
     label = paste0("tab:esses_example_", example_no, "_best"),
     digits = 0
@@ -226,7 +231,7 @@ sink()
 sink(file.path(example_folder, "esses_twin_gen.latex"))
 xtable::print.xtable(
   xtable::xtable(
-    esses_twin_gen,
+    df_esses_twin_gen,
     caption = paste0("ESSes of example ", example_no, " for generative model, twin tree"),
     label = paste0("tab:esses_example_", example_no, "_twin_gen"),
     digits = 0
@@ -238,7 +243,7 @@ sink()
 sink(file.path(example_folder, "esses_twin_best.latex"))
 xtable::print.xtable(
   xtable::xtable(
-    esses_twin_best,
+    df_esses_twin_best,
     caption = paste0("ESSes of example ", example_no, " for best candidate model, twin tree"),
     label = paste0("tab:esses_example_", example_no, "_twin__best"),
     digits = 0

### 2019-04-15 article.tex
Tree from unknown tree prior here

@@
 and the conditions $C$ to actually use it in the inference step.
-There are two main setups of condition $C$: either the attached
-inference model $I$ is the generative model and should always be selected,
+There are multiple setups of condition $C$: 
+either the attached inference model $I$ is the generative model and should always be selected,
 or the attached inference model is a candidate model, that will be
-be selected if it is the best model \giovanni{then why do you need the "run\_if" option?}.
+be selected if it is the best model, or will be selected always (regardless
+of it being the best). 
 When setting $C$ to be attached to the generative model,
@@
 user-defined error statistic can be used.
-\giovanni{How is the MCMC topic connected to what previously said?} An MCMC needs to 'warm up', as initially it will sample
+The error statistic is not applied on all posterior phylogenies 
+\giovanni{How is the MCMC topic connected to what previously said?} 
+XXXXXXXXXXXXXXXXXXXXXXXXXXXX33
+An MCMC needs to 'warm up', as initially it will sample
 an unrepresentative part of state space. This phase, called 
@@
 \end{lstlisting}
-\begin{figure}[ht]
+\begin{figure}[H]
   \includegraphics[width=\textwidth]{example_2/tree_unknown.png}
@@
 We choose again to use the default error measurement setup,
-as shown in listing~\ref{lst:create_error_measure_params},
-which is the default in \verb;create_pir_params;, so we'll
-leave it out for now.
+as shown in listing~\ref{lst:create_error_measure_params}.
 

### 2019-04-15 article.tex
Put creation of experiments in code
Describe burn-in better, thank @Giappo

@@
 
-The fourth step quantifies the inference error made.
-It creates an error distribution, by measuring the difference
+The fourth step quantifies the inference error made, by removing
+the burn-in, then applying an error statistic.
+The burn-in is the first phase of the MCMC,
+where it samples an unrepresentative part of state space,
+and is commonly discarded. By default, \verb;pirouette; 
+uses removes the first 10\% of the posterior.
+From the remaining posterior, \verb;pirouette; 
+creates an error distribution, by measuring the difference
 between the true tree and each of the posterior 
@@
 user-defined error statistic can be used.
-The error statistic is not applied on all posterior phylogenies 
-\giovanni{How is the MCMC topic connected to what previously said?} 
-XXXXXXXXXXXXXXXXXXXXXXXXXXXX33
-An MCMC needs to 'warm up', as initially it will sample
-an unrepresentative part of state space. This phase, called 
-the burn-in, is commonly set up to be a small part of the entire run, 
-after which it is discarded. By default, \verb;pirouette; 
-uses a cut-off value of 10\%.
 
@@
 
-YYYYYYYYYYYYYYYYYYYYYYYYYYYYYYYY
+\begin{lstlisting}[
+  language=R, 
+  floatplacement=ht, 
+  frame=single,
+  label = {lst:create_experiments},
+  caption = Create a list of one experiment.
+]
 experiments <- list(experiment)
+\end{lstlisting}
 
@@
 as we previously did in listing~\ref{lst:create_all_experiments}.
-
 We choose again to use the default error measurement setup,
 as shown in listing~\ref{lst:create_error_measure_params}.
-
 To get an idea of a baseline error made by BEAST2, 
 we also make use of the twinning option by using
-the same pir\_params XXXXXXXXXXXXXXXXXXXXXXXXXXXx as shown in 
+the same call to \verb;create_pir_params; as shown in 
 listing~\ref{lst:create_pir_params_with_twinning}.
@@
 We choose again to use the default error measurement parameters
-created in listing~\ref{lst:create_error_measure_params},
-with twinning, as alsready shown in 
-listing~\ref{lst:create_pir_params_with_twinning}. XXXXXXXXXXXXXXXXXXXXX
+created in listing~\ref{lst:create_error_measure_params}.
+We enable twinning, by calling \verb;create_pir_params;
+as shown in listing~\ref{lst:create_pir_params_with_twinning}.
 

### 2019-04-15 article.tex
No need to repeat what the alignment params are, thanks @Giappo

@@
 
-\subsection{pirouette's pipeline}
+\subsection{pirouette's pipeline}\label{subsec:pipeline}
 
@@
 The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny. 
-To do so, we must specify an alignment model $\mathit{A}$ which consists of a 
-DNA root sequence $\mathit{r_{s}}$, a DNA mutation rate $\mathit{m_{r}}$, a
-site model $\mathit{s_{m}}$ (in this context: a nucleotide substitution model \giovanni{We already specified this in 2.1, when explaining the first step. Do we need to repeat it?}) 
-and a clock model $\mathit{c_{m}}$. 
-In this example, the DNA root sequence consists in four blocks of 250 
+given phylogeny, as described in subsection~\ref{subsec:pipeline}.
+In this example, the DNA root sequence $\mathit{r_{s}}$ 
+consists in four blocks of 250 
 mononucleotides each, 
-where the per-nucleotide mutation rate is 0.1 mutations per unit time.
+where the per-nucleotide mutation rate $\mathit{m_{r}}$ 
+is 0.1 mutations per unit time.
 We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model

### 2019-04-15 article.tex
Moved @Giappo's feedback about the workflow figure to https://github.com/richelbilderbeek/pirouette_article/issues/26

@@
     resulting in a twin error distribution (5b)
-    \giovanni{Ok, that's my take on this: 1) all the capital letters should go 
-in a ball (or a rhombus, as you already did for C) which stands in the middle 
-of the arrows (as with C now); 2) The box in the middle should contain all the 
-inference models $I_{i}$; 3) We need another C at the intersection of the 
-arrows that point to T and A; 4) In between of the arrow 2a->4a there should be 
-another rhombus (or circle) with an I (or I' or I*, according on how we want to 
-call the selected inference model), which is pointed by the rhombus (of circle) 
-containing C; 5) The first arrow should start from a rhombus (or circle) with P 
-inside.}
-  }
   \label{fig:pipeline}

### 2019-04-15 article.tex
Fix references

@@
 
-\subsection{pirouette's pipeline}\label{subsec:pipeline}
+\subsection{pirouette's pipeline}
+\label{subsec:pipeline}
 
@@
     phylogeny (1a) using twinning parameters $T$, where $T$ may depend on
-    the experiments \richel{improve}.
+    the experiments.
     From a twin phylogeny, the same intermediate steps are taken,
     resulting in a twin error distribution (5b)
+  }
   \label{fig:pipeline}
@@
 \verb;pirouette; depends on multiple packages, which are:
-\richel{I think we can get rid of some dependencies, we just need to check}
-\verb;ape; (\cite{APE}), 
 \verb;babette; (\cite{bilderbeek2018babette}),

### 2019-04-15 article.tex
Use @Giappo's workflow figure
Added @Giappo's figure

@@
   \textbf{4. }
-    Last, we discuss the results obtained by the examples,
-    their interpretation. \\
+    Last, we discuss the results obtained by the examples and their interpretation. \\
 \end{abstract}
@@
 BEAST has built-in tree priors such as the Yule [\cite{yule}] and 
-(constant-rate) birth-death [\cite{nee1994reconstructed}] model.
+(constant-rate) birth-death [\cite{nee1994reconstructed}] models.
 These simple tree priors are most commonly used, as these 
@@
 it is not necessarily used in Bayesian inference.
-In Bayesian inference, the likelihood of a diversification 
-model, is computed millions of times.
+In these kind of procedures the likelihood of a diversification 
+model can be computed millions of times.
 Due to this, effective but computationally expensive tree priors are non-viable.
@@
 These investigations show how current diversification models compare
-to one another, but these do not help to assess the importance of a 
+to one another, but they do not help to assess the importance of a 
 new tree prior.
@@
   \centering
-  \includegraphics[width=\textwidth]{workflow.png}
+  \includegraphics[width=\textwidth]{workflow_alt.png}
   \caption{
@@
 and the conditions $C$ to actually use it in the inference step.
-There are multiple setups of condition $C$: 
+There are multiple setups of $C$: 
 either the attached inference model $I$ is the generative model and should always be selected,

### 2019-04-15 article.tex
Spellcheck

--- 2019-04-15 article.tex
@@
 the lower the influence the tree prior has.
-Due to this, computionally faster tree priors are more commonly used.
+Due to this, computationally faster tree priors are more commonly used.
 However, it is unknown if it is valid to assume that tree prior choice is 
@@
     TN93 = Tamura and Nei \cite{tamura1993estimation}, 
-    GTR = Generalised time-reversible model (\cite{tavare1986some}).
+    GTR = Generalized time-reversible model (\cite{tavare1986some}).
   }
@@
   language=R, floatplacement=ht, frame=single,
-  label = {lts:create_pir_params_defaults},
+  label = {lst:create_pir_params_defaults},
   caption = Create a \texttt{pir\_params} with many defaults.
@@
 so you'll enable twinning as well. This example is a combination
-of the previos two ERQs.
+of the previous two ERQs.
 
@@
 
-\verb;pirouette;'s is extensively documentated on its website,
+\verb;pirouette;'s is extensively documented on its website,
 its documentation and its vignettes.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgements}
+\section{Acknowledgments}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-04-15 article.tex
Updates from Overleaf

--- 2019-04-15 article.tex
@@
 and the conditions $C$ to actually use it in the inference step.
-There are multiple setups of $C$: 
-either the attached inference model $I$ is the generative model and should always be selected,
-or the attached inference model is a candidate model, that will be
-be selected if it is the best model, or will be selected always (regardless
+There are multiple possible setups for $C$. 
+If the attached inference model $I$ is the generative model, then it should always be selected.
+If, instead, the attached inference model is a candidate model the user can either let \verb;pirouette; select it as the best among all the candidate models, or force its selection (regardless
 of it being the best).
-When setting $C$ to be attached to the generative model,
-it is implied that attached inference model $I$ uses the 
+When $I$ is set to be the generative model,
+it will use the 
 same site and clock model used in the alignment simulation,
-and $I$ uses the same tree prior as used in the creation of the given phylogeny. 
+and the same tree prior used in the creation of the given phylogeny. 
 A 'best' model is defined as the inference model with

### 2019-04-15 article.tex
Scrapped lines

--- 2019-04-15 article.tex
@@
 as shown in figure~\ref{fig:example_2}.
-In this case, the inference model with a JC69 site model, 
-strict clock model and Yule tree prior has the highest likelihood
-amongst all candidates (table \ref{tab:evidences_example_2} shows
-the evidences).
 
@@
 
-The alignment
-generated for the true tree had most evidence for the JC site model,
-strict clock model and Yule tree prior, 
-with a weight of 0.981 (as 
-shown in table \ref{tab:evidences_example_4}). 
-The alignment
-generated for the twin tree had most evidence for the JC site model,
-RLN clock model and CBS tree prior,
-with a weight of 0.707 (as 
-shown in table \ref{tab:evidences_example_4}).
-
 \begin{figure}[H]
@@
 The output is shown in figure~\ref{fig:example_5}.
-The alignment
-generated from the true tree had most evidence for the JC site model,
-strict clock model and Yule tree prior, 
-with a weight of 0.981 \richel{again} (as 
-shown in table \ref{tab:evidences_example_5}). 
-The alignment
-generated for the twin tree had most evidence for the JC site model,
-RLN clock model and CBS tree prior,
-with a weight of 0.707 \richel{again} (as 
-shown in table \ref{tab:evidences_example_5_twin}).
 
@@
 \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Interpretation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-04-15 article.tex
Updates from Overleaf

--- 2019-04-15 article.tex
@@
 it will use the 
-same site and clock model used in the alignment simulation,
+same site and clock model used in the alignment simulation
 and the same tree prior used in the creation of the given phylogeny. 
@@
 the burn-in, then applying an error statistic.
-The burn-in is the first phase of the MCMC,
-where it samples an unrepresentative part of state space,
+The burn-in is the first phase of the Markov chain Monte Carlo (MCMC) run,
+in which it samples an unrepresentative part of state space,
 and is commonly discarded. By default, \verb;pirouette; 
-uses removes the first 10\% of the posterior.
-From the remaining posterior, \verb;pirouette; 
+removes the first 10\% of the posterior.
+From the remaining posterior \verb;pirouette; 
 creates an error distribution, by measuring the difference
@@
 but we cannot use the tree prior in Bayesian inference.
-The twinning process, T, in accordance with a chosen inference 
+The twinning process T, in accordance with a chosen inference 
 model $t_p \in \mathit{I}$, creates a tree $\tau_{\mathit{I}}$
-with branching times $\Vec{t}_{\mathit{I}}$ while preserving 
+with branching times $\Vec{t}_{\mathit{I}}$ while preserving the 
 topology $\psi_{\mathit{P}}$:

### 2019-04-17 article.tex
Updates from Overleaf

@@
 $\tau_{\mathit{I}}$ will produce an error distribution. 
-These errors unavoidably accumulate in the \verb;pirouette; pipeline,
+These errors unavoidably accumulate across the \verb;pirouette; pipeline,
 in both the true and twin tree: either tree may be an unlikely representative
 of the underlying tree prior, the alignment simulation may give different
-numbers of mutations for a same evolutionary time, 
+numbers of mutations for the same evolutionary time, 
 and the MCMC will sample different states.
-These errors can be equalized by imposing a same tree topology and by
-replicating the pipelines a sufficient number.
+These errors can be equalized by imposing the same tree topology and by
+replicating the pipelines a sufficient number of times.
 
 The left-over differences in error distributions will be 
-consequences of different phylogeny and tree priors for the two parallel 
-pipelines: the main pipeline uses tree $\tau_{\mathit{P}}$ and
+consequences of different phylogenies and tree priors for the two parallel 
+pipelines: the main pipeline uses the tree $\tau_{\mathit{P}}$ and a
 possibly mismatched prior $\mathit{P}$, where the twin pipeline
-uses tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
-So, for the twin route, the resulting error distribution cannot be 
-attributed to a mismatch between true tree prior and tree prior used in
+uses the tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
+So, for the twin route, if the process is repeated a sufficient number of times, the resulting error distribution cannot be 
+attributed to a mismatch between the true tree prior and the tree prior used in
 inference.
@@
 \end{lstlisting}
-For the option to select a best candidate model,
+For the option to select the best candidate model,
 \verb;pirouette; needs the "NS" BEAST2 package [\cite{maturana2018model}].
@@
 All \verb;pirouette;'s functions are documented,
-have a useful example, and have sensible defaults.
+have a useful example and sensible defaults.
 
@@
 used in the inference.
-There are two ways to deal with this problem: hand-pick a tree
+There are two ways to deal with this problem: hand-picking a tree
 prior, or let the best of a set of models be picked. 
@@
 
-We will create an experiment as in 
+We create an experiment as in 
 listing~\ref{lst:create_gen_experiment},
@@
 
-When you are moderately sure about hand-picking an inference model,
-but would also like to check how well your hand-picked model compares
-to others, we can do both. Of course, you are interested in the
+If we are moderately sure about hand-picking an inference model,
+but would also like to check how well the hand-picked model compares
+to the others, we can do both. Of course, as we are interested in the
 baseline error for both the hand-picked and best candidate model,
-so you'll enable twinning as well. This example is a combination
+we need to enable twinning as well. This example is a combination
 of the previous two ERQs.
@@
 In this example we use a tree generated by the Yule process.
-Instead of using the common Yule tree as in example 1, we will use
+Instead of using the common Yule tree as in ERQ 1, we will use
 a Yule tree of low likelihood, 
@@
 to be obtained from thousands of replicates.
-In that case, figure~\ref{fig:example_1} shows 
+In that case, figure~\ref{fig:example_1} would show 
 the minimum error BEAST2 when the
@@
 the comparison between the results obtained 
-on a new generative prior as compared to the error BEAST2 
+adopting a new generative prior as compared to the error BEAST2 
 would make when operating on a tree generated 

### 2019-04-24 article.tex
Changes @Giappo

@@
   \textbf{1. }
-    BEAST2 is a popular Bayesian phylogenetics software tool,
-    that takes an alignment of characters (usually nucleotides) 
-    and an inference model to create a
-    posterior of jointly-estimated phylogenies and model parameters. 
-    An important ingredient of BEAST2 is the species tree prior, 
-    due to which the tool can be used for the inference of 
-    macroevolutionary processes underlying the species tree. 
-    However, only models that allow for fast computation of 
-    the prior probability are possible in practice. 
-    This begs the question of how accurate the tree estimation is 
+    Phylogenetic trees are current routinely reconstructed from an alignment of character sequences (usually nucleotide sequences).
+Bayesian tools, such as MrBayes, RevBayes and BEAST2, have gained much popularity over the last decade, as they allow joint estimation of the posterior distribution of the phylogenetic tree and the parameters of the underlying inference model.
+    An important ingredient of these Bayesian approaches is the species tree prior.
+While in principle the Bayesian framework allows for comparing different species tree priors and hence may elucidate the macroevolutionary processes underlying the species tree, in practice only
+    macroevolutionary models that allow for fast computation of the prior probability are used. 
+    This begs the question how accurate the tree estimation is 
     when the real macroevolutionary processes are different 
-    from those assumed in BEAST2. \\
+    from those assumed in the tree prior. \\
   \textbf{2. }
@@
     a free, libre and open-source R package that assesses 
-    the inference error BEAST2 makes based on a 
-    phylogeny from any macroevolutionary diversification model. \\
+    the inference error made by Bayesian phylogenetics for a given macroevolutionary diversification model.
+\verb;pirouette; makes use of BEAST2, but its philosophy applies to any Bayesian phylogenetic inference tool\\
   \textbf{3. }
@@
 
-The development of new powerful inference tools, 
+The development of new powerful Bayesian phylogenetic inference tools, 
 such as BEAST [\cite{drummond2007beast}], 
@@
 or RevBayes [\cite{hohna2016revbayes}], 
-allows to build phylogenetic trees 
+has been a major advance in building phylogenetic trees 
 from character data (usually nucleotide sequences) extracted 
-from extant (but also extinct) organisms.
-This has constituted an important step forward 
-in our understanding of how species evolve.
-Such tools have been increasingly exploited to hypothesize 
-and test what the main drivers and modes for diversification are.
-
-BEAST [\cite{drummond2007beast}] is a Bayesian phylogenetics tool, 
-that needs both data and priors to infer a posterior.
-Of these priors, the tree prior describes how diversification occurs.
+from extant (but also extinct) organisms, and hence in our understanding of the main drivers and modes of diversification.
+
+BEAST [\cite{drummond2007beast}] is a typical Bayesian phylogenetics tool, 
+that needs both character data and priors to infer a posterior.
+Specifically, the species tree prior describes the process of diversification.
 BEAST has built-in tree priors such as the Yule [\cite{yule}] and 
@@
 functionalities.
-By writing a BEAST2 package, novel diversification models can be added.
-The key ingredient of such a package would be the likelihood equation
-of a phylogeny for the novel diversification model.
-
-\iffalse
-There have been plenty of these models (and their associated probability 
-algorithms) 
-developed, among others, those in which diversification is 
+Hence one can add novel diversification models by writing a BEAST2 package that contains the likelihood formula
+of a phylogeny under the novel diversification model, i.e. the prior probability of a species tree.
+
+Many such diversification models (and their associated probability 
+algorithms) have been developed
+developed, e.g., those in which diversification is 
 time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
 or diversity-dependent [\cite{etienne2011diversity}],
-or that can shift [\cite{etienne2012conceptual}, 
-\cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}].
-\fi
-
-Many of such models (and their associated probability 
-algorithms) have been
-developed to include diverse features: those in which diversification rates are 
-time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
-or diversity-dependent [\cite{etienne2011diversity}],
-or that can shift [\cite{etienne2012conceptual}, 
-\cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}].
-
-Other birth-death models treat speciation as a process that takes 
+or where diversification rates change for specific lineages and their descendants [\cite{etienne2012conceptual}, 
+\cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}], models that treat speciation as a process that takes 
 time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
-or that happens in bursts of simultaneous 
-events [\cite{laudanno2018mbd}], or that
+or as a burst of simultaneous branching
+events [\cite{laudanno2018mbd}], or where diversification rate
 depends on a trait that has two [\cite{maddison2007estimating}], 
 or more [\cite{fitzjohn2012diversitree}] states,
-or even allows concealed states [\cite{beaulieu2016detecting}] 
+even concealed states [\cite{beaulieu2016detecting}] 
 or a combination of all these [\cite{herrera2018detecting}].
@@
 When a novel diversification model is introduced,
-its performance is commonly tested as well.
-Part of a model's performance is its parameter 
-recovery (e.g. [\cite{etienne2014estimating}]),
-in which phylogenies are simulated from known parameters, 
-and then the model is used to infer back the parameters from the phylogenies.
+its performance in inference is commonly tested.
+Part of a model's performance is its ability to recover parameters from simulated data with known parameters (e.g. [\cite{etienne2014estimating}]),
 Ideally, the estimated parameter values closely match the known/true values.
@@
 it is not necessarily used in Bayesian inference.
-In these kind of procedures the likelihood of a diversification 
-model can be computed millions of times.
-Due to this, effective but computationally expensive tree priors are non-viable.
-Also, there may be no need to add such a tree prior to a Bayesian
-inference framework: the more informative the data, 
-the lower the influence the tree prior has.
-Due to this, computationally faster tree priors are more commonly used.
-However, it is unknown if it is valid to assume that tree prior choice is 
-of low importance.
+Bayesian phylogenetic inference often requires that the prior probability of the phylogeny according to the diversification model has to be computed millions of times.
+Therefore, biologically interesting but computationally expensive tree priors are non-viable.
+This is not necessarily problematic, when the data are very informative, as this will reduce the influence of the tree prior.
+However, the assumption that tree prior choice is of low importance must first be verified.
 
@@
 prior choice. For example, recently Sarver et al., [\cite{sarver2019choice}] 
-show that the choice of tree prior does not 
+showed that the choice of tree prior does not 
 substantially affect phylogenetic inferences of diversification rates.
@@
 recover a given tree when using the standard diversification models.
-These investigations show how current diversification models compare
-to one another, but they do not help to assess the importance of a 
-new tree prior.
+These studies show how current diversification models compare
+to one another, but they do not help to assess the importance of a new tree prior.
 
 Here we introduce a method to quantify the importance of a novel tree prior.
-The method starts with a phylogeny generated by the process of 
-a (usually novel) tree prior.
-Using the tree priors built-in into BEAST2, 
-a Bayesian posterior is inferred.
-The more similar the phylogenies in the posterior are to the original phylogeny,
-the less important the novel tree prior has been.
-Only impactful tree priors will be worth the effort and computational burden 
-to use in a Bayesian framework.
-
-This method is programmed as an R package called \verb;pirouette;.
+The method starts with a phylogeny generated by the new model for the tree prior.
+Next, nucleotide sequences are simulated on this phylogeny.
+Then, using the tree priors built-in into BEAST2, a Bayesian posterior distribution of phylogenies is inferred.
+We can then compare the inferred and simulated phylogenies.
+How to properly perform this comparison forms the heart of our method.
+Only new diversification models that result in a large discrepancy between inferred and simulated phylogeny will be worth the effort and computational burden 
+to implement a species tree prior for in a Bayesian framework.
+
+Our method is programmed as an R package called \verb;pirouette;.
 \verb;pirouette; is built on \verb;babette; [\cite{bilderbeek2018babette}], 
@@
 \verb;pirouette; is written in the R programming language (\cite{R}).
-The goal of \verb;pirouette; is to quantify the importance of a tree
-prior.
-It does so by measuring the inference error
-made for a given reconstructed phylogeny, 
-simulated under the mechanisms of a (usually novel) diversification model.
+The goal of \verb;pirouette; is to quantify the importance of a tree prior.
+It does so by measuring the inference error made for a given reconstructed phylogeny, 
+simulated under a (usually novel) diversification model.
 We refer to the true prior that generated the given tree
@@
 to a standard tree prior. For example, a protracted birth-death model
-becomes equivalent to a standard birth-death model when speciation-completion rate approaches infinity.
+reduces to a standard birth-death model when speciation-completion rate approaches infinity.
 When benchmarking a novel tree prior, 
-such a study will typically construct phylogenies 
+one will typically construct phylogenies 
 for different combinations of the diversification model's parameters, 
 to assess under which scenarios the inference error cannot be neglected. 
-Although a well-conducted theoretical study uses many replicates, 
-this article only shows examples with one replicate, for the sake of clarity.
+While we recommend many replicate simulations when assessing a novel tree prior, our examples contain only one replicate as they are for illustrative purposes only.
 
@@
 helps interpreting the resulting alignment.
-In the context of this research, a 'site model' is a nucleotide substitution
-model. Supported site models are JC, HKY, TN and GTR. Only the strict
+The nucleotide substitution
+model will be referred to as 'site model'. Supported site models are JC, HKY, TN and GTR. Only the strict
 clock model is supported in this step.
@@
 The second step (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a)
-selects inference models from a set of experiments. 
+selects an experiment from a set of experiments. 
 We define an experiment $X$ as the combination of an inference model $I$
@@
 If the attached inference model $I$ is the generative model, then it should always be selected.
-If, instead, the attached inference model is a candidate model the user can either let \verb;pirouette; select it as the best among all the candidate models, or force its selection (regardless
+If, instead, the attached inference model is a candidate model, the user can either let \verb;pirouette; select it as the best among all the candidate models, or force its selection (regardless
 of it being the best).
@@
 For each selected experiment a posterior is inferred, using the 
-\verb;babette; [\cite{bilderbeek2018babette}] R package. 
+\verb;babette; [\cite{bilderbeek2018babette}] R package which makes use of BEAST2. 
 This step usually takes up most of the full pipeline's computation time.
 
-The fourth step quantifies the inference error made, by removing
-the burn-in, then applying an error statistic.
-The burn-in is the first phase of the Markov chain Monte Carlo (MCMC) run,
-in which it samples an unrepresentative part of state space,
-and is commonly discarded. By default, \verb;pirouette; 
+The fourth step quantifies the inference error made. First the burn-in is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
+in which it samples an unrepresentative part of state space. By default, \verb;pirouette; 
 removes the first 10\% of the posterior.
@@
 that will be analyzed in the same way as the true tree.
-Any phylogeny $\tau$ is the combination of 
+A phylogeny $\tau$ is the combination of 
 branching times $\Vec{t}$ and topology $\psi$.
-Starting from a non-standard generative prior $t_p = \mathit{P}$,
-we can observe a given tree $\tau_{\mathit{P}}$,
+Assume that a non-standard generative prior $t_p = \mathit{P}$,
+produces a tree $\tau_{\mathit{P}}$,
 with branching times $\Vec{t}_{\mathit{P}}$ and 
-topology $\psi_{\mathit{P}}$, 
-but we cannot use the tree prior in Bayesian inference.
+topology $\psi_{\mathit{P}}$.
 The twinning process T, in accordance with a chosen inference 
@@
 
-We use $\theta^{*}_{\mathit{I}}$ to simulate an equal amount 
+We use $\theta^{*}_{\mathit{I}}$ to simulate an equal number 
 $n_{\mathit{I}} = n_{\mathit{P}}$ 
@@
 
-\verb;pirouette; can be installed easily from CRAN:
+\verb;pirouette; will be made available on CRAN from which it can then be easily installed easily from CRAN:
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]
@@
 
-For the most up-to-date version, 
+Until it is on CRAN, and for the most up-to-date version, 
 one can download and install the package from \verb;pirouette;'s GitHub 

### 2019-05-02 article.tex

@@
 We refer to the true prior that generated the given tree
-as the 'generative tree prior', $\mathit{P}$.
+as the 'generative tree prior' $\mathit{p_{G}}$.
 Many tree priors have a parameter setting in which these are equivalent
-to a standard tree prior. For example, a protracted birth-death model
-reduces to a standard birth-death model when speciation-completion rate approaches infinity.
+to a standard tree prior.
+For example, a protracted birth-death model reduces to a standard birth-death model when speciation-completion rate approaches infinity [i.e.
+rate $\lambda$ in \cite{etienne2014estimating}, eqs.
+(2b) and (2c)].
 When benchmarking a novel tree prior, 
@@
 \centering
-  \begin{tabular}{|@{}c|p{2.5cm}|p{9cm}|p{4.5cm}@{}|}
+  \begin{tabular}{|p{3.2cm}|p{9.7cm}|p{4.5cm}@{}|}
     \hline
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} 
-      & \textbf{Possible values} \\ 
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model 
-      & Yule [\cite{yule}], BD, CBS, CCP, CEP \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{c_{m}}$ & clock model & Clock for the DNA mutation rates 
-      & strict, relaxed log-normal \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{s_{m}}$ & site model & Nucleotide substitution model 
-      & JC, HKY, TN93, GTR \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{m_{r}}$ & mutation rate & Pace at which mutation occurs 
-      & $m_{r} \in \mathbb{R}_{>0}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{r_{s}}$ & root sequence & DNA sequence at the root of the tree 
-      & any combination of a, c, g, t \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{m_{p}}$ & MRCA prior & Prior knowledge of a most recent common 
-      ancestor & see documentation \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{m_{s}}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup 
+    \textbf{Sub-argument} & 
+    \textbf{Description} &
+    \textbf{Possible values} \\ 
+    \hline
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    tree\_prior &
+    Macroevolutionary diversification model &
+    Yule, BD, CBS, CCP, CEP \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    clock\_model &
+    Clock for the DNA mutation rates &
+    strict, relaxed log-normal \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    site\_model &
+    Nucleotide substitution model &
+    JC, HKY, TN93, GTR \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    mutation\_rate &
+    Pace at which mutation occurs &
+    mutation\_rate $\in \mathbb{R}_{>0}$\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    root\_sequence &
+    DNA sequence at the root of the tree &
+    any combination of a, c, g, t \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    model\_type &
+    Criterion to select an inference model &
+    Generative, Candidate \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    run\_if &
+    Condition under which an inference model is used &
+    Always, Best candidate \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    do\_measure\_evidence &
+    Sets whether or not the evidence of the model is to be computed &
+    TRUE, FALSE \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    error\_function &
+    Specifies how to measure the error &
+    nLTT, $|\gamma|$ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    burn\_in\_fraction &
+    Specifies the percentage of initial posterior trees to discard &
+    burn\_in\_fraction $\in [0, 1]$\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+    \iffalse
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{mc}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup 
       & see documentation \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{m_{t}}$ & model type & Criterion to select an inference model 
-      & Generative, Candidate \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{r_{i}}$ & run if & Condition under which an inference model is 
-      used & Always, Best candidate\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{e_{f}}$ & error function & Specifies how to measure the error 
-      & nLTT, $|\gamma|$ \\
+    $\mathit{mp}$ & MRCA prior & Prior knowledge of a most recent common ancestor & see documentation \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \fi
     \hline
@@
     Most important parameter options.
+    Yule = pure birth model (\cite{yule})
     BD = birth death (\cite{nee1994reconstructed}), 
@@
     HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
-    TN93 = Tamura and Nei \cite{tamura1993estimation}, 
+    TN93 = Tamura and Nei (\cite{tamura1993estimation}), 
     GTR = Generalized time-reversible model (\cite{tavare1986some}).
@@
 
-  \begin{tabular}{|@{}c|p{4cm}|p{9cm}|p{3cm}@{}|}
+  \begin{tabular}{|@{}c|p{4cm}|p{12cm}|}
     \hline
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{Symbol} & \textbf{Macro-argument} & \textbf{Description} 
-      & \textbf{Includes} \\
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{P}$ & Generative tree prior & Diversification model to generate 
-      the 'true' tree & $\mathit{t_{p}}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{A}$ & Alignment model & Model used to simulate the 
-      alignment ($\mathit{c_{m}}$ must be strict) 
-      & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{m_{r}}, \mathit{r_{s}}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{G}$ & Generative model & The full setting to produce BEAST2 input 
-      data & $\mathit{P}$, $\mathit{A}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{I}$ & Inference model & Phylogenetic inference model to run BEAST2 
-      & $\mathit{c_{m}}, \mathit{s_{m}}, \mathit{t_{p}}, 
-      \mathit{m_{p}}, \mathit{m_{s}}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ 
-      is used in the inference & $\mathit{m_{t}}, \mathit{r_{i}}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{X}$ & Experiment (candidate) 
-      & Full setting for a Bayesian inference 
-      & $\mathit{I},\mathit{C}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{E}$ & Error measure parameters & Errors measurement setup & 
-      $\mathit{e_{f}}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \textbf{Symbol} &
+    \textbf{Macro-argument} &
+    \textbf{Description} \\
+    \hline
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{G}$ &
+    Generative model &
+    The full setting to produce BEAST2 input data. Its core features are the tree\_prior $\mathit{p_{G}}$, the clock\_model $\mathit{c_{G}}$ and the site\_model $\mathit{s_{G}}$. \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
+    $\mathit{A}$ &
+    Alignment model &
+    It includes the parts of the generative model that directly affect the alignment generation, like the clock\_model $\mathit{c_{G}}$ and the site\_model $\mathit{s_{G}}$.
+Additional arguments can be provided, such as the mutation\_rate and the root\_sequence.
+\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{X_{i}}$ &
+    $i$-th candidate experiment &
+    Full setting for a Bayesian inference. It is made by a candidate inference model $\mathit{I_{i}}$ and its inference conditions $\mathit{C_{i}}$. \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{I}$ &
+    Inference model &
+    Phylogenetic inference model to run BEAST2.
+Likewise the generative model $G$, its main components are the tree\_prior $\mathit{p_{I}}$, the clock\_model $\mathit{c_{I}}$ and the site\_model $\mathit{s_{I}}$.
+\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference. They are composed by the model\_type, run\_if and do\_measure\_evidence sub-arguments.\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{E}$ & Error measure parameters & Errors measurement setup that can be specified providing an error\_function to measure the difference between the original phylogeny and the inferred posterior.
+The initial part of the posterior that is reckoned as not representative can be discarded using a burn\_in\_fraction.
+\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \iffalse
     $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & 
-      $\mathit{\tau_{p}}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+      $\mathit{\tau_{p}}$ \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \fi
     \hline 
@@
     Definitions of terms and relative symbols used in the main text and in 
-    Fig~\ref{fig:pipeline}. $Pr(D)$ denotes the evidence, $\omega$ denotes
-    a model's weight 
+    Fig~\ref{fig:pipeline}.
+To run the pipeline $\mathit{A}$, $\mathit{X}$ and $\mathit{E}$ must be specified.
+Examples can be found in listings~\ref{lst:create_alignment}, \ref{lst:create_gen_experiment_explicit} and \ref{lst:create_error_measure_params}.
   }
@@
   \centering
-  \includegraphics[width=\textwidth]{workflow_alt.png}
+  \includegraphics[width = \textwidth]{workflow3.png}
   \caption{
     \texttt{pirouette} pipeline.
-    The pipeline starts from a phylogeny (1a) simulated by generative prior 
-    $\mathit{P}$.
-    The phylogeny is converted to an alignment (2a) using alignment model 
-    $\mathit{A}$. 
+    The pipeline starts from a phylogeny (1a) simulated by the generative tree prior 
+    $\mathit{p_{G}}$.
+    The phylogeny is converted to an alignment (2a) using the generative alignment model 
+    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed by a clock and a site models. 
     The user supplies one or more experiments.
-    For each candidate experiment $\mathit{X}$ 
-    (a tuple of inference model $\mathit{I}$ and condition $\mathit{C}$),
-    if its condition $\mathit{C}$ is 
+    For each candidate experiment $\mathit{X_{i}}$ 
+    (a tuple of inference model $\mathit{I_{i}}$ and condition $\mathit{C_{i}}$),
+    if its condition $\mathit{C_{i}}$ is 
     satisfied (which can depend on the alignment), 
-    it becomes a selected experiment $\mathit{X'}$, 
-    to be used in the next step (3a).
-    The inference models of the selected experiment use the alignment (2a) 
+    the corresponding inference model $\mathit{I} = \mathit{I_{i}}$ is selected 
+    to be used in the next step.
+    The inference models (3) of the selected experiments use the alignment (2a) 
     to each create a Bayesian posterior of (parameter estimates and) 
@@
     resulting in an error distribution (5a). 
-    Optionally, a twin phylogeny (1b) can be generated from the original 
-    phylogeny (1a) using twinning parameters $T$, where $T$ may depend on
-    the experiments.
-    From a twin phylogeny, the same intermediate steps are taken,
-    resulting in a twin error distribution (5b)
+    Optionally, for each selected inference model a twin pipeline can be run.
+    A twin phylogeny (1b) can be generated from the original 
+    phylogeny (1a) using the tree prior $\mathit{p_{I}}$ of the selected inference model (3).
+    Analogously, a twin alignment is simulated from the twin phylogeny along with clock model $\mathit{c_{I}}$ and site model $\mathit{s_{I}}$ imported from the selected inference model.
+    The twin pipeline mimics the procedure of the main pipeline, resulting in a twin error distribution (5b).
   }
@@
     an inference model $\mathit{I}$ is picked (which may differ from the 
-    generative alignment model);
+    generative model);
   \item the inference model and the alignment are used 
@@
 using the DNA alignment parameters.
-The DNA alignment parameters consist of a DNA root sequence, a DNA mutation rate,
-a site model and a (strict) clock model.
-The DNA root sequence is the DNA sequence of the shared common ancestor,
+The DNA alignment parameters consist of a (DNA) root\_sequence, a (DNA) mutation\_rate, a clock\_model and a nucleotide substitution
+model, which will be referred to as site\_model.
+The root\_sequence is the DNA sequence of the shared common ancestor,
 and is set to four equally-sized mononucleotide blocks by default, as this
 helps interpreting the resulting alignment.
-The nucleotide substitution
-model will be referred to as 'site model'. Supported site models are JC, HKY, TN and GTR. Only the strict
-clock model is supported in this step.
-
-The second step (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a)
-selects an experiment from a set of experiments. 
-We define an experiment $X$ as the combination of an inference model $I$
-and the conditions $C$ to actually use it in the inference step.
-There are multiple possible setups for $C$. 
-If the attached inference model $I$ is the generative model, then it should always be selected.
-If, instead, the attached inference model is a candidate model, the user can either let \verb;pirouette; select it as the best among all the candidate models, or force its selection (regardless
-of it being the best).
-When $I$ is set to be the generative model,
-it will use the 
-same site and clock model used in the alignment simulation
-and the same tree prior used in the creation of the given phylogeny. 
+Supported site\_models are JC, HKY, TN and GTR. Only the strict
+clock\_model is supported in this step.
+
+The second step (Fig.~\ref{fig:pipeline}, 3)
+selects an inference model $I$ from a set of inference models $I_{1},\dots,I_{n}$. 
+We define an experiment $X_{i}$ as the combination of an inference model $I_{i}$ and the conditions $C_{i}$ to actually use it in the inference step.
+
+Any inference model is composed by a tree\_prior, a clock\_model and a site\_model. 
+The inference conditions are composed by two main elements called model\_type and run\_if.
+They determine how to choose the basic elements composing the inference model that is going to be used in the following of the \verb;pirouette;'s pipeline.
+The simplest setup is to use the same site\_model and clock\_model used in the alignment simulation and the same tree\_prior used in the creation of the given phylogeny.
+We label this model\_type as "generative".
+In any other case we are dealing with a "candidate" model\_type.
+The argument run\_if determines whether the proposed model has to be used or not.
+If the model\_type is generative it will always be used.
+If, instead, one or more candidate models are proposed, the user can either let \verb;pirouette; select it as the best among all the candidate models (run\_if == "best\_candidate"), or force its selection (run\_if == "always").
 A 'best' model is defined as the inference model with
-the highest evidence (a.k.a. marginal likelihood) for that inference model, 
-given the alignment simulated in the previous step.
-The evidence of an inference model is estimated by nested 
-sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package.
-Scripted use of BEAST2 packages is only possible under under Linux and Mac.
-Windows systems can do the model comparison for shorter DNA sequences
-using the web interface of \verb;mcbette; [\cite{mcbette}].
+the highest evidence (a.k.a. marginal likelihood), given the alignment simulated in the previous step.
+The evidence of an inference model is estimated by nested sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. Scripted use of BEAST2 packages is only possible under Linux and Mac.
+Windows systems can do the model comparison for shorter DNA sequences using the web interface of \verb;mcbette; [\cite{mcbette}].
+\iffalse
+Even if the model selection is not performed, the user can still estimate the evidence of each proposed inference model by setting "do\_measure\_evidence" to be TRUE.
+\fi
 
 The third step infers the posteriors,
-using the simulated alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a $\rightarrow$ 4a),
-and those inference models that were selected in the previous step. 
+using the simulated alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 4a),
+and those inference models that were selected in the previous step (3). 
 For each selected experiment a posterior is inferred, using the 
@@
 
-The fourth step quantifies the inference error made. First the burn-in is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
+The fourth step quantifies the inference error made. First the burn\_in\_fraction is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
 in which it samples an unrepresentative part of state space. By default, \verb;pirouette; 
@@
 branching times $\Vec{t}$ and topology $\psi$.
-Assume that a non-standard generative prior $t_p = \mathit{P}$,
-produces a tree $\tau_{\mathit{P}}$,
-with branching times $\Vec{t}_{\mathit{P}}$ and 
-topology $\psi_{\mathit{P}}$.
+Assume that a non-standard generative prior $\mathit{p_{G}}$
+produces a tree $\tau_{\mathit{G}}$,
+with branching times $\Vec{t}_{\mathit{G}}$ and 
+topology $\psi_{\mathit{G}}$.
 The twinning process T, in accordance with a chosen inference 
-model $t_p \in \mathit{I}$, creates a tree $\tau_{\mathit{I}}$
+model $\mathit{I}$, creates a tree $\tau_{\mathit{I}}$
 with branching times $\Vec{t}_{\mathit{I}}$ while preserving the 
-topology $\psi_{\mathit{P}}$:
+topology $\psi_{\mathit{G}}$:
 
 \begin{align}
-  \tau_{\mathit{P}} = (\Vec{t}_{\mathit{P}}, \psi_{\mathit{P}}) 
+  \tau_{\mathit{G}} = (\Vec{t}_{\mathit{G}}, \psi_{\mathit{G}}) 
   \xrightarrow[]{\mathit{T}} 
-  \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, \psi_{\mathit{P}})
+  \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, \psi_{\mathit{G}})
 \end{align}
@@
 that maximize the likelihood $L_{\mathit{I}}$ applied 
-to the true tree, conditioned on its number of tips $n_{\mathit{P}}$,
+to the true tree, conditioned on its number of tips $n_{\mathit{G}}$,
 where $L_{\mathit{I}}$ is the likelihood function  
@@
 \begin{align}
-    max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{P}}, n_{\mathit{P}})] 
+    max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{G}}, n_{\mathit{G}})] 
 \rightarrow \theta^{*}_{\mathit{I}}.
@@
 We use $\theta^{*}_{\mathit{I}}$ to simulate an equal number 
-$n_{\mathit{I}} = n_{\mathit{P}}$ 
+$n_{\mathit{I}} = n_{\mathit{G}}$ 
 of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree 
@@
 
-Using a twin tree serves as a control: both $\tau_{\mathit{P}}$ and 
+Using a twin tree serves as a control: both $\tau_{\mathit{G}}$ and 
 $\tau_{\mathit{I}}$ will produce an error distribution. 
 These errors unavoidably accumulate across the \verb;pirouette; pipeline,
-in both the true and twin tree: either tree may be an unlikely representative
+for both the true and twin tree, due to several possible reasons: either tree may be an unlikely representative
 of the underlying tree prior, the alignment simulation may give different
@@
 consequences of different phylogenies and tree priors for the two parallel 
-pipelines: the main pipeline uses the tree $\tau_{\mathit{P}}$ and a
-possibly mismatched prior $\mathit{P}$, where the twin pipeline
+pipelines: the main pipeline uses the tree $\tau_{\mathit{G}}$ and a
+possibly mismatched prior $\mathit{G}$, where the twin pipeline
 uses the tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
@@
 
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+\begin{lstlisting}[
+    language = R,
+    floatplacement = ht,
+    frame = single
+]
 usethis::install_github("richelbilderbeek/pirouette")
@@
 phylogenies in a posterior. Due to stochasticity, all posterior trees
-will differ from the given phylogeny $\tau_{\mathit{P}}$, 
-even when the generative tree prior $\mathit{P}$ 
+will differ from the given phylogeny $\tau_{\mathit{G}}$, 
+even when the generative tree prior $\mathit{p_{G}}$ 
 and alignment model $\mathit{A}$ are the same as 
@@
 
-In this example we use a standard generative tree prior $\mathit{P_{0}}$, 
+In this example we use a standard generative tree prior $\mathit{p_{G}^0}$, 
 namely the Yule (pure-birth) tree prior. 
@@
 \begin{lstlisting}[
-    language=R, floatplacement=ht, frame=single, 
+    language = R,
+    floatplacement = ht,
+    frame = single, 
     label = {lst:create_yule_tree}, 
@@
 given phylogeny, as described in subsection~\ref{subsec:pipeline}.
-In this example, the DNA root sequence $\mathit{r_{s}}$ 
-consists in four blocks of 250 
+In this example, the root\_sequence consists in four blocks of 250 
 mononucleotides each, 
-where the per-nucleotide mutation rate $\mathit{m_{r}}$ 
-is 0.1 mutations per unit time.
-We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
-and a strict clock model as these are the simplest.
-A JC site model
-assumes that mutation rates between nucleotides are equal and constant. 
-A strict clock model assumes that the mutation rates 
+where the per-nucleotide mutation\_rate is 0.1 mutations per unit time.
+We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site\_model
+and a strict clock\_model as these are the simplest.
+A JC site\_model assumes that mutation rates between nucleotides are equal and constant. 
+A strict clock\_model assumes that the mutation rates 
 of all species are equal and constant.
@@
 \begin{lstlisting}[
-    language=R,
-    floatplacement=ht, frame=single,
+    language = R,
+    floatplacement = ht,
+    frame = single,
     label = {lst:create_alignment}, 
@@
 alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
+  clock_model = create_strict_clock_model(),
+  site_model = create_jc69_site_model(),
   mutation_rate = 0.1,
-  site_model = create_jc69_site_model(),
-  clock_model = create_strict_clock_model()
+  root_sequence = create_blocked_dna(length = 1000)
 )
@@
 generative one,
-which is the chosen Yule tree prior $\mathit{P_{0}}$, as well as site and clock 
+which is the chosen Yule tree prior $\mathit{p_{G}^0}$, as well as site and clock 
 models defined in $\mathit{A}$, respectively JC and strict.
@@
     \hline
-    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & 
+    \textbf{model\_type} &
+    \textbf{run\_if} &
+    \textbf{do\_measure\_evidence} & 
     \textbf{inference model} \\ 
     \hline
-    generative & always & no & JC, strict, Yule \\
+    generative &
+    always &
+    FALSE &
+    JC, strict, Yule \\
     \hline
@@
     Inference conditions and inference model for ERQ 1.
-    JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
+    JC: Jukes-Cantor site\_model.
+    strict: strict clock\_model.
+    Yule: Yule (pure-birth) tree\_prior.
   }
@@
 \begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:create_gen_experiment_explicit},
@@
   inference_model = create_inference_model(
-   site_model = create_jc69_site_model(), 
+    tree_prior = create_yule_tree_prior(),
     clock_model = create_strict_clock_model(), 
-   tree_prior = create_yule_tree_prior()
+    site_model = create_jc69_site_model()
   )
@@
 \begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, frame=single,
+  language = R, 
+  floatplacement = ht,
+  frame = single,
   label = {lst:create_gen_experiment},
@@
 twinning parameters $\mathit{T}$.
+\giovanni{I suggest to refer to the twinning parameters only when strictly required and with the minimum possible amount of info provided, as we set it in such a way that the tree prior is imported from the inference model in an automatic way.}
 
 We use the default error measurement
-parameters, which uses a burn-in of 10\% and the nLTT statistic to
+parameters, which uses a burn\_in\_fraction of 10\% and the nLTT statistic to
 measure the difference between phylogenies. For clarity,
@@
 \begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:create_error_measure_params},
@@
 error_measure_params <- create_error_measure_params(
-  burn_in_fraction = 0.1, 
-  error_function = get_nltt_error_function()
+  error_function = get_nltt_error_function(),
+  burn_in_fraction = 0.1
 )
@@
 \begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:create_pir_params},
@@
 \begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:pir_run},
@@
 \begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:pir_plot},
@@
 
-The resulting error distribution, as shown figure~\ref{fig:example_1}),
+The resulting error distribution, as shown figure~\ref{fig:example_1},
 shows the inference error 
 when $\mathit{I}$ matches with the generative model given 
-by $\mathit{A}$ and $\mathit{P_{0}}$.
+by $\mathit{A}$ and $\mathit{p_{G}^0}$.
 This error distribution can serve as a control,
@@
 In the previous example we selected the inference model $\mathit{I}$ to match 
-with a known generative tree prior $\mathit{P_{0}}$.
-However, a novel tree prior $\mathit{P}$, by our definition, is non-standard, 
+with a known generative tree prior $\mathit{p_{G}^0}$.
+However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, 
 thus cannot be part of a standard inference 
@@
 In such a case, the user has to somehow select 
-a standard tree prior $\mathit{t_{p}}$ to be 
+a standard tree prior to be 
 used in the inference.
@@
 \begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, 
-  frame=single, 
+  language = R, 
+  floatplacement = ht,
+  frame = single, 
   label = {lst:unknown_phylogeny},
@@
 
-First step is to simulate a DNA alignment from the 
+The first step is to simulate a DNA alignment from the 
 given phylogeny. 
@@
     \hline
-    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & 
+    \textbf{model\_type} &
+    \textbf{run\_if} & \textbf{do\_measure\_evidence} & 
     \textbf{inference model} \\ 
     \hline
-    candidate & best candidate & yes & JC, strict, Yule \\
-    candidate & best candidate & yes & JC, strict, BD   \\
+    candidate & best candidate & TRUE & JC, strict, Yule \\
+    candidate & best candidate & TRUE & JC, strict, BD   \\
     ...       & ...            & ... & ...              \\
-    candidate & best candidate & yes & GTR, RLN, CCP    \\
-    candidate & best candidate & yes & GTR, RLN, CEP    \\
+    candidate & best candidate & TRUE & GTR, RLN, CCP    \\
+    candidate & best candidate & TRUE & GTR, RLN, CEP    \\
     \hline
@@
     Inference conditions and inference model for ERQ 2.
-    JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-    BD: birth-death tree prior.
-    GTR: GTR site model.
-    RLN: relaxed log-normal clock model.
-    CCP: coalescent constant-population tree prior.
-    CEP: coalescent exponential-population tree prior.
+    JC: Jukes-Cantor site\_model.
+    strict: strict clock\_model.
+    Yule: Yule (pure-birth) tree\_prior.
+    BD: birth-death tree\_prior.
+    GTR: GTR site\_model.
+    RLN: relaxed log-normal clock\_model.
+    CCP: coalescent constant-population tree\_prior.
+    CEP: coalescent exponential-population tree\_prior.
   }
@@
 
-In the second step we set up our experiments. Unlike ERQ 1, here we provide a 
+In the second step, unlike ERQ 1, we provide a 
 set of experiments $\{\mathit{X_{i}}\}_{i=1}^{40}$. All the experiments have 
@@
 
-As model selection is commonly performed on the full list of available candidate models,
-\verb;pirouette; has a dedicated function to provide it:
-\verb;create_all_experiments; will create the entire set of 40 experiments,
-containing the inference models of all combinations of 4 site 
-models, 2 clock models and 5 tree priors.
+As model selection is commonly performed on the full list of available candidate models, \verb;pirouette; has a dedicated function to provide it:
+\verb;create_all_experiments; will create a full set of 40 experiments,
+containing the inference models of all combinations of 4 site\_models, 2 clock\_models and 5 tree\_priors.
 
 \begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, 
-  frame=single, 
+  language = R, 
+  floatplacement = ht, 
+  frame = single, 
   label = {lst:create_all_experiments},
@@
 \begin{lstlisting}[
-  language=R, floatplacement=ht, frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:create_pir_params_defaults},
@@
 by using a twin tree (see subsection~\ref{subsec:twinning}).
-In this ERQ, we hand-pick a standard tree prior $\mathit{P}$.
+In this ERQ, we hand-pick a standard tree prior $\mathit{p_{G}}$.
 
@@
 \begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, 
-  frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:create_gen_experiment_defaults},
@@
 \begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, 
-  frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:create_experiments},
@@
 \begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, 
-  frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:create_twinning_params},
@@
 \end{lstlisting}
+\giovanni{We want to import the "twin\_model" from the inference model.}
 
@@
 \begin{lstlisting}[
-  language=R, 
-  floatplacement=ht, 
-  frame=single,
+  language = R,
+  floatplacement = ht,
+  frame = single,
   label = {lst:create_pir_params_with_twinning},
@@
 as shown in ERQ2.
-The obtained error distribution contains the error accumulated in the pipeline,
+The resulting error distribution contains the error accumulated in the pipeline,
 without quantifying to what extent this error is due to the mismatch between 
-the generative tree prior $\mathit{P}$ and the selected inference model 
+the generative model $\mathit{G}$ and the selected inference model 
 $\mathit{I}$.
-As twinning (see section \ref{subsec:twinning}: 'Twinning') shows us the baseline error,
-we can quantify the contribution of the tree prior to the error.
+As twinning (see section \ref{subsec:twinning}: 'Twinning') shows us the baseline error, we can quantify the contribution of the tree prior to the error.
 
@@
     \hline
-    \textbf{model type} & \textbf{run if} & \textbf{measure evidence} & 
+    \textbf{model\_type} & 
+    \textbf{run\_if} & 
+    \textbf{do\_measure\_evidence} & 
     \textbf{inference model} \\ 
     \hline
-    generative & always         & yes & JC, strict, Yule \\
-    candidate  & best candidate & yes & JC, strict, BD \\
-    candidate  & best candidate & yes & JC, strict, CCP \\
-    candidate  & best candidate & yes & JC, strict, CEP \\
+    generative & always         & TRUE & JC, strict, Yule \\
+    candidate  & best candidate & TRUE & JC, strict, BD \\
+    candidate  & best candidate & TRUE & JC, strict, CCP \\
+    candidate  & best candidate & TRUE & JC, strict, CEP \\
     \hline
@@
     JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-    BD: birth-death tree prior.
-    CCP: coalescent constant-population tree prior.
-    CEP: coalescent exponential-population tree prior.
+    strict: strict clock\_model.
+    Yule: Yule (pure-birth) tree\_prior.
+    BD: birth-death tree\_prior.
+    CCP: coalescent constant-population tree\_prior.
+    CEP: coalescent exponential-population tree\_prior.
   }
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[
+    language = R,
+    floatplacement = H,
+    frame = single
+]
 candidate_experiments <- create_all_experiments(
@@
 
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
+\begin{lstlisting}[
+    language = R,
+    floatplacement = H,
+    frame = single
+]
 experiments <- c(list(generative_experiment), candidate_experiments)
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% RJCB will add: in six percent did the twin do worse, 
+% see https://github.com/richelbilderbeek/pirouette_article/issues/41
 

### 2019-05-02 article.tex
Fix typo

@@
     The phylogeny is converted to an alignment (2a) using the generative alignment model 
-    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed by a clock and a site models. 
+    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed by a clock and a site model. 
     The user supplies one or more experiments.

### 2019-05-02 article.tex
Revised paragraph about experiment selection

@@
 The second step (Fig.~\ref{fig:pipeline}, 3)
-selects an inference model $I$ from a set of inference models $I_{1},\dots,I_{n}$. 
-We define an experiment $X_{i}$ as the combination of an inference model $I_{i}$ and the conditions $C_{i}$ to actually use it in the inference step.
-
-Any inference model is composed by a tree\_prior, a clock\_model and a site\_model. 
-The inference conditions are composed by two main elements called model\_type and run\_if.
-They determine how to choose the basic elements composing the inference model that is going to be used in the following of the \verb;pirouette;'s pipeline.
-The simplest setup is to use the same site\_model and clock\_model used in the alignment simulation and the same tree\_prior used in the creation of the given phylogeny.
-We label this model\_type as "generative".
-In any other case we are dealing with a "candidate" model\_type.
-The argument run\_if determines whether the proposed model has to be used or not.
-If the model\_type is generative it will always be used.
-If, instead, one or more candidate models are proposed, the user can either let \verb;pirouette; select it as the best among all the candidate models (run\_if == "best\_candidate"), or force its selection (run\_if == "always").
+selects one or more inference models $I$ from a set of inference 
+models $I_{1},\dots,I_{n}$. 
+We define an experiment $X_{i}$ as the combination of 
+an inference model $I_{i}$ and the conditions $C_{i}$ 
+to actually use it in the inference step.
+For example, we can express that an inference
+model (a combination of a tree prior, clock model and site model) 
+is known to include the generative/true tree prior and that 
+we definitely want to use that model in our inference. 
+As a second example, wwe can also express that we've submitted a set of 
+candidate inference models,
+of which only the best should be used in the actual inference.
+In the first example, we specified the condition $C_{i}$ that this
+generative model should always be run, where in the second example,
+we specified condition $C_{i}$ that a candidate model should only be run
+when being the best.
 A 'best' model is defined as the inference model with
-the highest evidence (a.k.a. marginal likelihood), given the alignment simulated in the previous step.
-The evidence of an inference model is estimated by nested sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. Scripted use of BEAST2 packages is only possible under Linux and Mac.
-Windows systems can do the model comparison for shorter DNA sequences using the web interface of \verb;mcbette; [\cite{mcbette}].
-\iffalse
-Even if the model selection is not performed, the user can still estimate the evidence of each proposed inference model by setting "do\_measure\_evidence" to be TRUE.
-\fi
+the highest evidence (a.k.a. marginal likelihood), given the alignment 
+simulated in the previous step.
+The evidence of an inference model is estimated by nested 
+sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. 
+Scripted use of BEAST2 packages is only possible under Linux and Mac.
+Windows systems can do the model comparison for shorter DNA sequences 
+using the web interface of \verb;mcbette; [\cite{mcbette}].
 

### 2019-05-02 article.tex
Use no underscores in spoken text, do use monospace font when using underscores in code

@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    tree\_prior &
+    \verb;tree_prior; &
     Macroevolutionary diversification model &
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    clock\_model &
+    \verb;clock_model; &
     Clock for the DNA mutation rates &
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    site\_model &
+    \verb;site_model; &
     Nucleotide substitution model &
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    mutation\_rate &
+    \verb;mutation_rate; &
     Pace at which mutation occurs &
-    mutation\_rate $\in \mathbb{R}_{>0}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    root\_sequence &
+    mutation rate $\in \mathbb{R}_{>0}$\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \verb;root_sequence; &
     DNA sequence at the root of the tree &
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    model\_type &
+    \verb;model_type; &
     Criterion to select an inference model &
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    run\_if &
+    \verb;run_if; &
     Condition under which an inference model is used &
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    do\_measure\_evidence &
+    \verb;do_measure_evidence; &
     Sets whether or not the evidence of the model is to be computed &
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    error\_function &
+    \verb;error_function; &
     Specifies how to measure the error &
@@
     Specifies the percentage of initial posterior trees to discard &
-    burn\_in\_fraction $\in [0, 1]$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-    \iffalse
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{mc}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup 
-      & see documentation \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{mp}$ & MRCA prior & Prior knowledge of a most recent common ancestor & see documentation \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{\tau_{p}}$ & twin tree prior 
-      & Sets the tree prior for the twin tree 
-      & Yule, BD \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \fi
+    \verb;burn_in_fraction; $\in [0, 1]$\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \hline
@@
     Generative model &
-    The full setting to produce BEAST2 input data. Its core features are the tree\_prior $\mathit{p_{G}}$, the clock\_model $\mathit{c_{G}}$ and the site\_model $\mathit{s_{G}}$. \\
+    The full setting to produce BEAST2 input data. 
+    Its core features are the tree prior $\mathit{p_{G}}$, the clock 
+    model $\mathit{c_{G}}$ and the site model $\mathit{s_{G}}$. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
@@
     Alignment model &
-    It includes the parts of the generative model that directly affect the alignment generation, like the clock\_model $\mathit{c_{G}}$ and the site\_model $\mathit{s_{G}}$.
-Additional arguments can be provided, such as the mutation\_rate and the root\_sequence.
-\\
+    It includes the parts of the generative model that directly affect the 
+    alignment generation, like the clock model $\mathit{c_{G}}$ and 
+    the site model $\mathit{s_{G}}$. Additional arguments can be provided, 
+    such as the mutation rate and the root sequence. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     $i$-th candidate experiment &
-    Full setting for a Bayesian inference. It is made by a candidate inference model $\mathit{I_{i}}$ and its inference conditions $\mathit{C_{i}}$. \\
+    Full setting for a Bayesian inference. It is made by a 
+    candidate inference model $\mathit{I_{i}}$ and its 
+    inference conditions $\mathit{C_{i}}$. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Inference model &
-    Phylogenetic inference model to run BEAST2.
-Likewise the generative model $G$, its main components are the tree\_prior $\mathit{p_{I}}$, the clock\_model $\mathit{c_{I}}$ and the site\_model $\mathit{s_{I}}$.
-\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ is used in the inference. They are composed by the model\_type, run\_if and do\_measure\_evidence sub-arguments.\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{E}$ & Error measure parameters & Errors measurement setup that can be specified providing an error\_function to measure the difference between the original phylogeny and the inferred posterior.
-The initial part of the posterior that is reckoned as not representative can be discarded using a burn\_in\_fraction.
-\\
+    Phylogenetic inference model to run BEAST2. Likewise the generative model $G$, 
+    its main components are the tree prior $\mathit{p_{I}}$, the 
+    clock model $\mathit{c_{I}}$ and the site model $\mathit{s_{I}}$. \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ 
+    is used in the inference. 
+    They are composed by the model type, run condition and 
+    wether to measure the evidence. \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    $\mathit{E}$ & Error measure parameters & 
+    Errors measurement setup that can be specified providing an 
+    error function to measure the difference between the original phylogeny 
+    and the inferred posterior. The initial part of the posterior that is 
+    reckoned as not representative can be discarded using a burn-in fraction. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Definitions of terms and relative symbols used in the main text and in 
-    Fig~\ref{fig:pipeline}.
-To run the pipeline $\mathit{A}$, $\mathit{X}$ and $\mathit{E}$ must be specified.
-Examples can be found in listings~\ref{lst:create_alignment}, \ref{lst:create_gen_experiment_explicit} and \ref{lst:create_error_measure_params}.
+    Fig~\ref{fig:pipeline}. To run the pipeline $\mathit{A}$, $\mathit{X}$ 
+    and $\mathit{E}$ must be specified. Examples can be found in 
+    listings~\ref{lst:create_alignment}, 
+    \ref{lst:create_gen_experiment_explicit} and 
+    \ref{lst:create_error_measure_params}.
   }
@@
 using the DNA alignment parameters.
-The DNA alignment parameters consist of a (DNA) root\_sequence, a (DNA) mutation\_rate, a clock\_model and a nucleotide substitution
-model, which will be referred to as site\_model.
-The root\_sequence is the DNA sequence of the shared common ancestor,
+The DNA alignment parameters consist of a (DNA) root sequence, 
+a (DNA) mutation rate, a clock model and a nucleotide substitution
+model, which will be referred to as site model.
+The root sequence is the DNA sequence of the shared common ancestor,
 and is set to four equally-sized mononucleotide blocks by default, as this
 helps interpreting the resulting alignment.
-Supported site\_models are JC, HKY, TN and GTR. Only the strict
-clock\_model is supported in this step.
+Supported site models are JC, HKY, TN and GTR. Only the strict
+clock model is supported in this step.
 
@@
 
-The fourth step quantifies the inference error made. First the burn\_in\_fraction is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
+The fourth step quantifies the inference error made. First the 
+burn-in fraction is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
 in which it samples an unrepresentative part of state space. By default, \verb;pirouette; 
@@
 given phylogeny, as described in subsection~\ref{subsec:pipeline}.
-In this example, the root\_sequence consists in four blocks of 250 
+In this example, the root sequence consists in four blocks of 250 
 mononucleotides each, 
-where the per-nucleotide mutation\_rate is 0.1 mutations per unit time.
-We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site\_model
-and a strict clock\_model as these are the simplest.
-A JC site\_model assumes that mutation rates between nucleotides are equal and constant. 
-A strict clock\_model assumes that the mutation rates 
+where the per-nucleotide mutation rate is 0.1 mutations per unit time.
+We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
+and a strict clock model as these are the simplest.
+A JC site model assumes that mutation rates between nucleotides are equal and constant. 
+A strict clock model assumes that the mutation rates 
 of all species are equal and constant.
@@
     Inference conditions and inference model for ERQ 1.
-    JC: Jukes-Cantor site\_model.
-    strict: strict clock\_model.
-    Yule: Yule (pure-birth) tree\_prior.
+    JC: Jukes-Cantor site model.
+    strict: strict clock model.
+    Yule: Yule (pure-birth) tree prior.
   }
@@
 We use the default error measurement
-parameters, which uses a burn\_in\_fraction of 10\% and the nLTT statistic to
+parameters, which uses a burn-in fraction of 10\% and the nLTT statistic to
 measure the difference between phylogenies. For clarity,
@@
     Inference conditions and inference model for ERQ 2.
-    JC: Jukes-Cantor site\_model.
-    strict: strict clock\_model.
-    Yule: Yule (pure-birth) tree\_prior.
-    BD: birth-death tree\_prior.
-    GTR: GTR site\_model.
-    RLN: relaxed log-normal clock\_model.
-    CCP: coalescent constant-population tree\_prior.
-    CEP: coalescent exponential-population tree\_prior.
+    JC: Jukes-Cantor site model.
+    strict: strict clock model.
+    Yule: Yule (pure-birth) tree prior.
+    BD: birth-death tree prior.
+    GTR: GTR site model.
+    RLN: relaxed log-normal clock model.
+    CCP: coalescent constant-population tree prior.
+    CEP: coalescent exponential-population tree prior.
   }
@@
 
-As model selection is commonly performed on the full list of available candidate models, \verb;pirouette; has a dedicated function to provide it:
+As model selection is commonly performed on the full list of available 
+candidate models, \verb;pirouette; has a dedicated function to provide it:
 \verb;create_all_experiments; will create a full set of 40 experiments,
-containing the inference models of all combinations of 4 site\_models, 2 clock\_models and 5 tree\_priors.
+containing the inference models of all combinations of 
+4 site models, 2 clock models and 5 tree priors.
 
@@
 \giovanni{We want to import the "twin\_model" from the inference model.}
+\richel{Yes, that is an option, but according to Rampal not important enough
+to add now}
 
@@
     JC: Jukes-Cantor site model.
-    strict: strict clock\_model.
-    Yule: Yule (pure-birth) tree\_prior.
-    BD: birth-death tree\_prior.
-    CCP: coalescent constant-population tree\_prior.
-    CEP: coalescent exponential-population tree\_prior.
+    strict: strict clock model.
+    Yule: Yule (pure-birth) tree prior.
+    BD: birth-death tree prior.
+    CCP: coalescent constant-population tree prior.
+    CEP: coalescent exponential-population tree prior.
   }

### 2019-05-06 article.tex
Changes @Giappo

@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    burn\_in\_fraction &
+    \verb;burn\_in\_fraction; &
     Specifies the percentage of initial posterior trees to discard &
@@
 we definitely want to use that model in our inference. 
-As a second example, wwe can also express that we've submitted a set of 
+As a second example, we can also express that we've submitted a set of 
 candidate inference models,

### 2019-05-07 article.tex
Split up in article.tex (big picture), content.tex (the content) and (supplement.tex) supplementary materials

--- 2019-05-02 article.tex
@@
 
-\begin{abstract}
+\include{content}
 
-  \textbf{1. }
-    Phylogenetic trees are current routinely reconstructed from an alignment of character sequences (usually nucleotide sequences).
-Bayesian tools, such as MrBayes, RevBayes and BEAST2, have gained much popularity over the last decade, as they allow joint estimation of the posterior distribution of the phylogenetic tree and the parameters of the underlying inference model.
-    An important ingredient of these Bayesian approaches is the species tree prior.
-While in principle the Bayesian framework allows for comparing different species tree priors and hence may elucidate the macroevolutionary processes underlying the species tree, in practice only
-    macroevolutionary models that allow for fast computation of the prior probability are used. 
-    This begs the question how accurate the tree estimation is 
-    when the real macroevolutionary processes are different 
-    from those assumed in the tree prior. \\
-  \textbf{2. }
-    Here we present \verb;pirouette;, 
-    a free, libre and open-source R package that assesses 
-    the inference error made by Bayesian phylogenetics for a given macroevolutionary diversification model.
-\verb;pirouette; makes use of BEAST2, but its philosophy applies to any Bayesian phylogenetic inference tool\\
-  \textbf{3. }
-    We describe \verb;pirouette;'s usage and the biological scientific
-    question it can answer, including full examples. \\
-  \textbf{4. }
-    Last, we discuss the results obtained by the examples and their interpretation. \\
-\end{abstract}
+\include{supplement}
 
-{\bf Keywords:} computational biology, evolution, phylogenetics, tree prior,
-  Bayesian model selection, BEAST2, babette, pirouette, R
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The development of new powerful Bayesian phylogenetic inference tools, 
-such as BEAST [\cite{drummond2007beast}], 
-MrBayes [\cite{huelsenbeck2001mrbayes}]
-or RevBayes [\cite{hohna2016revbayes}], 
-has been a major advance in building phylogenetic trees 
-from character data (usually nucleotide sequences) extracted 
-from extant (but also extinct) organisms, and hence in our understanding of the main drivers and modes of diversification.
-
-BEAST [\cite{drummond2007beast}] is a typical Bayesian phylogenetics tool, 
-that needs both character data and priors to infer a posterior.
-Specifically, the species tree prior describes the process of diversification.
-BEAST has built-in tree priors such as the Yule [\cite{yule}] and 
-(constant-rate) birth-death [\cite{nee1994reconstructed}] models.
-These simple tree priors are most commonly used, as these 
-have sufficient biological complexity, while being computationally fast.
-BEAST's successor, BEAST2 [\cite{bouckaert2014beast}],
-has a package manager, that allows third-party users to extend existing 
-functionalities.
-Hence one can add novel diversification models by writing a BEAST2 package that contains the likelihood formula
-of a phylogeny under the novel diversification model, i.e. the prior probability of a species tree.
-
-Many such diversification models (and their associated probability 
-algorithms) have been developed
-developed, e.g., those in which diversification is 
-time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
-or diversity-dependent [\cite{etienne2011diversity}],
-or where diversification rates change for specific lineages and their descendants [\cite{etienne2012conceptual}, 
-\cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}], models that treat speciation as a process that takes 
-time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
-or as a burst of simultaneous branching
-events [\cite{laudanno2018mbd}], or where diversification rate
-depends on a trait that has two [\cite{maddison2007estimating}], 
-or more [\cite{fitzjohn2012diversitree}] states,
-even concealed states [\cite{beaulieu2016detecting}] 
-or a combination of all these [\cite{herrera2018detecting}].
-Only few of these diversification models are available as a BEAST2 package.
-
-When a novel diversification model is introduced,
-its performance in inference is commonly tested.
-Part of a model's performance is its ability to recover parameters from simulated data with known parameters (e.g. [\cite{etienne2014estimating}]),
-Ideally, the estimated parameter values closely match the known/true values.
-
-Even when a diversification model passes the procedure described above,
-it is not necessarily used in Bayesian inference.
-Bayesian phylogenetic inference often requires that the prior probability of the phylogeny according to the diversification model has to be computed millions of times.
-Therefore, biologically interesting but computationally expensive tree priors are non-viable.
-This is not necessarily problematic, when the data are very informative, as this will reduce the influence of the tree prior.
-However, the assumption that tree prior choice is of low importance must first be verified.
-
-There have been multiple attempts to investigate the importance of tree
-prior choice. For example, recently Sarver et al., [\cite{sarver2019choice}] 
-showed that the choice of tree prior does not 
-substantially affect phylogenetic inferences of diversification rates.
-Also recently, Duchene et al. [\cite{duchene2018phylodynamic}] released
-a BEAST2 package to assess how well posterior predictive simulations
-recover a given tree when using the standard diversification models.
-These studies show how current diversification models compare
-to one another, but they do not help to assess the importance of a new tree prior.
-
-Here we introduce a method to quantify the importance of a novel tree prior.
-The method starts with a phylogeny generated by the new model for the tree prior.
-Next, nucleotide sequences are simulated on this phylogeny.
-Then, using the tree priors built-in into BEAST2, a Bayesian posterior distribution of phylogenies is inferred.
-We can then compare the inferred and simulated phylogenies.
-How to properly perform this comparison forms the heart of our method.
-Only new diversification models that result in a large discrepancy between inferred and simulated phylogeny will be worth the effort and computational burden 
-to implement a species tree prior for in a Bayesian framework.
-
-Our method is programmed as an R package called \verb;pirouette;.
-\verb;pirouette; is built on \verb;babette; [\cite{bilderbeek2018babette}], 
-which calls BEAST2 [\cite{bouckaert2014beast}].
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Description}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\verb;pirouette; is written in the R programming language (\cite{R}).
-The goal of \verb;pirouette; is to quantify the importance of a tree prior.
-It does so by measuring the inference error made for a given reconstructed phylogeny, 
-simulated under a (usually novel) diversification model.
-We refer to the true prior that generated the given tree
-as the 'generative tree prior' $\mathit{p_{G}}$.
-Many tree priors have a parameter setting in which these are equivalent
-to a standard tree prior.
-For example, a protracted birth-death model reduces to a standard birth-death model when speciation-completion rate approaches infinity [i.e.
-rate $\lambda$ in \cite{etienne2014estimating}, eqs.
-(2b) and (2c)].
-When benchmarking a novel tree prior, 
-one will typically construct phylogenies 
-for different combinations of the diversification model's parameters, 
-to assess under which scenarios the inference error cannot be neglected. 
-While we recommend many replicate simulations when assessing a novel tree prior, our examples contain only one replicate as they are for illustrative purposes only.
-
-\verb;pirouette; is very flexible and allows the user 
-to specify a wide variety of custom settings. 
-These settings can be grouped in macro-sections, 
-according to how they operate in the pipeline. 
-We summarize them in tables~\ref{tab:options} and~\ref{tab:definitions}.
-
-Although many possible tests can be performed, 
-we present here what we think are the five most important 
-example research questions (ERQ) that \verb;pirouette; can answer. 
-These ERQs gradually build up in complexity.
-For each of these ERQs we will quantify the importance of a tree prior.
-
-\begin{sidewaystable}
-\centering
-  \begin{tabular}{|p{3.2cm}|p{9.7cm}|p{4.5cm}@{}|}
-    \hline
-    \centering
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{Sub-argument} & 
-    \textbf{Description} &
-    \textbf{Possible values} \\ 
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;tree_prior; &
-    Macroevolutionary diversification model &
-    Yule, BD, CBS, CCP, CEP \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;clock_model; &
-    Clock for the DNA mutation rates &
-    strict, relaxed log-normal \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;site_model; &
-    Nucleotide substitution model &
-    JC, HKY, TN93, GTR \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;mutation_rate; &
-    Pace at which mutation occurs &
-    mutation rate $\in \mathbb{R}_{>0}$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;root_sequence; &
-    DNA sequence at the root of the tree &
-    any combination of a, c, g, t \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;model_type; &
-    Criterion to select an inference model &
-    Generative, Candidate \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;run_if; &
-    Condition under which an inference model is used &
-    Always, Best candidate \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;do_measure_evidence; &
-    Sets whether or not the evidence of the model is to be computed &
-    TRUE, FALSE \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;error_function; &
-    Specifies how to measure the error &
-    nLTT, $|\gamma|$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    burn\_in\_fraction &
-    Specifies the percentage of initial posterior trees to discard &
-    \verb;burn_in_fraction; $\in [0, 1]$\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-  \end{tabular}
-  \caption{
-    Most important parameter options.
-    Yule = pure birth model (\cite{yule})
-    BD = birth death (\cite{nee1994reconstructed}), 
-    CBS = coalescent Bayesian skyline (\cite{drummond2005bayesian}), 
-    CCP = coalescent constant-population, 
-    CEP = coalescent exponential-population,
-    JC = Jukes and Cantor (\cite{jukes1969evolution}), 
-    HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
-    TN93 = Tamura and Nei (\cite{tamura1993estimation}), 
-    GTR = Generalized time-reversible model (\cite{tavare1986some}).
-  }
-  \label{tab:options}
-\bigskip
-
-  \begin{tabular}{|@{}c|p{4cm}|p{12cm}|}
-    \hline
-    \centering
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{Symbol} &
-    \textbf{Macro-argument} &
-    \textbf{Description} \\
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{G}$ &
-    Generative model &
-    The full setting to produce BEAST2 input data. 
-    Its core features are the tree prior $\mathit{p_{G}}$, the clock 
-    model $\mathit{c_{G}}$ and the site model $\mathit{s_{G}}$. \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
-    $\mathit{A}$ &
-    Alignment model &
-    It includes the parts of the generative model that directly affect the 
-    alignment generation, like the clock model $\mathit{c_{G}}$ and 
-    the site model $\mathit{s_{G}}$. Additional arguments can be provided, 
-    such as the mutation rate and the root sequence. \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{X_{i}}$ &
-    $i$-th candidate experiment &
-    Full setting for a Bayesian inference. It is made by a 
-    candidate inference model $\mathit{I_{i}}$ and its 
-    inference conditions $\mathit{C_{i}}$. \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{I}$ &
-    Inference model &
-    Phylogenetic inference model to run BEAST2. Likewise the generative model $G$, 
-    its main components are the tree prior $\mathit{p_{I}}$, the 
-    clock model $\mathit{c_{I}}$ and the site model $\mathit{s_{I}}$. \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ 
-    is used in the inference. 
-    They are composed by the model type, run condition and 
-    wether to measure the evidence. \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    $\mathit{E}$ & Error measure parameters & 
-    Errors measurement setup that can be specified providing an 
-    error function to measure the difference between the original phylogeny 
-    and the inferred posterior. The initial part of the posterior that is 
-    reckoned as not representative can be discarded using a burn-in fraction. \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \iffalse
-    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & 
-      $\mathit{\tau_{p}}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \fi
-    \hline 
-  \end{tabular}
-  \caption{
-    Definitions of terms and relative symbols used in the main text and in 
-    Fig~\ref{fig:pipeline}. To run the pipeline $\mathit{A}$, $\mathit{X}$ 
-    and $\mathit{E}$ must be specified. Examples can be found in 
-    listings~\ref{lst:create_alignment}, 
-    \ref{lst:create_gen_experiment_explicit} and 
-    \ref{lst:create_error_measure_params}.
-  }
-  \label{tab:definitions}
-\end{sidewaystable}
-
-\subsection{pirouette's pipeline}
-\label{subsec:pipeline}
-
-\begin{figure}
-  \centering
-  \includegraphics[width = \textwidth]{workflow3.png}
-  \caption{
-    \texttt{pirouette} pipeline.
-    The pipeline starts from a phylogeny (1a) simulated by the generative tree prior 
-    $\mathit{p_{G}}$.
-    The phylogeny is converted to an alignment (2a) using the generative alignment model 
-    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed by a clock and a site model. 
-    The user supplies one or more experiments.
-    For each candidate experiment $\mathit{X_{i}}$ 
-    (a tuple of inference model $\mathit{I_{i}}$ and condition $\mathit{C_{i}}$),
-    if its condition $\mathit{C_{i}}$ is 
-    satisfied (which can depend on the alignment), 
-    the corresponding inference model $\mathit{I} = \mathit{I_{i}}$ is selected 
-    to be used in the next step.
-    The inference models (3) of the selected experiments use the alignment (2a) 
-    to each create a Bayesian posterior of (parameter estimates and) 
-    phylogenies (4a). 
-    Each of the posteriors' trees is compared to the true phylogeny (1a) 
-    using error measure $\mathit{E}$, 
-    resulting in an error distribution (5a). 
-    Optionally, for each selected inference model a twin pipeline can be run.
-    A twin phylogeny (1b) can be generated from the original 
-    phylogeny (1a) using the tree prior $\mathit{p_{I}}$ of the selected inference model (3).
-    Analogously, a twin alignment is simulated from the twin phylogeny along with clock model $\mathit{c_{I}}$ and site model $\mathit{s_{I}}$ imported from the selected inference model.
-    The twin pipeline mimics the procedure of the main pipeline, resulting in a twin error distribution (5b).
-  }
-  \label{fig:pipeline}
-\end{figure}
-
-The pipeline is summarized by the following steps:
-\begin{enumerate}
-  \item from a given phylogeny an alignment is simulated 
-    under a known alignment model $\mathit{A}$;
-  \item from this alignment, according to the specified inference conditions $\mathit{C}$, 
-    an inference model $\mathit{I}$ is picked (which may differ from the 
-    generative model);
-  \item the inference model and the alignment are used 
-    to infer a posterior distribution of phylogenies;
-  \item the phylogenies in the posterior are compared with the given phylogeny 
-    to estimate the error made, according to the error measure $\mathit{E}$ specified 
-    by the user;
-\end{enumerate}
-The pipeline is visualized in Fig.~\ref{fig:pipeline}. 
-There is also the option to generate a 'twin tree', 
-that goes through the same pipeline. 
-The utility of this twin tree will be explained below.
-
-The first step simulates a DNA alignment from a given 
-phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a)
-using the DNA alignment parameters.
-The DNA alignment parameters consist of a (DNA) root sequence, 
-a (DNA) mutation rate, a clock model and a nucleotide substitution
-model, which will be referred to as site model.
-The root sequence is the DNA sequence of the shared common ancestor,
-and is set to four equally-sized mononucleotide blocks by default, as this
-helps interpreting the resulting alignment.
-Supported site models are JC, HKY, TN and GTR. Only the strict
-clock model is supported in this step.
-
-The second step (Fig.~\ref{fig:pipeline}, 3)
-selects one or more inference models $I$ from a set of inference 
-models $I_{1},\dots,I_{n}$. 
-We define an experiment $X_{i}$ as the combination of 
-an inference model $I_{i}$ and the conditions $C_{i}$ 
-to actually use it in the inference step.
-For example, we can express that an inference
-model (a combination of a tree prior, clock model and site model) 
-is known to include the generative/true tree prior and that 
-we definitely want to use that model in our inference. 
-As a second example, wwe can also express that we've submitted a set of 
-candidate inference models,
-of which only the best should be used in the actual inference.
-In the first example, we specified the condition $C_{i}$ that this
-generative model should always be run, where in the second example,
-we specified condition $C_{i}$ that a candidate model should only be run
-when being the best.
-A 'best' model is defined as the inference model with
-the highest evidence (a.k.a. marginal likelihood), given the alignment 
-simulated in the previous step.
-The evidence of an inference model is estimated by nested 
-sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. 
-Scripted use of BEAST2 packages is only possible under Linux and Mac.
-Windows systems can do the model comparison for shorter DNA sequences 
-using the web interface of \verb;mcbette; [\cite{mcbette}].
-
-The third step infers the posteriors,
-using the simulated alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 4a),
-and those inference models that were selected in the previous step (3). 
-For each selected experiment a posterior is inferred, using the 
-\verb;babette; [\cite{bilderbeek2018babette}] R package which makes use of BEAST2. 
-This step usually takes up most of the full pipeline's computation time.
-
-The fourth step quantifies the inference error made. First the 
-burn-in fraction is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
-in which it samples an unrepresentative part of state space. By default, \verb;pirouette; 
-removes the first 10\% of the posterior.
-From the remaining posterior \verb;pirouette; 
-creates an error distribution, by measuring the difference
-between the true tree and each of the posterior 
-trees (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
-The default way to quantify the difference between two phylogenies
-is the nLTT statistic (\cite{janzen2015approximate}), but any 
-user-defined error statistic can be used.
-
-\subsection{Twin tree}\label{subsec:twinning}
-
-An optional step is to generate a 'twin tree' $\tau_{I}$
-(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
-that will be analyzed in the same way as the true tree.
-A phylogeny $\tau$ is the combination of 
-branching times $\Vec{t}$ and topology $\psi$.
-Assume that a non-standard generative prior $\mathit{p_{G}}$
-produces a tree $\tau_{\mathit{G}}$,
-with branching times $\Vec{t}_{\mathit{G}}$ and 
-topology $\psi_{\mathit{G}}$.
-The twinning process T, in accordance with a chosen inference 
-model $\mathit{I}$, creates a tree $\tau_{\mathit{I}}$
-with branching times $\Vec{t}_{\mathit{I}}$ while preserving the 
-topology $\psi_{\mathit{G}}$:
-
-\begin{align}
-  \tau_{\mathit{G}} = (\Vec{t}_{\mathit{G}}, \psi_{\mathit{G}}) 
-  \xrightarrow[]{\mathit{T}} 
-  \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, \psi_{\mathit{G}})
-\end{align}
-
-To achieve this, we find the parameters $\theta^{*}_{\mathit{I}}$ 
-(e.g. speciation and extinction rates, in case of a birth-death model) 
-that maximize the likelihood $L_{\mathit{I}}$ applied 
-to the true tree, conditioned on its number of tips $n_{\mathit{G}}$,
-where $L_{\mathit{I}}$ is the likelihood function  
-associated with the chosen inference model $\mathit{I}$:
-
-\begin{align}
-    max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{G}}, n_{\mathit{G}})] 
-\rightarrow \theta^{*}_{\mathit{I}}.
-\end{align}
-
-We use $\theta^{*}_{\mathit{I}}$ to simulate an equal number 
-$n_{\mathit{I}} = n_{\mathit{G}}$ 
-of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree 
-$\tau_{\mathit{I}}$, under the process $\mathit{I}$, 
-while preserving the same topology.
-
-Using a twin tree serves as a control: both $\tau_{\mathit{G}}$ and 
-$\tau_{\mathit{I}}$ will produce an error distribution. 
-These errors unavoidably accumulate across the \verb;pirouette; pipeline,
-for both the true and twin tree, due to several possible reasons: either tree may be an unlikely representative
-of the underlying tree prior, the alignment simulation may give different
-numbers of mutations for the same evolutionary time, 
-and the MCMC will sample different states.
-These errors can be equalized by imposing the same tree topology and by
-replicating the pipelines a sufficient number of times.
-
-The left-over differences in error distributions will be 
-consequences of different phylogenies and tree priors for the two parallel 
-pipelines: the main pipeline uses the tree $\tau_{\mathit{G}}$ and a
-possibly mismatched prior $\mathit{G}$, where the twin pipeline
-uses the tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
-So, for the twin route, if the process is repeated a sufficient number of times, the resulting error distribution cannot be 
-attributed to a mismatch between the true tree prior and the tree prior used in
-inference.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Installation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\verb;pirouette; will be made available on CRAN from which it can then be easily installed easily from CRAN:
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install.packages("pirouette")
-\end{lstlisting}
-
-Until it is on CRAN, and for the most up-to-date version, 
-one can download and install the package from \verb;pirouette;'s GitHub 
-repository:
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = ht,
-    frame = single
-]
-usethis::install_github("richelbilderbeek/pirouette")
-\end{lstlisting}
-
-To start using \verb;pirouette;, load its functions in the global namespace 
-first:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-library(pirouette)
-\end{lstlisting}
-Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
-This can be done from within R, using:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install_beast2()
-\end{lstlisting}
-For the option to select the best candidate model,
-\verb;pirouette; needs the "NS" BEAST2 package [\cite{maturana2018model}].
-It can be installed from within R, using:
-
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install_beast2_pkg("NS")
-\end{lstlisting}
-
-An overview of \verb;pirouette;'s main functions is shown in 
-table~\ref{tab:functions}. 
-Their usage is demonstrated in the example code below.
-All \verb;pirouette;'s functions are documented,
-have a useful example and sensible defaults.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[h]
-  \centering
-  \begin{tabular}{ | l | l | }
-    \hline
-    \textbf{Name} & \textbf{Description} \\
-    \hline
-    \verb;pir_run; & Run \verb;pirouette; \\
-    \verb;pir_plot; & Show the \verb;pirouette; results as a plot \\
-    \verb;create_pir_params; & Create the \verb;pirouette; parameters \\
-    \hline
-    \verb;create_alignment_params; & Create the alignment parameters \\
-    \verb;create_twinning_params; & Create the twinning parameters \\
-    \verb;create_experiment; & Create one experiment \\
-    \verb;create_error_measure_params; & Create the error measurement parameters \\
-    \hline
-  \end{tabular}
-  \caption{pirouette's main functions.}
-  \label{tab:functions}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 1}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's 
-inference,
-by measuring the difference between a given tree and 
-phylogenies in a posterior. Due to stochasticity, all posterior trees
-will differ from the given phylogeny $\tau_{\mathit{G}}$, 
-even when the generative tree prior $\mathit{p_{G}}$ 
-and alignment model $\mathit{A}$ are the same as 
-those in the inference model $\mathit{I}$.
-Measuring this difference allows us to know the baseline error
-of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' 
-all the tree priors that are available to be used within an inference model 
-(see Table~\ref{tab:options}).
-
-Now we can formulate the first example research question that \verb;pirouette; 
-can answer.
-
-\subsection{Question}
-
-"What is the inference error made on phylogenies
-created by a standard diversification model?"
-
-\subsection{Answer}
-
-In this example we use a standard generative tree prior $\mathit{p_{G}^0}$, 
-namely the Yule (pure-birth) tree prior. 
-We choose to use a small tree with six taxa to keep
-the calculations short and the figure more readable.
-We pick a crown age of ten time units. This value is 
-completely arbitrary, but it ties in with the mutation rate 
-used in simulating an alignment in the next step.
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = ht,
-    frame = single, 
-    label = {lst:create_yule_tree}, 
-    caption = {
-      Create a Yule tree. 
-      The resulting tree is shown in figure~\ref{fig:yule_tree}.
-    }
-  ]
-phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
-\end{lstlisting}
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_1/true_tree.png}
-  \caption{The Yule tree, as created by listing~\ref{lst:create_yule_tree}.}
-  \label{fig:yule_tree}
-\end{figure}
-
-The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny, as described in subsection~\ref{subsec:pipeline}.
-In this example, the root sequence consists in four blocks of 250 
-mononucleotides each, 
-where the per-nucleotide mutation rate is 0.1 mutations per unit time.
-We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
-and a strict clock model as these are the simplest.
-A JC site model assumes that mutation rates between nucleotides are equal and constant. 
-A strict clock model assumes that the mutation rates 
-of all species are equal and constant.
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = ht,
-    frame = single,
-    label = {lst:create_alignment}, 
-    caption = {Create an alignment.}
-  ]
-alignment_params <- create_alignment_params(
-  clock_model = create_strict_clock_model(),
-  site_model = create_jc69_site_model(),
-  mutation_rate = 0.1,
-  root_sequence = create_blocked_dna(length = 1000)
-)
-\end{lstlisting}
-
-As the site and clock models used here are also the defaults, 
-these function arguments can be safely omitted: we just show these
-explicitly for clarity.
-
-In the second step we state our experiment.
-We define an experiment $\mathit{X}$ as a combination of an inference model 
-$\mathit{I}$
-and conditions $\mathit{C}$.
-In this example we pick $\mathit{I}$ to be the same inference model as the 
-generative one,
-which is the chosen Yule tree prior $\mathit{p_{G}^0}$, as well as site and clock 
-models defined in $\mathit{A}$, respectively JC and strict.
-We specify in $\mathit{C}$ that the experiment will always be run.
-
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \textbf{model\_type} &
-    \textbf{run\_if} &
-    \textbf{do\_measure\_evidence} & 
-    \textbf{inference model} \\ 
-    \hline
-    generative &
-    always &
-    FALSE &
-    JC, strict, Yule \\
-    \hline
-  \end{tabular}
-  \caption{
-    Inference conditions and inference model for ERQ 1.
-    JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-  }
-  \label{tab:RQ1}
-\end{table}
-
-Listing~\ref{lst:create_gen_experiment_explicit} shows how to
-set up this experiment:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_gen_experiment_explicit},
-  caption = {
-    Create the experiments to answer research question 1, 
-    using explicit arguments.
-  }
-]
-experiment <- create_experiment(
-  inference_conditions = create_inference_conditions(
-    model_type = "generative", 
-    run_if = "always"
-  ), 
-  inference_model = create_inference_model(
-    tree_prior = create_yule_tree_prior(),
-    clock_model = create_strict_clock_model(), 
-    site_model = create_jc69_site_model()
-  )
-)
-\end{lstlisting}
-
-Experiments must be bundled in a list to work, even if only one is provided, as 
-in this case:
-
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_gen_experiment},
-  caption = {
-    Create the experiments to answer research question 1.
-  }
-]
-experiments <- list(experiment)
-\end{lstlisting}
-
-The objects we just created must be bundled
-in a \verb;pirouette; parameters structure,
-together with error measurement parameters $\mathit{E}$ and
-twinning parameters $\mathit{T}$.
-\giovanni{I suggest to refer to the twinning parameters only when strictly required and with the minimum possible amount of info provided, as we set it in such a way that the tree prior is imported from the inference model in an automatic way.}
-
-We use the default error measurement
-parameters, which uses a burn-in fraction of 10\% and the nLTT statistic to
-measure the difference between phylogenies. For clarity,
-we create this setup explicitly here:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_error_measure_params},
-  caption = Calling \texttt{create\_error\_measure\_params}.
-]
-error_measure_params <- create_error_measure_params(
-  error_function = get_nltt_error_function(),
-  burn_in_fraction = 0.1
-)
-\end{lstlisting}
-
-We will not use twinning here yet.  
-All these arguments are bundled
-and checked by \verb;create_pir_params;:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_pir_params},
-  caption = Calling \texttt{create\_pir\_params}.
-]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  error_measure_params = error_measure_params,
-  twinning_params = NA
-)
-\end{lstlisting}
-
-Now we can use our Yule tree and \verb;pir_params; to measure 
-the inference error made on phylogenies
-created by a standard diversification model:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:pir_run},
-  caption = Calling \texttt{pir\_run}.
-]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-\end{lstlisting}
-
-The error distribution can be plotted directly using \verb;pir_plot;:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:pir_plot},
-  caption = Calling \texttt{pir\_plot}.
-]
-pir_plot(errors)
-\end{lstlisting}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_1/errors.png}
-  \caption{
-    ERQ 1: the inference error made from a phylogeny 
-    when the generative and inference model are the same.
-    The vertical dashed line indicates the median error.
-  }
-  \label{fig:example_1}
-\end{figure}
-
-The resulting error distribution, as shown figure~\ref{fig:example_1},
-shows the inference error 
-when $\mathit{I}$ matches with the generative model given 
-by $\mathit{A}$ and $\mathit{p_{G}^0}$.
-This error distribution can serve as a control,
-as it is obtained from a tree of which the tree prior is standard and known.
-Obtaining this control is an inherent part of \verb;pirouette;. 
-We call it 'twinning' and we use it from ERQ 3.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In the previous example we selected the inference model $\mathit{I}$ to match 
-with a known generative tree prior $\mathit{p_{G}^0}$.
-However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, 
-thus cannot be part of a standard inference 
-model (i.e. an inference model using a standard tree prior).
-In such a case, the user has to somehow select 
-a standard tree prior to be 
-used in the inference.
-There are two ways to deal with this problem: hand-picking a tree
-prior, or let the best of a set of models be picked. 
-Hand-picking a tree prior can be rational if one standard tree prior 
-is more suitable than others for the inference, for example, when
-the novel tree prior is most similar to a certain tree prior.
-Instead of hand-picking, the best inference model from a list of candidate 
-models can be used in inference.
-\verb;pirouette; selects the 'best' model, that is, 
-the model with the highest evidence (also known as marginal likelihood).
-
-We show the procedure using \verb;pirouette; to answer the second research 
-question.
-
-\subsection{Question}
-
-"What is the inference error made on a novel phylogeny when
-using the best inference model?"
-
-\subsection{Answer}
-
-This time we start with a tree generated from an unknown 
-diversification model, that has six taxa and a crown age of ten:
-
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht,
-  frame = single, 
-  label = {lst:unknown_phylogeny},
-  caption = A phylogeny generated by an unknown diversification model.
-]
-phylogeny  <- ape::read.tree(
-  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
-)
-\end{lstlisting}
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_2/tree_unknown.png}
-  \caption{The tree derived from an unknown diversification process, 
-    as created by listing~\ref{lst:unknown_phylogeny}.
-  }
-\end{figure}
-
-The first step is to simulate a DNA alignment from the 
-given phylogeny. 
-We will re-use the alignment parameters of the previous example 
-as shown in listing~\ref{lst:create_alignment}.
-
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \textbf{model\_type} &
-    \textbf{run\_if} & \textbf{do\_measure\_evidence} & 
-    \textbf{inference model} \\ 
-    \hline
-    candidate & best candidate & TRUE & JC, strict, Yule \\
-    candidate & best candidate & TRUE & JC, strict, BD   \\
-    ...       & ...            & ... & ...              \\
-    candidate & best candidate & TRUE & GTR, RLN, CCP    \\
-    candidate & best candidate & TRUE & GTR, RLN, CEP    \\
-    \hline
-  \end{tabular}
-  \caption{
-    Inference conditions and inference model for ERQ 2.
-    JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-    BD: birth-death tree prior.
-    GTR: GTR site model.
-    RLN: relaxed log-normal clock model.
-    CCP: coalescent constant-population tree prior.
-    CEP: coalescent exponential-population tree prior.
-  }
-  \label{tab:RQ2}
-\end{table}
-
-In the second step, unlike ERQ 1, we provide a 
-set of experiments $\{\mathit{X_{i}}\}_{i=1}^{40}$. All the experiments have 
-the same inference conditions $\mathit{C_{i}} = \mathit{C}$ but different 
-candidate inference models $\mathit{I_{i}}$. Only one of them will be used 
-to run the actual inference. According to the specified $\mathit{C}$, 
-\verb;pirouette; will measure the evidence for each of them and select only 
-the one with the highest evidence as the best candidate.
-
-As model selection is commonly performed on the full list of available 
-candidate models, \verb;pirouette; has a dedicated function to provide it:
-\verb;create_all_experiments; will create a full set of 40 experiments,
-containing the inference models of all combinations of 
-4 site models, 2 clock models and 5 tree priors.
-
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht, 
-  frame = single, 
-  label = {lst:create_all_experiments},
-  caption = Create all 40 candidate experiments.
-]
-experiments <- create_all_experiments()
-\end{lstlisting}
-
-For all the other steps we use the default setups. 
-We can create the complete
-\verb;pirouette; parameter set like this (which is the
-same as listing~\ref{lst:create_pir_params}, but using the defaults):
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_pir_params_defaults},
-  caption = Create a \texttt{pir\_params} with many defaults.
-]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments
-)
-\end{lstlisting}
-
-We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
-and plot the results (listing~\ref{lst:pir_plot}),
-as shown in figure~\ref{fig:example_2}.
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_2/errors.png}
-  \caption{
-    ERQ 2: the inference error from
-    when selecting the best inference model.
-  }
-  \label{fig:example_2}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 3}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In ERQ 1 we measured the baseline error of the \verb;pirouette; method,
-by using the same standard tree prior in both the generation of 
-the original tree, as well as in the inference.
-This baseline error distribution serves as a control,
-to quantify the impact of a new (non-standard) tree prior.
-To measure that impact, we compare the inference error 
-from the new tree prior to the baseline error,
-by using a twin tree (see subsection~\ref{subsec:twinning}).
-In this ERQ, we hand-pick a standard tree prior $\mathit{p_{G}}$.
-
-\subsection{Question}
-
-"What is the inference error made from a phylogeny, 
-when hand-picking a tree prior, compared to the background noise?"
-
-\subsection{Answer}
-
-Most of the settings for this ERQ are
-the same as in ERQ 2: we will re-use the phylogeny from an 
-unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
-the alignment parameters (listing~\ref{lst:create_alignment}).
-
-We create an experiment as in 
-listing~\ref{lst:create_gen_experiment},
-but with all defaults:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_gen_experiment_defaults},
-  caption = Create a list of one default generative experiment.
-]
-experiment <- create_gen_experiment()
-\end{lstlisting}
-
-Creating the complete set of experiments:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_experiments},
-  caption = Create a list of one experiment.
-]
-experiments <- list(experiment)
-\end{lstlisting}
-
-This time, however, we measure both the baseline and complete inference error.
-We set the twinning parameters $\mathit{T}$ to create a twin tree,
-as described in the 'Twinning' section \ref{subsec:twinning}.
-
-Creating the twinning parameters is trivial with the default settings.
-For clarity, however, we explicitly show the most
-important arguments:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_twinning_params},
-  caption = Create the default twinning parameters.
-]
-twinning_params <- create_twinning_params(
-  twin_model = "bd", 
-  method = "random_tree"
-)
-\end{lstlisting}
-\giovanni{We want to import the "twin\_model" from the inference model.}
-\richel{Yes, that is an option, but according to Rampal not important enough
-to add now}
-
-We combine all the parameters using \verb;create_pir_params;:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_pir_params_with_twinning},
-  caption = Create the default twinning parameters.
-]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  twinning_params = twinning_params
-)
-\end{lstlisting}
-
-We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
-and plot the results (listing~\ref{lst:pir_plot}),
-The output is shown in figure~\ref{fig:example_3}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_3/errors.png}
-  \caption{
-    ERQ 3: the inference error made 
-    when hand-picking a generative tree prior versus the control.
-    Here, the 'twin' column shows the baseline inference error,
-    as measured on the twin tree.
-  }
-  \label{fig:example_3}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-When no inference model can be hand-picked, 
-one can select the best model from a set,
-as shown in ERQ2.
-The resulting error distribution contains the error accumulated in the pipeline,
-without quantifying to what extent this error is due to the mismatch between 
-the generative model $\mathit{G}$ and the selected inference model 
-$\mathit{I}$.
-As twinning (see section \ref{subsec:twinning}: 'Twinning') shows us the baseline error, we can quantify the contribution of the tree prior to the error.
-
-\subsection{Question}
-
-"What is the inference error from a phylogeny 
-derived from an unknown speciation model,
-when selecting the best inference model, 
-in relation to the baseline error?"
-
-\subsection{Answer}
-
-Here, as in ERQs 2 and 3, we start with a phylogeny 
-generated by an unknown speciation model (listing~\ref{lst:unknown_phylogeny}). 
-The first step is to simulate a DNA alignment 
-from the given phylogeny, for which we will use the same code 
-as shown in listing~\ref{lst:create_alignment}.
-
-We specify that we want all 40 inference models to compete 
-by calling \verb;create_all_experiments;,
-as we previously did in listing~\ref{lst:create_all_experiments}.
-We choose again to use the default error measurement setup,
-as shown in listing~\ref{lst:create_error_measure_params}.
-To get an idea of a baseline error made by BEAST2, 
-we also make use of the twinning option by using
-the same call to \verb;create_pir_params; as shown in 
-listing~\ref{lst:create_pir_params_with_twinning}.
-We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
-and plot the results (listing~\ref{lst:pir_plot}).
-The output is shown in figure~\ref{fig:example_4}.
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_4/errors.png}
-  \caption{
-    Example 4: the error made by BEAST2 from a phylogeny, when the generative 
-    model is known, compared to candidates with similar inference models, in 
-    relationship to the background error.
-  }
-  \label{fig:example_4}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 5}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-If we are moderately sure about hand-picking an inference model,
-but would also like to check how well the hand-picked model compares
-to the others, we can do both. Of course, as we are interested in the
-baseline error for both the hand-picked and best candidate model,
-we need to enable twinning as well. This example is a combination
-of the previous two ERQs.
-
-\subsection{Question}
-
-"What is the error made by BEAST2 from a phylogeny, 
-when the generative model is hand-picked, 
-compared to candidates with similar inference models, 
-in relation to the background error?"
-
-\subsection{Answer}
-
-In this example we use a tree generated by the Yule process.
-Instead of using the common Yule tree as in ERQ 1, we will use
-a Yule tree of low likelihood, 
-i.e. the one already used in ERQs 2-4, 
-by using the code shown in listing~\ref{lst:unknown_phylogeny}.
-The first step is to simulate a DNA alignment 
-from the given phylogeny, for which we will use the same code 
-as shown in listing~\ref{lst:create_alignment}.
-
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \textbf{model\_type} & 
-    \textbf{run\_if} & 
-    \textbf{do\_measure\_evidence} & 
-    \textbf{inference model} \\ 
-    \hline
-    generative & always         & TRUE & JC, strict, Yule \\
-    candidate  & best candidate & TRUE & JC, strict, BD \\
-    candidate  & best candidate & TRUE & JC, strict, CCP \\
-    candidate  & best candidate & TRUE & JC, strict, CEP \\
-    \hline
-  \end{tabular}
-  \caption{
-    Experimental setup to answer the fourth research question.
-    JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-    BD: birth-death tree prior.
-    CCP: coalescent constant-population tree prior.
-    CEP: coalescent exponential-population tree prior.
-  }
-  \label{tab:experiment_5}
-\end{table}
-
-In the second step we state our experiments. 
-In this example we must specify both the generative and candidate experiments. 
-Here we specify the generative model:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-generative_experiment <- create_gen_experiment()
-\end{lstlisting}
-
-We can specify all the candidates simply by calling 
-\verb;create_all_experiments; and excluding the generative inference model,
-resulting in 39 candidate experiments:
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = H,
-    frame = single
-]
-candidate_experiments <- create_all_experiments(
-  exclude_model = generative_experiment$inference_model
-)
-\end{lstlisting}
-
-We need to combine these experiments into one set of experiments:
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = H,
-    frame = single
-]
-experiments <- c(list(generative_experiment), candidate_experiments)
-\end{lstlisting}
-
-We choose again to use the default error measurement parameters
-created in listing~\ref{lst:create_error_measure_params}.
-We enable twinning, by calling \verb;create_pir_params;
-as shown in listing~\ref{lst:create_pir_params_with_twinning}.
-
-We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
-and plot the results (listing~\ref{lst:pir_plot}).
-The output is shown in figure~\ref{fig:example_5}.
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_5/errors.png}
-  \caption{
-    Example 5: the inference error made, 
-    when the generative model is hand-picked, 
-    compared to the best of 39 candidate models, 
-    each in relationship to the background error.
-  }
-  \label{fig:example_5}
-\end{figure}
-
-\begin{table}
-  \begin{tabular}{ | r | c | c | c | }
-    \hline
-    \textbf{Compare with} & \textbf{TB} & \textbf{WG} & \textbf{WB} \\ 
-    \hline
-    \textbf{TG} & Y & Y & N \\
-    \textbf{TB} & . & N & Y \\
-    \textbf{WG} & . & . & Y \\
-    \textbf{WB} & . & . & . \\
-    \hline
-  \end{tabular}
-  \caption{
-    Error distributions that are relevant to compare.
-    TG: true tree, generative model.
-    TB: true tree, best candidate model.
-    WG: twin tree, generative model.
-    WB: twin tree, best candidate model.
-  }
-  \label{tab:relevant_comparisions}
-\end{table}
-
-\begin{table}
-  \begin{tabular}{ | r | c | p{9cm} | }
-    \hline
-    \textbf{Condition} & \textbf{Expectation} & \textbf{Interpretation} \\ 
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG > TB$}       & Unexpected & Novel tree prior is more related to 
-      best candidate than hand-picked tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG \approx TB$} & Possible   & Hand-picked tree prior is just as 
-      suitable as the best candidate tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG < TB$}       & Expected   & Hand-picked tree prior is the most 
-      related tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG > WG$}       & Unknown    & Novel tree prior important \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG \approx WG$} & Unknown    & Novel tree prior unimportant \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG < WG$}       & Unexpected & Twinning procedure increases 
-      inference errors when using hand-picked tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TB > WB$}       & Expected   & Impact of novel tree prior cannot 
-      be compensated for by model selection: twin tree with low likelihood? \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TB \approx WB$} & Possible   & Best candidate tree priors perform 
-      equally well in true and twin tree: true and twin tree similar? \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TB < WB$}       & Unexpected & Twinning procedure increases 
-      inference errors  when using best tree prior candidate \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$WG > WB$}       & Unexpected & Hand-picked tree prior (that equals 
-      the twinning tree prior!) worse than best candidate tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$WG \approx WB$} & Possible   & Twin tree fits equally well to the 
-      hand-picked and best candidate tree prior  \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$WG < WB$}       & Expected   & Hand-pick tree prior (that equals 
-      the twinning tree prior) performs as expected \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-  \end{tabular}
-  \caption{
-    Interpretation of the different error distributions.
-    This assumes that the operators ($<$, $\approx$ and $>$) to compare
-    distributions are defined.
-    TG: true tree, generative model.
-    TB: true tree, best candidate model.
-    WG: true tree, generative model.
-    WB: true tree, best candidate model.
-  }
-  \label{tab:interpretation}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% RJCB will add: in six percent did the twin do worse, 
-% see https://github.com/richelbilderbeek/pirouette_article/issues/41
-
-In the five example research questions, 
-we showed how to use \verb;pirouette; 
-to quantify the importance of a tree prior in Bayesian phylogenetics.
-In all ERQs we used and assumed the most simple tree prior possible,
-which is the Yule tree prior.
-In principle any other (more complex) generative tree prior 
-can be tested, but we chose to provide the 
-simplest (and fastest to run) examples.
-All tests were performed on a single original tree, 
-hence not providing us enough statistical power to 
-evaluate the level of accuracy of the inference.
-
-However, if the same procedure were repeated and performed on a distribution 
-with a sufficient number of generative trees, 
-it could actually constitute a quantitative assessment 
-of the quality of the inference.
-To give an idea of the insights \verb;pirouette; can give,
-we pretend the results of the example research questions
-to be obtained from thousands of replicates.
-In that case, figure~\ref{fig:example_1} would show 
-the minimum error BEAST2 when the
-generative model is known, which can serve as 
-a minimum noise level assessment in a bigger experiment.  
-Figure~\ref{fig:example_2} shows that 
-the best candidate model gives an error distribution similar to the unknown 
-tree prior,
-hinting that the candidate model would be just as good in inference.
-Figure~\ref{fig:example_3} shows that 
-the twin model gives an error distribution similar to the unknown tree prior,
-hinting that the shape of the phylogeny with unknown tree prior
-does not effect the error made by BEAST2.
-Figure~\ref{fig:example_4} shows that 
-the twin model and best candidate model give 
-an error distribution similar to the unknown tree prior,
-hinting that the shape of the phylogeny with unknown tree prior
-does not effect the error made by BEAST2, 
-and that the candidate model would be just as good in inference.
-
-In conclusion, \verb;pirouette; can directly show 
-the comparison between the results obtained 
-adopting a new generative prior as compared to the error BEAST2 
-would make when operating on a tree generated 
-under a known, right and existing tree prior (the twin tree).
-From such results an user can estimate whether or not a new tree prior, 
-tailored on the generative process, is needed. 
-If this would indeed be the case, 
-one could implement their own tree prior as a new module in BEAST2.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{pirouette resources}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\verb;pirouette; is free, libre and open source software available at 
-\url{http://github.com/richelbilderbeek/pirouette},
-licensed under the GNU General Public License version 3.
-\verb;pirouette; depends on multiple packages, which are:
-\verb;babette; (\cite{bilderbeek2018babette}),
-\verb;becosys; (\cite{becosys}),
-\verb;DDD; (\cite{DDD}),
-\verb;devtools; (\cite{devtools}),
-\verb;geiger; (\cite{geiger}),
-\verb;ggplot2; (\cite{ggplot2}),
-\verb;knitr; (\cite{knitr}),
-\verb;lintr; (\cite{lintr}),
-\verb;mcbette; (\cite{mcbette}),
-\verb;nLTT; (\cite{nLTT}),
-\verb;PBD; (\cite{PBD}),
-\verb;phangorn; (\cite{phangorn}),
-\verb;phytools; (\cite{phytools}),
-\verb;rappdirs; (\cite{rappdirs}),
-\verb;rmarkdown; (\cite{rmarkdown}),
-\verb;Rmpfr; (\cite{Rmpfr}),
-\verb;stringr; (\cite{stringr}),
-\verb;subplex; (\cite{subplex}),
-\verb;TESS; (\cite{TESS}),
-\verb;testit; (\cite{testit}), 
-\verb;testthat; (\cite{testthat}) and
-\verb;tidyr; (\cite{tidyr}).
-
-\verb;pirouette;'s development takes place on GitHub,
-\url{https://github.com/richelbilderbeek/pirouette},
-which allows submitting bug reports, requesting features, 
-or adding code. To ensure a high quality, \verb;pirouette; 
-uses a continuous integration service, has a code coverage of 100\%
-and enforces the most commonly used R style guide (\cite{style_guide}).
-
-\verb;pirouette;'s is extensively documented on its website,
-its documentation and its vignettes.
-The \verb;pirouette; website is a good starting point to learn
-how to use \verb;pirouette;, as it links to tutorials and videos.
-The \verb;pirouette; package documentation describes
-all functions and liberally links to related functions.
-All exported functions show a minimal example as part of their documentation.
-The \verb;pirouette; vignette demonstrates extensively how 
-to use \verb;pirouette; in a more lightly written way.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Citation of pirouette}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Professionals using \verb;pirouette; can cite this
-article. To obtain the citation from within R, use:
-
-\begin{lstlisting}[language=R]
-> citation("pirouette")
-\end{lstlisting}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgments}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We would like to thank the Center for Information Technology of the University 
-of Groningen for their support and for providing access to the Peregrine 
-high performance computing cluster. 
-We thank the Netherlands 
-Organization for Scientific Research (NWO) for financial support 
-through a VICI grant awarded to RSE.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Data Accessibility}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-All code is archived at 
-\url{http://github.com/richelbilderbeek/pirouette_article},
-with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Authors' contributions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-RJCB, GL and RSE conceived the idea for the package. 
-RJCB created and tested the package, and wrote the first draft of the 
-manuscript.
-GL tested the package and contributed substantially to revisions.
-RSE contributed to revisions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% MEE style
-\bibliographystyle{mee}
-\bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results in detail}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Errors}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 1}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.25\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1/true_alignment.png}
-    };   
-    \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1/true_posterior_gen.png}
-    };   
-    \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1/true_error_violin_gen.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (C)
-      (C) edge [anchor = south] node {} (D)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_1_full_pipeline}
-  \caption{Example 1: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_1/esses.latex}
-% has label tab:esses_example_1
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.25\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_alignment.png}
-    };   
-    \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_posterior_best.png}
-    };   
-    \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_error_violin_best.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (C)
-      (C) edge [anchor = south] node {} (D)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_2_full_pipeline}
-  \caption{Example 2: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_2/esses.latex}
-% has label tab:esses_example_2
-
-\input{example_2/evidences.latex}
-% has label tab:evidences_example_2
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 3}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.4\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3/true_alignment.png}
-    };   
-    \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_posterior_gen.png}
-    };   
-    \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_error_violin_gen.png}
-    };   
-    \node[state] (AT) [right of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/twin_tree.png}
-    };   
-    \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3/twin_alignment.png}
-    };   
-    \node[state] (CT) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/twin_posterior_gen.png}
-    };   
-    \node[state] (DT) [below of = CT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/twin_error_violin_gen.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (C)
-      (C) edge [anchor = south] node {} (D)
-      (A) edge [anchor = east] node {} (AT)
-      (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = south] node {} (CT)
-      (CT) edge [anchor = south] node {} (DT)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_3_full_pipeline}
-  \caption{Example 3: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_3/esses.latex}
-% has label tab:esses_example_3
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {1.0\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.5\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{example_4/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_4/true_alignment.png}
-    };   
-    \node[state] (CB) [below of = B, rectangle, node distance=0.8\textheight] {
-
-\includegraphics[height=0.95\textheight]{example_4/true_posterior_best.png}
-    };   
-    \node[state] (DB) [below of = CB, rectangle, node distance=1.0\textheight] {
-      \includegraphics[height=0.65\textheight]{example_4/true_error_violin_best.png}
-    };   
-    \node[state] (AT) [right of = A, rectangle, node distance=2.0\textwidth] {
-      \includegraphics[height=0.4\textheight]{example_4/twin_tree.png}
-    };   
-    \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_4/twin_alignment.png}
-    };   
-    \node[state] (CTB) [below of = BT, rectangle, node distance=0.8\textheight] 
-{
-
-\includegraphics[height=0.95\textheight]{example_4/twin_posterior_best.png}
-    };   
-    \node[state] (DTB) [below of = CTB, rectangle, node 
-distance=1.0\textheight] {
-
-\includegraphics[height=0.65\textheight]{example_4/twin_error_violin_best.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south east] node {} (CB)
-      (CB) edge [anchor = south] node {} (DB)
-      (A) edge [anchor = east] node {} (AT)
-      (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = south] node {} (CTB)
-      (CTB) edge [anchor = south] node {} (DTB)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_4_full_pipeline}
-  \caption{Example 4: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_4/esses_best.latex}
-% has label tab:esses_example_4_best
-
-\input{example_4/esses_twin_best.latex}
-% has label tab:esses_example_4_twin_best
-
-\input{example_4/evidence_true.latex}
-% has label tab:evidences_example_4
-
-\input{example_4/evidence_twin.latex}
-% has label tab:evidences_example_4
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 5}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.4\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5/true_alignment.png}
-    };   
-    \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_posterior_gen.png}
-    };   
-    \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_error_violin_gen.png}
-    };   
-    \node[state] (CB) [right of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_posterior_best.png}
-    };   
-    \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_error_violin_best.png}
-    };   
-    \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.2\textheight]{example_5/twin_tree.png}
-    };   
-    \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5/twin_alignment.png}
-    };   
-    \node[state] (CTG) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/twin_posterior_gen.png}
-    };   
-    \node[state] (DTG) [below of = CTG, rectangle] {
-
-\includegraphics[height=0.2\textheight]{example_5/twin_error_violin_gen.png}
-    };   
-    \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/twin_posterior_best.png}
-    };   
-    \node[state] (DTB) [below of = CTB, rectangle] {
-
-\includegraphics[height=0.2\textheight]{example_5/twin_error_violin_best.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (CG)
-      (CG) edge [anchor = south] node {} (DG)
-      (B) edge [anchor = south east] node {} (CB)
-      (CB) edge [anchor = south] node {} (DB)
-      (A) edge [anchor = east] node {} (AT)
-      (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = south east] node {} (CTG)
-      (CTG) edge [anchor = south] node {} (DTG)
-      (BT) edge [anchor = south] node {} (CTB)
-      (CTB) edge [anchor = south] node {} (DTB)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_5_full_pipeline}
-  \caption{Example 5: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_5/esses_gen.latex}
-% has label tab:esses_example_5_gen
-
-\input{example_5/esses_best.latex}
-% has label tab:esses_example_5_best
-
-\input{example_5/esses_twin_gen.latex}
-% has label tab:esses_example_5_twin_gen
-
-\input{example_5/esses_twin_best.latex}
-% has label tab:esses_example_5_twin_best
-
-\input{example_5/evidence_true.latex}
-% has label tab:evidences_example_5
-
-\input{example_5/evidence_twin.latex}
-% has label tab:evidences_example_5
-
-\begin{table}
-  \begin{tabular}{ | r | c | }
-    \hline
-    \textbf{ERQ} & \textbf{run-time} \\ 
-    \hline
-      1 & 00:10:12 \\
-      2 & 00:34:10 \\
-      3 & 00:24:53 \\
-      4 & 01:04:50 \\
-      5 & 01:29:40 \\
-    \hline
-  \end{tabular}
-  \caption{
-    Time for an example research question to finish.
-    Time is in in hh:mm::ss.
-  }
-  \label{tab:runtime}
-\end{table}
 \end{document}

### 2019-05-07 content.tex
Split up in article.tex (big picture), content.tex (the content) and (supplement.tex) supplementary materials

\begin{abstract}

  \textbf{1. }
    Phylogenetic trees are current routinely reconstructed from an alignment of character sequences (usually nucleotide sequences). Bayesian tools, such as MrBayes, RevBayes and BEAST2, have gained much popularity over the last decade, as they allow joint estimation of the posterior distribution of the phylogenetic tree and the parameters of the underlying inference model.  
    An important ingredient of these Bayesian approaches is the species tree prior. While in principle the Bayesian framework allows for comparing different species tree priors and hence may elucidate the macroevolutionary processes underlying the species tree, in practice only 
    macroevolutionary models that allow for fast computation of the prior probability are used. 
    This begs the question how accurate the tree estimation is 
    when the real macroevolutionary processes are different 
    from those assumed in the tree prior. \\
  \textbf{2. }
    Here we present \verb;pirouette;, 
    a free, libre and open-source R package that assesses 
    the inference error made by Bayesian phylogenetics for a given macroevolutionary diversification model. \verb;pirouette; makes use of BEAST2, but its philosophy applies to any Bayesian phylogenetic inference tool\\
  \textbf{3. }
    We describe \verb;pirouette;'s usage and the biological scientific
    question it can answer, including full examples. \\
  \textbf{4. }
    Last, we discuss the results obtained by the examples and their interpretation. \\
\end{abstract}

{\bf Keywords:} computational biology, evolution, phylogenetics, tree prior,
  Bayesian model selection, BEAST2, babette, pirouette, R

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The development of new powerful Bayesian phylogenetic inference tools, 
such as BEAST [\cite{drummond2007beast}], 
MrBayes [\cite{huelsenbeck2001mrbayes}]
or RevBayes [\cite{hohna2016revbayes}], 
has been a major advance in building phylogenetic trees 
from character data (usually nucleotide sequences) extracted 
from extant (but also extinct) organisms, and hence in our understanding of the main drivers and modes of diversification.

BEAST [\cite{drummond2007beast}] is a typical Bayesian phylogenetics tool, 
that needs both character data and priors to infer a posterior.
Specifically, the species tree prior describes the process of diversification.
BEAST has built-in tree priors such as the Yule [\cite{yule}] and 
(constant-rate) birth-death [\cite{nee1994reconstructed}] models.
These simple tree priors are most commonly used, as these 
have sufficient biological complexity, while being computationally fast.
BEAST's successor, BEAST2 [\cite{bouckaert2014beast}],
has a package manager, that allows third-party users to extend existing 
functionalities.
Hence one can add novel diversification models by writing a BEAST2 package that contains the likelihood formula
of a phylogeny under the novel diversification model, i.e. the prior probability of a species tree.

Many such diversification models (and their associated probability 
algorithms) have been developed
developed, e.g., those in which diversification is 
time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
or diversity-dependent [\cite{etienne2011diversity}],
or where diversification rates change for specific lineages and their descendants [\cite{etienne2012conceptual}, 
\cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}], models that treat speciation as a process that takes 
time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
or as a burst of simultaneous branching
events [\cite{laudanno2018mbd}], or where diversification rate
depends on a trait that has two [\cite{maddison2007estimating}], 
or more [\cite{fitzjohn2012diversitree}] states,
even concealed states [\cite{beaulieu2016detecting}] 
or a combination of all these [\cite{herrera2018detecting}].
Only few of these diversification models are available as a BEAST2 package.

When a novel diversification model is introduced,
its performance in inference is commonly tested.
Part of a model's performance is its ability to recover parameters from simulated data with known parameters (e.g. [\cite{etienne2014estimating}]),
Ideally, the estimated parameter values closely match the known/true values.

Even when a diversification model passes the procedure described above,
it is not necessarily used in Bayesian inference.
Bayesian phylogenetic inference often requires that the prior probability of the phylogeny according to the diversification model has to be computed millions of times. Therefore, biologically interesting but computationally expensive tree priors are non-viable. This is not necessarily problematic, when the data are very informative, as this will reduce the influence of the tree prior.
However, the assumption that tree prior choice is of low importance must first be verified.

There have been multiple attempts to investigate the importance of tree
prior choice. For example, recently Sarver et al., [\cite{sarver2019choice}] 
showed that the choice of tree prior does not 
substantially affect phylogenetic inferences of diversification rates.
Also recently, Duchene et al. [\cite{duchene2018phylodynamic}] released
a BEAST2 package to assess how well posterior predictive simulations
recover a given tree when using the standard diversification models.
These studies show how current diversification models compare
to one another, but they do not help to assess the importance of a new tree prior.

Here we introduce a method to quantify the importance of a novel tree prior.
The method starts with a phylogeny generated by the new model for the tree prior. Next, nucleotide sequences are simulated on this phylogeny. Then, using the tree priors built-in into BEAST2, a Bayesian posterior distribution of phylogenies is inferred. We can then compare the inferred and simulated phylogenies. How to properly perform this comparison forms the heart of our method.
Only new diversification models that result in a large discrepancy between inferred and simulated phylogeny will be worth the effort and computational burden 
to implement a species tree prior for in a Bayesian framework.

Our method is programmed as an R package called \verb;pirouette;.
\verb;pirouette; is built on \verb;babette; [\cite{bilderbeek2018babette}], 
which calls BEAST2 [\cite{bouckaert2014beast}].

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Description}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\verb;pirouette; is written in the R programming language (\cite{R}).
The goal of \verb;pirouette; is to quantify the importance of a tree prior.
It does so by measuring the inference error made for a given reconstructed phylogeny, 
simulated under a (usually novel) diversification model.
We refer to the true prior that generated the given tree
as the 'generative tree prior' $\mathit{p_{G}}$.
Many tree priors have a parameter setting in which these are equivalent
to a standard tree prior. For example, a protracted birth-death model reduces to a standard birth-death model when speciation-completion rate approaches infinity [i.e. rate $\lambda$ in \cite{etienne2014estimating}, eqs. (2b) and (2c)].
When benchmarking a novel tree prior, 
one will typically construct phylogenies 
for different combinations of the diversification model's parameters, 
to assess under which scenarios the inference error cannot be neglected. 
While we recommend many replicate simulations when assessing a novel tree prior, our examples contain only one replicate as they are for illustrative purposes only.

\verb;pirouette; is very flexible and allows the user 
to specify a wide variety of custom settings. 
These settings can be grouped in macro-sections, 
according to how they operate in the pipeline. 
We summarize them in tables~\ref{tab:options} and~\ref{tab:definitions}.

Although many possible tests can be performed, 
we present here what we think are the five most important 
example research questions (ERQ) that \verb;pirouette; can answer. 
These ERQs gradually build up in complexity.
For each of these ERQs we will quantify the importance of a tree prior.

\begin{sidewaystable}
\centering
  \begin{tabular}{|p{3.2cm}|p{9.7cm}|p{4.5cm}@{}|}
    \hline
    \centering
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Sub-argument} & 
    \textbf{Description} &
    \textbf{Possible values} \\ 
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;tree_prior; &
    Macroevolutionary diversification model &
    Yule, BD, CBS, CCP, CEP \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;clock_model; &
    Clock for the DNA mutation rates &
    strict, relaxed log-normal \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;site_model; &
    Nucleotide substitution model &
    JC, HKY, TN93, GTR \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;mutation_rate; &
    Pace at which mutation occurs &
    mutation rate $\in \mathbb{R}_{>0}$\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;root_sequence; &
    DNA sequence at the root of the tree &
    any combination of a, c, g, t \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;model_type; &
    Criterion to select an inference model &
    Generative, Candidate \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;run_if; &
    Condition under which an inference model is used &
    Always, Best candidate \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;do_measure_evidence; &
    Sets whether or not the evidence of the model is to be computed &
    TRUE, FALSE \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;error_function; &
    Specifies how to measure the error &
    nLTT, $|\gamma|$ \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;burn\_in\_fraction; &
    Specifies the percentage of initial posterior trees to discard &
    \verb;burn_in_fraction; $\in [0, 1]$\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
  \end{tabular}
  \caption{
    Most important parameter options.
    Yule = pure birth model (\cite{yule})
    BD = birth death (\cite{nee1994reconstructed}), 
    CBS = coalescent Bayesian skyline (\cite{drummond2005bayesian}), 
    CCP = coalescent constant-population, 
    CEP = coalescent exponential-population,
    JC = Jukes and Cantor (\cite{jukes1969evolution}), 
    HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
    TN93 = Tamura and Nei (\cite{tamura1993estimation}), 
    GTR = Generalized time-reversible model (\cite{tavare1986some}).
  }
  \label{tab:options}
\bigskip

  \begin{tabular}{|@{}c|p{4cm}|p{12cm}|}
    \hline
    \centering
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Symbol} &
    \textbf{Macro-argument} &
    \textbf{Description} \\
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{G}$ &
    Generative model &
    The full setting to produce BEAST2 input data. 
    Its core features are the tree prior $\mathit{p_{G}}$, the clock 
    model $\mathit{c_{G}}$ and the site model $\mathit{s_{G}}$. \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
    $\mathit{A}$ &
    Alignment model &
    It includes the parts of the generative model that directly affect the 
    alignment generation, like the clock model $\mathit{c_{G}}$ and 
    the site model $\mathit{s_{G}}$. Additional arguments can be provided, 
    such as the mutation rate and the root sequence. \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{X_{i}}$ &
    $i$-th candidate experiment &
    Full setting for a Bayesian inference. It is made by a 
    candidate inference model $\mathit{I_{i}}$ and its 
    inference conditions $\mathit{C_{i}}$. \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{I}$ &
    Inference model &
    Phylogenetic inference model to run BEAST2. Likewise the generative model $G$, 
    its main components are the tree prior $\mathit{p_{I}}$, the 
    clock model $\mathit{c_{I}}$ and the site model $\mathit{s_{I}}$. \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{C}$ & Inference conditions & Conditions under which $\mathit{I}$ 
    is used in the inference. 
    They are composed by the model type, run condition and 
    wether to measure the evidence. \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{E}$ & Error measure parameters & 
    Errors measurement setup that can be specified providing an 
    error function to measure the difference between the original phylogeny 
    and the inferred posterior. The initial part of the posterior that is 
    reckoned as not representative can be discarded using a burn-in fraction. \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \iffalse
    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & 
      $\mathit{\tau_{p}}$ \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \fi
    \hline 
  \end{tabular}
  \caption{
    Definitions of terms and relative symbols used in the main text and in 
    Fig~\ref{fig:pipeline}. To run the pipeline $\mathit{A}$, $\mathit{X}$ 
    and $\mathit{E}$ must be specified. Examples can be found in 
    listings~\ref{lst:create_alignment}, 
    \ref{lst:create_gen_experiment_explicit} and 
    \ref{lst:create_error_measure_params}.
  }
  \label{tab:definitions}
\end{sidewaystable}

\subsection{pirouette's pipeline}
\label{subsec:pipeline}

\begin{figure}
  \centering
  \includegraphics[width = \textwidth]{workflow3.png}
  \caption{
    \texttt{pirouette} pipeline.
    The pipeline starts from a phylogeny (1a) simulated by the generative tree prior 
    $\mathit{p_{G}}$.
    The phylogeny is converted to an alignment (2a) using the generative alignment model 
    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed by a clock and a site model. 
    The user supplies one or more experiments.
    For each candidate experiment $\mathit{X_{i}}$ 
    (a tuple of inference model $\mathit{I_{i}}$ and condition $\mathit{C_{i}}$),
    if its condition $\mathit{C_{i}}$ is 
    satisfied (which can depend on the alignment), 
    the corresponding inference model $\mathit{I} = \mathit{I_{i}}$ is selected 
    to be used in the next step.
    The inference models (3) of the selected experiments use the alignment (2a) 
    to each create a Bayesian posterior of (parameter estimates and) 
    phylogenies (4a). 
    Each of the posteriors' trees is compared to the true phylogeny (1a) 
    using error measure $\mathit{E}$, 
    resulting in an error distribution (5a). 
    Optionally, for each selected inference model a twin pipeline can be run.
    A twin phylogeny (1b) can be generated from the original 
    phylogeny (1a) using the tree prior $\mathit{p_{I}}$ of the selected inference model (3).
    Analogously, a twin alignment is simulated from the twin phylogeny along with clock model $\mathit{c_{I}}$ and site model $\mathit{s_{I}}$ imported from the selected inference model.
    The twin pipeline mimics the procedure of the main pipeline, resulting in a twin error distribution (5b).
  }
  \label{fig:pipeline}
\end{figure}

The pipeline is summarized by the following steps:
\begin{enumerate}
  \item from a given phylogeny an alignment is simulated 
    under a known alignment model $\mathit{A}$;
  \item from this alignment, according to the specified inference conditions $\mathit{C}$, 
    an inference model $\mathit{I}$ is picked (which may differ from the 
    generative model);
  \item the inference model and the alignment are used 
    to infer a posterior distribution of phylogenies;
  \item the phylogenies in the posterior are compared with the given phylogeny 
    to estimate the error made, according to the error measure $\mathit{E}$ specified 
    by the user;
\end{enumerate}
The pipeline is visualized in Fig.~\ref{fig:pipeline}. 
There is also the option to generate a 'twin tree', 
that goes through the same pipeline. 
The utility of this twin tree will be explained below.

The first step simulates a DNA alignment from a given 
phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a)
using the DNA alignment parameters.
The DNA alignment parameters consist of a (DNA) root sequence, 
a (DNA) mutation rate, a clock model and a nucleotide substitution
model, which will be referred to as site model.
The root sequence is the DNA sequence of the shared common ancestor,
and is set to four equally-sized mononucleotide blocks by default, as this
helps interpreting the resulting alignment.
Supported site models are JC, HKY, TN and GTR. Only the strict
clock model is supported in this step.

The second step (Fig.~\ref{fig:pipeline}, 3)
selects one or more inference models $I$ from a set of inference 
models $I_{1},\dots,I_{n}$. 
We define an experiment $X_{i}$ as the combination of 
an inference model $I_{i}$ and the conditions $C_{i}$ 
to actually use it in the inference step.
For example, we can express that an inference
model (a combination of a tree prior, clock model and site model) 
is known to include the generative/true tree prior and that 
we definitely want to use that model in our inference. 
As a second example, we can also express that we've submitted a set of 
candidate inference models,
of which only the best should be used in the actual inference.
In the first example, we specified the condition $C_{i}$ that this
generative model should always be run, where in the second example,
we specified condition $C_{i}$ that a candidate model should only be run
when being the best.
A 'best' model is defined as the inference model with
the highest evidence (a.k.a. marginal likelihood), given the alignment 
simulated in the previous step.
The evidence of an inference model is estimated by nested 
sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. 
Scripted use of BEAST2 packages is only possible under Linux and Mac.
Windows systems can do the model comparison for shorter DNA sequences 
using the web interface of \verb;mcbette; [\cite{mcbette}].

The third step infers the posteriors,
using the simulated alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 4a),
and those inference models that were selected in the previous step (3). 
For each selected experiment a posterior is inferred, using the 
\verb;babette; [\cite{bilderbeek2018babette}] R package which makes use of BEAST2. 
This step usually takes up most of the full pipeline's computation time.

The fourth step quantifies the inference error made. First the 
burn-in fraction is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
in which it samples an unrepresentative part of state space. By default, \verb;pirouette; 
removes the first 10\% of the posterior.
From the remaining posterior \verb;pirouette; 
creates an error distribution, by measuring the difference
between the true tree and each of the posterior 
trees (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
The default way to quantify the difference between two phylogenies
is the nLTT statistic (\cite{janzen2015approximate}), but any 
user-defined error statistic can be used.

\subsection{Twin tree}\label{subsec:twinning}

An optional step is to generate a 'twin tree' $\tau_{I}$
(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
that will be analyzed in the same way as the true tree.
A phylogeny $\tau$ is the combination of 
branching times $\Vec{t}$ and topology $\psi$.
Assume that a non-standard generative prior $\mathit{p_{G}}$
produces a tree $\tau_{\mathit{G}}$,
with branching times $\Vec{t}_{\mathit{G}}$ and 
topology $\psi_{\mathit{G}}$.
The twinning process T, in accordance with a chosen inference 
model $\mathit{I}$, creates a tree $\tau_{\mathit{I}}$
with branching times $\Vec{t}_{\mathit{I}}$ while preserving the 
topology $\psi_{\mathit{G}}$:

\begin{align}
  \tau_{\mathit{G}} = (\Vec{t}_{\mathit{G}}, \psi_{\mathit{G}}) 
  \xrightarrow[]{\mathit{T}} 
  \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, \psi_{\mathit{G}})
\end{align}

To achieve this, we find the parameters $\theta^{*}_{\mathit{I}}$ 
(e.g. speciation and extinction rates, in case of a birth-death model) 
that maximize the likelihood $L_{\mathit{I}}$ applied 
to the true tree, conditioned on its number of tips $n_{\mathit{G}}$,
where $L_{\mathit{I}}$ is the likelihood function  
associated with the chosen inference model $\mathit{I}$:

\begin{align}
    max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{G}}, n_{\mathit{G}})] 
\rightarrow \theta^{*}_{\mathit{I}}.
\end{align}

We use $\theta^{*}_{\mathit{I}}$ to simulate an equal number 
$n_{\mathit{I}} = n_{\mathit{G}}$ 
of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree 
$\tau_{\mathit{I}}$, under the process $\mathit{I}$, 
while preserving the same topology.

Using a twin tree serves as a control: both $\tau_{\mathit{G}}$ and 
$\tau_{\mathit{I}}$ will produce an error distribution. 
These errors unavoidably accumulate across the \verb;pirouette; pipeline,
for both the true and twin tree, due to several possible reasons: either tree may be an unlikely representative
of the underlying tree prior, the alignment simulation may give different
numbers of mutations for the same evolutionary time, 
and the MCMC will sample different states.
These errors can be equalized by imposing the same tree topology and by
replicating the pipelines a sufficient number of times.

The left-over differences in error distributions will be 
consequences of different phylogenies and tree priors for the two parallel 
pipelines: the main pipeline uses the tree $\tau_{\mathit{G}}$ and a
possibly mismatched prior $\mathit{G}$, where the twin pipeline
uses the tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
So, for the twin route, if the process is repeated a sufficient number of times, the resulting error distribution cannot be 
attributed to a mismatch between the true tree prior and the tree prior used in
inference.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Installation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\verb;pirouette; will be made available on CRAN from which it can then be easily installed easily from CRAN:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
install.packages("pirouette")
\end{lstlisting}

Until it is on CRAN, and for the most up-to-date version, 
one can download and install the package from \verb;pirouette;'s GitHub 
repository:

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single
]
usethis::install_github("richelbilderbeek/pirouette")
\end{lstlisting}

To start using \verb;pirouette;, load its functions in the global namespace 
first:

\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
library(pirouette)
\end{lstlisting}
Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
This can be done from within R, using:

\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
install_beast2()
\end{lstlisting}
For the option to select the best candidate model,
\verb;pirouette; needs the "NS" BEAST2 package [\cite{maturana2018model}].
It can be installed from within R, using:

\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
install_beast2_pkg("NS")
\end{lstlisting}

An overview of \verb;pirouette;'s main functions is shown in 
table~\ref{tab:functions}. 
Their usage is demonstrated in the example code below.
All \verb;pirouette;'s functions are documented,
have a useful example and sensible defaults.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \begin{tabular}{ | l | l | }
    \hline
    \textbf{Name} & \textbf{Description} \\
    \hline
    \verb;pir_run; & Run \verb;pirouette; \\
    \verb;pir_plot; & Show the \verb;pirouette; results as a plot \\
    \verb;create_pir_params; & Create the \verb;pirouette; parameters \\
    \hline
    \verb;create_alignment_params; & Create the alignment parameters \\
    \verb;create_twinning_params; & Create the twinning parameters \\
    \verb;create_experiment; & Create one experiment \\
    \verb;create_error_measure_params; & Create the error measurement parameters \\
    \hline
  \end{tabular}
  \caption{pirouette's main functions.}
  \label{tab:functions}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: Example Research Question 1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's 
inference,
by measuring the difference between a given tree and 
phylogenies in a posterior. Due to stochasticity, all posterior trees
will differ from the given phylogeny $\tau_{\mathit{G}}$, 
even when the generative tree prior $\mathit{p_{G}}$ 
and alignment model $\mathit{A}$ are the same as 
those in the inference model $\mathit{I}$.
Measuring this difference allows us to know the baseline error
of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' 
all the tree priors that are available to be used within an inference model 
(see Table~\ref{tab:options}).

Now we can formulate the first example research question that \verb;pirouette; 
can answer.

\subsection{Question}

"What is the inference error made on phylogenies
created by a standard diversification model?"

\subsection{Answer}

In this example we use a standard generative tree prior $\mathit{p_{G}^0}$, 
namely the Yule (pure-birth) tree prior. 
We choose to use a small tree with six taxa to keep
the calculations short and the figure more readable.
We pick a crown age of ten time units. This value is 
completely arbitrary, but it ties in with the mutation rate 
used in simulating an alignment in the next step.

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single, 
    label = {lst:create_yule_tree}, 
    caption = {
      Create a Yule tree. 
      The resulting tree is shown in figure~\ref{fig:yule_tree}.
    }
  ]
phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
\end{lstlisting}

\begin{figure}[ht]
  \includegraphics[width=\textwidth]{example_1/true_tree.png}
  \caption{The Yule tree, as created by listing~\ref{lst:create_yule_tree}.}
  \label{fig:yule_tree}
\end{figure}

The first step in \verb;pirouette; is to simulate a DNA alignment from the 
given phylogeny, as described in subsection~\ref{subsec:pipeline}.
In this example, the root sequence consists in four blocks of 250 
mononucleotides each, 
where the per-nucleotide mutation rate is 0.1 mutations per unit time.
We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
and a strict clock model as these are the simplest.
A JC site model assumes that mutation rates between nucleotides are equal and constant. 
A strict clock model assumes that the mutation rates 
of all species are equal and constant.

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single,
    label = {lst:create_alignment}, 
    caption = {Create an alignment.}
  ]
alignment_params <- create_alignment_params(
  clock_model = create_strict_clock_model(),
  site_model = create_jc69_site_model(),
  mutation_rate = 0.1,
  root_sequence = create_blocked_dna(length = 1000)
)
\end{lstlisting}

As the site and clock models used here are also the defaults, 
these function arguments can be safely omitted: we just show these
explicitly for clarity.

In the second step we state our experiment.
We define an experiment $\mathit{X}$ as a combination of an inference model 
$\mathit{I}$
and conditions $\mathit{C}$.
In this example we pick $\mathit{I}$ to be the same inference model as the 
generative one,
which is the chosen Yule tree prior $\mathit{p_{G}^0}$, as well as site and clock 
models defined in $\mathit{A}$, respectively JC and strict.
We specify in $\mathit{C}$ that the experiment will always be run.

\begin{table}
  \begin{tabular}{ | c | c | c | l | }
    \hline
    \textbf{model\_type} &
    \textbf{run\_if} &
    \textbf{do\_measure\_evidence} & 
    \textbf{inference model} \\ 
    \hline
    generative &
    always &
    FALSE &
    JC, strict, Yule \\
    \hline
  \end{tabular}
  \caption{
    Inference conditions and inference model for ERQ 1.
    JC: Jukes-Cantor site model.
    strict: strict clock model.
    Yule: Yule (pure-birth) tree prior.
  }
  \label{tab:RQ1}
\end{table}

Listing~\ref{lst:create_gen_experiment_explicit} shows how to
set up this experiment:

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_gen_experiment_explicit},
  caption = {
    Create the experiments to answer research question 1, 
    using explicit arguments.
  }
]
experiment <- create_experiment(
  inference_conditions = create_inference_conditions(
    model_type = "generative", 
    run_if = "always"
  ), 
  inference_model = create_inference_model(
    tree_prior = create_yule_tree_prior(),
    clock_model = create_strict_clock_model(), 
    site_model = create_jc69_site_model()
  )
)
\end{lstlisting}

Experiments must be bundled in a list to work, even if only one is provided, as 
in this case:

\begin{lstlisting}[
  language = R, 
  floatplacement = ht,
  frame = single,
  label = {lst:create_gen_experiment},
  caption = {
    Create the experiments to answer research question 1.
  }
]
experiments <- list(experiment)
\end{lstlisting}

The objects we just created must be bundled
in a \verb;pirouette; parameters structure,
together with error measurement parameters $\mathit{E}$ and
twinning parameters $\mathit{T}$. \giovanni{I suggest to refer to the twinning parameters only when strictly required and with the minimum possible amount of info provided, as we set it in such a way that the tree prior is imported from the inference model in an automatic way.}

We use the default error measurement
parameters, which uses a burn-in fraction of 10\% and the nLTT statistic to
measure the difference between phylogenies. For clarity,
we create this setup explicitly here:

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_error_measure_params},
  caption = Calling \texttt{create\_error\_measure\_params}.
]
error_measure_params <- create_error_measure_params(
  error_function = get_nltt_error_function(),
  burn_in_fraction = 0.1
)
\end{lstlisting}

We will not use twinning here yet.  
All these arguments are bundled
and checked by \verb;create_pir_params;:

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_pir_params},
  caption = Calling \texttt{create\_pir\_params}.
]
pir_params <- create_pir_params(
  alignment_params = alignment_params,
  experiments = experiments,
  error_measure_params = error_measure_params,
  twinning_params = NA
)
\end{lstlisting}

Now we can use our Yule tree and \verb;pir_params; to measure 
the inference error made on phylogenies
created by a standard diversification model:

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:pir_run},
  caption = Calling \texttt{pir\_run}.
]
errors <- pir_run(
  phylogeny = phylogeny,
  pir_params = pir_params
)
\end{lstlisting}

The error distribution can be plotted directly using \verb;pir_plot;:

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:pir_plot},
  caption = Calling \texttt{pir\_plot}.
]
pir_plot(errors)
\end{lstlisting}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_1/errors.png}
  \caption{
    ERQ 1: the inference error made from a phylogeny 
    when the generative and inference model are the same.
    The vertical dashed line indicates the median error.
  }
  \label{fig:example_1}
\end{figure}

The resulting error distribution, as shown figure~\ref{fig:example_1},
shows the inference error 
when $\mathit{I}$ matches with the generative model given 
by $\mathit{A}$ and $\mathit{p_{G}^0}$.
This error distribution can serve as a control,
as it is obtained from a tree of which the tree prior is standard and known.
Obtaining this control is an inherent part of \verb;pirouette;. 
We call it 'twinning' and we use it from ERQ 3.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: Example Research Question 2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the previous example we selected the inference model $\mathit{I}$ to match 
with a known generative tree prior $\mathit{p_{G}^0}$.
However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, 
thus cannot be part of a standard inference 
model (i.e. an inference model using a standard tree prior).
In such a case, the user has to somehow select 
a standard tree prior to be 
used in the inference.
There are two ways to deal with this problem: hand-picking a tree
prior, or let the best of a set of models be picked. 
Hand-picking a tree prior can be rational if one standard tree prior 
is more suitable than others for the inference, for example, when
the novel tree prior is most similar to a certain tree prior.
Instead of hand-picking, the best inference model from a list of candidate 
models can be used in inference.
\verb;pirouette; selects the 'best' model, that is, 
the model with the highest evidence (also known as marginal likelihood).

We show the procedure using \verb;pirouette; to answer the second research 
question.

\subsection{Question}

"What is the inference error made on a novel phylogeny when
using the best inference model?"

\subsection{Answer}

This time we start with a tree generated from an unknown 
diversification model, that has six taxa and a crown age of ten:

\begin{lstlisting}[
  language = R, 
  floatplacement = ht,
  frame = single, 
  label = {lst:unknown_phylogeny},
  caption = A phylogeny generated by an unknown diversification model.
]
phylogeny  <- ape::read.tree(
  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
)
\end{lstlisting}
\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_2/tree_unknown.png}
  \caption{The tree derived from an unknown diversification process, 
    as created by listing~\ref{lst:unknown_phylogeny}.
  }
\end{figure}

The first step is to simulate a DNA alignment from the 
given phylogeny. 
We will re-use the alignment parameters of the previous example 
as shown in listing~\ref{lst:create_alignment}.

\begin{table}
  \begin{tabular}{ | c | c | c | l | }
    \hline
    \textbf{model\_type} &
    \textbf{run\_if} & \textbf{do\_measure\_evidence} & 
    \textbf{inference model} \\ 
    \hline
    candidate & best candidate & TRUE & JC, strict, Yule \\
    candidate & best candidate & TRUE & JC, strict, BD   \\
    ...       & ...            & ... & ...              \\
    candidate & best candidate & TRUE & GTR, RLN, CCP    \\
    candidate & best candidate & TRUE & GTR, RLN, CEP    \\
    \hline
  \end{tabular}
  \caption{
    Inference conditions and inference model for ERQ 2.
    JC: Jukes-Cantor site model.
    strict: strict clock model.
    Yule: Yule (pure-birth) tree prior.
    BD: birth-death tree prior.
    GTR: GTR site model.
    RLN: relaxed log-normal clock model.
    CCP: coalescent constant-population tree prior.
    CEP: coalescent exponential-population tree prior.
  }
  \label{tab:RQ2}
\end{table}

In the second step, unlike ERQ 1, we provide a 
set of experiments $\{\mathit{X_{i}}\}_{i=1}^{40}$. All the experiments have 
the same inference conditions $\mathit{C_{i}} = \mathit{C}$ but different 
candidate inference models $\mathit{I_{i}}$. Only one of them will be used 
to run the actual inference. According to the specified $\mathit{C}$, 
\verb;pirouette; will measure the evidence for each of them and select only 
the one with the highest evidence as the best candidate.

As model selection is commonly performed on the full list of available 
candidate models, \verb;pirouette; has a dedicated function to provide it:
\verb;create_all_experiments; will create a full set of 40 experiments,
containing the inference models of all combinations of 
4 site models, 2 clock models and 5 tree priors.

\begin{lstlisting}[
  language = R, 
  floatplacement = ht, 
  frame = single, 
  label = {lst:create_all_experiments},
  caption = Create all 40 candidate experiments.
]
experiments <- create_all_experiments()
\end{lstlisting}

For all the other steps we use the default setups. 
We can create the complete
\verb;pirouette; parameter set like this (which is the
same as listing~\ref{lst:create_pir_params}, but using the defaults):

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_pir_params_defaults},
  caption = Create a \texttt{pir\_params} with many defaults.
]
pir_params <- create_pir_params(
  alignment_params = alignment_params,
  experiments = experiments
)
\end{lstlisting}

We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
and plot the results (listing~\ref{lst:pir_plot}),
as shown in figure~\ref{fig:example_2}.

\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_2/errors.png}
  \caption{
    ERQ 2: the inference error from
    when selecting the best inference model.
  }
  \label{fig:example_2}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: Example Research Question 3}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In ERQ 1 we measured the baseline error of the \verb;pirouette; method,
by using the same standard tree prior in both the generation of 
the original tree, as well as in the inference.
This baseline error distribution serves as a control,
to quantify the impact of a new (non-standard) tree prior.
To measure that impact, we compare the inference error 
from the new tree prior to the baseline error,
by using a twin tree (see subsection~\ref{subsec:twinning}).
In this ERQ, we hand-pick a standard tree prior $\mathit{p_{G}}$.

\subsection{Question}

"What is the inference error made from a phylogeny, 
when hand-picking a tree prior, compared to the background noise?"

\subsection{Answer}

Most of the settings for this ERQ are
the same as in ERQ 2: we will re-use the phylogeny from an 
unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
the alignment parameters (listing~\ref{lst:create_alignment}).

We create an experiment as in 
listing~\ref{lst:create_gen_experiment},
but with all defaults:

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_gen_experiment_defaults},
  caption = Create a list of one default generative experiment.
]
experiment <- create_gen_experiment()
\end{lstlisting}

Creating the complete set of experiments:

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_experiments},
  caption = Create a list of one experiment.
]
experiments <- list(experiment)
\end{lstlisting}

This time, however, we measure both the baseline and complete inference error.
We set the twinning parameters $\mathit{T}$ to create a twin tree,
as described in the 'Twinning' section \ref{subsec:twinning}.

Creating the twinning parameters is trivial with the default settings.
For clarity, however, we explicitly show the most
important arguments:

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_twinning_params},
  caption = Create the default twinning parameters.
]
twinning_params <- create_twinning_params(
  twin_model = "bd", 
  method = "random_tree"
)
\end{lstlisting}
\giovanni{We want to import the "twin\_model" from the inference model.}
\richel{Yes, that is an option, but according to Rampal not important enough
to add now}

We combine all the parameters using \verb;create_pir_params;:

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_pir_params_with_twinning},
  caption = Create the default twinning parameters.
]
pir_params <- create_pir_params(
  alignment_params = alignment_params,
  experiments = experiments,
  twinning_params = twinning_params
)
\end{lstlisting}

We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
and plot the results (listing~\ref{lst:pir_plot}),
The output is shown in figure~\ref{fig:example_3}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_3/errors.png}
  \caption{
    ERQ 3: the inference error made 
    when hand-picking a generative tree prior versus the control.
    Here, the 'twin' column shows the baseline inference error,
    as measured on the twin tree.
  }
  \label{fig:example_3}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: Example Research Question 4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

When no inference model can be hand-picked, 
one can select the best model from a set,
as shown in ERQ2.
The resulting error distribution contains the error accumulated in the pipeline,
without quantifying to what extent this error is due to the mismatch between 
the generative model $\mathit{G}$ and the selected inference model 
$\mathit{I}$.
As twinning (see section \ref{subsec:twinning}: 'Twinning') shows us the baseline error, we can quantify the contribution of the tree prior to the error.

\subsection{Question}

"What is the inference error from a phylogeny 
derived from an unknown speciation model,
when selecting the best inference model, 
in relation to the baseline error?"

\subsection{Answer}

Here, as in ERQs 2 and 3, we start with a phylogeny 
generated by an unknown speciation model (listing~\ref{lst:unknown_phylogeny}). 
The first step is to simulate a DNA alignment 
from the given phylogeny, for which we will use the same code 
as shown in listing~\ref{lst:create_alignment}.

We specify that we want all 40 inference models to compete 
by calling \verb;create_all_experiments;,
as we previously did in listing~\ref{lst:create_all_experiments}.
We choose again to use the default error measurement setup,
as shown in listing~\ref{lst:create_error_measure_params}.
To get an idea of a baseline error made by BEAST2, 
we also make use of the twinning option by using
the same call to \verb;create_pir_params; as shown in 
listing~\ref{lst:create_pir_params_with_twinning}.
We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
and plot the results (listing~\ref{lst:pir_plot}).
The output is shown in figure~\ref{fig:example_4}.

\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_4/errors.png}
  \caption{
    Example 4: the error made by BEAST2 from a phylogeny, when the generative 
    model is known, compared to candidates with similar inference models, in 
    relationship to the background error.
  }
  \label{fig:example_4}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage: Example Research Question 5}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

If we are moderately sure about hand-picking an inference model,
but would also like to check how well the hand-picked model compares
to the others, we can do both. Of course, as we are interested in the
baseline error for both the hand-picked and best candidate model,
we need to enable twinning as well. This example is a combination
of the previous two ERQs.

\subsection{Question}

"What is the error made by BEAST2 from a phylogeny, 
when the generative model is hand-picked, 
compared to candidates with similar inference models, 
in relation to the background error?"

\subsection{Answer}

In this example we use a tree generated by the Yule process.
Instead of using the common Yule tree as in ERQ 1, we will use
a Yule tree of low likelihood, 
i.e. the one already used in ERQs 2-4, 
by using the code shown in listing~\ref{lst:unknown_phylogeny}.
The first step is to simulate a DNA alignment 
from the given phylogeny, for which we will use the same code 
as shown in listing~\ref{lst:create_alignment}.

\begin{table}
  \begin{tabular}{ | c | c | c | l | }
    \hline
    \textbf{model\_type} & 
    \textbf{run\_if} & 
    \textbf{do\_measure\_evidence} & 
    \textbf{inference model} \\ 
    \hline
    generative & always         & TRUE & JC, strict, Yule \\
    candidate  & best candidate & TRUE & JC, strict, BD \\
    candidate  & best candidate & TRUE & JC, strict, CCP \\
    candidate  & best candidate & TRUE & JC, strict, CEP \\
    \hline
  \end{tabular}
  \caption{
    Experimental setup to answer the fourth research question.
    JC: Jukes-Cantor site model.
    strict: strict clock model.
    Yule: Yule (pure-birth) tree prior.
    BD: birth-death tree prior.
    CCP: coalescent constant-population tree prior.
    CEP: coalescent exponential-population tree prior.
  }
  \label{tab:experiment_5}
\end{table}

In the second step we state our experiments. 
In this example we must specify both the generative and candidate experiments. 
Here we specify the generative model:

\begin{lstlisting}[language=R, floatplacement=H, frame=single]
generative_experiment <- create_gen_experiment()
\end{lstlisting}

We can specify all the candidates simply by calling 
\verb;create_all_experiments; and excluding the generative inference model,
resulting in 39 candidate experiments:

\begin{lstlisting}[
    language = R,
    floatplacement = H,
    frame = single
]
candidate_experiments <- create_all_experiments(
  exclude_model = generative_experiment$inference_model
)
\end{lstlisting}

We need to combine these experiments into one set of experiments:

\begin{lstlisting}[
    language = R,
    floatplacement = H,
    frame = single
]
experiments <- c(list(generative_experiment), candidate_experiments)
\end{lstlisting}

We choose again to use the default error measurement parameters
created in listing~\ref{lst:create_error_measure_params}.
We enable twinning, by calling \verb;create_pir_params;
as shown in listing~\ref{lst:create_pir_params_with_twinning}.

We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
and plot the results (listing~\ref{lst:pir_plot}).
The output is shown in figure~\ref{fig:example_5}.

\begin{figure}[H]
  \includegraphics[width=\textwidth]{example_5/errors.png}
  \caption{
    Example 5: the inference error made, 
    when the generative model is hand-picked, 
    compared to the best of 39 candidate models, 
    each in relationship to the background error.
  }
  \label{fig:example_5}
\end{figure}

\begin{table}
  \begin{tabular}{ | r | c | c | c | }
    \hline
    \textbf{Compare with} & \textbf{TB} & \textbf{WG} & \textbf{WB} \\ 
    \hline
    \textbf{TG} & Y & Y & N \\
    \textbf{TB} & . & N & Y \\
    \textbf{WG} & . & . & Y \\
    \textbf{WB} & . & . & . \\
    \hline
  \end{tabular}
  \caption{
    Error distributions that are relevant to compare.
    TG: true tree, generative model.
    TB: true tree, best candidate model.
    WG: twin tree, generative model.
    WB: twin tree, best candidate model.
  }
  \label{tab:relevant_comparisions}
\end{table}

\begin{table}
  \begin{tabular}{ | r | c | p{9cm} | }
    \hline
    \textbf{Condition} & \textbf{Expectation} & \textbf{Interpretation} \\ 
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$TG > TB$}       & Unexpected & Novel tree prior is more related to 
      best candidate than hand-picked tree prior \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$TG \approx TB$} & Possible   & Hand-picked tree prior is just as 
      suitable as the best candidate tree prior \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$TG < TB$}       & Expected   & Hand-picked tree prior is the most 
      related tree prior \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$TG > WG$}       & Unknown    & Novel tree prior important \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$TG \approx WG$} & Unknown    & Novel tree prior unimportant \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$TG < WG$}       & Unexpected & Twinning procedure increases 
      inference errors when using hand-picked tree prior \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$TB > WB$}       & Expected   & Impact of novel tree prior cannot 
      be compensated for by model selection: twin tree with low likelihood? \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$TB \approx WB$} & Possible   & Best candidate tree priors perform 
      equally well in true and twin tree: true and twin tree similar? \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$TB < WB$}       & Unexpected & Twinning procedure increases 
      inference errors  when using best tree prior candidate \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$WG > WB$}       & Unexpected & Hand-picked tree prior (that equals 
      the twinning tree prior!) worse than best candidate tree prior \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$WG \approx WB$} & Possible   & Twin tree fits equally well to the 
      hand-picked and best candidate tree prior  \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{$WG < WB$}       & Expected   & Hand-pick tree prior (that equals 
      the twinning tree prior) performs as expected \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
  \end{tabular}
  \caption{
    Interpretation of the different error distributions.
    This assumes that the operators ($<$, $\approx$ and $>$) to compare
    distributions are defined.
    TG: true tree, generative model.
    TB: true tree, best candidate model.
    WG: true tree, generative model.
    WB: true tree, best candidate model.
  }
  \label{tab:interpretation}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% RJCB will add: in six percent did the twin do worse, 
% see https://github.com/richelbilderbeek/pirouette_article/issues/41

In the five example research questions, 
we showed how to use \verb;pirouette; 
to quantify the importance of a tree prior in Bayesian phylogenetics.
In all ERQs we used and assumed the most simple tree prior possible,
which is the Yule tree prior.
In principle any other (more complex) generative tree prior 
can be tested, but we chose to provide the 
simplest (and fastest to run) examples.
All tests were performed on a single original tree, 
hence not providing us enough statistical power to 
evaluate the level of accuracy of the inference.

However, if the same procedure were repeated and performed on a distribution 
with a sufficient number of generative trees, 
it could actually constitute a quantitative assessment 
of the quality of the inference.
To give an idea of the insights \verb;pirouette; can give,
we pretend the results of the example research questions
to be obtained from thousands of replicates.
In that case, figure~\ref{fig:example_1} would show 
the minimum error BEAST2 when the
generative model is known, which can serve as 
a minimum noise level assessment in a bigger experiment.  
Figure~\ref{fig:example_2} shows that 
the best candidate model gives an error distribution similar to the unknown 
tree prior,
hinting that the candidate model would be just as good in inference.
Figure~\ref{fig:example_3} shows that 
the twin model gives an error distribution similar to the unknown tree prior,
hinting that the shape of the phylogeny with unknown tree prior
does not effect the error made by BEAST2.
Figure~\ref{fig:example_4} shows that 
the twin model and best candidate model give 
an error distribution similar to the unknown tree prior,
hinting that the shape of the phylogeny with unknown tree prior
does not effect the error made by BEAST2, 
and that the candidate model would be just as good in inference.

In conclusion, \verb;pirouette; can directly show 
the comparison between the results obtained 
adopting a new generative prior as compared to the error BEAST2 
would make when operating on a tree generated 
under a known, right and existing tree prior (the twin tree).
From such results an user can estimate whether or not a new tree prior, 
tailored on the generative process, is needed. 
If this would indeed be the case, 
one could implement their own tree prior as a new module in BEAST2.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{pirouette resources}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\verb;pirouette; is free, libre and open source software available at 
\url{http://github.com/richelbilderbeek/pirouette},
licensed under the GNU General Public License version 3.
\verb;pirouette; depends on multiple packages, which are:
\verb;babette; (\cite{bilderbeek2018babette}),
\verb;becosys; (\cite{becosys}),
\verb;DDD; (\cite{DDD}),
\verb;devtools; (\cite{devtools}),
\verb;geiger; (\cite{geiger}),
\verb;ggplot2; (\cite{ggplot2}),
\verb;knitr; (\cite{knitr}),
\verb;lintr; (\cite{lintr}),
\verb;mcbette; (\cite{mcbette}),
\verb;nLTT; (\cite{nLTT}),
\verb;PBD; (\cite{PBD}),
\verb;phangorn; (\cite{phangorn}),
\verb;phytools; (\cite{phytools}),
\verb;rappdirs; (\cite{rappdirs}),
\verb;rmarkdown; (\cite{rmarkdown}),
\verb;Rmpfr; (\cite{Rmpfr}),
\verb;stringr; (\cite{stringr}),
\verb;subplex; (\cite{subplex}),
\verb;TESS; (\cite{TESS}),
\verb;testit; (\cite{testit}), 
\verb;testthat; (\cite{testthat}) and
\verb;tidyr; (\cite{tidyr}).

\verb;pirouette;'s development takes place on GitHub,
\url{https://github.com/richelbilderbeek/pirouette},
which allows submitting bug reports, requesting features, 
or adding code. To ensure a high quality, \verb;pirouette; 
uses a continuous integration service, has a code coverage of 100\%
and enforces the most commonly used R style guide (\cite{style_guide}).

\verb;pirouette;'s is extensively documented on its website,
its documentation and its vignettes.
The \verb;pirouette; website is a good starting point to learn
how to use \verb;pirouette;, as it links to tutorials and videos.
The \verb;pirouette; package documentation describes
all functions and liberally links to related functions.
All exported functions show a minimal example as part of their documentation.
The \verb;pirouette; vignette demonstrates extensively how 
to use \verb;pirouette; in a more lightly written way.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Citation of pirouette}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Professionals using \verb;pirouette; can cite this
article. To obtain the citation from within R, use:

\begin{lstlisting}[language=R]
> citation("pirouette")
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We would like to thank the Center for Information Technology of the University 
of Groningen for their support and for providing access to the Peregrine 
high performance computing cluster. 
We thank the Netherlands 
Organization for Scientific Research (NWO) for financial support 
through a VICI grant awarded to RSE.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Accessibility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

All code is archived at 
\url{http://github.com/richelbilderbeek/pirouette_article},
with DOI \url{https://doi.org/12.3456/zenodo.1234567}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Authors' contributions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

RJCB, GL and RSE conceived the idea for the package. 
RJCB created and tested the package, and wrote the first draft of the 
manuscript.
GL tested the package and contributed substantially to revisions.
RSE contributed to revisions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{mee}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-05-07 supplement.tex
Split up in article.tex (big picture), content.tex (the content) and (supplement.tex) supplementary materials

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results in detail}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Errors}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example 1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht]
  \centering
  \resizebox {0.8\columnwidth} {!} {
    \begin{tikzpicture}[
      ->,>=stealth',shorten >=1pt,auto,
      node distance=0.25\textheight, 
      semithick
    ]   
    \tikzstyle{every state}=[]
    \node[state, draw=none] (O) [] {
    };   
    \node[state] (A) [right of = O, rectangle] {
      \includegraphics[height=0.2\textheight]{example_1/true_tree.png}
    };   
    \node[state] (B) [below of = A, rectangle] {
      \includegraphics[height=0.2\textheight]{example_1/true_alignment.png}
    };   
    \node[state] (C) [below of = B, rectangle] {
      \includegraphics[height=0.2\textheight]{example_1/true_posterior_gen.png}
    };   
    \node[state] (D) [below of = C, rectangle] {
      \includegraphics[height=0.2\textheight]{example_1/true_error_violin_gen.png}
    };   
    \path 
      (O) edge [anchor = south] node {} (A)
      (A) edge [anchor = south] node {} (B)
      (B) edge [anchor = south] node {} (C)
      (C) edge [anchor = south] node {} (D)
    ; 
    \end{tikzpicture}
  }
  \label{fig:example_1_full_pipeline}
  \caption{Example 1: full pipeline}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{example_1/esses.latex}
% has label tab:esses_example_1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example 2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht]
  \centering
  \resizebox {0.8\columnwidth} {!} {
    \begin{tikzpicture}[
      ->,>=stealth',shorten >=1pt,auto,
      node distance=0.25\textheight, 
      semithick
    ]   
    \tikzstyle{every state}=[]
    \node[state, draw=none] (O) [] {
    };   
    \node[state] (A) [right of = O, rectangle] {
      \includegraphics[height=0.2\textheight]{example_2/true_tree.png}
    };   
    \node[state] (B) [below of = A, rectangle] {
      \includegraphics[height=0.2\textheight]{example_2/true_alignment.png}
    };   
    \node[state] (C) [below of = B, rectangle] {
      \includegraphics[height=0.2\textheight]{example_2/true_posterior_best.png}
    };   
    \node[state] (D) [below of = C, rectangle] {
      \includegraphics[height=0.2\textheight]{example_2/true_error_violin_best.png}
    };   
    \path 
      (O) edge [anchor = south] node {} (A)
      (A) edge [anchor = south] node {} (B)
      (B) edge [anchor = south] node {} (C)
      (C) edge [anchor = south] node {} (D)
    ; 
    \end{tikzpicture}
  }
  \label{fig:example_2_full_pipeline}
  \caption{Example 2: full pipeline}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{example_2/esses.latex}
% has label tab:esses_example_2

\input{example_2/evidences.latex}
% has label tab:evidences_example_2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example 3}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht]
  \centering
  \resizebox {0.8\columnwidth} {!} {
    \begin{tikzpicture}[
      ->,>=stealth',shorten >=1pt,auto,
      node distance=0.4\textheight, 
      semithick
    ]   
    \tikzstyle{every state}=[]
    \node[state, draw=none] (O) [] {
    };   
    \node[state] (A) [right of = O, rectangle] {
      \includegraphics[height=0.2\textheight]{example_3/true_tree.png}
    };   
    \node[state] (B) [below of = A, rectangle] {
      \includegraphics[height=0.13\textheight]{example_3/true_alignment.png}
    };   
    \node[state] (C) [below of = B, rectangle] {
      \includegraphics[height=0.2\textheight]{example_3/true_posterior_gen.png}
    };   
    \node[state] (D) [below of = C, rectangle] {
      \includegraphics[height=0.2\textheight]{example_3/true_error_violin_gen.png}
    };   
    \node[state] (AT) [right of = A, rectangle] {
      \includegraphics[height=0.2\textheight]{example_3/twin_tree.png}
    };   
    \node[state] (BT) [below of = AT, rectangle] {
      \includegraphics[height=0.13\textheight]{example_3/twin_alignment.png}
    };   
    \node[state] (CT) [below of = BT, rectangle] {
      \includegraphics[height=0.2\textheight]{example_3/twin_posterior_gen.png}
    };   
    \node[state] (DT) [below of = CT, rectangle] {
      \includegraphics[height=0.2\textheight]{example_3/twin_error_violin_gen.png}
    };   
    \path 
      (O) edge [anchor = south] node {} (A)
      (A) edge [anchor = south] node {} (B)
      (B) edge [anchor = south] node {} (C)
      (C) edge [anchor = south] node {} (D)
      (A) edge [anchor = east] node {} (AT)
      (AT) edge [anchor = south] node {} (BT)
      (BT) edge [anchor = south] node {} (CT)
      (CT) edge [anchor = south] node {} (DT)
    ; 
    \end{tikzpicture}
  }
  \label{fig:example_3_full_pipeline}
  \caption{Example 3: full pipeline}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{example_3/esses.latex}
% has label tab:esses_example_3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example 4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht]
  \centering
  \resizebox {1.0\columnwidth} {!} {
    \begin{tikzpicture}[
      ->,>=stealth',shorten >=1pt,auto,
      node distance=0.5\textheight, 
      semithick
    ]   
    \tikzstyle{every state}=[]
    \node[state, draw=none] (O) [] {
    };   
    \node[state] (A) [right of = O, rectangle] {
      \includegraphics[height=0.4\textheight]{example_4/true_tree.png}
    };   
    \node[state] (B) [below of = A, rectangle] {
      \includegraphics[height=0.45\textheight]{example_4/true_alignment.png}
    };   
    \node[state] (CB) [below of = B, rectangle, node distance=0.8\textheight] {

\includegraphics[height=0.95\textheight]{example_4/true_posterior_best.png}
    };   
    \node[state] (DB) [below of = CB, rectangle, node distance=1.0\textheight] {
      \includegraphics[height=0.65\textheight]{example_4/true_error_violin_best.png}
    };   
    \node[state] (AT) [right of = A, rectangle, node distance=2.0\textwidth] {
      \includegraphics[height=0.4\textheight]{example_4/twin_tree.png}
    };   
    \node[state] (BT) [below of = AT, rectangle] {
      \includegraphics[height=0.45\textheight]{example_4/twin_alignment.png}
    };   
    \node[state] (CTB) [below of = BT, rectangle, node distance=0.8\textheight] 
{

\includegraphics[height=0.95\textheight]{example_4/twin_posterior_best.png}
    };   
    \node[state] (DTB) [below of = CTB, rectangle, node 
distance=1.0\textheight] {

\includegraphics[height=0.65\textheight]{example_4/twin_error_violin_best.png}
    };   
    \path 
      (O) edge [anchor = south] node {} (A)
      (A) edge [anchor = south] node {} (B)
      (B) edge [anchor = south east] node {} (CB)
      (CB) edge [anchor = south] node {} (DB)
      (A) edge [anchor = east] node {} (AT)
      (AT) edge [anchor = south] node {} (BT)
      (BT) edge [anchor = south] node {} (CTB)
      (CTB) edge [anchor = south] node {} (DTB)
    ; 
    \end{tikzpicture}
  }
  \label{fig:example_4_full_pipeline}
  \caption{Example 4: full pipeline}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{example_4/esses_best.latex}
% has label tab:esses_example_4_best

\input{example_4/esses_twin_best.latex}
% has label tab:esses_example_4_twin_best

\input{example_4/evidence_true.latex}
% has label tab:evidences_example_4

\input{example_4/evidence_twin.latex}
% has label tab:evidences_example_4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example 5}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht]
  \centering
  \resizebox {0.8\columnwidth} {!} {
    \begin{tikzpicture}[
      ->,>=stealth',shorten >=1pt,auto,
      node distance=0.4\textheight, 
      semithick
    ]   
    \tikzstyle{every state}=[]
    \node[state, draw=none] (O) [] {
    };   
    \node[state] (A) [right of = O, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5/true_tree.png}
    };   
    \node[state] (B) [below of = A, rectangle] {
      \includegraphics[height=0.13\textheight]{example_5/true_alignment.png}
    };   
    \node[state] (CG) [below of = B, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5/true_posterior_gen.png}
    };   
    \node[state] (DG) [below of = CG, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5/true_error_violin_gen.png}
    };   
    \node[state] (CB) [right of = C, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5/true_posterior_best.png}
    };   
    \node[state] (DB) [below of = CB, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5/true_error_violin_best.png}
    };   
    \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
      \includegraphics[height=0.2\textheight]{example_5/twin_tree.png}
    };   
    \node[state] (BT) [below of = AT, rectangle] {
      \includegraphics[height=0.13\textheight]{example_5/twin_alignment.png}
    };   
    \node[state] (CTG) [below of = BT, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5/twin_posterior_gen.png}
    };   
    \node[state] (DTG) [below of = CTG, rectangle] {

\includegraphics[height=0.2\textheight]{example_5/twin_error_violin_gen.png}
    };   
    \node[state] (CTB) [right of = CTG, rectangle] {
      \includegraphics[height=0.2\textheight]{example_5/twin_posterior_best.png}
    };   
    \node[state] (DTB) [below of = CTB, rectangle] {

\includegraphics[height=0.2\textheight]{example_5/twin_error_violin_best.png}
    };   
    \path 
      (O) edge [anchor = south] node {} (A)
      (A) edge [anchor = south] node {} (B)
      (B) edge [anchor = south] node {} (CG)
      (CG) edge [anchor = south] node {} (DG)
      (B) edge [anchor = south east] node {} (CB)
      (CB) edge [anchor = south] node {} (DB)
      (A) edge [anchor = east] node {} (AT)
      (AT) edge [anchor = south] node {} (BT)
      (BT) edge [anchor = south east] node {} (CTG)
      (CTG) edge [anchor = south] node {} (DTG)
      (BT) edge [anchor = south] node {} (CTB)
      (CTB) edge [anchor = south] node {} (DTB)
    ; 
    \end{tikzpicture}
  }
  \label{fig:example_5_full_pipeline}
  \caption{Example 5: full pipeline}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{example_5/esses_gen.latex}
% has label tab:esses_example_5_gen

\input{example_5/esses_best.latex}
% has label tab:esses_example_5_best

\input{example_5/esses_twin_gen.latex}
% has label tab:esses_example_5_twin_gen

\input{example_5/esses_twin_best.latex}
% has label tab:esses_example_5_twin_best

\input{example_5/evidence_true.latex}
% has label tab:evidences_example_5

\input{example_5/evidence_twin.latex}
% has label tab:evidences_example_5

\begin{table}
  \begin{tabular}{ | r | c | }
    \hline
    \textbf{ERQ} & \textbf{run-time} \\ 
    \hline
      1 & 00:10:12 \\
      2 & 00:34:10 \\
      3 & 00:24:53 \\
      4 & 01:04:50 \\
      5 & 01:29:40 \\
    \hline
  \end{tabular}
  \caption{
    Time for an example research question to finish.
    Time is in in hh:mm::ss.
  }
  \label{tab:runtime}
\end{table}

### 2019-05-13 content.tex
Improve Table 3, use 'Table' and 'Figure' when referring to tables and figures

@@
 according to how they operate in the pipeline. 
-We summarize them in tables~\ref{tab:options} and~\ref{tab:definitions}.
+We summarize them in Table~\ref{tab:options} and Table~\ref{tab:definitions}.
 
@@
     and $\mathit{E}$ must be specified. Examples can be found in 
-    listings~\ref{lst:create_alignment}, 
+    listings~\ref{lst:create_alignment_params}, 
     \ref{lst:create_gen_experiment_explicit} and 
@@
 An overview of \verb;pirouette;'s main functions is shown in 
-table~\ref{tab:functions}. 
+Table~\ref{tab:functions}. 
 Their usage is demonstrated in the example code below.
@@
   \centering
-  \begin{tabular}{ | l | l | }
-    \hline
-    \textbf{Name} & \textbf{Description} \\
-    \hline
-    \verb;pir_run; & Run \verb;pirouette; \\
-    \verb;pir_plot; & Show the \verb;pirouette; results as a plot \\
-    \verb;create_pir_params; & Create the \verb;pirouette; parameters \\
-    \hline
-    \verb;create_alignment_params; & Create the alignment parameters \\
-    \verb;create_twinning_params; & Create the twinning parameters \\
-    \verb;create_experiment; & Create one experiment \\
-    \verb;create_error_measure_params; & Create the error measurement parameters \\
+  \begin{tabular}{ | l | l | l | }
+    \hline
+    \textbf{Name} & \textbf{Description} & \textbf{Listing} \\
+    \hline
+    \verb;pir_run; & Run \verb;pirouette; & \ref{lst:pir_run} \\
+    \verb;pir_plot; & Show the \verb;pirouette; results as a plot & \ref{lst:pir_plot} \\
+    \verb;create_pir_params; & Create the \verb;pirouette; parameters & \ref{lst:create_pir_params} \\
+    \hline
+    \verb;create_alignment_params; & Create the alignment parameters & \ref{lst:create_alignment_params} \\
+    \verb;create_twinning_params; & Create the twinning parameters & \ref{lst:create_twinning_params} \\
+    \verb;create_experiment; & Create one experiment & \ref{lst:create_gen_experiment_explicit} \\
+    \verb;create_error_measure_params; & Create the error measurement parameters & \ref{lst:create_error_measure_params} \\
     \hline
   \end{tabular}
-  \caption{pirouette's main functions.}
+  \caption{
+    \texttt{pirouette}'s main functions, description and the number of the 
+    listing in which it is used. 
+  }
   \label{tab:functions}
@@
       Create a Yule tree. 
-      The resulting tree is shown in figure~\ref{fig:yule_tree}.
+      The resulting tree is shown in Figure~\ref{fig:yule_tree}.
     }
@@
     frame = single,
-    label = {lst:create_alignment}, 
+    label = {lst:create_alignment_params}, 
     caption = {Create an alignment.}
@@
 
-The resulting error distribution, as shown figure~\ref{fig:example_1},
+The resulting error distribution, as shown Figure~\ref{fig:example_1},
 shows the inference error 
@@
 We will re-use the alignment parameters of the previous example 
-as shown in listing~\ref{lst:create_alignment}.
+as shown in listing~\ref{lst:create_alignment_params}.
 
@@
 and plot the results (listing~\ref{lst:pir_plot}),
-as shown in figure~\ref{fig:example_2}.
+as shown in Figure~\ref{fig:example_2}.
 
@@
 unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
-the alignment parameters (listing~\ref{lst:create_alignment}).
+the alignment parameters (listing~\ref{lst:create_alignment_params}).
 
@@
 and plot the results (listing~\ref{lst:pir_plot}),
-The output is shown in figure~\ref{fig:example_3}
+The output is shown in Figure~\ref{fig:example_3}
 
@@
 from the given phylogeny, for which we will use the same code 
-as shown in listing~\ref{lst:create_alignment}.
+as shown in listing~\ref{lst:create_alignment_params}.
 
@@
 and plot the results (listing~\ref{lst:pir_plot}).
-The output is shown in figure~\ref{fig:example_4}.
+The output is shown in Figure~\ref{fig:example_4}.
 
@@
 and plot the results (listing~\ref{lst:pir_plot}).
-The output is shown in figure~\ref{fig:example_5}.
+The output is shown in Figure~\ref{fig:example_5}.
 
@@
 to be obtained from thousands of replicates.
-In that case, figure~\ref{fig:example_1} would show 
+In that case, Figure~\ref{fig:example_1} would show 
 the minimum error BEAST2 when the

### 2019-05-13 content.tex
Updates from Overleaf

--- 2019-05-07 content.tex
@@
     This begs the question how accurate the tree estimation is
-    when the real macroevolutionary processes are different 
+    when the real macroevolutionary processes are substantially different 
     from those assumed in the tree prior. \\
@@
 or RevBayes [\cite{hohna2016revbayes}], 
-has been a major advance in building phylogenetic trees 
+has been a major advance in constructing phylogenetic trees 
 from character data (usually nucleotide sequences) extracted 
@@
 BEAST [\cite{drummond2007beast}] is a typical Bayesian phylogenetics tool, 
-that needs both character data and priors to infer a posterior.
-Specifically, the species tree prior describes the process of diversification.
-BEAST has built-in tree priors such as the Yule [\cite{yule}] and 
+that needs both character data and priors to infer a posterior distribution of phylogenies.
+Specifically, for the species tree prior - which describes the process of diversification - 
+BEAST has built-in priors such as the Yule [\cite{yule}] and 
 (constant-rate) birth-death [\cite{nee1994reconstructed}] models.
@@
 BEAST's successor, BEAST2 [\cite{bouckaert2014beast}],
-has a package manager, that allows third-party users to extend existing 
-functionalities.
-Hence one can add novel diversification models by writing a BEAST2 package that contains the likelihood formula
+has a package manager, that allows third-party users to extend existing functionalities.
+For example, one can add novel diversification models by writing a BEAST2 package that contains the likelihood formula
 of a phylogeny under the novel diversification model, i.e. the prior probability of a species tree.
-
-Many such diversification models (and their associated probability 
-algorithms) have been developed
-developed, e.g., those in which diversification is 
+Many such diversification models (and their associated probability algorithms) have been developed, e.g., models in which diversification is 
 time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
@@
 or a combination of all these [\cite{herrera2018detecting}].
-Only few of these diversification models are available as a BEAST2 package.
+Only a few of these diversification models are available as a BEAST2 package.
 
 When a novel diversification model is introduced,
-its performance in inference is commonly tested.
-Part of a model's performance is its ability to recover parameters from simulated data with known parameters (e.g. [\cite{etienne2014estimating}]),
-Ideally, the estimated parameter values closely match the known/true values.
+its performance in inference should be tested.
+Part of a model's performance is its ability to recover parameters from simulated data with known parameters (e.g.
+[\cite{etienne2014estimating}]), where ideally the estimated parameter values closely match the known/true values.
 
@@
 How to properly perform this comparison forms the heart of our method.
-Only new diversification models that result in a large discrepancy between inferred and simulated phylogeny will be worth the effort and computational burden 
+Only new diversification models that result in a large discrepancy between inferred and simulated phylogenies will be worth the effort and computational burden 
 to implement a species tree prior for in a Bayesian framework.
@@
 as the 'generative tree prior' $\mathit{p_{G}}$.
-Many tree priors have a parameter setting in which these are equivalent
-to a standard tree prior.
+Many tree priors have a parameter setting for which it reduces to a standard tree prior.
 For example, a protracted birth-death model reduces to a standard birth-death model when speciation-completion rate approaches infinity [i.e.
@@
 according to how they operate in the pipeline. 
-We summarize them in tables~\ref{tab:options} and~\ref{tab:definitions}.
+We summarize them in Tables~\ref{tab:options} and~\ref{tab:definitions}.
 
@@
 we present here what we think are the five most important 
-example research questions (ERQ) that \verb;pirouette; can answer. 
-These ERQs gradually build up in complexity.
+example research questions (ERQ) that \verb;pirouette; can answer, from simple questions to complex ones.
 For each of these ERQs we will quantify the importance of a tree prior.
@@
 
-The pipeline is summarized by the following steps:
+We assume the user has a phylogeny simulated with the new diversification model. The pipeline to assess the error BEAST2 makes in inferring this phylogeny then contains the following steps:
 \begin{enumerate}
-  \item from a given phylogeny an alignment is simulated 
+  \item from the given phylogeny an alignment is simulated 
     under a known alignment model $\mathit{A}$;
@@
 using the DNA alignment parameters.
-The DNA alignment parameters consist of a (DNA) root sequence, 
-a (DNA) mutation rate, a clock model and a nucleotide substitution
+The DNA alignment parameters consist of a (DNA) root sequence, a (DNA) mutation rate, a clock model and a nucleotide substitution
 model, which will be referred to as site model.
@@
 Supported site models are JC, HKY, TN and GTR. Only the strict
-clock model is supported in this step.
+clock model is currently supported in this step.
 
@@
 to actually use it in the inference step.
-For example, we can express that an inference
+For example, we may require that an inference
 model (a combination of a tree prior, clock model and site model) 
-is known to include the generative/true tree prior and that 
-we definitely want to use that model in our inference. 
-As a second example, we can also express that we've submitted a set of 
+should include the generative/true tree prior. 
+As a second example, we may require that we have selected a set of 
 candidate inference models,
@@
 In the first example, we specified the condition $C_{i}$ that this
-generative model should always be run, where in the second example,
+generative model should always be run, whereas in the second example,
 we specified condition $C_{i}$ that a candidate model should only be run
@@
 The evidence of an inference model is estimated by nested 
-sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. 
-Scripted use of BEAST2 packages is only possible under Linux and Mac.
+sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. We note that scripted use of BEAST2 packages is only possible under Linux and Mac.
 Windows systems can do the model comparison for shorter DNA sequences 
@@
 using the simulated alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 4a),
-and those inference models that were selected in the previous step (3). 
+and the inference models that were selected in the previous step (3). 
 For each selected experiment a posterior is inferred, using the 
@@
 
-The fourth step quantifies the inference error made. First the 
-burn-in fraction is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
+The fourth step quantifies the inference error made. First the burn-in fraction is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
 in which it samples an unrepresentative part of state space. By default, \verb;pirouette; 
@@
 that will be analyzed in the same way as the true tree.
-A phylogeny $\tau$ is the combination of 
+We define a phylogeny $\tau$ is the combination of 
 branching times $\Vec{t}$ and topology $\psi$.
-Assume that a non-standard generative prior $\mathit{p_{G}}$
+Assuming that a non-standard generative prior $\mathit{p_{G}}$
 produces a tree $\tau_{\mathit{G}}$,
 with branching times $\Vec{t}_{\mathit{G}}$ and 
-topology $\psi_{\mathit{G}}$.
-The twinning process T, in accordance with a chosen inference 
+topology $\psi_{\mathit{G}}$, the twinning process T, in accordance with a chosen inference 
 model $\mathit{I}$, creates a tree $\tau_{\mathit{I}}$
@@
 
-The left-over differences in error distributions will be 
+The remaining differences in error distributions will be 
 consequences of different phylogenies and tree priors for the two parallel 
@@
 uses the tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
-So, for the twin route, if the process is repeated a sufficient number of times, the resulting error distribution cannot be 
+So, for the twin trees, if the process is repeated a sufficient number of times, the resulting error distribution cannot be 
 attributed to a mismatch between the true tree prior and the tree prior used in
@@
 
-\verb;pirouette; will be made available on CRAN from which it can then be easily installed easily from CRAN:
+\verb;pirouette; will be made available on CRAN from which it can then be easily installed easily:
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]
@@
 \end{lstlisting}
-
-To start using \verb;pirouette;, load its functions in the global namespace 
-first:
-
+To start using \verb;pirouette;, load its functions in the global namespace first:
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]
@@
 This can be done from within R, using:
-
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]
@@
 It can be installed from within R, using:
-
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]
@@
 given phylogeny, as described in subsection~\ref{subsec:pipeline}.
-In this example, the root sequence consists in four blocks of 250 
-mononucleotides each, 
+In this example, the root sequence consists of four blocks of 250 mononucleotides each, 
 where the per-nucleotide mutation rate is 0.1 mutations per unit time.
@@
 A strict clock model assumes that the mutation rates 
-of all species are equal and constant.
+of all lineages are equal and constant.
 
@@
 As the site and clock models used here are also the defaults, 
-these function arguments can be safely omitted: we just show these
-explicitly for clarity.
+the function arguments can be safely omitted: we just show these explicitly for clarity.
 
@@
 and conditions $\mathit{C}$.
-In this example we pick $\mathit{I}$ to be the same inference model as the 
-generative one,
+In this example we pick $\mathit{I}$ to be the same inference model as the generative one,
 which is the chosen Yule tree prior $\mathit{p_{G}^0}$, as well as site and clock 
@@
   \caption{
-    ERQ 1: the inference error made from a phylogeny 
+    ERQ 1: the inference error made 
     when the generative and inference model are the same.
@@
 as it is obtained from a tree of which the tree prior is standard and known.
-Obtaining this control is an inherent part of \verb;pirouette;. 
+Obtaining this control is in fact an inherent part of \verb;pirouette;. 
 We call it 'twinning' and we use it from ERQ 3.
@@
 with a known generative tree prior $\mathit{p_{G}^0}$.
-However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, 
-thus cannot be part of a standard inference 
+However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, and hence 
+usually it is not part of a standard inference 
 model (i.e. an inference model using a standard tree prior).
-In such a case, the user has to somehow select 
-a standard tree prior to be 
+In such a case, the question is what standard tree prior should be 
 used in the inference.
 There are two ways to deal with this problem: hand-picking a tree
-prior, or let the best of a set of models be picked. 
+prior, or picking the best of a set of tree priors. 
 Hand-picking a tree prior can be rational if one standard tree prior 
 is more suitable than others for the inference, for example, when
-the novel tree prior is most similar to a certain tree prior.
+the novel tree prior is most similar to a specific standard tree prior.
 Instead of hand-picking, the best inference model from a list of candidate 
-models can be used in inference.
-\verb;pirouette; selects the 'best' model, that is, 
-the model with the highest evidence (also known as marginal likelihood).
-
-We show the procedure using \verb;pirouette; to answer the second research 
-question.
+models can also be used in inference. Selecting the 'best' model in
+\verb;pirouette; means selecting the model with the highest evidence (also known as marginal likelihood).
+
+We show the procedure using \verb;pirouette; to answer the second research question.
 
@@
 
-In the second step, unlike ERQ 1, we provide a 
+In the second step, we deviate from the answer to ERQ 1 by providing a 
 set of experiments $\{\mathit{X_{i}}\}_{i=1}^{40}$. All the experiments have 
@@
 candidate inference models $\mathit{I_{i}}$. Only one of them will be used 
-to run the actual inference. According to the specified $\mathit{C}$, 
+to run the actual full inference. According to the specified $\mathit{C}$, 
 \verb;pirouette; will measure the evidence for each of them and select only 
@@
 As model selection is commonly performed on the full list of available 
-candidate models, \verb;pirouette; has a dedicated function to provide it:
+candidate models, \verb;pirouette; has a dedicated function for this choice:
 \verb;create_all_experiments; will create a full set of 40 experiments,
@@
   \caption{
-    ERQ 2: the inference error from
+    ERQ 2: the inference error 
     when selecting the best inference model.
@@
 to add now}
+\rampal{did I say that? I think it's OK not to do it in the paper, but perhaps it can be noted.}
 
@@
     ERQ 3: the inference error made 
-    when hand-picking a generative tree prior versus the control.
-    Here, the 'twin' column shows the baseline inference error,
+    when hand-picking a generative tree prior compared with the error obtained for the twin tree.
+    Here, the 'twin' tree shows the baseline inference error,
     as measured on the twin tree.
@@
 
-When no inference model can be hand-picked, 
+Rather than hand-picking a tree prior, 
 one can select the best model from a set,
-as shown in ERQ2.
-The resulting error distribution contains the error accumulated in the pipeline,
-without quantifying to what extent this error is due to the mismatch between 
+as in ERQ 2.
+The resulting error distribution contains the full error accumulated in the pipeline, To quantify to what extent this error is due to the mismatch between 
 the generative model $\mathit{G}$ and the selected inference model 
-$\mathit{I}$.
-As twinning (see section \ref{subsec:twinning}: 'Twinning') shows us the baseline error, we can quantify the contribution of the tree prior to the error.
+$\mathit{I}$, we can use twinning (see section \ref{subsec:twinning}: 'Twinning') to provide the baseline error.
 
@@
 as we previously did in listing~\ref{lst:create_all_experiments}.
-We choose again to use the default error measurement setup,
+We use again the default error measurement setup,
 as shown in listing~\ref{lst:create_error_measure_params}.
@@
   \caption{
-    Example 4: the error made by BEAST2 from a phylogeny, when the generative 
-    model is known, compared to candidates with similar inference models, in 
-    relationship to the background error.
+    ERQ 4: the inference error made 
+    when selecting the 'best' tree prior compared with the error obtained for the twin tree.
+    Here, the 'twin' tree shows the baseline inference error,
+    as measured on the twin tree.
   }
@@
 
-We choose again to use the default error measurement parameters
+We use again the default error measurement parameters
 created in listing~\ref{lst:create_error_measure_params}.
@@
     compared to the best of 39 candidate models, 
-    each in relationship to the background error.
+    each in relationship to the background error measured from the twin tree.
   }
@@
 to quantify the importance of a tree prior in Bayesian phylogenetics.
-In all ERQs we used and assumed the most simple tree prior possible,
+In all ERQs we used the most simple generative tree prior possible,
 which is the Yule tree prior.
@@
 with a sufficient number of generative trees, 
-it could actually constitute a quantitative assessment 
+it constitutes a quantitative assessment 
 of the quality of the inference.
@@
 we pretend the results of the example research questions
-to be obtained from thousands of replicates.
-In that case, figure~\ref{fig:example_1} would show 
+were obtained from thousands of replicates.
+In that case, Figure~\ref{fig:example_1} would show 
 the minimum error BEAST2 when the
 generative model is known, which can serve as 
-a minimum noise level assessment in a bigger experiment.  
+a minimum noise level assessment in a larger experiment.  
 Figure~\ref{fig:example_2} shows that 
@@
 tailored on the generative process, is needed. 
-If this would indeed be the case, 
+If this were indeed the case, 
 one could implement their own tree prior as a new module in BEAST2.
@@
 
-Professionals using \verb;pirouette; can cite this
-article. To obtain the citation from within R, use:
+To cite \verb;pirouette; this article from within R, use:
 
@@
 We would like to thank the Center for Information Technology of the University 
-of Groningen for their support and for providing access to the Peregrine 
+of Groningen for its support and for providing access to the Peregrine 
 high performance computing cluster. 
@@
 RJCB, GL and RSE conceived the idea for the package. 
-RJCB created and tested the package, and wrote the first draft of the 
-manuscript.
-GL tested the package and contributed substantially to revisions.
+RJCB created and tested the package, and wrote the first draft of the manuscript.
+GL tested the package and contributed to revisions.
 RSE contributed to revisions.

### 2019-05-13 content.tex

@@
 according to how they operate in the pipeline. 
-We summarize them in Tables~\ref{tab:options} and~\ref{tab:definitions}.
+We summarize them in Table~\ref{tab:options} and Table~\ref{tab:definitions}.
 
@@
 we pretend the results of the example research questions
-were obtained from thousands of replicates.
+to be obtained from thousands of replicates.
 In that case, Figure~\ref{fig:example_1} would show 

### 2019-05-13 content.tex
Add ERQ 6

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 1}
+\section{Usage}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Generative model only}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Now we can formulate the first example research question that \verb;pirouette; 
-can answer.
-
-\subsection{Question}
-
-"What is the inference error made on phylogenies
+can answer: "What is the inference error made on phylogenies
 created by a standard diversification model?"
-
-\subsection{Answer}
 
@@
   \includegraphics[width=\textwidth]{example_1/true_tree.png}
-  \caption{The Yule tree, as created by listing~\ref{lst:create_yule_tree}.}
+  \caption{The Yule tree, as created by Listing~\ref{lst:create_yule_tree}.}
   \label{fig:yule_tree}
@@
 ]
-experiment <- create_experiment(
+gen_experiment <- create_experiment(
   inference_conditions = create_inference_conditions(
@@
 ]
-experiments <- list(experiment)
+experiments <- list(gen_experiment)
 \end{lstlisting}
@@
 
-We will not use twinning here yet.  
 All these arguments are bundled
@@
 Obtaining this control is in fact an inherent part of \verb;pirouette;. 
-We call it 'twinning' and we use it from ERQ 3.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 2}
+We call it 'twinning' and we use it in the last part of this example.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Comparing to other candidate models}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 with a known generative tree prior $\mathit{p_{G}^0}$.
-However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, and hence 
-usually it is not part of a standard inference 
+However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, 
+and hence usually it is not part of a standard inference 
 model (i.e. an inference model using a standard tree prior).
@@
 used in the inference.
-There are two ways to deal with this problem: hand-picking a tree
-prior, or picking the best of a set of tree priors. 
+There are three ways to deal with this problem: hand-picking a tree
+prior (as we did above), or picking the best of a set of tree priors, 
+or to do both. 
 Hand-picking a tree prior can be rational if one standard tree prior 
@@
 models can also be used in inference. Selecting the 'best' model in
-\verb;pirouette; means selecting the model with the highest evidence (also known as marginal likelihood).
-
-We show the procedure using \verb;pirouette; to answer the second research question.
-
-\subsection{Question}
-
-"What is the inference error made on a novel phylogeny when
-using the best inference model?"
-
-\subsection{Answer}
-
-This time we start with a tree generated from an unknown 
-diversification model, that has six taxa and a crown age of ten:
-
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht,
-  frame = single, 
-  label = {lst:unknown_phylogeny},
-  caption = A phylogeny generated by an unknown diversification model.
-]
-phylogeny  <- ape::read.tree(
-  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
-)
-\end{lstlisting}
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_2/tree_unknown.png}
-  \caption{The tree derived from an unknown diversification process, 
-    as created by listing~\ref{lst:unknown_phylogeny}.
-  }
-\end{figure}
+\verb;pirouette; means selecting the model with the highest 
+evidence (also known as marginal likelihood).
+
+We show the procedure using \verb;pirouette; to answer the second research 
+question: "What is the inference error made on a novel phylogeny when
+using the best inference model, in comparison to a hand-picked model?
+
+We will use the same tree as generated in \ref{lst:create_yule_tree}.
 
@@
 We will re-use the alignment parameters of the previous example 
-as shown in listing~\ref{lst:create_alignment_params}.
+as shown in Listing~\ref{lst:create_alignment_params}.
 
@@
     \hline
-    candidate & best candidate & TRUE & JC, strict, Yule \\
+    generative & always         & FALSE & JC, strict, Yule \\
     candidate  & best candidate & TRUE  & JC, strict, BD   \\
+    candidate  & best candidate & TRUE  & JC, strict, CBS  \\
     ...        & ...            & ...   & ...              \\
@@
     RLN: relaxed log-normal clock model.
+    CBS: coalescent Bayesian Skyline tree prior.
     CCP: coalescent constant-population tree prior.
@@
   label = {lst:create_all_experiments},
-  caption = Create all 40 candidate experiments.
-]
-experiments <- create_all_experiments()
+  caption = {
+    Create all 40 candidate experiments, except for the inference model of the generative model.
+  }
+]
+cand_experiments <- create_all_experiments(
+  exclude_model = gen_experiment$inference_model
+)
+\end{lstlisting}
+
+\begin{lstlisting}[
+  language = R, 
+  floatplacement = ht, 
+  frame = single, 
+  label = {lst:create_all_experiments},
+  caption = {
+    Create a collection of experiments, with 1
+    generative model, and 39 candidate models
+  }
+]
+experiments <- list(
+  list(gen_experiment),
+  cand_experiments
+)
 \end{lstlisting}
@@
 \verb;pirouette; parameter set like this (which is the
-same as listing~\ref{lst:create_pir_params}, but using the defaults):
+same as Listing~\ref{lst:create_pir_params}, but using the defaults):
 
@@
 
-We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
-and plot the results (listing~\ref{lst:pir_plot}),
-as shown in Figure~\ref{fig:example_2}.
+We run \verb;pirouette; (Listing~\ref{lst:pir_run}) 
+and plot the results (Listing~\ref{lst:pir_plot}),
+as shown in Figure~\ref{fig:example_6}.
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_2/errors.png}
+  \includegraphics[width=\textwidth]{example_6/errors.png}
   \caption{
-    ERQ 2: the inference error 
-    when selecting the best inference model.
-  }
-  \label{fig:example_2}
+    ERQ 6: the inference error
+    for both the generative and best candidate inference models.
+  }
+  \label{fig:example_6}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 3}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In ERQ 1 we measured the baseline error of the \verb;pirouette; method,
+\subsection{Comparing to background noise}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Up until now, we measured the baseline error of the \verb;pirouette; method,
 by using the same standard tree prior in both the generation of 
@@
 from the new tree prior to the baseline error,
-by using a twin tree (see subsection~\ref{subsec:twinning}).
-In this ERQ, we hand-pick a standard tree prior $\mathit{p_{G}}$.
-
-\subsection{Question}
-
+by using a twin tree (see subsection~\ref{subsec:twinning}) for
+both the generative and best candidate model.
+
+The research question this example answers is:
 "What is the inference error made from a phylogeny,
-when hand-picking a tree prior, compared to the background noise?"
-
-\subsection{Answer}
-
-Most of the settings for this ERQ are
-the same as in ERQ 2: we will re-use the phylogeny from an 
-unknown speciation model (listing~\ref{lst:unknown_phylogeny}), 
-the alignment parameters (listing~\ref{lst:create_alignment_params}).
-
-We create an experiment as in 
-listing~\ref{lst:create_gen_experiment},
-but with all defaults:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_gen_experiment_defaults},
-  caption = Create a list of one default generative experiment.
-]
-experiment <- create_gen_experiment()
-\end{lstlisting}
-
-Creating the complete set of experiments:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_experiments},
-  caption = Create a list of one experiment.
-]
-experiments <- list(experiment)
-\end{lstlisting}
+for both a generative model and a best candidate model, 
+compared to the background noise?"
+
+Most of the settings for this example are
+the same as in ERQ 2: we will re-use the phylogeny as 
+before (Listing~\ref{lst:create_yule_tree}), 
+as well as the alignment parameters (Listing~\ref{lst:create_alignment_params}).
+Also the experiments are the same as in 
+Listing~\ref{lst:create_all_experiments}.
 
@@
 
-We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
-and plot the results (listing~\ref{lst:pir_plot}),
-The output is shown in Figure~\ref{fig:example_3}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_3/errors.png}
-  \caption{
-    ERQ 3: the inference error made 
-    when hand-picking a generative tree prior compared with the error obtained for the twin tree.
-    Here, the 'twin' tree shows the baseline inference error,
-    as measured on the twin tree.
-  }
-  \label{fig:example_3}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Rather than hand-picking a tree prior, 
-one can select the best model from a set,
-as in ERQ 2.
-The resulting error distribution contains the full error accumulated in the pipeline, To quantify to what extent this error is due to the mismatch between 
-the generative model $\mathit{G}$ and the selected inference model 
-$\mathit{I}$, we can use twinning (see section \ref{subsec:twinning}: 'Twinning') to provide the baseline error.
-
-\subsection{Question}
-
-"What is the inference error from a phylogeny 
-derived from an unknown speciation model,
-when selecting the best inference model, 
-in relation to the baseline error?"
-
-\subsection{Answer}
-
-Here, as in ERQs 2 and 3, we start with a phylogeny 
-generated by an unknown speciation model (listing~\ref{lst:unknown_phylogeny}). 
-The first step is to simulate a DNA alignment 
-from the given phylogeny, for which we will use the same code 
-as shown in listing~\ref{lst:create_alignment_params}.
-
-We specify that we want all 40 inference models to compete 
-by calling \verb;create_all_experiments;,
-as we previously did in listing~\ref{lst:create_all_experiments}.
-We use again the default error measurement setup,
-as shown in listing~\ref{lst:create_error_measure_params}.
-To get an idea of a baseline error made by BEAST2, 
-we also make use of the twinning option by using
-the same call to \verb;create_pir_params; as shown in 
-listing~\ref{lst:create_pir_params_with_twinning}.
-We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
-and plot the results (listing~\ref{lst:pir_plot}).
-The output is shown in Figure~\ref{fig:example_4}.
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_4/errors.png}
-  \caption{
-    ERQ 4: the inference error made 
-    when selecting the 'best' tree prior compared with the error obtained for the twin tree.
-    Here, the 'twin' tree shows the baseline inference error,
-    as measured on the twin tree.
-  }
-  \label{fig:example_4}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Usage: Example Research Question 5}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-If we are moderately sure about hand-picking an inference model,
-but would also like to check how well the hand-picked model compares
-to the others, we can do both. Of course, as we are interested in the
-baseline error for both the hand-picked and best candidate model,
-we need to enable twinning as well. This example is a combination
-of the previous two ERQs.
-
-\subsection{Question}
-
-"What is the error made by BEAST2 from a phylogeny, 
-when the generative model is hand-picked, 
-compared to candidates with similar inference models, 
-in relation to the background error?"
-
-\subsection{Answer}
-
-In this example we use a tree generated by the Yule process.
-Instead of using the common Yule tree as in ERQ 1, we will use
-a Yule tree of low likelihood, 
-i.e. the one already used in ERQs 2-4, 
-by using the code shown in listing~\ref{lst:unknown_phylogeny}.
-The first step is to simulate a DNA alignment 
-from the given phylogeny, for which we will use the same code 
-as shown in listing~\ref{lst:create_alignment_params}.
-
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \textbf{model\_type} & 
-    \textbf{run\_if} & 
-    \textbf{do\_measure\_evidence} & 
-    \textbf{inference model} \\ 
-    \hline
-    generative & always         & TRUE & JC, strict, Yule \\
-    candidate  & best candidate & TRUE & JC, strict, BD \\
-    candidate  & best candidate & TRUE & JC, strict, CCP \\
-    candidate  & best candidate & TRUE & JC, strict, CEP \\
-    \hline
-  \end{tabular}
-  \caption{
-    Experimental setup to answer the fourth research question.
-    JC: Jukes-Cantor site model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-    BD: birth-death tree prior.
-    CCP: coalescent constant-population tree prior.
-    CEP: coalescent exponential-population tree prior.
-  }
-  \label{tab:experiment_5}
-\end{table}
-
-In the second step we state our experiments. 
-In this example we must specify both the generative and candidate experiments. 
-Here we specify the generative model:
-
-\begin{lstlisting}[language=R, floatplacement=H, frame=single]
-generative_experiment <- create_gen_experiment()
-\end{lstlisting}
-
-We can specify all the candidates simply by calling 
-\verb;create_all_experiments; and excluding the generative inference model,
-resulting in 39 candidate experiments:
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = H,
-    frame = single
-]
-candidate_experiments <- create_all_experiments(
-  exclude_model = generative_experiment$inference_model
-)
-\end{lstlisting}
-
-We need to combine these experiments into one set of experiments:
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = H,
-    frame = single
-]
-experiments <- c(list(generative_experiment), candidate_experiments)
-\end{lstlisting}
-
-We use again the default error measurement parameters
-created in listing~\ref{lst:create_error_measure_params}.
-We enable twinning, by calling \verb;create_pir_params;
-as shown in listing~\ref{lst:create_pir_params_with_twinning}.
-
-We run \verb;pirouette; (listing~\ref{lst:pir_run}) 
-and plot the results (listing~\ref{lst:pir_plot}).
-The output is shown in Figure~\ref{fig:example_5}.
+We run \verb;pirouette; (Listing~\ref{lst:pir_run}) 
+and plot the results (Listing~\ref{lst:pir_plot}),
+The output is shown in Figure~\ref{fig:example_5}
 
@@
   \caption{
-    Example 5: the inference error made, 
-    when the generative model is hand-picked, 
-    compared to the best of 39 candidate models, 
-    each in relationship to the background error measured from the twin tree.
+    The inference error made 
+    for both a generative tree prior and best candidate model
+    compared with the error obtained for the twin tree.
+    Here, the 'twin' tree shows the baseline inference error
   }
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Interpretation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-05-13 scripts/example_6.R

@@ -1,0 +1,262 @@
+# Code of example 6
+#
+# Works under Linux and MacOS only
+library(pirouette)
+library(ggplot2)
+library(ggtree)
+
+root_folder <- path.expand("~/GitHubs/pirouette_article")
+example_no <- 6
+example_folder <- file.path(root_folder, paste0("example_", example_no))
+dir.create(example_folder, showWarnings = FALSE)
+setwd(example_folder)
+set.seed(314)
+
+testit::assert(is_beast2_installed())
+
+# An infintely rare Yule tree
+phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
+
+alignment_params <- create_alignment_params(
+  root_sequence = create_blocked_dna(length = 1000),
+  mutation_rate = 0.1
+)
+
+# JC69, strict, Yule
+generative_experiment <- create_experiment(
+  inference_conditions = create_inference_conditions(
+    model_type = "generative",
+    run_if = "always",
+    do_measure_evidence = TRUE
+  ),
+  inference_model = create_inference_model(
+    site_model = create_jc69_site_model(),
+    clock_model = create_strict_clock_model(),
+    tree_prior = create_yule_tree_prior(),
+    mcmc = create_mcmc(chain_length = 1e+07, store_every = 1000)
+  )
+)
+generative_experiment <- create_gen_experiment()
+check_experiment(generative_experiment)
+
+# All non-Yule tree priors
+candidate_experiments <- create_all_experiments(
+  exclude_model = generative_experiment$inference_model
+)
+check_experiments(candidate_experiments)
+
+experiments <- c(list(generative_experiment), candidate_experiments)
+check_experiments(experiments)
+
+# Testing
+if (1 == 2) {
+  experiments <- experiments[1:2]
+  for (i in seq_along(experiments)) {
+    experiments[[i]]$inference_model$mcmc <- create_mcmc(chain_length = 10000, store_every = 1000)
+    experiments[[i]]$est_evidence_mcmc <- create_mcmc_nested_sampling(
+      chain_length = 10000,
+      store_every = 1000,
+      epsilon = 100.0
+    )
+  }
+}
+
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments
+)
+
+print("#######################################################################")
+print("Settings to run on Peregrine cluster")
[…]

### 2019-05-13 scripts/example_1.R
Add script for new ERQ

@@ -12,9 +12,7 @@ example_folder <- file.path(root_folder, paste0("example_", example_no))
 dir.create(example_folder, showWarnings = FALSE)
 setwd(example_folder)
 set.seed(314)
-
 testit::assert(is_beast2_installed())
-
 phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
 
 alignment_params <- create_alignment_params(

### 2019-05-13 scripts/example_2.R

@@ -12,12 +12,8 @@ example_folder <- file.path(root_folder, paste0("example_", example_no))
 dir.create(example_folder, showWarnings = FALSE)
 setwd(example_folder)
 set.seed(314)
-
 testit::assert(is_beast2_installed())
-
-phylogeny  <- ape::read.tree(
-  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
-)
+phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),

### 2019-05-13 scripts/example_3.R

@@ -27,7 +27,6 @@ print(paste("Number of arguments:", length(args)))
 example_no <- 3
 root_folder <- path.expand("~/GitHubs/pirouette_article")
 example_folder <- file.path(root_folder, paste0("example_", example_no))
-rng_seed <- 314
 
 if (length(args) == 1) {
   rng_seed <- as.numeric(args[1])
@@ -45,11 +44,9 @@ library(ggtree)
 
 dir.create(example_folder, showWarnings = FALSE)
 setwd(example_folder)
-# No need to do 'set.seed(rng_seed)': we use 'rng_seed' arguments instead
-
+set.seed(314)
 testit::assert(is_beast2_installed())
-
-phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
+phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),

### 2019-05-13 scripts/example_4.R

@@ -11,10 +11,8 @@ example_folder <- file.path(root_folder, paste0("example_", example_no))
 dir.create(example_folder, showWarnings = FALSE)
 setwd(example_folder)
 set.seed(314)
-
 testit::assert(is_beast2_installed())
-
-phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
+phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),

### 2019-05-13 scripts/example_6.R

@@ -11,11 +11,8 @@ example_folder <- file.path(root_folder, paste0("example_", example_no))
 dir.create(example_folder, showWarnings = FALSE)
 setwd(example_folder)
 set.seed(314)
-
 testit::assert(is_beast2_installed())
-
-# An infintely rare Yule tree
-phylogeny  <- ape::read.tree(text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);")
+phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
 
 alignment_params <- create_alignment_params(
   root_sequence = create_blocked_dna(length = 1000),

### 2019-05-13 supplement.tex
Remove ERQ 2 to and including 4, added ERQ 6 (which belongs between 1 and 5)

@@
 % has label tab:esses_example_1
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 2}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.25\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_alignment.png}
-    };   
-    \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_posterior_best.png}
-    };   
-    \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_error_violin_best.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (C)
-      (C) edge [anchor = south] node {} (D)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_2_full_pipeline}
-  \caption{Example 2: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_2/esses.latex}
-% has label tab:esses_example_2
-
-\input{example_2/evidences.latex}
-% has label tab:evidences_example_2
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 3}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.4\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3/true_alignment.png}
-    };   
-    \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_posterior_gen.png}
-    };   
-    \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_error_violin_gen.png}
-    };   
-    \node[state] (AT) [right of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/twin_tree.png}
-    };   
-    \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3/twin_alignment.png}
-    };   
-    \node[state] (CT) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/twin_posterior_gen.png}
-    };   
-    \node[state] (DT) [below of = CT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/twin_error_violin_gen.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (C)
-      (C) edge [anchor = south] node {} (D)
-      (A) edge [anchor = east] node {} (AT)
-      (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = south] node {} (CT)
-      (CT) edge [anchor = south] node {} (DT)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_3_full_pipeline}
-  \caption{Example 3: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_3/esses.latex}
-% has label tab:esses_example_3
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 4}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {1.0\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.5\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{example_4/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_4/true_alignment.png}
-    };   
-    \node[state] (CB) [below of = B, rectangle, node distance=0.8\textheight] {
-
-\includegraphics[height=0.95\textheight]{example_4/true_posterior_best.png}
-    };   
-    \node[state] (DB) [below of = CB, rectangle, node distance=1.0\textheight] {
-      \includegraphics[height=0.65\textheight]{example_4/true_error_violin_best.png}
-    };   
-    \node[state] (AT) [right of = A, rectangle, node distance=2.0\textwidth] {
-      \includegraphics[height=0.4\textheight]{example_4/twin_tree.png}
-    };   
-    \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_4/twin_alignment.png}
-    };   
-    \node[state] (CTB) [below of = BT, rectangle, node distance=0.8\textheight] 
-{
-
-\includegraphics[height=0.95\textheight]{example_4/twin_posterior_best.png}
-    };   
-    \node[state] (DTB) [below of = CTB, rectangle, node 
-distance=1.0\textheight] {
-
-\includegraphics[height=0.65\textheight]{example_4/twin_error_violin_best.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south east] node {} (CB)
-      (CB) edge [anchor = south] node {} (DB)
-      (A) edge [anchor = east] node {} (AT)
-      (AT) edge [anchor = south] node {} (BT)
-      (BT) edge [anchor = south] node {} (CTB)
-      (CTB) edge [anchor = south] node {} (DTB)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_4_full_pipeline}
-  \caption{Example 4: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_4/esses_best.latex}
-% has label tab:esses_example_4_best
-
-\input{example_4/esses_twin_best.latex}
-% has label tab:esses_example_4_twin_best
-
-\input{example_4/evidence_true.latex}
-% has label tab:evidences_example_4
-
-\input{example_4/evidence_twin.latex}
-% has label tab:evidences_example_4
 
@@
 \end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Example 6}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[ht]
+  \centering
+  \resizebox {0.8\columnwidth} {!} {
+    \begin{tikzpicture}[
+      ->,>=stealth',shorten >=1pt,auto,
+      node distance=0.4\textheight, 
+      semithick
+    ]   
+    \tikzstyle{every state}=[]
+    \node[state, draw=none] (O) [] {
+    };   
+    \node[state] (A) [right of = O, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_6/true_tree.png}
+    };   
+    \node[state] (B) [below of = A, rectangle] {
+      \includegraphics[height=0.13\textheight]{example_6/true_alignment.png}
+    };   
+    \node[state] (CG) [below of = B, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_6/true_posterior_gen.png}
+    };   
+    \node[state] (DG) [below of = CG, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_6/true_error_violin_gen.png}
+    };   
+    \node[state] (CB) [right of = C, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_6/true_posterior_best.png}
+    };   
+    \node[state] (DB) [below of = CB, rectangle] {
+      \includegraphics[height=0.2\textheight]{example_6/true_error_violin_best.png}
+    };   
+    \path 
+      (O) edge [anchor = south] node {} (A)
+      (A) edge [anchor = south] node {} (B)
+      (B) edge [anchor = south] node {} (CG)
+      (CG) edge [anchor = south] node {} (DG)
+      (B) edge [anchor = south east] node {} (CB)
+      (CB) edge [anchor = south] node {} (DB)
+    ; 
+    \end{tikzpicture}
+  }
+  \label{fig:example_6_full_pipeline}
+  \caption{Example 6: full pipeline}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\input{example_6/esses_gen.latex}
+% has label tab:esses_example_6_gen
+
+\input{example_6/esses_best.latex}
+% has label tab:esses_example_6_best
+
+\input{example_6/evidence_true.latex}
+% has label tab:evidences_example_6
+
+\begin{table}
+  \begin{tabular}{ | r | c | }
+    \hline
+    \textbf{ERQ} & \textbf{run-time} \\ 
+    \hline
+      1 & 00:10:12 \\
+      2 & 00:34:10 \\
+      3 & 00:24:53 \\
+      4 & 01:04:50 \\
+      5 & 01:29:40 \\
+    \hline
+  \end{tabular}
+  \caption{
+    Time for an example research question to finish.
+    Time is in in hh:mm::ss.
+  }
+  \label{tab:runtime}
+\end{table}

### 2019-05-13 scripts/example_2.R

@@ -1,147 +1,0 @@
-# Code of example 2
-#
-# Works under Linux and MacOS only
-library(pirouette)
-library(ggplot2)
-library(ggthemes)
-library(ggtree)
-
-root_folder <- path.expand("~/GitHubs/pirouette_article")
-example_no <- 2
-example_folder <- file.path(root_folder, paste0("example_", example_no))
-dir.create(example_folder, showWarnings = FALSE)
-setwd(example_folder)
-set.seed(314)
-testit::assert(is_beast2_installed())
-phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
-
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
-)
-
-experiments <- create_all_experiments()
-
-# Testing
-if (1 == 2) {
-  experiments <- create_all_experiments()[1:2]
-  for (i in seq_along(experiments)) {
-    experiments[[i]]$inference_model$mcmc <- create_mcmc(chain_length = 10000, store_every = 1000)
-    experiments[[i]]$est_evidence_mcmc <- create_mcmc_nested_sampling(
-      chain_length = 10000,
-      store_every = 1000,
-      epsilon = 100.0
-    )
-  }
-}
-
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments
-)
-
-################################################################################
-# Settings to run on Peregrine cluster
-################################################################################
-pir_params$alignment_params$fasta_filename <- file.path(example_folder, "true.fasta")
-for (i in seq_along(pir_params$experiments)) {
-  pir_params$experiments[[i]]$beast2_options$input_filename <- file.path(example_folder, "beast2_input.xml")
-  pir_params$experiments[[i]]$beast2_options$output_log_filename <- file.path(example_folder, "beast2_output.log")
-  pir_params$experiments[[i]]$beast2_options$output_trees_filenames <- file.path(example_folder, "beast2_output.trees")
-  pir_params$experiments[[i]]$beast2_options$output_state_filename <- file.path(example_folder, "beast2_output.xml.state")
-  pir_params$experiments[[i]]$beast2_options$beast2_working_dir <- example_folder
-  pir_params$experiments[[i]]$errors_filename <- file.path(example_folder, "error.csv")
-}
[…]

### 2019-05-13 scripts/example_3.R

@@ -1,161 +1,0 @@
-# Code of example 3
-#
-# Works under Windows
-#
-# Usage #1:
-#
-# Rscript example_3.R
-#
-# Uses default RNG seed.
-# Puts result in pirouette_article/example_3 folder
-#
-# Usage #2:
-#
-# Rscript example_3.R [rng_seed]
-#
-# For example:
-#
-# Rscript example_3.R 314
-#
-# Uses [rng_seed] as the RNG seed.
-# Puts result in 'pirouette_article/example_3_[rng_seed]' folder,
-# e.g 'pirouette_article/example_3_314'
-
-args <- commandArgs(TRUE)
-print(paste("Number of arguments:", length(args)))
-
-example_no <- 3
-root_folder <- path.expand("~/GitHubs/pirouette_article")
-example_folder <- file.path(root_folder, paste0("example_", example_no))
-
-if (length(args) == 1) {
-  rng_seed <- as.numeric(args[1])
-  example_folder <- file.path(root_folder, paste0("example_", example_no, "_", rng_seed))
-} else {
-  stop("Need argument")
-}
-
-print(rng_seed)
-print(example_folder)
-
-library(pirouette)
-library(ggplot2)
-library(ggtree)
-
-dir.create(example_folder, showWarnings = FALSE)
-setwd(example_folder)
-set.seed(314)
-testit::assert(is_beast2_installed())
-phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
-
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1,
-  rng_seed = rng_seed
-)
-
-experiments <- list(create_gen_experiment(beast2_options = create_beast2_options(rng_seed = rng_seed)))
-
-# Testing
-if (1 == 2) {
-  for (i in seq_along(experiments)) {
-    experiments[[i]]$inference_model$mcmc <- create_mcmc(chain_length = 10000, store_every = 1000)
-    experiments[[i]]$est_evidence_mcmc <- create_mcmc_nested_sampling(
-      chain_length = 10000,
-      store_every = 1000,
-      epsilon = 100.0
-    )
-  }
-}
-
-twinning_params <- create_twinning_params(
-  twin_model = "birth_death",
-  method = "random_tree",
- rng_seed = rng_seed
-)
-
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  twinning_params = twinning_params
-)
-
[…]

### 2019-05-13 scripts/example_4.R

@@ -1,203 +1,0 @@
-# Code of example 4
-#
-# Works under Linux and MacOS only
-library(pirouette)
-library(ggplot2)
-library(ggtree)
-
-root_folder <- path.expand("~/GitHubs/pirouette_article")
-example_no <- 4
-example_folder <- file.path(root_folder, paste0("example_", example_no))
-dir.create(example_folder, showWarnings = FALSE)
-setwd(example_folder)
-set.seed(314)
-testit::assert(is_beast2_installed())
-phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
-
-alignment_params <- create_alignment_params(
-  root_sequence = create_blocked_dna(length = 1000),
-  mutation_rate = 0.1
-)
-
-experiments <- create_all_experiments()
-check_experiments(experiments)
-
-# Testing
-if (1 == 2) {
-  experiments <- experiments[1:2]
-  for (i in seq_along(experiments)) {
-    experiments[[i]]$inference_model$mcmc <- create_mcmc(chain_length = 10000, store_every = 1000)
-    experiments[[i]]$est_evidence_mcmc <- create_mcmc_nested_sampling(
-      chain_length = 10000,
-      store_every = 1000,
-      epsilon = 100.0
-    )
-  }
-}
-
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  twinning_params = create_twinning_params()
-)
-
-print("#######################################################################")
-print("Settings to run on Peregrine cluster")
-print("#######################################################################")
-pir_params$alignment_params$fasta_filename <- file.path(example_folder, "true.fasta")
-for (i in seq_along(pir_params$experiments)) {
-  pir_params$experiments[[i]]$beast2_options$input_filename <- file.path(example_folder, "beast2_input_best.xml")
-  pir_params$experiments[[i]]$beast2_options$output_log_filename <- file.path(example_folder, "beast2_output_best.log")
-  pir_params$experiments[[i]]$beast2_options$output_trees_filenames <- file.path(example_folder, "beast2_output_best.trees")
-  pir_params$experiments[[i]]$beast2_options$output_state_filename <- file.path(example_folder, "beast2_output_best.xml.state")
-  pir_params$experiments[[i]]$beast2_options$beast2_working_dir <- example_folder
[…]

### 2019-05-14 content.tex
Review manuscript, progress #46

@@
 Although many possible tests can be performed, 
-we present here what we think are the five most important 
-example research questions (ERQ) that \verb;pirouette; can answer, from simple questions to complex ones.
-For each of these ERQs we will quantify the importance of a tree prior.
+we show the usage of \verb;pirouette; by a story that introduces
+its features gradually, yet ending in quantifying the impact a tree prior has
+in Bayesian inference.
 
@@
 
+We show the usage of \verb;pirouette; by a story that introduces
+its features gradually. First we measure the Bayesian inference error
+when we have a phylogeny with a known and standard tree prior,
+to get an idea of the baseline error.
+Second, we'll additionally measure the inference error made by a
+best candidate model, to establish that the true tree prior is indeed the best.
+Lastly, we will seperate the baseline error and complete error in
+a tree generated by an unknown or non-standard tree prior, 
+to quantify the impact the tree prior has in Bayesian inference.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \hline
-    \textbf{model\_type} &
-    \textbf{run\_if} &
-    \textbf{do\_measure\_evidence} & 
-    \textbf{inference model} \\ 
+    \texttt{model\_type} &
+    \texttt{run\_if} &
+    \texttt{do\_measure\_evidence} & 
+    \texttt{inference model} \\ 
     \hline
@@
   \caption{
-    Inference conditions and inference model for ERQ 1.
+    Inference conditions and model.
     JC: Jukes-Cantor site model.
@@
   caption = {
-    Create the experiments to answer research question 1, 
+    Create an experiment with the generative model,
+    that will always be used in the actual inference, 
     using explicit arguments.
@@
 ]
-gen_experiment <- create_experiment(
+generative_experiment <- create_experiment(
   inference_conditions = create_inference_conditions(
@@
   caption = {
-    Create the experiments to answer research question 1.
-  }
-]
-experiments <- list(gen_experiment)
-\end{lstlisting}
-
-The objects we just created must be bundled
-in a \verb;pirouette; parameters structure,
-together with error measurement parameters $\mathit{E}$ and
-twinning parameters $\mathit{T}$.
-\giovanni{I suggest to refer to the twinning parameters only when strictly required and with the minimum possible amount of info provided, as we set it in such a way that the tree prior is imported from the inference model in an automatic way.}
-
-We use the default error measurement
+    Create the experiments. In this case, we use only
+    an experiment with the generative model,
+    that will always be used in the actual inference.
+  }
+]
+experiments <- list(generative_experiment)
+\end{lstlisting}
+
+We also need to specify the error measurement parameters $\mathit{E}$.
+Here we use the default error measurement
 parameters, which uses a burn-in fraction of 10\% and the nLTT statistic to
@@
 
-All these arguments are bundled
-and checked by \verb;create_pir_params;:
+We now have create the alignment parameters,
+our experiments and our error measure parameters.
+These objects need to be bundled in a bigger
+\verb;pirouette; parameter structure, 
+using \verb;create_pir_params;:
 
@@
   experiments = experiments,
-  error_measure_params = error_measure_params,
-  twinning_params = NA
+  error_measure_params = error_measure_params
 )
@@
   \caption{
-    ERQ 1: the inference error made 
+    The inference error made 
     when the generative and inference model are the same.
@@
 with a known generative tree prior $\mathit{p_{G}^0}$.
-However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, 
-and hence usually it is not part of a standard inference 
+However, a novel tree prior $\mathit{p_{G}}$, by our definition, 
+is non-standard, and hence not part of a standard inference 
 model (i.e. an inference model using a standard tree prior).
-In such a case, the question is what standard tree prior should be 
+In such a case, the question is which standard tree prior should be 
 used in the inference.
-There are three ways to deal with this problem: hand-picking a tree
-prior (as we did above), or picking the best of a set of tree priors, 
-or to do both. 
-Hand-picking a tree prior can be rational if one standard tree prior 
-is more suitable than others for the inference, for example, when
-the novel tree prior is most similar to a specific standard tree prior.
-Instead of hand-picking, the best inference model from a list of candidate 
-models can also be used in inference. Selecting the 'best' model in
-\verb;pirouette; means selecting the model with the highest 
-evidence (also known as marginal likelihood).
-
-We show the procedure using \verb;pirouette; to answer the second research 
+In this example, we re-use our hand-picked inference model,
+but additionally pick a best inference model from a set of inference models.
+With 'best' inference model, we mean the model with the highest 
+evidence (also known as marginal likelihood), based on the alignment.
+We show the procedure using \verb;pirouette; to answer a second research 
 question: "What is the inference error made on a novel phylogeny when
 using the best inference model, in comparison to a hand-picked model?
-
-We will use the same tree as generated in \ref{lst:create_yule_tree}.
-
-The first step is to simulate a DNA alignment from the 
-given phylogeny. 
-We will re-use the alignment parameters of the previous example 
-as shown in Listing~\ref{lst:create_alignment_params}.
 
@@
   \caption{
-    Inference conditions and inference model for ERQ 2.
+    Inference conditions and model.
     JC: Jukes-Cantor site model.
@@
 
-In the second step, we deviate from the answer to ERQ 1 by providing a 
-set of experiments $\{\mathit{X_{i}}\}_{i=1}^{40}$. All the experiments have 
-the same inference conditions $\mathit{C_{i}} = \mathit{C}$ but different 
-candidate inference models $\mathit{I_{i}}$. Only one of them will be used 
-to run the actual full inference. According to the specified $\mathit{C}$, 
-\verb;pirouette; will measure the evidence for each of them and select only 
-the one with the highest evidence as the best candidate.
-
+We will use the same tree as generated in \ref{lst:create_yule_tree},
+as well as the same alignment parameters as shown in 
+Listing~\ref{lst:create_alignment_params}.
+
+We do need to specify a different set of experiments: we
+need to state that we already have an experiment for
+the generative model, as well as that we want all other inference
+models to compete. We call the competing models 'candidate models',
+as only the best will be used in the actual inference.
 As model selection is commonly performed on the full list of available 
@@
 containing the inference models of all combinations of 
-4 site models, 2 clock models and 5 tree priors.
+4 site models, 2 clock models and 5 tree priors. All we need to add is
+to exclude the inference model in the generative experiment:
 
@@
 ]
-cand_experiments <- create_all_experiments(
-  exclude_model = gen_experiment$inference_model
+candidate_experiments <- create_all_experiments(
+  exclude_model = generative_experiment$inference_model
 )
 \end{lstlisting}
+
+We combine the generative and all candidate models into one set of experiments:
 
@@
 experiments <- list(
-  list(gen_experiment),
-  cand_experiments
+  list(generative_experiment),
+  candidate_experiments
 )
@@
 
-For all the other steps we use the default setups. 
-We can create the complete
+We can now create the complete
 \verb;pirouette; parameter set like this (which is the
@@
   \caption{
-    ERQ 6: the inference error
+    The inference error
     for both the generative and best candidate inference models.
@@
 
-Up until now, we measured the baseline error of the \verb;pirouette; method,
-by using the same standard tree prior in both the generation of 
-the original tree, as well as in the inference.
-This baseline error distribution serves as a control,
-to quantify the impact of a new (non-standard) tree prior.
-To measure that impact, we compare the inference error 
-from the new tree prior to the baseline error,
-by using a twin tree (see subsection~\ref{subsec:twinning}) for
+Up until now, we measured the inference error on a tree
+of which we knew the (standard) tree prior. The goal of \verb;pirouette;
+is to measure the impact of a novel tree prior.
+To do so, in this example, we'll use a tree generated by a non-standard
+tree prior, of which we assume is closely related to the Yule tree prior.
+To measure the impact of the tree prior, we measure both the
+baseline error (that serves as a control) 
+and the full inference error: their difference shows
+the impact of the tree prior.
+To do so, we use a twin tree (see Subsection~\ref{subsec:twinning}) for
 both the generative and best candidate model.
-
 The research question this example answers is:
@@
 
-Most of the settings for this example are
-the same as in ERQ 2: we will re-use the phylogeny as 
-before (Listing~\ref{lst:create_yule_tree}), 
-as well as the alignment parameters (Listing~\ref{lst:create_alignment_params}).
-Also the experiments are the same as in 
-Listing~\ref{lst:create_all_experiments}.
-
-This time, however, we measure both the baseline and complete inference error.
-We set the twinning parameters $\mathit{T}$ to create a twin tree,
-as described in the 'Twinning' section \ref{subsec:twinning}.
-
-Creating the twinning parameters is trivial with the default settings.
+This time we start with a tree generated from an unknown 
+diversification model, that has six taxa and a crown age of ten:
+\begin{lstlisting}[
+  language = R, 
+  floatplacement = ht,
+  frame = single, 
+  label = {lst:unknown_phylogeny},
+  caption = A phylogeny generated by an unknown diversification model.
+]
+phylogeny  <- ape::read.tree(
+  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
+)
+\end{lstlisting}
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{example_6/true_tree.png}
+  \caption{The tree derived from an unknown diversification process, 
+    as created by listing~\ref{lst:unknown_phylogeny}.
+  }
+\end{figure}
+
+Most of the other settings are the same as before:
+we re-use the alignment parameters (Listing~\ref{lst:create_alignment_params}),
+as well as the experiments (Listing~\ref{lst:create_all_experiments}).
+This time, however, we enable twinning (see Subsection~\ref{subsec:twinning}),
+by creating twinning parameters.
+Creating these parameters is trivial with the default settings.
 For clarity, however, we explicitly show the most
@@
 \end{lstlisting}
-\giovanni{We want to import the "twin\_model" from the inference model.}
-\richel{Yes, that is an option, but according to Rampal not important enough
-to add now}
-\rampal{did I say that? I think it's OK not to do it in the paper, but perhaps it can be noted.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Interpretation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We showed how to use \verb;pirouette; 
+to quantify the importance of a tree prior in Bayesian phylogenetics,
+using the most simple generative tree prior possible.
+In principle any other (more complex) generative tree prior 
+can be tested, but we chose to provide the 
+simplest (and fastest to run) examples.
+All tests were performed on a single original tree, 
+hence not providing us enough statistical power to 
+evaluate the level of accuracy of the inference.
+
+However, if the same procedure were repeated and performed on a distribution 
+with a sufficient number of generative trees, 
+it constitutes a quantitative assessment 
+of the quality of the inference.
+
+A twin tree does not always result in a lower error distribution,
+due to the stochasticity in generating the twin tree.
+For example, of the second tree used in the 'Usage' section,
+94\% of all twin trees resulted in an error distribution with a lower median.
+
+% Add: the effect of the alignment on this
+% https://github.com/richelbilderbeek/pirouette_article/issues/45
+
+To give an idea of the insights \verb;pirouette; can give,
+we pretend the final result was to be obtained from thousands of replicates.
+We ...
+
+In conclusion, \verb;pirouette; can directly show 
+the comparison between the results obtained 
+adopting a new generative prior as compared to the Bayesian inference error
+when operating on a tree generated under a known and standard tree 
+prior (the twin tree).
+From such results an user can estimate whether or not a new tree prior, 
+tailored on the generative process, is needed. 
+If this were indeed the case, 
+one could rightly implement their novel tree prior as an addition to 
+their favorite Bayesian inference tool.
 
@@
 \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% RJCB will add: in six percent did the twin do worse, 
-% see https://github.com/richelbilderbeek/pirouette_article/issues/41
-
-In the five example research questions, 
-we showed how to use \verb;pirouette; 
-to quantify the importance of a tree prior in Bayesian phylogenetics.
-In all ERQs we used the most simple generative tree prior possible,
-which is the Yule tree prior.
-In principle any other (more complex) generative tree prior 
-can be tested, but we chose to provide the 
-simplest (and fastest to run) examples.
-All tests were performed on a single original tree, 
-hence not providing us enough statistical power to 
-evaluate the level of accuracy of the inference.
-
-However, if the same procedure were repeated and performed on a distribution 
-with a sufficient number of generative trees, 
-it constitutes a quantitative assessment 
-of the quality of the inference.
-To give an idea of the insights \verb;pirouette; can give,
-we pretend the results of the example research questions
-to be obtained from thousands of replicates.
-In that case, Figure~\ref{fig:example_1} would show 
-the minimum error BEAST2 when the
-generative model is known, which can serve as 
-a minimum noise level assessment in a larger experiment.  
-Figure~\ref{fig:example_2} shows that 
-the best candidate model gives an error distribution similar to the unknown 
-tree prior,
-hinting that the candidate model would be just as good in inference.
-Figure~\ref{fig:example_3} shows that 
-the twin model gives an error distribution similar to the unknown tree prior,
-hinting that the shape of the phylogeny with unknown tree prior
-does not effect the error made by BEAST2.
-Figure~\ref{fig:example_4} shows that 
-the twin model and best candidate model give 
-an error distribution similar to the unknown tree prior,
-hinting that the shape of the phylogeny with unknown tree prior
-does not effect the error made by BEAST2, 
-and that the candidate model would be just as good in inference.
-
-In conclusion, \verb;pirouette; can directly show 
-the comparison between the results obtained 
-adopting a new generative prior as compared to the error BEAST2 
-would make when operating on a tree generated 
-under a known, right and existing tree prior (the twin tree).
-From such results an user can estimate whether or not a new tree prior, 
-tailored on the generative process, is needed. 
-If this were indeed the case, 
-one could implement their own tree prior as a new module in BEAST2.
 

### 2019-05-16 content.tex
Changes @Giappo

@@
 \centering
-  \begin{tabular}{|p{3.2cm}|p{9.7cm}|p{4.5cm}@{}|}
+  \begin{tabular}{|p{3.4cm}|p{9.7cm}|p{4.5cm}@{}|}
     \hline
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;burn\_in\_fraction; &
+    \verb;burn_in_fraction; &
     Specifies the percentage of initial posterior trees to discard &
@@
 
-  \begin{tabular}{|@{}c|p{4cm}|p{12cm}|}
+  \begin{tabular}{|@{}c|p{4cm}|p{12.2cm}|}
     \hline
@@
     Phylogenetic inference model to run BEAST2. Likewise the generative model $G$, 
-    its main components are the tree prior $\mathit{p_{I}}$, the 
-    clock model $\mathit{c_{I}}$ and the site model $\mathit{s_{I}}$. \\
+    its main components are the tree prior $\mathit{p_{I}}$, the clock model $\mathit{c_{I}}$ and the site model $\mathit{s_{I}}$. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     They are composed by the model type, run condition and 
-    wether to measure the evidence. \\
+    whether to measure the evidence. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{table}
+\giovanni{I think the reference to the listings is a great idea. I don't know what to add more to this table. Probably (not 100\% sure) we should make all tables fit within the page limits.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We show the usage of \verb;pirouette; by a story that introduces
-its features gradually. First we measure the Bayesian inference error
-when we have a phylogeny with a known and standard tree prior,
-to get an idea of the baseline error.
-Second, we'll additionally measure the inference error made by a
-best candidate model, to establish that the true tree prior is indeed the best.
-Lastly, we will seperate the baseline error and complete error in
-a tree generated by an unknown or non-standard tree prior, 
-to quantify the impact the tree prior has in Bayesian inference.
+We show the usage of \verb;pirouette; by gradually introducing
+its features.
+First, to get an idea of the baseline error, we measure the Bayesian inference error when we start from a phylogeny generated under a known and standard tree prior.
+Second, to establish that the true tree prior is indeed the best, we also measure the inference error made adopting a
+best candidate inference model.
+Lastly, we quantify the impact of the tree prior in the Bayesian inference.
+We do so by separating the baseline error and the complete error running the pipeline from a tree generated by an unknown or non-standard tree prior.
 
@@
 
-\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's 
-inference,
-by measuring the difference between a given tree and 
-phylogenies in a posterior. Due to stochasticity, all posterior trees
-will differ from the given phylogeny $\tau_{\mathit{G}}$, 
-even when the generative tree prior $\mathit{p_{G}}$ 
-and alignment model $\mathit{A}$ are the same as 
-those in the inference model $\mathit{I}$.
+\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's inference by measuring the discrepancy between a given tree and a posterior distribution of phylogenies obtained as result of the inference process.
+Due to stochasticity, posterior trees will generally differ from the given phylogeny $\tau_{\mathit{G}}$, even when the generative tree prior $\mathit{p_{G}}$ and alignment model $\mathit{A}$ are the same as those in the inference model $\mathit{I}$.
 Measuring this difference allows us to know the baseline error
-of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' 
-all the tree priors that are available to be used within an inference model 
-(see Table~\ref{tab:options}).
-
-Now we can formulate the first example research question that \verb;pirouette; 
-can answer: "What is the inference error made on phylogenies
-created by a standard diversification model?"
-
-In this example we use a standard generative tree prior $\mathit{p_{G}^0}$, 
-namely the Yule (pure-birth) tree prior. 
+of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' all the tree priors that are available to be used within an inference model (see Table~\ref{tab:options}).
+
+Now we can formulate the first example research question that \verb;pirouette; can answer: "What is the inference error made on phylogenies created by a standard diversification model?"
+
+In this example we use a standard generative tree prior $\mathit{p_{G}^0}$, namely the Yule (pure-birth) tree prior. 
 We choose to use a small tree with six taxa to keep
@@
 given phylogeny, as described in subsection~\ref{subsec:pipeline}.
-In this example, the root sequence consists of four blocks of 250 mononucleotides each, 
-where the per-nucleotide mutation rate is 0.1 mutations per unit time.
+In this example, the root sequence consists of four blocks of 250 mononucleotides each, while the per-nucleotide mutation rate is 0.1 mutations per unit time.
 We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
@@
 As the site and clock models used here are also the defaults, 
-the function arguments can be safely omitted: we just show these explicitly for clarity.
+the function arguments can be safely omitted: we just explicitly show them for the sake of clarity.
 
@@
 In this example we pick $\mathit{I}$ to be the same inference model as the generative one,
-which is the chosen Yule tree prior $\mathit{p_{G}^0}$, as well as site and clock 
-models defined in $\mathit{A}$, respectively JC and strict.
+which is the chosen Yule tree prior $\mathit{p_{G}^0}$ as well as site and clock models defined in $\mathit{A}$, respectively JC and strict.
 We specify in $\mathit{C}$ that the experiment will always be run.
@@
 We also need to specify the error measurement parameters $\mathit{E}$.
-Here we use the default error measurement
-parameters, which uses a burn-in fraction of 10\% and the nLTT statistic to
+Here we use the default $\mathit{E}$, which uses a burn-in fraction of 10\% and the nLTT statistic to
 measure the difference between phylogenies. For clarity,
@@
 
-We now have create the alignment parameters,
-our experiments and our error measure parameters.
-These objects need to be bundled in a bigger
-\verb;pirouette; parameter structure, 
-using \verb;create_pir_params;:
+We now have all the needed \verb;pirouette; parameters: the alignment parameters, the experiments and the error measure parameters.
+These objects need to be bundled in a bigger parameter structure, using \verb;create_pir_params;:
 
@@
 
-Now we can use our Yule tree and \verb;pir_params; to measure 
-the inference error made on phylogenies
+We can finally use the given Yule tree and \verb;pir_params; to measure the inference error made on phylogenies
 created by a standard diversification model:
@@
     The inference error made 
-    when the generative and inference model are the same.
+    when the generative and inference models are the same.
     The vertical dashed line indicates the median error.
@@
 Obtaining this control is in fact an inherent part of \verb;pirouette;. 
-We call it 'twinning' and we use it in the last part of this example.
+We call it 'twinning' and we will use it starting from \ref{Comparing to background noise}.
 
@@
 with a known generative tree prior $\mathit{p_{G}^0}$.
-However, a novel tree prior $\mathit{p_{G}}$, by our definition, 
-is non-standard, and hence not part of a standard inference 
+However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, and hence not part of a standard inference 
 model (i.e. an inference model using a standard tree prior).
-In such a case, the question is which standard tree prior should be 
-used in the inference.
-In this example, we re-use our hand-picked inference model,
-but additionally pick a best inference model from a set of inference models.
-With 'best' inference model, we mean the model with the highest 
-evidence (also known as marginal likelihood), based on the alignment.
-We show the procedure using \verb;pirouette; to answer a second research 
-question: "What is the inference error made on a novel phylogeny when
-using the best inference model, in comparison to a hand-picked model?
+In such a case, the question is which standard tree prior should be used in the inference.
+In this example, not only we re-use our hand-picked inference model, but we also pick the best inference model from a set of inference models.
+With the 'best' inference model, we mean the model with the highest evidence (also known as marginal likelihood), based on the alignment.
+We show the procedure using \verb;pirouette; to answer a second research question: "What is the inference error made on a novel phylogeny when using the best inference model, in comparison to a hand-picked model?
 
@@
     \textbf{model\_type} &
-    \textbf{run\_if} & \textbf{do\_measure\_evidence} & 
+    \textbf{run\_if} &
+    \textbf{do\_measure\_evidence} & 
     \textbf{inference model} \\ 
@@
 
-We will use the same tree as generated in \ref{lst:create_yule_tree},
-as well as the same alignment parameters as shown in 
-Listing~\ref{lst:create_alignment_params}.
-
-We do need to specify a different set of experiments: we
+We will use the same tree as generated in \ref{lst:create_yule_tree}, as well as the same alignment parameters as shown in Listing~\ref{lst:create_alignment_params}.
+
+Here we specify a different set of experiments: we
 need to state that we already have an experiment for
-the generative model, as well as that we want all other inference
-models to compete. We call the competing models 'candidate models',
-as only the best will be used in the actual inference.
-As model selection is commonly performed on the full list of available 
-candidate models, \verb;pirouette; has a dedicated function for this choice:
-\verb;create_all_experiments; will create a full set of 40 experiments,
-containing the inference models of all combinations of 
-4 site models, 2 clock models and 5 tree priors. All we need to add is
-to exclude the inference model in the generative experiment:
+the generative model, as well as that we want all the other inference models to compete. We call the competing models 'candidate models'.
+As model selection is commonly performed on the full list of available candidate models, \verb;pirouette; has a dedicated function for this choice: \verb;create_all_experiments; creates a full set of 40 experiments, containing the inference models of all combinations of 4 site models, 2 clock models and 5 tree priors.
+All we need to add is to exclude the inference model in the generative experiment:
 
@@
 
-We combine the generative and all candidate models into one set of experiments:
+We combine the generative and all the candidate models into one set of experiments:
 
@@
     Create a collection of experiments, with 1
-    generative model, and 39 candidate models
+    generative model, and 39 candidate models.
   }
@@
 
-We can now create the complete
-\verb;pirouette; parameter set like this (which is the
-same as Listing~\ref{lst:create_pir_params}, but using the defaults):
+We can now create the complete \verb;pirouette; parameter set in the usual way (which is the same as Listing~\ref{lst:create_pir_params}, but using the defaults):
 
@@
 We run \verb;pirouette; (Listing~\ref{lst:pir_run}) 
-and plot the results (Listing~\ref{lst:pir_plot}),
-as shown in Figure~\ref{fig:example_6}.
+and plot the results (Listing~\ref{lst:pir_plot}) in Figure~\ref{fig:example_6}.
 
@@
   \caption{
-    The inference error
-    for both the generative and best candidate inference models.
+    The inference error for both generative and best candidate inference models.
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparing to background noise}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Up until now, we measured the inference error on a tree
-of which we knew the (standard) tree prior. The goal of \verb;pirouette;
-is to measure the impact of a novel tree prior.
-To do so, in this example, we'll use a tree generated by a non-standard
-tree prior, of which we assume is closely related to the Yule tree prior.
-To measure the impact of the tree prior, we measure both the
-baseline error (that serves as a control) 
-and the full inference error: their difference shows
-the impact of the tree prior.
-To do so, we use a twin tree (see Subsection~\ref{subsec:twinning}) for
-both the generative and best candidate model.
+\subsection{Comparing to background noise}\label{Comparing to background noise}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+So far we have measured the inference error on a tree
+generated according to a known (and standard) tree prior. The goal of \verb;pirouette; is to measure the impact of a novel tree prior.
+To do so, in this example, we will use a tree generated by a non-standard tree prior, assumed to be closely related to the Yule tree prior.
+To measure the impact of the tree prior, we measure both the baseline error (that serves as a control) and the full inference error: their difference shows the impact of the tree prior.
+To do so, we use a twin tree (see Subsection~\ref{subsec:twinning}) for each of the generative and best candidate model.
 The research question this example answers is:
-"What is the inference error made from a phylogeny,
-for both a generative model and a best candidate model, 
-compared to the background noise?"
-
-This time we start with a tree generated from an unknown 
-diversification model, that has six taxa and a crown age of ten:
+"What is the inference error made from a phylogeny, for both a generative model and a best candidate model, compared to the background noise?"
+
+We start from the tree generated by a non-standard (and unknown) tree prior, having six taxa and a crown age of ten:
 \begin{lstlisting}[
@@
 
-Most of the other settings are the same as before:
-we re-use the alignment parameters (Listing~\ref{lst:create_alignment_params}),
-as well as the experiments (Listing~\ref{lst:create_all_experiments}).
+Most of the other settings are the same as before: we re-use the alignment parameters (Listing~\ref{lst:create_alignment_params}), as well as the experiments (Listing~\ref{lst:create_all_experiments}).
 This time, however, we enable twinning (see Subsection~\ref{subsec:twinning}),
@@
 Creating these parameters is trivial with the default settings.
-For clarity, however, we explicitly show the most
-important arguments:
+For clarity, however, we explicitly show the most important arguments:
 
@@
 We run \verb;pirouette; (Listing~\ref{lst:pir_run}) 
-and plot the results (Listing~\ref{lst:pir_plot}),
-The output is shown in Figure~\ref{fig:example_5}
+and plot the results (Listing~\ref{lst:pir_plot}).
+The output is shown in Figure~\ref{fig:example_5}.
 
@@
     compared with the error obtained for the twin tree.
-    Here, the 'twin' tree shows the baseline inference error
+    Here, the 'twin' tree shows the baseline inference error.
   }
@@
 \end{figure}
+\giovanni{I find the figure to be too confused. What if we use instead different plots, and in each of them we just show one inference model's results plus its twin counterpart?}
 
@@
 
-We showed how to use \verb;pirouette; 
-to quantify the importance of a tree prior in Bayesian phylogenetics,
-using the most simple generative tree prior possible.
-In principle any other (more complex) generative tree prior 
-can be tested, but we chose to provide the 
-simplest (and fastest to run) examples.
+We showed how to use \verb;pirouette; to quantify the importance of a tree prior in Bayesian phylogenetics, using the simplest generative tree prior possible.
+In principle any other (more complex) generative tree prior can be tested, but we chose to provide the simplest (and fastest to run) examples.
 All tests were performed on a single original tree, 
@@
 
-However, if the same procedure were repeated and performed on a distribution 
-with a sufficient number of generative trees, 
-it constitutes a quantitative assessment 
-of the quality of the inference.
+However, if the same procedure were repeated and performed on a distribution with a sufficient number of generative trees, it would constitute a quantitative and effective assessment of the quality of the inference.
 
@@
 94\% of all twin trees resulted in an error distribution with a lower median.
+\giovanni{If we write this some reviewer could argue: where do you see this? Do you have a plot to show it? Why don't you present such data about a bigger distribution, at least in an appendix?}
 
@@
 
+\iffalse
+\giovanni{We can probably just remove this.}
 To give an idea of the insights \verb;pirouette; can give,
@@
 We ...
-
-In conclusion, \verb;pirouette; can directly show 
-the comparison between the results obtained 
+\fi
+
+In conclusion, \verb;pirouette; can directly show the comparison between the results obtained 
 adopting a new generative prior as compared to the Bayesian inference error
@@
 tailored on the generative process, is needed. 
-If this were indeed the case, 
-one could rightly implement their novel tree prior as an addition to 
+If this were indeed the case, one could rightly implement their novel tree prior as an addition to 
 their favorite Bayesian inference tool.

### 2019-05-16 content.tex
Use 'workflow.png' instead of keeping 'workflow[version].png', git can control versions
Processed Rampal's feedback, fix #46

@@
   \centering
-  \includegraphics[width = \textwidth]{workflow3.png}
+  \includegraphics[width = \textwidth]{workflow.png}
   \caption{
@@
 An optional step is to generate a 'twin tree' $\tau_{I}$
-(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
-that will be analyzed in the same way as the true tree.
+(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b)
+and 'twin alignment'
+that will be analyzed in the same way as the true tree (and alignment).
 We define a phylogeny $\tau$ is the combination of
@@
 
+From the twin tree, we simulate an alignment following the same
+history as the twin tree, with the addition that 
+the same total number of mutations have
+taken place from root sequence to the sequences in the present.
+This addition assures there is the same amount of historical information
+stored in the alignments of both the true and twin alignment.
+We achieve this by simply simulating twin alignments up until we
+obtain one that has the desired number of mutations.
+
 Using a twin tree serves as a control: both $\tau_{\mathit{G}}$ and 
@@
 \end{table}
-\giovanni{I think the reference to the listings is a great idea. I don't know what to add more to this table. Probably (not 100\% sure) we should make all tables fit within the page limits.}
+\giovanni{Probably (not 100\% sure) we should make all tables fit within the page limits.}
+\richel{Nope, the publisher will do this for us in the end}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     The inference error for both generative and best candidate inference models.
+    Vertical dashed lines show the median error value per distribution.
   }
@@
     Here, the 'twin' tree shows the baseline inference error.
+    Vertical dashed lines show the median error value per distribution.
   }
@@
 \end{figure}
-\giovanni{I find the figure to be too confused. What if we use instead different plots, and in each of them we just show one inference model's results plus its twin counterpart?}
+\giovanni{
+  I find the figure to be too confused. 
+  What if we use instead different plots, 
+  and in each of them we just show one inference model's results plus its twin counterpart?
+}
+\richel{Will do}
 
@@
 
-However, if the same procedure were repeated and performed on a distribution with a sufficient number of generative trees, it would constitute a quantitative and effective assessment of the quality of the inference.
-
-A twin tree does not always result in a lower error distribution,
+However, if the same procedure were repeated and performed on a distribution with a sufficient number of generative trees, 
+it would constitute a quantitative and effective assessment of the quality of the inference.
+Also a twin tree does not always result in a lower error distribution,
 due to the stochasticity in generating the twin tree.
-For example, of the second tree used in the 'Usage' section,
-94\% of all twin trees resulted in an error distribution with a lower median.
-\giovanni{If we write this some reviewer could argue: where do you see this? Do you have a plot to show it? Why don't you present such data about a bigger distribution, at least in an appendix?}
-
-% Add: the effect of the alignment on this
-% https://github.com/richelbilderbeek/pirouette_article/issues/45
-
-\iffalse
-\giovanni{We can probably just remove this.}
-To give an idea of the insights \verb;pirouette; can give,
-we pretend the final result was to be obtained from thousands of replicates.
-We ...
-\fi
 

### 2019-05-16 supplement.tex
Renamed 'Generative and Best Candidate' setup to 'example_2' (from 'example_6')

@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_6/true_tree.png}
+      \includegraphics[height=0.2\textheight]{example_2/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_6/true_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_2/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_6/true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_2/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_6/true_error_violin_gen.png}
+      \includegraphics[height=0.2\textheight]{example_2/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_6/true_posterior_best.png}
+      \includegraphics[height=0.2\textheight]{example_2/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_6/true_error_violin_best.png}
+      \includegraphics[height=0.2\textheight]{example_2/true_error_violin_best.png}
     };   
@@
   }
-  \label{fig:example_6_full_pipeline}
+  \label{fig:example_2_full_pipeline}
   \caption{Example 6: full pipeline}
@@
 
-\input{example_6/esses_gen.latex}
-% has label tab:esses_example_6_gen
+\input{example_2/esses_gen.latex}
+% has label tab:esses_example_2_gen
 
-\input{example_6/esses_best.latex}
-% has label tab:esses_example_6_best
+\input{example_2/esses_best.latex}
+% has label tab:esses_example_2_best
 
-\input{example_6/evidence_true.latex}
-% has label tab:evidences_example_6
+\input{example_2/evidence_true.latex}
+% has label tab:evidences_example_2
 

### 2019-05-16 content.tex
Renamed 'Generative and Best Candidate' setup to 'example_2' (from 'example_6')
Renamed 'Generative and Best Candidate and Twinning' setup to 'example_3' (from 'example_5')

@@
 We run \verb;pirouette; (Listing~\ref{lst:pir_run}) 
-and plot the results (Listing~\ref{lst:pir_plot}) in Figure~\ref{fig:example_6}.
+and plot the results (Listing~\ref{lst:pir_plot}) in Figure~\ref{fig:example_2}.
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_6/errors.png}
+  \includegraphics[width=\textwidth]{example_2/errors.png}
   \caption{
@@
   }
-  \label{fig:example_6}
+  \label{fig:example_2}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_6/true_tree.png}
+  \includegraphics[width=\textwidth]{example_2/true_tree.png}
   \caption{The tree derived from an unknown diversification process, 
@@
 and plot the results (Listing~\ref{lst:pir_plot}).
-The output is shown in Figure~\ref{fig:example_5}.
+The output is shown in Figure~\ref{fig:example_3}.
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_5/errors.png}
+  \includegraphics[width=\textwidth]{example_3/errors.png}
   \caption{
@@
   }
-  \label{fig:example_5}
+  \label{fig:example_3}
 \end{figure}

### 2019-05-16 supplement.tex
Renamed 'Generative and Best Candidate and Twinning' setup to 'example_3' (from 'example_5')

@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_tree.png}
+      \includegraphics[height=0.2\textheight]{example_3/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5/true_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_3/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_3/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_error_violin_gen.png}
+      \includegraphics[height=0.2\textheight]{example_3/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_posterior_best.png}
+      \includegraphics[height=0.2\textheight]{example_3/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/true_error_violin_best.png}
+      \includegraphics[height=0.2\textheight]{example_3/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.2\textheight]{example_5/twin_tree.png}
+      \includegraphics[height=0.2\textheight]{example_3/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_5/twin_alignment.png}
+      \includegraphics[height=0.13\textheight]{example_3/twin_alignment.png}
     };   
     \node[state] (CTG) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/twin_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_3/twin_posterior_gen.png}
     };   
@@
 
-\includegraphics[height=0.2\textheight]{example_5/twin_error_violin_gen.png}
+\includegraphics[height=0.2\textheight]{example_3/twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_5/twin_posterior_best.png}
+      \includegraphics[height=0.2\textheight]{example_3/twin_posterior_best.png}
     };   
@@
 
-\includegraphics[height=0.2\textheight]{example_5/twin_error_violin_best.png}
+\includegraphics[height=0.2\textheight]{example_3/twin_error_violin_best.png}
     };   
@@
   }
-  \label{fig:example_5_full_pipeline}
+  \label{fig:example_3_full_pipeline}
   \caption{Example 5: full pipeline}
@@
 
-\input{example_5/esses_gen.latex}
-% has label tab:esses_example_5_gen
+\input{example_3/esses_gen.latex}
+% has label tab:esses_example_3_gen
 
-\input{example_5/esses_best.latex}
-% has label tab:esses_example_5_best
+\input{example_3/esses_best.latex}
+% has label tab:esses_example_3_best
 
-\input{example_5/esses_twin_gen.latex}
-% has label tab:esses_example_5_twin_gen
+\input{example_3/esses_twin_gen.latex}
+% has label tab:esses_example_3_twin_gen
 
-\input{example_5/esses_twin_best.latex}
-% has label tab:esses_example_5_twin_best
+\input{example_3/esses_twin_best.latex}
+% has label tab:esses_example_3_twin_best
 
-\input{example_5/evidence_true.latex}
-% has label tab:evidences_example_5
+\input{example_3/evidence_true.latex}
+% has label tab:evidences_example_3
 
-\input{example_5/evidence_twin.latex}
-% has label tab:evidences_example_5
+\input{example_3/evidence_twin.latex}
+% has label tab:evidences_example_3
 

### 2019-05-21 content.tex
Review

@@
   \textbf{1. }
-    Phylogenetic trees are current routinely reconstructed from an alignment of character sequences (usually nucleotide sequences).
-Bayesian tools, such as MrBayes, RevBayes and BEAST2, have gained much popularity over the last decade, as they allow joint estimation of the posterior distribution of the phylogenetic tree and the parameters of the underlying inference model.
-    An important ingredient of these Bayesian approaches is the species tree prior.
-While in principle the Bayesian framework allows for comparing different species tree priors and hence may elucidate the macroevolutionary processes underlying the species tree, in practice only
-    macroevolutionary models that allow for fast computation of the prior probability are used. 
-    This begs the question how accurate the tree estimation is
+    Phylogenetic trees are current routinely reconstructed from an alignment 
+    of character sequences (usually nucleotide sequences). 
+    Bayesian tools, such as MrBayes, RevBayes and BEAST2, 
+    have gained much popularity over the last decade, 
+    as they allow joint estimation of the posterior distribution of the 
+    phylogenetic tree and the parameters of the underlying inference model.  
+    An important ingredient of these Bayesian approaches is the species tree 
+    prior. While in principle the Bayesian framework allows for comparing 
+    different species tree priors and hence may elucidate the macroevolutionary 
+    processes underlying the species tree, in practice only 
+    macroevolutionary models that allow for fast computation of the prior 
+    probability are used. 
+    An open question is, how accurate the tree estimation is
     when the real macroevolutionary processes are substantially different 
@@
     a free, libre and open-source R package that assesses 
-    the inference error made by Bayesian phylogenetics for a given macroevolutionary diversification model.
-\verb;pirouette; makes use of BEAST2, but its philosophy applies to any Bayesian phylogenetic inference tool\\
+    the inference error made by Bayesian phylogenetics for a given 
+    macroevolutionary diversification model. \verb;pirouette; makes use of 
+    BEAST2, but its philosophy applies to any Bayesian phylogenetic inference 
+    tool. \\
   \textbf{3. }
@@
     Pace at which mutation occurs &
-    mutation rate $\in \mathbb{R}_{>0}$\\
+    \verb;mutation_rate; $\in \mathbb{R}_{>0}$\\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{table}
-\giovanni{Probably (not 100\% sure) we should make all tables fit within the page limits.}
-\richel{Nope, the publisher will do this for us in the end}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-05-21 supplement.tex
Review

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results in detail}
+\section{Supplementary materials}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Errors}
+\subsection{Usage with generative model}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 1}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 5}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Usage with generative and candidate models}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
+\begin{figure}[H]
   \centering
@@
       ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.4\textheight, 
+      node distance=0.7\textheight, 
       semithick
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_tree.png}
+      \includegraphics[height=0.5\textheight]{example_2/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3/true_alignment.png}
+      \includegraphics[height=0.5\textheight]{example_2/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_posterior_gen.png}
+      \includegraphics[height=0.45\textheight]{example_2/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{example_2/true_error_violin_gen.png}
     };   
-    \node[state] (CB) [right of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_posterior_best.png}
+    \node[state] (CB) [right of = CG, rectangle] {
+      \includegraphics[height=0.45\textheight]{example_2/true_posterior_best.png}
+    };   
+    \node[state] (DB) [right of = DG, rectangle] {
+      \includegraphics[height=0.3\textheight]{example_2/true_error_violin_best.png}
+    };   
+    \path 
+      (O) edge [anchor = south] node {} (A)
+      (A) edge [anchor = south] node {} (B)
+      (B) edge [anchor = south] node {} (CG)
+      (CG) edge [anchor = south] node {} (DG)
+      (B) edge [anchor = south] node {} (CB)
+      (CB) edge [anchor = south] node {} (DB)
+    ; 
+    \end{tikzpicture}
+  }
+  \label{fig:example_2_full_pipeline}
+  \caption{Usage with generative and candidate models: full pipeline}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\input{example_2/esses_gen.latex}
+% has label tab:esses_example_2_gen
+
+\input{example_2/esses_best.latex}
+% has label tab:esses_example_2_best
+
+\input{example_2/evidence_true.latex}
+% has label tab:evidences_example_2
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Usage with generative and candidate models, compared to the background noise}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[H]
+  \centering
+  \resizebox {1.0\columnwidth} {!} {
+    \begin{tikzpicture}[
+      ->,>=stealth',shorten >=1pt,auto,
+      node distance=0.5\textheight, 
+      semithick
+    ]   
+    \tikzstyle{every state}=[]
+    \node[state, draw=none] (O) [] {
+    };   
+    \node[state] (A) [right of = O, rectangle] {
+      \includegraphics[height=0.4\textheight]{example_3/true_tree.png}
+    };   
+    \node[state] (B) [below of = A, rectangle] {
+      \includegraphics[height=0.25\textheight]{example_3/true_alignment.png}
+    };   
+    \node[state] (CG) [below of = B, rectangle] {
+      \includegraphics[height=0.3\textheight]{example_3/true_posterior_gen.png}
+    };   
+    \node[state] (DG) [below of = CG, rectangle] {
+      \includegraphics[height=0.3\textheight]{example_3/true_error_violin_gen.png}
+    };   
+    \node[state] (CB) [right of = CG, rectangle] {
+      \includegraphics[height=0.3\textheight]{example_3/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/true_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{example_3/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.2\textheight]{example_3/twin_tree.png}
+      \includegraphics[height=0.4\textheight]{example_3/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_3/twin_alignment.png}
+      \includegraphics[height=0.25\textheight]{example_3/twin_alignment.png}
     };   
-    \node[state] (CTG) [below of = BT, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/twin_posterior_gen.png}
+    \node[state] (CTG) [right of = CB, rectangle] {
+      \includegraphics[height=0.3\textheight]{example_3/twin_posterior_gen.png}
     };   
     \node[state] (DTG) [below of = CTG, rectangle] {
-
-\includegraphics[height=0.2\textheight]{example_3/twin_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{example_3/twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_3/twin_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{example_3/twin_posterior_best.png}
     };   
     \node[state] (DTB) [below of = CTB, rectangle] {
-
-\includegraphics[height=0.2\textheight]{example_3/twin_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{example_3/twin_error_violin_best.png}
     };   
@@
   \label{fig:example_3_full_pipeline}
-  \caption{Example 5: full pipeline}
+  \caption{Usage with generative and candidate models, compared to the background noise: full pipeline}
 \end{figure}
@@
 
-\begin{table}
-  \begin{tabular}{ | r | c | }
-    \hline
-    \textbf{ERQ} & \textbf{run-time} \\ 
-    \hline
-      1 & 00:10:12 \\
-      2 & 00:34:10 \\
-      3 & 00:24:53 \\
-      4 & 01:04:50 \\
-      5 & 01:29:40 \\
-    \hline
-  \end{tabular}
-  \caption{
-    Time for an example research question to finish.
-    Time is in in hh:mm::ss.
-  }
-  \label{tab:runtime}
-\end{table}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Example 6}
+\subsection{Runtime}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[ht]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.4\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.13\textheight]{example_2/true_alignment.png}
-    };   
-    \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_posterior_gen.png}
-    };   
-    \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_error_violin_gen.png}
-    };   
-    \node[state] (CB) [right of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_posterior_best.png}
-    };   
-    \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_2/true_error_violin_best.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (CG)
-      (CG) edge [anchor = south] node {} (DG)
-      (B) edge [anchor = south east] node {} (CB)
-      (CB) edge [anchor = south] node {} (DB)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_2_full_pipeline}
-  \caption{Example 6: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{example_2/esses_gen.latex}
-% has label tab:esses_example_2_gen
-
-\input{example_2/esses_best.latex}
-% has label tab:esses_example_2_best
-
-\input{example_2/evidence_true.latex}
-% has label tab:evidences_example_2
 

### 2019-05-21 content.tex
Add note for Rampal

@@
 
+\richel{New section about equal number of mutations}
 From the twin tree, we simulate an alignment following the same

### 2019-06-03 content.tex

@@
 that will be analyzed in the same way as the true tree (and alignment).
-We define a phylogeny $\tau$ is the combination of
+We define a phylogeny $\tau$ as the combination of
 branching times $\Vec{t}$ and topology $\psi$.
@@
 topology $\psi_{\mathit{G}}$:
-
 \begin{align}
@@
 \end{align}
-
 To achieve this, we find the parameters $\theta^{*}_{\mathit{I}}$ 
@@
 associated with the chosen inference model $\mathit{I}$:
-
 \begin{align}
-    max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{G}}, n_{\mathit{G}})] 
+    \max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{G}}, n_{\mathit{G}})] 
 \rightarrow \theta^{*}_{\mathit{I}}.
 \end{align}
-
 We use $\theta^{*}_{\mathit{I}}$ to simulate an equal number 
@@
 This addition assures there is the same amount of historical information
-stored in the alignments of both the true and twin alignment.
-We achieve this by simply simulating twin alignments up until we
+stored in both the true and twin alignments.
+We achieve this by simply simulating twin alignments until we
 obtain one that has the desired number of mutations.
 
-Using a twin tree serves as a control: both $\tau_{\mathit{G}}$ and 
-$\tau_{\mathit{I}}$ will produce an error distribution. 
-These errors unavoidably accumulate across the \verb;pirouette; pipeline,
-for both the true and twin tree, due to several possible reasons: either tree may be an unlikely representative
-of the underlying tree prior, the alignment simulation may give different
-numbers of mutations for the same evolutionary time, 
-and the MCMC will sample different states.
-These errors can be equalized by imposing the same tree topology and by
-replicating the pipelines a sufficient number of times.
-
-The remaining differences in error distributions will be 
-consequences of different phylogenies and tree priors for the two parallel 
+The twin tree serves as a control: even when the generating and inference models are identical, the posterior distribution of inferred trees will differ from the true tree due to stochasticity in producing an alignment and in MCMC sampling of the posterior; the twin tree provides this minimum error.
+We therefore need to compare the error distributions of $\tau_{\mathit{G}}$ and
+$\tau_{\mathit{I}}$. 
+By imposing the same topology and the same number of mutations in the alignments, the differences in error distributions of the true tree and the twin tree are the consequence of different phylogenies (produced by different processes) and tree priors for the two parallel
 pipelines: the main pipeline uses the tree $\tau_{\mathit{G}}$ and a
-possibly mismatched prior $\mathit{G}$, where the twin pipeline
+possibly mismatched prior $\mathit{G}$, whereas the twin pipeline
 uses the tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
@@
 
-\verb;pirouette; will be made available on CRAN from which it can then be easily installed easily:
+\verb;pirouette; will be made available on CRAN from which it can then be easily installed:
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]

### 2019-06-12 supplement.tex
Use figures from example 6

@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{example_3/true_tree.png}
+      \includegraphics[height=0.4\textheight]{example_6/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.25\textheight]{example_3/true_alignment.png}
+      \includegraphics[height=0.25\textheight]{example_6/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_3/true_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{example_6/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_3/true_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{example_6/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_3/true_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{example_6/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_3/true_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{example_6/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.4\textheight]{example_3/twin_tree.png}
+      \includegraphics[height=0.4\textheight]{example_6/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.25\textheight]{example_3/twin_alignment.png}
+      \includegraphics[height=0.25\textheight]{example_6/twin_alignment.png}
     };   
     \node[state] (CTG) [right of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_3/twin_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{example_6/twin_posterior_gen.png}
     };   
     \node[state] (DTG) [below of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_3/twin_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{example_6/twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_3/twin_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{example_6/twin_posterior_best.png}
     };   
     \node[state] (DTB) [below of = CTB, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_3/twin_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{example_6/twin_error_violin_best.png}
     };   
@@
 
-\input{example_3/esses_gen.latex}
+\input{example_6/esses_gen.latex}
 % has label tab:esses_example_3_gen
 
-\input{example_3/esses_best.latex}
+\input{example_6/esses_best.latex}
 % has label tab:esses_example_3_best
 
-\input{example_3/esses_twin_gen.latex}
+\input{example_6/esses_twin_gen.latex}
 % has label tab:esses_example_3_twin_gen
 
-\input{example_3/esses_twin_best.latex}
+\input{example_6/esses_twin_best.latex}
 % has label tab:esses_example_3_twin_best
 
-\input{example_3/evidence_true.latex}
+\input{example_6/evidence_true.latex}
 % has label tab:evidences_example_3
 
-\input{example_3/evidence_twin.latex}
+\input{example_6/evidence_twin.latex}
 % has label tab:evidences_example_3

### 2019-06-13 content.tex
Show result of example_6

@@
 For example, a protracted birth-death model reduces to a standard birth-death model when speciation-completion rate approaches infinity [i.e.
-rate $\lambda$ in \cite{etienne2014estimating}, eqs.
-(2b) and (2c)].
+rate $\lambda$ in \cite{etienne2014estimating}, eqs.~(2b) and (2c)].
 When benchmarking a novel tree prior, 
@@
 
-\richel{New section about equal number of mutations}
 From the twin tree, we simulate an alignment following the same
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_2/true_tree.png}
+  \includegraphics[width=\textwidth]{example_6/true_tree.png}
   \caption{The tree derived from an unknown diversification process, 
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_3/errors.png}
+  \includegraphics[width=\textwidth]{example_6/errors.png}
   \caption{

### 2019-06-13 content.tex
Revised Examples

@@
 its features.
-First, to get an idea of the baseline error, we measure the Bayesian inference error when we start from a phylogeny generated under a known and standard tree prior.
-Second, to establish that the true tree prior is indeed the best, we also measure the inference error made adopting a
+First, to get an idea of the baseline error, 
+we measure the Bayesian inference error when we start from a phylogeny 
+generated under a known and standard tree prior.
+Second, to establish that the true tree prior is indeed the best, 
+we also measure the inference error made adopting a
 best candidate inference model.
 Lastly, we quantify the impact of the tree prior in the Bayesian inference. 
-We do so by separating the baseline error and the complete error running the pipeline from a tree generated by an unknown or non-standard tree prior.
+We do so by separating the baseline error from the complete error 
+running the pipeline starting from a tree generated by an unknown 
+or non-standard tree prior.
 
@@
 
-\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's inference by measuring the discrepancy between a given tree and a posterior distribution of phylogenies obtained as result of the inference process.
-Due to stochasticity, posterior trees will generally differ from the given phylogeny $\tau_{\mathit{G}}$, even when the generative tree prior $\mathit{p_{G}}$ and alignment model $\mathit{A}$ are the same as those in the inference model $\mathit{I}$.
+\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's 
+inference by measuring the discrepancy between a given tree and a posterior 
+distribution of phylogenies obtained as result of the inference process. 
+Due to stochasticity, posterior trees will generally differ from the given 
+phylogeny $\tau_{\mathit{G}}$, even when the generative tree 
+prior $\mathit{p_{G}}$ and alignment model $\mathit{A}$ are the same as those 
+in the inference model $\mathit{I}$.
 Measuring this difference allows us to know the baseline error
-of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' all the tree priors that are available to be used within an inference model (see Table~\ref{tab:options}).
-
-Now we can formulate the first example research question that \verb;pirouette; can answer: "What is the inference error made on phylogenies created by a standard diversification model?"
+of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' 
+all the tree priors that are available to be used within an inference model (see Table~\ref{tab:options}).
+
+Now we can formulate the first example research question that \verb;pirouette; 
+can answer: "What is the inference error made on phylogenies created by a 
+standard diversification model?"
 
 In this example we use a standard generative tree prior $\mathit{p_{G}^0}$, namely the Yule (pure-birth) tree prior. 
-We choose to use a small tree with six taxa to keep
+We choose to use a small tree with six taxa, to keep
 the calculations short and the figure more readable.
@@
 The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny, as described in subsection~\ref{subsec:pipeline}.
-In this example, the root sequence consists of four blocks of 250 mononucleotides each, while the per-nucleotide mutation rate is 0.1 mutations per unit time.
+given phylogeny, as described in Subsection~\ref{subsec:pipeline}.
+In this example, the root sequence consists of four blocks of 250 
+mononucleotides each, while the per-nucleotide mutation rate is 
+0.1 mutations per unit time.
 We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
 and a strict clock model as these are the simplest.
-A JC site model assumes that mutation rates between nucleotides are equal and constant. 
+A JC site model assumes that mutation rates between nucleotides are equal and 
+constant. 
 A strict clock model assumes that the mutation rates 
@@
 As the site and clock models used here are also the defaults, 
-the function arguments can be safely omitted: we just explicitly show them for the sake of clarity.
+the function arguments can be safely omitted: we just explicitly show 
+them for the sake of clarity.
 
@@
 We define an experiment $\mathit{X}$ as a combination of an inference model 
-$\mathit{I}$
-and conditions $\mathit{C}$.
-In this example we pick $\mathit{I}$ to be the same inference model as the generative one,
-which is the chosen Yule tree prior $\mathit{p_{G}^0}$ as well as site and clock models defined in $\mathit{A}$, respectively JC and strict.
+$\mathit{I}$ and conditions $\mathit{C}$.
+In this example we pick $\mathit{I}$ to be the same inference model as 
+the generative one,
+which is the chosen Yule tree prior $\mathit{p_{G}^0}$ as well as site 
+and clock models defined in $\mathit{A}$, respectively JC and strict.
 We specify in $\mathit{C}$ that the experiment will always be run.
@@
 We also need to specify the error measurement parameters $\mathit{E}$.
-Here we use the default $\mathit{E}$, which uses a burn-in fraction of 10\% and the nLTT statistic to
+Here we use the default $\mathit{E}$, which uses a burn-in fraction 
+of 10\% and the nLTT statistic to
 measure the difference between phylogenies. For clarity,
@@
 
-We now have all the needed \verb;pirouette; parameters: the alignment parameters, the experiments and the error measure parameters.
-These objects need to be bundled in a bigger parameter structure, using \verb;create_pir_params;:
+We now have all the needed \verb;pirouette; parameters: the alignment 
+parameters, the experiments and the error measure parameters.
+These objects need to be bundled in a bigger parameter structure, 
+using \verb;create_pir_params;:
 
@@
 
-We can finally use the given Yule tree and \verb;pir_params; to measure the inference error made on phylogenies
+We can finally use the given Yule tree and \verb;pir_params; to measure the 
+inference error made on phylogenies
 created by a standard diversification model:
@@
 Obtaining this control is in fact an inherent part of \verb;pirouette;. 
-We call it 'twinning' and we will use it starting from \ref{Comparing to background noise}.
+We call it 'twinning' and we will demonstrate it at section
+ \ref{Comparing to background noise}.
 
@@
 with a known generative tree prior $\mathit{p_{G}^0}$.
-However, a novel tree prior $\mathit{p_{G}}$, by our definition, is non-standard, and hence not part of a standard inference 
+However, a novel tree prior $\mathit{p_{G}}$, by our definition, is 
+non-standard, and hence not part of a standard inference 
 model (i.e. an inference model using a standard tree prior).
-In such a case, the question is which standard tree prior should be used in the inference.
-In this example, not only we re-use our hand-picked inference model, but we also pick the best inference model from a set of inference models.
-With the 'best' inference model, we mean the model with the highest evidence (also known as marginal likelihood), based on the alignment.
-We show the procedure using \verb;pirouette; to answer a second research question: "What is the inference error made on a novel phylogeny when using the best inference model, in comparison to a hand-picked model?
+In such a case, the question is which standard tree prior should be used in 
+the inference. In this example, not only we re-use our hand-picked 
+inference model, but we also pick the best inference model from a set of 
+inference models. With the 'best' inference model, we mean the model with 
+the highest evidence (also known as marginal likelihood), based on the 
+alignment. We show the procedure using \verb;pirouette; to answer a second 
+research question: "What is the inference error made on a novel phylogeny when using the best inference model, in comparison to a hand-picked model?
 
@@
 
-We will use the same tree as generated in \ref{lst:create_yule_tree}, as well as the same alignment parameters as shown in Listing~\ref{lst:create_alignment_params}.
+We will use the same tree as generated in \ref{lst:create_yule_tree}, as well 
+as the same alignment parameters as shown 
+in Listing~\ref{lst:create_alignment_params}.
 
@@
 need to state that we already have an experiment for
-the generative model, as well as that we want all the other inference models to compete. We call the competing models 'candidate models'.
-As model selection is commonly performed on the full list of available candidate models, \verb;pirouette; has a dedicated function for this choice: \verb;create_all_experiments; creates a full set of 40 experiments, containing the inference models of all combinations of 4 site models, 2 clock models and 5 tree priors.
-All we need to add is to exclude the inference model in the generative experiment:
+the generative model, as well as that we want all the other inference models 
+to compete. We call the competing models 'candidate models'.
+As model selection is commonly performed on the full list of available 
+candidate models, \verb;pirouette; has a dedicated function for this choice: \verb;create_all_experiments; creates a full set of 40 experiments, 
+containing the inference models of all combinations of 4 site models, 
+2 clock models and 5 tree priors. All we need to add is to exclude the 
+inference model in the generative experiment:
 
@@
 
-We combine the generative and all the candidate models into one set of experiments:
+We combine the generative and all the candidate models into one set of 
+experiments:
 
@@
 
-We can now create the complete \verb;pirouette; parameter set in the usual way (which is the same as Listing~\ref{lst:create_pir_params}, but using the defaults):
+We can now create the complete \verb;pirouette; parameter set in the usual 
+way (which is the same as Listing~\ref{lst:create_pir_params}, but using the 
+defaults):
 
@@
 So far we have measured the inference error on a tree
-generated according to a known (and standard) tree prior. The goal of \verb;pirouette; is to measure the impact of a novel tree prior.
-To do so, in this example, we will use a tree generated by a non-standard tree prior, assumed to be closely related to the Yule tree prior.
-To measure the impact of the tree prior, we measure both the baseline error (that serves as a control) and the full inference error: their difference shows the impact of the tree prior.
-To do so, we use a twin tree (see Subsection~\ref{subsec:twinning}) for each of the generative and best candidate model.
+generated according to a known (and standard) tree prior. 
+The goal of \verb;pirouette; is to measure the impact of a novel tree prior.
+To do so, in this example, we will use a tree generated by a non-standard 
+tree prior, assumed to be closely related to the Yule tree prior.
+To measure the impact of the tree prior, we measure both the baseline 
+error (that serves as a control) and the full inference error: it is their 
+difference that shows the impact of the tree prior.
+To do so, we use a twin tree (see Subsection~\ref{subsec:twinning}) for 
+each of the generative and best candidate model.
 The research question this example answers is:
-"What is the inference error made from a phylogeny, for both a generative model and a best candidate model, compared to the background noise?"
+"What is the inference error made from a phylogeny, 
+for both a generative model and a best candidate model, compared to the 
+background noise?"
 
@@
 \end{figure}
-\giovanni{
-  I find the figure to be too confused. 
-  What if we use instead different plots, 
-  and in each of them we just show one inference model's results plus its twin counterpart?
-}
-\richel{Will do}
 
@@
 
-We showed how to use \verb;pirouette; to quantify the importance of a tree prior in Bayesian phylogenetics, using the simplest generative tree prior possible.
-In principle any other (more complex) generative tree prior can be tested, but we chose to provide the simplest (and fastest to run) examples.
-All tests were performed on a single original tree, 
-hence not providing us enough statistical power to 
-evaluate the level of accuracy of the inference.
-
-However, if the same procedure were repeated and performed on a distribution with a sufficient number of generative trees, 
-it would constitute a quantitative and effective assessment of the quality of the inference.
+We showed how to use \verb;pirouette; to quantify the importance of a 
+tree prior in Bayesian phylogenetics, using the simplest generative tree 
+prior possible.
+In principle any other (more complex) generative tree prior can be tested, 
+but we chose to provide the simplest (and fastest to run) examples.
+All examples used only one original tree,
+where any speciation process produces a whole range of trees.
+One tree is not enough to determine the impact of a tree prior 
+on Bayesian inference.
+However, if the same procedure were repeated and performed on a 
+distribution with a sufficient number of generative trees, 
+it would constitute a quantitative and effective assessment of the 
+quality of the inference.
 Also a twin tree does not always result in a lower error distribution,
@@
 
-In conclusion, \verb;pirouette; can directly show the comparison between the results obtained 
+In conclusion, \verb;pirouette; can directly show the comparison 
+between the results obtained 
 adopting a new generative prior as compared to the Bayesian inference error
@@
 tailored on the generative process, is needed. 
-If this were indeed the case, one could rightly implement their novel tree prior as an addition to 
+If this were indeed the case, one could rightly implement 
+their novel tree prior as an addition to 
 their favorite Bayesian inference tool.

### 2019-06-13 content.tex
Moved interpretation to pirouette repo

@@
 
-\begin{table}
-  \begin{tabular}{ | r | c | c | c | }
-    \hline
-    \textbf{Compare with} & \textbf{TB} & \textbf{WG} & \textbf{WB} \\ 
-    \hline
-    \textbf{TG} & Y & Y & N \\
-    \textbf{TB} & . & N & Y \\
-    \textbf{WG} & . & . & Y \\
-    \textbf{WB} & . & . & . \\
-    \hline
-  \end{tabular}
-  \caption{
-    Error distributions that are relevant to compare.
-    TG: true tree, generative model.
-    TB: true tree, best candidate model.
-    WG: twin tree, generative model.
-    WB: twin tree, best candidate model.
-  }
-  \label{tab:relevant_comparisions}
-\end{table}
-
-\begin{table}
-  \begin{tabular}{ | r | c | p{9cm} | }
-    \hline
-    \textbf{Condition} & \textbf{Expectation} & \textbf{Interpretation} \\ 
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG > TB$}       & Unexpected & Novel tree prior is more related to 
-      best candidate than hand-picked tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG \approx TB$} & Possible   & Hand-picked tree prior is just as 
-      suitable as the best candidate tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG < TB$}       & Expected   & Hand-picked tree prior is the most 
-      related tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG > WG$}       & Unknown    & Novel tree prior important \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG \approx WG$} & Unknown    & Novel tree prior unimportant \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TG < WG$}       & Unexpected & Twinning procedure increases 
-      inference errors when using hand-picked tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TB > WB$}       & Expected   & Impact of novel tree prior cannot 
-      be compensated for by model selection: twin tree with low likelihood? \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TB \approx WB$} & Possible   & Best candidate tree priors perform 
-      equally well in true and twin tree: true and twin tree similar? \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$TB < WB$}       & Unexpected & Twinning procedure increases 
-      inference errors  when using best tree prior candidate \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$WG > WB$}       & Unexpected & Hand-picked tree prior (that equals 
-      the twinning tree prior!) worse than best candidate tree prior \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$WG \approx WB$} & Possible   & Twin tree fits equally well to the 
-      hand-picked and best candidate tree prior  \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \textbf{$WG < WB$}       & Expected   & Hand-pick tree prior (that equals 
-      the twinning tree prior) performs as expected \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \hline
-  \end{tabular}
-  \caption{
-    Interpretation of the different error distributions.
-    This assumes that the operators ($<$, $\approx$ and $>$) to compare
-    distributions are defined.
-    TG: true tree, generative model.
-    TB: true tree, best candidate model.
-    WG: true tree, generative model.
-    WB: true tree, best candidate model.
-  }
-  \label{tab:interpretation}
-\end{table}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-13 content.tex
Revised discussion

@@
 Also a twin tree does not always result in a lower error distribution,
-due to the stochasticity in generating the twin tree.
-
-In conclusion, \verb;pirouette; can directly show the comparison 
-between the results obtained 
-adopting a new generative prior as compared to the Bayesian inference error
-when operating on a tree generated under a known and standard tree 
-prior (the twin tree).
+as the stochasticity in generating a twin tree will occasionally
+yield a tree of very low likelihood.
+
+In conclusion, \verb;pirouette; can show the errors
+when adopting a new generative prior
+in comparison to the errors made by a standard prior.
 From such results an user can estimate whether or not a new tree prior, 

### 2019-06-13 content.tex
Update packages used
Code is checked to work

@@
     clock_model = create_strict_clock_model(), 
-    site_model = create_jc69_site_model()
+    site_model = create_jc69_site_model(),
+    mcmc = create_mcmc(store_every = 1000)
   )
@@
 ]
-experiments <- list(
+experiments <- c(
   list(generative_experiment),
@@
 twinning_params <- create_twinning_params(
-  twin_model = "bd", 
+  twin_model = "birth_death",
   method = "random_tree"
@@
 \verb;pirouette; depends on multiple packages, which are:
+\verb;ape; (\cite{ape}),
 \verb;babette; (\cite{bilderbeek2018babette}),
@@
 \verb;devtools; (\cite{devtools}),
+\verb;dplyr; (\cite{dplyr}),
 \verb;geiger; (\cite{geiger}),
@@
 \verb;lintr; (\cite{lintr}),
+\verb;magrittr; (\cite{magrittr}),
 \verb;mcbette; (\cite{mcbette}),
@@
 \verb;phytools; (\cite{phytools}),
+\verb;plyr; (\cite{plyr}),
 \verb;rappdirs; (\cite{rappdirs}),
@@
 \verb;stringr; (\cite{stringr}),
-\verb;subplex; (\cite{subplex}),
 \verb;TESS; (\cite{TESS}),

### 2019-06-13 content.tex
Link to examples

@@
 or adding code. To ensure a high quality, \verb;pirouette; 
-uses a continuous integration service, has a code coverage of 100\%
+uses a continuous integration service, has a code coverage of above 95\%
 and enforces the most commonly used R style guide (\cite{style_guide}).
@@
 The \verb;pirouette; vignette demonstrates extensively how 
-to use \verb;pirouette; in a more lightly written way.
+to use \verb;pirouette; in a more informally written way.
+
+The code used in this article and more examples that are periodically 
+tested to work, can be found at 
+\url{https://github.com/richelbilderbeek/pirouette}.
 

### 2019-06-13 supplement.tex
Link to examples

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Usage with generative model}
+\subsection{Generative model only}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_1}.
 
@@
   \label{fig:example_1_full_pipeline}
-  \caption{Example 1: full pipeline}
+  \caption{Generative model only: full pipeline}
 \end{figure}
@@
 \input{example_1/esses.latex}
-% has label tab:esses_example_1
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Usage with generative and candidate models}
+\subsection{Comparing to other candidate models}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_2}.
 
@@
   \label{fig:example_2_full_pipeline}
-  \caption{Usage with generative and candidate models: full pipeline}
+  \caption{Comparing to other candidate models: full pipeline}
 \end{figure}
@@
 \input{example_2/esses_gen.latex}
-% has label tab:esses_example_2_gen
 
 \input{example_2/esses_best.latex}
-% has label tab:esses_example_2_best
 
 \input{example_2/evidence_true.latex}
-% has label tab:evidences_example_2
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Usage with generative and candidate models, compared to the background noise}
+\subsection{Comparing to background noise}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_6}.
 
@@
   \label{fig:example_3_full_pipeline}
-  \caption{Usage with generative and candidate models, compared to the background noise: full pipeline}
+  \caption{Comparing to background noise: full pipeline}
 \end{figure}
@@
 \input{example_6/esses_gen.latex}
-% has label tab:esses_example_3_gen
 
 \input{example_6/esses_best.latex}
-% has label tab:esses_example_3_best
 
 \input{example_6/esses_twin_gen.latex}
-% has label tab:esses_example_3_twin_gen
 
 \input{example_6/esses_twin_best.latex}
-% has label tab:esses_example_3_twin_best
 
 \input{example_6/evidence_true.latex}
-% has label tab:evidences_example_3
 
 \input{example_6/evidence_twin.latex}
-% has label tab:evidences_example_3
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Runtime}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}
-  \begin{tabular}{ | r | c | }
-    \hline
-    \textbf{ERQ} & \textbf{run-time} \\ 
-    \hline
-      1 & 00:10:12 \\
-      2 & 00:34:10 \\
-      3 & 00:24:53 \\
-      4 & 01:04:50 \\
-      5 & 01:29:40 \\
-    \hline
-  \end{tabular}
-  \caption{
-    Time for an example research question to finish.
-    Time is in in hh:mm::ss.
-  }
-  \label{tab:runtime}
-\end{table}

### 2019-07-01 content.tex
Fix URL
Spellcheck (there were no spelling errors found)
Updates from Overleaf

@@
 Bayesian phylogenetic inference often requires that the prior probability of the phylogeny according to the diversification model has to be computed millions of times.
-Therefore, biologically interesting but computationally expensive tree priors are non-viable.
+Therefore, biologically interesting but computationally expensive tree priors are often not implemented, and simpler priors are used instead.
 This is not necessarily problematic, when the data are very informative, as this will reduce the influence of the tree prior.
@@
 Many tree priors have a parameter setting for which it reduces to a standard tree prior.
-For example, a protracted birth-death model reduces to a standard birth-death model when speciation-completion rate approaches infinity [i.e.
+For example, a protracted birth-death model \citep{etienne2012prolonging} reduces to a standard birth-death model when the speciation-completion rate is infinite [i.e.
 rate $\lambda$ in \cite{etienne2014estimating}, eqs.~(2b) and (2c)].
@@
 Although many possible tests can be performed, 
-we show the usage of \verb;pirouette; by a story that introduces
+we show the usage of \verb;pirouette; by introducing 
 its features gradually, yet ending in quantifying the impact a tree prior has
@@
   \item from this alignment, according to the specified inference conditions $\mathit{C}$, 
-    an inference model $\mathit{I}$ is picked (which may differ from the 
+    an inference model $\mathit{I}$ is chosen (which may differ from the 
     generative model);
@@
 we specified condition $C_{i}$ that a candidate model should only be run
-when being the best.
-A 'best' model is defined as the inference model with
+when it is the best.
+The 'best' model is defined as the inference model with
 the highest evidence (a.k.a. marginal likelihood), given the alignment 
 simulated in the previous step.
-The evidence of an inference model is estimated by nested 
+The evidence for an inference model is estimated by nested 
 sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. We note that scripted use of BEAST2 packages is only possible under Linux and Mac.
@@
 
-The third step infers the posteriors,
+The third step infers the posterior distributions,
 using the simulated alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 4a),
 and the inference models that were selected in the previous step (3). 
-For each selected experiment a posterior is inferred, using the 
+For each selected experiment a posterior distribution is inferred, using the 
 \verb;babette; [\cite{bilderbeek2018babette}] R package which makes use of BEAST2. 
-This step usually takes up most of the full pipeline's computation time.
+This step usually takes up most of the pipeline's computation time.
 
 The fourth step quantifies the inference error made. First the burn-in fraction is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
-in which it samples an unrepresentative part of state space. By default, \verb;pirouette; 
+which samples an unrepresentative part of state space. By default, \verb;pirouette; 
 removes the first 10\% of the posterior.
@@
 From the twin tree, we simulate an alignment following the same
-history as the twin tree, with the addition that 
+history as the twin tree, with the constraint that 
 the same total number of mutations have
-taken place from root sequence to the sequences in the present.
-This addition assures there is the same amount of historical information
+taken place.
+This constraint ensures there is the same amount of historical information
 stored in both the true and twin alignments.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Generative model only}
+\subsection{The generative and inference models are equal and standard.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 $\mathit{I}$ and conditions $\mathit{C}$.
-In this example we pick $\mathit{I}$ to be the same inference model as 
+In this example we choose $\mathit{I}$ to be the same inference model as 
 the generative one,
-which is the chosen Yule tree prior $\mathit{p_{G}^0}$ as well as site 
-and clock models defined in $\mathit{A}$, respectively JC and strict.
+i.e. the Yule tree prior $\mathit{p_{G}^0}$ and site 
+and clock models defined in $\mathit{A}$, respectively Jukes-Cantor and strict clock.
 We specify in $\mathit{C}$ that the experiment will always be run.
@@
 We also need to specify the error measurement parameters $\mathit{E}$.
-Here we use the default $\mathit{E}$, which uses a burn-in fraction 
-of 10\% and the nLTT statistic to
+Here we choose the default $\mathit{E}$, which has a burn-in fraction 
+of 10\% and uses the nLTT statistic to
 measure the difference between phylogenies. For clarity,
@@
 parameters, the experiments and the error measure parameters.
-These objects need to be bundled in a bigger parameter structure, 
+These objects need to be bundled in a larger parameter structure, 
 using \verb;create_pir_params;:
@@
 Obtaining this control is in fact an inherent part of \verb;pirouette;. 
-We call it 'twinning' and we will demonstrate it at section
+We call it 'twinning' and we will demonstrate it in section
  \ref{Comparing to background noise}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparing to other candidate models}
+\subsection{The inference model may differ from the generative model}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 with a known generative tree prior $\mathit{p_{G}^0}$.
-However, a novel tree prior $\mathit{p_{G}}$, by our definition, is 
+However, a novel tree prior $\mathit{p_{G}}$ is by (our) definition
 non-standard, and hence not part of a standard inference 
@@
 In such a case, the question is which standard tree prior should be used in 
-the inference. In this example, not only we re-use our hand-picked 
+the inference. In this example, we do not only re-use our hand-picked 
 inference model, but we also pick the best inference model from a set of 
-inference models. With the 'best' inference model, we mean the model with 
-the highest evidence (also known as marginal likelihood), based on the 
-alignment. We show the procedure using \verb;pirouette; to answer a second 
+inference models, i.e.  the model with 
+the highest evidence measured by its marginal likelihood. We show the procedure using \verb;pirouette; to answer a second 
 research question: "What is the inference error made on a novel phylogeny when using the best inference model, in comparison to a hand-picked model?
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparing to background noise}\label{Comparing to background noise}
+\subsection{The generative model is non-standard}\label{Comparing to background noise}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 To do so, we use a twin tree (see Subsection~\ref{subsec:twinning}) for 
-each of the generative and best candidate model.
+both the generative and best candidate models.
 The research question this example answers is:
 "What is the inference error made from a phylogeny, 
-for both a generative model and a best candidate model, compared to the 
-background noise?"
+for both a generative model and the best candidate model?"
 
@@
 The output is shown in Figure~\ref{fig:example_3}.
+While the error distributions using the best or generative model as inference model are very similar, the error distributions of the true tree are substantially larger than those of the twin tree.
+This is the error made in inference by not using the generating species tree prior.
 
@@
 but we chose to provide the simplest (and fastest to run) examples.
-All examples used only one original tree,
+
+Figure~\ref{fig:example_3} illustrates the primary result of our pipeline: it shows the error distributions for the true tree and the twin tree when either the generating model or the best candidate model is used.
+For this specific case it does not seem to matter that the site and clock models used in inference are different from those used in generating the data, but the difference between the error distributions for the true tree and the twin tree suggest that the choice of tree prior does matter.
+
+We note, however, that all examples used only one original tree,
 where any speciation process produces a whole range of trees.
-One tree is not enough to determine the impact of a tree prior 
-on Bayesian inference.
-However, if the same procedure were repeated and performed on a 
-distribution with a sufficient number of generative trees, 
+One tree is not enough to determine the impact of a tree prior on Bayesian inference.
+However, if the same procedure were repeated and performed on a distribution with a sufficient number of generative trees, 
 it would constitute a quantitative and effective assessment of the 
@@
 as the stochasticity in generating a twin tree will occasionally
-yield a tree of very low likelihood.
+yield a tree of very low probability.
 
@@
 in comparison to the errors made by a standard prior.
-From such results an user can estimate whether or not a new tree prior, 
+The user can then judge whether or not a new tree prior, 
 tailored on the generative process, is needed. 
-If this were indeed the case, one could rightly implement 
-their novel tree prior as an addition to 
-their favorite Bayesian inference tool.
+If this is indeed the case, one should implement 
+the novel tree prior as an addition to 
+the favorite Bayesian inference tool.
 
@@
 The code used in this article and more examples that are periodically 
-tested to work, can be found at 
-\url{https://github.com/richelbilderbeek/pirouette}.
+tested, can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_examples}.
 

### 2019-07-01 content.tex
No duplicate label 'lst:create_all_experiments', renamed one to 'lst:combine_all_experiments'

@@
   frame = single, 
-  label = {lst:create_all_experiments},
+  label = {lst:combine_all_experiments},
   caption = {

### 2019-07-23 content.tex
Reviewed

@@
   \textbf{4. }
-    Last, we discuss the results obtained by the examples and their interpretation. \\
+    Last, we discuss the results obtained by the examples and their 
+    interpretation. \\
 \end{abstract}
@@
 from character data (usually nucleotide sequences) extracted 
-from extant (but also extinct) organisms, and hence in our understanding of the main drivers and modes of diversification.
+from extant (but also extinct) organisms, 
+and hence in our understanding of the main drivers 
+and modes of diversification.
 
 BEAST [\cite{drummond2007beast}] is a typical Bayesian phylogenetics tool, 
-that needs both character data and priors to infer a posterior distribution of phylogenies.
-Specifically, for the species tree prior - which describes the process of diversification - 
+that needs both character data and priors to infer 
+a posterior distribution of phylogenies.
+Specifically, for the species tree prior - which describes 
+the process of diversification - 
 BEAST has built-in priors such as the Yule [\cite{yule}] and 
@@
 BEAST's successor, BEAST2 [\cite{bouckaert2014beast}],
-has a package manager, that allows third-party users to extend existing functionalities.
-For example, one can add novel diversification models by writing a BEAST2 package that contains the likelihood formula
-of a phylogeny under the novel diversification model, i.e. the prior probability of a species tree.
-Many such diversification models (and their associated probability algorithms) have been developed, e.g., models in which diversification is 
+has a package manager, that allows third-party users 
+to extend existing functionalities. For example, 
+one can add novel diversification models by writing a BEAST2 package 
+that contains the likelihood formula of a phylogeny 
+under the novel diversification model, 
+i.e. the prior probability of a species tree.
+Many such diversification models (and their associated probability algorithms) 
+have been developed, e.g., models in which diversification is 
 time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
 or diversity-dependent [\cite{etienne2011diversity}],
-or where diversification rates change for specific lineages and their descendants [\cite{etienne2012conceptual}, 
-\cite{rabosky2014automatic}, \cite{alfaro2009nine}, \cite{laudanno2018sls}], models that treat speciation as a process that takes 
+or where diversification rates change for specific lineages 
+and their descendants [\cite{etienne2012conceptual}, 
+\cite{rabosky2014automatic}, \cite{alfaro2009nine}, 
+\cite{laudanno2018sls}], models that treat speciation as a process that takes 
 time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
@@
 its performance in inference should be tested.
-Part of a model's performance is its ability to recover parameters from simulated data with known parameters (e.g.
-[\cite{etienne2014estimating}]), where ideally the estimated parameter values closely match the known/true values.
+Part of a model's performance is its ability to recover parameters 
+from simulated data with known 
+parameters (e.g. [\cite{etienne2014estimating}]), 
+where ideally the estimated parameter values closely match 
+the known/true values.
 
@@
 it is not necessarily used in Bayesian inference.
-Bayesian phylogenetic inference often requires that the prior probability of the phylogeny according to the diversification model has to be computed millions of times.
-Therefore, biologically interesting but computationally expensive tree priors are often not implemented, and simpler priors are used instead.
-This is not necessarily problematic, when the data are very informative, as this will reduce the influence of the tree prior.
-However, the assumption that tree prior choice is of low importance must first be verified.
+Bayesian phylogenetic inference often requires 
+that the prior probability of the phylogeny 
+according to the diversification model has to be computed millions of times. 
+Therefore, biologically interesting but computationally expensive tree priors 
+are often not implemented, and simpler priors are used instead. 
+This is not necessarily problematic, when the data are very informative, 
+as this will reduce the influence of the tree prior.
+However, the assumption that tree prior choice is of low importance 
+must first be verified.
 
@@
 These studies show how current diversification models compare
-to one another, but they do not help to assess the importance of a new tree prior.
+to one another, but they do not help to assess 
+the importance of a new tree prior.
 
 Here we introduce a method to quantify the importance of a novel tree prior.
-The method starts with a phylogeny generated by the new model for the tree prior.
+The method starts with a phylogeny generated 
+by the new model. 
 Next, nucleotide sequences are simulated on this phylogeny. 
-Then, using the tree priors built-in into BEAST2, a Bayesian posterior distribution of phylogenies is inferred.
-We can then compare the inferred and simulated phylogenies.
+Then, using the tree priors built-in into BEAST2, 
+a Bayesian posterior distribution of phylogenies is inferred. 
+We then compare the inferred and simulated phylogenies. 
 How to properly perform this comparison forms the heart of our method.
-Only new diversification models that result in a large discrepancy between inferred and simulated phylogenies will be worth the effort and computational burden 
+Only new diversification models that result 
+in a large discrepancy between inferred and simulated phylogenies 
+will be worth the effort and computational burden 
 to implement a species tree prior for in a Bayesian framework.
@@
 The goal of \verb;pirouette; is to quantify the importance of a tree prior.
-It does so by measuring the inference error made for a given reconstructed phylogeny, 
+It does so by measuring the inference error made 
+for a given reconstructed phylogeny, 
 simulated under a (usually novel) diversification model.
@@
 as the 'generative tree prior' $\mathit{p_{G}}$.
-Many tree priors have a parameter setting for which it reduces to a standard tree prior.
-For example, a protracted birth-death model \citep{etienne2012prolonging} reduces to a standard birth-death model when the speciation-completion rate is infinite [i.e.
-rate $\lambda$ in \cite{etienne2014estimating}, eqs.~(2b) and (2c)].
+Many tree priors have a parameter setting 
+for which it reduces to a standard tree prior. 
+For example, a protracted birth-death model \citep{etienne2012prolonging} 
+reduces to a standard birth-death model when the speciation-completion rate 
+is infinite [i.e. rate $\lambda$ in \cite{etienne2014estimating}, 
+eqs.~(2b) and (2c)].
 When benchmarking a novel tree prior, 
@@
 to assess under which scenarios the inference error cannot be neglected. 
-While we recommend many replicate simulations when assessing a novel tree prior, our examples contain only one replicate as they are for illustrative purposes only.
+While we recommend many replicate simulations 
+when assessing a novel tree prior, 
+our examples contain only one replicate 
+as they are for illustrative purposes only.
 
@@
 
-We assume the user has a phylogeny simulated with the new diversification model. The pipeline to assess the error BEAST2 makes in inferring this phylogeny then contains the following steps:
+We assume the user has a phylogeny simulated with the new diversification model. 
+The pipeline to assess the error BEAST2 makes in inferring this phylogeny 
+then contains the following steps:
 \begin{enumerate}
@@
     under a known alignment model $\mathit{A}$;
-  \item from this alignment, according to the specified inference conditions $\mathit{C}$, 
+  \item from this alignment, according to the specified 
+    inference conditions $\mathit{C}$, 
     an inference model $\mathit{I}$ is chosen (which may differ from the 
@@
   \item the phylogenies in the posterior are compared with the given phylogeny 
-    to estimate the error made, according to the error measure $\mathit{E}$ specified 
-    by the user;
+    to estimate the error made, according to 
+    the error measure $\mathit{E}$ specified by the user;
 \end{enumerate}
@@
 using the DNA alignment parameters.
-The DNA alignment parameters consist of a (DNA) root sequence, a (DNA) mutation rate, a clock model and a nucleotide substitution
-model, which will be referred to as site model.
+The DNA alignment parameters consist of 
+a (DNA) root sequence, a (DNA) mutation rate, a clock model 
+and a nucleotide substitution model.
 The root sequence is the DNA sequence of the shared common ancestor,
@@
 helps interpreting the resulting alignment.
-Supported site models are JC, HKY, TN and GTR. Only the strict
+Supported nucleotide substitution model, which
+we will refer to as site models, 
+are JC, HKY, TN and GTR. Only the strict
 clock model is currently supported in this step.
@@
 The evidence for an inference model is estimated by nested 
-sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. We note that scripted use of BEAST2 packages is only possible under Linux and Mac.
+sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. 
+We note that scripted use of BEAST2 packages is only possible under Linux and Mac.
 Windows systems can do the model comparison for shorter DNA sequences 
@@
 
-The fourth step quantifies the inference error made. First the burn-in fraction is removed, i.e. the first phase of the Markov chain Monte Carlo (MCMC) run,
-which samples an unrepresentative part of state space. By default, \verb;pirouette; 
+The fourth step quantifies the inference error made. 
+First the burn-in fraction is removed, i.e. the first phase of the 
+Markov chain Monte Carlo (MCMC) run,
+which samples an unrepresentative part of state space. 
+By default, \verb;pirouette; 
 removes the first 10\% of the posterior.
@@
 with branching times $\Vec{t}_{\mathit{G}}$ and 
-topology $\psi_{\mathit{G}}$, the twinning process T, in accordance with a chosen inference 
+topology $\psi_{\mathit{G}}$, the twinning process T, 
+in accordance with a chosen inference 
 model $\mathit{I}$, creates a tree $\tau_{\mathit{I}}$
@@
 
-The twin tree serves as a control: even when the generating and inference models are identical, the posterior distribution of inferred trees will differ from the true tree due to stochasticity in producing an alignment and in MCMC sampling of the posterior; the twin tree provides this minimum error.
+The twin tree serves as a control: 
+even when the generating and inference models are identical, 
+the posterior distribution of inferred trees will differ 
+from the true tree due to stochasticity in producing an alignment 
+and in MCMC sampling of the posterior; 
+the twin tree provides this minimum error. 
 We therefore need to compare the error distributions of $\tau_{\mathit{G}}$ and 
 $\tau_{\mathit{I}}$. 
-By imposing the same topology and the same number of mutations in the alignments, the differences in error distributions of the true tree and the twin tree are the consequence of different phylogenies (produced by different processes) and tree priors for the two parallel
+By imposing the same topology and the same number of mutations 
+in the alignments, the differences in error distributions of 
+the true tree and the twin tree are the consequence 
+of different phylogenies (produced by different processes) 
+and tree priors for the two parallel 
 pipelines: the main pipeline uses the tree $\tau_{\mathit{G}}$ and a
@@
 uses the tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
-So, for the twin trees, if the process is repeated a sufficient number of times, the resulting error distribution cannot be 
-attributed to a mismatch between the true tree prior and the tree prior used in
-inference.
+So, for the twin trees, if the process is repeated 
+a sufficient number of times, the resulting error distribution cannot be 
+attributed to a mismatch between the true tree prior 
+and the tree prior used in inference.
 
@@
 
-\verb;pirouette; will be made available on CRAN from which it can then be easily installed:
+\verb;pirouette; will be made available on CRAN from which 
+it can then be easily installed:
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]
@@
 \end{lstlisting}
-To start using \verb;pirouette;, load its functions in the global namespace first:
+To start using \verb;pirouette;, 
+load its functions in the global namespace first:
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]
@@
 Second, to establish that the true tree prior is indeed the best, 
-we also measure the inference error made adopting a
+we also measure the inference error made by a
 best candidate inference model.
@@
 
+All the figures shown in this section are shown as-is, without any
+aesthetical modifications..
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \verb;pirouette; quantifies the influence of a new tree prior on BEAST2's 
-inference by measuring the discrepancy between a given tree and a posterior 
+inference by measuring the discrepancy between a given/true tree and a posterior 
 distribution of phylogenies obtained as result of the inference process. 
@@
 of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' 
-all the tree priors that are available to be used within an inference model (see Table~\ref{tab:options}).
+all the tree priors that are available 
+to be used within an inference model (see Table~\ref{tab:options}).
 
@@
 
-In this example we use a standard generative tree prior $\mathit{p_{G}^0}$, namely the Yule (pure-birth) tree prior. 
+In this example we use a standard generative tree prior $\mathit{p_{G}^0}$, 
+namely the Yule (pure-birth) tree prior. 
 We choose to use a small tree with six taxa, to keep
@@
 i.e. the Yule tree prior $\mathit{p_{G}^0}$ and site 
-and clock models defined in $\mathit{A}$, respectively Jukes-Cantor and strict clock.
+and clock models defined in $\mathit{A}$, 
+respectively Jukes-Cantor and strict clock.
 We specify in $\mathit{C}$ that the experiment will always be run.
@@
 We call it 'twinning' and we will demonstrate it in section
- \ref{Comparing to background noise}.
+\ref{Comparing to background noise} below.
 
@@
 inference models, i.e.  the model with 
-the highest evidence measured by its marginal likelihood. We show the procedure using \verb;pirouette; to answer a second 
-research question: "What is the inference error made on a novel phylogeny when using the best inference model, in comparison to a hand-picked model?
+the highest evidence measured by its marginal likelihood. 
+We show the procedure using \verb;pirouette; to answer a second 
+research question: "What is the inference error made on a novel phylogeny 
+when using the best inference model, in comparison to a hand-picked model?
 
@@
 As model selection is commonly performed on the full list of available 
-candidate models, \verb;pirouette; has a dedicated function for this choice: \verb;create_all_experiments; creates a full set of 40 experiments, 
+candidate models, \verb;pirouette; has a dedicated function 
+for this choice: \verb;create_all_experiments; creates a full set 
+of 40 experiments, 
 containing the inference models of all combinations of 4 site models, 
@@
 
-We start from the tree generated by a non-standard (and unknown) tree prior, having six taxa and a crown age of ten:
+We start from the tree generated by a non-standard (and unknown) tree prior, 
+having six taxa and a crown age of ten:
 \begin{lstlisting}[
@@
     as created by listing~\ref{lst:unknown_phylogeny}.
+    The scale overlaps with the lower branch, which is the default 
+    behavior for the plotting function used.
   }
@@
 
-Most of the other settings are the same as before: we re-use the alignment parameters (Listing~\ref{lst:create_alignment_params}), as well as the experiments (Listing~\ref{lst:create_all_experiments}).
+Most of the other settings are the same as before: 
+we reuse the alignment parameters (Listing~\ref{lst:create_alignment_params}), 
+as well as the experiments (Listing~\ref{lst:create_all_experiments}).
 This time, however, we enable twinning (see Subsection~\ref{subsec:twinning}),
@@
 twinning_params <- create_twinning_params(
-  twin_model = "birth_death",
+  twin_model = "yule",
   method = "random_tree"
@@
 The output is shown in Figure~\ref{fig:example_3}. 
-While the error distributions using the best or generative model as inference model are very similar, the error distributions of the true tree are substantially larger than those of the twin tree.
-This is the error made in inference by not using the generating species tree prior.
+While the error distributions using the best or generative model 
+as inference model are very similar, 
+the error distributions of the true tree are substantially 
+larger than those of the twin tree. 
+This is the error made in inference 
+by not using the generating species tree prior.
 
@@
 
-Figure~\ref{fig:example_3} illustrates the primary result of our pipeline: it shows the error distributions for the true tree and the twin tree when either the generating model or the best candidate model is used.
-For this specific case it does not seem to matter that the site and clock models used in inference are different from those used in generating the data, but the difference between the error distributions for the true tree and the twin tree suggest that the choice of tree prior does matter.
+Figure~\ref{fig:example_3} illustrates the primary result of our pipeline: 
+it shows the error distributions for the true tree and the twin tree 
+when either the generating model or the best candidate model is used. 
+The clear difference between the error distributions 
+for the true tree and the twin tree suggest 
+that the choice of tree prior does matter.
 
@@
 where any speciation process produces a whole range of trees.
-One tree is not enough to determine the impact of a tree prior on Bayesian inference.
-However, if the same procedure were repeated and performed on a distribution with a sufficient number of generative trees, 
+One tree is not enough to determine the impact 
+of a tree prior on Bayesian inference.
+However, if the same procedure were repeated and 
+performed on a distribution with a sufficient number of generative trees, 
 it would constitute a quantitative and effective assessment of the 
@@
 Also a twin tree does not always result in a lower error distribution,
-as the stochasticity in generating a twin tree will occasionally
-yield a tree of very low probability.
+as the stochasticity in generating a twin tree will - with
+very low probability - yield a tree of that same very low probability.
 
@@
 tailored on the generative process, is needed. 
-If this is indeed the case, one should implement 
+If this is indeed the case, one is warranted to implement 
 the novel tree prior as an addition to 
-the favorite Bayesian inference tool.
+his/her favorite Bayesian inference tool.
 
@@
 which allows submitting bug reports, requesting features, 
-or adding code. To ensure a high quality, \verb;pirouette; 
+and adding code. To ensure a high quality, \verb;pirouette; 
 uses a continuous integration service, has a code coverage of above 95\%

### 2019-07-23 supplement.tex
Refresh data
Reviewed

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+These supplementary materials show the full pipeline and
+added diagnostics for the examples in the main article.
+
+As in the main article, all the figures shown in this section are shown as-is, 
+without any aesthetical modifications, with the exception that the arrangement
+of the sub-figures (for example, aligning true and twin tree
+horiztonally) is done manually.
 
@@
 
-\input{example_1/esses.latex}
+\input{example_1/esses_gen.latex}
 

### 2019-07-23 content.tex
Use preferred new BEAST2 citation

@@
 have sufficient biological complexity, while being computationally fast.
-BEAST's successor, BEAST2 [\cite{bouckaert2014beast}],
+BEAST's successor, BEAST2 [\cite{bouckaert2019beast}],
 has a package manager, that allows third-party users 
@@
 \verb;pirouette; is built on \verb;babette; [\cite{bilderbeek2018babette}], 
-which calls BEAST2 [\cite{bouckaert2014beast}].
+which calls BEAST2 [\cite{bouckaert2019beast}].
 

### 2019-07-23 content.tex
Cite RLN

@@
     Clock for the DNA mutation rates &
-    strict, relaxed log-normal \\
+    strict, RLN \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     TN93 = Tamura and Nei (\cite{tamura1993estimation}), 
-    GTR = Generalized time-reversible model (\cite{tavare1986some}).
+    GTR = Generalized time-reversible model (\cite{tavare1986some})
+    RLN = relaxed log-normal clock model (\cite{drummond2006relaxed}).
   }

### 2019-07-23 content.tex
Point to supplementary materials

@@
 All the figures shown in this section are shown as-is, without any
-aesthetical modifications..
+aesthetical modifications. Figures showing the full workflow 
+and tables showing the effective sample sized (a measure
+of inference quality) can be found in the supplementary materials.
 

### 2019-07-24 content.tex
Twinning does use birth_death by default, code also uses birth_death
Simulating alignments uses the same mutation rate

@@
 From the twin tree, we simulate an alignment following the same
-history as the twin tree, with the constraint that 
+history as the twin tree, using the same mutation rate,
+with the constraint that 
 the same total number of mutations have
@@
 twinning_params <- create_twinning_params(
-  twin_model = "yule",
+  twin_model = "birth_death",
   method = "random_tree"

### 2019-08-19 content.tex
Use newer figures
Minor rephrasing in twin alignment construction

@@
 with the constraint that 
-the same total number of mutations have
-taken place.
-This constraint ensures there is the same amount of historical information
-stored in both the true and twin alignments.
+the exact same total number of mutations have
+taken place from the phylogeny's root sequence and all sequences at the tips. 
 We achieve this by simply simulating twin alignments until we
 obtain one that has the desired number of mutations.
+This constraint ensures there is approximately 
+the same amount of historical information
+stored in both the true and twin alignments.
 
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_6/true_tree.png}
+  \includegraphics[width=\textwidth]{example_3/true_tree.png}
   \caption{The tree derived from an unknown diversification process, 
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_6/errors.png}
+  \includegraphics[width=\textwidth]{example_3/errors.png}
   \caption{

### 2019-09-30 content.tex
Preparing for submission

@@
 
-{\bf Keywords:} computational biology, evolution, phylogenetics, tree prior, 
-Bayesian model selection, BEAST2, babette, pirouette, R
+{\bf Keywords:} Bayesian model selection, BEAST2, computational biology, evolution, phylogenetics, R, tree prior
 
@@
   \centering
-  \includegraphics[width = \textwidth]{workflow.png}
+  \includegraphics[width = \textwidth]{workflow4.png}
   \caption{
@@
     For each candidate experiment $\mathit{X_{i}}$ 
-    (a tuple of inference model $\mathit{I_{i}}$ and condition $\mathit{C_{i}}$),
+    (a 2-tuple of inference model $\mathit{I_{i}}$ and condition $\mathit{C_{i}}$),
     if its condition $\mathit{C_{i}}$ is 
@@
     to be used in the next step.
-    The inference models (3) of the selected experiments use the alignment (2a) 
+    The inference models (3a) of the selected experiments use the alignment (2a) 
     to each create a Bayesian posterior of (parameter estimates and) 
@@
     Each of the posteriors' trees is compared to the true phylogeny (1a) 
-    using error measure $\mathit{E}$, 
+    using the error measure $\mathit{E}$, 
     resulting in an error distribution (5a). 
@@
     A twin phylogeny (1b) can be generated from the original 
-    phylogeny (1a) using the tree prior $\mathit{p_{I}}$ of the selected inference model (3).
-    Analogously, a twin alignment is simulated from the twin phylogeny along with clock model $\mathit{c_{I}}$ and site model $\mathit{s_{I}}$ imported from the selected inference model.
+    phylogeny (1a) using the twin tree prior $\mathit{p_{t}}$, selected among standard tree priors; the default option is the standard birth-death prior.
+    A twin alignment (2b) is then simulated from the twin phylogeny using clock model $\mathit{c_{G}}$ and site model $\mathit{s_{G}}$ imported from the generative model.
+The twin alignment has the same amount of mutations as the original alignment.
     The twin pipeline mimics the procedure of the main pipeline, resulting in a twin error distribution (5b).
@@
 
-\subsection{Twin tree}\label{subsec:twinning}
-
-An optional step is to generate a 'twin tree' $\tau_{I}$
-(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b)
-and 'twin alignment'
-that will be analyzed in the same way as the true tree (and alignment).
+\subsection{Twinning}\label{subsec:twinning}
+
+An optional step is to use the 'twinning process'. Such process generates (T1) a 'twin tree' (Fig.~\ref{fig:pipeline}, 1b) and (T2) a 'twin alignment' (Fig.~\ref{fig:pipeline}, 2b)
+that will be analyzed in the same way as the true tree and alignment.
+
 We define a phylogeny $\tau$ as the combination of
-branching times $\Vec{t}$ and topology $\psi$.
-Assuming that a non-standard generative prior $\mathit{p_{G}}$
-produces a tree $\tau_{\mathit{G}}$,
-with branching times $\Vec{t}_{\mathit{G}}$ and 
-topology $\psi_{\mathit{G}}$, the twinning process T, 
-in accordance with a chosen inference 
-model $\mathit{I}$, creates a tree $\tau_{\mathit{I}}$
-with branching times $\Vec{t}_{\mathit{I}}$ while preserving the 
+branching times $\Vec{t}$ and topology $\psi$, and denote as $\tau_{\mathit{G}}$ the phylogeny produced by a non-standard generative prior, having branching times $\Vec{t}_{\mathit{G}}$ and 
+topology $\psi_{\mathit{G}}$.
+
+The first step (T1) of the twinning process creates a tree $\tau_{\mathit{T}}$
+with branching times $\Vec{t}_{\mathit{T}}$ while preserving the original
 topology $\psi_{\mathit{G}}$:
@@
   \xrightarrow[]{\mathit{T}} 
-  \tau_{\mathit{I}} = (\Vec{t}_{\mathit{I}}, \psi_{\mathit{G}})
+  \tau_{\mathit{T}} = (\Vec{t}_{\mathit{T}}, \psi_{\mathit{G}})
 \end{align}
-To achieve this, we find the parameters $\theta^{*}_{\mathit{I}}$ 
+To do so, a standard tree prior $p_T$ for the twin tree must be specified. The default option here is the standard birth-death prior.
+It is then possible to use the likelihood function $L_{\mathit{T}}$ defined by such tree prior to find the parameters $\theta^{*}_{\mathit{T}}$ 
 (e.g. speciation and extinction rates, in case of a birth-death model) 
-that maximize the likelihood $L_{\mathit{I}}$ applied 
-to the true tree, conditioned on its number of tips $n_{\mathit{G}}$,
-where $L_{\mathit{I}}$ is the likelihood function  
-associated with the chosen inference model $\mathit{I}$:
+that maximize the likelihood applied 
+to the true tree, conditioned on its number of tips $n_{\mathit{G}}$:
 \begin{align}
-    \max[L_{\mathit{I}}(\theta_{\mathit{I}}|\tau_{\mathit{G}}, n_{\mathit{G}})] 
-\rightarrow \theta^{*}_{\mathit{I}}.
+    \max[L_{\mathit{T}}(\theta_{\mathit{T}}|\tau_{\mathit{G}}, n_{\mathit{G}})] 
+\rightarrow \theta^{*}_{\mathit{T}}.
 \end{align}
-We use $\theta^{*}_{\mathit{I}}$ to simulate an equal number 
-$n_{\mathit{I}} = n_{\mathit{G}}$ 
-of branching times $\Vec{t}_{\mathit{I}}$ for the twin tree 
-$\tau_{\mathit{I}}$, under the process $\mathit{I}$, 
+We use $\theta^{*}_{\mathit{T}}$ to simulate an equal number 
+$n_{\mathit{T}} = n_{\mathit{G}}$ 
+of branching times $\Vec{t}_{\mathit{T}}$ for the twin tree 
+$\tau_{\mathit{T}}$, under the process $p_{T}$, 
 while preserving the same topology.
 
-From the twin tree, we simulate an alignment following the same
-history as the twin tree, using the same mutation rate,
-with the constraint that 
-the exact same total number of mutations have
-taken place from the phylogeny's root sequence and all sequences at the tips. 
+The second step (T2) of the twinning process simulates the twin alignment. To do so, we use the same clock model, site model and mutation rate used to simulate the original alignment. 
+We also impose that, in the twin alignment, the total number of mutations respect to the root sequence must be the same as in the true alignment.
+
 We achieve this by simply simulating twin alignments until we
@@
 
-The twin tree serves as a control: 
-even when the generating and inference models are identical, 
-the posterior distribution of inferred trees will differ 
-from the true tree due to stochasticity in producing an alignment 
-and in MCMC sampling of the posterior; 
-the twin tree provides this minimum error. 
-We therefore need to compare the error distributions of $\tau_{\mathit{G}}$ and 
-$\tau_{\mathit{I}}$. 
-By imposing the same topology and the same number of mutations 
-in the alignments, the differences in error distributions of 
-the true tree and the twin tree are the consequence 
-of different phylogenies (produced by different processes) 
-and tree priors for the two parallel 
-pipelines: the main pipeline uses the tree $\tau_{\mathit{G}}$ and a
-possibly mismatched prior $\mathit{G}$, whereas the twin pipeline
-uses the tree $\tau_{\mathit{I}}$ for a matching prior $\mathit{I}$.
-So, for the twin trees, if the process is repeated 
-a sufficient number of times, the resulting error distribution cannot be 
-attributed to a mismatch between the true tree prior 
-and the tree prior used in inference.
+The twin pipeline serves as a control: even when the generating and inference models are identical, the inferred trees from the posterior distribution will still differ from the true tree due to stochasticity in producing an alignment and to the MCMC sampling of the posterior; the twin pipeline provides this minimum error.
+We therefore need to compare the error distributions at the end of the two parallel pipelines.
+By imposing the same topology and the same number of mutations in the two alignments, any differences found in the two error distributions will be the consequence of different distributions of branching times in the starting phylogenies.
+This is due to the fact that true and twin phylogenies are realizations of different processes: one non-standard, $p_G$, and one standard, $p_T$ (see Fig~\ref{fig:pipeline}).
+
+This can be seen for both the "generative" and "candidate" model types (see Table~\ref{tab:options}).
+If the chosen model type is "generative", the tree prior chosen for the twin inference will be exactly the same used to generate the tree.
+In the main pipeline, being the tree prior $p_{G}$ non-standard, this is not possible.
+If, instead, the chosen model type is "candidate", the twin tree prior will be included in the pool of examined models during the process of selection of the inference model.
+Being the generative tree prior non-standard, this is again not possible for the main pipeline.
+
+Finally, if the goal is to evaluate BEAST2's performance on a non-standard tree prior, the last source of stochasticity comes from phylogenies.
+In fact, a single phylogeny cannot be considered as fully representative of the prior.
+For this reason multiple phylogenies, as well as an equal amount of twins, must be considered.
+Being the error measure normalized (i.e.
+comprised in the interval $[0, 1]$), it is possible to considerate the aggregated versions of the errors distributions across all the runs.
+Therefore, if the number of considered phylogenies is high enough, the comparison between the main pipeline's aggregated error distribution and its twin counterpart leads to a fair evaluation of the new tree prior respect to the baseline error.
 

### 2019-09-30 content.tex
Proofread twinning figure, caption and section, fix https://github.com/richelbilderbeek/razzo/issues/291

@@
     The phylogeny is converted to an alignment (2a) using the generative alignment model 
-    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed by a clock and a site model. 
+    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed of a clock and a site model. 
     The user supplies one or more experiments.
@@
     A twin phylogeny (1b) can be generated from the original 
-    phylogeny (1a) using the twin tree prior $\mathit{p_{t}}$, selected among standard tree priors; the default option is the standard birth-death prior.
-    A twin alignment (2b) is then simulated from the twin phylogeny using clock model $\mathit{c_{G}}$ and site model $\mathit{s_{G}}$ imported from the generative model.
+    phylogeny (1a) using the twin tree prior $\mathit{p_{t}}$, 
+    selected among standard tree priors; 
+    the default option is the standard birth-death prior.
+    A twin alignment (2b) is then simulated from the twin phylogeny 
+    using clock model $\mathit{c_{G}}$ and site model $\mathit{s_{G}}$ 
+    imported from the generative model. 
     The twin alignment has the same amount of mutations as the original alignment.
-    The twin pipeline mimics the procedure of the main pipeline, resulting in a twin error distribution (5b).
+    The twin pipeline mimics the procedure of the main pipeline, 
+    resulting in a twin error distribution (5b).
   }
@@
 
-An optional step is to use the 'twinning process'. Such process generates (T1) a 'twin tree' (Fig.~\ref{fig:pipeline}, 1b) and (T2) a 'twin alignment' (Fig.~\ref{fig:pipeline}, 2b)
-that will be analyzed in the same way as the true tree and alignment.
+An optional step is to use the 'twinning process'.
+This process, $T$, encompasses two steps:
+$T_1$, that generates a 'twin tree' (Fig.~\ref{fig:pipeline}, 1b) 
+and $T_2$, which generates a 'twin alignment' (Fig.~\ref{fig:pipeline}, 2b).
+Both twin tree and alignment will be analyzed in the same way 
+as the true tree and alignment.
 
 We define a phylogeny $\tau$ as the combination of
-branching times $\Vec{t}$ and topology $\psi$, and denote as $\tau_{\mathit{G}}$ the phylogeny produced by a non-standard generative prior, having branching times $\Vec{t}_{\mathit{G}}$ and 
+branching times $\Vec{t}$ and topology $\psi$, 
+and denote as $\tau_{\mathit{G}}$ the phylogeny 
+produced by a (possibly non-standard) generative prior, 
+having branching times $\Vec{t}_{\mathit{G}}$ and 
 topology $\psi_{\mathit{G}}$.
 
-The first step (T1) of the twinning process creates a tree $\tau_{\mathit{T}}$
+The first step ($T_1$) of the twinning process creates a tree $\tau_{\mathit{T}}$
 with branching times $\Vec{t}_{\mathit{T}}$ while preserving the original
@@
   \tau_{\mathit{G}} = (\Vec{t}_{\mathit{G}}, \psi_{\mathit{G}}) 
-  \xrightarrow[]{\mathit{T}} 
+  \xrightarrow[]{\mathit{T_1}} 
   \tau_{\mathit{T}} = (\Vec{t}_{\mathit{T}}, \psi_{\mathit{G}})
 \end{align}
-To do so, a standard tree prior $p_T$ for the twin tree must be specified. The default option here is the standard birth-death prior.
-It is then possible to use the likelihood function $L_{\mathit{T}}$ defined by such tree prior to find the parameters $\theta^{*}_{\mathit{T}}$ 
+To do so, a standard tree prior $p_T$ for the twin tree must be specified. 
+The default option here is the standard birth-death prior.
+It is then possible to use the likelihood function 
+$L_{\mathit{T}}$ defined by such tree prior to find the parameters $\theta^{*}_{\mathit{T}}$ 
 (e.g. speciation and extinction rates, in case of a birth-death model) 
@@
 
-The second step (T2) of the twinning process simulates the twin alignment. To do so, we use the same clock model, site model and mutation rate used to simulate the original alignment. 
-We also impose that, in the twin alignment, the total number of mutations respect to the root sequence must be the same as in the true alignment.
+The second step, $T_2$m of the twinning process simulates the twin alignment. 
+To do so, we use the same clock model, site model and mutation rate 
+used to simulate the original alignment. 
+We also impose that, in the twin alignment, 
+the total number of mutations with respect to the root sequence 
+must be the same as in the true alignment.
 
@@
 
-The twin pipeline serves as a control: even when the generating and inference models are identical, the inferred trees from the posterior distribution will still differ from the true tree due to stochasticity in producing an alignment and to the MCMC sampling of the posterior; the twin pipeline provides this minimum error.
-We therefore need to compare the error distributions at the end of the two parallel pipelines.
-By imposing the same topology and the same number of mutations in the two alignments, any differences found in the two error distributions will be the consequence of different distributions of branching times in the starting phylogenies.
-This is due to the fact that true and twin phylogenies are realizations of different processes: one non-standard, $p_G$, and one standard, $p_T$ (see Fig~\ref{fig:pipeline}).
-
-This can be seen for both the "generative" and "candidate" model types (see Table~\ref{tab:options}).
-If the chosen model type is "generative", the tree prior chosen for the twin inference will be exactly the same used to generate the tree.
-In the main pipeline, being the tree prior $p_{G}$ non-standard, this is not possible.
-If, instead, the chosen model type is "candidate", the twin tree prior will be included in the pool of examined models during the process of selection of the inference model.
-Being the generative tree prior non-standard, this is again not possible for the main pipeline.
-
-Finally, if the goal is to evaluate BEAST2's performance on a non-standard tree prior, the last source of stochasticity comes from phylogenies.
+The twin pipeline serves as a control: 
+even when the generating and inference models are identical
+(as is the case in the twin pipeline),
+the inferred trees from the posterior distribution will 
+still differ from the true tree, 
+due to stochasticity in producing an alignment 
+and to the MCMC sampling of the posterior.
+The twin pipeline provides this minimum error possible,
+because the generating and inference model match exactly.
+When comparing the true and twin error distribution,
+any differences will be due to the fact that true and twin phylogenies are realizations 
+of different processes: one (possibly) non-standard, $p_G$, 
+and one standard, $p_T$ (see Fig~\ref{fig:pipeline}).
+
+This can be seen for both the "generative" and "candidate" model 
+types (see Table~\ref{tab:options}).
+If the chosen model type is "generative", 
+the tree prior chosen for the twin inference will be 
+exactly the same used to generate the tree.
+In the main pipeline, as the tree prior $p_{G}$ is non-standard, 
+it cannot be used in inference.
+If, instead, the chosen model type is "candidate", 
+the twin tree prior will be included in the pool of examined models 
+during the process of selection of the inference model.
+
+Finally, if the goal is to evaluate BEAST2's performance on a non-standard tree prior, 
+the last source of stochasticity comes from phylogenies. 
 In fact, a single phylogeny cannot be considered as fully representative of the prior. 
-For this reason multiple phylogenies, as well as an equal amount of twins, must be considered.
-Being the error measure normalized (i.e.
-comprised in the interval $[0, 1]$), it is possible to considerate the aggregated versions of the errors distributions across all the runs.
-Therefore, if the number of considered phylogenies is high enough, the comparison between the main pipeline's aggregated error distribution and its twin counterpart leads to a fair evaluation of the new tree prior respect to the baseline error.
+For this reason multiple phylogenies, 
+as well as an equal amount of twins, 
+must be considered. 
+Having the error measure normalized (i.e. comprised in the interval $[0, 1]$), 
+it is possible to considerate the aggregated versions of the errors distributions 
+across all the runs.
+Therefore, if the number of considered phylogenies is high enough, 
+the comparison between the main pipeline's aggregated error distribution 
+and its twin counterpart leads to a fair evaluation 
+of the new tree prior respect to the baseline error.
 

### 2019-10-02 content.tex
Updates from Overleaf

@@
 simulated under a (usually novel) diversification model.
-We refer to the true prior that generated the given tree
-as the 'generative tree prior' $\mathit{p_{G}}$.
+We refer to the true model that generated the given tree
+as the 'generative tree model' $\mathit{p_{G}}$.
 Many tree priors have a parameter setting 
-for which it reduces to a standard tree prior. 
+for which they reduce to a standard tree prior. 
 For example, a protracted birth-death model \citep{etienne2012prolonging} 
@@
     \texttt{pirouette} pipeline.
-    The pipeline starts from a phylogeny (1a) simulated by the generative tree prior 
+    The pipeline starts from a phylogeny (1a) simulated by the generative tree model 
     $\mathit{p_{G}}$.
@@
     $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed of a clock and a site model. 
-    The user supplies one or more experiments.
+    The user defines one or more experiments.
     For each candidate experiment $\mathit{X_{i}}$ 
-    (a 2-tuple of inference model $\mathit{I_{i}}$ and condition $\mathit{C_{i}}$),
+    (a combination of inference model $\mathit{I_{i}}$ and condition $\mathit{C_{i}}$),
     if its condition $\mathit{C_{i}}$ is 
@@
     A twin phylogeny (1b) can be generated from the original 
-    phylogeny (1a) using the twin tree prior $\mathit{p_{t}}$, 
-    selected among standard tree priors; 
-    the default option is the standard birth-death prior.
+    phylogeny (1a) using the twin tree model $\mathit{p_{t}}$, 
+    selected among standard diversification models; 
+    the default option is the standard birth-death model, with parameters estimated from the original phylogeny.
     A twin alignment (2b) is then simulated from the twin phylogeny 
@@
     imported from the generative model. 
-    The twin alignment has the same amount of mutations as the original alignment.
-    The twin pipeline mimics the procedure of the main pipeline, 
+    The twin alignment has the same number of mutations as the original alignment.
+    The twin pipeline follows the procedure of the main pipeline, 
     resulting in a twin error distribution (5b).
@@
 The root sequence is the DNA sequence of the shared common ancestor,
-and is set to four equally-sized mononucleotide blocks by default, as this
+and is set to four different equally-sized mononucleotide blocks by default, as this
 helps interpreting the resulting alignment.
@@
 and denote as $\tau_{\mathit{G}}$ the phylogeny 
-produced by a (possibly non-standard) generative prior, 
+produced by a (possibly non-standard) generative diversification model, 
 having branching times $\Vec{t}_{\mathit{G}}$ and 
@@
 \end{align}
-To do so, a standard tree prior $p_T$ for the twin tree must be specified. 
-The default option here is the standard birth-death prior.
+The default option for the diversification model $p_T$ is the standard birth-death model.
 It is then possible to use the likelihood function 
-$L_{\mathit{T}}$ defined by such tree prior to find the parameters $\theta^{*}_{\mathit{T}}$ 
+$L_{\mathit{T}}$ for this diversification model to find the parameters $\theta^{*}_{\mathit{T}}$ 
 (e.g. speciation and extinction rates, in case of a birth-death model) 
-that maximize the likelihood applied 
+that maximize this likelihood applied 
 to the true tree, conditioned on its number of tips $n_{\mathit{G}}$:
@@
 \end{align}
-We use $\theta^{*}_{\mathit{T}}$ to simulate an equal number 
+We use $\theta^{*}_{\mathit{T}}$ to simulate a number 
 $n_{\mathit{T}} = n_{\mathit{G}}$ 
@@
 $\tau_{\mathit{T}}$, under the process $p_{T}$, 
-while preserving the same topology.
-
-The second step, $T_2$m of the twinning process simulates the twin alignment. 
-To do so, we use the same clock model, site model and mutation rate 
+while preserving the topology.
+
+The second step ($T_2$) of the twinning process simulates the twin alignment with the same clock model, site model and mutation rate 
 used to simulate the original alignment. 
@@
 the total number of mutations with respect to the root sequence 
-must be the same as in the true alignment.
-
-We achieve this by simply simulating twin alignments until we
+must be the same as in the true alignment in order to keep the information content 
+stored in both the true and twin alignments as similar as possible. We achieve this by simply simulating twin alignments until we
 obtain one that has the desired number of mutations.
-This constraint ensures there is approximately 
-the same amount of historical information
-stored in both the true and twin alignments.
 
@@
 and to the MCMC sampling of the posterior.
-The twin pipeline provides this minimum error possible,
+The twin pipeline provides this minimum error,
 because the generating and inference model match exactly.
@@
 exactly the same used to generate the tree.
-In the main pipeline, as the tree prior $p_{G}$ is non-standard, 
+In the main pipeline, as the tree model $p_{G}$ is non-standard, 
 it cannot be used in inference.
@@
 For this reason multiple phylogenies, 
-as well as an equal amount of twins, 
+as well as an equal number of twins, 
 must be considered. 
@@
 and its twin counterpart leads to a fair evaluation 
-of the new tree prior respect to the baseline error.
+of the new tree prior with respect to the baseline error.
 
@@
 we measure the Bayesian inference error when we start from a phylogeny 
-generated under a known and standard tree prior.
+generated under a known and standard tree model.
 Second, to establish that the true tree prior is indeed the best, 
@@
 running the pipeline starting from a tree generated by an unknown 
-or non-standard tree prior.
+or non-standard tree model.
 
@@
 Due to stochasticity, posterior trees will generally differ from the given 
-phylogeny $\tau_{\mathit{G}}$, even when the generative tree 
-prior $\mathit{p_{G}}$ and alignment model $\mathit{A}$ are the same as those 
-in the inference model $\mathit{I}$.
+phylogeny $\tau_{\mathit{G}}$, even when the tree 
+prior and alignment model used for inference are the same as those 
+used to generate the alignment.
 Measuring this difference allows us to know the baseline error
@@
 
-In this example we use a standard generative tree prior $\mathit{p_{G}^0}$, 
+In this example we use a standard generative tree model $\mathit{p_{G}^0}$, 
 namely the Yule (pure-birth) tree prior. 
@@
 This error distribution can serve as a control,
-as it is obtained from a tree of which the tree prior is standard and known.
+as it is obtained from a tree of which the generating tree model is standard and known.
 Obtaining this control is in fact an inherent part of \verb;pirouette;. 
@@
 So far we have measured the inference error on a tree
-generated according to a known (and standard) tree prior. 
-The goal of \verb;pirouette; is to measure the impact of a novel tree prior.
+generated according to a known (and standard) tree diversification model. 
+The goal of \verb;pirouette; is to measure the expected impact of a novel tree prior.
 To do so, in this example, we will use a tree generated by a non-standard 
-tree prior, assumed to be closely related to the Yule tree prior.
-To measure the impact of the tree prior, we measure both the baseline 
-error (that serves as a control) and the full inference error: it is their 
-difference that shows the impact of the tree prior.
-To do so, we use a twin tree (see Subsection~\ref{subsec:twinning}) for 
+tree diversification model, assumed to be closely related to the Yule tree prior.
+We measure both the baseline 
+error (that serves as a control) and the full inference error, because their 
+difference shows the impact of the tree model.
+We obtain the baseline error by using a twin tree (see Subsection~\ref{subsec:twinning}) for 
 both the generative and best candidate models.
@@
 
-We start from the tree generated by a non-standard (and unknown) tree prior, 
+We start from the tree generated by a non-standard (and unknown) diversification model, 
 having six taxa and a crown age of ten:
@@
 tree prior in Bayesian phylogenetics, using the simplest generative tree 
-prior possible.
-In principle any other (more complex) generative tree prior can be tested, 
+model possible.
+In principle any other (more complex) generative tree model can be tested, 
 but we chose to provide the simplest (and fastest to run) examples.
@@
 it shows the error distributions for the true tree and the twin tree 
-when either the generating model or the best candidate model is used. 
+when either the generating model or the best candidate model is used in inference. 
 The clear difference between the error distributions 
-for the true tree and the twin tree suggest 
+for the true tree and the twin tree suggests 
 that the choice of tree prior does matter.
@@
 
-In conclusion, \verb;pirouette; can show the errors
-when adopting a new generative prior
-in comparison to the errors made by a standard prior.
+In conclusion, \verb;pirouette; can show the errors to be expected
+when the tree prior used in inference is different from the generating model.
 The user can then judge whether or not a new tree prior, 
 tailored on the generative process, is needed. 
-If this is indeed the case, one is warranted to implement 
-the novel tree prior as an addition to 
+If this is indeed the case, one can implement the novel tree prior as an addition to 
 his/her favorite Bayesian inference tool.

### 2019-10-02 content.tex
Proper glory to @Giappo

@@
 RJCB, GL and RSE conceived the idea for the package. 
-RJCB created and tested the package, and wrote the first draft of the manuscript.
-GL tested the package and contributed to revisions.
-RSE contributed to revisions.
+RJCB created, tested and revised the package.
+GL done major contributions to the package.
+RJCB wrote the first draft of the manuscript, 
+GL and RSE contributed to revisions.
 

### 2019-10-07 content.tex
Processed feedback @rsetienne, fix https://github.com/richelbilderbeek/razzo/issues/313

@@
 For example, we may require that an inference
-model (a combination of a tree prior, clock model and site model) 
-should include the generative/true tree prior. 
+model (a combination of a tree model, clock model and site model) 
+should include the generative/true tree model. 
 As a second example, we may require that we have selected a set of 
@@
 If the chosen model type is "generative", 
-the tree prior chosen for the twin inference will be 
-exactly the same used to generate the tree.
+the tree prior chosen for the twin inference will  
+exactly match the model to generate the tree.
 In the main pipeline, as the tree model $p_{G}$ is non-standard, 
@@
 If, instead, the chosen model type is "candidate", 
-the twin tree prior will be included in the pool of examined models 
+the twin tree model will be included in the pool of examined models 
 during the process of selection of the inference model.
@@
 the last source of stochasticity comes from phylogenies. 
-In fact, a single phylogeny cannot be considered as fully representative of the prior. 
+In fact, a single phylogeny cannot be considered 
+as fully representative of the model. 
 For this reason multiple phylogenies, 
@@
 In this example we use a standard generative tree model $\mathit{p_{G}^0}$, 
-namely the Yule (pure-birth) tree prior. 
+namely the Yule (pure-birth) tree model. 
 We choose to use a small tree with six taxa, to keep
@@
 In the previous example we selected the inference model $\mathit{I}$ to match 
-with a known generative tree prior $\mathit{p_{G}^0}$.
+with a known generative tree model $\mathit{p_{G}^0}$.
 However, a novel tree prior $\mathit{p_{G}}$ is by (our) definition
@@
 larger than those of the twin tree. 
-This is the error made in inference 
-by not using the generating species tree prior.
+This is the error made by the mismatch
+between the generating species tree model and the tree prior used in inference.
 

### 2019-11-18 content.tex
EOD

@@
 alignment_params <- create_alignment_params(
-  clock_model = create_strict_clock_model(),
-  site_model = create_jc69_site_model(),
+  sim_true_alignment_fun = get_sim_true_alignment_with_std_site_model_fun(
     mutation_rate = 0.1,
-  root_sequence = create_blocked_dna(length = 1000)
+    site_model = create_jc69_site_model()
+  ),
+  root_sequence = create_blocked_dna(length = 1000),
+  rng_seed = rng_seed
 )
@@
 RJCB created, tested and revised the package.
-GL done major contributions to the package.
+GL provided major contributions to the package.
 RJCB wrote the first draft of the manuscript, 

### 2019-12-01 content.tex
Verify all code works

@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \verb;error_function; &
+    \verb;error_fun; &
     Specifies how to measure the error &
@@
 ]
-usethis::install_github("richelbilderbeek/pirouette")
+remotes::install_github("richelbilderbeek/pirouette")
 \end{lstlisting}
@@
   ),
-  root_sequence = create_blocked_dna(length = 1000),
-  rng_seed = rng_seed
+  root_sequence = create_blocked_dna(length = 1000)
 )
@@
     site_model = create_jc69_site_model(),
-    mcmc = create_mcmc(store_every = 1000)
+    mcmc = create_mcmc()
   )
@@
 error_measure_params <- create_error_measure_params(
-  error_function = get_nltt_error_function(),
+  error_fun = get_nltt_error_fun(),
   burn_in_fraction = 0.1
@@
 twinning_params <- create_twinning_params(
-  twin_model = "birth_death",
-  method = "random_tree"
+  sim_twin_tree_fun = get_sim_bd_twin_tree_fun(),
+  sim_twin_alignment_fun = get_sim_twin_alignment_with_std_site_model_fun()
 )

### 2019-12-01 scripts/code.R

@@ -1,0 +1,62 @@
+remotes::install_github("richelbilderbeek/pirouette")
+library(pirouette)
+install_beast2()
+install_beast2_pkg("NS")
+phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
+alignment_params <- create_alignment_params(
+  sim_true_alignment_fun = get_sim_true_alignment_with_std_site_model_fun(
+    mutation_rate = 0.1,
+    site_model = create_jc69_site_model()
+  ),
+  root_sequence = create_blocked_dna(length = 1000)
+)
+generative_experiment <- create_experiment(
+  inference_conditions = create_inference_conditions(
+    model_type = "generative", 
+    run_if = "always"
+  ), 
+  inference_model = create_inference_model(
+    tree_prior = create_yule_tree_prior(),
+    clock_model = create_strict_clock_model(), 
+    site_model = create_jc69_site_model(),
+    mcmc = create_mcmc()
+  )
+)
+experiments <- list(generative_experiment)
+error_measure_params <- create_error_measure_params(
+  error_fun = get_nltt_error_fun(),
+  burn_in_fraction = 0.1
+)
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments,
+  error_measure_params = error_measure_params
+)
+errors <- pir_run(
+  phylogeny = phylogeny,
+  pir_params = pir_params
+)
+pir_plot(errors)
+candidate_experiments <- create_all_experiments(
+  exclude_model = generative_experiment$inference_model
+)
+experiments <- c(
+  list(generative_experiment),
+  candidate_experiments
+)
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments
+)
+phylogeny  <- ape::read.tree(
+  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
+)
+twinning_params <- create_twinning_params(
+  sim_twin_tree_fun = get_sim_bd_twin_tree_fun(),
+  sim_twin_alignment_fun = get_sim_twin_alignment_with_std_site_model_fun()
+)
+pir_params <- create_pir_params(
+  alignment_params = alignment_params,
+  experiments = experiments,
+  twinning_params = twinning_params
+)

### 2019-12-01 scripts/run_code.R

@@ -1,0 +1,46 @@
+setwd("~/GitHubs/pirouette_article")
+filename <- list.files(pattern = "content.tex", full.names = TRUE)
+
+# Extract the code from the TeX file
+text <- readLines(filename)
+start_lines <- which(grepl(pattern = "begin.*lstlisting", x = text))
+text[start_lines]
+end_lines <- which(grepl(pattern = "end.*lstlisting", x = text))
+text[end_lines]
+testit::assert(length(start_lines) == length(end_lines))
+
+code <- c()
+
+# Extract the code, with LaTeX stuff
+for (i in seq_along(start_lines)) {
+  from <- start_lines[i] + 1
+  to <- end_lines[i] - 1
+  code <- c(code, text[seq(from, to)])
+}
+code
+
+remove_lines <- which(grepl(pattern = "(]|  })", x = code))
+code <- code[-remove_lines]
+remove_lines <- which(grepl(pattern = "  (language|floatplacement|frame|label|caption) = .*", x = code))
+code <- code[-remove_lines]
+remove_lines <- which(grepl(pattern = "> .*", x = code))
+code <- code[-remove_lines]
+remove_lines <- which(
+  grepl(
+    pattern = "(resulting tree|Create a|generative model|Create the|that will|using explicit)",
+    x = code
+  )
+)
+code <- code[-remove_lines]
+# Do not assume pirouette is on CRAN yet
+remove_lines <- which(grepl(pattern = "install.packages", x = code))
+code <- code[-remove_lines]
+code
+
+writeLines(text = code, con = "~/GitHubs/pirouette_article/scripts/code.R")
+
+# Do so, because the script will install it
+library(pirouette)
+uninstall_beast2_pkg("NS")
+uninstall_beast2()
+source("~/GitHubs/pirouette_article/scripts/code.R")

### 2019-12-01 scripts/run_code.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+Rscript scripts/run_code.R

### 2019-12-17 content.tex
Use new naming

@@
 alignment_params <- create_alignment_params(
-  sim_true_alignment_fun = get_sim_true_alignment_with_std_site_model_fun(
+  sim_tral_fun = get_sim_tral_with_std_nsm_fun(
     mutation_rate = 0.1,
@@
   sim_twin_tree_fun = get_sim_bd_twin_tree_fun(),
-  sim_twin_alignment_fun = get_sim_twin_alignment_with_std_site_model_fun()
+  sim_twal_fun = get_sim_twal_with_std_nsm_fun()
 )

### 2019-12-17 scripts/code.R

@@ -4,7 +4,7 @@ install_beast2()
 install_beast2_pkg("NS")
 phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
 alignment_params <- create_alignment_params(
-  sim_true_alignment_fun = get_sim_true_alignment_with_std_nsm_fun(
+  sim_trall_fun = get_sim_trall_with_std_nsm_fun(
     mutation_rate = 0.1,
     site_model = create_jc69_site_model()
   ),
@@ -53,7 +53,7 @@ phylogeny  <- ape::read.tree(
 )
 twinning_params <- create_twinning_params(
   sim_twin_tree_fun = get_sim_bd_twin_tree_fun(),
-  sim_twin_alignment_fun = get_sim_twin_alignment_with_std_nsm_fun()
+  sim_twal_fun = get_sim_twal_with_std_nsm_fun()
 )
 pir_params <- create_pir_params(
   alignment_params = alignment_params,

### 2020-01-24 content.tex
Use pirouette_example pictures

@@
 \begin{figure}[ht]
-  \includegraphics[width=\textwidth]{example_1/true_tree.png}
+  \includegraphics[width=\textwidth]{pirouette_example_1/example_1_314/true_tree.png}
   \caption{The Yule tree, as created by Listing~\ref{lst:create_yule_tree}.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_1/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_1/example_1_314/errors.png}
   \caption{
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_2/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_2/example_2_314/errors.png}
   \caption{
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_3/true_tree.png}
+  \includegraphics[width=\textwidth]{pirouette_example_3/example_3_314/true_tree.png}
   \caption{The tree derived from an unknown diversification process, 
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{example_3/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_3/example_3_314/errors.png}
   \caption{

### 2020-01-24 supplement.tex
Use pirouette_example pictures

@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1/true_tree.png}
+      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1/true_alignment.png}
+      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_alignment.png}
     };   
     \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1/true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_posterior_gen.png}
     };   
     \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{example_1/true_error_violin_gen.png}
+      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_error_violin_gen.png}
     };   
@@
 
-\input{example_1/esses_gen.latex}
+\input{pirouette_example_1/example_1_314/esses_gen.latex}
 
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.5\textheight]{example_2/true_tree.png}
+      \includegraphics[height=0.5\textheight]{pirouette_example_2/example_2_314/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.5\textheight]{example_2/true_alignment.png}
+      \includegraphics[height=0.5\textheight]{pirouette_example_2/example_2_314/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_2/true_posterior_gen.png}
+      \includegraphics[height=0.45\textheight]{pirouette_example_2/example_2_314/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_2/true_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_2/example_2_314/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = CG, rectangle] {
-      \includegraphics[height=0.45\textheight]{example_2/true_posterior_best.png}
+      \includegraphics[height=0.45\textheight]{pirouette_example_2/example_2_314/true_posterior_best.png}
     };   
     \node[state] (DB) [right of = DG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_2/true_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_2/example_2_314/true_error_violin_best.png}
     };   
@@
 
-\input{example_2/esses_gen.latex}
+\input{pirouette_example_2/example_2_314/esses_gen.latex}
 
-\input{example_2/esses_best.latex}
+\input{pirouette_example_2/example_2_314/esses_best.latex}
 
-\input{example_2/evidence_true.latex}
+\input{pirouette_example_2/example_2_314/evidence_true.latex}
 
@@
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_6}.
+\url{https://github.com/richelbilderbeek/pirouette_example_3}.
 
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{example_6/true_tree.png}
+      \includegraphics[height=0.4\textheight]{pirouette_example_3/example_3_314/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.25\textheight]{example_6/true_alignment.png}
+      \includegraphics[height=0.25\textheight]{pirouette_example_3/example_3_314/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_6/true_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_6/true_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_6/true_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_6/true_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.4\textheight]{example_6/twin_tree.png}
+      \includegraphics[height=0.4\textheight]{pirouette_example_3/example_3_314/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.25\textheight]{example_6/twin_alignment.png}
+      \includegraphics[height=0.25\textheight]{pirouette_example_3/example_3_314/twin_alignment.png}
     };   
     \node[state] (CTG) [right of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_6/twin_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_posterior_gen.png}
     };   
     \node[state] (DTG) [below of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_6/twin_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_6/twin_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_posterior_best.png}
     };   
     \node[state] (DTB) [below of = CTB, rectangle] {
-      \includegraphics[height=0.3\textheight]{example_6/twin_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_error_violin_best.png}
     };   
@@
 
-\input{example_6/esses_gen.latex}
+\input{pirouette_example_3/example_3_314/esses_gen.latex}
 
-\input{example_6/esses_best.latex}
+\input{pirouette_example_3/example_3_314/esses_best.latex}
 
-\input{example_6/esses_twin_gen.latex}
+\input{pirouette_example_3/example_3_314/esses_twin_gen.latex}
 
-\input{example_6/esses_twin_best.latex}
+\input{pirouette_example_3/example_3_314/esses_twin_best.latex}
 
-\input{example_6/evidence_true.latex}
+\input{pirouette_example_3/example_3_314/evidence_true.latex}
 
-\input{example_6/evidence_twin.latex}
+\input{pirouette_example_3/example_3_314/evidence_twin.latex}

### 2020-01-24 supplement.tex
Made sections to put results figures

@@
 \input{pirouette_example_3/example_3_314/evidence_twin.latex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The effect of the number of taxa}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+pirouette_example_20
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The effect of DNA sequence length}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+pirouette_example_21
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The effect of non-clock like models}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+pirouette_example_17
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The effect of assuming a Yule tree prior on a Yule tree}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+pirouette_example_22
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The effect of assuming a Yule tree prior on a BD tree}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+pirouette_example_26
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The effect of differently common diversity-dependent trees}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+pirouette_example_23
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The effect of equal or equalized mutation rate in the twin alignment}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+pirouette_example_3 vs pirouette_example_18
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The effect of mutation rate}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+pirouette_example_24

### 2020-01-24 supplement.tex
Add all figures

@@
 
-pirouette_example_20
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_20}.
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_20/example_20_314/errors.png}
+  \caption{10 taxa}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_20/example_20_315/errors.png}
+  \caption{20 taxa}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_20/example_20_316/errors.png}
+  \caption{30 taxa}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_20/example_20_317/errors.png}
+  \caption{40 taxa}
+\end{figure}
 
@@
 
-pirouette_example_21
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_21}.
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_314/errors.png}
+  \caption{100 nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_315/errors.png}
+  \caption{248 nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_316/errors.png}
+  \caption{500 nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_317/errors.png}
+  \caption{1000 nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_318/errors.png}
+  \caption{2000 nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_319/errors.png}
+  \caption{4000 nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_320/errors.png}
+  \caption{8000 nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_321/errors.png}
+  \caption{16000 nucleotides}
+\end{figure}
 
@@
 
-pirouette_example_17
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_17}.
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_17/example_17_314/errors.png}
+  \caption{An unliked node substitution model}
+\end{figure}
 
@@
 
-pirouette_example_22
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_22}.
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_22/example_22_314/errors.png}
+  \caption{Assuming a Yule tree prior on a Yule tree}
+\end{figure}
 
@@
 
-pirouette_example_26
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_26}.
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_26/example_26_314/errors.png}
+  \caption{Assuming a Yule tree prior on a BD tree}
+\end{figure}
 
@@
 
-pirouette_example_23
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_23}.
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_314/errors.png}
+  \caption{Lowest likelihood}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_315/errors.png}
+  \caption{Between lowest and median likelihood}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_316/errors.png}
+  \caption{Median likelihood}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_317/errors.png}
+  \caption{Between median and highest likelihood}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_318/errors.png}
+  \caption{Highest likelihood}
+\end{figure}
 
@@
 
-pirouette_example_3 vs pirouette_example_18
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_3/example_3_314/errors.png}
+  \caption{Equal number of mutations}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_18/example_18_314/errors.png}
+  \caption{Equal mutation rate}
+\end{figure}
 
@@
 
-pirouette_example_24
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_24}.
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_314/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.0125}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_315/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.025}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_316/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.05}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_317/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.1}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_318/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.2}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_319/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.4}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_320/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.8}
+\end{figure}

### 2020-01-24 supplement.tex
Can create PDF
Updates from Overleaf

--- 2020-01-24 supplement.tex
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_tree.png}
+      \includegraphics[height=0.2\textheight]{example_1/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_alignment.png}
+      \includegraphics[height=0.2\textheight]{example_1/true_alignment.png}
     };   
     \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_posterior_gen.png}
+      \includegraphics[height=0.2\textheight]{example_1/true_posterior_gen.png}
     };   
     \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_error_violin_gen.png}
+      \includegraphics[height=0.2\textheight]{example_1/true_error_violin_gen.png}
     };   
@@
 
-\input{pirouette_example_1/example_1_314/esses_gen.latex}
+\input{example_1/esses_gen.latex}
 
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.5\textheight]{pirouette_example_2/example_2_314/true_tree.png}
+      \includegraphics[height=0.5\textheight]{example_2/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.5\textheight]{pirouette_example_2/example_2_314/true_alignment.png}
+      \includegraphics[height=0.5\textheight]{example_2/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.45\textheight]{pirouette_example_2/example_2_314/true_posterior_gen.png}
+      \includegraphics[height=0.45\textheight]{example_2/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_2/example_2_314/true_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{example_2/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = CG, rectangle] {
-      \includegraphics[height=0.45\textheight]{pirouette_example_2/example_2_314/true_posterior_best.png}
+      \includegraphics[height=0.45\textheight]{example_2/true_posterior_best.png}
     };   
     \node[state] (DB) [right of = DG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_2/example_2_314/true_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{example_2/true_error_violin_best.png}
     };   
@@
 
-\input{pirouette_example_2/example_2_314/esses_gen.latex}
+\input{example_2/esses_gen.latex}
 
-\input{pirouette_example_2/example_2_314/esses_best.latex}
+\input{example_2/esses_best.latex}
 
-\input{pirouette_example_2/example_2_314/evidence_true.latex}
+\input{example_2/evidence_true.latex}
 
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{pirouette_example_3/example_3_314/true_tree.png}
+      \includegraphics[height=0.4\textheight]{example_3/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.25\textheight]{pirouette_example_3/example_3_314/true_alignment.png}
+      \includegraphics[height=0.25\textheight]{example_3/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{example_3/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{example_3/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{example_3/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{example_3/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.4\textheight]{pirouette_example_3/example_3_314/twin_tree.png}
+      \includegraphics[height=0.4\textheight]{example_3/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.25\textheight]{pirouette_example_3/example_3_314/twin_alignment.png}
+      \includegraphics[height=0.25\textheight]{example_3/twin_alignment.png}
     };   
     \node[state] (CTG) [right of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{example_3/twin_posterior_gen.png}
     };   
     \node[state] (DTG) [below of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{example_3/twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{example_3/twin_posterior_best.png}
     };   
     \node[state] (DTB) [below of = CTB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{example_3/twin_error_violin_best.png}
     };   
@@
 
-\input{pirouette_example_3/example_3_314/esses_gen.latex}
+\input{example_3/esses_gen.latex}
 
-\input{pirouette_example_3/example_3_314/esses_best.latex}
+\input{example_3/esses_best.latex}
 
-\input{pirouette_example_3/example_3_314/esses_twin_gen.latex}
+\input{example_3/esses_twin_gen.latex}
 
-\input{pirouette_example_3/example_3_314/esses_twin_best.latex}
+\input{example_3/esses_twin_best.latex}
 
-\input{pirouette_example_3/example_3_314/evidence_true.latex}
+\input{example_3/evidence_true.latex}
 
-\input{pirouette_example_3/example_3_314/evidence_twin.latex}
+\input{example_3/evidence_twin.latex}

### 2020-01-24 scripts/copy_figs_from_examples_to_here.R

@@ -1,30 +1,0 @@
-# Copy files from the pirouette example to this folder
-#
-# [from_root/]pirouette_example_9/example_9_314/errors.png
-#                          ->
-# [to_root/]pirouette_example_9/example_9_314/errors.png
-#
-# Usage, from any folder:
-#
-#  Rscript copy_figs_from_examples_to_here.R
-#
-
-from_root <- "~/GitHubs"
-to_root <- "~/GitHubs/pirouette_article"
-
-# Rough search
-files <- list.files(path = from_root, pattern = "*.(png|latex)", recursive = TRUE)
-files
-
-# Only pirouette examples
-files <- stringr::str_match(string = files, pattern = "pirouette_example_.*(true|twin|errors|esses).*")
-files
-
-# Remove NA's
-files <- as.character(na.omit(files[, 1]))
-files
-
-from_files <- file.path(from_root, files)
-to_files <- file.path(to_root, files)
-for (dir_name in dirname(to_files)) dir.create(dir_name, recursive = TRUE, showWarnings = FALSE)
-file.copy(from_files, to_files, overwrite = TRUE)

### 2020-01-28 content.tex
Describe DD model, fix #79

@@
 We do so by separating the baseline error from the complete error 
-running the pipeline starting from a tree generated by an unknown 
-or non-standard tree model.
+running the pipeline starting from a tree generated by a non-standard tree model.
 
@@
 tree diversification model, assumed to be closely related to the Yule tree prior.
+We'll use the diversity-dependent tree model \cite{etienne2011diversity},
+which is a birth-death model with a speciation that is dependent on the
+number of species.
 We measure both the baseline 
@@
 
-We start from the tree generated by a non-standard (and unknown) diversification model, 
+We start from the tree generated by a diversity-dependent model,
 having six taxa and a crown age of ten:
@@
 ]
-phylogeny  <- ape::read.tree(
-  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
-)
+# NEW CODE HERE, 
+# depends on https://github.com/richelbilderbeek/pirouette_article/issues/75
+#
+# phylogeny  <- ape::read.tree(
+#   text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
+# )
 \end{lstlisting}

### 2020-01-28 content.tex
Described nLTT statistic (too?) briefly, fix #80

@@
 trees (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
-The default way to quantify the difference between two phylogenies
-is the nLTT statistic (\cite{janzen2015approximate}), but any 
-user-defined error statistic can be used.
+The user can specify a function to quantify the differences between
+the true and posterior trees. By default, the package uses the nLTT 
+statistic (\cite{janzen2015approximate}), which is the absolute difference
+between the normalized lineages-through-time plots of two trees.
 

### 2020-01-28 content.tex
Call a site model a site model, call a nucleotide substitution model as such, fix #84

@@
     Phylogenetic inference model to run BEAST2. Likewise the generative model $G$, 
-    its main components are the tree prior $\mathit{p_{I}}$, the clock model $\mathit{c_{I}}$ and the site model $\mathit{s_{I}}$. \\
+    its main components are the tree prior $\mathit{p_{I}}$, 
+    the clock model $\mathit{c_{I}}$ and the site model $\mathit{s_{I}}$. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 helps interpreting the resulting alignment.
-Supported nucleotide substitution model, which
-we will refer to as site models, 
+Supported nucleotide substitution model
 are JC, HKY, TN and GTR. Only the strict
@@
 
-The second step ($T_2$) of the twinning process simulates the twin alignment with the same clock model, site model and mutation rate 
+The second step ($T_2$) of the twinning process simulates the twin alignment 
+with the same clock model, site model and mutation rate 
 used to simulate the original alignment. 
@@
 0.1 mutations per unit time.
-We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) site model
+We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) nucleotide substitution model
 and a strict clock model as these are the simplest.
-A JC site model assumes that mutation rates between nucleotides are equal and 
+A JC model assumes that mutation rates between nucleotides are equal and 
 constant. 
@@
     Inference conditions and model.
-    JC: Jukes-Cantor site model.
+    JC: Jukes-Cantor nucleotide substitution model.
     strict: strict clock model.
@@
     BD: birth-death tree prior.
-    GTR: GTR site model.
+    GTR: GTR nucleotide substitution model.
     RLN: relaxed log-normal clock model.
@@
 of 40 experiments, 
-containing the inference models of all combinations of 4 site models, 
+containing the inference models of all combinations of 4 nucleotide substitution model, 
 2 clock models and 5 tree priors. All we need to add is to exclude the 

### 2020-01-28 scripts/copy_figs_from_examples_to_here.R
Use newer script to copy figures

@@ -1,0 +1,30 @@
+# Copy files from the pirouette example to this folder
+#
+# [from_root/]pirouette_example_9/example_9_314/errors.png
+#                          ->
+# [to_root/]pirouette_example_9/example_9_314/errors.png
+#
+# Usage, from any folder:
+#
+#  Rscript copy_figs_from_examples_to_here.R
+#
+
+from_root <- "~/GitHubs"
+to_root <- "~/GitHubs/pirouette_article"
+
+# Rough search
+files <- list.files(path = from_root, pattern = "*.(png|latex)", recursive = TRUE)
+files
+
+# Only pirouette examples
+files <- stringr::str_match(string = files, pattern = "pirouette_example_.*(true|twin|errors|esses).*")
+files
+
+# Remove NA's
+files <- as.character(na.omit(files[, 1]))
+files
+
+from_files <- file.path(from_root, files)
+to_files <- file.path(to_root, files)
+for (dir_name in dirname(to_files)) dir.create(dir_name, recursive = TRUE, showWarnings = FALSE)
+file.copy(from_files, to_files, overwrite = TRUE)

### 2020-01-28 content.tex
Updates from Overleaf

--- 2020-01-28 content.tex
@@
     CEP = coalescent exponential-population,
+    strict = strict clock model (\cite{zuckerkandl1965molecules}),
+    RLN = relaxed log-normal clock model (\cite{drummond2006relaxed}),
     JC = Jukes and Cantor (\cite{jukes1969evolution}), 
@@
     TN93 = Tamura and Nei (\cite{tamura1993estimation}), 
-    GTR = Generalized time-reversible model (\cite{tavare1986some})
-    RLN = relaxed log-normal clock model (\cite{drummond2006relaxed}).
+    GTR = Generalized time-reversible model (\cite{tavare1986some}).
   }

### 2020-01-28 content.tex

--- 2020-01-28 content.tex
@@
     Macroevolutionary diversification model &
-    Yule, BD, CBS, CCP, CEP \\
+    BD, CBS, CCP, CEP, Yule \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Clock for the DNA mutation rates &
-    strict, RLN \\
+    RLN, strict \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Nucleotide substitution model &
-    JC, HKY, TN93, GTR \\
+    GTR, HKY, JC, TN93 \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Most important parameter options.
-    Yule = pure birth model (\cite{yule})
     BD = birth death (\cite{nee1994reconstructed}), 
@@
     CEP = coalescent exponential-population,
+    Yule = pure birth model (\cite{yule}),
+    RLN = relaxed log-normal clock model (\cite{drummond2006relaxed}),
+    strict = strict clock model (\cite{zuckerkandl1965molecules}), 
+    GTR = Generalized time-reversible model (\cite{tavare1986some}), 
+    HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
     JC = Jukes and Cantor (\cite{jukes1969evolution}), 
-    HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
-    TN93 = Tamura and Nei (\cite{tamura1993estimation}), 
-    GTR = Generalized time-reversible model (\cite{tavare1986some})
-    RLN = relaxed log-normal clock model (\cite{drummond2006relaxed}).
+    TN93 = Tamura and Nei (\cite{tamura1993estimation}).
   }

### 2020-02-06 supplement.tex
I hope everything is OK before the Overleaf merge
Changes Giappo

--- 2020-01-24 supplement.tex
@@
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_26/example_26_314/errors.png}
-  \caption{Assuming a Yule tree prior on a BD tree}
-\end{figure}
+\richel{depends on https://github.com/richelbilderbeek/pirouette/issues/378}
+
+% \begin{figure}[H]
+%   \includegraphics[width=\textwidth]{pirouette_example_26/example_26_314/errors.png}
+%   \caption{Assuming a Yule tree prior on a BD tree}
+% \end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Here we show the results of a \verb;pirouette; run on a dataset of $N = 100$ \giovanni{N can be changed if 100 is too much} trees simulated under a diversity-dependent process.
+All the trees are simulated using the DDD package (\cite{DDD}).
+Our goal is to show how trees simulated by a process that is sensibly different from a standard tree prior will result in a significant divergence between true and twin distribution.
+To do so we set the simulation parameters (in this case: absolute speciation rate $\lambda_0 = 1.6$, extinction rate $\mu = 0.2$ and carrying capacity $K = 20$) in order to express clear characteristics of diversity dependence in the phylogenies.
+In fact, for a high speciation rate and a low carrying capacity we expect trees to exhibit high unbalance, compared with trees generated by a standard birth-death model.
+We can measure how unbalanced is a tree using the $\gamma$ statistics (\cite{pybus2000testing}).
+For this reason, we choose to measure the error made by the inference process using the $\gamma$ error statistics, which is currently implemented in \verb;pirouette;.
 

### 2020-02-06 content.tex

@@
 $\tau_{\mathit{T}}$, under the process $p_{T}$, 
-while preserving the topology.
+while preserving the topology. We simulate the new branching times using the TESS package (\cite{TESS, hohna2016tess}).
 
@@
 tree diversification model, assumed to be closely related to the Yule tree prior.
-We'll use the diversity-dependent tree model \cite{etienne2011diversity},
+We will use the diversity-dependent tree model \cite{DDD, etienne2011diversity},
 which is a birth-death model with a speciation that is dependent on the

### 2020-02-06 content.tex
EOD

@@
 
-We show the usage of \verb;pirouette; by gradually introducing
+We show the usage of \verb;pirouette; on an tree generated by a non-standard 
+tree model.
+
+\begin{lstlisting}[
+    language = R,
+    floatplacement = ht,
+    frame = single, 
+    label = {lst:create_dd_tree}, 
+    caption = {
+      Create a Yule tree. 
+      The resulting tree is shown in Figure~\ref{fig:yule_tree}.
+    }
+  ]
+phylogeny <- create_dd_tree(n_taxa = 6, crown_age = 10)
+\end{lstlisting}
+
+by gradually introducing
 its features.
@@
 
-\begin{lstlisting}[
-    language = R,
-    floatplacement = ht,
-    frame = single, 
-    label = {lst:create_yule_tree}, 
-    caption = {
-      Create a Yule tree. 
-      The resulting tree is shown in Figure~\ref{fig:yule_tree}.
-    }
-  ]
-phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
-\end{lstlisting}
-
 \begin{figure}[ht]
   \includegraphics[width=\textwidth]{pirouette_example_1/example_1_314/true_tree.png}
-  \caption{The Yule tree, as created by Listing~\ref{lst:create_yule_tree}.}
+  \caption{The Yule tree, as created by Listing~\ref{lst:create_dd_tree}.}
   \label{fig:yule_tree}
@@
 
-We will use the same tree as generated in \ref{lst:create_yule_tree}, as well 
+We will use the same tree as generated in \ref{lst:create_dd_tree}, as well 
 as the same alignment parameters as shown 
@@
 ]
-# NEW CODE HERE, 
-# depends on https://github.com/richelbilderbeek/pirouette_article/issues/75
-#
-# phylogeny  <- ape::read.tree(
-#   text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
-# )
+phylogeny <- create_dd_tree(n_taxa = 6, crown_age = 10)
 \end{lstlisting}

### 2020-02-06 content.tex
EOD

@@
 
-We show the usage of \verb;pirouette; on an tree generated by a non-standard 
-tree model.
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = ht,
-    frame = single, 
-    label = {lst:create_dd_tree}, 
-    caption = {
-      Create a Yule tree. 
-      The resulting tree is shown in Figure~\ref{fig:yule_tree}.
-    }
-  ]
-phylogeny <- create_dd_tree(n_taxa = 6, crown_age = 10)
-\end{lstlisting}
-
-by gradually introducing
-its features.
-First, to get an idea of the baseline error, 
-we measure the Bayesian inference error when we start from a phylogeny 
-generated under a known and standard tree model.
-Second, to establish that the true tree prior is indeed the best, 
-we also measure the inference error made by a
-best candidate inference model.
-Lastly, we quantify the impact of the tree prior in the Bayesian inference. 
-We do so by separating the baseline error from the complete error 
-running the pipeline starting from a tree generated by a non-standard tree model.
-
-All the figures shown in this section are shown as-is, without any
-aesthetical modifications. Figures showing the full workflow 
-and tables showing the effective sample sized (a measure
-of inference quality) can be found in the supplementary materials.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The generative and inference models are equal and standard.}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\richel{
 
@@
 used to generate the alignment.
+
 Measuring this difference allows us to know the baseline error
@@
 
-Now we can formulate the first example research question that \verb;pirouette; 
-can answer: "What is the inference error made on phylogenies created by a 
-standard diversification model?"
-
-In this example we use a standard generative tree model $\mathit{p_{G}^0}$, 
-namely the Yule (pure-birth) tree model. 
-We choose to use a small tree with six taxa, to keep
-the calculations short and the figure more readable.
-We pick a crown age of ten time units. This value is 
-completely arbitrary, but it ties in with the mutation rate 
-used in simulating an alignment in the next step.
-
-\begin{figure}[ht]
-  \includegraphics[width=\textwidth]{pirouette_example_1/example_1_314/true_tree.png}
-  \caption{The Yule tree, as created by Listing~\ref{lst:create_dd_tree}.}
-  \label{fig:yule_tree}
-\end{figure}
-
-The first step in \verb;pirouette; is to simulate a DNA alignment from the 
-given phylogeny, as described in Subsection~\ref{subsec:pipeline}.
-In this example, the root sequence consists of four blocks of 250 
-mononucleotides each, while the per-nucleotide mutation rate is 
-0.1 mutations per unit time.
-We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) nucleotide substitution model
-and a strict clock model as these are the simplest.
-A JC model assumes that mutation rates between nucleotides are equal and 
-constant. 
-A strict clock model assumes that the mutation rates 
-of all lineages are equal and constant.
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = ht,
-    frame = single,
-    label = {lst:create_alignment_params}, 
-    caption = {Create an alignment.}
-  ]
-alignment_params <- create_alignment_params(
-  sim_tral_fun = get_sim_tral_with_std_nsm_fun(
-    mutation_rate = 0.1,
-    site_model = create_jc69_site_model()
-  ),
-  root_sequence = create_blocked_dna(length = 1000)
-)
-\end{lstlisting}
-
-As the site and clock models used here are also the defaults, 
-the function arguments can be safely omitted: we just explicitly show 
-them for the sake of clarity.
-
-In the second step we state our experiment.
-We define an experiment $\mathit{X}$ as a combination of an inference model 
-$\mathit{I}$ and conditions $\mathit{C}$.
-In this example we choose $\mathit{I}$ to be the same inference model as 
-the generative one,
-i.e. the Yule tree prior $\mathit{p_{G}^0}$ and site 
-and clock models defined in $\mathit{A}$, 
-respectively Jukes-Cantor and strict clock.
-We specify in $\mathit{C}$ that the experiment will always be run.
-
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \texttt{model\_type} &
-    \texttt{run\_if} &
-    \texttt{do\_measure\_evidence} & 
-    \texttt{inference model} \\ 
-    \hline
-    generative &
-    always &
-    FALSE &
-    JC, strict, Yule \\
-    \hline
-  \end{tabular}
-  \caption{
-    Inference conditions and model.
-    JC: Jukes-Cantor nucleotide substitution model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-  }
-  \label{tab:RQ1}
-\end{table}
-
-Listing~\ref{lst:create_gen_experiment_explicit} shows how to
-set up this experiment:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_gen_experiment_explicit},
-  caption = {
-    Create an experiment with the generative model,
-    that will always be used in the actual inference, 
-    using explicit arguments.
-  }
-]
-generative_experiment <- create_experiment(
-  inference_conditions = create_inference_conditions(
-    model_type = "generative", 
-    run_if = "always"
-  ), 
-  inference_model = create_inference_model(
-    tree_prior = create_yule_tree_prior(),
-    clock_model = create_strict_clock_model(), 
-    site_model = create_jc69_site_model(),
-    mcmc = create_mcmc()
-  )
-)
-\end{lstlisting}
-
-Experiments must be bundled in a list to work, even if only one is provided, as 
-in this case:
-
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_gen_experiment},
-  caption = {
-    Create the experiments. In this case, we use only
-    an experiment with the generative model,
-    that will always be used in the actual inference.
-  }
-]
-experiments <- list(generative_experiment)
-\end{lstlisting}
-
-We also need to specify the error measurement parameters $\mathit{E}$.
-Here we choose the default $\mathit{E}$, which has a burn-in fraction 
-of 10\% and uses the nLTT statistic to
-measure the difference between phylogenies. For clarity,
-we create this setup explicitly here:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_error_measure_params},
-  caption = Calling \texttt{create\_error\_measure\_params}.
-]
-error_measure_params <- create_error_measure_params(
-  error_fun = get_nltt_error_fun(),
-  burn_in_fraction = 0.1
-)
-\end{lstlisting}
-
-We now have all the needed \verb;pirouette; parameters: the alignment 
-parameters, the experiments and the error measure parameters.
-These objects need to be bundled in a larger parameter structure, 
-using \verb;create_pir_params;:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_pir_params},
-  caption = Calling \texttt{create\_pir\_params}.
-]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  error_measure_params = error_measure_params
-)
-\end{lstlisting}
-
-We can finally use the given Yule tree and \verb;pir_params; to measure the 
-inference error made on phylogenies
-created by a standard diversification model:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:pir_run},
-  caption = Calling \texttt{pir\_run}.
-]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-\end{lstlisting}
-
-The error distribution can be plotted directly using \verb;pir_plot;:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:pir_plot},
-  caption = Calling \texttt{pir\_plot}.
-]
-pir_plot(errors)
-\end{lstlisting}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_1/example_1_314/errors.png}
-  \caption{
-    The inference error made 
-    when the generative and inference models are the same.
-    The vertical dashed line indicates the median error.
-  }
-  \label{fig:example_1}
-\end{figure}
-
-The resulting error distribution, as shown Figure~\ref{fig:example_1},
-shows the inference error 
-when $\mathit{I}$ matches with the generative model given 
-by $\mathit{A}$ and $\mathit{p_{G}^0}$.
-This error distribution can serve as a control,
-as it is obtained from a tree of which the generating tree model is standard and known.
-Obtaining this control is in fact an inherent part of \verb;pirouette;. 
-We call it 'twinning' and we will demonstrate it in section
-\ref{Comparing to background noise} below.
+...
+
+[first step]
+
+...
+
+...
 
@@
 
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \textbf{model\_type} &
-    \textbf{run\_if} &
-    \textbf{do\_measure\_evidence} & 
-    \textbf{inference model} \\ 
-    \hline
-    generative & always         & FALSE & JC, strict, Yule \\
-    candidate  & best candidate & TRUE  & JC, strict, BD   \\
-    candidate  & best candidate & TRUE  & JC, strict, CBS  \\
-    ...        & ...            & ...   & ...              \\
-    candidate  & best candidate & TRUE  & GTR, RLN, CCP    \\
-    candidate  & best candidate & TRUE  & GTR, RLN, CEP    \\
-    \hline
-  \end{tabular}
-  \caption{
-    Inference conditions and model.
-    JC: Jukes-Cantor nucleotide substitution model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-    BD: birth-death tree prior.
-    GTR: GTR nucleotide substitution model.
-    RLN: relaxed log-normal clock model.
-    CBS: coalescent Bayesian Skyline tree prior.
-    CCP: coalescent constant-population tree prior.
-    CEP: coalescent exponential-population tree prior.
-  }
-  \label{tab:RQ2}
-\end{table}
-
 We will use the same tree as generated in \ref{lst:create_dd_tree}, as well 
@@
 
-So far we have measured the inference error on a tree
-generated according to a known (and standard) tree diversification model. 
-The goal of \verb;pirouette; is to measure the expected impact of a novel tree prior.
-To do so, in this example, we will use a tree generated by a non-standard 
-tree diversification model, assumed to be closely related to the Yule tree prior.
-We will use the diversity-dependent tree model \cite{DDD, etienne2011diversity},
-which is a birth-death model with a speciation that is dependent on the
+} % richel
+
+We show the usage of \verb;pirouette; on an tree generated by a non-standard 
+tree model \richel{code at \url{https://github.com/richelbilderbeek/pirouette_example_30}}. 
+We use the diversity-dependent tree model \cite{DDD, etienne2011diversity},
+which is a birth-death model with a speciation rate that is dependent on the
 number of species.
-We measure both the baseline 
-error (that serves as a control) and the full inference error, because their 
-difference shows the impact of the tree model.
-We obtain the baseline error by using a twin tree (see Subsection~\ref{subsec:twinning}) for 
-both the generative and best candidate models.
-The research question this example answers is:
-"What is the inference error made from a phylogeny, 
-for both a generative model and the best candidate model?"
-
-We start from the tree generated by a diversity-dependent model,
-having six taxa and a crown age of ten:
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht,
-  frame = single, 
-  label = {lst:unknown_phylogeny},
+
+We choose to use a small tree with six taxa, to keep
+the calculations short and the figure more readable.
+We pick a crown age of ten time units. This value is 
+completely arbitrary, but it ties in with the mutation rate 
+used in simulating an alignment in the next step.
+
+\begin{lstlisting}[
+  language = R, 
+  floatplacement = ht,
+  frame = single, 
+  label = {lst:create_dd_tree},
   caption = A phylogeny generated by an unknown diversification model.
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_3/example_3_314/true_tree.png}
-  \caption{The tree derived from an unknown diversification process, 
-    as created by listing~\ref{lst:unknown_phylogeny}.
-    The scale overlaps with the lower branch, which is the default 
-    behavior for the plotting function used.
+  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/true_tree.png}
+  \caption{
+    The tree derived from a DD process, 
+    as created by listing~\ref{lst:create_dd_tree}.
   }
@@
 
-Most of the other settings are the same as before: 
-we reuse the alignment parameters (Listing~\ref{lst:create_alignment_params}), 
-as well as the experiments (Listing~\ref{lst:create_all_experiments}).
-This time, however, we enable twinning (see Subsection~\ref{subsec:twinning}),
+The first step in \verb;pirouette; is to simulate a DNA alignment from the 
+given phylogeny, as described in Subsection~\ref{subsec:pipeline}.
+In this example, the root sequence consists of four blocks of 250 
+mononucleotides each, while the per-nucleotide mutation rate is 
+0.1 mutations per unit time.
+We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) nucleotide substitution model
+and a strict clock model as these are the simplest.
+A JC model assumes that mutation rates between nucleotides are equal and 
+constant. 
+A strict clock model assumes that the mutation rates 
+of all lineages are equal and constant.
+
+\begin{lstlisting}[
+    language = R,
+    floatplacement = ht,
+    frame = single,
+    label = {lst:create_alignment_params}, 
+    caption = {Create an alignment.}
+  ]
+alignment_params <- create_alignment_params(
+  sim_tral_fun = get_sim_tral_with_std_nsm_fun(
+    mutation_rate = 0.1,
+    site_model = create_jc69_site_model()
+  ),
+  root_sequence = create_blocked_dna(length = 1000)
+)
+\end{lstlisting}
+
+As the site and clock models used here are also the defaults, 
+the function arguments can be safely omitted: we just explicitly show 
+them for the sake of clarity.
+
+In the second step we state our experiment.
+We define an experiment $\mathit{X}$ as a combination of an inference model 
+$\mathit{I}$ and conditions $\mathit{C}$.
+In this example we choose $\mathit{I}$ to be the same inference model as 
+the generative one,
+i.e. the Yule tree prior $\mathit{p_{G}^0}$ and site 
+and clock models defined in $\mathit{A}$, 
+respectively Jukes-Cantor and strict clock.
+We specify in $\mathit{C}$ that the experiment will always be run.
+
+\begin{table}
+  \begin{tabular}{ | c | c | c | l | }
+    \hline
+    \texttt{model\_type} &
+    \texttt{run\_if} &
+    \texttt{do\_measure\_evidence} & 
+    \texttt{inference model} \\ 
+    \hline
+    generative &
+    always &
+    FALSE &
+    JC, strict, Yule \\
+    \hline
+  \end{tabular}
+  \caption{
+    Inference conditions and model.
+    JC: Jukes-Cantor nucleotide substitution model.
+    strict: strict clock model.
+    Yule: Yule (pure-birth) tree prior.
+  }
+  \label{tab:RQ1}
+\end{table}
+
+Listing~\ref{lst:create_gen_experiment_explicit} shows how to
+set up this experiment:
+
+\begin{lstlisting}[
+  language = R,
+  floatplacement = ht,
+  frame = single,
+  label = {lst:create_gen_experiment_explicit},
+  caption = {
+    Create an experiment with the generative model,
+    that will always be used in the actual inference, 
+    using explicit arguments.
+  }
+]
+generative_experiment <- create_experiment(
+  inference_conditions = create_inference_conditions(
+    model_type = "generative", 
+    run_if = "always"
+  ), 
+  inference_model = create_inference_model(
+    tree_prior = create_yule_tree_prior(),
+    clock_model = create_strict_clock_model(), 
+    site_model = create_jc69_site_model(),
+    mcmc = create_mcmc()
+  )
+)
+\end{lstlisting}
+
+Here we specify a different set of experiments: we
+need to state that we already have an experiment for
+the generative model, as well as that we want all the other inference models 
+to compete. We call the competing models 'candidate models'.
+As model selection is commonly performed on the full list of available 
+candidate models, \verb;pirouette; has a dedicated function 
+for this choice: \verb;create_all_experiments; creates a full set 
+of 40 experiments, 
+containing the inference models of all combinations of 4 nucleotide substitution model, 
+2 clock models and 5 tree priors. All we need to add is to exclude the 
+inference model in the generative experiment:
+
+\begin{lstlisting}[
+  language = R, 
+  floatplacement = ht, 
+  frame = single, 
+  label = {lst:create_all_experiments},
+  caption = {
+    Create all 40 candidate experiments, except for the inference model of the generative model.
+  }
+]
+candidate_experiments <- create_all_experiments(
+  exclude_model = generative_experiment$inference_model
+)
+\end{lstlisting}
+
+We combine the generative and all the candidate models into one set of 
+experiments:
+
+\begin{lstlisting}[
+  language = R, 
+  floatplacement = ht, 
+  frame = single, 
+  label = {lst:combine_all_experiments},
+  caption = {
+    Create a collection of experiments, with 1
+    generative model, and 39 candidate models.
+  }
+]
+experiments <- c(
+  list(generative_experiment),
+  candidate_experiments
+)
+\end{lstlisting}
+
+We also need to specify the error measurement parameters $\mathit{E}$.
+Here we choose the default $\mathit{E}$, which has a burn-in fraction 
+of 10\% and uses the nLTT statistic to
+measure the difference between phylogenies. For clarity,
+we create this setup explicitly here:
+
+\begin{lstlisting}[
+  language = R,
+  floatplacement = ht,
+  frame = single,
+  label = {lst:create_error_measure_params},
+  caption = Calling \texttt{create\_error\_measure\_params}.
+]
+error_measure_params <- create_error_measure_params(
+  error_fun = get_nltt_error_fun(),
+  burn_in_fraction = 0.1
+)
+\end{lstlisting}
+
+We now have all the needed \verb;pirouette; parameters: the alignment 
+parameters, the experiments and the error measure parameters.
+
+Next, we enable twinning (see Subsection~\ref{subsec:twinning}),
 by creating twinning parameters.
@@
 
-We run \verb;pirouette; (Listing~\ref{lst:pir_run}) 
-and plot the results (Listing~\ref{lst:pir_plot}).
-The output is shown in Figure~\ref{fig:example_3}. 
+We can finally use the DD tree and \verb;pir_params; to measure the 
+inference error made on phylogenies
+created by a standard diversification model:
+
+\begin{lstlisting}[
+  language = R,
+  floatplacement = ht,
+  frame = single,
+  label = {lst:pir_run},
+  caption = Calling \texttt{pir\_run}.
+]
+errors <- pir_run(
+  phylogeny = phylogeny,
+  pir_params = pir_params
+)
+\end{lstlisting}
+
+The error distribution can be plotted directly using \verb;pir_plot;:
+
+\begin{lstlisting}[
+  language = R,
+  floatplacement = ht,
+  frame = single,
+  label = {lst:pir_plot},
+  caption = Calling \texttt{pir\_plot}.
+]
+pir_plot(errors)
+\end{lstlisting}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/errors.png}
+  \caption{
+    The inference error made 
+    for both a generative tree prior and best candidate model
+    compared with the error obtained for the twin tree.
+    Here, the 'twin' tree shows the baseline inference error.
+    Vertical dashed lines show the median error value per distribution.
+  }
+  \label{fig:example_30}
+\end{figure}
+
 While the error distributions using the best or generative model 
@@
 between the generating species tree model and the tree prior used in inference.
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_3/example_3_314/errors.png}
-  \caption{
-    The inference error made 
-    for both a generative tree prior and best candidate model
-    compared with the error obtained for the twin tree.
-    Here, the 'twin' tree shows the baseline inference error.
-    Vertical dashed lines show the median error value per distribution.
-  }
-  \label{fig:example_3}
-\end{figure}
 

### 2020-02-10 content.tex
Add comment

@@
 i.e. the prior probability of a species tree.
+\richel{break up next sentence}
 Many such diversification models (and their associated probability algorithms) 

### 2020-02-10 content.tex
Updated figures

@@
 
-\richel{
-
-\verb;pirouette; quantifies the influence of a new tree prior on BEAST2's 
-inference by measuring the discrepancy between a given/true tree and a posterior 
-distribution of phylogenies obtained as result of the inference process. 
-Due to stochasticity, posterior trees will generally differ from the given 
-phylogeny $\tau_{\mathit{G}}$, even when the tree 
-prior and alignment model used for inference are the same as those 
-used to generate the alignment.
-
-Measuring this difference allows us to know the baseline error
-of the \verb;pirouette; pipeline. We therefore define as 'standard tree priors' 
-all the tree priors that are available 
-to be used within an inference model (see Table~\ref{tab:options}).
-
-...
-
-[first step]
-
-...
-
-...
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The inference model may differ from the generative model}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In the previous example we selected the inference model $\mathit{I}$ to match 
-with a known generative tree model $\mathit{p_{G}^0}$.
-However, a novel tree prior $\mathit{p_{G}}$ is by (our) definition
-non-standard, and hence not part of a standard inference 
-model (i.e. an inference model using a standard tree prior).
-In such a case, the question is which standard tree prior should be used in 
-the inference. In this example, we do not only re-use our hand-picked 
-inference model, but we also pick the best inference model from a set of 
-inference models, i.e.  the model with 
-the highest evidence measured by its marginal likelihood. 
-We show the procedure using \verb;pirouette; to answer a second 
-research question: "What is the inference error made on a novel phylogeny 
-when using the best inference model, in comparison to a hand-picked model?
-
-We will use the same tree as generated in \ref{lst:create_dd_tree}, as well 
-as the same alignment parameters as shown 
-in Listing~\ref{lst:create_alignment_params}.
-
-Here we specify a different set of experiments: we
-need to state that we already have an experiment for
-the generative model, as well as that we want all the other inference models 
-to compete. We call the competing models 'candidate models'.
-As model selection is commonly performed on the full list of available 
-candidate models, \verb;pirouette; has a dedicated function 
-for this choice: \verb;create_all_experiments; creates a full set 
-of 40 experiments, 
-containing the inference models of all combinations of 4 nucleotide substitution model, 
-2 clock models and 5 tree priors. All we need to add is to exclude the 
-inference model in the generative experiment:
-
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht, 
-  frame = single, 
-  label = {lst:create_all_experiments},
-  caption = {
-    Create all 40 candidate experiments, except for the inference model of the generative model.
-  }
-]
-candidate_experiments <- create_all_experiments(
-  exclude_model = generative_experiment$inference_model
-)
-\end{lstlisting}
-
-We combine the generative and all the candidate models into one set of 
-experiments:
-
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht, 
-  frame = single, 
-  label = {lst:combine_all_experiments},
-  caption = {
-    Create a collection of experiments, with 1
-    generative model, and 39 candidate models.
-  }
-]
-experiments <- c(
-  list(generative_experiment),
-  candidate_experiments
-)
-\end{lstlisting}
-
-We can now create the complete \verb;pirouette; parameter set in the usual 
-way (which is the same as Listing~\ref{lst:create_pir_params}, but using the 
-defaults):
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_pir_params_defaults},
-  caption = Create a \texttt{pir\_params} with many defaults.
-]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments
-)
-\end{lstlisting}
-
-We run \verb;pirouette; (Listing~\ref{lst:pir_run}) 
-and plot the results (Listing~\ref{lst:pir_plot}) in Figure~\ref{fig:example_2}.
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_2/example_2_314/errors.png}
-  \caption{
-    The inference error for both generative and best candidate inference models.
-    Vertical dashed lines show the median error value per distribution.
-  }
-  \label{fig:example_2}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The generative model is non-standard}\label{Comparing to background noise}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-} % richel
-
 We show the usage of \verb;pirouette; on an tree generated by a non-standard 

### 2020-02-11 supplement.tex
Weave a story in the supplement

@@
 
-As in the main article, all the figures shown in this section are shown as-is, 
+All the figures shown in this section are shown as-is, 
 without any aesthetical modifications, with the exception that the arrangement
 of the sub-figures (for example, aligning true and twin tree
-horiztonally) is done manually.
+horizontally) is done manually.
+
+First, subsection \ref{subsec:main_example} shows extra figures and
+ diagnostics regarding the main example. Because the main example
+only uses one exemplary tree, subsection \ref{subsec:distribution}
+shows the result of using multiple trees, as generated by the
+same stochastic process as the main example.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Generative model only}
+\subsection{Main example}
+\label{subsec:main_example}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\richel{Issue \url{https://github.com/richelbilderbeek/pirouette_article/issues/75}}
+
+This subsection shows the diagnostics from the main example, which
+uses one tree.
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_1}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[H]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.25\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_alignment.png}
-    };   
-    \node[state] (C) [below of = B, rectangle] {
-      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_posterior_gen.png}
-    };   
-    \node[state] (D) [below of = C, rectangle] {
-      \includegraphics[height=0.2\textheight]{pirouette_example_1/example_1_314/true_error_violin_gen.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (C)
-      (C) edge [anchor = south] node {} (D)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_1_full_pipeline}
-  \caption{Generative model only: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{pirouette_example_1/example_1_314/esses_gen.latex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparing to other candidate models}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_2}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[H]
-  \centering
-  \resizebox {0.8\columnwidth} {!} {
-    \begin{tikzpicture}[
-      ->,>=stealth',shorten >=1pt,auto,
-      node distance=0.7\textheight, 
-      semithick
-    ]   
-    \tikzstyle{every state}=[]
-    \node[state, draw=none] (O) [] {
-    };   
-    \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.5\textheight]{pirouette_example_2/example_2_314/true_tree.png}
-    };   
-    \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.5\textheight]{pirouette_example_2/example_2_314/true_alignment.png}
-    };   
-    \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.45\textheight]{pirouette_example_2/example_2_314/true_posterior_gen.png}
-    };   
-    \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_2/example_2_314/true_error_violin_gen.png}
-    };   
-    \node[state] (CB) [right of = CG, rectangle] {
-      \includegraphics[height=0.45\textheight]{pirouette_example_2/example_2_314/true_posterior_best.png}
-    };   
-    \node[state] (DB) [right of = DG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_2/example_2_314/true_error_violin_best.png}
-    };   
-    \path 
-      (O) edge [anchor = south] node {} (A)
-      (A) edge [anchor = south] node {} (B)
-      (B) edge [anchor = south] node {} (CG)
-      (CG) edge [anchor = south] node {} (DG)
-      (B) edge [anchor = south] node {} (CB)
-      (CB) edge [anchor = south] node {} (DB)
-    ; 
-    \end{tikzpicture}
-  }
-  \label{fig:example_2_full_pipeline}
-  \caption{Comparing to other candidate models: full pipeline}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\input{pirouette_example_2/example_2_314/esses_gen.latex}
-
-\input{pirouette_example_2/example_2_314/esses_best.latex}
-
-\input{pirouette_example_2/example_2_314/evidence_true.latex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparing to background noise}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_3}.
+\url{https://github.com/richelbilderbeek/pirouette_example_30}.
 
@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{pirouette_example_3/example_3_314/true_tree.png}
+      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30_314/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.25\textheight]{pirouette_example_3/example_3_314/true_alignment.png}
+      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30_314/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/true_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.4\textheight]{pirouette_example_3/example_3_314/twin_tree.png}
+      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30_314/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.25\textheight]{pirouette_example_3/example_3_314/twin_alignment.png}
+      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30_314/twin_alignment.png}
     };   
     \node[state] (CTG) [right of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_posterior_gen.png}
     };   
     \node[state] (DTG) [below of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_posterior_best.png}
     };   
     \node[state] (DTB) [below of = CTB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_3/example_3_314/twin_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_error_violin_best.png}
     };   
@@
   }
-  \label{fig:example_3_full_pipeline}
+  \label{fig:example_30_full_pipeline}
   \caption{Comparing to background noise: full pipeline}
@@
 
-\input{pirouette_example_3/example_3_314/esses_gen.latex}
+\input{pirouette_example_30/example_30_314/esses_gen.latex}
 
-\input{pirouette_example_3/example_3_314/esses_best.latex}
+\input{pirouette_example_30/example_30_314/esses_best.latex}
 
-\input{pirouette_example_3/example_3_314/esses_twin_gen.latex}
+\input{pirouette_example_30/example_30_314/esses_twin_gen.latex}
 
-\input{pirouette_example_3/example_3_314/esses_twin_best.latex}
+\input{pirouette_example_30/example_30_314/esses_twin_best.latex}
 
-\input{pirouette_example_3/example_3_314/evidence_true.latex}
+\input{pirouette_example_30/example_30_314/evidence_true.latex}
 
-\input{pirouette_example_3/example_3_314/evidence_twin.latex}
+\input{pirouette_example_30/example_30_314/evidence_twin.latex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Using a distribution of trees}
+\label{subsec:distribution}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{Issue \url{https://github.com/richelbilderbeek/pirouette_article/issues/89}}
+
+This subsection extends the main example, by using multiple (instead of
+one) trees. These trees are produced by running a DD tree simulation
+with the same parameters as the main example, which are
+speciation rate \richel{?}, extinction rate $0.1$, and carrying
+capacity 40.
+
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_28}.
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_28/example_28_314/errors.png}
+  \caption{Error distribution from a distribution of trees}
+\end{figure}
 
@@
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_17}.
+\url{https://github.com/richelbilderbeek/pirouette_example_3007}.
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_17/example_17_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_3007/example_17_314/errors.png}
   \caption{An unliked node substitution model}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_3/example_3_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/errors.png}
   \caption{Equal number of mutations}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_18/example_18_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_3008/example_18_314/errors.png}
   \caption{Equal mutation rate}

### 2020-02-11 supplement.tex
Weave a better story at the supplementary materials

@@
 
-First, subsection \ref{subsec:main_example} shows extra figures and
- diagnostics regarding the main example. Because the main example
-only uses one exemplary tree, subsection \ref{subsec:distribution}
-shows the result of using multiple trees, as generated by the
-same stochastic process as the main example.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Overview}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For theoreticians:
+
+\itemize{
+  \item subsection \ref{subsec:main_example}: extra figures and
+    diagnostics regarding the main example.
+  \item subsection \ref{subsec:distribution}: the result of using 
+    multiple trees, as generated by the
+    same stochastic process as the main example
+}
+
+For empiricists:
+
+\itemize{
+  \item subsection \ref{subsec:n_taxa}: the effect of the number
+    of taxa
+  \item subsection \ref{subsec:n_nucleotides}: the effect of the DNA
+    alignment sequence length
+  \item subsection \ref{subsec:non_clock}: the effect when the alignment
+    is generated by a non-clock like nucleotide substitution model
+  \item subsection \ref{subsec:simplest_correct_parameterization} shows the
+    effect when doing an inference in the simplest use case
+  \item subsection \ref{subsec:under_parameterization} shows the
+    effect when doing an inference with an under-parameterization
+  \item subsection \ref{subsec:better_label_needed} shows 
+    \richel{better description needed}
+
+}
 
@@
 
-\richel{Issue \url{https://github.com/richelbilderbeek/pirouette_article/issues/75}}
+\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/75}}
 
@@
 \subsection{The effect of the number of taxa}
+\label{subsec:n_taxa}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/63}}
+
+The main example uses \richel{?} taxa.
+One expects that the more taxa are present in a phylogeny,
+the better the inference will be, 
+thus the lower the errors will be.
+Here, we show the same results as the main example,
+except for a varying number of taxa.
+
+Note that, \richel{[some authors]} found that an increase in the number
+of taxa may counterintuitively decrease the inference.
 
@@
 \subsection{The effect of DNA sequence length}
+\label{subsec:n_nucleotides}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/64}}
+
+The main example uses a DNA alignment length of \richel{?} nucleotides.
+One expects that the more nucleotides are present in the alignment,
+the more information is available,
+thus the better the inference will be, 
+thus the lower the errors will be.
+Here, we show the same results as the main example,
+except for a varying DNA alignment sequence length.
 
@@
 \subsection{The effect of non-clock like models}
+\label{subsec:non_clock}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/87}}
+
+The main example generates an alignment using a strict clock model (that is,
+all species have an equal mutation rate).
+One may wonder, how well the inference will be for a different non-clock like
+model.
+Here, we show the same results as the main example,
+except that instead of generating the alignment with a strict clock
+model, we use an unlinked node substitution model (by Thijs Janzen,
+unpublished, code available at \url{https://github.com/thijsjanzen/nodeSub}),
+with a branch mutation rate of $0.1$ and a node mutation rate of $0.1$.
+
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_3007}.
+\url{https://github.com/richelbilderbeek/pirouette_example_17}.
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_3007/example_17_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_17/example_17_314/errors.png}
   \caption{An unliked node substitution model}
@@
 \subsection{The effect of assuming a Yule tree prior on a Yule tree}
+\label{subsec:simplest_correct_parameterization}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/66}}
+
+The main example uses a tree generated by a non-standard tree model.
+Here, we show the same results, with the only difference that
+the tree used is generated by simplest tree model (the Yule model),
+which we also assume as the (correct) tree prior.
+This example thus shows a parameterization at the correct level for the
+simplest case possible.
 
@@
 \subsection{The effect of assuming a Yule tree prior on a BD tree}
+\label{subsec:under_parameterization}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/67}}
+
+The main example uses a tree generated by a non-standard tree model.
+
+Here, we show the same results, with the difference that
+the tree used is generated by a birth-death (BD) tree model,
+where we assume it is generated by a Yule (or pure-birth) model.
+This example thus shows the effect of underparameterization.
 
@@
 \subsection{The effect of differently common diversity-dependent trees}
+\label{subsec:better_label_needed}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Here we show the results of a \verb;pirouette; run on a dataset of $N = 100$ \giovanni{N can be changed if 100 is too much} trees simulated under a diversity-dependent process.
+\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/68}}
+
+Here we show the results of a \verb;pirouette; run on a dataset 
+of $N = 100$ \giovanni{N can be changed if 100 is too much} 
+trees simulated under a diversity-dependent process. 
 All the trees are simulated using the DDD package (\cite{DDD}). 
-Our goal is to show how trees simulated by a process that is sensibly different from a standard tree prior will result in a significant divergence between true and twin distribution.
-To do so we set the simulation parameters (in this case: absolute speciation rate $\lambda_0 = 1.6$, extinction rate $\mu = 0.2$ and carrying capacity $K = 20$) in order to express clear characteristics of diversity dependence in the phylogenies.
-In fact, for a high speciation rate and a low carrying capacity we expect trees to exhibit high unbalance, compared with trees generated by a standard birth-death model.
-We can measure how unbalanced is a tree using the $\gamma$ statistics (\cite{pybus2000testing}).
-For this reason, we choose to measure the error made by the inference process using the $\gamma$ error statistics, which is currently implemented in \verb;pirouette;.
+Our goal is to show how trees simulated by a process that is 
+sensibly different from a standard tree prior will result in a 
+significant divergence between true and twin distribution. 
+To do so we set the simulation parameters (in this case: 
+absolute speciation rate $\lambda_0 = 1.6$, 
+extinction rate $\mu = 0.2$ and carrying capacity $K = 20$) 
+in order to express clear characteristics of diversity dependence in the 
+phylogenies. In fact, for a high speciation rate and a low carrying capacity, 
+we expect trees to exhibit high unbalance, compared with trees generated 
+by a standard birth-death model. 
+We can measure how unbalanced is a tree using the $\gamma$ 
+statistics (\cite{pybus2000testing}). For this reason, we choose to 
+measure the error made by the inference process using the $\gamma$ error statistics, 
+which is currently implemented in \verb;pirouette;.
 
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_3008/example_18_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_18/example_18_314/errors.png}
   \caption{Equal mutation rate}

### 2020-02-11 supplement.tex
Better layout

@@
 
-For theoreticians:
-
 \itemize{
-  \item subsection \ref{subsec:main_example}: extra figures and
+  \item{
+    subsection \ref{subsec:main_example}: extra figures and
     diagnostics regarding the main example.
-  \item subsection \ref{subsec:distribution}: the result of using 
+  }
+  \item{
+    subsection \ref{subsec:distribution}: the result of using 
     multiple trees, as generated by the
@@
   }
-
-For empiricists:
-
-\itemize{
-  \item subsection \ref{subsec:n_taxa}: the effect of the number
+  \item{
+    subsection \ref{subsec:n_taxa}: the effect of the number
     of taxa
-  \item subsection \ref{subsec:n_nucleotides}: the effect of the DNA
+  }
+  \item{
+    subsection \ref{subsec:n_nucleotides}: the effect of the DNA
     alignment sequence length
-  \item subsection \ref{subsec:non_clock}: the effect when the alignment
+  }
+  \item{
+    subsection \ref{subsec:non_clock}: the effect when the alignment
     is generated by a non-clock like nucleotide substitution model
-  \item subsection \ref{subsec:simplest_correct_parameterization} shows the
+  }
+  \item{
+    subsection \ref{subsec:simplest_correct_parameterization} shows the
     effect when doing an inference in the simplest use case
-  \item subsection \ref{subsec:under_parameterization} shows the
+  }
+  \item{
+    subsection \ref{subsec:under_parameterization} shows the
     effect when doing an inference with an under-parameterization
-  \item subsection \ref{subsec:better_label_needed} shows 
+  }
+  \item{
+    subsection \ref{subsec:better_label_needed} shows 
     \richel{better description needed}
-
+  }
 }

### 2020-02-11 supplement.tex
Nice story in all supmat subsections

@@
     effect when doing an inference with an under-parameterization
+  }
+  \item{
+    subsection \ref{subsec:different_n_mutations} shows the
+    effect when the twin alignment is allowed to have a different
+    number of mutations
+  }
+  \item{
+    subsection \ref{subsec:mutation_rate} shows the
+    effect of different mutation rates
   }
@@
 \subsection{The effect of equal or equalized mutation rate in the twin alignment}
+\label{subsec:different_n_mutations}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/errors.png}
-  \caption{Equal number of mutations}
-\end{figure}
+\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/69}}
+
+The main example uses a twin alignment that has the same number
+of mutations (as measured from the ancestral sequence) as the true alignment.
+
+Here, we show the same results, with the difference that
+the twin alignment uses the same mutation rate, yet is not guaranteed
+to have the same number of mutations.
+
+The code used in this part of the article can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_example_18}.
 
@@
 \subsection{The effect of mutation rate}
+\label{subsec:mutation_rate}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/70}}
+
+The main example uses a mutation rate such that all nucleotides,
+on average, mutate once over the history going from the
+ancestral sequence at the crown to the alignments at the tips.
+In this way, the alignment is expected to contain the maximum
+amount of information.
+
+Here, we show the same results, with the only difference that
+the mutations rates are changed.
+For a lower mutation rate, less mutations will be observed, thus
+genetic information will be less, thus the inference is expected
+to go worse, thus the errors are expected to increase.
+For a lower mutation rate, more double-mutations will occur,
+scrambling the alignments, resulting in less genetic information, 
+thus the inference is expected to go worse, 
+thus the errors are expected to increase.
 

### 2020-02-13 content.tex

@@
     as they allow joint estimation of the posterior distribution of the 
-    phylogenetic tree and the parameters of the underlying inference model.  
+    phylogenetic trees and the parameters of the underlying inference model.  
     An important ingredient of these Bayesian approaches is the species tree 
-    prior. While in principle the Bayesian framework allows for comparing 
+    prior.
+    \iffalse 
+    While in principle the Bayesian framework allows for comparing 
     different species tree priors and hence may elucidate the macroevolutionary 
@@
     probability are used. 
-    An open question is, how accurate the tree estimation is
-    when the real macroevolutionary processes are substantially different 
-    from those assumed in the tree prior. \\
+    \fi
+    In principle the Bayesian framework allows for comparing different tree priors, which may elucidate the macroevolutionary processes underlying the species tree.
+    In practice, however, only macroevolutionary models that allow for fast computation of the prior 
+    probability are used.
+    \iffalse
+    An open question is, how accurate the tree estimation is when the real macroevolutionary processes are substantially different from those assumed in the tree prior. \\
+    \fi
+    It is still unclear how accurate is the tree estimation when the real \giovanni{People will ask what we mean with "real".} macroevolutionary processes are substantially different from those assumed in the tree prior.
+\\
     \textbf{2. }
@@
 MrBayes [\cite{huelsenbeck2001mrbayes}]
-or RevBayes [\cite{hohna2016revbayes}], 
+or RevBayes [\cite{hohna2016revbayes}] 
 has been a major advance in constructing phylogenetic trees 
 from character data (usually nucleotide sequences) extracted 
-from extant (but also extinct) organisms, 
-and hence in our understanding of the main drivers 
+from extant (but also extinct, where possible) organisms, and hence in our understanding of the main drivers 
 and modes of diversification.
@@
 (constant-rate) birth-death [\cite{nee1994reconstructed}] models.
-These simple tree priors are most commonly used, as these 
-have sufficient biological complexity, while being computationally fast.
+These simple tree priors are the most commonly used, as they have sufficient biological complexity, while being computationally fast.
 BEAST's successor, BEAST2 [\cite{bouckaert2019beast}],
 has a package manager, that allows third-party users 
-to extend existing functionalities. For example, 
-one can add novel diversification models by writing a BEAST2 package 
-that contains the likelihood formula of a phylogeny 
-under the novel diversification model, 
-i.e. the prior probability of a species tree.
-\richel{break up next sentence}
-Many such diversification models (and their associated probability algorithms) 
-have been developed, e.g., models in which diversification is 
-time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
+to extend existing functionalities.
+\iffalse
+For example, one can add novel diversification models by writing a BEAST2 package that contains the likelihood formula of a phylogeny under the novel diversification model, i.e.
+the prior probability of a species tree.
+\fi
+For example, one can add a novel diversification model by implementing a BEAST2 package featuring a novel likelihood formula to assign a prior probability to the species tree, given model's macroevolutionary parameters.
+\richel{break up next sentence}\giovanni{did it}
+\iffalse
+Many diversification models (and their associated probability algorithms) have been developed, e.g., models in which diversification is time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}],
 or diversity-dependent [\cite{etienne2011diversity}],
@@
 or a combination of all these [\cite{herrera2018detecting}].
-Only a few of these diversification models are available as a BEAST2 package.
+\fi
+Many diversification models (and their associated probability algorithms) have been developed to account for possible mechanisms.
+There are models in which diversification rates are time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
+or diversity-dependent [\cite{etienne2011diversity}],
+or not homogeneous across the entire phylogeny [\cite{etienne2012conceptual}, 
+\cite{rabosky2014automatic}, \cite{alfaro2009nine}, 
+\cite{laudanno2018sls}], models that treat speciation as a process that takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
+or as a burst of simultaneous branching
+events [\cite{laudanno2018mbd}].
+Others account for trait-dependent regimes, in several ways: where diversification rates
+depend on a trait that has two [\cite{maddison2007estimating}], 
+or more [\cite{fitzjohn2012diversitree}] states,
+even concealed states [\cite{beaulieu2016detecting}] 
+or a combination of all these [\cite{herrera2018detecting}].
+Despite such a rich theoretical landscape, only a few of these diversification models are available as a BEAST2 package.
 
@@
 its performance in inference should be tested.
-Part of a model's performance is its ability to recover parameters 
-from simulated data with known 
+Part of a model's performance is its ability to recover parameters from simulated data with known 
 parameters (e.g. [\cite{etienne2014estimating}]), 
-where ideally the estimated parameter values closely match 
-the known/true values.
-
-Even when a diversification model passes the procedure described above,
-it is not necessarily used in Bayesian inference.
+where ideally the estimated parameter values closely match the known/true values.
+
+Even when a diversification model passes the procedure described above, it is not necessarily used in Bayesian inference.
 Bayesian phylogenetic inference often requires 
@@
 as this will reduce the influence of the tree prior.
-However, the assumption that tree prior choice is of low importance 
-must first be verified.
+However, the assumption that tree prior choice is of low importance must first be verified.
 
@@
 Also recently, Duchene et al. [\cite{duchene2018phylodynamic}] released
-a BEAST2 package to assess how well posterior predictive simulations
-recover a given tree when using the standard diversification models.
-These studies show how current diversification models compare
-to one another, but they do not help to assess 
+a BEAST2 package to assess how well posterior predictive simulations recover a given tree when using the standard diversification models.
+These studies show how current diversification models compare to one another, but they do not help to assess 
 the importance of a new tree prior.
@@
 by the new model. 
-Next, nucleotide sequences are simulated on this phylogeny. 
-Then, using the tree priors built-in into BEAST2, 
+Next, nucleotide sequences are simulated on \giovanni{from?} this phylogeny. 
+Then,
+\iffalse
+using the tree priors built-in into BEAST2, 
+\fi
+using BEAST2's built-in tree priors,
 a Bayesian posterior distribution of phylogenies is inferred. 
-We then compare the inferred and simulated phylogenies. 
+We then compare the inferred and simulated \giovanni{original?} phylogenies. 
 How to properly perform this comparison forms the heart of our method.
 Only new diversification models that result 
-in a large discrepancy between inferred and simulated phylogenies 
-will be worth the effort and computational burden 
-to implement a species tree prior for in a Bayesian framework.
-
-Our method is programmed as an R package called \verb;pirouette;.
+in a large discrepancy between inferred and simulated phylogenies will be worth the effort and computational burden to implement a species tree prior for in a Bayesian framework.
+
+Our method is programmed as an R package (\cite{R}) called \verb;pirouette;.
 \verb;pirouette; is built on \verb;babette; [\cite{bilderbeek2018babette}], 
@@
 
-\verb;pirouette; is written in the R programming language (\cite{R}).
 The goal of \verb;pirouette; is to quantify the importance of a tree prior.
@@
 for which they reduce to a standard tree prior. 
-For example, a protracted birth-death model \citep{etienne2012prolonging} 
-reduces to a standard birth-death model when the speciation-completion rate 
+For example, a protracted birth-death model [\cite{etienne2012prolonging}] reduces to a standard birth-death model when the speciation-completion rate 
 is infinite [i.e. rate $\lambda$ in \cite{etienne2014estimating}, 
@@
 our examples contain only one replicate 
-as they are for illustrative purposes only.
+as they are for illustrative purposes only. \giovanni{Is this true also for the post-review state of the manuscript?}
 
@@
 We summarize them in Table~\ref{tab:options} and Table~\ref{tab:definitions}.
+\giovanni{Pirouette has become very complex.
+A user might feel lost.
+What if we build a function to help the user to build the pir-params he or she needs?
+Maybe we can even program it in such a way the function asks direct questions to the user.
+If you think this could be a good idea I candidate to write a first version of such function.}
 
@@
 we show the usage of \verb;pirouette; by introducing 
-its features gradually, yet ending in quantifying the impact a tree prior has
-in Bayesian inference.
+its features gradually \giovanni{Is this true in the new version of the manuscript?}, yet ending in quantifying the impact a tree prior has in Bayesian inference.
 
@@
 We assume the user has a phylogeny simulated with the new diversification model.
+
+\giovanni{long comment start}
+A nice thing to do would be to exploit the novel pir\_runs function to create a wrapper that could work directly from the diversification model simulation function.
+
+Something like this:
+\begin{lstlisting}
+sim_fun <- DDD::dd_sim
+sim_pars <- list(
+    pars1 = c(0.2, 0.1, 1.5, 0.1),
+    age = 8
+)
+pir_paramses <- pirouette::create_pir_paramses(
+    seeds = seeds,
+    ...
+)
+seeds <- 1:100
+pir_out <- pirouette::pir_run_from_sim_model(
+    sim_model = sim_fun,
+    sim_pars = sim_pars,
+    pir_paramses = pir_paramses,
+    seeds = seeds
+)
+\end{lstlisting}
+In case you deem this is a good idea I candidate to write this function. Then you can test it and see if it works in a satisfying way.
+\giovanni{long comment end}
+
 The pipeline to assess the error BEAST2 makes in inferring this phylogeny 
@@
 a (DNA) root sequence, a (DNA) mutation rate, a clock model 
-and a nucleotide substitution model.
+and a nucleotide substitution model (which, for the sake of brevity, we will often refer to as "site model").
 The root sequence is the DNA sequence of the shared common ancestor,
@@
 helps interpreting the resulting alignment.
-Supported nucleotide substitution model
+Supported nucleotide substitution models
 are JC, HKY, TN and GTR. Only the strict
@@
 
-The second step (Fig.~\ref{fig:pipeline}, 3)
-selects one or more inference models $I$ from a set of inference 
-models $I_{1},\dots,I_{n}$. 
+The second step (Fig.~\ref{fig:pipeline}, 3a)
+selects one (or more) inference model(s) $I$ from a set of inference models $I_{1},\dots,I_{n}$. 
 We define an experiment $X_{i}$ as the combination of 
@@
 The evidence for an inference model is estimated by nested 
-sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package. 
+sampling [\cite{russel2019model}], using the \verb;NS; BEAST2 package. 
 We note that scripted use of BEAST2 packages is only possible under Linux and Mac.
@@
 using the simulated alignment (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 4a),
-and the inference models that were selected in the previous step (3). 
+and the inference models that were selected in the previous step (3a). 
 For each selected experiment a posterior distribution is inferred, using the 
@@
 For the option to select the best candidate model,
-\verb;pirouette; needs the "NS" BEAST2 package [\cite{maturana2018model}].
+\verb;pirouette; needs the "NS" BEAST2 package [\cite{russel2019model}].
 It can be installed from within R, using:

### 2020-02-13 content.tex
Use 'create_all_bd_experiments'

@@
 containing the inference models of all combinations of 4 nucleotide substitution model, 
-2 clock models and 5 tree priors. All we need to add is to exclude the 
+2 clock models and 2 birth-death tree priors. All we need to add is to exclude the 
 inference model in the generative experiment:
@@
   caption = {
-    Create all 40 candidate experiments, except for the inference model of the generative model.
+    Create all 16 candidate experiments, 
+    except for the inference model of the generative model.
   }
 ]
-candidate_experiments <- create_all_experiments(
+candidate_experiments <- create_all_bd_experiments(
   exclude_model = generative_experiment$inference_model

### 2020-02-14 content.tex
EOD

@@
 candidate models, \verb;pirouette; has a dedicated function 
-for this choice: \verb;create_all_experiments; creates a full set 
-of 40 experiments, 
+for this choice: \verb;create_all_bd_experiments; creates a full set 
+of 16 experiments, 
 containing the inference models of all combinations of 4 nucleotide substitution model, 
@@
     Create a collection of experiments, with 1
-    generative model, and 39 candidate models.
+    generative model, and 15 candidate models.
   }

### 2020-03-18 content.tex
Compiles, with stub figures for now

@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/true_tree.png}
+  \includegraphics[width=\textwidth]{pirouette_example_30/example_30/true_tree.png}
   \caption{
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_30/example_30/errors.png}
   \caption{

### 2020-03-18 supplement.tex
Compiles, with stub figures for now

@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30_314/true_tree.png}
+      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30_314/true_alignment.png}
+      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30_314/twin_tree.png}
+      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30_314/twin_alignment.png}
+      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30/twin_alignment.png}
     };   
     \node[state] (CTG) [right of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/twin_posterior_gen.png}
     };   
     \node[state] (DTG) [below of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/twin_posterior_best.png}
     };   
     \node[state] (DTB) [below of = CTB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/twin_error_violin_best.png}
     };   
@@
 
-\input{pirouette_example_30/example_30_314/esses_gen.latex}
+\input{pirouette_example_30/example_30/esses_gen.latex}
 
-\input{pirouette_example_30/example_30_314/esses_best.latex}
+\input{pirouette_example_30/example_30/esses_best.latex}
 
-\input{pirouette_example_30/example_30_314/esses_twin_gen.latex}
+\input{pirouette_example_30/example_30/esses_twin_gen.latex}
 
-\input{pirouette_example_30/example_30_314/esses_twin_best.latex}
+\input{pirouette_example_30/example_30/esses_twin_best.latex}
 
-\input{pirouette_example_30/example_30_314/evidence_true.latex}
+\input{pirouette_example_30/example_30/evidence_true.latex}
 
-\input{pirouette_example_30/example_30_314/evidence_twin.latex}
+\input{pirouette_example_30/example_30/evidence_twin.latex}
 
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_28/example_28_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_28/example_28/errors.png}
   \caption{Error distribution from a distribution of trees}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/example_20_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_20/example_20/errors.png}
   \caption{10 taxa}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/example_20_315/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_20/example_20/315/errors.png}
   \caption{20 taxa}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/example_20_316/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_20/example_20/316/errors.png}
   \caption{30 taxa}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/example_20_317/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_20/example_20/317/errors.png}
   \caption{40 taxa}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/errors.png}
+  \caption{10 taxa}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/314/errors.png}
   \caption{100 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_315/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/315/errors.png}
   \caption{248 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_316/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/316/errors.png}
   \caption{500 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_317/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/317/errors.png}
   \caption{1000 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_318/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/318/errors.png}
   \caption{2000 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_319/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/319/errors.png}
   \caption{4000 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_320/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/320/errors.png}
   \caption{8000 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21_321/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/321/errors.png}
   \caption{16000 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_17/example_17_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_17/example_17/errors.png}
   \caption{An unliked node substitution model}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_22/example_22_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_22/example_22/314/errors.png}
   \caption{Assuming a Yule tree prior on a Yule tree}
@@
 % \begin{figure}[H]
-%   \includegraphics[width=\textwidth]{pirouette_example_26/example_26_314/errors.png}
+%   \includegraphics[width=\textwidth]{pirouette_example_26/example_26/314/errors.png}
 %   \caption{Assuming a Yule tree prior on a BD tree}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/314/errors.png}
   \caption{Lowest likelihood}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_315/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/315/errors.png}
   \caption{Between lowest and median likelihood}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_316/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/316/errors.png}
   \caption{Median likelihood}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_317/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/317/errors.png}
   \caption{Between median and highest likelihood}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23_318/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/318/errors.png}
   \caption{Highest likelihood}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_18/example_18_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_18/example_18/314/errors.png}
   \caption{Equal mutation rate}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/314/errors.png}
   \caption{Per-nucleotide mutation rate of 0.0125}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_315/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/315/errors.png}
   \caption{Per-nucleotide mutation rate of 0.025}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_316/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/316/errors.png}
   \caption{Per-nucleotide mutation rate of 0.05}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_317/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/317/errors.png}
   \caption{Per-nucleotide mutation rate of 0.1}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_318/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/318/errors.png}
   \caption{Per-nucleotide mutation rate of 0.2}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_319/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/319/errors.png}
   \caption{Per-nucleotide mutation rate of 0.4}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24_320/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/320/errors.png}
   \caption{Per-nucleotide mutation rate of 0.8}

### 2020-03-18 content.tex
Changes @Giappo

--- 2020-02-14 content.tex
@@
 For example, one can add a novel diversification model by implementing a BEAST2 package featuring a novel likelihood formula to assign a prior probability to the species tree, given model's macroevolutionary parameters.
-\richel{break up next sentence}\giovanni{did it}
 \iffalse
@@
 \cite{rabosky2014automatic}, \cite{alfaro2009nine}, 
-\cite{laudanno2018sls}], models that treat speciation as a process that takes 
+\cite{laudanno2020sls}], models that treat speciation as a process that takes 
 time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
@@
 \cite{rabosky2014automatic}, \cite{alfaro2009nine}, 
-\cite{laudanno2018sls}], models that treat speciation as a process that takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
+\cite{laudanno2020sls}], models that treat speciation as a process that takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
 or as a burst of simultaneous branching

### 2020-03-23 supplement.tex
Update
Fill in supmat

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/75}}
 
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_28/example_28/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
   \caption{Error distribution from a distribution of trees}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/example_20/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_20/errors_5.png}
+  \caption{5 taxa}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_20/errors_10.png}
   \caption{10 taxa}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/example_20/315/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_20/errors_20.png}
   \caption{20 taxa}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/example_20/316/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_20/errors_30.png}
   \caption{30 taxa}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/example_20/317/errors.png}
-  \caption{40 taxa}
 \end{figure}
@@
 
-The main example uses a DNA alignment length of \richel{?} nucleotides.
+The main example uses a DNA alignment length of 1000 nucleotides.
 One expects that the more nucleotides are present in the alignment,

### 2020-03-23 supplement.tex
Show effect of mutation rate

@@
 
-\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/64}}
-
 The main example uses a DNA alignment length of 1000 nucleotides.
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/errors.png}
-  \caption{10 taxa}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/errors_100.png}
   \caption{100 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/315/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/errors_248.png}
   \caption{248 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/316/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/errors_500.png}
   \caption{500 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/317/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/errors_1000.png}
   \caption{1000 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/318/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/errors_2000.png}
   \caption{2000 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/319/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/errors_4000.png}
   \caption{4000 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/320/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/errors_8000.png}
   \caption{8000 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/example_21/321/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_21/errors_16000.png}
   \caption{16000 nucleotides}

### 2020-03-23 supplement.tex
More pics in supmat

@@
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_17}.
+\url{https://github.com/richelbilderbeek/pirouette_example_27}.
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_17/example_17/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_27/errors.png}
   \caption{An unliked node substitution model}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/66}}
 
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_22/example_22/314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_22/errors.png}
   \caption{Assuming a Yule tree prior on a Yule tree}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/67}}
 
@@
 
-\richel{depends on https://github.com/richelbilderbeek/pirouette/issues/378}
-
-% \begin{figure}[H]
-%   \includegraphics[width=\textwidth]{pirouette_example_26/example_26/314/errors.png}
-%   \caption{Assuming a Yule tree prior on a BD tree}
-% \end{figure}
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_26/errors.png}
+  \caption{Assuming a Yule tree prior on a BD tree}
+\end{figure}
 

### 2020-03-23 supplement.tex
More pics in supmat

@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_23/errors_low.png}
   \caption{Lowest likelihood}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/315/errors.png}
-  \caption{Between lowest and median likelihood}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/316/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_23/errors_mid.png}
   \caption{Median likelihood}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/317/errors.png}
-  \caption{Between median and highest likelihood}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/example_23/318/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_23/errors_high.png}
   \caption{Highest likelihood}

### 2020-03-23 supplement.tex
Put all supmat pics in place

@@
 
-\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/69}}
-
 The main example uses a twin alignment that has the same number
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_18/example_18/314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_18/errors.png}
   \caption{Equal mutation rate}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/70}}
 
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/314/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/errors_1.png}
   \caption{Per-nucleotide mutation rate of 0.0125}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/315/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/errors_2.png}
   \caption{Per-nucleotide mutation rate of 0.025}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/316/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/errors_3.png}
   \caption{Per-nucleotide mutation rate of 0.05}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/317/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/errors_4.png}
   \caption{Per-nucleotide mutation rate of 0.1}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/318/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/errors_5.png}
   \caption{Per-nucleotide mutation rate of 0.2}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/319/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/errors_6.png}
   \caption{Per-nucleotide mutation rate of 0.4}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/example_24/320/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_24/errors_7.png}
   \caption{Per-nucleotide mutation rate of 0.8}

### 2020-03-23 content.tex
Process feedback @Giappo (and accepted it in most of the cases). Thanks @Giappo!

@@
     prior.
-    \iffalse 
-    While in principle the Bayesian framework allows for comparing 
-    different species tree priors and hence may elucidate the macroevolutionary 
-    processes underlying the species tree, in practice only 
-    macroevolutionary models that allow for fast computation of the prior 
-    probability are used. 
-    \fi
-    In principle the Bayesian framework allows for comparing different tree priors, which may elucidate the macroevolutionary processes underlying the species tree.
-    In practice, however, only macroevolutionary models that allow for fast computation of the prior 
-    probability are used.
-    \iffalse
-    An open question is, how accurate the tree estimation is when the real macroevolutionary processes are substantially different from those assumed in the tree prior. \\
-    \fi
-    It is still unclear how accurate is the tree estimation when the real \giovanni{People will ask what we mean with "real".} macroevolutionary processes are substantially different from those assumed in the tree prior.
-\\
+    In principle, the Bayesian framework allows for comparing 
+    different tree priors, which may elucidate the macroevolutionary 
+    processes underlying the species tree.
+    In practice, however, only macroevolutionary models that allow for 
+    fast computation of the prior probability are used.
+    An open question is, how accurate the tree estimation 
+    is when the real macroevolutionary processes are substantially 
+    different from those assumed in the tree prior. \\
     \textbf{2. }
@@
 to extend existing functionalities.
-\iffalse
-For example, one can add novel diversification models by writing a BEAST2 package that contains the likelihood formula of a phylogeny under the novel diversification model, i.e.
-the prior probability of a species tree.
-\fi
-For example, one can add a novel diversification model by implementing a BEAST2 package featuring a novel likelihood formula to assign a prior probability to the species tree, given model's macroevolutionary parameters.
-\iffalse
+For example, one can add novel diversification models 
+by writing a BEAST2 package that contains the likelihood 
+formula of a phylogeny under the novel diversification model, 
+i.e. the prior probability of a species tree.
 Many diversification models (and their associated probability algorithms) have been developed, e.g., models in which diversification is time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}],
@@
 by the new model. 
-Next, nucleotide sequences are simulated on \giovanni{from?} this phylogeny. 
-Then,
-\iffalse
-using the tree priors built-in into BEAST2, 
-\fi
-using BEAST2's built-in tree priors,
+Next, nucleotide sequences are simulated that follow the same
+evolutionary history as the given phylogeny.
+
+Then, using BEAST2's built-in tree priors,
 a Bayesian posterior distribution of phylogenies is inferred. 
-We then compare the inferred and simulated \giovanni{original?} phylogenies. 
+We then compare the inferred with the original phylogenies. 
 How to properly perform this comparison forms the heart of our method.
@@
 when assessing a novel tree prior, 
-our examples contain only one replicate 
-as they are for illustrative purposes only. \giovanni{Is this true also for the post-review state of the manuscript?}
+our examples contain only one replicate,
+as the goal is to show the workings of \verb;pirouette;,
+instead of doing an extensive analysis.
+The supplementary materials, however, do show the results
+of relicated runs under multiple settings.
 
@@
 We summarize them in Table~\ref{tab:options} and Table~\ref{tab:definitions}.
-\giovanni{Pirouette has become very complex.
-A user might feel lost.
-What if we build a function to help the user to build the pir-params he or she needs?
-Maybe we can even program it in such a way the function asks direct questions to the user.
-If you think this could be a good idea I candidate to write a first version of such function.}
 
 Although many possible tests can be performed, 
-we show the usage of \verb;pirouette; by introducing 
-its features gradually \giovanni{Is this true in the new version of the manuscript?}, yet ending in quantifying the impact a tree prior has in Bayesian inference.
+we show the usage of \verb;pirouette; by 
+showcasing one single run in high detail,
+ending in quantifying the impact a tree prior has in Bayesian inference.
 
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \iffalse
-    $\mathit{T}$ & Twinning parameters & Twin tree creation procedure & 
-      $\mathit{\tau_{p}}$ \\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \fi
     \hline 
@@
 
-\giovanni{long comment start}
-A nice thing to do would be to exploit the novel pir\_runs function to create a wrapper that could work directly from the diversification model simulation function.
-
-Something like this:
-\begin{lstlisting}
-sim_fun <- DDD::dd_sim
-sim_pars <- list(
-    pars1 = c(0.2, 0.1, 1.5, 0.1),
-    age = 8
-)
-pir_paramses <- pirouette::create_pir_paramses(
-    seeds = seeds,
-    ...
-)
-seeds <- 1:100
-pir_out <- pirouette::pir_run_from_sim_model(
-    sim_model = sim_fun,
-    sim_pars = sim_pars,
-    pir_paramses = pir_paramses,
-    seeds = seeds
-)
-\end{lstlisting}
-In case you deem this is a good idea I candidate to write this function. Then you can test it and see if it works in a satisfying way.
-\giovanni{long comment end}
-
 The pipeline to assess the error BEAST2 makes in inferring this phylogeny 

### 2020-03-23 supplement.tex
Process feedback @Giappo (and accepted it in most of the cases). Thanks @Giappo!

@@
     \node[state] (A) [right of = O, rectangle] {
-      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30/true_tree.png}
+      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30_314/true_tree.png}
     };   
     \node[state] (B) [below of = A, rectangle] {
-      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30/true_alignment.png}
+      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30_314/true_alignment.png}
     };   
     \node[state] (CG) [below of = B, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/true_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_posterior_gen.png}
     };   
     \node[state] (DG) [below of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/true_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_error_violin_gen.png}
     };   
     \node[state] (CB) [right of = CG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/true_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_posterior_best.png}
     };   
     \node[state] (DB) [below of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/true_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/true_error_violin_best.png}
     };   
     \node[state] (AT) [right of = A, rectangle, node distance=0.8\textheight] {
-      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30/twin_tree.png}
+      \includegraphics[height=0.4\textheight]{pirouette_example_30/example_30_314/twin_tree.png}
     };   
     \node[state] (BT) [below of = AT, rectangle] {
-      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30/twin_alignment.png}
+      \includegraphics[height=0.25\textheight]{pirouette_example_30/example_30_314/twin_alignment.png}
     };   
     \node[state] (CTG) [right of = CB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/twin_posterior_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_posterior_gen.png}
     };   
     \node[state] (DTG) [below of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/twin_error_violin_gen.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_error_violin_gen.png}
     };   
     \node[state] (CTB) [right of = CTG, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/twin_posterior_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_posterior_best.png}
     };   
     \node[state] (DTB) [below of = CTB, rectangle] {
-      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30/twin_error_violin_best.png}
+      \includegraphics[height=0.3\textheight]{pirouette_example_30/example_30_314/twin_error_violin_best.png}
     };   
@@
 
-\input{pirouette_example_30/example_30/esses_gen.latex}
+\input{pirouette_example_30/example_30_314/esses_gen.latex}
 
-\input{pirouette_example_30/example_30/esses_best.latex}
+\input{pirouette_example_30/example_30_314/esses_best.latex}
 
-\input{pirouette_example_30/example_30/esses_twin_gen.latex}
+\input{pirouette_example_30/example_30_314/esses_twin_gen.latex}
 
-\input{pirouette_example_30/example_30/esses_twin_best.latex}
+\input{pirouette_example_30/example_30_314/esses_twin_best.latex}
 
-\input{pirouette_example_30/example_30/evidence_true.latex}
+\input{pirouette_example_30/example_30_314/evidence_true.latex}
 
-\input{pirouette_example_30/example_30/evidence_twin.latex}
+\input{pirouette_example_30/example_30_314/evidence_twin.latex}
 
@@
 Here we show the results of a \verb;pirouette; run on a dataset 
-of $N = 100$ \giovanni{N can be changed if 100 is too much} 
-trees simulated under a diversity-dependent process. 
+of 2 trees simulated under a diversity-dependent process. 
 All the trees are simulated using the DDD package (\cite{DDD}). 

### 2020-03-23 content.tex
Fix LaTeX
Builds

@@
 or a combination of all these [\cite{herrera2018detecting}].
-\fi
+
 Many diversification models (and their associated probability algorithms) have been developed to account for possible mechanisms.
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_30/example_30/true_tree.png}
+  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/true_tree.png}
   \caption{
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_30/example_30/errors.png}
+  \includegraphics[width=\textwidth]{pirouette_example_30/errors.png}
   \caption{

### 2020-03-23 content.tex
3187 words

@@
 a (DNA) root sequence, a (DNA) mutation rate, a clock model 
-and a nucleotide substitution model (which, for the sake of brevity, we will often refer to as "site model").
+and a (DNA) site model.
 The root sequence is the DNA sequence of the shared common ancestor,
@@
 helps interpreting the resulting alignment.
-Supported nucleotide substitution models
-are JC, HKY, TN and GTR. Only the strict
+Supported nucleotide substitution models (part of a DNA
+site model) are JC, HKY, TN and GTR. Only the strict
 clock model is currently supported in this step.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Installation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\verb;pirouette; will be made available on CRAN from which 
-it can then be easily installed:
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install.packages("pirouette")
-\end{lstlisting}
-
-Until it is on CRAN, and for the most up-to-date version, 
-one can download and install the package from \verb;pirouette;'s GitHub 
-repository:
-
-\begin{lstlisting}[
-    language = R,
-    floatplacement = ht,
-    frame = single
-]
-remotes::install_github("richelbilderbeek/pirouette")
-\end{lstlisting}
-To start using \verb;pirouette;, 
-load its functions in the global namespace first:
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-library(pirouette)
-\end{lstlisting}
-Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
-This can be done from within R, using:
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install_beast2()
-\end{lstlisting}
-For the option to select the best candidate model,
-\verb;pirouette; needs the "NS" BEAST2 package [\cite{russel2019model}].
-It can be installed from within R, using:
-\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install_beast2_pkg("NS")
-\end{lstlisting}
-
-An overview of \verb;pirouette;'s main functions is shown in 
-Table~\ref{tab:functions}. 
-Their usage is demonstrated in the example code below.
-All \verb;pirouette;'s functions are documented,
-have a useful example and sensible defaults.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[h]
-  \centering
-  \begin{tabular}{ | l | l | l | }
-    \hline
-    \textbf{Name} & \textbf{Description} & \textbf{Listing} \\
-    \hline
-    \verb;pir_run; & Run \verb;pirouette; & \ref{lst:pir_run} \\
-    \verb;pir_plot; & Show the \verb;pirouette; results as a plot & \ref{lst:pir_plot} \\
-    \verb;create_pir_params; & Create the \verb;pirouette; parameters & \ref{lst:create_pir_params} \\
-    \hline
-    \verb;create_alignment_params; & Create the alignment parameters & \ref{lst:create_alignment_params} \\
-    \verb;create_twinning_params; & Create the twinning parameters & \ref{lst:create_twinning_params} \\
-    \verb;create_experiment; & Create one experiment & \ref{lst:create_gen_experiment_explicit} \\
-    \verb;create_error_measure_params; & Create the error measurement parameters & \ref{lst:create_error_measure_params} \\
-    \hline
-  \end{tabular}
-  \caption{
-    \texttt{pirouette}'s main functions, description and the number of the 
-    listing in which it is used. 
-  }
-  \label{tab:functions}
-\end{table}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Usage}
@@
 
-We show the usage of \verb;pirouette; on an tree generated by a non-standard 
+We show the usage of \verb;pirouette; on an tree generated 
+by a non-standard 
 tree model \richel{code at \url{https://github.com/richelbilderbeek/pirouette_example_30}}. 
@@
 the calculations short and the figure more readable.
-We pick a crown age of ten time units. This value is 
-completely arbitrary, but it ties in with the mutation rate 
-used in simulating an alignment in the next step.
+We pick a arbitrary crown age of ten time units.
 
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht,
-  frame = single, 
-  label = {lst:create_dd_tree},
-  caption = A phylogeny generated by an unknown diversification model.
-]
-phylogeny <- create_dd_tree(n_taxa = 6, crown_age = 10)
-\end{lstlisting}
 \begin{figure}[H]
@@
 
-The first step in \verb;pirouette; is to simulate a DNA alignment from the 
+The first step in \verb;pirouette; is to 
+simulate a DNA alignment from the 
 given phylogeny, as described in Subsection~\ref{subsec:pipeline}.
@@
 0.1 mutations per unit time.
+Together with the crown age of 10 time units, this results
+in an expected per-locus mutation probability of 1.
 We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) nucleotide substitution model
@@
 
-\begin{lstlisting}[
-    language = R,
-    floatplacement = ht,
-    frame = single,
-    label = {lst:create_alignment_params}, 
-    caption = {Create an alignment.}
-  ]
-alignment_params <- create_alignment_params(
-  sim_tral_fun = get_sim_tral_with_std_nsm_fun(
-    mutation_rate = 0.1,
-    site_model = create_jc69_site_model()
-  ),
-  root_sequence = create_blocked_dna(length = 1000)
-)
-\end{lstlisting}
+In the second step we state our experiment.
+We define an experiment $\mathit{X}$ 
+as a combination of an inference model 
+$\mathit{I}$ and conditions $\mathit{C}$.
+As a generative model, we assume a
+site model with a JC69 NSM,
+a strict clock model and a Yule tree prior.
+For our candidate models,
+we select all 16 experiments
+that are a combination of the 4 NSMs,
+2 clock models and 2 BD tree priors.
+From this set, we exclude the generative model.
 
-As the site and clock models used here are also the defaults, 
-the function arguments can be safely omitted: we just explicitly show 
-them for the sake of clarity.
-
-In the second step we state our experiment.
-We define an experiment $\mathit{X}$ as a combination of an inference model 
-$\mathit{I}$ and conditions $\mathit{C}$.
-In this example we choose $\mathit{I}$ to be the same inference model as 
-the generative one,
-i.e. the Yule tree prior $\mathit{p_{G}^0}$ and site 
-and clock models defined in $\mathit{A}$, 
-respectively Jukes-Cantor and strict clock.
-We specify in $\mathit{C}$ that the experiment will always be run.
-
-\begin{table}
-  \begin{tabular}{ | c | c | c | l | }
-    \hline
-    \texttt{model\_type} &
-    \texttt{run\_if} &
-    \texttt{do\_measure\_evidence} & 
-    \texttt{inference model} \\ 
-    \hline
-    generative &
-    always &
-    FALSE &
-    JC, strict, Yule \\
-    \hline
-  \end{tabular}
-  \caption{
-    Inference conditions and model.
-    JC: Jukes-Cantor nucleotide substitution model.
-    strict: strict clock model.
-    Yule: Yule (pure-birth) tree prior.
-  }
-  \label{tab:RQ1}
-\end{table}
-
-Listing~\ref{lst:create_gen_experiment_explicit} shows how to
-set up this experiment:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_gen_experiment_explicit},
-  caption = {
-    Create an experiment with the generative model,
-    that will always be used in the actual inference, 
-    using explicit arguments.
-  }
-]
-generative_experiment <- create_experiment(
-  inference_conditions = create_inference_conditions(
-    model_type = "generative", 
-    run_if = "always"
-  ), 
-  inference_model = create_inference_model(
-    tree_prior = create_yule_tree_prior(),
-    clock_model = create_strict_clock_model(), 
-    site_model = create_jc69_site_model(),
-    mcmc = create_mcmc()
-  )
-)
-\end{lstlisting}
-
-Here we specify a different set of experiments: we
-need to state that we already have an experiment for
-the generative model, as well as that we want all the other inference models 
-to compete. We call the competing models 'candidate models'.
-As model selection is commonly performed on the full list of available 
-candidate models, \verb;pirouette; has a dedicated function 
-for this choice: \verb;create_all_bd_experiments; creates a full set 
-of 16 experiments, 
-containing the inference models of all combinations of 4 nucleotide substitution model, 
-2 clock models and 2 birth-death tree priors. All we need to add is to exclude the 
-inference model in the generative experiment:
-
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht, 
-  frame = single, 
-  label = {lst:create_all_experiments},
-  caption = {
-    Create all 16 candidate experiments, 
-    except for the inference model of the generative model.
-  }
-]
-candidate_experiments <- create_all_bd_experiments(
-  exclude_model = generative_experiment$inference_model
-)
-\end{lstlisting}
-
-We combine the generative and all the candidate models into one set of 
-experiments:
-
-\begin{lstlisting}[
-  language = R, 
-  floatplacement = ht, 
-  frame = single, 
-  label = {lst:combine_all_experiments},
-  caption = {
-    Create a collection of experiments, with 1
-    generative model, and 15 candidate models.
-  }
-]
-experiments <- c(
-  list(generative_experiment),
-  candidate_experiments
-)
-\end{lstlisting}
-
-We also need to specify the error measurement parameters $\mathit{E}$.
-Here we choose the default $\mathit{E}$, which has a burn-in fraction 
-of 10\% and uses the nLTT statistic to
-measure the difference between phylogenies. For clarity,
-we create this setup explicitly here:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_error_measure_params},
-  caption = Calling \texttt{create\_error\_measure\_params}.
-]
-error_measure_params <- create_error_measure_params(
-  error_fun = get_nltt_error_fun(),
-  burn_in_fraction = 0.1
-)
-\end{lstlisting}
-
-We now have all the needed \verb;pirouette; parameters: the alignment 
-parameters, the experiments and the error measure parameters.
-
-Next, we enable twinning (see Subsection~\ref{subsec:twinning}),
-by creating twinning parameters.
-Creating these parameters is trivial with the default settings.
-For clarity, however, we explicitly show the most important arguments:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_twinning_params},
-  caption = Create the default twinning parameters.
-]
-twinning_params <- create_twinning_params(
-  sim_twin_tree_fun = get_sim_bd_twin_tree_fun(),
-  sim_twal_fun = get_sim_twal_with_std_nsm_fun()
-)
-\end{lstlisting}
-
-We combine all the parameters using \verb;create_pir_params;:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:create_pir_params_with_twinning},
-  caption = Create the default twinning parameters.
-]
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  twinning_params = twinning_params
-)
-\end{lstlisting}
-
-We can finally use the DD tree and \verb;pir_params; to measure the 
-inference error made on phylogenies
-created by a standard diversification model:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:pir_run},
-  caption = Calling \texttt{pir\_run}.
-]
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-\end{lstlisting}
-
-The error distribution can be plotted directly using \verb;pir_plot;:
-
-\begin{lstlisting}[
-  language = R,
-  floatplacement = ht,
-  frame = single,
-  label = {lst:pir_plot},
-  caption = Calling \texttt{pir\_plot}.
-]
-pir_plot(errors)
-\end{lstlisting}
+We use the nLTT statistic to
+measure the difference between phylogenies,
+with a burn-in of 10\%.
 

### 2020-03-23 supplement.tex
3187 words

@@
 \itemize{
+  \item{
+    subsection \ref{subsec:installation}: installation
+  }
+  \item{
+    subsection \ref{subsec:main_functions}: \verb;pirouette;'s 
+    main functions
+  }
   \item{
@@
 }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Installation}
+\label{subsec:installation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\verb;pirouette; will be made available on CRAN from which 
+it can then be easily installed:
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+install.packages("pirouette")
+\end{lstlisting}
+
+Until it is on CRAN, and for the most up-to-date version, 
+one can download and install the package from \verb;pirouette;'s GitHub 
+repository:
+
+\begin{lstlisting}[
+    language = R,
+    floatplacement = ht,
+    frame = single
+]
+remotes::install_github("richelbilderbeek/pirouette")
+\end{lstlisting}
+To start using \verb;pirouette;, 
+load its functions in the global namespace first:
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+library(pirouette)
+\end{lstlisting}
+Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
+This can be done from within R, using:
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+install_beast2()
+\end{lstlisting}
+For the option to select the best candidate model,
+\verb;pirouette; needs the "NS" BEAST2 package [\cite{russel2019model}].
+It can be installed from within R, using:
+\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
+install_beast2_pkg("NS")
+\end{lstlisting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Main functions}
+\label{subsec:main_functions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+An overview of \verb;pirouette;'s main functions is shown in 
+Table~\ref{tab:functions}. 
+All \verb;pirouette;'s functions are documented,
+have a useful example and sensible defaults.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[h]
+  \centering
+  \begin{tabular}{ | l | l | l | }
+    \hline
+    \textbf{Name} & \textbf{Description} & \textbf{Listing} \\
+    \hline
+    \verb;pir_run; & Run \verb;pirouette; & \ref{lst:pir_run} \\
+    \verb;pir_plot; & Show the \verb;pirouette; results as a plot & \ref{lst:pir_plot} \\
+    \verb;create_pir_params; & Create the \verb;pirouette; parameters & \ref{lst:create_pir_params} \\
+    \hline
+    \verb;create_alignment_params; & Create the alignment parameters & \ref{lst:create_alignment_params} \\
+    \verb;create_twinning_params; & Create the twinning parameters & \ref{lst:create_twinning_params} \\
+    \verb;create_experiment; & Create one experiment & \ref{lst:create_gen_experiment_explicit} \\
+    \verb;create_error_measure_params; & Create the error measurement parameters & \ref{lst:create_error_measure_params} \\
+    \hline
+  \end{tabular}
+  \caption{
+    \texttt{pirouette}'s main functions, description and the number of the 
+    listing in which it is used. 
+  }
+  \label{tab:functions}
+\end{table}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2020-03-23 content.tex
3222 words

@@
 as the 'generative tree model' $\mathit{p_{G}}$.
-Many tree priors have a parameter setting 
-for which they reduce to a standard tree prior. 
-For example, a protracted birth-death model [\cite{etienne2012prolonging}] reduces to a standard birth-death model when the speciation-completion rate 
-is infinite [i.e. rate $\lambda$ in \cite{etienne2014estimating}, 
-eqs.~(2b) and (2c)].
-When benchmarking a novel tree prior, 
+We define the birth-death (BD) model (\cite{nee1994reconstructed}) as
+the standard tree model, as many (non-standard) tree models 
+have a parameter setting for which they reduce to this model. 
+For example, a protracted birth-death (PBD) model [\cite{etienne2012prolonging}] 
+reduces to a BD model when the speciation-completion rate 
+is infinite [i.e. rate $\lambda$ in \cite{etienne2014estimating}, eqs.~(2b) and (2c)].
+When benchmarking a novel tree model, 
 one will typically construct phylogenies 
@@
 The supplementary materials, however, do show the results
-of relicated runs under multiple settings.
+of replicated runs under multiple settings.
 
-\verb;pirouette; is very flexible and allows the user 
-to specify a wide variety of custom settings. 
+\verb;pirouette; allows the user to specify a wide variety of custom settings. 
 These settings can be grouped in macro-sections, 
@@
 We summarize them in Table~\ref{tab:options} and Table~\ref{tab:definitions}.
-
-Although many possible tests can be performed, 
-we show the usage of \verb;pirouette; by 
-showcasing one single run in high detail,
-ending in quantifying the impact a tree prior has in Bayesian inference.
 
@@
     selected among standard diversification models; 
-    the default option is the standard birth-death model, with parameters estimated from the original phylogeny.
+    the default option is the standard BD model, 
+    with parameters estimated from the original phylogeny.
     A twin alignment (2b) is then simulated from the twin phylogeny 
@@
 
-We assume the user has a phylogeny simulated with the new diversification model.
-
 The pipeline to assess the error BEAST2 makes in inferring this phylogeny 
@@
 \begin{enumerate}
+  \item the user supplies one or (ideally) more phylogenies from a 
+    new diversification model
   \item from the given phylogeny an alignment is simulated 
@@
 \end{enumerate}
+
 The pipeline is visualized in Fig.~\ref{fig:pipeline}. 
@@
 
-The first step simulates a DNA alignment from a given 
-phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a)
-using the DNA alignment parameters.
-The DNA alignment parameters consist of 
-a (DNA) root sequence, a (DNA) mutation rate, a clock model 
-and a (DNA) site model.
+The first step simulates an alignment from the given 
+phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
+For the sake of clarity, here we will assume the alignment consists out
+of DNA sequences, but other heritable material such as amino acids also works.
+The user must specify a root sequence, a mutation rate and a site model.
 The root sequence is the DNA sequence of the shared common ancestor,
-and is set to four different equally-sized mononucleotide blocks by default, as this
-helps interpreting the resulting alignment.
+and is set to four different equally-sized mononucleotide blocks by default, 
+which helps interpreting the resulting alignment.
 Supported nucleotide substitution models (part of a DNA
-site model) are JC, HKY, TN and GTR. Only the strict
-clock model is currently supported in this step.
+site model) are JC, HKY, TN and GTR (see \ref{tab:options} for
+the meaning of these abbreviatons).
+JC is the default nucleotide subsitution mode (NSM),
+in which the nucleotide substition rates between all
+nucleotides are equal and constant over time.
+Note that the user cannot specify a clock model, as only the strict clock model 
+is implemented and thus always assumed.
 
@@
 The 'best' model is defined as the inference model with
-the highest evidence (a.k.a. marginal likelihood), given the alignment 
+the highest evidence (a.k.a. marginal likelihood), based on the alignment 
 simulated in the previous step.
@@
 between the normalized lineages-through-time plots of two trees.
+The nLTT statistic is chosen, as it can operate on any two trees (regardless
+of their crown ages and number of taxa) and its results have a clear range
+from zero to one. This normalized result makes it possible to compare trees 
+from a distribution of trees from any tree model.
 
@@
 \end{align}
-The default option for the diversification model $p_T$ is the standard birth-death model.
+The default option for the diversification model $p_T$ is the standard BD model.
 It is then possible to use the likelihood function 
 $L_{\mathit{T}}$ for this diversification model to find the parameters $\theta^{*}_{\mathit{T}}$ 
-(e.g. speciation and extinction rates, in case of a birth-death model) 
+(e.g. speciation and extinction rates, in case of a BD model) 
 that maximize this likelihood applied 
@@
 $\tau_{\mathit{T}}$, under the process $p_{T}$, 
-while preserving the topology. We simulate the new branching times using the TESS package (\cite{TESS, hohna2016tess}).
+while preserving the topology. 
+We simulate the new branching times using the TESS package (\cite{TESS, hohna2016tess}).
 
@@
 must be the same as in the true alignment in order to keep the information content 
-stored in both the true and twin alignments as similar as possible. We achieve this by simply simulating twin alignments until we
+stored in both the true and twin alignments as similar as possible. 
+We achieve this by simply simulating twin alignments until we
 obtain one that has the desired number of mutations.
@@
 Finally, if the goal is to evaluate BEAST2's performance on a non-standard tree prior, 
-the last source of stochasticity comes from phylogenies. 
-In fact, a single phylogeny cannot be considered 
-as fully representative of the model. 
-For this reason multiple phylogenies, 
-as well as an equal number of twins, 
-must be considered. 
-Having the error measure normalized (i.e. comprised in the interval $[0, 1]$), 
-it is possible to considerate the aggregated versions of the errors distributions 
-across all the runs.
-Therefore, if the number of considered phylogenies is high enough, 
+one must also consider the last source of stochasticity: 
+the different phylogenies a tree model generates.
+A single phylogeny cannot be considered as fully representative of the model. 
+For this reason multiple phylogenies, as well as an equal number of twins, must be considered. 
+If the number of considered phylogenies is high enough, 
 the comparison between the main pipeline's aggregated error distribution 
@@
 We show the usage of \verb;pirouette; on an tree generated 
-by a non-standard 
-tree model \richel{code at \url{https://github.com/richelbilderbeek/pirouette_example_30}}. 
-We use the diversity-dependent tree model \cite{DDD, etienne2011diversity},
-which is a birth-death model with a speciation rate that is dependent on the
-number of species.
+by the non-standard diversity-dependent tree model \cite{DDD, etienne2011diversity},
+which is a birth-death model with a speciation rate that is dependent on the number of species.
 
-We choose to use a small tree with six taxa, to keep
-the calculations short and the figure more readable.
-We pick a arbitrary crown age of ten time units.
+We use a small tree with six taxa, to keep the figure more readable,
+with an arbitrary crown age of ten time units.
 
@@
 in an expected per-locus mutation probability of 1.
-We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) nucleotide substitution model
+We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) NSM
 and a strict clock model as these are the simplest.

### 2020-03-23 content.tex
3248 words

@@
 have a parameter setting for which they reduce to this model. 
-For example, a protracted birth-death (PBD) model [\cite{etienne2012prolonging}] 
-reduces to a BD model when the speciation-completion rate 
-is infinite [i.e. rate $\lambda$ in \cite{etienne2014estimating}, eqs.~(2b) and (2c)].
+One such example is the diversity-dependent (DD) tree model (\cite{DDD, etienne2011diversity}):
+the DD model is a BD model with a speciation rate that is dependent 
+on the number of species and a carrying capacity. For a carrying capacity
+of infinity, the DD model reduces to a BD model.
 When benchmarking a novel tree model, 
@@
 We show the usage of \verb;pirouette; on an tree generated 
-by the non-standard diversity-dependent tree model \cite{DDD, etienne2011diversity},
-which is a birth-death model with a speciation rate that is dependent on the number of species.
+by the non-standard diversity-dependent (DD) tree model \cite{DDD, etienne2011diversity},
+which is a BD model with a speciation rate that is dependent on the number of species.
 
-We use a small tree with six taxa, to keep the figure more readable,
+We use the tree as shown in figure \ref{fig:dd_tree}.
+The with six taxa, to keep the figure more readable,
 with an arbitrary crown age of ten time units.
@@
   \caption{
-    The tree derived from a DD process, 
-    as created by listing~\ref{lst:create_dd_tree}.
+    The tree derived from a DD process.
   }
+  \label{fig:dd_tree}
 \end{figure}

### 2020-03-23 content.tex
3123 words

@@
 
-We use the tree as shown in figure \ref{fig:dd_tree}.
-The with six taxa, to keep the figure more readable,
-with an arbitrary crown age of ten time units.
+We use the tree as shown in figure \ref{fig:dd_tree},
+which has an arbitrary crown age of ten time units
+and six taxa, to keep the figure more readable.
 
@@
 
-The first step in \verb;pirouette; is to 
-simulate a DNA alignment from the 
-given phylogeny, as described in Subsection~\ref{subsec:pipeline}.
-In this example, the root sequence consists of four blocks of 250 
-mononucleotides each, while the per-nucleotide mutation rate is 
-0.1 mutations per unit time.
-Together with the crown age of 10 time units, this results
-in an expected per-locus mutation probability of 1.
-We use a Jukes-Cantor (JC, \cite{jukes1969evolution}) NSM
-and a strict clock model as these are the simplest.
-A JC model assumes that mutation rates between nucleotides are equal and 
-constant. 
-A strict clock model assumes that the mutation rates 
-of all lineages are equal and constant.
-
-In the second step we state our experiment.
-We define an experiment $\mathit{X}$ 
-as a combination of an inference model 
-$\mathit{I}$ and conditions $\mathit{C}$.
-As a generative model, we assume a
-site model with a JC69 NSM,
-a strict clock model and a Yule tree prior.
-For our candidate models,
-we select all 16 experiments
-that are a combination of the 4 NSMs,
-2 clock models and 2 BD tree priors.
-From this set, we exclude the generative model.
-
-We use the nLTT statistic to
-measure the difference between phylogenies,
-with a burn-in of 10\%.
+Using the default \verb;pirouette; settings,
+the error distribution shown in figure \label{fig:example_30}
+is produced.
 
@@
 
-While the error distributions using the best or generative model 
-as inference model are very similar, 
-the error distributions of the true tree are substantially 
-larger than those of the twin tree. 
-This is the error made by the mismatch
-between the generating species tree model and the tree prior used in inference.
+Focussing on the upper panel of figure \ref{fig:example_30},
+we can see that the error distributions of the generative model
+differ strongly between the true and twin tree. 
+This difference shows the extent of the mismatch between
+the actual tree model (which is DD) and the (BD) tree prior used.
+Because these distributions are distinctively different,
+the inference error we make when using an 
+incorrect (that is, BD) tree prior on a DD tree
+is profound.
+
+Comparing the upper and lower panel of figure \ref{fig:example_30}, 
+we can see that the best
+candidate model is only slightly better at inferring the true tree,
+compared to the assumed generative model, confirming that the 
+picked generative model was a good choice.
 
@@
 
-Figure~\ref{fig:example_3} illustrates the primary result of our pipeline: 
+Figure~\ref{fig:example_30} illustrates the primary result of our pipeline: 
 it shows the error distributions for the true tree and the twin tree 
@@
 \verb;ape; (\cite{ape}),
+\verb;assertive; (\cite{assertive}),
 \verb;babette; (\cite{bilderbeek2018babette}),
-\verb;becosys; (\cite{becosys}),
 \verb;DDD; (\cite{DDD}),
@@
 \verb;dplyr; (\cite{dplyr}),
-\verb;geiger; (\cite{geiger}),
 \verb;ggplot2; (\cite{ggplot2}),
@@
 \verb;nLTT; (\cite{nLTT}),
-\verb;PBD; (\cite{PBD}),
 \verb;phangorn; (\cite{phangorn}),

### 2020-03-25 content.tex
Done for today
Follow ?rsetienne's style of citing, using square brackets

@@
 
-Even when a diversification model passes the procedure described above, it is not necessarily used in Bayesian inference.
+Even when a diversification model passes the procedure described above, 
+it is not necessarily used in Bayesian inference.
 Bayesian phylogenetic inference often requires 
@@
 as this will reduce the influence of the tree prior.
-However, the assumption that tree prior choice is of low importance must first be verified.
+However, the assumption that tree prior choice is of low importance 
+must first be verified.
 
@@
 Also recently, Duchene et al. [\cite{duchene2018phylodynamic}] released
-a BEAST2 package to assess how well posterior predictive simulations recover a given tree when using the standard diversification models.
-These studies show how current diversification models compare to one another, but they do not help to assess 
-the importance of a new tree prior.
+a BEAST2 package to assess how well posterior predictive simulations 
+recover a given tree when using the standard diversification models.
+These studies show how current diversification models compare to one another, 
+but they do not help to assess the importance of a new tree prior.
 
@@
 
-Our method is programmed as an R package (\cite{R}) called \verb;pirouette;.
+Our method is programmed as an R package [\cite{R}] called \verb;pirouette;.
 \verb;pirouette; is built on \verb;babette; [\cite{bilderbeek2018babette}], 
@@
 as the 'generative tree model' $\mathit{p_{G}}$.
-We define the birth-death (BD) model (\cite{nee1994reconstructed}) as
+We define the birth-death (BD) model [\cite{nee1994reconstructed}] as
 the standard tree model, as many (non-standard) tree models 
 have a parameter setting for which they reduce to this model. 
-One such example is the diversity-dependent (DD) tree model (\cite{DDD, etienne2011diversity}):
+One such example is the diversity-dependent (DD) tree 
+model [\cite{DDD, etienne2011diversity}]:
 the DD model is a BD model with a speciation rate that is dependent 
@@
     Nucleotide substitution model &
-    GTR, HKY, JC, TN93 \\
+    GTR, HKY, JC, TN \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Most important parameter options.
-    BD = birth death (\cite{nee1994reconstructed}), 
-    CBS = coalescent Bayesian skyline (\cite{drummond2005bayesian}), 
+    BD = birth death [\cite{nee1994reconstructed}], 
+    CBS = coalescent Bayesian skyline [\cite{drummond2005bayesian}], 
     CCP = coalescent constant-population, 
     CEP = coalescent exponential-population,
-    Yule = pure birth model (\cite{yule}),
-    RLN = relaxed log-normal clock model (\cite{drummond2006relaxed}),
-    strict = strict clock model (\cite{zuckerkandl1965molecules}), 
-    GTR = Generalized time-reversible model (\cite{tavare1986some}), 
-    HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
-    JC = Jukes and Cantor (\cite{jukes1969evolution}), 
-    TN93 = Tamura and Nei (\cite{tamura1993estimation}).
+    Yule = pure birth model [\cite{yule}],
+    RLN = relaxed log-normal clock model [\cite{drummond2006relaxed}],
+    strict = strict clock model [\cite{zuckerkandl1965molecules}], 
+    GTR = Generalized time-reversible model [\cite{tavare1986some}], 
+    HKY = Hasegawa, Kishino and Yano [\cite{hasegawa1985dating}], 
+    JC = Jukes and Cantor [\cite{jukes1969evolution}], 
+    TN = Tamura and Nei [\cite{tamura1993estimation}].
   }
@@
 The user must specify a root sequence, a mutation rate and a site model.
-The root sequence is the DNA sequence of the shared common ancestor,
-and is set to four different equally-sized mononucleotide blocks by default, 
+The root sequence is the DNA sequence of the shared common ancestor of 
+all species, and is set to four different equally-sized 
+mononucleotide blocks by default, 
 which helps interpreting the resulting alignment.
 Supported nucleotide substitution models (part of a DNA
-site model) are JC, HKY, TN and GTR (see \ref{tab:options} for
+site model) are JC, HKY, TN and GTR (see Table \ref{tab:options} for
 the meaning of these abbreviatons).
@@
 nucleotides are equal and constant over time.
-Note that the user cannot specify a clock model, as only the strict clock model 
-is implemented and thus always assumed.
 
@@
 sampling [\cite{russel2019model}], using the \verb;NS; BEAST2 package.
-We note that scripted use of BEAST2 packages is only possible under Linux and Mac.
-Windows systems can do the model comparison for shorter DNA sequences 
-using the web interface of \verb;mcbette; [\cite{mcbette}].
 
@@
 the true and posterior trees. By default, the package uses the nLTT 
-statistic (\cite{janzen2015approximate}), which is the absolute difference
+statistic [\cite{janzen2015approximate}], which is the absolute difference
 between the normalized lineages-through-time plots of two trees.
@@
 while preserving the topology. 
-We simulate the new branching times using the TESS package (\cite{TESS, hohna2016tess}).
+We simulate the new branching times using the TESS 
+package [\cite{TESS, hohna2016tess}].
 
@@
 
-We use the tree as shown in figure \ref{fig:dd_tree},
+We use the tree as shown in Figure \ref{fig:dd_tree},
 which has an arbitrary crown age of ten time units
@@
 Using the default \verb;pirouette; settings,
-the error distribution shown in figure \label{fig:example_30}
+the error distribution shown in Figure \label{fig:example_30}
 is produced.
@@
 
-Focussing on the upper panel of figure \ref{fig:example_30},
+Focussing on the upper panel of Figure \ref{fig:example_30},
 we can see that the error distributions of the generative model
@@
 
-Comparing the upper and lower panel of figure \ref{fig:example_30}, 
+Comparing the upper and lower panel of Figure \ref{fig:example_30}, 
 we can see that the best
@@
 but we chose to provide the simplest (and fastest to run) examples.
+
+HIERO
+Note that the user cannot specify a clock model, as only the strict clock model 
+is implemented and thus always assumed.
+
+HIERO
+We note that scripted use of BEAST2 packages is only possible under Linux and Mac.
+Windows systems can do the model comparison for shorter DNA sequences 
+using the web interface of \verb;mcbette; [\cite{mcbette}].
 
@@
 \verb;pirouette; depends on multiple packages, which are:
-\verb;ape; (\cite{ape}),
-\verb;assertive; (\cite{assertive}),
-\verb;babette; (\cite{bilderbeek2018babette}),
-\verb;DDD; (\cite{DDD}),
-\verb;devtools; (\cite{devtools}),
-\verb;dplyr; (\cite{dplyr}),
-\verb;ggplot2; (\cite{ggplot2}),
-\verb;knitr; (\cite{knitr}),
-\verb;lintr; (\cite{lintr}),
-\verb;magrittr; (\cite{magrittr}),
-\verb;mcbette; (\cite{mcbette}),
-\verb;nLTT; (\cite{nLTT}),
-\verb;phangorn; (\cite{phangorn}),
-\verb;phytools; (\cite{phytools}),
-\verb;plyr; (\cite{plyr}),
-\verb;rappdirs; (\cite{rappdirs}),
-\verb;rmarkdown; (\cite{rmarkdown}),
-\verb;Rmpfr; (\cite{Rmpfr}),
-\verb;stringr; (\cite{stringr}),
-\verb;TESS; (\cite{TESS}),
-\verb;testit; (\cite{testit}), 
-\verb;testthat; (\cite{testthat}) and
-\verb;tidyr; (\cite{tidyr}).
+\verb;ape; [\cite{ape}],
+\verb;assertive; [\cite{assertive}],
+\verb;babette; [\cite{bilderbeek2018babette}],
+\verb;DDD; [\cite{DDD}],
+\verb;devtools; [\cite{devtools}],
+\verb;dplyr; [\cite{dplyr}],
+\verb;ggplot2; [\cite{ggplot2}],
+\verb;knitr; [\cite{knitr}],
+\verb;lintr; [\cite{lintr}],
+\verb;magrittr; [\cite{magrittr}],
+\verb;mcbette; [\cite{mcbette}],
+\verb;nLTT; [\cite{nLTT}],
+\verb;phangorn; [\cite{phangorn}],
+\verb;phytools; [\cite{phytools}],
+\verb;plyr; [\cite{plyr}],
+\verb;rappdirs; [\cite{rappdirs}],
+\verb;rmarkdown; [\cite{rmarkdown}],
+\verb;Rmpfr; [\cite{Rmpfr}],
+\verb;stringr; [\cite{stringr}],
+\verb;TESS; [\cite{TESS}],
+\verb;testit; [\cite{testit}], 
+\verb;testthat; [\cite{testthat}] and
+\verb;tidyr; [\cite{tidyr}].
 
@@
 uses a continuous integration service, has a code coverage of above 95\%
-and enforces the most commonly used R style guide (\cite{style_guide}).
+and enforces the most commonly used R style guide [\cite{style_guide}].
 

### 2020-03-25 content.tex
2856 words

@@
 removes the first 10\% of the posterior.
-From the remaining posterior \verb;pirouette; 
+From the remaining posterior, \verb;pirouette; 
 creates an error distribution, by measuring the difference
@@
 It is then possible to use the likelihood function 
-$L_{\mathit{T}}$ for this diversification model to find the parameters $\theta^{*}_{\mathit{T}}$ 
+$L_{\mathit{T}}$ for this diversification model to find 
+the parameters $\theta^{*}_{\mathit{T}}$ 
 (e.g. speciation and extinction rates, in case of a BD model) 
@@
 used to simulate the original alignment. 
-We also impose that, in the twin alignment, 
-the total number of mutations with respect to the root sequence 
-must be the same as in the true alignment in order to keep the information content 
-stored in both the true and twin alignments as similar as possible. 
-We achieve this by simply simulating twin alignments until we
-obtain one that has the desired number of mutations.
+The twin alignment can be simulated in any user-defined way.
+\verb;pirouette; supplies the option to have it simulated with
+the same mutation rate as the true alignment. By default, however,
+not only the same mutation rate is used, but also the total number of mutations
+matches the true alignment. The total number of mutations is defines
+as the amount of different nucleotides between the (known) root sequence
+compared to the sequences at the tips.
 
 The twin pipeline serves as a control: 
-even when the generating and inference models are identical
+even when the generating model matches the assumed inference model
 (as is the case in the twin pipeline),
@@
 When comparing the true and twin error distribution,
-any differences will be due to the fact that true and twin phylogenies are realizations 
-of different processes: one (possibly) non-standard, $p_G$, 
+any differences will be due to the fact that true and twin phylogenies 
+are realizations of different processes: one (possibly) non-standard, $p_G$, 
 and one standard, $p_T$ (see Fig~\ref{fig:pipeline}).
 
-This can be seen for both the "generative" and "candidate" model 
-types (see Table~\ref{tab:options}).
-If the chosen model type is "generative", 
-the tree prior chosen for the twin inference will  
-exactly match the model to generate the tree.
-In the main pipeline, as the tree model $p_{G}$ is non-standard, 
-it cannot be used in inference.
-If, instead, the chosen model type is "candidate", 
-the twin tree model will be included in the pool of examined models 
-during the process of selection of the inference model.
-
-Finally, if the goal is to evaluate BEAST2's performance on a non-standard tree prior, 
+Finally, if the goal is to evaluate BEAST2's performance 
+on a non-standard tree model, 
 one must also consider the last source of stochasticity: 
@@
 A single phylogeny cannot be considered as fully representative of the model. 
-For this reason multiple phylogenies, as well as an equal number of twins, must be considered. 
+For this reason multiple phylogenies must be considered. 
 If the number of considered phylogenies is high enough, 
@@
 and its twin counterpart leads to a fair evaluation 
-of the new tree prior with respect to the baseline error.
+of the new tree model with respect to the baseline error.
+
+\subsection{Candidate models}\label{subsec:candidates}
+
+The user has to specify exactly one standard inference model,
+but may be insure which one to pick. To account for this, the user can
+specify a set of candidate inference models. After a true alignment
+has been simulated, the candidate model with the highest evidence will be
+used in another inference run, resulting in another error distribution.
+If twinning is used, a candidate model that has the highest evidence for
+the twin alignment, this model is also used to create another error
+distribution. Thanks to the use of candidate models, one can check
+if the generative model picked, has been a good choice.
 
@@
 
-We use the tree as shown in Figure \ref{fig:dd_tree},
+We use the DD tree as shown in Figure \ref{fig:dd_tree},
 which has an arbitrary crown age of ten time units
@@
 
-HIERO
-Note that the user cannot specify a clock model, as only the strict clock model 
-is implemented and thus always assumed.
-
-HIERO
-We note that scripted use of BEAST2 packages is only possible under Linux and Mac.
-Windows systems can do the model comparison for shorter DNA sequences 
-using the web interface of \verb;mcbette; [\cite{mcbette}].
-
 Figure~\ref{fig:example_30} illustrates the primary result of our pipeline: 
@@
 that the choice of tree prior does matter.
-
-We note, however, that all examples used only one original tree,
-where any speciation process produces a whole range of trees.
-One tree is not enough to determine the impact 
-of a tree prior on Bayesian inference.
-However, if the same procedure were repeated and 
-performed on a distribution with a sufficient number of generative trees, 
-it would constitute a quantitative and effective assessment of the 
-quality of the inference.
-Also a twin tree does not always result in a lower error distribution,
-as the stochasticity in generating a twin tree will - with
-very low probability - yield a tree of that same very low probability.
+We note, however, that only one tree from a novel tree model
+is not enough to determine the impact of using an incorrect
+tree prior. Instead, a distribution 
+of multiple trees, generated by the novel tree model, should be used.
 
@@
 tailored on the generative process, is needed. 
-If this is indeed the case, one can implement the novel tree prior as an addition to 
+If this is indeed the case, one can implement 
+the novel tree prior as an addition to 
 his/her favorite Bayesian inference tool.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{pirouette resources}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\verb;pirouette; is free, libre and open source software available at 
-\url{http://github.com/richelbilderbeek/pirouette},
-licensed under the GNU General Public License version 3.
-\verb;pirouette; depends on multiple packages, which are:
-\verb;ape; [\cite{ape}],
-\verb;assertive; [\cite{assertive}],
-\verb;babette; [\cite{bilderbeek2018babette}],
-\verb;DDD; [\cite{DDD}],
-\verb;devtools; [\cite{devtools}],
-\verb;dplyr; [\cite{dplyr}],
-\verb;ggplot2; [\cite{ggplot2}],
-\verb;knitr; [\cite{knitr}],
-\verb;lintr; [\cite{lintr}],
-\verb;magrittr; [\cite{magrittr}],
-\verb;mcbette; [\cite{mcbette}],
-\verb;nLTT; [\cite{nLTT}],
-\verb;phangorn; [\cite{phangorn}],
-\verb;phytools; [\cite{phytools}],
-\verb;plyr; [\cite{plyr}],
-\verb;rappdirs; [\cite{rappdirs}],
-\verb;rmarkdown; [\cite{rmarkdown}],
-\verb;Rmpfr; [\cite{Rmpfr}],
-\verb;stringr; [\cite{stringr}],
-\verb;TESS; [\cite{TESS}],
-\verb;testit; [\cite{testit}], 
-\verb;testthat; [\cite{testthat}] and
-\verb;tidyr; [\cite{tidyr}].
-
-\verb;pirouette;'s development takes place on GitHub,
-\url{https://github.com/richelbilderbeek/pirouette},
-which allows submitting bug reports, requesting features, 
-and adding code. To ensure a high quality, \verb;pirouette; 
-uses a continuous integration service, has a code coverage of above 95\%
-and enforces the most commonly used R style guide [\cite{style_guide}].
-
-\verb;pirouette;'s is extensively documented on its website,
-its documentation and its vignettes.
-The \verb;pirouette; website is a good starting point to learn
-how to use \verb;pirouette;, as it links to tutorials and videos.
-The \verb;pirouette; package documentation describes
-all functions and liberally links to related functions.
-All exported functions show a minimal example as part of their documentation.
-The \verb;pirouette; vignette demonstrates extensively how 
-to use \verb;pirouette; in a more informally written way.
-
-The code used in this article and more examples that are periodically 
-tested, can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_examples}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Citation of pirouette}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-To cite \verb;pirouette; this article from within R, use:
-
-\begin{lstlisting}[language=R]
-> citation("pirouette")
-\end{lstlisting}
 

### 2020-03-25 supplement.tex
3123 words
2856 words

@@
   \item{
+    subsection \ref{subsec:guidelines}: guidelines for empiricists
+  }
+  \item{
     subsection \ref{subsec:installation}: installation
@@
   \item{
-    subsection \ref{subsec:main_functions}: \verb;pirouette;'s 
-    main functions
+    subsection \ref{subsec:resources}: resources, such as 
+    website, tutorials, packages used, bug reporting and contributing
+  }
+  \item{
+    subsection \ref{subsec:citation}: citation of pirouette
+  }
+  \item{
+    subsection \ref{subsec:main_functions}: main functions
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Guidelines for empiricists}
+\label{subsec:guidelines}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\itemize{
+  \item follow the guidelines as given by BEAST2
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Installation}
@@
 install_beast2_pkg("NS")
+\end{lstlisting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Resources}
+\label{subsec:resources}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\verb;pirouette; is free, libre and open source software available at 
+\url{http://github.com/richelbilderbeek/pirouette},
+licensed under the GNU General Public License version 3.
+\verb;pirouette; depends on multiple packages, which are:
+\verb;ape; [\cite{ape}],
+\verb;assertive; [\cite{assertive}],
+\verb;babette; [\cite{bilderbeek2018babette}],
+\verb;DDD; [\cite{DDD}],
+\verb;devtools; [\cite{devtools}],
+\verb;dplyr; [\cite{dplyr}],
+\verb;ggplot2; [\cite{ggplot2}],
+\verb;knitr; [\cite{knitr}],
+\verb;lintr; [\cite{lintr}],
+\verb;magrittr; [\cite{magrittr}],
+\verb;mcbette; [\cite{mcbette}],
+\verb;nLTT; [\cite{nLTT}],
+\verb;phangorn; [\cite{phangorn}],
+\verb;phytools; [\cite{phytools}],
+\verb;plyr; [\cite{plyr}],
+\verb;rappdirs; [\cite{rappdirs}],
+\verb;rmarkdown; [\cite{rmarkdown}],
+\verb;Rmpfr; [\cite{Rmpfr}],
+\verb;stringr; [\cite{stringr}],
+\verb;TESS; [\cite{TESS}],
+\verb;testit; [\cite{testit}], 
+\verb;testthat; [\cite{testthat}] and
+\verb;tidyr; [\cite{tidyr}].
+
+\verb;pirouette;'s development takes place on GitHub,
+\url{https://github.com/richelbilderbeek/pirouette},
+which allows submitting bug reports, requesting features, 
+and adding code. To ensure a high quality, \verb;pirouette; 
+uses a continuous integration service, has a code coverage of above 95\%
+and enforces the most commonly used R style guide [\cite{style_guide}].
+
+\verb;pirouette;'s is extensively documented on its website,
+its documentation and its vignettes.
+The \verb;pirouette; website is a good starting point to learn
+how to use \verb;pirouette;, as it links to tutorials and videos.
+The \verb;pirouette; package documentation describes
+all functions and liberally links to related functions.
+All exported functions show a minimal example as part of their documentation.
+The \verb;pirouette; vignette demonstrates extensively how 
+to use \verb;pirouette; in a more informally written way.
+
+The code used in this article and more examples that are periodically 
+tested, can be found at 
+\url{https://github.com/richelbilderbeek/pirouette_examples}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Citation of pirouette}
+\label{subsec:citation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To cite \verb;pirouette; this article from within R, use:
+
+\begin{lstlisting}[language=R]
+> citation("pirouette")
 \end{lstlisting}
@@
 uses one tree.
+
 The code used in this part of the article can be found at 

### 2020-03-25 content.tex
Spellcheck

@@
 There have been multiple attempts to investigate the importance of tree
-prior choice. For example, recently Sarver et al., [\cite{sarver2019choice}] 
+prior choice. For example, recently Sarver and colleagues, [\cite{sarver2019choice}] 
 showed that the choice of tree prior does not 
 substantially affect phylogenetic inferences of diversification rates.
-Also recently, Duchene et al. [\cite{duchene2018phylodynamic}] released
+Also recently, Duchêne and colleagues [\cite{duchene2018phylodynamic}] released
 a BEAST2 package to assess how well posterior predictive simulations 
@@
 site model) are JC, HKY, TN and GTR (see Table \ref{tab:options} for
-the meaning of these abbreviatons).
-JC is the default nucleotide subsitution mode (NSM),
-in which the nucleotide substition rates between all
+the meaning of these abbreviations).
+JC is the default nucleotide substitution mode (NSM),
+in which the nucleotide substitution rates between all
 nucleotides are equal and constant over time.

### 2020-03-25 scripts/spell_check.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+#
+# Usage:
+#
+#   ./scripts/spell_check.sh
+#
+
+Rscript -e 'spelling::spell_check_files("content.tex", ignore = readLines("inst/WORDLIST"))'
+

### 2020-03-25 content.tex
Implementing feedback reviewer: references should have correct capitalization

@@
 i.e. the prior probability of a species tree.
-Many diversification models (and their associated probability algorithms) have been developed, e.g., models in which diversification is time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}],
+Many diversification models (and their associated probability algorithms) 
+have been developed, e.g., models in which diversification is time-dependent [\cite{nee1994reconstructed,rabosky2008explosive}],
 or diversity-dependent [\cite{etienne2011diversity}],
 or where diversification rates change for specific lineages 
-and their descendants [\cite{etienne2012conceptual}, 
-\cite{rabosky2014automatic}, \cite{alfaro2009nine}, 
-\cite{laudanno2020sls}], models that treat speciation as a process that takes 
-time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
-or as a burst of simultaneous branching
-events [\cite{laudanno2018mbd}], or where diversification rate
-depends on a trait that has two [\cite{maddison2007estimating}], 
-or more [\cite{fitzjohn2012diversitree}] states,
-even concealed states [\cite{beaulieu2016detecting}] 
-or a combination of all these [\cite{herrera2018detecting}].
-
-Many diversification models (and their associated probability algorithms) have been developed to account for possible mechanisms.
-There are models in which diversification rates are time-dependent [\cite{nee1994reconstructed}, \cite{rabosky2008explosive}], 
-or diversity-dependent [\cite{etienne2011diversity}],
-or not homogeneous across the entire phylogeny [\cite{etienne2012conceptual}, 
-\cite{rabosky2014automatic}, \cite{alfaro2009nine}, 
-\cite{laudanno2020sls}], models that treat speciation as a process that takes time [\cite{rosindell2010protracted}][\cite{etienne2012prolonging}][\cite{lambert2015reconstructed}], 
-or as a burst of simultaneous branching
-events [\cite{laudanno2018mbd}].
-Others account for trait-dependent regimes, in several ways: where diversification rates
-depend on a trait that has two [\cite{maddison2007estimating}], 
-or more [\cite{fitzjohn2012diversitree}] states,
-even concealed states [\cite{beaulieu2016detecting}] 
-or a combination of all these [\cite{herrera2018detecting}].
-Despite such a rich theoretical landscape, only a few of these diversification models are available as a BEAST2 package.
+and their descendants [\cite{etienne2012conceptual, rabosky2014automatic, alfaro2009nine}].
+Other models treat speciation as a process that takes 
+time [\cite{rosindell2010protracted, etienne2012prolonging, lambert2015reconstructed}],
+or where diversification rate
+depends on one or more traits [\cite{maddison2007estimating, fitzjohn2012diversitree}].
+Despite such a rich theoretical landscape, only a few of these 
+diversification models are available as a BEAST2 package.
 
@@
 Here we introduce a method to quantify the importance of a novel tree prior.
-The method starts with a phylogeny generated 
-by the new model. 
+The method starts with a phylogeny generated by the new model. 
 Next, nucleotide sequences are simulated that follow the same
@@
 simulated under a (usually novel) diversification model.
-We refer to the true model that generated the given tree
+We refer to the model that has generated the given tree
 as the 'generative tree model' $\mathit{p_{G}}$.
@@
     Alignment model &
-    It includes the parts of the generative model that directly affect the 
-    alignment generation, like the clock model $\mathit{c_{G}}$ and 
-    the site model $\mathit{s_{G}}$. Additional arguments can be provided, 
-    such as the mutation rate and the root sequence. \\
+    Specifies the alignment generation, such as the clock model 
+    $\mathit{c_{G}}$, site model $\mathit{s_{G}}$ and root sequence. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     Inference model &
-    Phylogenetic inference model to run BEAST2. Likewise the generative model $G$, 
-    its main components are the tree prior $\mathit{p_{I}}$, 
-    the clock model $\mathit{c_{I}}$ and the site model $\mathit{s_{I}}$. \\
+    The assumed phylogenetic inference model, of which the main components
+    are the tree prior $\mathit{p_{I}}$, assumed clock model $\mathit{c_{I}}$ 
+    and assumed site model $\mathit{s_{I}}$. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     \texttt{pirouette} pipeline.
-    The pipeline starts from a phylogeny (1a) simulated by the generative tree model 
+    The pipeline starts from a phylogeny (1a) simulated by the 
+    generative tree model 
     $\mathit{p_{G}}$.
-    The phylogeny is converted to an alignment (2a) using the generative alignment model 
+    The phylogeny is converted to an alignment (2a) using the generative 
+    alignment model 
     $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed of a clock and a site model. 
@@
     imported from the generative model. 
-    The twin alignment has the same number of mutations as the original alignment.
     The twin pipeline follows the procedure of the main pipeline, 
@@
 to actually use it in the inference step.
-For example, we may require that an inference
-model (a combination of a tree model, clock model and site model) 
-should include the generative/true tree model. 
-As a second example, we may require that we have selected a set of 
-candidate inference models,
-of which only the best should be used in the actual inference.
-In the first example, we specified the condition $C_{i}$ that this
-generative model should always be run, whereas in the second example,
-we specified condition $C_{i}$ that a candidate model should only be run
-when it is the best.
-The 'best' model is defined as the inference model with
-the highest evidence (a.k.a. marginal likelihood), based on the alignment 
-simulated in the previous step.
-The evidence for an inference model is estimated by nested 
-sampling [\cite{russel2019model}], using the \verb;NS; BEAST2 package.
+As a first example, if the generative model is known and standard,
+one can specify this as such.
+As a second example, if we want to run only the inference
+model that fits best to an alignment from a set of candidates,
+one can specify this as well (see section 'Candidate models').
 
@@
 
-An optional step is to use the 'twinning process'.
-This process, $T$, encompasses two steps:
+Similar to other experiments, \verb;pirouette; allows to do
+a control measurement, by use of a process we call twinning. 
+In this context, this control results in an error distribution
+that is the baseline error of the pipeline. The difference
+between the 'true' and 'twin' error distributions is caused only
+by the mismatch between the generative tree model and the assumed
+tree prior.
+
+The twinning process, $T$, encompasses two steps:
 $T_1$, that generates a 'twin tree' (Fig.~\ref{fig:pipeline}, 1b) 
@@
 
-The twin pipeline serves as a control: 
-even when the generating model matches the assumed inference model
-(as is the case in the twin pipeline),
-the inferred trees from the posterior distribution will 
-still differ from the true tree, 
-due to stochasticity in producing an alignment 
-and to the MCMC sampling of the posterior.
-The twin pipeline provides this minimum error,
-because the generating and inference model match exactly.
-When comparing the true and twin error distribution,
-any differences will be due to the fact that true and twin phylogenies 
-are realizations of different processes: one (possibly) non-standard, $p_G$, 
-and one standard, $p_T$ (see Fig~\ref{fig:pipeline}).
+\subsection{Candidate models}\label{subsec:candidates}
+
+The user has to specify exactly one standard inference model,
+but may be insure which one to pick. To account for this, the user can
+specify a set of candidate inference models. After a true alignment
+has been simulated, the candidate model with the highest 
+evidence (a.k.a. marginal likelihood) will be
+used in another inference run, resulting in another error distribution.
+The evidence for an inference model is estimated by nested 
+sampling [\cite{russel2019model}], using the \verb;NS; BEAST2 package.
+
+If twinning is used, a candidate model that has the highest evidence for
+the twin alignment, this model is also used to create another error
+distribution. Thanks to the use of candidate models, one can check
+if the assumed generative model, has been a good choice.
+
+\subsection{Stochasticity caused by simulating phylogenies}
 
@@
 of the new tree model with respect to the baseline error.
-
-\subsection{Candidate models}\label{subsec:candidates}
-
-The user has to specify exactly one standard inference model,
-but may be insure which one to pick. To account for this, the user can
-specify a set of candidate inference models. After a true alignment
-has been simulated, the candidate model with the highest evidence will be
-used in another inference run, resulting in another error distribution.
-If twinning is used, a candidate model that has the highest evidence for
-the twin alignment, this model is also used to create another error
-distribution. Thanks to the use of candidate models, one can check
-if the generative model picked, has been a good choice.
 
@@
     The inference error made 
-    for both a generative tree prior and best candidate model
+    for an assumed generative tree model and best candidate model,
     compared with the error obtained for the twin tree.
-    Here, the 'twin' tree shows the baseline inference error.
+    The assumed generative tree model is the model assumed to be closest to
+    the actual generative tree model that generated the true tree.
+    The twin distributions show the baseline inference error.
     Vertical dashed lines show the median error value per distribution.
@@
 Focussing on the upper panel of Figure \ref{fig:example_30},
-we can see that the error distributions of the generative model
+we can see that the error distributions of the assumed generative model
 differ strongly between the true and twin tree. 
@@
 
+The candidate model that had highest evidence given the simulated alignment,
+was JC, RLN, BD (see Table \ref{tab:options} for the meaning of these 
+abbreviations). JC was indeed the site model used when simulating the
+alignments. The RLN clock model is an unexpected choice: the RLN clock
+model assumes that mutation rates differ between species, where the
+alignment is simulated with equal mutation rates for all species.
+The BD model is indeed closer to the DD model than Yule, as both BD and DD
+model allow for extinction, where Yule does not.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We showed how to use \verb;pirouette; to quantify the importance of a 
-tree prior in Bayesian phylogenetics, using the simplest generative tree 
-model possible.
-In principle any other (more complex) generative tree model can be tested, 
-but we chose to provide the simplest (and fastest to run) examples.
+tree prior in Bayesian phylogenetics, assuming the simplest standard 
+tree model possible.
+In principle any other standard tree model can be assume, 
+but we chose to provide the simplest example.
 
@@
 of multiple trees, generated by the novel tree model, should be used.
+Also, we think that judging a new tree model by its recovery when using
+a standard tree prior is an oversimplification: a new tree model may
+still tell us something about an evolutionary aspect such as 
+a trait or geography. \verb;pirouette;'s results encourage
+to convert this novel tree model into a standard tree prior, 
+so that the inference is (as predicted to be) improved strongly.
+
+Interpreting the results of \verb;pirouette; remains a subject 
+of study. \verb;pirouette; can quantify these, which is a good
+first step, but does not give a unequivocal answer to the question 
+if the inference error is profound. Using the nLTT statistic is
+a promising starting, as it can compare any two trees and 
+results in an error distribution of known range. If a researcher
+would be more interested in other aspects such as tree balance,
+\verb;pirouette; allows for this.
 
 In conclusion, \verb;pirouette; can show the errors to be expected
-when the tree prior used in inference is different from the generating model.
+when the tree model assumed in inference is different from the 
+actual generative tree model.
 The user can then judge whether or not a new tree prior, 
-tailored on the generative process, is needed. 
-If this is indeed the case, one can implement 
-the novel tree prior as an addition to 
+tailored to the novel tree model, is needed. 
+If this is indeed the case, one has a reason
+to implement the novel tree prior as an addition to 
 his/her favorite Bayesian inference tool.

### 2020-03-25 content.tex
Second round of processing feedback from reviewers

@@
 The second step (Fig.~\ref{fig:pipeline}, 3a)
-selects one (or more) inference model(s) $I$ from a set of inference models $I_{1},\dots,I_{n}$. 
-We define an experiment $X_{i}$ as the combination of 
-an inference model $I_{i}$ and the conditions $C_{i}$ 
-to actually use it in the inference step.
-As a first example, if the generative model is known and standard,
-one can specify this as such.
-As a second example, if we want to run only the inference
+selects one or more inference model(s) $I$ from 
+a set of standard inference models $I_{1},\dots,I_{n}$.
+For example, if the generative model is known and standard,
+one can specify an inference model as such.
+If the tree model is unknown or non-standard, one can pick
+an (assumed) closest standard inference model.
+Also, if we want to run only the inference
 model that fits best to an alignment from a set of candidates,
-one can specify this as well (see section 'Candidate models').
+one can specify these inference models as 
+well (see section 'Candidate models').
 
@@
 
+Like most phylogenetic experiments, the setup of \verb;pirouette;
+is determined by many arbitrary choices. A prime example is the
+length of simulated DNA sequence. We predict that the inference error
+decreases for longer DNA sequences, but we only investigated this
+superficially (see the supplementary materials). The strength
+of \verb;pirouette; is, however, to supply a pipeline that is uniform
+regardless these choices.
+
 Interpreting the results of \verb;pirouette; remains a subject 

### 2020-03-26 content.tex
All feedback is in

@@
     fast computation of the prior probability are used.
-    An open question is, how accurate the tree estimation 
+    Question is, how accurate the tree estimation 
     is when the real macroevolutionary processes are substantially 
@@
 substantially affect phylogenetic inferences of diversification rates.
-Also recently, Duchêne and colleagues [\cite{duchene2018phylodynamic}] released
-a BEAST2 package to assess how well posterior predictive simulations 
-recover a given tree when using the standard diversification models.
-These studies show how current diversification models compare to one another, 
+This study shows how current diversification models compare to one another, 
 but they do not help to assess the importance of a new tree prior.
@@
 evolutionary history as the given phylogeny. 
-
 Then, using BEAST2's built-in tree priors,
@@
 Only new diversification models that result 
-in a large discrepancy between inferred and simulated phylogenies will be worth the effort and computational burden to implement a species tree prior for in a Bayesian framework.
+in a large discrepancy between inferred and simulated phylogenies 
+will be worth the effort and computational burden to implement 
+a species tree prior for in a Bayesian framework.
 
@@
 
+One interesting avenue to find out some guidelines is to compare
+the results of \verb;pirouette; with those from a recent study 
+by Duchêne and colleagues [\cite{duchene2018phylodynamic}],
+where they also created a method to assess the adequacy of a tree model
+on empirical trees. A feature of their setup, is that they bypass
+the simulation of alignments, which definitely has its advantages,
+for example, in terms of run-time speed and the need for less assumptions.
+On the other hand, \verb;pirouette; uniquely allows to estimate the
+adequacy of a tree prior in combination with novel clock and site models.
+
 In conclusion, \verb;pirouette; can show the errors to be expected

### 2020-03-26 supplement.tex
Second round of processing feedback from reviewers
Changes @Giappo
Supplementary material complete

@@
 horizontally) is done manually.
+
+We limited the amount of replicates by time, aiming at a duration of
+24 hours per setting, when run on the computer cluster of the
+University of Groningen. Due to this, for example, a run of 40 taxa only
+has 6 replicates, as one run takes 4 hours.
 
@@
 
-\richel{Issue \url{https://github.com/richelbilderbeek/pirouette_article/issues/89}}
-
 This subsection extends the main example, by using multiple (instead of
 one) trees. These trees are produced by running a DD tree simulation
-with the same parameters as the main example, which are
-speciation rate \richel{?}, extinction rate $0.1$, and carrying
-capacity 40.
+with the same parameters as the main example.
 
@@
 
-\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/63}}
-
-The main example uses \richel{?} taxa.
-One expects that the more taxa are present in a phylogeny,
-the better the inference will be, 
-thus the lower the errors will be.
-Here, we show the same results as the main example,
+The main example uses 6 taxa. Here, we show the same results as the main example,
 except for a varying number of taxa.
 
-Note that, \richel{[some authors]} found that an increase in the number
-of taxa may counterintuitively decrease the inference.
-
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_20}.
+\url{https://github.com/richelbilderbeek/pirouette_example_28} (6 taxa, main example),
+\url{https://github.com/richelbilderbeek/pirouette_example_32} (12 taxa),
+\url{https://github.com/richelbilderbeek/pirouette_example_33} (24 taxa),
+\url{https://github.com/richelbilderbeek/pirouette_example_41} (32 taxa),
+\url{https://github.com/richelbilderbeek/pirouette_example_42} (40 taxa).
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/errors_5.png}
-  \caption{5 taxa}
+  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+  \caption{6 taxa, 16 hours, main example}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/errors_10.png}
-  \caption{10 taxa}
+  \includegraphics[width=\textwidth]{pirouette_example_32/errors.png}
+  \caption{12 taxa, 19 hours}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/errors_20.png}
-  \caption{20 taxa}
+  \includegraphics[width=\textwidth]{pirouette_example_33/errors.png}
+  \caption{24 taxa, 28 hours}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_20/errors_30.png}
-  \caption{30 taxa}
+  \includegraphics[width=\textwidth]{pirouette_example_41/errors.png}
+  \caption{32 taxa, 7 hours}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_42/errors.png}
+  \caption{40 taxa, 8 hours}
 \end{figure}
@@
 The main example uses a DNA alignment length of 1000 nucleotides.
-One expects that the more nucleotides are present in the alignment,
-the more information is available,
-thus the better the inference will be, 
-thus the lower the errors will be.
 Here, we show the same results as the main example,
@@
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_21}.
+\url{https://github.com/richelbilderbeek/pirouette_example_19} (500 nucleotides),
+\url{https://github.com/richelbilderbeek/pirouette_example_28} (1000 nucleotides, main example),
+and \url{https://github.com/richelbilderbeek/pirouette_example_34} (2000 nucleotides).
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/errors_100.png}
-  \caption{100 nucleotides}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/errors_248.png}
-  \caption{248 nucleotides}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/errors_500.png}
+  \includegraphics[width=\textwidth]{pirouette_example_19/errors.png}
   \caption{500 nucleotides}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/errors_1000.png}
-  \caption{1000 nucleotides}
+  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+  \caption{1000 nucleotides (main example)}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/errors_2000.png}
+  \includegraphics[width=\textwidth]{pirouette_example_34/errors.png}
   \caption{2000 nucleotides}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/errors_4000.png}
-  \caption{4000 nucleotides}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/errors_8000.png}
-  \caption{8000 nucleotides}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_21/errors_16000.png}
-  \caption{16000 nucleotides}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/87}}
 
@@
   \includegraphics[width=\textwidth]{pirouette_example_27/errors.png}
-  \caption{An unliked node substitution model}
+  \caption{An unlinked node substitution model, 20 replicates}
 \end{figure}
@@
 
-\richel{Issue at \url{https://github.com/richelbilderbeek/pirouette_article/issues/68}}
-
 Here we show the results of a \verb;pirouette; run on a dataset 
@@
 
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_23/likelihoods.png}
+  \caption{Distribution of likelihoods, as well as which ones are picked}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_18}.
+\url{https://github.com/richelbilderbeek/pirouette_example_18} and
+\url{https://github.com/richelbilderbeek/pirouette_example_28}.
 
@@
   \caption{Equal mutation rate}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+  \caption{Equal mutation rate and number of mutations (main example)}
 \end{figure}
@@
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_24}.
+\url{https://github.com/richelbilderbeek/pirouette_example_35} (0.25 / crown age),
+\url{https://github.com/richelbilderbeek/pirouette_example_36} (0.50 / crown age),
+\url{https://github.com/richelbilderbeek/pirouette_example_37} (0.75 / crown age),
+\url{https://github.com/richelbilderbeek/pirouette_example_28} (1.00 / crown age, main example),
+\url{https://github.com/richelbilderbeek/pirouette_example_38} (1.25 / crown age),
+\url{https://github.com/richelbilderbeek/pirouette_example_39} (1.50 / crown age),
+\url{https://github.com/richelbilderbeek/pirouette_example_40} (2.00 / crown age),
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/errors_1.png}
-  \caption{Per-nucleotide mutation rate of 0.0125}
+  \includegraphics[width=\textwidth]{pirouette_example_35/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.25 / crown age}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/errors_2.png}
-  \caption{Per-nucleotide mutation rate of 0.025}
+  \includegraphics[width=\textwidth]{pirouette_example_36/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.50 / crown age}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/errors_3.png}
-  \caption{Per-nucleotide mutation rate of 0.05}
+  \includegraphics[width=\textwidth]{pirouette_example_37/errors.png}
+  \caption{Per-nucleotide mutation rate of 0.75 / crown age}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/errors_4.png}
-  \caption{Per-nucleotide mutation rate of 0.1}
+  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+  \caption{Per-nucleotide mutation rate of 1.0 / crown age (main example)}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/errors_5.png}
-  \caption{Per-nucleotide mutation rate of 0.2}
+  \includegraphics[width=\textwidth]{pirouette_example_38/errors.png}
+  \caption{Per-nucleotide mutation rate of 1.25 / crown age}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/errors_6.png}
-  \caption{Per-nucleotide mutation rate of 0.4}
+  \includegraphics[width=\textwidth]{pirouette_example_39/errors.png}
+  \caption{Per-nucleotide mutation rate of 1.50 / crown age}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_24/errors_7.png}
-  \caption{Per-nucleotide mutation rate of 0.8}
+  \includegraphics[width=\textwidth]{pirouette_example_40/errors.png}
+  \caption{Per-nucleotide mutation rate of 2.0 / crown age}
 \end{figure}

### 2020-03-26 scripts/copy_figs_from_examples_to_here.R

@@ -17,7 +17,7 @@ files <- list.files(path = from_root, pattern = "*.(png|latex)", recursive = TRU
 files
 
 # Only pirouette examples
-files <- stringr::str_match(string = files, pattern = "pirouette_example_.*(true|twin|errors|esses).*")
+files <- stringr::str_match(string = files, pattern = "pirouette_example_.*(likilihoods|true|twin|errors|esses).*")
 files
 
 # Remove NA's

### 2020-03-26 supplement.tex
Create guidelines

@@
 
-These supplementary materials show the full pipeline and
-added diagnostics for the examples in the main article.
+These supplementary materials show facets of \verb;pirouette; that did not
+fit the main article, such as the installation and an overview of
+pirouette's main functions.
+
+Additionally, there is a guide for empicists, based on multiple experiments
+that are shown here as well.
+
+For these experiments, we limited the amount of replicates by time, 
+aiming at a duration of
+24 hours per setting, when run on the computer cluster of the
+University of Groningen. Due to this, for example, a run of 40 taxa only
+has 6 replicates, as one run takes 4 hours.
 
@@
 without any aesthetical modifications, with the exception that the arrangement
-of the sub-figures (for example, aligning true and twin tree
-horizontally) is done manually.
-
-We limited the amount of replicates by time, aiming at a duration of
-24 hours per setting, when run on the computer cluster of the
-University of Groningen. Due to this, for example, a run of 40 taxa only
-has 6 replicates, as one run takes 4 hours.
+of the sub-figures in subsection \ref{subsec:main_example},
+where we aligned parts of the figure by hand.
 
@@
 
+From the experiments shown below, we composed some rough guidelines.
+These guidelines should be treated as preliminary results, as
+the total runtime of these experiments is 'only' 19 days.
+
 \itemize{
-  \item follow the guidelines as given by BEAST2
+  \item the use of 20 replicates results in decent plots
+  \item the use of more taxa, counterintuitively, increases the inference error
+  \item the use of longer DNA sequences, as expected, decreases the inference error
+  \item {
+    there is a difference when a twin alignment is selected to have
+    an equal amount of mutations as the true alignment. The consequences
+    of this is unknown
+  }
+  \item using a mutation rate less than 1.0 / crown age, decreases the
+    inference error. We predict this will increase the error in the parameter
+    estimation
+  } 
 }

### 2020-03-26 supplement.tex
Create guidelines
Add hours

@@
   }
-  \item{
-    subsection \ref{subsec:better_label_needed} shows 
-    \richel{better description needed}
-  }
 }
@@
   }
-  \item using a mutation rate less than 1.0 / crown age, decreases the
+  \item {
+    using a mutation rate less than 1.0 / crown age, decreases the
     inference error. We predict this will increase the error in the parameter
@@
 ]
-remotes::install_github("richelbilderbeek/pirouette")
+remotes::install_github(
+  "richelbilderbeek/pirouette"
+)
 \end{lstlisting}
@@
     \hline
-    \textbf{Name} & \textbf{Description} & \textbf{Listing} \\
+    \textbf{Name} & \textbf{Description} \\
     \hline
-    \verb;pir_run; & Run \verb;pirouette; & \ref{lst:pir_run} \\
-    \verb;pir_plot; & Show the \verb;pirouette; results as a plot & \ref{lst:pir_plot} \\
-    \verb;create_pir_params; & Create the \verb;pirouette; parameters & \ref{lst:create_pir_params} \\
+    \verb;pir_run; & Run \verb;pirouette; \\
+    \verb;pir_plot; & Show the \verb;pirouette; results as a plot  \\
+    \verb;create_pir_params; & Create the \verb;pirouette; parameters  \\
     \hline
-    \verb;create_alignment_params; & Create the alignment parameters & \ref{lst:create_alignment_params} \\
-    \verb;create_twinning_params; & Create the twinning parameters & \ref{lst:create_twinning_params} \\
-    \verb;create_experiment; & Create one experiment & \ref{lst:create_gen_experiment_explicit} \\
-    \verb;create_error_measure_params; & Create the error measurement parameters & \ref{lst:create_error_measure_params} \\
+    \verb;create_alignment_params; & Create the alignment parameters  \\
+    \verb;create_twinning_params; & Create the twinning parameters  \\
+    \verb;create_experiment; & Create one experiment  \\
+    \verb;create_error_measure_params; & Create the error measurement parameters  \\
     \hline
@@
   \caption{
-    \texttt{pirouette}'s main functions, description and the number of the 
-    listing in which it is used. 
+    \texttt{pirouette}'s main functions and description. 
   }
@@
   \includegraphics[width=\textwidth]{pirouette_example_19/errors.png}
-  \caption{500 nucleotides}
+  \caption{500 nucleotides, 12 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{1000 nucleotides (main example)}
+  \caption{1000 nucleotides, 17 hours (main example)}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_34/errors.png}
-  \caption{2000 nucleotides}
+  \caption{2000 nucleotides, 11 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_27/errors.png}
-  \caption{An unlinked node substitution model, 20 replicates}
+  \caption{An unlinked node substitution model, 20 replicates, 8 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_22/errors.png}
-  \caption{Assuming a Yule tree prior on a Yule tree}
+  \caption{Assuming a Yule tree prior on a Yule tree, 9 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_26/errors.png}
-  \caption{Assuming a Yule tree prior on a BD tree}
+  \caption{Assuming a Yule tree prior on a BD tree, 9 hours}
 \end{figure}
@@
 The code used in this part of the article can be found at 
-\url{https://github.com/richelbilderbeek/pirouette_example_23}.
+\url{https://github.com/richelbilderbeek/pirouette_example_23}
+and took 120 hours in total.
 
@@
   \includegraphics[width=\textwidth]{pirouette_example_18/errors.png}
-  \caption{Equal mutation rate}
+  \caption{Equal mutation rate, 10 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Equal mutation rate and number of mutations (main example)}
+  \caption{Equal mutation rate and number of mutations, 19 hours (main example)}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_35/errors.png}
-  \caption{Per-nucleotide mutation rate of 0.25 / crown age}
+  \caption{Per-nucleotide mutation rate of 0.25 / crown age, 11 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_36/errors.png}
-  \caption{Per-nucleotide mutation rate of 0.50 / crown age}
+  \caption{Per-nucleotide mutation rate of 0.50 / crown age, 14 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_37/errors.png}
-  \caption{Per-nucleotide mutation rate of 0.75 / crown age}
+  \caption{Per-nucleotide mutation rate of 0.75 / crown age, 16 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Per-nucleotide mutation rate of 1.0 / crown age (main example)}
+  \caption{Per-nucleotide mutation rate of 1.0 / crown age, 19 hours (main example)}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_38/errors.png}
-  \caption{Per-nucleotide mutation rate of 1.25 / crown age}
+  \caption{Per-nucleotide mutation rate of 1.25 / crown age, 17 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_39/errors.png}
-  \caption{Per-nucleotide mutation rate of 1.50 / crown age}
+  \caption{Per-nucleotide mutation rate of 1.50 / crown age, 10 hours}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_40/errors.png}
-  \caption{Per-nucleotide mutation rate of 2.0 / crown age}
+  \caption{Per-nucleotide mutation rate of 2.0 / crown age, 19 hours}
 \end{figure}

### 2020-03-26 supplement.tex
Finish response letter

@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{1000 nucleotides, 17 hours (main example)}
+  \caption{1000 nucleotides, 16 hours (main example)}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Equal mutation rate and number of mutations, 19 hours (main example)}
+  \caption{Equal mutation rate and number of mutations, 16 hours (main example)}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Per-nucleotide mutation rate of 1.0 / crown age, 19 hours (main example)}
+  \caption{Per-nucleotide mutation rate of 1.0 / crown age, 16 hours (main example)}
 \end{figure}

### 2020-03-26 content.tex

@@
     processes underlying the species tree.
-    In practice, however, only macroevolutionary models that allow for 
-    fast computation of the prior probability are used.
-    Question is, how accurate the tree estimation 
-    is when the real macroevolutionary processes are substantially 
-    different from those assumed in the tree prior. \\
+    In practice, however, only macroevolutionary models that allow for fast computation of the prior probability are used.
+    \iffalse
+    This will inevitably introduce a bias on the inference, due to the choice of the tree prior. A way to estimate this error is to choose an inference tree prior
+    \fi
+    Question is, how accurate the tree estimation is when the real macroevolutionary processes are substantially different from those assumed in the tree prior. \\
     \textbf{2. }
@@
     a free, libre and open-source R package that assesses 
-    the inference error made by Bayesian phylogenetics for a given 
-    macroevolutionary diversification model. \verb;pirouette; makes use of 
+    the inference error made by Bayesian phylogenetics for a given macroevolutionary diversification model. \verb;pirouette; makes use of 
     BEAST2, but its philosophy applies to any Bayesian phylogenetic inference 

### 2020-03-26 supplement.tex
Updates from Overleaf
Remove the likelihood shape
Update figures

@@
 has 6 replicates, as one run takes 4 hours.
+
+For all experiments, the intermediate results can all be downloaded 
+from their respective websites.
 
@@
 
-The code used in this part of the article can be found at 
+The code to reproduce this can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_30}.
@@
 
-The code used in this part of the article can be found at 
+The code to reproduce this figure can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_28}.
@@
 The main example uses 6 taxa. Here, we show the same results as the main example,
-except for a varying number of taxa.
+except for a varying number of taxa. We did so, by setting the DD model's
+carrying capacity to the desired number of taxa.
 
-The code used in this part of the article can be found at 
+The code to reproduce these figure can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_28} (6 taxa, main example),
@@
 
-The code used in this part of the article can be found at 
+The code to reproduce these figures can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_19} (500 nucleotides),
@@
 
-The code used in this part of the article can be found at 
+The code to reproduce this figure can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_27}.
@@
 
-The code used in this part of the article can be found at 
+The code to reproduce this figure can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_22}.
@@
 
-The code used in this part of the article can be found at 
+The code to reproduce this figure can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_26}.
@@
 Here we show the results of a \verb;pirouette; run on a dataset 
-of 2 trees simulated under a diversity-dependent process. 
-All the trees are simulated using the DDD package (\cite{DDD}). 
-Our goal is to show how trees simulated by a process that is 
-sensibly different from a standard tree prior will result in a 
-significant divergence between true and twin distribution. 
-To do so we set the simulation parameters (in this case: 
-absolute speciation rate $\lambda_0 = 1.6$, 
-extinction rate $\mu = 0.2$ and carrying capacity $K = 20$) 
-in order to express clear characteristics of diversity dependence in the 
-phylogenies. In fact, for a high speciation rate and a low carrying capacity, 
-we expect trees to exhibit high unbalance, compared with trees generated 
-by a standard birth-death model. 
-We can measure how unbalanced is a tree using the $\gamma$ 
-statistics (\cite{pybus2000testing}). For this reason, we choose to 
-measure the error made by the inference process using the $\gamma$ error statistics, 
-which is currently implemented in \verb;pirouette;.
+of multiple DD trees that we selected for having a low, median
+and high likelihood. In that way, we effectively selected for trees
+that are rare, uncommon and common respectively.
 
-The code used in this part of the article can be found at 
+The code to reproduce these figure can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_23}
@@
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/likelihoods.png}
-  \caption{Distribution of likelihoods, as well as which ones are picked}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The code used in this part of the article can be found at 
+The code to reproduce this figure can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_18} and
@@
 ancestral sequence at the crown to the alignments at the tips.
+This value equals `1.0 / crown age`.
 In this way, the alignment is expected to contain the maximum
@@
 
-Here, we show the same results, with the only difference that
-the mutations rates are changed.
-For a lower mutation rate, less mutations will be observed, thus
-genetic information will be less, thus the inference is expected
-to go worse, thus the errors are expected to increase.
-For a lower mutation rate, more double-mutations will occur,
-scrambling the alignments, resulting in less genetic information, 
-thus the inference is expected to go worse, 
-thus the errors are expected to increase.
+Here, we show the same results for different mutation rates.
 
-The code used in this part of the article can be found at 
+The code to reproduce this figure can be found at  
 \url{https://github.com/richelbilderbeek/pirouette_example_35} (0.25 / crown age),

### 2020-03-26 supplement.tex
Ready for feeback by @Giappo
Fix itemization in supmat

@@
 For all experiments, the intermediate results can all be downloaded 
-from their respective websites.
+from their respective websites, which is approximately 5 gigabyte in total.
 
@@
 \itemize{
-  \item the use of 20 replicates results in decent plots
-  \item the use of more taxa, counterintuitively, increases the inference error
-  \item the use of longer DNA sequences, as expected, decreases the inference error
-  \item {
+  \item{
+    the use of 20 replicates results in decent plots
+  }
+  \item{
+    the use of more taxa, counterintuitively, increases the inference error
+  }
+  \item{
+    the use of longer DNA sequences, as expected, decreases the inference error
+  }
+  \item{
     there is a difference when a twin alignment is selected to have
@@
   }
-  \item {
+  \item{
     using a mutation rate less than 1.0 / crown age, decreases the

### 2020-03-28 supplement.tex
Itemize correctly

@@
 
-\itemize{
+\begin{itemize}
   \item{
@@
   }
-}
+\end{itemize}
 
@@
 
-\itemize{
+\begin{itemize}
+  \item the use of 20 replicates results in decent plots
   \item{

### 2020-03-30 content.tex
Let's hope things keep working...
Remove 'iffalse' comment, 2991 words
Feedback @rsetienne and changes @Giappo

--- 2020-03-25 content.tex
@@
   \textbf{1. }
-    Phylogenetic trees are current routinely reconstructed from an alignment 
+    Phylogenetic trees are currently routinely reconstructed from an alignment 
     of character sequences (usually nucleotide sequences). 
@@
     processes underlying the species tree.
-    In practice, however, only macroevolutionary models that allow for 
-    fast computation of the prior probability are used.
-    An open question is, how accurate the tree estimation 
-    is when the real macroevolutionary processes are substantially 
-    different from those assumed in the tree prior. \\
+    In practice, however, only macroevolutionary models that allow for fast computation of the prior probability are used.
+    The question is how accurate the tree estimation is when the real macroevolutionary processes are substantially different from those assumed in the tree prior. \\
     \textbf{2. }
@@
 (constant-rate) birth-death [\cite{nee1994reconstructed}] models.
-These simple tree priors are the most commonly used, as they have sufficient biological complexity, while being computationally fast.
+These simple tree priors are the most commonly used, as they represent some biologically realistic processes (i.e. branching at per species constant rates), while being computationally fast.
 BEAST's successor, BEAST2 [\cite{bouckaert2019beast}],
@@
 have been developed, e.g., models in which diversification is time-dependent [\cite{nee1994reconstructed,rabosky2008explosive}],
-or diversity-dependent [\cite{etienne2011diversity}],
+or diversity-dependent [\cite{etienne2012diversity}],
 or where diversification rates change for specific lineages 
@@
 Despite such a rich theoretical landscape, only a few of these 
-diversification models are available as a BEAST2 package.
+diversification models are available as tree priors in BEAST2.
 
@@
 where ideally the estimated parameter values closely match the known/true values.
-
-Even when a diversification model passes the procedure described above, 
-it is not necessarily used in Bayesian inference.
+Even when a diversification model passes this procedure, 
+it is not necessarily used as tree prior in Bayesian inference.
 Bayesian phylogenetic inference often requires 
@@
 There have been multiple attempts to investigate the importance of tree
-prior choice. For example, recently Sarver and colleagues, [\cite{sarver2019choice}] 
+prior choice. For example, Sarver and colleagues, [\cite{sarver2019choice}] 
 showed that the choice of tree prior does not 
 substantially affect phylogenetic inferences of diversification rates.
-Also recently, Duchêne and colleagues [\cite{duchene2018phylodynamic}] released
-a BEAST2 package to assess how well posterior predictive simulations 
-recover a given tree when using the standard diversification models.
-These studies show how current diversification models compare to one another, 
-but they do not help to assess the importance of a new tree prior.
+However, they only compared current diversification models to one another, 
+and thus this does not inform us on the importance of a new tree prior.
 
@@
 The method starts with a phylogeny generated by the new model. 
-Next, nucleotide sequences are simulated that follow the same
-evolutionary history as the given phylogeny.
-
+Next, nucleotide sequences are simulated that follow the evolutionary history of the given phylogeny. 
 Then, using BEAST2's built-in tree priors,
@@
 have a parameter setting for which they reduce to this model. 
-One such example is the diversity-dependent (DD) tree 
-model [\cite{DDD, etienne2011diversity}]:
-the DD model is a BD model with a speciation rate that is dependent 
-on the number of species and a carrying capacity. For a carrying capacity
-of infinity, the DD model reduces to a BD model.
+One such example is the diversity-dependent (DD) diversification 
+model [\cite{etienne2020package, etienne2012diversity}]:
+the DD model is a BD model with a speciation or extinction rate that depends 
+on the number of species and a clade-level carrying capacity. For a carrying capacity
+of infinity, the DD model reduces to the BD model.
 When benchmarking a novel tree model, 
@@
 instead of doing an extensive analysis.
-The supplementary materials, however, do show the results
+The supplementary material, however, does show the results
 of replicated runs under multiple settings.
@@
     \verb;do_measure_evidence; &
-    Sets whether or not the evidence of the model is to be computed &
+    Sets whether or not the evidence of the model must be computed &
     TRUE, FALSE \\
@@
     is used in the inference. 
-    They are composed by the model type, run condition and 
+    They are composed of the model type, run condition and 
     whether to measure the evidence. \\
@@
     error function to measure the difference between the original phylogeny 
-    and the inferred posterior. The initial part of the posterior that is 
-    reckoned as not representative can be discarded using a burn-in fraction. \\
+    and the inferred posterior. The first iterations of the MCMC chain of the posterior may not be representative and can be discarded using a burn-in fraction. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
     alignment model 
-    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed of a clock and a site model. 
+    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed of a clock model and a site model. 
     The user defines one or more experiments.
@@
 The pipeline to assess the error BEAST2 makes in inferring this phylogeny 
-then contains the following steps:
+contains the following steps:
 \begin{enumerate}
-  \item the user supplies one or (ideally) more phylogenies from a 
-    new diversification model
-  \item from the given phylogeny an alignment is simulated 
-    under a known alignment model $\mathit{A}$;
-  \item from this alignment, according to the specified 
+  \item The user supplies one or (ideally) more phylogenies from a 
+    new diversification model.
+  \item From the given phylogeny an alignment is simulated 
+    under a known alignment model $\mathit{A}$.
+  \item From this alignment, according to the specified 
     inference conditions $\mathit{C}$, 
     an inference model $\mathit{I}$ is chosen (which may differ from the 
-    generative model);
-  \item the inference model and the alignment are used 
-    to infer a posterior distribution of phylogenies;
-  \item the phylogenies in the posterior are compared with the given phylogeny 
+    generative model).
+  \item The inference model and the alignment are used 
+    to infer a posterior distribution of phylogenies.
+  \item The phylogenies in the posterior are compared with the given phylogeny 
     to estimate the error made, according to 
-    the error measure $\mathit{E}$ specified by the user;
+    the error measure $\mathit{E}$ specified by the user.
 \end{enumerate}
@@
 phylogeny (Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 2a).
-For the sake of clarity, here we will assume the alignment consists out
-of DNA sequences, but other heritable material such as amino acids also works.
+For the sake of clarity, here we will assume the alignment consists
+of DNA sequences, but one can also use other heritable material such as amino acids.
 The user must specify a root sequence, a mutation rate and a site model.
@@
 the meaning of these abbreviations).
-JC is the default nucleotide substitution mode (NSM),
+JC is the default nucleotide substitution model (NSM),
 in which the nucleotide substitution rates between all
@@
 For example, if the generative model is known and standard,
-one can specify an inference model as such.
-If the tree model is unknown or non-standard, one can pick
-an (assumed) closest standard inference model.
+one can specify the inference model to be the same as the generative model.
+If the tree model is unknown or non-standard - which is the primary motivation for this paper -, one can pick
+a standard inference model which is considered to be closest to the true tree model.
 Also, if we want to run only the inference
@@
 Markov chain Monte Carlo (MCMC) run,
-which samples an unrepresentative part of state space. 
+which samples an unrepresentative part of paraneter and tree space. 
 By default, \verb;pirouette; 
@@
 
-Similar to other experiments, \verb;pirouette; allows to do
+Similar to lab experiments, \verb;pirouette; allows to perform
 a control measurement, by use of a process we call twinning. 
@@
 not only the same mutation rate is used, but also the total number of mutations
-matches the true alignment. The total number of mutations is defines
+matches the true alignment. The total number of mutations is defined
 as the amount of different nucleotides between the (known) root sequence
@@
 The user has to specify exactly one standard inference model,
-but may be insure which one to pick. To account for this, the user can
-specify a set of candidate inference models. After a true alignment
-has been simulated, the candidate model with the highest 
-evidence (a.k.a. marginal likelihood) will be
-used in another inference run, resulting in another error distribution.
+but may be unsure which one to pick. To account for this, the user can
+specify a set of candidate inference models. Each of these candidate inference models is run in an initial, relatively short, analysis; the candidate model with the highest 
+evidence (i.e., marginal likelihood) will then be
+used in another, longer, inference run, resulting in another error distribution.
 The evidence for an inference model is estimated by nested 
@@
 If twinning is used, a candidate model that has the highest evidence for
-the twin alignment, this model is also used to create another error
-distribution. Thanks to the use of candidate models, one can check
-if the assumed generative model, has been a good choice.
+the twin alignment is also used to create another error
+distribution.
 
@@
 We show the usage of \verb;pirouette; on an tree generated 
-by the non-standard diversity-dependent (DD) tree model \cite{DDD, etienne2011diversity},
+by the non-standard diversity-dependent (DD) tree model [\citep{etienne2020package, etienne2012diversity}],
 which is a BD model with a speciation rate that is dependent on the number of species.
@@
 which has an arbitrary crown age of ten time units
-and six taxa, to keep the figure more readable.
+and six taxa.
 
@@
 Using the default \verb;pirouette; settings,
-the error distribution shown in Figure \label{fig:example_30}
+the error distribution shown in Figure \ref{fig:example_30}
 is produced.
@@
 
-Focussing on the upper panel of Figure \ref{fig:example_30},
+In the upper panel of Figure \ref{fig:example_30},
 we can see that the error distributions of the assumed generative model
@@
 candidate model is only slightly better at inferring the true tree,
-compared to the assumed generative model, confirming that the 
-picked generative model was a good choice.
+than the assumed generative model, thereby showing that they cannot compensate for the true generative tree model not being among the inference models.
 
@@
 alignments. The RLN clock model is an unexpected choice: the RLN clock
-model assumes that mutation rates differ between species, where the
-alignment is simulated with equal mutation rates for all species.
+model assumes that mutation rates differ between species, whereas the
+alignment was simulated with equal mutation rates for all species.
 The BD model is indeed closer to the DD model than Yule, as both BD and DD
-model allow for extinction, where Yule does not.
+model allow for extinction, while Yule does not.
 
@@
 tree model possible.
-In principle any other standard tree model can be assume, 
+In principle any other standard tree model can be assumed, 
 but we chose to provide the simplest example.
@@
 it shows the error distributions for the true tree and the twin tree 
-when either the generating model or the best candidate model is used in inference. 
+when either the (assumed) generating model or the best candidate model is used in inference. 
 The clear difference between the error distributions 
@@
 tree prior. Instead, a distribution 
-of multiple trees, generated by the novel tree model, should be used.
-Also, we think that judging a new tree model by its recovery when using
-a standard tree prior is an oversimplification: a new tree model may
-still tell us something about an evolutionary aspect such as 
-a trait or geography. \verb;pirouette;'s results encourage
-to convert this novel tree model into a standard tree prior, 
-so that the inference is (as predicted to be) improved strongly.
+of multiple trees, generated by the novel tree model, should be used. In the supplementary material we have provided some examples.
 
 Like most phylogenetic experiments, the setup of \verb;pirouette;
-is determined by many arbitrary choices. A prime example is the
-length of simulated DNA sequence. We predict that the inference error
-decreases for longer DNA sequences, but we only investigated this
-superficially (see the supplementary materials). The strength
-of \verb;pirouette; is, however, to supply a pipeline that is uniform
-regardless these choices.
+involves many choices. A prime example is the
+length of the simulated DNA sequence. One expects that the inference error
+decreases for longer DNA sequences. We investigated this
+superficially and confirmed this prediction (see the supplementary materials).
+However, we note that for longer DNA sequences, the assumption of constant substitution rates may become less realistic and hence longer sequences may require more parameters.
+Hence, simply getting longer sequences will not always lead to a drastic reduction of the influence of the species tree prior.
+Fortunately, \verb;pirouette; provides a pipeline that works for all choices.
 
-Interpreting the results of \verb;pirouette; remains a subject 
-of study. \verb;pirouette; can quantify these, which is a good
-first step, but does not give a unequivocal answer to the question 
-if the inference error is profound. Using the nLTT statistic is
-a promising starting, as it can compare any two trees and 
-results in an error distribution of known range. If a researcher
-would be more interested in other aspects such as tree balance,
-\verb;pirouette; allows for this.
+Interpreting the results of \verb;pirouette; is up to the user; 
+\verb;pirouette; does not answer the question 
+whether the inference error is too large to trust the inferred tree. The user is encouraged to use different statistics to measure the error. The nLTT statistic is
+a promising starting point, as it can compare any two trees and 
+results in an error distribution of known range, but one may also explore other statistics. One such statistic is phylogenetic tree balance. In principle,
+\verb;pirouette; allows for this, but in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.
+
+As noted in the introduction, Duch\^{e}ne and colleagues [\cite{duchene2018phylodynamic}],
+also developed a method to assess the adequacy of a tree model
+on empirical trees.
+They simulate trees from the posterior distribution of the parameters and then compare this to the originally inferred tree using tree statistics, to determine whether the assumed tree model in inference indeed generates the tree as inferred.
+This is useful if these trees match, but when they do not, this does not mean that the inferred tree is incorrect; if sufficient data is available the species tree prior may not be important, and hence the inference may be adequate even though the assumed species tree prior is not.
+In short, the approach is applied to empirical trees and compares the posterior and prior distribution of trees (with the latter generated with the posterior parameters!).
+\verb;pirouette; aims to identify when assuming standard priors, not only for the species tree - which was our main focus here -, but also for the site and clock model, leads to incorrect inference if one believes more complex diversification, site or clock models are operating than can be currently accommodated in inference.
+In short, our approach applies to simulated trees and compares the posterior distributions of trees generated with a standard and non-standard model, but inferred with a standard one.
+The two methods therefore complement one another.
 
@@
 The user can then judge whether or not a new tree prior, 
-tailored to the novel tree model, is needed. 
-If this is indeed the case, one has a reason
-to implement the novel tree prior as an addition to 
-his/her favorite Bayesian inference tool.
+tailored to the novel tree model, is needed in our Bayesian phylogenetic tool.
 
@@
 
-We would like to thank the Center for Information Technology of the University 
+We thank the Center for Information Technology of the University 
 of Groningen for its support and for providing access to the Peregrine 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Authors' contributions}
+\section{Author contributions}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-03-30 supplement.tex
Feedback @rsetienne and changes @Giappo

--- 2020-03-26 supplement.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Supplementary materials}
+\section{Supplementary material}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-These supplementary materials show facets of \verb;pirouette; that did not
-fit the main article, such as the installation and an overview of
-pirouette's main functions.
-
-Additionally, there is a guide for empicists, based on multiple experiments
+This supplementary material contains additional facets of \verb;pirouette;, such as the installation of the package, an overview of
+pirouette's main functions and a guide for empiricists, based on multiple experiments
 that are shown here as well.
 
-For these experiments, we limited the amount of replicates by time, 
+For these experiments, we limited the number of replicates by time, 
 aiming at a duration of
-24 hours per setting, when run on the computer cluster of the
+24 hours per setting, when run on the Peregrine computer cluster of the
 University of Groningen. Due to this, for example, a run of 40 taxa only
-has 6 replicates, as one run takes 4 hours.
-
-For all experiments, the intermediate results can all be downloaded 
+has 6 replicates, because one run takes 4 hours. For all experiments, the intermediate results can all be downloaded 
 from their respective websites, which is approximately 5 gigabyte in total.
 
-All the figures shown in this section are shown as-is, 
-without any aesthetical modifications, with the exception that the arrangement
+All the figures shown in this section are shown without any aesthetical modifications, with the exception that the arrangement
 of the sub-figures in subsection \ref{subsec:main_example},
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Overview}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Here is an overview of the various sections:
 
@@
   \item{
-    the use of 20 replicates results in decent plots
+    The use of 20 replicates results in decent plots.
   }
   \item{
-    the use of more taxa, counterintuitively, increases the inference error
+    The use of more taxa increases the inference error
   }
   \item{
-    the use of longer DNA sequences, as expected, decreases the inference error
+    The use of longer DNA sequences decreases the inference error.
   }
   \item{
-    there is a difference when a twin alignment is selected to have
-    an equal amount of mutations as the true alignment. The consequences
-    of this is unknown
+    When we do not impose the same number of mutations between true and twin alignment, we observe a difference in the error distributions with respect to the standard case (presented in the main text) where they are forced to have the same number of mutations.
+    The consequences of this are unknown.
   }
   \item{
-    using a mutation rate less than 1.0 / crown age, decreases the
+    Using a mutation rate less than 1.0 / crown age, decreases the
     inference error. We predict this will increase the error in the parameter
-    estimation
+    estimation.
   }
@@
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install_beast2()
+beastier::install_beast2()
 \end{lstlisting}
@@
 \begin{lstlisting}[language=R, floatplacement=ht, frame=single]
-install_beast2_pkg("NS")
+mauricer::install_beast2_pkg("NS")
 \end{lstlisting}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Resources}
+\subsection{Resources}
 \label{subsec:resources}
@@
 which allows submitting bug reports, requesting features, 
-and adding code. To ensure a high quality, \verb;pirouette; 
+and adding code. To improve quality, \verb;pirouette; 
 uses a continuous integration service, has a code coverage of above 95\%
@@
 
+\newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 uses one tree.
+To assess if the results of the inference are meaningful one important parameter is the Effective Sample Size (ESS).
+This quantity describes how many independent trees are sampled from the posterior distributions.
+For reliable results it is good practice to have at least $ESS = 200$ (see \url{https://beast.community/ess_tutorial}).
+In the following we present the ESS for the posterior distributions of the 4 cases shown in Fig.
+\ref{fig:example_30_full_pipeline}: "true" pipeline with generative model (Table \ref{tab:esses_true_gen}), "true" pipeline with best candidate model (Table \ref{tab:esses_true_best}), "twin" pipeline with generative model (Table \ref{tab:esses_twin_gen}), "twin" pipeline with best candidate model (Table \ref{tab:esses_twin_best}).
+We also report the marginal likelihood (or evidence) data for model selection performed both in the "true" (Table \ref{tab:true_evidence}) and "twin" pipeline (Table \ref{tab:twin_evidence}).
 
@@
   \label{fig:example_30_full_pipeline}
-  \caption{Comparing to background noise: full pipeline}
+  \caption{Comparing to background noise: full pipeline.
+The true tree (top right) is used to simulate an alignment.
+From this alignment two posterior distributions of trees are created: one using the generative model and another one using the inference model with the highest marginal likelihood.
+From each of the two, exploiting the nltt error statistics, a distribution of errors between the posterior trees and the main trees is drawn.
+From the true tree also a twin tree is created (right side of the figure) which follows the same pipeline, leading to two additional error distributions to use as baseline errors for their equivalents.}
 \end{figure}
@@
 
+\clearpage
+%\newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Error distribution from a distribution of trees}
+  \caption{Error distribution from a distribution of trees.
+What is shown in the figure is similar to Fig.
+\ref{fig:example_30} for the main example.
+The main difference is that here the original tree (and its twin counterpart) is not only one, but instead there are 20 replicated.
+Each replicate has been simulated independently.
+For each setting (true generative, true best candidate, twin generative and twin candidate), the resulting errors from each replicated pipeline have been merged in a single dataset.}
 \end{figure}
+
+\newpage
 
@@
 
-The main example uses 6 taxa. Here, we show the same results as the main example,
+The main example uses 6 taxa. Here we show the same results as the main example,
 except for a varying number of taxa. We did so, by setting the DD model's
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The effect of differently common diversity-dependent trees}
+\subsection{The effect of diversity-dependent trees differing in how likely they are under the DD process}
 \label{subsec:better_label_needed}
@@
 of multiple DD trees that we selected for having a low, median
-and high likelihood. In that way, we effectively selected for trees
+and high likelihood. In this way, we effectively selected for trees
 that are rare, uncommon and common respectively.
@@
 The main example uses a twin alignment that has the same number
-of mutations (as measured from the ancestral sequence) as the true alignment.
-
-Here, we show the same results, with the difference that
+of mutations (as measured from the ancestral sequence) as the true alignment. Here, we show the same results, with the difference that
 the twin alignment uses the same mutation rate, yet is not guaranteed
-to have the same number of mutations.
+to have the same number of mutations. Any difference in the error distributions can therefore also stem from a difference in the amount of information contained in the alignments.
 
@@
   \includegraphics[width=\textwidth]{pirouette_example_35/errors.png}
-  \caption{Per-nucleotide mutation rate of 0.25 / crown age, 11 hours}
+  \caption{Error distribution for the example tree of the main text but
+with a per-nucleotide mutation rate of 0.25 / crown age. Run time of 11 hours.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_36/errors.png}
-  \caption{Per-nucleotide mutation rate of 0.50 / crown age, 14 hours}
+  \caption{Error distribution for the example tree of the main text but
+with a per-nucleotide mutation rate of 0.50 / crown age. Run time of 14 hours.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_37/errors.png}
-  \caption{Per-nucleotide mutation rate of 0.75 / crown age, 16 hours}
+  \caption{Error distribution for the example tree of the main text but
+with a per-nucleotide mutation rate of 0.75 / crown age. Run time of 16 hours.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Per-nucleotide mutation rate of 1.0 / crown age, 16 hours (main example)}
+  \caption{Error distribution for the example tree of the main text but
+with a per-nucleotide mutation rate of 1.0 / crown age. Run time of 16 hours (main example).}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_38/errors.png}
-  \caption{Per-nucleotide mutation rate of 1.25 / crown age, 17 hours}
+  \caption{Error distribution for the example tree of the main text but
+with a per-nucleotide mutation rate of 1.25 / crown age. Run time of 17 hours.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_39/errors.png}
-  \caption{Per-nucleotide mutation rate of 1.50 / crown age, 10 hours}
+  \caption{Error distribution for the example tree of the main text but
+with a per-nucleotide mutation rate of 1.50 / crown age. Run time of 10 hours.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_40/errors.png}
-  \caption{Per-nucleotide mutation rate of 2.0 / crown age, 19 hours}
+  \caption{Error distribution for the example tree of the main text but
+with a per-nucleotide mutation rate of 2.0 / crown age. Run time of 19 hours.}
 \end{figure}

### 2020-03-30 supplement.tex
Correct capitalization of nLTT

@@
 From this alignment two posterior distributions of trees are created: one using the generative model and another one using the inference model with the highest marginal likelihood.
-From each of the two, exploiting the nltt error statistics, a distribution of errors between the posterior trees and the main trees is drawn.
+From each of the two, exploiting the nLTT error statistics, a distribution of errors between the posterior trees and the main trees is drawn.
 From the true tree also a twin tree is created (right side of the figure) which follows the same pipeline, leading to two additional error distributions to use as baseline errors for their equivalents.}

### 2020-03-30 supplement.tex
Changes @Giappo

--- 2020-03-30 supplement.tex
@@
   \includegraphics[width=\textwidth]{pirouette_example_18/errors.png}
-  \caption{Equal mutation rate, 10 hours}
+  \caption{Error distribution for the example tree of the main text but,
+in this case, the number of mutations is not imposed to be the same between true and twin alignment. Instead we just use an equal mutation rate. Run time of 10 hours.}
 \end{figure}

### 2020-03-30 content.tex

@@
 a promising starting point, as it can compare any two trees and 
-results in an error distribution of known range, but one may also explore other statistics. One such statistic is phylogenetic tree balance. In principle,
+results in an error distribution of known range, but one may also explore other statistics.
+\giovanni{tree imbalance is NOT measured with gamma but with beta among other metrics.
+If we want to mention gamma we should talk about tree balance.
+I suggest to remove the last sentence here, in order to avoid talking about topology.}
+One such statistic is phylogenetic tree balance. In principle,
 \verb;pirouette; allows for this, but in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.

### 2020-03-30 content.tex
Remove sentence, as suggested by @Giappo

--- 2020-03-30 content.tex
@@
 a promising starting point, as it can compare any two trees and 
-results in an error distribution of known range, but one may also explore other statistics. One such statistic is phylogenetic tree balance. In principle,
-\verb;pirouette; allows for this, but in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.
+results in an error distribution of known range, but one may also explore other statistics.
+In principle, \verb;pirouette; allows for this, but in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.
 

### 2020-03-30 content.tex
Changes @Giappo

@@
 One such example is the diversity-dependent (DD) diversification 
-model [\cite{etienne2020package, etienne2012diversity}]:
+model [\cite{DDD, etienne2012diversity}]:
 the DD model is a BD model with a speciation or extinction rate that depends 
@@
   \centering
-  \includegraphics[width = \textwidth]{workflow4.png}
+  \includegraphics[width = 0.99\textwidth]{workflow4.png}
   \caption{
@@
 We show the usage of \verb;pirouette; on an tree generated 
-by the non-standard diversity-dependent (DD) tree model [\citep{etienne2020package, etienne2012diversity}],
+by the non-standard diversity-dependent (DD) tree model [\citep{DDD, etienne2012diversity}],
 which is a BD model with a speciation rate that is dependent on the number of species.
@@
 alignments. The RLN clock model is an unexpected choice: the RLN clock
-model assumes that mutation rates differ between species, whereas the
-alignment was simulated with equal mutation rates for all species.
+model assumes that mutation rates differ between lineages, whereas the
+alignment was simulated with equal mutation rates for all lineages.
 The BD model is indeed closer to the DD model than Yule, as both BD and DD
@@
 In short, the approach is applied to empirical trees and compares the posterior and prior distribution of trees (with the latter generated with the posterior parameters!).
-\verb;pirouette; aims to identify when assuming standard priors, not only for the species tree - which was our main focus here -, but also for the site and clock model, leads to incorrect inference if one believes more complex diversification, site or clock models are operating than can be currently accommodated in inference.
+\verb;pirouette; aims to identify when assuming standard priors for the species tree leads to incorrect inference if one believes more complex diversification models are operating than can be currently accommodated in inference.
 In short, our approach applies to simulated trees and compares the posterior distributions of trees generated with a standard and non-standard model, but inferred with a standard one.
@@
 
+However, we note that the \verb;pirouette; pipeline is not restricted to exploring the effects of a new species tree model.
+The pipeline can also be used to explore the effects of non-standard clock or site models, such as relaxed clock models with a non-standard distribution, correlated substitutions on sister lineages, or elevated substitutions rates during speciation events.
+It is, however, beyond the scope of this paper to discuss all these options in more detail.
+
 In conclusion, \verb;pirouette; can show the errors to be expected
-when the tree model assumed in inference is different from the 
-actual generative tree model.
-The user can then judge whether or not a new tree prior, 
-tailored to the novel tree model, is needed in our Bayesian phylogenetic tool.
+when the model assumed in inference is different from the 
+actual generative model.
+The user can then judge whether or not this new model should be implemented in our Bayesian phylogenetic tool.
 

### 2020-03-30 supplement.tex
Changes @Giappo

--- 2020-03-30 supplement.tex
@@
   }
-  \item{
-    subsection \ref{subsec:non_clock}: the effect when the alignment
-    is generated by a non-clock like nucleotide substitution model
-  }
+%  \item{
+%    subsection \ref{subsec:non_clock}: the %effect when the alignment
+%    is generated by a non-clock like nucleotide substitution model
+%  }
   \item{
@@
 \begin{itemize}
-  \item the use of 20 replicates results in decent plots
   \item{
@@
     When we do not impose the same number of mutations between true and twin alignment, we observe a difference in the error distributions with respect to the standard case (presented in the main text) where they are forced to have the same number of mutations.
-    The consequences of this are unknown.
   }
@@
   \label{fig:example_30_full_pipeline}
-  \caption{Comparing to background noise: full pipeline.
+  \caption{Full pirouette pipeline, including comparison to baseline error.
 The true tree (top right) is used to simulate an alignment.
 From this alignment two posterior distributions of trees are created: one using the generative model and another one using the inference model with the highest marginal likelihood.
-From each of the two, exploiting the nLTT error statistics, a distribution of errors between the posterior trees and the main trees is drawn.
-From the true tree also a twin tree is created (right side of the figure) which follows the same pipeline, leading to two additional error distributions to use as baseline errors for their equivalents.}
+For each distribution of trees, a distribution of errors, measured with the nLTT statistic, between the posterior trees and the main trees is drawn.
+From the true tree also a twin tree is created (right side of the figure) which follows the same pipeline, leading to two additional error distributions to use as baseline errors.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Error distribution from a distribution of trees.
-What is shown in the figure is similar to Fig.
-\ref{fig:example_30} for the main example.
-The main difference is that here the original tree (and its twin counterpart) is not only one, but instead there are 20 replicated.
-Each replicate has been simulated independently.
-For each setting (true generative, true best candidate, twin generative and twin candidate), the resulting errors from each replicated pipeline have been merged in a single dataset.}
+  \caption{Error distributions of trees, similar to Fig.
+\ref{fig:example_30} for the main example, but now for a collection of 20 replicate trees.
+For each setting (true generative, true best candidate, twin generative and twin best candidate), the resulting errors from each replicate pipeline have been merged into a single distribution.}
+  \label{fig:replicate_trees}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{6 taxa, 16 hours, main example}
+  %\caption{Error distributions of trees for 6 taxa, using 20 replicates. This took 16 hours to compute.}
+  \caption{Aggregate error distributions for 20 replicates. Here each each true tree has 6 taxa. This took 16 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_32/errors.png}
-  \caption{12 taxa, 19 hours}
+  %\caption{Error distributions of trees for 12 taxa, using 10 replicates. This took 19 hours to compute.}
+  \caption{Aggregate error distributions for 10 replicates. Here each each true tree has 12 taxa. This took 19 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_33/errors.png}
-  \caption{24 taxa, 28 hours}
+  %\caption{Error distributions of trees for 24 taxa, using 20 replicates. This took 28 hours to compute.}
+  \caption{Aggregate error distributions for 20 replicates. Here each each true tree has 24 taxa. This took 28 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_41/errors.png}
-  \caption{32 taxa, 7 hours}
+  %\caption{Error distributions of trees for 32 taxa, using 5 replicates. This took 7 hours to compute.}
+  \caption{Aggregate error distributions for 5 replicates. Here each each true tree has 32 taxa. This took 7 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_42/errors.png}
-  \caption{40 taxa, 8 hours}
+  %\caption{Error distributions of trees for 40 taxa, using 2 replicates. This took 8 hours to compute.}
+  \caption{Aggregate error distributions for 2 replicates. Here each each true tree has 40 taxa. This took 8 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_19/errors.png}
-  \caption{500 nucleotides, 12 hours}
+  %\caption{Error distributions of trees for a sequence length of 500 nucleotides, using 20 replicates. This took 12 hours to compute.}
+  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{1000 nucleotides, 16 hours (main example)}
+  %\caption{Error distributions of trees for a sequence length of 1000 nucleotides (as used in the main example), using 20 replicates. This took 16 hours.}
+  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_34/errors.png}
-  \caption{2000 nucleotides, 11 hours}
+  %\caption{Error distributions of trees for a sequence length of 2000 nucleotides (as used in the main example), using 10 replicates. This took 11 hours.}
+  \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The effect of non-clock like models}
-\label{subsec:non_clock}
+%\subsection{The effect of non-clock like %models}
+%\label{subsec:non_clock}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-The main example generates an alignment using a strict clock model (that is,
-all species have an equal mutation rate).
-One may wonder, how well the inference will be for a different non-clock like
-model.
-Here, we show the same results as the main example,
-except that instead of generating the alignment with a strict clock
-model, we use an unlinked node substitution model (by Thijs Janzen,
-unpublished, code available at \url{https://github.com/thijsjanzen/nodeSub}),
-with a branch mutation rate of $0.1$ and a node mutation rate of $0.1$.
+%The main example generates an alignment using a %strict clock model (that is, all species have an equal mutation rate).
+%One may wonder how well the inference will be for a different non-clock like model.
+%Here, we show the same results as the main example, except that instead of generating the alignment with a strict clock model, we use an unlinked node substitution model (by Thijs Janzen, unpublished, code available at \url{https://github.com/thijsjanzen/nodeSub}), with a branch mutation rate of $0.1$ and a node mutation rate of $0.1$.
 
-The code to reproduce this figure can be found at  
-\url{https://github.com/richelbilderbeek/pirouette_example_27}.
+%The code to reproduce this figure can be found at \url{https://github.com/richelbilderbeek/pirouette_example_27}.
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_27/errors.png}
-  \caption{An unlinked node substitution model, 20 replicates, 8 hours}
-\end{figure}
+%\begin{figure}[H]
+%  \includegraphics[width=\textwidth]{pirouette_example_27/errors.png}
+%  \caption{An unlinked node substitution model, 20 replicates, 8 hours}
+%\end{figure}
 
@@
   \includegraphics[width=\textwidth]{pirouette_example_22/errors.png}
-  \caption{Assuming a Yule tree prior on a Yule tree, 9 hours}
+  %\caption{Error distributions of trees when using a Yule prior on a Yule tree, for 10 replicates. This took 9 hours to compute.}
+  \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_26/errors.png}
-  \caption{Assuming a Yule tree prior on a BD tree, 9 hours}
+  %\caption{Error distributions of trees when using a Yule prior on a BD tree, for 20 replicates. This took 9 hours to compute.}
+  \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_23/errors_low.png}
-  \caption{Lowest likelihood}
+  %\caption{Error distributions of trees using a true tree a DD tree with low likelihood.}
+  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with low likelihood.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_23/errors_mid.png}
-  \caption{Median likelihood}
+  %\caption{Error distributions of trees using a true tree a DD tree with median likelihood.}
+  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with median likelihood.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_23/errors_high.png}
-  \caption{Highest likelihood}
+  %\caption{Error distributions of trees using a true tree a DD tree with high likelihood.}
+  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with high likelihood.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_18/errors.png}
-  \caption{Equal mutation rate, 10 hours}
+  \caption{Aggregate error distributions similar to Fig.
+\ref{fig:replicate_trees}, but here the number of mutations is not imposed to be the same between true and twin alignment.
+Instead we just use an equal mutation rate.
+This took 10 hours to compute.}
 \end{figure}
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Equal mutation rate and number of mutations, 16 hours (main example)}
-\end{figure}
+%\begin{figure}[H]
+%  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+%  \caption{Equal mutation rate and number of mutations, 16 hours (main example)}
+%\end{figure}
+
+\clearpage
 
@@
 \url{https://github.com/richelbilderbeek/pirouette_example_37} (0.75 / crown age),
-\url{https://github.com/richelbilderbeek/pirouette_example_28} (1.00 / crown age, main example),
+\url{https://github.com/richelbilderbeek/pirouette_example_28} (1.00 / crown age, example reported in \ref{subsec:distribution}, see Fig. \ref{fig:replicate_trees}),
 \url{https://github.com/richelbilderbeek/pirouette_example_38} (1.25 / crown age),
@@
   \includegraphics[width=\textwidth]{pirouette_example_35/errors.png}
-  \caption{Error distribution for the example tree of the main text but
-with a per-nucleotide mutation rate of 0.25 / crown age. Run time of 11 hours.}
+  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.25 / crown age. This took 11 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_36/errors.png}
-  \caption{Error distribution for the example tree of the main text but
-with a per-nucleotide mutation rate of 0.50 / crown age. Run time of 14 hours.}
+  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.50 / crown age. This took 14 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_37/errors.png}
-  \caption{Error distribution for the example tree of the main text but
-with a per-nucleotide mutation rate of 0.75 / crown age. Run time of 16 hours.}
+  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.75 / crown age. This took 16 hours to compute.}
 \end{figure}
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Error distribution for the example tree of the main text but
-with a per-nucleotide mutation rate of 1.0 / crown age. Run time of 16 hours (main example).}
-\end{figure}
+%begin{figure}[H]
+% \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+% \caption{Error distribution for the example tree of the main text which has a per-nucleotide mutation rate of 1.0 / crown age. This took 16 hours to compute.}
+%end{figure}
 
@@
   \includegraphics[width=\textwidth]{pirouette_example_38/errors.png}
-  \caption{Error distribution for the example tree of the main text but
-with a per-nucleotide mutation rate of 1.25 / crown age. Run time of 17 hours.}
+  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.25 / crown age.This took 17 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_39/errors.png}
-  \caption{Error distribution for the example tree of the main text but
-with a per-nucleotide mutation rate of 1.50 / crown age. Run time of 10 hours.}
+  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.50 / crown age. This took 10 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_40/errors.png}
-  \caption{Error distribution for the example tree of the main text but
-with a per-nucleotide mutation rate of 2.0 / crown age. Run time of 19 hours.}
+  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 2.0 / crown age. This took 19 hours to compute.}
 \end{figure}

### 2020-03-30 supplement.tex
Added simplified version of pirouette example 30 code

@@
   \item{
-    subsection \ref{subsec:main_example}: extra figures and
+    subsection \ref{subsec:main_example}: code, extra figures and
     diagnostics regarding the main example.
@@
 The code to reproduce this can be found at  
-\url{https://github.com/richelbilderbeek/pirouette_example_30}.
+\url{https://github.com/richelbilderbeek/pirouette_example_30}
+and a simplified version is shown for convenience here as well:
+
+\begin{lstlisting}[language=R]
+library(pirouette)
+
+# Create phylogeny
+set.seed(314)
+phylogeny <- create_exemplary_dd_tree(
+  n_taxa = 6, 
+  crown_age = 10
+)
+
+# Setup pirouette
+pir_params <- create_std_pir_params(
+  folder_name = "example_30"
+)
+
+# Do the runs
+pir_out <- pir_run(
+  phylogeny = phylogeny,
+  pir_params = pir_params
+)
+
+# Save
+library(ggplot2)
+pir_outs <- list()
+pir_outs[[1]] <- pir_out
+pir_plots(pir_outs) +
+  ggtitle("pirouette example 30") +
+  ggsave("errors.png", width = 7, height = 7)
+
+pir_save(
+  phylogeny = phylogeny,
+  pir_params = pir_params,
+  pir_out = pir_out,
+  folder_name = "example_30"
+)
+\end{lstlisting}
 

### 2020-03-30 supplement.tex
No nodeSub
User also needs to install mcbette

@@
   }
-%  \item{
-%    subsection \ref{subsec:non_clock}: the %effect when the alignment
-%    is generated by a non-clock like nucleotide substitution model
-%  }
   \item{
@@
 ]
+remotes::install_github(
+  "richelbilderbeek/mcbette"
+)
 remotes::install_github(
@@
 \url{https://github.com/richelbilderbeek/pirouette_example_30}
-and a simplified version is shown for convenience here as well:
+and a simplified version is shown here for convenience:
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{The effect of non-clock like %models}
-%\label{subsec:non_clock}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%The main example generates an alignment using a %strict clock model (that is, all species have an equal mutation rate).
-%One may wonder how well the inference will be for a different non-clock like model.
-%Here, we show the same results as the main example, except that instead of generating the alignment with a strict clock model, we use an unlinked node substitution model (by Thijs Janzen, unpublished, code available at \url{https://github.com/thijsjanzen/nodeSub}), with a branch mutation rate of $0.1$ and a node mutation rate of $0.1$.
-
-%The code to reproduce this figure can be found at \url{https://github.com/richelbilderbeek/pirouette_example_27}.
-
-%\begin{figure}[H]
-%  \includegraphics[width=\textwidth]{pirouette_example_27/errors.png}
-%  \caption{An unlinked node substitution model, 20 replicates, 8 hours}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{The effect of assuming a Yule tree prior on a Yule tree}

### 2020-03-30 scripts/create_article.sh
Renamed files [WIP]

@@ -8,7 +8,7 @@
 #
 #
 
-pdflatex article.tex --shell-escape
-bibtex article.aux
-pdflatex article.tex --shell-escape
-pdflatex article.tex --shell-escape
+pdflatex article --shell-escape
+bibtex article
+pdflatex article --shell-escape
+pdflatex article --shell-escape

### 2020-03-30 scripts/create_article.sh
Renamed files for thesis

@@ -8,7 +8,8 @@
 #
 #
 
-pdflatex article --shell-escape
-bibtex article
-pdflatex article --shell-escape
-pdflatex article --shell-escape
+pdflatex pirouette_header --shell-escape
+bibtex pirouette_header
+pdflatex pirouette_header --shell-escape
+pdflatex pirouette_header --shell-escape
+mv pirouette_header.pdf pirouette_article.pdf

### 2020-03-30 pirouette_article.tex
Renamed files for thesis
Changes @Giappo

--- 2020-03-30 content.tex
@@
     Fig~\ref{fig:pipeline}. To run the pipeline $\mathit{A}$, $\mathit{X}$ 
-    and $\mathit{E}$ must be specified. Examples can be found in 
-    listings~\ref{lst:create_alignment_params}, 
-    \ref{lst:create_gen_experiment_explicit} and 
-    \ref{lst:create_error_measure_params}.
+    and $\mathit{E}$ must be specified.
   }
@@
 
-We use the DD tree as shown in Figure \ref{fig:dd_tree},
-which has an arbitrary crown age of ten time units
-and six taxa.
+The code to reproduce our results can be found at  
+\url{https://github.com/richelbilderbeek/pirouette_example_30}
+and a simplified version is shown here for convenience:
+
+\begin{lstlisting}[language=R]
+library(pirouette)
+
+# Create phylogeny
+set.seed(314)
+phylogeny <- create_exemplary_dd_tree(
+  n_taxa = 6, 
+  crown_age = 10
+)
+
+# Setup pirouette
+pir_params <- create_std_pir_params(
+  folder_name = "example_30"
+)
+
+# Do the runs
+pir_out <- pir_run(
+  phylogeny = phylogeny,
+  pir_params = pir_params
+)
+
+# Save
+library(ggplot2)
+pir_outs <- list()
+pir_outs[[1]] <- pir_out
+pir_plots(pir_outs) +
+  ggtitle("pirouette example 30") +
+  ggsave("errors.png", width = 7, height = 7)
+
+pir_save(
+  phylogeny = phylogeny,
+  pir_params = pir_params,
+  pir_out = pir_out,
+  folder_name = "example_30"
+)
+\end{lstlisting}
+
+The DD tree that is generated by this code is as shown in Figure \ref{fig:dd_tree},
+which has an arbitrarily chosen crown age of ten time units and six taxa.
 
@@
   \caption{
-    The tree derived from a DD process.
+    The example tree resulting from a diversity-dependent (DD) simulation.
   }
@@
 % MEE style
-\bibliographystyle{mee}
-\bibliography{article}
+\bibliographystyle{pirouette_mee}
+\bibliography{pirouette_article}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-03-30 pirouette_supplement.tex
Renamed files [WIP]
Changes @Giappo

--- 2020-03-30 supplement.tex
@@
 one can download and install the package from \verb;pirouette;'s GitHub 
-repository:
-
+repository. We first need the \verb;mcbette; package:
 \begin{lstlisting}[
@@
 )
+\end{lstlisting}
+
+Now we can install \verb;pirouette;:
+\begin{lstlisting}[
+    language = R,
+    floatplacement = ht,
+    frame = single
+]
 remotes::install_github(
@@
 \end{lstlisting}
+which also installs its dependencies from CRAN.
+
 To start using \verb;pirouette;, 
@@
 
-The code to reproduce this can be found at  
-\url{https://github.com/richelbilderbeek/pirouette_example_30}
-and a simplified version is shown here for convenience:
-
-\begin{lstlisting}[language=R]
-library(pirouette)
-
-# Create phylogeny
-set.seed(314)
-phylogeny <- create_exemplary_dd_tree(
-  n_taxa = 6, 
-  crown_age = 10
-)
-
-# Setup pirouette
-pir_params <- create_std_pir_params(
-  folder_name = "example_30"
-)
-
-# Do the runs
-pir_out <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-
-# Save
-library(ggplot2)
-pir_outs <- list()
-pir_outs[[1]] <- pir_out
-pir_plots(pir_outs) +
-  ggtitle("pirouette example 30") +
-  ggsave("errors.png", width = 7, height = 7)
-
-pir_save(
-  phylogeny = phylogeny,
-  pir_params = pir_params,
-  pir_out = pir_out,
-  folder_name = "example_30"
-)
-\end{lstlisting}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Error distributions of trees, similar to Fig.
+  \caption{Aggregate error distributions, similar to Fig.
 \ref{fig:example_30} for the main example, but now for a collection of 20 replicate trees.
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  %\caption{Error distributions of trees for 6 taxa, using 20 replicates. This took 16 hours to compute.}
-  \caption{Aggregate error distributions for 20 replicates. Here each each true tree has 6 taxa. This took 16 hours to compute.}
+  \caption{Aggregate error distributions for 20 replicates. Here each true tree has 6 taxa. This took 16 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_32/errors.png}
-  %\caption{Error distributions of trees for 12 taxa, using 10 replicates. This took 19 hours to compute.}
-  \caption{Aggregate error distributions for 10 replicates. Here each each true tree has 12 taxa. This took 19 hours to compute.}
+  \caption{Aggregate error distributions for 10 replicates. Here each true tree has 12 taxa. This took 19 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_33/errors.png}
-  %\caption{Error distributions of trees for 24 taxa, using 20 replicates. This took 28 hours to compute.}
-  \caption{Aggregate error distributions for 20 replicates. Here each each true tree has 24 taxa. This took 28 hours to compute.}
+    \caption{Aggregate error distributions for 20 replicates. Here each true tree has 24 taxa. This took 28 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_41/errors.png}
-  %\caption{Error distributions of trees for 32 taxa, using 5 replicates. This took 7 hours to compute.}
-  \caption{Aggregate error distributions for 5 replicates. Here each each true tree has 32 taxa. This took 7 hours to compute.}
+  \caption{Aggregate error distributions for 5 replicates. Here each true tree has 32 taxa. This took 7 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_42/errors.png}
-  %\caption{Error distributions of trees for 40 taxa, using 2 replicates. This took 8 hours to compute.}
-  \caption{Aggregate error distributions for 2 replicates. Here each each true tree has 40 taxa. This took 8 hours to compute.}
+  \caption{Aggregate error distributions for 2 replicates. Here each true tree has 40 taxa. This took 8 hours to compute.}
 \end{figure}
@@
   \includegraphics[width=\textwidth]{pirouette_example_19/errors.png}
-  %\caption{Error distributions of trees for a sequence length of 500 nucleotides, using 20 replicates. This took 12 hours to compute.}
   \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  %\caption{Error distributions of trees for a sequence length of 1000 nucleotides (as used in the main example), using 20 replicates. This took 16 hours.}
   \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
@@
   \includegraphics[width=\textwidth]{pirouette_example_34/errors.png}
-  %\caption{Error distributions of trees for a sequence length of 2000 nucleotides (as used in the main example), using 10 replicates. This took 11 hours.}
   \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
@@
   \includegraphics[width=\textwidth]{pirouette_example_22/errors.png}
-  %\caption{Error distributions of trees when using a Yule prior on a Yule tree, for 10 replicates. This took 9 hours to compute.}
   \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
@@
   \includegraphics[width=\textwidth]{pirouette_example_26/errors.png}
-  %\caption{Error distributions of trees when using a Yule prior on a BD tree, for 20 replicates. This took 9 hours to compute.}
   \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
@@
   \includegraphics[width=\textwidth]{pirouette_example_23/errors_low.png}
-  %\caption{Error distributions of trees using a true tree a DD tree with low likelihood.}
   \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with low likelihood.}
@@
   \includegraphics[width=\textwidth]{pirouette_example_23/errors_mid.png}
-  %\caption{Error distributions of trees using a true tree a DD tree with median likelihood.}
   \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with median likelihood.}
@@
   \includegraphics[width=\textwidth]{pirouette_example_23/errors_high.png}
-  %\caption{Error distributions of trees using a true tree a DD tree with high likelihood.}
   \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with high likelihood.}
@@
 \end{figure}
-
-%\begin{figure}[H]
-%  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-%  \caption{Equal mutation rate and number of mutations, 16 hours (main example)}
-%\end{figure}
 
@@
 
-%begin{figure}[H]
-% \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-% \caption{Error distribution for the example tree of the main text which has a per-nucleotide mutation rate of 1.0 / crown age. This took 16 hours to compute.}
-%end{figure}
-
 \begin{figure}[H]
   \includegraphics[width=\textwidth]{pirouette_example_38/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.25 / crown age.This took 17 hours to compute.}
+  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.25 / crown age. This took 17 hours to compute.}
 \end{figure}

### 2020-03-30 pirouette_supplement.tex
Indeed need to install the nodeSub package. Thanks @rsetienne

@@
 one can download and install the package from \verb;pirouette;'s GitHub 
-repository. We first need the \verb;mcbette; package:
+repository. We do first need the \verb;mcbette; and \verb;nodeSub; packages:
 \begin{lstlisting}[
@@
   "richelbilderbeek/mcbette"
+)
+remotes::install_github(
+  "thijsjanzen/nodeSub"
 )

### 2020-03-31 scripts/split_pdf.sh
Supplementary materials also have references

@@ -13,5 +13,5 @@
 #   sudo apt-get install pdftk
 #
 
-pdftk article.pdf cat 1-35 output main.pdf
-pdftk article.pdf cat 36-45 output supp_mat.pdf
+pdftk pirouette_article.pdf cat 1-24 output main.pdf
+pdftk pirouette_article.pdf cat 25-end output supp_mat.pdf

### 2020-04-24 pirouette_article.tex

@@
   n_taxa = 6, 
-  crown_age = 10
+  crown_age = 10,
+  extinction_rate = 0.1
 )
@@
 The DD tree that is generated by this code is as shown in Figure \ref{fig:dd_tree},
-which has an arbitrarily chosen crown age of ten time units and six taxa.
+which has an arbitrarily chosen crown age of 10 time units and 6 tips for an extinction rate of 0.1.
+The carrying-capacity is also set at 6.
+The initial speciation rate $\lambda_0$ is chosen such that the expected number of species in a constant-rate BD model would be equal to the number of tips, which amounts to $\lambda_0 = 0.63$.
 

### 2020-04-24 pirouette_supplement.tex
Supplementary materials also have references
Use sloppypar to make it look tidier :confused:

@@
 \verb;pirouette; is free, libre and open source software available at 
+\begin{sloppypar}
   \url{http://github.com/richelbilderbeek/pirouette},
+\end{sloppypar}
 licensed under the GNU General Public License version 3.
@@
 \verb;pirouette;'s development takes place on GitHub,
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette},
+\end{sloppypar}
 which allows submitting bug reports, requesting features, 
@@
 tested, can be found at
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette_examples}. 
+\end{sloppypar}
 
@@
 uses one tree.
-To assess if the results of the inference are meaningful one important parameter is the Effective Sample Size (ESS).
-This quantity describes how many independent trees are sampled from the posterior distributions.
-For reliable results it is good practice to have at least $ESS = 200$ (see \url{https://beast.community/ess_tutorial}).
+To assess if the results of the inference are meaningful one important 
+parameter is the Effective Sample Size (ESS). This quantity describes how 
+many independent trees are sampled from the posterior distributions. 
+For reliable results it is good practice to have at 
+least $ESS = 200$ (see 
+\begin{sloppypar}
+  \url{https://beast.community/ess_tutorial}).
+\end{sloppypar}
 In the following we present the ESS for the posterior distributions of the 4 cases shown in Fig.
@@
   }
-  \label{fig:example_30_full_pipeline}
   \caption{Full pirouette pipeline, including comparison to baseline error.
@@
 From the true tree also a twin tree is created (right side of the figure) which follows the same pipeline, leading to two additional error distributions to use as baseline errors.}
+  \label{fig:example_30_full_pipeline}
 \end{figure}
@@
 The code to reproduce this figure can be found at  
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette_example_28}.
+\end{sloppypar}
 
@@
 The code to reproduce these figure can be found at  
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette_example_28} (6 taxa, main example),
@@
   \url{https://github.com/richelbilderbeek/pirouette_example_42} (40 taxa). 
+\end{sloppypar}
 
@@
 The code to reproduce these figures can be found at  
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette_example_19} (500 nucleotides),
@@
   and \url{https://github.com/richelbilderbeek/pirouette_example_34} (2000 nucleotides).
+\end{sloppypar}
 
@@
 The code to reproduce this figure can be found at  
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette_example_22}.
+\end{sloppypar}
 
@@
 The code to reproduce this figure can be found at
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette_example_26}.
+\end{sloppypar}
 
@@
 The code to reproduce these figure can be found at  
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette_example_23}
+\end{sloppypar}
 and took 120 hours in total.
@@
 The code to reproduce this figure can be found at
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette_example_18} and
   \url{https://github.com/richelbilderbeek/pirouette_example_28}.
+\end{sloppypar}
 
@@
 The code to reproduce this figure can be found at
+\begin{sloppypar}
   \url{https://github.com/richelbilderbeek/pirouette_example_35} (0.25 / crown age),
@@
   \url{https://github.com/richelbilderbeek/pirouette_example_40} (2.00 / crown age),
+\end{sloppypar}
 
@@
 \end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Bibliography
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% MEE style
+\bibliographystyle{pirouette_mee}
+\bibliography{pirouette_supplement}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-04-24 scripts/merge_files.sh
Merge into one file

@@ -1,0 +1,16 @@
+#!/bin/bash
+#
+# Merge all .tex files to one file 'pirouette_article.tex'.
+# 'pirouette_article.tex' is ignored by git
+#
+# Usage:
+#
+#   ./scripts/merge_files.sh
+#
+#
+# 
+# Follow same structure as article_header.tex
+
+cat pirouette_content.tex \
+  pirouette_supplement.tex > pirouette_article.tex
+

### 2020-04-24 scripts/merge_files.sh
Revert "Merge into one file"
This reverts commit c8d136b3d49907e7975ff76cbefd346080018006.

@@ -1,16 +1,0 @@
-#!/bin/bash
-#
-# Merge all .tex files to one file 'pirouette_article.tex'.
-# 'pirouette_article.tex' is ignored by git
-#
-# Usage:
-#
-#   ./scripts/merge_files.sh
-#
-#
-# 
-# Follow same structure as article_header.tex
-
-cat pirouette_content.tex \
-  pirouette_supplement.tex > pirouette_article.tex
-

### 2020-05-27 pirouette_article.tex
Progress response letter 2

@@
     Here we present \verb;pirouette;, 
-    a free, libre and open-source R package that assesses 
+    a free and open-source R package that assesses 
     the inference error made by Bayesian phylogenetics for a given macroevolutionary diversification model. \verb;pirouette; makes use of 

### 2020-05-27 pirouette_supplement.tex
Only one bib style and source
pirouette has refs to citations and figs
Stick to 80 chars per line
Progress response letter 2

@@
 This supplementary material contains additional facets of \verb;pirouette;, such as the installation of the package, an overview of
-pirouette's main functions and a guide for empiricists, based on multiple experiments
+pirouette's main functions and a guide for users, based on multiple experiments
 that are shown here as well.
@@
   \item{
-    subsection \ref{subsec:guidelines}: guidelines for empiricists
+    subsection \ref{subsec:guidelines}: guidelines for users
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Guidelines for empiricists}
+\subsection{Guidelines for users}
 \label{subsec:guidelines}
@@
 \end{sloppypar}
-In the following we present the ESS for the posterior distributions of the 4 cases shown in Fig.
-\ref{fig:example_30_full_pipeline}: "true" pipeline with generative model (Table \ref{tab:esses_true_gen}), "true" pipeline with best candidate model (Table \ref{tab:esses_true_best}), "twin" pipeline with generative model (Table \ref{tab:esses_twin_gen}), "twin" pipeline with best candidate model (Table \ref{tab:esses_twin_best}).
-We also report the marginal likelihood (or evidence) data for model selection performed both in the "true" (Table \ref{tab:true_evidence}) and "twin" pipeline (Table \ref{tab:twin_evidence}).
+In the following we present the ESS for the posterior distributions of 
+the 4 cases shown in Fig. \ref{fig:example_30_full_pipeline}: 
+"true" pipeline with generative model (Table \ref{tab:esses_true_gen}), 
+"true" pipeline with best candidate model (Table \ref{tab:esses_true_best}), 
+"twin" pipeline with generative model (Table \ref{tab:esses_twin_gen}), 
+"twin" pipeline with best candidate model (Table \ref{tab:esses_twin_best}). 
+We also report the marginal likelihood (or evidence) data for model selection 
+performed both in the "true" (Table \ref{tab:true_evidence}) and 
+"twin" pipeline (Table \ref{tab:twin_evidence}).
 
@@
   }
-  \caption{Full pirouette pipeline, including comparison to baseline error.
-The true tree (top right) is used to simulate an alignment.
-From this alignment two posterior distributions of trees are created: one using the generative model and another one using the inference model with the highest marginal likelihood.
+  \caption{
+    Full pirouette pipeline, including comparison to baseline error. 
+    The true tree (top left) is used to simulate an alignment. 
+    From this alignment two posterior distributions of trees are created: 
+    one using the generative model and another one using the inference model with the highest marginal likelihood.
 For each distribution of trees, a distribution of errors, measured with the nLTT statistic, between the posterior trees and the main trees is drawn.
-From the true tree also a twin tree is created (right side of the figure) which follows the same pipeline, leading to two additional error distributions to use as baseline errors.}
+From the true tree also a twin tree is created (right side of the figure) which follows the same pipeline, leading to two additional error distributions to use as baseline errors.
+  }
   \label{fig:example_30_full_pipeline}
@@
 \bibliographystyle{pirouette_mee}
-\bibliography{pirouette_supplement}
+% \bibliography{pirouette_supplement}
+\bibliography{pirouette_article}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-05-29 pirouette_supplement.tex
Changes @Giappo

@@
 
-The code to reproduce this figure can be found at  
-\begin{sloppypar}
-  \url{https://github.com/richelbilderbeek/pirouette_example_28}.
-\end{sloppypar}
-
 \begin{figure}[H]
@@
 \end{figure}
+
+\new{
+The resulting error distributions are showed in Fig.
+\ref{fig:replicate_trees}.
+We present results for cases where (1) the generative model has been used or (2) the model with highest evidence has been selected for the inference.
+From the plots we can see that in both cases there is a clear distinction between the two distributions (true and twin), suggesting that the inference models that have been used cannot capture in an accurate way the features of the diversity-dependent tree prior used to simulate the original trees.
+}
+
+The code to reproduce Fig. \ref{fig:replicate_trees} can be found at  
+\begin{sloppypar}
+  \url{https://github.com/richelbilderbeek/pirouette_example_28}.
+\end{sloppypar}
 
@@
 
-The code to reproduce these figure can be found at  
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+  \caption{Aggregate error distributions for 20 replicates. Here each true tree has 6 taxa. This took 16 hours to compute.}
+  \label{fig:example_6_taxa}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_32/errors.png}
+  \caption{Aggregate error distributions for 10 replicates. Here each true tree has 12 taxa. This took 19 hours to compute.}
+  \label{fig:example_12_taxa}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_33/errors.png}
+    \caption{Aggregate error distributions for 20 replicates. Here each true tree has 24 taxa. This took 28 hours to compute.}
+    \label{fig:example_24_taxa}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_41/errors.png}
+  \caption{Aggregate error distributions for 5 replicates. Here each true tree has 32 taxa. This took 7 hours to compute.}
+  \label{fig:example_32_taxa}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_42/errors.png}
+  \caption{Aggregate error distributions for 2 replicates. Here each true tree has 40 taxa. This took 8 hours to compute.}
+  \label{fig:example_40_taxa}
+\end{figure}
+
+\new{
+We show in figures \ref{fig:example_6_taxa}, \ref{fig:example_12_taxa}, \ref{fig:example_24_taxa}, \ref{fig:example_32_taxa} and \ref{fig:example_40_taxa} what are the errors obtained when starting from phylogenies with, respectively, 6, 12, 24, 32 and 40 taxa.
+Again we can see that in each case errors tend to be greater in the true distribution than in the twin distribution, confirming the result of the previous paragraph.
+However we cannot detect a clear trend between the number of taxa of the original phylogeny and the extent of such discrepancy.
+Due to long computational times we have been able to provide data only for a limited amount of replicates.
+We cannot exclude that a clearer trend could be observed increasing the number of replicates.
+}
+
+The code to reproduce these figures can be found at  
 \begin{sloppypar}
@@
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each true tree has 6 taxa. This took 16 hours to compute.}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_32/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. Here each true tree has 12 taxa. This took 19 hours to compute.}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_33/errors.png}
-    \caption{Aggregate error distributions for 20 replicates. Here each true tree has 24 taxa. This took 28 hours to compute.}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_41/errors.png}
-  \caption{Aggregate error distributions for 5 replicates. Here each true tree has 32 taxa. This took 7 hours to compute.}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_42/errors.png}
-  \caption{Aggregate error distributions for 2 replicates. Here each true tree has 40 taxa. This took 8 hours to compute.}
-\end{figure}
+\newpage
 
@@
 
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_19/errors.png}
+  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
+  \label{fig:example_500_nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
+  \label{fig:example_1000_nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_34/errors.png}
+  \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
+  \label{fig:example_2000_nucleotides}
+\end{figure}
+
+\new{
+From figures \ref{fig:example_500_nucleotides}, \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} we can observe a that the discrepancy between the true and twin error distributions tend to become smaller as the number of nucleotides increase.
+This occurred for both the generative and best candidate cases.
+}
+\giovanni{richel do you have an idea why this happens?
+Also, do we have the raw data stored somewhere?
+From what I can judge from the plots the median discrepancy is 0.012 for 500 nucs, 0.008 for 1000 nucs and 0.003 for 2000 nucs.}
+
 The code to reproduce these figures can be found at  
@@
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_19/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_34/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
-\end{figure}
+\newpage
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The effect of assuming a Yule tree prior on a Yule tree}
+\subsection{Effect of assuming a Yule tree prior on a Yule tree}
 \label{subsec:simplest_correct_parameterization}
@@
 which we also assume as the (correct) tree prior.
-This example thus shows a parameterization at the correct level for the
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_22/errors.png}
+  \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
+  \label{fig:example_yule}
+\end{figure}
+
+This example shows a parameterization at the correct level for the
 simplest case possible.
+\new{
+As expected the twin and true distributions in Fig. \ref{fig:example_yule} are extremely similar for both the generative and the best candidate case.
+}
 
@@
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_22/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
-\end{figure}
+\newpage
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The effect of assuming a Yule tree prior on a BD tree}
+\subsection{Effect of assuming a Yule tree prior on a BD tree}
 \label{subsec:under_parameterization}
@@
 The main example uses a tree generated by a non-standard tree model.
-
 Here, we show the same results, with the difference that
@@
 This example thus shows the effect of underparameterization.
+
+\begin{figure}[H]
+  \includegraphics[width=\textwidth]{pirouette_example_26/errors.png}
+  \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
+\end{figure}
+
+\giovanni{This is the example one of the reviewer was referring to.
+The figure that was present before (still present in the overleaf files) looked actually quite different from the one on the repo (\url{https://raw.githubusercontent.com/richelbilderbeek/pirouette_example_26/master/errors.png}).
+I changed it with the one found there.
+In case you think this is wrong you can revert it back.}
+
+\new{
+Since the two models are very similar to each other (the BD model can be turned into a Yule model just by setting the extinction parameter to zero \giovanni{Cite Louca and Pennell here?}) the median discrepancy it's almost negligible.
+However, respect to the previous case when a Yule tree prior has been considered, here the distributions exhibit a greater difference in shapes.
+}
 
@@
 
-\begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_26/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
-\end{figure}
+\newpage
 
@@
 that are rare, uncommon and common respectively.
-
-The code to reproduce these figure can be found at  
-\begin{sloppypar}
-  \url{https://github.com/richelbilderbeek/pirouette_example_23}
-\end{sloppypar}
-and took 120 hours in total.
 
@@
 
+\new{
+Here the median errors are similar in the three settings and similar to ones relative to the full dataset of \ref{subsec:distribution}.
+We can also notice that in the case of median likelihood, the twin median error appear to be lower than the true mean error.
+This usually is a sign that the number of replicates (in this case 10) is too low to allow us to draw precise conclusions from this test.
+We decided to not explore further in this direction using more simulations because computational times turned to be extremely high.
+The entire run took 120 hours in total.
+}
+
+The code to reproduce these figure can be found at  
+\begin{sloppypar}
+  \url{https://github.com/richelbilderbeek/pirouette_example_23}
+\end{sloppypar}.
+
+\newpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The code to reproduce this figure can be found at
-\begin{sloppypar}
-  \url{https://github.com/richelbilderbeek/pirouette_example_18} and
-  \url{https://github.com/richelbilderbeek/pirouette_example_28}.
-\end{sloppypar}
-
 \begin{figure}[H]
@@
 This took 10 hours to compute.}
+  \label{fig:example_random_mutations}
 \end{figure}
+
+\new{
+Comparing Fig.
+\ref{fig:example_random_mutations} to  \label{fig:replicate_trees} we can see that the extent of the error tend to increase both for the true and the twin distributions.
+This is probably due to the fact that
+}
+
+The code to reproduce this figure can be found at
+\begin{sloppypar}
+  \url{https://github.com/richelbilderbeek/pirouette_example_18} and
+  \url{https://github.com/richelbilderbeek/pirouette_example_28}.
+\end{sloppypar}
 

### 2020-05-29 pirouette_supplement.tex
Use '\newpage' instead of '\clearpage' as the latter destroys the layout and adds a lot of comma's for some reason :confused:

@@
 
-\clearpage
-%\newpage
+\newpage
 
@@
   \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
-  \caption{Aggregate error distributions, similar to Fig.
-\ref{fig:example_30} for the main example, but now for a collection of 20 replicate trees.
-For each setting (true generative, true best candidate, twin generative and twin best candidate), the resulting errors from each replicate pipeline have been merged into a single distribution.}
+  \caption{
+    Aggregate error distributions, similar to Fig. \ref{fig:example_30} 
+    for the main example, but now for a collection of 20 replicate trees. 
+    For each setting (true generative, true best candidate, twin generative 
+    and twin best candidate), the resulting errors from each replicate 
+    pipeline have been merged into a single distribution.
+  }
   \label{fig:replicate_trees}

### 2020-05-29 pirouette_article.tex
1 hour of work

@@
 (constant-rate) birth-death [\cite{nee1994reconstructed}] models.
-These simple tree priors are the most commonly used, as they represent some biologically realistic processes (i.e. branching at per species constant rates), while being computationally fast.
-BEAST's successor, BEAST2 [\cite{bouckaert2019beast}],
-has a package manager, that allows third-party users 
-to extend existing functionalities.
+These simple tree priors are the most commonly used, 
+as they represent some biologically realistic processes (i.e. 
+branching at per species constant rates), while being computationally fast.
+
+\new{
+  % 00002
+  In this paper, we focus on BD models, instead of coalescent models [Explain
+  these a bit],
+  although these are just as important. 
+}
+
+\new{
+  % 00001
+  To allow users to extend the functionalities of BEAST
+  by using plug-ins, BEAST2 was written [\cite{bouckaert2019beast}]
+  and both tools are still developed independently.
+}
+
 For example, one can add novel diversification models 
@@
 its performance in inference should be tested.
-Part of a model's performance is its ability to recover parameters from simulated data with known 
+Part of a model's performance is its ability to 
+recover parameters from simulated data with known 
 parameters (e.g. [\cite{etienne2014estimating}]), 
@@
 The method starts with a phylogeny generated by the new model. 
-Next, nucleotide sequences are simulated that follow the evolutionary history of the given phylogeny. 
+Next, nucleotide sequences are simulated that follow the evolutionary 
+history of the given phylogeny. 
 Then, using BEAST2's built-in tree priors,
@@
 superficially and confirmed this prediction (see the supplementary materials). 
-However, we note that for longer DNA sequences, the assumption of constant substitution rates may become less realistic and hence longer sequences may require more parameters.
-Hence, simply getting longer sequences will not always lead to a drastic reduction of the influence of the species tree prior.
+However, we note that for longer DNA sequences, the assumption 
+of constant substitution rates may become less realistic 
+and hence longer sequences may require more parameters. 
+Hence, simply getting longer sequences will not always lead to a drastic 
+reduction of the influence of the species tree prior.
 Fortunately, \verb;pirouette; provides a pipeline that works for all choices.
@@
 
-However, we note that the \verb;pirouette; pipeline is not restricted to exploring the effects of a new species tree model.
-The pipeline can also be used to explore the effects of non-standard clock or site models, such as relaxed clock models with a non-standard distribution, correlated substitutions on sister lineages, or elevated substitutions rates during speciation events.
-It is, however, beyond the scope of this paper to discuss all these options in more detail.
+However, we note that the \verb;pirouette; pipeline is not restricted 
+to exploring the effects of a new species tree model. 
+The pipeline can also be used to explore the effects of non-standard 
+clock or site models, such as relaxed clock models with a non-standard 
+distribution, correlated substitutions on sister lineages, or elevated 
+substitutions rates during speciation events. 
+It is, however, beyond the scope of this paper to discuss all these options 
+in more detail.
 

### 2020-05-29 pirouette_article.tex
Complete caption of table 1

@@
     JC = Jukes and Cantor [\cite{jukes1969evolution}], 
-    TN = Tamura and Nei [\cite{tamura1993estimation}].
+    TN = Tamura and Nei [\cite{tamura1993estimation}],
+    nLTT = normalized lineages-through-time [\cite{janzen2015approximate}],
+    $|\gamma|$ = absolute value of the gamma statistic [\cite{pybus2000testing}].
   }

### 2020-05-29 pirouette_article.tex
2nd hour

@@
 from character data (usually nucleotide sequences) extracted
-from extant (but also extinct, where possible) organisms, and hence in our understanding of the main drivers 
-and modes of diversification.
+% 00004
+from organisms (usually extant, but extinction events 
+and/or time-stamped data can be added), 
+and hence in our understanding of the main drivers 
+and modes of diversification \richel{Reword this awkwardly long sentence}.
 
@@
 Despite such a rich theoretical landscape, only a few of these 
-diversification models are available as tree priors in BEAST2.
+diversification models are available as tree priors in BEAST2
+% 00005
+\richel{
+  This is awkward: most of these models are available as plug-ins,
+  which was the whole idea when making BEAST2. This is a great moment
+  to mention the tree models available in those BEAST2 packages
+}
+.
 
@@
 \end{align}
-The default option for the diversification model $p_T$ is the standard BD model.
+
+\new{
+  % 00003
+  The default option for the twin diversification model $p_T$ 
+  is to use the standard BD model.
+}
 It is then possible to use the likelihood function 

### 2020-06-01 edit.sh
Add 'edit.sh', text in 'subsec:distribution' is 80 chars per line max

@@ -1,0 +1,10 @@
+#!/bin/bash
+#
+# Edit the article.
+#
+# Usage:
+#
+#   ./edit.sh
+#
+#
+mousepad pirouette_article.bib pirouette_article.tex pirouette_supplement.tex

### 2020-06-01 pirouette_supplement.tex
Add 'edit.sh', text in 'subsec:distribution' is 80 chars per line max
Interpret section 8.14, progress #98

@@
 \new{
-The resulting error distributions are showed in Fig.
-\ref{fig:replicate_trees}.
-We present results for cases where (1) the generative model has been used or (2) the model with highest evidence has been selected for the inference.
-From the plots we can see that in both cases there is a clear distinction between the two distributions (true and twin), suggesting that the inference models that have been used cannot capture in an accurate way the features of the diversity-dependent tree prior used to simulate the original trees.
+  The resulting error distributions are shown in Fig. \ref{fig:replicate_trees}. 
+  We present results for cases where (1) the generative model has been used 
+  or (2) the model with highest evidence has been selected for the inference.
+  From the plots we can see that in both cases there is a clear distinction 
+  between the two distributions (true and twin), suggesting that the inference 
+  models that have been used cannot capture in an accurate way the features of 
+  the diversity-dependent tree prior used to simulate the original trees. 
 }
@@
 \new{
-Comparing Fig.
-\ref{fig:example_random_mutations} to  \label{fig:replicate_trees} we can see that the extent of the error tend to increase both for the true and the twin distributions.
-This is probably due to the fact that
+  Comparing Fig. \ref{fig:example_random_mutations} to 
+  \label{fig:replicate_trees} we can see that the extent of the error 
+  tend to increase both for the true and the twin distributions. This is 
+  probably due to the fact that 
 }
@@
 Here, we show the same results for different mutation rates.
+
+\new{
+  We can observe that the error increases for mutation rates above
+  `1.0 / crown age`. Below this value, the errors are the same.
+  For these lower mutation rates, however, the shape of the error
+  distribution changes for the baseline/generative models,
+  deviating more and more from the gamma-distribution-like shape
+  as is stil present in the twin distributions.
+}
 

### 2020-06-01 compare_mut_rates/create_compare_mut_rates_fig.R

@@ -67,4 +67,4 @@ ggplot(df, aes(x = mutation_rate, y = value, col = name)) +
   geom_smooth(method = "loess") +
   coord_cartesian(ylim = c(0.0, 0.1)) +
   facet_grid(cols = vars(name)) +
-  ggsave("compare_mut_rates_graphs.png", width = 7, height = 7)
+  ggsave("compare_mut_rates.png", width = 7, height = 7)

### 2020-06-08 pirouette_supplement.tex
Changes @Giappo

--- 2020-05-29 pirouette_supplement.tex
@@
     effect when the twin alignment is allowed to have a different
-    number of mutations
+    number of \new{substitution}s
   }
@@
   \item{
-    When we do not impose the same number of mutations between true and twin alignment, we observe a difference in the error distributions with respect to the standard case (presented in the main text) where they are forced to have the same number of mutations.
+    When we do not impose the same number of \new{substitution}s between true and twin alignment, we observe a difference in the error distributions with respect to the standard case (presented in the main text) where they are forced to have the same number of \new{substitution}s.
   }
@@
 
+\clearpage
 \newpage
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
   \caption{
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
   \caption{Aggregate error distributions for 20 replicates. Here each true tree has 6 taxa. This took 16 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_32/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_32/errors.png}
   \caption{Aggregate error distributions for 10 replicates. Here each true tree has 12 taxa. This took 19 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_33/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_33/errors.png}
     \caption{Aggregate error distributions for 20 replicates. Here each true tree has 24 taxa. This took 28 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_41/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_41/errors.png}
   \caption{Aggregate error distributions for 5 replicates. Here each true tree has 32 taxa. This took 7 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_42/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_42/errors.png}
   \caption{Aggregate error distributions for 2 replicates. Here each true tree has 40 taxa. This took 8 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_19/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_19/errors.png}
   \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
   \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_34/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_34/errors.png}
   \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Effect of assuming a Yule tree prior on a Yule tree}
+\subsection{The effect of assuming a Yule tree prior on a Yule tree}
 \label{subsec:simplest_correct_parameterization}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_22/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_22/errors.png}
   \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Effect of assuming a Yule tree prior on a BD tree}
+\subsection{The effect of assuming a Yule tree prior on a BD tree}
 \label{subsec:under_parameterization}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_26/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_26/errors.png}
   \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/errors_low.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_low.png}
   \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with low likelihood.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/errors_mid.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_mid.png}
   \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with median likelihood.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_23/errors_high.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_high.png}
   \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with high likelihood.}
@@
 The main example uses a twin alignment that has the same number
-of mutations (as measured from the ancestral sequence) as the true alignment. Here, we show the same results, with the difference that
+of \new{substitution}s (as measured from the ancestral sequence) as the true alignment. Here, we show the same results, with the difference that
 the twin alignment uses the same mutation rate, yet is not guaranteed
-to have the same number of mutations. Any difference in the error distributions can therefore also stem from a difference in the amount of information contained in the alignments.
+to have the same number of \new{substitution}s.
+\iffalse
+Any difference in the error distributions can therefore also stem from a difference in the amount of information contained in the alignments.
+\fi
 
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_18/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_18/errors.png}
   \caption{Aggregate error distributions similar to Fig.
-\ref{fig:replicate_trees}, but here the number of mutations is not imposed to be the same between true and twin alignment.
+\ref{fig:replicate_trees}, but here the number of \new{substitution}s is not imposed to be the same between true and twin alignment.
 Instead we just use an equal mutation rate.
@@
 \new{
-Comparing Fig.
-\ref{fig:example_random_mutations} to  \label{fig:replicate_trees} we can see that the extent of the error tend to increase both for the true and the twin distributions.
-This is probably due to the fact that
+Comparing figures \ref{fig:example_random_mutations} and \ref{fig:replicate_trees} we can see that the discrepancy between true and twin distributions tend to increase.
+This is probably due to the fact that letting mutation rates induces a difference in the amount of information contained in the alignments and this is reflected in the error distributions.
 }
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_35/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_35/errors.png}
   \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.25 / crown age. This took 11 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_36/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_36/errors.png}
   \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.50 / crown age. This took 14 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_37/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_37/errors.png}
   \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.75 / crown age. This took 16 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_38/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_38/errors.png}
   \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.25 / crown age. This took 17 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_39/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_39/errors.png}
   \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.50 / crown age. This took 10 hours to compute.}
@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_40/errors.png}
+  \includegraphics[width=0.98\textwidth]{pirouette_example_40/errors.png}
   \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 2.0 / crown age. This took 19 hours to compute.}
 \end{figure}
+
+\giovanni{I am not able to give an interpretation to these figures.
+I would expect a trend between the difference of true and twin medians and mutation rate, but I cannot see anything.
+The fact that sometimes true is smaller and some other times twin is, makes me think that these results are probably not very reliable, probably due to a not sufficient number of replicates.}
 

### 2020-06-09 pirouette_article.tex

@@
   \textbf{3. }
+    \iffalse
     We describe \verb;pirouette;'s usage and the biological scientific
-    question it can answer, including full examples. \\
+    question it can answer, including full examples.
+    \fi
+    We describe \verb;pirouette;'s usage \new{providing full examples in which we interrogate a model for its power to describe another.
+    }
+    \\
   \textbf{4. }
@@
 (constant-rate) birth-death [\cite{nee1994reconstructed}] models.
-These simple tree priors are the most commonly used, 
+These simple tree priors are \new{among} the most commonly used, 
 as they represent some biologically realistic processes (i.e. 
@@
 are often not implemented, and simpler priors are used instead. 
-This is not necessarily problematic, when the data are very informative, 
+This is not necessarily problematic, when the data are very informative 
+\new{or when the prior is truly uninformative}, 
 as this will reduce the influence of the tree prior.
-However, the assumption that tree prior choice is of low importance 
+However, the assumption that tree prior choice is of low \new{impact} 
 must first be verified.
 
-There have been multiple attempts to investigate the importance of tree
+There have been multiple attempts to investigate the \new{impact} of tree
 prior choice. For example, Sarver and colleagues, [\cite{sarver2019choice}] 
@@
 However, they only compared current diversification models to one another, 
-and thus this does not inform us on the importance of a new tree prior.
+and thus this does not inform us on the \new{impact} of a new tree prior.
 
-Here we introduce a method to quantify the importance of a novel tree prior.
+Here we introduce a method to quantify the \new{impact} of a novel tree prior.
 The method starts with a phylogeny generated by the new model. 
@@
 
-The goal of \verb;pirouette; is to quantify the importance of a tree prior.
+The goal of \verb;pirouette; is to quantify the \new{impact} of a tree prior.
 It does so by measuring the inference error made 
@@
 as the 'generative tree model' $\mathit{p_{G}}$.
+\new{
+% 00006
+The error we aim to quantify is not of stochastic nature.
+Stochastic errors are usually non-directional.
+We, instead, aim to expose the bias due to the mismatch between generative and inference models.
+}
 We define the birth-death (BD) model [\cite{nee1994reconstructed}] as
@@
 model [\cite{DDD, etienne2012diversity}]:
-the DD model is a BD model with a speciation or extinction rate that depends 
-on the number of species and a clade-level carrying capacity. For a carrying capacity
+the DD model is a \new{diversification model} with a speciation or extinction rate that depends 
+on the number of species and a clade-level carrying capacity.
+\new{The BD model can be seen as a special case for the DD model. In fact, for} a carrying capacity
 of infinity, the DD model reduces to the BD model.
@@
     \verb;mutation_rate; &
-    Pace at which mutation occurs &
+    Pace at which \new{substitutions occur} &
     \verb;mutation_rate; $\in \mathbb{R}_{>0}$\\
@@
 the same mutation rate as the true alignment. By default, however,
-not only the same mutation rate is used, but also the total number of mutations
-matches the true alignment. The total number of mutations is defined
+not only the same mutation rate is used, but also the total number of \new{substitution}s
+matches the true alignment. The total number of \new{substitution}s is defined
 as the amount of different nucleotides between the (known) root sequence
@@
 
-We showed how to use \verb;pirouette; to quantify the importance of a 
+We showed how to use \verb;pirouette; to quantify the \new{impact} of a 
 tree prior in Bayesian phylogenetics, assuming the simplest standard 
@@
 We note, however, that only one tree from a novel tree model
-is not enough to determine the impact of using an incorrect
+is not enough to determine the \new{impact} of using an incorrect
 tree prior. Instead, a distribution 

### 2020-06-09 pirouette_supplement.tex

@@
 
+\new{
+  We can observe that the error increases for mutation rates above
+  `1.0 / crown age`. Below this value, the errors are the same.
+  For these lower mutation rates, however, the shape of the error
+  distribution changes for the baseline/generative models,
+  deviating more and more from the gamma-distribution-like shape
+  as is stil present in the twin distributions.
+}
+
 The code to reproduce this figure can be found at

### 2020-06-09 pirouette_supplement.tex
@Giappo was right: figure 26 in section 8.11 was not updated yet. Let @richelbilderbeek check all other figures, see #103

@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_26/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
+  \caption{
+    Aggregate error distributions for 20 replicates. Here each true 
+    tree is generated by a BD process. For the inference we used instead a 
+    Yule tree prior. This took 9 hours to compute.
+  }
 \end{figure}
 
-\giovanni{This is the example one of the reviewer was referring to.
-The figure that was present before (still present in the overleaf files) looked actually quite different from the one on the repo (\url{https://raw.githubusercontent.com/richelbilderbeek/pirouette_example_26/master/errors.png}).
-I changed it with the one found there.
-In case you think this is wrong you can revert it back.}
-
 \new{
-Since the two models are very similar to each other (the BD model can be turned into a Yule model just by setting the extinction parameter to zero \giovanni{Cite Louca and Pennell here?}) the median discrepancy it's almost negligible.
-However, respect to the previous case when a Yule tree prior has been considered, here the distributions exhibit a greater difference in shapes.
+  Since the two models are very similar to each 
+  other (the BD model can be turned into a Yule model just by setting the 
+  extinction parameter to zero \giovanni{Cite Louca and Pennell here?}) 
+  the median discrepancy it's almost negligible. However, respect to the 
+  previous case when a Yule tree prior has been considered, here the 
+  distributions exhibit a greater difference in shapes.
 }

### 2020-06-10 pirouette_supplement.tex
Reviewed interprations in appendix

--- 2020-06-09 pirouette_supplement.tex
@@
   or (2) the model with highest evidence has been selected for the inference.
-  From the plots we can see that in both cases there is a clear distinction 
-  between the two distributions (true and twin), suggesting that the inference 
-  models that have been used cannot capture in an accurate way the features of 
-  the diversity-dependent tree prior used to simulate the original trees. 
+  From the plots we can see that in both cases the two distributions 
+  (true and twin) are mostly overlapping, but not everywhere.
+  This suggests that the inference 
+  models that have been used can to a reasonable extent capture in an accurate way 
+  the features of the diversity-dependent tree prior used to simulate 
+  the original trees. 
 }
@@
 \new{
-We show in figures \ref{fig:example_6_taxa}, \ref{fig:example_12_taxa}, \ref{fig:example_24_taxa}, \ref{fig:example_32_taxa} and \ref{fig:example_40_taxa} what are the errors obtained when starting from phylogenies with, respectively, 6, 12, 24, 32 and 40 taxa.
-Again we can see that in each case errors tend to be greater in the true distribution than in the twin distribution, confirming the result of the previous paragraph.
-However we cannot detect a clear trend between the number of taxa of the original phylogeny and the extent of such discrepancy.
-Due to long computational times we have been able to provide data only for a limited amount of replicates.
-We cannot exclude that a clearer trend could be observed increasing the number of replicates.
+  We show in figures \ref{fig:example_6_taxa}, \ref{fig:example_12_taxa}, 
+  \ref{fig:example_24_taxa}, \ref{fig:example_32_taxa} and 
+  \ref{fig:example_40_taxa} what are the errors obtained when starting from 
+  phylogenies with, respectively, 6, 12, 24, 32 and 40 taxa. Again we can see 
+  that in each case errors tend to be greater in the true distribution than 
+  in the twin distribution, similar to the result of the previous 
+  subsection, subsection\ref{subsec:distribution}.
+  In addition to this, we cannot detect a clear trend between the number of 
+  taxa of the original phylogeny and the extent of such discrepancy. 
+  Due to long computational times we have been able to provide data only 
+  for a limited amount of replicates. 
+  We cannot exclude that a clearer trend could be observed increasing the 
+  number of replicates.
 }
@@
 \new{
-From figures \ref{fig:example_500_nucleotides}, \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} we can observe a that the discrepancy between the true and twin error distributions tend to become smaller as the number of nucleotides increase.
+  From figures \ref{fig:example_500_nucleotides}, 
+  \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} 
+  we can observe a that the discrepancy between the true and twin error 
+  distributions tend to become smaller as the number of nucleotides increase. 
   This occurred for both the generative and best candidate cases.
+  This follows the expectation that a prior becomes less important when
+  more information becomes available.
 }
-\giovanni{richel do you have an idea why this happens?
-Also, do we have the raw data stored somewhere?
-From what I can judge from the plots the median discrepancy is 0.012 for 500 nucs, 0.008 for 1000 nucs and 0.003 for 2000 nucs.}
+\giovanni{
+  Also, do we have the raw data 
+  stored somewhere? From what I can judge from the plots the median discrepancy 
+   is 0.012 for 500 nucs, 0.008 for 1000 nucs and 0.003 for 2000 nucs.
+}
+\richel{
+  Sure, raw data is at
+  \url{https://www.richelbilderbeek.nl/pirouette_example_19.zip},
+  \url{https://www.richelbilderbeek.nl/pirouette_example_28.zip}
+  and 
+  \url{https://www.richelbilderbeek.nl/pirouette_example_34.zip}.
+  These URLS can be found on the GitHub site.
+  What would you like to with the raw data?
+}
 
@@
 simplest case possible.
+
 \new{
-As expected the twin and true distributions in Fig. \ref{fig:example_yule} are extremely similar for both the generative and the best candidate case.
+  As expected the twin and true distributions in Fig. \ref{fig:example_yule} 
+  are extremely similar for both the generative and the best candidate case.
 }
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_26/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
+  \caption{Aggregate error distributions for 10 replicates. 
+  Here each true tree is generated by a BD process. 
+  For the inference we used instead a Yule tree prior. This took 
+  \richel{unknown} hours to compute.
+}
 \end{figure}
 
-\giovanni{This is the example one of the reviewer was referring to.
-The figure that was present before (still present in the overleaf files) looked actually quite different from the one on the repo (\url{https://raw.githubusercontent.com/richelbilderbeek/pirouette_example_26/master/errors.png}).
-I changed it with the one found there.
-In case you think this is wrong you can revert it back.}
-
+\giovanni{
+  This is the example one of the reviewer was referring to. 
+  The figure that was present before (still present in the overleaf files) 
+  looked actually quite different from the one on the 
+  repo (\url{https://raw.githubusercontent.com/richelbilderbeek/pirouette_example_26/master/errors.png}). 
+  I changed it with the one found there. In case you think this is wrong you can revert it back.
+}
+\richel{ 
+  Well spotted! Thanks to you, the figure is correct now
+}
 \new{
-Since the two models are very similar to each other (the BD model can be turned into a Yule model just by setting the extinction parameter to zero \giovanni{Cite Louca and Pennell here?}) the median discrepancy it's almost negligible.
-However, respect to the previous case when a Yule tree prior has been considered, here the distributions exhibit a greater difference in shapes.
+  Since the two models are very similar to each other (the BD model can be 
+  turned into a Yule model just by setting the extinction parameter to 
+  zero [\cite{nee1994reconstructed}]) the median discrepancy is
+  almost negligible. 
+  However, with respect to the previous 
+  case (subsection \ref{subsec:simplest_correct_parameterization}), 
+  where a Yule tree prior has been used, 
+  the distributions here exhibit a greater difference.
 }
@@
 \new{
-Here the median errors are similar in the three settings and similar to ones relative to the full dataset of \ref{subsec:distribution}.
-We can also notice that in the case of median likelihood, the twin median error appear to be lower than the true mean error.
-This usually is a sign that the number of replicates (in this case 10) is too low to allow us to draw precise conclusions from this test.
-We decided to not explore further in this direction using more simulations because computational times turned to be extremely high.
-The entire run took 120 hours in total.
+  Here the median errors are similar in the three settings and similar to 
+  ones relative to the full dataset of \ref{subsec:distribution}. We can also 
+  notice that in the case of median likelihood, the twin median error appear 
+  to be lower than the true mean error. This usually is a sign that the number 
+  of replicates (in this case 10) is too low to allow us to draw precise 
+  conclusions from this test. We decided to not explore further in this 
+  direction using more simulations because computational times turned to be 
+  extremely high. The entire run took 120 hours in total.
 }
@@
 to have the same number of \new{substitution}s.
-\iffalse
-Any difference in the error distributions can therefore also stem from a difference in the amount of information contained in the alignments.
-\fi
 
@@
 \new{
-  Comparing figures \ref{fig:example_random_mutations} and \ref{fig:replicate_trees} we can see that the discrepancy between true and twin distributions tend to increase.
-This is probably due to the fact that letting mutation rates induces a difference in the amount of information contained in the alignments and this is reflected in the error distributions.
+  Comparing figures \ref{fig:example_random_mutations} 
+  and \ref{fig:replicate_trees} we can see that the discrepancy between 
+  true and twin distributions tend to increase. 
+  This is probably due to the fact that letting mutation rates 
+  induces a difference in the amount of information contained in the 
+  alignments and this is reflected in the error distributions.
 }
@@
 
-\giovanni{I am not able to give an interpretation to these figures.
-I would expect a trend between the difference of true and twin medians and mutation rate, but I cannot see anything.
-The fact that sometimes true is smaller and some other times twin is, makes me think that these results are probably not very reliable, probably due to a not sufficient number of replicates.}
+\giovanni{
+  I am not able to give an interpretation to these figures. 
+  I would expect a trend between the difference of true and twin medians 
+  and mutation rate, but I cannot see anything. The fact that sometimes true 
+  is smaller and some other times twin is, makes me think that these results 
+  are probably not very reliable, probably due to a not sufficient number 
+  of replicates.
+}
+\richel{
+  I interpreted our curvy friends :-)
+}
 

### 2020-06-19 pirouette_article.tex
Updates from Overleaf

@@
 % 00005
+\giovanni{
+  PROPOSAL:
+  Replace
+  "Despite such a rich theoretical landscape, only a few of these diversification models are available as tree priors in BEAST2"
+  with
+  "Despite the fact that some models are already included as tree priors in BEAST2 (e.g.
+coalescent models, BD skyline, etc.) many others for which a likelihood formula is already available in literature, are not implemented yet"
+}
 \richel{
@@
 The user must specify a root sequence, a mutation rate and a site model.
+\iffalse
 The root sequence is the DNA sequence of the shared common ancestor of 
@@
 which helps interpreting the resulting alignment.
+\fi
+\new{ % 00008
+The root sequence is the DNA sequence of the shared common ancestor of 
+all species. By default it is set to four different equally-sized 
+mononucleotide blocks (a block of length $n$ is a sequence in which the same character is repeated $n$ times), which helps interpreting the resulting alignment.
+}
 Supported nucleotide substitution models (part of a DNA
@@
 Markov chain Monte Carlo (MCMC) run,
-which samples an unrepresentative part of paraneter and tree space. 
+which samples an unrepresentative part of parameter and tree space. 
 By default, \verb;pirouette; 
@@
 \end{align}
+
+\new{
+    % 00007
+    We choose to preserve the original topology to increase the similarity between the twin to the original tree.
+This works well in the cases of the birth-death or diversity-dependent models we consider in our examples.
+    However this might not be suited for models in which branching times are strongly influenced by topology.
+}
 
@@
 In principle, \verb;pirouette; allows for this, but in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.
+\new{
+    % 00007
+    For models that make different predictions on topology, the twinning process should be modified in line with it.
+}
 

### 2020-06-19 pirouette_supplement.tex
Updates from Overleaf

@@
   \ref{fig:example_1000_nucleotides} and \ref{fig:example_2000_nucleotides} 
-  we can observe a that the discrepancy between the true and twin error 
+  we can observe that the discrepancy between the true and twin error 
   distributions tend to become smaller as the number of nucleotides increase. 
@@
   These URLS can be found on the GitHub site.
-  What would you like to with the raw data?
+  What would you like to do with the raw data?
 }
+\giovanni{I think it might be nice to have a simple lot of error medians versus number of nucleotides.
+If we want to be a bit fancier, we can think of including 0.25 and 0.75 percentiles as error bars.}
 

### 2020-06-19 pirouette_article.tex
Accept @Giappo's suggestions, simplify example code, progress in answering response letter

@@
   \textbf{3. }
-    \iffalse
-    We describe \verb;pirouette;'s usage and the biological scientific
-    question it can answer, including full examples.
-    \fi
-    We describe \verb;pirouette;'s usage \new{providing full examples in which we interrogate a model for its power to describe another.
-    }
+    We describe \verb;pirouette;'s usage \new{providing full examples in which 
+    we interrogate a model for its power to describe another.
     \\
@@
   % 00002
-  In this paper, we focus on BD models, instead of coalescent models [Explain
-  these a bit],
-  although these are just as important. 
+  In this paper, we focus on BD models, instead of coalescent models,
+  although these are just as important. Coalescent models ... 
+  [RJCB: Describe coalescent models]
 }
@@
 % 00005
-\giovanni{
-  PROPOSAL:
-  Replace
-  "Despite such a rich theoretical landscape, only a few of these diversification models are available as tree priors in BEAST2"
-  with
-  "Despite the fact that some models are already included as tree priors in BEAST2 (e.g.
-coalescent models, BD skyline, etc.) many others for which a likelihood formula is already available in literature, are not implemented yet"
+\new{
+  Despite the fact that some models are already included as tree priors 
+  in BEAST2 (e.g. coalescent models, BD skyline, etc.) 
+  many others for which a likelihood formula is already available 
+  in literature, are not implemented yet.
 }
-\richel{
-  This is awkward: most of these models are available as plug-ins,
-  which was the whole idea when making BEAST2. This is a great moment
-  to mention the tree models available in those BEAST2 packages
+\new{
+  BEAST2 allows to extend its functionality by plug-ins, such
+  as [RJCB: name some plug-ins with refs to literature].
 }
-.
 
@@
   % 00007
-    We choose to preserve the original topology to increase the similarity between the twin to the original tree.
-This works well in the cases of the birth-death or diversity-dependent models we consider in our examples.
-    However this might not be suited for models in which branching times are strongly influenced by topology.
+  We choose to preserve the original topology to 
+  increase the similarity between the twin to the original tree. 
+  This works well in the cases of the birth-death or diversity-dependent models 
+  we consider in our examples.
+  However, this might not be suited for models in which branching times 
+  are strongly influenced by topology.
 }
@@
 }
+\new{
+  % 00009
+  \verb;pirouette; has a built-in function to use a Yule model as well.
+  Additionally, a user can specify a function to generate a twin tree
+  from any speciation model, such as, for example, a coalescent model.
+}
+
 It is then possible to use the likelihood function 
@@
 library(ggplot2)
-pir_outs <- list()
-pir_outs[[1]] <- pir_out
-pir_plots(pir_outs) +
+pir_plot(pir_out) +
   ggtitle("pirouette example 30") +

### 2020-06-19 pirouette_article.tex
Process feedback

@@
   Despite the fact that some models are already included as tree priors 
-  in BEAST2 (e.g. coalescent models, BD skyline, etc.) 
+  in BEAST2 (e.g. birth-death models, coalescent models, coalescent skyline, etc.) 
   many others for which a likelihood formula is already available 
@@
   BEAST2 allows to extend its functionality by plug-ins, such
-  as [RJCB: name some plug-ins with refs to literature].
+  as BD skyline (ref) [RJCB: name some other plug-ins with 
+  refs to literature].
 }
@@
 This is not necessarily problematic, when the data are very informative 
+% 00010
 \new{or when the prior is truly uninformative}, 
@@
 model [\cite{DDD, etienne2012diversity}]:
-the DD model is a \new{diversification model} with a speciation or extinction rate that depends 
+the DD model is a \new{diversification model} with a speciation 
+or extinction rate that depends 
 on the number of species and a clade-level carrying capacity.
@@
     error function to measure the difference between the original phylogeny 
-    and the inferred posterior. The first iterations of the MCMC chain of the posterior may not be representative and can be discarded using a burn-in fraction. \\
+    and the inferred posterior. The first iterations of the MCMC chain of the 
+    posterior may not be representative and can be discarded using a burn-in 
+    fraction. \\
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We show the usage of \verb;pirouette; on an tree generated 
-by the non-standard diversity-dependent (DD) tree model [\citep{DDD, etienne2012diversity}],
+by the non-standard diversity-dependent (DD) tree 
+model [\citep{DDD, etienne2012diversity}],
 which is a BD model with a speciation rate that is dependent on the number of species.
@@
 \begin{lstlisting}[language=R]
+library(ggplot2)
 library(pirouette)
@@
 # Create phylogeny
-set.seed(314)
 phylogeny <- create_exemplary_dd_tree(
@@
 
-# Setup pirouette
-pir_params <- create_std_pir_params(
-  folder_name = "example_30"
-)
+# Use standard pirouette setup
+pir_params <- create_std_pir_params()
 
@@
 
+# Plot
+pir_plot(pir_out)
+
 # Save
-library(ggplot2)
-pir_plot(pir_out) +
-  ggtitle("pirouette example 30") +
-  ggsave("errors.png", width = 7, height = 7)
-
 pir_save(
@@
 The DD tree that is generated by this code is as shown in Figure \ref{fig:dd_tree},
-which has an arbitrarily chosen crown age of 10 time units and 6 tips for an extinction rate of 0.1.
+which has an arbitrarily chosen crown age of 10 time units and 6 tips 
+for an extinction rate of 0.1.
 The carrying-capacity is also set at 6.

### 2020-06-19 pirouette_supplement.tex
Accept @Giappo's suggestions, simplify example code, progress in answering response letter
Process feedback

@@
   \item{
-    When we do not impose the same number of \new{substitution}s between true and twin alignment, we observe a difference in the error distributions with respect to the standard case (presented in the main text) where they are forced to have the same number of \new{substitution}s.
+    When we do not impose the same number of \new{substitution}s 
+    between true and twin alignment, we observe a difference in the error 
+    distributions with respect to the standard case (presented in the main 
+    text) where they are forced to have the same number of \new{substitution}s.
   }
@@
 
-\giovanni{
-  This is the example one of the reviewer was referring to. 
-  The figure that was present before (still present in the overleaf files) 
-  looked actually quite different from the one on the 
-  repo (\url{https://raw.githubusercontent.com/richelbilderbeek/pirouette_example_26/master/errors.png}). 
-  I changed it with the one found there. In case you think this is wrong you can revert it back.
-}
-\richel{ 
-  Well spotted! Thanks to you, the figure is correct now
-}
 \new{
@@
   the distributions here exhibit a greater difference.
+}
+\new{
+  As we use only extant trees, it is reasonable that the method
+  is slightly weaker in distinguishing between the Yule and BD
+  models. It is unknown what the discriminatory power would be
+  when comparing trees with extinction events.
 }
@@
 
-\giovanni{
-  I am not able to give an interpretation to these figures. 
-  I would expect a trend between the difference of true and twin medians 
-  and mutation rate, but I cannot see anything. The fact that sometimes true 
-  is smaller and some other times twin is, makes me think that these results 
-  are probably not very reliable, probably due to a not sufficient number 
-  of replicates.
-}
-\richel{
-  I interpreted our curvy friends :-)
-}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-06-19 pirouette_article.tex
Can build PDF

@@
     We describe \verb;pirouette;'s usage \new{providing full examples in which 
-    we interrogate a model for its power to describe another.
-    \\
+    we interrogate a model for its power to describe another}. \\
   \textbf{4. }
@@
   % 00009
-  \verb;pirouette; has a built-in function to use a Yule model as well.
+  pirouette has a built-in function to use a Yule model as well.
   Additionally, a user can specify a function to generate a twin tree

### 2020-06-19 create_article.sh

@@ -10,7 +10,7 @@
 
 pdflatex pirouette_header --shell-escape
 bibtex pirouette_header
-pdflatex pirouette_header --shell-escape > /dev/null
+pdflatex pirouette_header --shell-escape
 pdflatex pirouette_header --shell-escape
 mv pirouette_header.pdf pirouette_article.pdf
 

### 2020-06-24 pirouette_article.tex
Updates from Overleaf

@@
 on the number of species and a clade-level carrying capacity.
-\new{The BD model can be seen as a special case for the DD model. In fact, for} a carrying capacity
+\new{ %00012
+The BD model can be seen as a special case for the DD model. In fact, for} a carrying capacity
 of infinity, the DD model reduces to the BD model.
@@
 package [\cite{TESS, hohna2016tess}].
+\new{ %00013
+For simplicity, when simulating phylogenies we assumed a sampling fraction of $100\%$. A different choice might have an effect on model performance.
+}
 
@@
 
-We show the usage of \verb;pirouette; on an tree generated 
+We show the usage of \verb;pirouette; on a tree generated 
 by the non-standard diversity-dependent (DD) tree 
@@
 which is a BD model with a speciation rate that is dependent on the number of species. 
+\new{ %00011
+The example we propose here has only an instructional purpose.
+To perform a full analysis we suggest to repeat the same procedure for at least 100 independent true and twin trees (we show the outcome of an analysis involving replicates in the supplementary material at subsection \ref{subsec:distribution}).
+}
 

### 2020-06-24 pirouette_article.tex
Process response letter

@@
   In this paper, we focus on BD models, instead of coalescent models,
-  although these are just as important. Coalescent models ... 
-  [RJCB: Describe coalescent models]
+  although these are just as important.
 }
@@
 }
+% 00014
 \new{
   BEAST2 allows to extend its functionality by plug-ins, such
-  as BD skyline (ref) [RJCB: name some other plug-ins with 
-  refs to literature].
+  as the calibrated Yule model (\cite{heled2015calibrated}),
+  BD with incomplete sampling (\cite{stadler2009incomplete}),
+  BD serial sampling (\cite{stadler2012estimating}),
+  BD serial skyline (\cite{stadler2013birth})
+  Fossilized BD process (\cite{gavryushkina2014bayesian}),
+  BD SIR (\cite{kuhnert2014simultaneous}) and many more.
 }

### 2020-06-24 pirouette_article.tex
4 hours of time, @Giappo's turn :+1:

@@
 % 00010
-\new{or when the prior is truly uninformative}, 
+\new{
+  or when the prior is truly uninformative
+}, 
 as this will reduce the influence of the tree prior.
@@
 
-Here we introduce a method to quantify the \new{impact} of a novel tree prior.
+\new{
+  A recent approach to test the impact of tree prior choice, is developed
+  by \cite{duchene2018phylodynamic}, which allows to measure
+  a phylodynamic model's adequacy for models that are
+  mathematically described (i.e. have a known likelihood equation)
+}
+
+Here we introduce a method to quantify the \new{impact} 
+of a novel tree prior,
+% 00016
+\new{
+  that is, a tree model that can simulate phylogenies, but not yet
+  calculate their likelihoods.
+}
+% 00017
+\new{
+  This new method simultaneously assesses the substitution, clock and tree
+  models and thus is a hierarchical model.
+  We learn from \cite{duchene2015evaluating} that 
+  such a method may be sensitive to other aspects of the hierarchical model.
+}
 The method starts with a phylogeny generated by the new model. 
@@
 as the 'generative tree model' $\mathit{p_{G}}$.
+
+\new{
+  % 00015
+  A 'generative tree model', in this paper, is always a standard tree model.
+  If the actual model that generated the tree is non-standard, 
+  the 'generative tree model' is the (standard) tree model that is (assumed
+  to be) closest to the actual (non-standard) model. In that case, the
+  'generative tree model' can also be called 'baseline model' 
+  or 'comparison model'.
+}
+
 \new{
@@
   Stochastic errors are usually non-directional.
-  We, instead, aim to expose the bias due to the mismatch between generative and inference models.
+  We, instead, aim to expose the bias due to the mismatch between 
+  a generative model (that has either generated the phyogeny, or
+  is assumed to have do so) and the model(s) used in the actual inference.
 }
@@
     using clock model $\mathit{c_{G}}$ and site model $\mathit{s_{G}}$ 
-    imported from the generative model. 
+    imported from the generative model (that is, the model that has
+    actually generated the phylogeny, are is assumed to have done so). 
     The twin pipeline follows the procedure of the main pipeline, 
@@
     inference conditions $\mathit{C}$, 
-    an inference model $\mathit{I}$ is chosen (which may differ from the 
-    generative model).
+    an inference model $\mathit{I}$ is chosen (which may or may not differ 
+    from the actual model that generated the tree).
   \item The inference model and the alignment are used 
@@
 one can specify the inference model to be the same as the generative model.
-If the tree model is unknown or non-standard - which is the primary motivation for this paper -, one can pick
-a standard inference model which is considered to be closest to the true tree model.
+If the tree model is unknown or non-standard - which is the primary 
+motivation for this paper -, one can pick
+a standard inference model which is considered 
+to be closest to the true tree model.
 Also, if we want to run only the inference
@@
 between the 'true' and 'twin' error distributions is caused only
-by the mismatch between the generative tree model and the assumed
-tree prior.
+by the mismatch between the (known-to-be-true or assumed) 
+generative tree model and the 
+tree prior used in the actual inference.
 

### 2020-06-25 pirouette_article.tex
Spellcheck

@@
   We, instead, aim to expose the bias due to the mismatch between 
-  a generative model (that has either generated the phyogeny, or
+  a generative model (that has either generated the phylogeny, or
   is assumed to have do so) and the model(s) used in the actual inference.

### 2020-06-25 pirouette_supplement.tex
Spellcheck

@@
   deviating more and more from the gamma-distribution-like shape
-  as is stil present in the twin distributions.
+  as is still present in the twin distributions.
 }

### 2020-06-25 pirouette_article.tex
Removing the location references, such as 00016

@@
 has been a major advance in constructing phylogenetic trees 
-from character data (usually nucleotide sequences) extracted
-% 00004
-from organisms (usually extant, but extinction events 
+from character data (usually nucleotide sequences) extracted from organisms
+\new{
+  (usually extant, but extinction events
   and/or time-stamped data can be added), 
+}
 and hence in our understanding of the main drivers 
-and modes of diversification \richel{Reword this awkwardly long sentence}.
+and modes of diversification.
 
@@
 (constant-rate) birth-death [\cite{nee1994reconstructed}] models.
-These simple tree priors are \new{among} the most commonly used, 
+These simple tree priors are 
+\new{
+  among
+} 
+the most commonly used, 
 as they represent some biologically realistic processes (i.e. 
 branching at per species constant rates), while being computationally fast.
-
 \new{
-  % 00002
   In this paper, we focus on BD models, instead of coalescent models,
@@
 \new{
-  % 00001
   To allow users to extend the functionalities of BEAST
@@
 }
-
 For example, one can add novel diversification models 
@@
 diversification models are available as tree priors in BEAST2
-% 00005
 \new{
@@
 }
-% 00014
 \new{
   BEAST2 allows to extend its functionality by plug-ins, such
-  as the calibrated Yule model (\cite{heled2015calibrated}),
-  BD with incomplete sampling (\cite{stadler2009incomplete}),
-  BD serial sampling (\cite{stadler2012estimating}),
-  BD serial skyline (\cite{stadler2013birth})
-  Fossilized BD process (\cite{gavryushkina2014bayesian}),
-  BD SIR (\cite{kuhnert2014simultaneous}) and many more.
+  as the calibrated Yule model [\cite{heled2015calibrated}],
+  BD with incomplete sampling [\cite{stadler2009incomplete}],
+  BD serial sampling [\cite{stadler2012estimating}],
+  BD serial skyline [\cite{stadler2013birth}]
+  Fossilized BD process [\cite{gavryushkina2014bayesian}],
+  BD SIR [\cite{kuhnert2014simultaneous}] and many more.
 }
@@
 This is not necessarily problematic, when the data are very informative 
-% 00010
 \new{
@@
 \new{
+  Another way to benchmark a diversification model, is by doing
+  a model comparison, in which the best model is determined
+  from a set of models. A good early example is \cite{goldman1993statistical} 
+  in which Goldman compared DNA substitution models.
   A recent approach to test the impact of tree prior choice, is developed
@@
   a phylodynamic model's adequacy for models that are
-  mathematically described (i.e. have a known likelihood equation)
+  mathematically described (i.e. have a known likelihood equation).
 }
@@
 of a novel tree prior,
-% 00016
 \new{
@@
 }
-% 00017
 \new{
@@
 \new{
-  % 00015
   A 'generative tree model', in this paper, is always a standard tree model.
@@
 \new{
-  % 00006
   The error we aim to quantify is not of stochastic nature.
@@
 on the number of species and a clade-level carrying capacity.
-\new{ %00012
-The BD model can be seen as a special case for the DD model. In fact, for} a carrying capacity
+\new{
+  The BD model can be seen as a special case of the DD model. 
+  In fact, for
+} 
+a carrying capacity
 of infinity, the DD model reduces to the BD model.
@@
 The user must specify a root sequence, a mutation rate and a site model.
-\iffalse
-The root sequence is the DNA sequence of the shared common ancestor of 
-all species, and is set to four different equally-sized 
-mononucleotide blocks by default, 
-which helps interpreting the resulting alignment.
-\fi
-\new{ % 00008
+\new{
   The root sequence is the DNA sequence of the shared common ancestor of 
   all species. By default it is set to four different equally-sized 
-mononucleotide blocks (a block of length $n$ is a sequence in which the same character is repeated $n$ times), which helps interpreting the resulting alignment.
+  mononucleotide blocks (a block of length $n$ is a sequence in which the 
+  same character is repeated $n$ times), which helps interpreting the 
+  resulting alignment.
 }
@@
 from a distribution of trees from any tree model.
+new{
+  The nLTT statistic is unsuitable, however, to distinguish
+  between a constant-rate BD model and a family of time-dependent 
+  models [\cite{louca2020extant}].
+}
 
@@
 \new{
-  % 00007
   We choose to preserve the original topology to 
@@
 }
-
 \new{
-  % 00003
   The default option for the twin diversification model $p_T$ 
@@
 \new{
-  % 00009
   pirouette has a built-in function to use a Yule model as well.
@@
 package [\cite{TESS, hohna2016tess}].
-\new{ %00013
-For simplicity, when simulating phylogenies we assumed a sampling fraction of $100\%$. A different choice might have an effect on model performance.
+\new{
+  For simplicity, when simulating phylogenies we assumed a sampling fraction 
+  of $100\%$. A different choice might have an effect on model performance.
 }
@@
 which is a BD model with a speciation rate that is dependent on the number of species. 
-\new{ %00011
+\new{
   The example we propose here has only an instructional purpose. 
-To perform a full analysis we suggest to repeat the same procedure for at least 100 independent true and twin trees (we show the outcome of an analysis involving replicates in the supplementary material at subsection \ref{subsec:distribution}).
+  To perform a full analysis we suggest to repeat the same procedure 
+  for at least 100 independent true and twin trees (we show the outcome of 
+  an analysis involving replicates in the supplementary material at 
+  subsection \ref{subsec:distribution}).
 }
+\richel{Check if we really do so}
 
@@
 \new{
-    % 00007
-    For models that make different predictions on topology, the twinning process should be modified in line with it.
+  For models that make different predictions on topology, 
+  the twinning process should be modified in line with it.
 }
 
-As noted in the introduction, Duch\^{e}ne and colleagues [\cite{duchene2018phylodynamic}],
+As noted in the introduction, Duch\^{e}ne and 
+colleagues [\cite{duchene2018phylodynamic}],
 also developed a method to assess the adequacy of a tree model
-on empirical trees.
-They simulate trees from the posterior distribution of the parameters and then compare this to the originally inferred tree using tree statistics, to determine whether the assumed tree model in inference indeed generates the tree as inferred.
-This is useful if these trees match, but when they do not, this does not mean that the inferred tree is incorrect; if sufficient data is available the species tree prior may not be important, and hence the inference may be adequate even though the assumed species tree prior is not.
-In short, the approach is applied to empirical trees and compares the posterior and prior distribution of trees (with the latter generated with the posterior parameters!).
-\verb;pirouette; aims to identify when assuming standard priors for the species tree leads to incorrect inference if one believes more complex diversification models are operating than can be currently accommodated in inference.
-In short, our approach applies to simulated trees and compares the posterior distributions of trees generated with a standard and non-standard model, but inferred with a standard one.
+on empirical trees. They simulate trees from the posterior distribution of 
+the parameters and then compare this to the originally inferred tree using 
+tree statistics, to determine whether the assumed tree model in inference 
+indeed generates the tree as inferred. This is useful if these trees match, 
+but when they do not, this does not mean that the inferred tree is incorrect; 
+if sufficient data is available the species tree prior may not be important, 
+and hence the inference may be adequate 
+even though the assumed species tree prior is not. 
+In short, the approach is applied to empirical trees and 
+compares the posterior and prior distribution of trees (with the latter 
+generated with the posterior parameters!).
+\verb;pirouette; aims to identify when assuming standard priors for the 
+species tree leads to incorrect inference if one believes more complex 
+diversification models are operating than can be currently accommodated in 
+inference. In short, our approach applies to simulated trees and compares 
+the posterior distributions of trees generated with 
+a standard and non-standard model, but inferred with a standard one. 
 The two methods therefore complement one another.

### 2020-06-26 pirouette_article.tex
Only two things left
Changes @Giappo

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\new{
 The development of new powerful Bayesian phylogenetic inference tools, 
@@
 from character   data (usually nucleotide sequences) extracted from organisms
-\new{
-  (usually extant, but extinction events
-  and/or time-stamped data can be added), 
-}
+(usually extant, but extinction events and/or time-stamped data can also be added), 
 and hence in our understanding of the main drivers 
@@
 BEAST has built-in priors such as the Yule [\cite{yule}] and 
-(constant-rate) birth-death [\cite{nee1994reconstructed}] models.
-These simple tree priors are 
-\new{
-  among
-} 
-the most commonly used, 
-as they represent some biologically realistic processes (i.e. 
-branching at per species constant rates), while being computationally fast.
-\new{
-  In this paper, we focus on BD models, instead of coalescent models,
-  although these are just as important.
-}
+(constant-rate) birth-death [\cite{nee1994reconstructed}] models
+\new{as well as coalescent priors.}
+These simple tree priors are among the most commonly used, 
+as they represent some biologically realistic processes (e.g. 
+viewing diversification as a branching process), while being computationally fast.
 
-\new{
 To allow users to extend the functionalities of BEAST
-  by using plug-ins, BEAST2 was written [\cite{bouckaert2019beast}]
-  and both tools are still developed independently.
-}
+using plug-ins, BEAST2 was written [\cite{bouckaert2019beast}]
+(with BEAST and BEAST2 still independently being developed further).
 For example, one can add novel diversification models 
-by writing a BEAST2 package that contains the likelihood 
-formula of a phylogeny under the novel diversification model, 
+by writing a BEAST2 plugin that contains the likelihood formula of a phylogeny under the novel diversification model, 
 i.e. the prior probability of a species tree.
-Many diversification models (and their associated probability algorithms) 
+Plugins have been provided, for instance, for the calibrated Yule model [\cite{heled2015calibrated}],
+  BD with incomplete sampling [\cite{stadler2009incomplete}],
+  BD serial sampling [\cite{stadler2012estimating}],
+  BD serial skyline [\cite{stadler2013birth}]
+  Fossilized BD process [\cite{gavryushkina2014bayesian}], and
+  BD SIR [\cite{kuhnert2014simultaneous}].
+Many other diversification models (and their associated likelihood algorithms) 
 have been developed, e.g., models in which diversification is time-dependent [\cite{nee1994reconstructed,rabosky2008explosive}],
@@
 or where diversification rates change for specific lineages 
-and their descendants [\cite{etienne2012conceptual, rabosky2014automatic, alfaro2009nine}].
+and their descendants [\cite{etienne2012conceptual, rabosky2014automatic, alfaro2009nine, laudanno2020sls}].
 Other models treat speciation as a process that takes 
 time [\cite{rosindell2010protracted, etienne2012prolonging, lambert2015reconstructed}],
-or where diversification rate
-depends on one or more traits [\cite{maddison2007estimating, fitzjohn2012diversitree}].
-Despite such a rich theoretical landscape, only a few of these 
-diversification models are available as tree priors in BEAST2
-\new{
-  Despite the fact that some models are already included as tree priors 
-  in BEAST2 (e.g. birth-death models, coalescent models, coalescent skyline, etc.) 
-  many others for which a likelihood formula is already available 
-  in literature, are not implemented yet.
+or where diversification rates
+depends on one or more traits [\cite{maddison2007estimating, fitzjohn2012diversitree, herrera2018detecting} ].
+These are, however, not yet available as tree priors in BEAST2, for reasons explained below.
+In this paper, we present methodology to determine whether such new plug-ins are needed, or whether currently available plug-ins are sufficient.
+We show this using the Yule and BD species tree priors, but our methods can be used with other built-in tree priors as well.
+
+The rationale of our paper is as follows. 
 }
-\new{
-  BEAST2 allows to extend its functionality by plug-ins, such
-  as the calibrated Yule model [\cite{heled2015calibrated}],
-  BD with incomplete sampling [\cite{stadler2009incomplete}],
-  BD serial sampling [\cite{stadler2012estimating}],
-  BD serial skyline [\cite{stadler2013birth}]
-  Fossilized BD process [\cite{gavryushkina2014bayesian}],
-  BD SIR [\cite{kuhnert2014simultaneous}] and many more.
-}
-
 When a novel diversification model is introduced,
@@
 where ideally the estimated parameter values closely match the known/true values.
-Even when a diversification model passes this procedure, 
+Even when a diversification model passes this test, 
 it is not necessarily used as tree prior in Bayesian inference.
@@
   in which Goldman compared DNA substitution models.
-  A recent approach to test the impact of tree prior choice, is developed
-  by \cite{duchene2018phylodynamic}, which allows to measure
-  a phylodynamic model's adequacy for models that are
-  mathematically described (i.e. have a known likelihood equation).
+  A recent approach to test the impact of tree prior choice, proposed
+  by \cite{duchene2018phylodynamic}, allows to measure
+  model adequacy for phylodynamic models that are
+  mathematically described (i.e. having a known likelihood equation).
 }
@@
 \new{
-  that is, a tree model that can simulate phylogenies, but not yet
+  i.e., a tree model, for which we can simulate phylogenies, but not yet
   calculate their likelihoods.
@@
 
-The goal of \verb;pirouette; is to quantify the \new{impact} of a tree prior.
+The goal of \verb;pirouette; is to quantify the \new{impact of a new} tree prior.
 It does so by measuring the inference error made 
@@
 as the 'generative tree model' $\mathit{p_{G}}$.
-
+\rampal{This only increases the confusion; maybe replace by baseline model everywhere}
 \new{
@@
 the standard tree model, as many (non-standard) tree models 
-have a parameter setting for which they reduce to this model. 
+have a parameter setting that can be reduced to this model. 
 One such example is the diversity-dependent (DD) diversification 
-model [\cite{DDD, etienne2012diversity}]:
-the DD model is a \new{diversification model} with a speciation 
-or extinction rate that depends 
+model [\cite{DDD, etienne2012diversity}] in which speciation 
+or extinction rate depends 
 on the number of species and a clade-level carrying capacity.
@@
 } 
-a carrying capacity
-of infinity, the DD model reduces to the BD model.
+an infinite carrying capacity, the DD model reduces to the BD model.
 When benchmarking a novel tree model, 
@@
 when assessing a novel tree prior, 
-our examples contain only one replicate,
+our example contain only one replicate,
 as the goal is to show the workings of \verb;pirouette;,
 instead of doing an extensive analysis.
-The supplementary material, however, does show the results
-of replicated runs under multiple settings.
+Supplementary material includes results of replicated runs under multiple settings.
 
@@
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
+    $\mathit{s_{G}}$ &
+    Site model &
+    Both the substitution model and rate variation across sites. \\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     $\mathit{A}$ &
@@
     alignment model 
-    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed of a clock model and a site model. 
+    $\mathit{A} = (\mathit{c_{G}}, \mathit{s_{G}})$, composed of a clock model 
+    and a site model. 
     The user defines one or more experiments.
@@
     using clock model $\mathit{c_{G}}$ and site model $\mathit{s_{G}}$ 
-    imported from the generative model (that is, the model that has
-    actually generated the phylogeny, are is assumed to have done so). 
+    imported from the generative model (the model that has
+    actually generated the phylogeny, or is assumed to have done so). 
     The twin pipeline follows the procedure of the main pipeline, 
@@
 There is also the option to generate a 'twin tree', 
-that goes through the same pipeline. 
-The utility of this twin tree will be explained below.
+that goes through the same pipeline (see section \ref{subsec:twinning}).
 
@@
 For the sake of clarity, here we will assume the alignment consists
-of DNA sequences, but one can also use other heritable material such as amino acids.
+of DNA sequences, but one can also use other heritable materials such as amino acids.
 The user must specify a root sequence, a mutation rate and a site model.
@@
 the meaning of these abbreviations).
-JC is the default nucleotide substitution model (NSM),
+JC is the default nucleotide substitution model,
 in which the nucleotide substitution rates between all
@@
 one can specify these inference models as 
-well (see section 'Candidate models').
+well (see section \ref{subsec:candidates}).
 
@@
 from a distribution of trees from any tree model.
-new{
+\new{
   The nLTT statistic is unsuitable, however, to distinguish
@@
   increase the similarity between the twin to the original tree. 
-  This works well in the cases of the birth-death or diversity-dependent models 
-  we consider in our examples.
-  However, this might not be suited for models in which branching times 
-  are strongly influenced by topology.
+  This works well in the cases of BD or DD models 
+  we consider in our example, because all these models make the same
+  assumption about topology (all topologies are equally likely).
+  However, this might not be suited for new models that assign different probabilities
+  to trees with the same branching times but different topologies.
 }
@@
   The default option for the twin diversification model $p_T$ 
-  is to use the standard BD model.
+  is the standard BD model.
 }
 \new{
-  pirouette has a built-in function to use a Yule model as well.
+  \verb;pirouette; has a built-in function to use a Yule model as well.
   Additionally, a user can specify a function to generate a twin tree
@@
 A single phylogeny cannot be considered as fully representative of the model. 
-For this reason multiple phylogenies must be considered. 
+For this reason multiple phylogenies must be considered \new{(at least 100 independent true and twin trees)}.
 If the number of considered phylogenies is high enough, 
@@
 model [\citep{DDD, etienne2012diversity}],
-which is a BD model with a speciation rate that is dependent on the number of species. 
-\new{
-  The example we propose here has only an instructional purpose. 
-  To perform a full analysis we suggest to repeat the same procedure 
-  for at least 100 independent true and twin trees (we show the outcome of 
-  an analysis involving replicates in the supplementary material at 
-  subsection \ref{subsec:distribution}).
-}
-\richel{Check if we really do so}
+which is a BD model with a speciation rate that depends on the number of species.
 
@@
 
+\giovanni{Do we need the library calls in the listing?}
 \begin{lstlisting}[language=R]
@@
 
-The DD tree that is generated by this code is as shown in Figure \ref{fig:dd_tree},
-which has an arbitrarily chosen crown age of 10 time units and 6 tips 
+The DD tree generated by this code is shown in Figure \ref{fig:dd_tree}. \iffalse % wordcut (this info is already in listing)
+,which has an arbitrarily chosen crown age of 10 time units and 6 tips 
 for an extinction rate of 0.1.
@@
 The initial speciation rate $\lambda_0$ is chosen such that the expected number of species in a constant-rate BD model would be equal to the number of tips, which amounts to $\lambda_0 = 0.63$.
+\fi
 
@@
 Because these distributions are distinctively different,
-the inference error we make when using an 
+the inference error made when using an 
 incorrect (that is, BD) tree prior on a DD tree
@@
   For models that make different predictions on topology, 
-  the twinning process should be modified in line with it.
+  the twinning process should be modified accordingly.
 }
-
+\new{
 As noted in the introduction, Duch\^{e}ne and 
-colleagues [\cite{duchene2018phylodynamic}],
+colleagues [\cite{duchene2018phylodynamic}]
 also developed a method to assess the adequacy of a tree model
@@
 The two methods therefore complement one another.
-
+}
+\giovanni{I don't understand this sentence: "pirouette aims to identify when assuming standard priors for the 
+species tree leads to incorrect inference if one believes more complex 
+diversification models are operating than can be currently accommodated in 
+inference". Maybe shortening it, it can become clearer}
 However, we note that the \verb;pirouette; pipeline is not restricted 

### 2020-06-26 pirouette_article.tex
Find bracket error

@@
 	(constant-rate) birth-death [\cite{nee1994reconstructed}] models
-\new{as well as coalescent priors.}
+	as well as coalescent priors.
 	These simple tree priors are among the most commonly used, 
@@
 	or where diversification rates
-depends on one or more traits [\cite{maddison2007estimating, fitzjohn2012diversitree, herrera2018detecting} ].
-These are, however, not yet available as tree priors in BEAST2, for reasons explained below.
-In this paper, we present methodology to determine whether such new plug-ins are needed, or whether currently available plug-ins are sufficient.
-We show this using the Yule and BD species tree priors, but our methods can be used with other built-in tree priors as well.
+	depends on one or more traits [\cite{maddison2007estimating, fitzjohn2012diversitree, herrera2018detecting}].
+	These are, however, not yet available as tree priors in BEAST2, 
+	for reasons explained below. In this paper, 
+	we present methodology to determine whether such new plug-ins are needed, 
+	or whether currently available plug-ins are sufficient. 
+	We show this using the Yule and BD species tree priors, 
+	but our methods can be used with other built-in tree priors as well.
 
@@
 the same mutation rate as the true alignment. By default, however,
-not only the same mutation rate is used, but also the total number of \new{substitution}s
+not only the same mutation rate is used, but also the total number of 
+\new{substitution}s
 matches the true alignment. The total number of \new{substitution}s is defined
@@
 A single phylogeny cannot be considered as fully representative of the model. 
-For this reason multiple phylogenies must be considered \new{(at least 100 independent true and twin trees)}.
+For this reason multiple phylogenies must be 
+considered \new{(at least 100 independent true and twin trees)}.
 If the number of considered phylogenies is high enough, 
@@
 
-\giovanni{Do we need the library calls in the listing?}
+\giovanni{
+  Do we need the library calls in the listing?
+}
+\richel{
+  I simplified the code and only call pirouette now. 
+  I think this is cleaner then using a scope operator.
+}
 \begin{lstlisting}[language=R]
-library(ggplot2)
 library(pirouette)
@@
 pir_plot(pir_out)
-
-# Save
-pir_save(
-  phylogeny = phylogeny,
-  pir_params = pir_params,
-  pir_out = pir_out,
-  folder_name = "example_30"
-)
 \end{lstlisting}
 
-The DD tree generated by this code is shown in Figure \ref{fig:dd_tree}. \iffalse % wordcut (this info is already in listing)
-,which has an arbitrarily chosen crown age of 10 time units and 6 tips 
-for an extinction rate of 0.1.
-The carrying-capacity is also set at 6.
-The initial speciation rate $\lambda_0$ is chosen such that the expected number of species in a constant-rate BD model would be equal to the number of tips, which amounts to $\lambda_0 = 0.63$.
-\fi
+The DD tree generated by this code is shown in Figure \ref{fig:dd_tree}.
 
@@
 candidate model is only slightly better at inferring the true tree,
-than the assumed generative model, thereby showing that they cannot compensate for the true generative tree model not being among the inference models.
+than the assumed generative model, 
+thereby showing that they cannot compensate 
+for the true generative tree model not being among the inference models.
 
@@
 }
-\giovanni{I don't understand this sentence: "pirouette aims to identify when assuming standard priors for the 
+\giovanni{
+	I don't understand this sentence: "pirouette aims to identify when assuming standard priors for the 
 	species tree leads to incorrect inference if one believes more complex 
 	diversification models are operating than can be currently accommodated in 
-inference". Maybe shortening it, it can become clearer}
+	inference". Maybe shortening it, it can become clearer
+}
 However, we note that the \verb;pirouette; pipeline is not restricted 
@@
 actual generative model.
-The user can then judge whether or not this new model should be implemented in our Bayesian phylogenetic tool.
+The user can then judge whether or not this new model 
+should be implemented in our Bayesian phylogenetic tool.
 

### 2020-06-26 pirouette_article.tex
Cannot do whitespace in 'new' macro :confused:

@@
 	and modes of diversification.
+}
 
+\new{
 	BEAST [\cite{drummond2007beast}] is a typical Bayesian phylogenetics tool, 
@@
 	viewing diversification as a branching process), while being computationally fast.
+}
 
+\new{
 	To allow users to extend the functionalities of BEAST
@@
 	i.e. the prior probability of a species tree.
-	Plugins have been provided, for instance, for the calibrated Yule model [\cite{heled2015calibrated}],
+	Plugins have been provided, for instance, for the calibrated 
+	Yule model [\cite{heled2015calibrated}],
 	  BD with incomplete sampling [\cite{stadler2009incomplete}],
@@
 	  BD SIR [\cite{kuhnert2014simultaneous}].
+}
+
+\new{
 	Many other diversification models (and their associated likelihood algorithms) 
-	have been developed, e.g., models in which diversification is time-dependent [\cite{nee1994reconstructed,rabosky2008explosive}],
+	have been developed, e.g., models in which diversification 
+	is time-dependent [\cite{nee1994reconstructed,rabosky2008explosive}],
 	or diversity-dependent [\cite{etienne2012diversity}],
@@
 	or where diversification rates
-	depends on one or more traits [\cite{maddison2007estimating, fitzjohn2012diversitree, herrera2018detecting}].
+	depends on one or more traits 
+	[\cite{maddison2007estimating, fitzjohn2012diversitree, herrera2018detecting}].
+}
+
+\new{
 	These are, however, not yet available as tree priors in BEAST2, 
@@
 	but our methods can be used with other built-in tree priors as well.
+}
 
+\new{
 	The rationale of our paper is as follows. 
@@
 }
+\verb;pirouette;
 \new{
-  \verb;pirouette; has a built-in function to use a Yule model as well.
+   has a built-in function to use a Yule model as well.
   Additionally, a user can specify a function to generate a twin tree
@@
 	generated with the posterior parameters!).
-	\verb;pirouette; aims to identify when assuming standard priors for the 
+}
+\verb;pirouette;
+new{
+	aims to identify when assuming standard priors for the 
 	species tree leads to incorrect inference if one believes more complex 

### 2020-06-26 pirouette_article.tex
Process feedback Giovanni

@@
 new{
-	aims to identify when assuming standard priors for the 
-	species tree leads to incorrect inference if one believes more complex 
-	diversification models are operating than can be currently accommodated in 
-	inference. In short, our approach applies to simulated trees and compares 
+	aims to expose when assuming standard priors for the 
+	species tree are a mis- or underparameterization. 
+	Our approach applies to simulated trees and compares 
 	the posterior distributions of trees generated with 
@@
 	inference". Maybe shortening it, it can become clearer
+}
+\richel{
+	Reworded, thanks!
 }

### 2020-06-26 pirouette_article.tex
Process feedback @rsetienne

@@
 as the 'generative tree model' $\mathit{p_{G}}$.
-\rampal{This only increases the confusion; maybe replace by baseline model everywhere}
+\rampal{
+	This only increases the confusion; maybe replace by baseline model everywhere
+}
+\richel{
+	I feel the new wording below is good enough. Replacing 'generative'
+	by 'baseline' add to confustion when the generative model is the
+	generative model. Giovanni, what do you think?
+}
 \new{
@@
   If the actual model that generated the tree is non-standard, 
-  the 'generative tree model' is the (standard) tree model that is (assumed
-  to be) closest to the actual (non-standard) model. In that case, the
-  'generative tree model' can also be called 'baseline model' 
-  or 'comparison model'.
+  the 'generative tree model' is the standard tree model that is 
+  closest to the actual model. In that case, the
+  'generative tree model' serves as a baseline or comparison model.
 }

### 2020-06-26 pirouette_article.tex
Ruthlessly moved the mathematics behind the twinning process to the supplementary materials, still 3464 words

@@
 
-The twinning process, $T$, encompasses two steps:
-$T_1$, that generates a 'twin tree' (Fig.~\ref{fig:pipeline}, 1b) 
-and $T_2$, which generates a 'twin alignment' (Fig.~\ref{fig:pipeline}, 2b).
-Both twin tree and alignment will be analyzed in the same way 
-as the true tree and alignment.
-
-We define a phylogeny $\tau$ as the combination of
-branching times $\Vec{t}$ and topology $\psi$, 
-and denote as $\tau_{\mathit{G}}$ the phylogeny 
-produced by a (possibly non-standard) generative diversification model, 
-having branching times $\Vec{t}_{\mathit{G}}$ and 
-topology $\psi_{\mathit{G}}$.
-
-The first step ($T_1$) of the twinning process creates a tree $\tau_{\mathit{T}}$
-with branching times $\Vec{t}_{\mathit{T}}$ while preserving the original
-topology $\psi_{\mathit{G}}$:
-\begin{align}
-  \tau_{\mathit{G}} = (\Vec{t}_{\mathit{G}}, \psi_{\mathit{G}}) 
-  \xrightarrow[]{\mathit{T_1}} 
-  \tau_{\mathit{T}} = (\Vec{t}_{\mathit{T}}, \psi_{\mathit{G}})
-\end{align}
-
 \new{
-  We choose to preserve the original topology to 
-  increase the similarity between the twin to the original tree. 
-  This works well in the cases of BD or DD models 
-  we consider in our example, because all these models make the same
-  assumption about topology (all topologies are equally likely).
-  However, this might not be suited for new models that assign different probabilities
-  to trees with the same branching times but different topologies.
+	See the supplementary materials section \ref{subsec:twinning_mathematics}
+	for more details.
 }
-\new{
-  The default option for the twin diversification model $p_T$ 
-  is the standard BD model.
-}
-\verb;pirouette;
-\new{
-   has a built-in function to use a Yule model as well.
-  Additionally, a user can specify a function to generate a twin tree
-  from any speciation model, such as, for example, a coalescent model.
-}
-
-It is then possible to use the likelihood function 
-$L_{\mathit{T}}$ for this diversification model to find 
-the parameters $\theta^{*}_{\mathit{T}}$ 
-(e.g. speciation and extinction rates, in case of a BD model) 
-that maximize this likelihood applied 
-to the true tree, conditioned on its number of tips $n_{\mathit{G}}$:
-\begin{align}
-    \max[L_{\mathit{T}}(\theta_{\mathit{T}}|\tau_{\mathit{G}}, n_{\mathit{G}})] 
-\rightarrow \theta^{*}_{\mathit{T}}.
-\end{align}
-We use $\theta^{*}_{\mathit{T}}$ to simulate a number 
-$n_{\mathit{T}} = n_{\mathit{G}}$ 
-of branching times $\Vec{t}_{\mathit{T}}$ for the twin tree 
-$\tau_{\mathit{T}}$, under the process $p_{T}$, 
-while preserving the topology. 
-We simulate the new branching times using the TESS 
-package [\cite{TESS, hohna2016tess}].
-\new{
-  For simplicity, when simulating phylogenies we assumed a sampling fraction 
-  of $100\%$. A different choice might have an effect on model performance.
-}
-
-The second step ($T_2$) of the twinning process simulates the twin alignment 
-with the same clock model, site model and mutation rate 
-used to simulate the original alignment. 
-The twin alignment can be simulated in any user-defined way.
-\verb;pirouette; supplies the option to have it simulated with
-the same mutation rate as the true alignment. By default, however,
-not only the same mutation rate is used, but also the total number of 
-\new{substitution}s
-matches the true alignment. The total number of \new{substitution}s is defined
-as the amount of different nucleotides between the (known) root sequence
-compared to the sequences at the tips.
 

### 2020-06-26 pirouette_supplement.tex
Find bracket error
Ruthlessly moved the mathematics behind the twinning process to the supplementary materials, still 3464 words

@@
 
-This supplementary material contains additional facets of \verb;pirouette;, such as the installation of the package, an overview of
+This supplementary material contains additional facets of \verb;pirouette;, 
+such as the installation of the package, an overview of
 pirouette's main functions and a guide for users, based on multiple experiments
@@
     subsection \ref{subsec:citation}: citation of pirouette
+  }
+  \item{
+    subsection \ref{subsec:twinning_mathematics}: the mathematics of the
+    twinning process
   }
@@
 \end{lstlisting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The mathematics of the twinning process}
+\label{subsec:twinning_mathematics}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The twinning process, $T$, encompasses two steps:
+$T_1$, that generates a 'twin tree' (Fig.~\ref{fig:pipeline}, 1b) 
+and $T_2$, which generates a 'twin alignment' (Fig.~\ref{fig:pipeline}, 2b).
+Both twin tree and alignment will be analyzed in the same way 
+as the true tree and alignment.
+
+We define a phylogeny $\tau$ as the combination of
+branching times $\Vec{t}$ and topology $\psi$, 
+and denote as $\tau_{\mathit{G}}$ the phylogeny 
+produced by a (possibly non-standard) generative diversification model, 
+having branching times $\Vec{t}_{\mathit{G}}$ and 
+topology $\psi_{\mathit{G}}$.
+
+The first step ($T_1$) of the twinning process creates a tree $\tau_{\mathit{T}}$
+with branching times $\Vec{t}_{\mathit{T}}$ while preserving the original
+topology $\psi_{\mathit{G}}$:
+\begin{align}
+  \tau_{\mathit{G}} = (\Vec{t}_{\mathit{G}}, \psi_{\mathit{G}}) 
+  \xrightarrow[]{\mathit{T_1}} 
+  \tau_{\mathit{T}} = (\Vec{t}_{\mathit{T}}, \psi_{\mathit{G}})
+\end{align}
+
+\new{
+  We choose to preserve the original topology to 
+  increase the similarity between the twin to the original tree. 
+  This works well in the cases of BD or DD models 
+  we consider in our example, because all these models make the same
+  assumption about topology (all topologies are equally likely).
+  However, this might not be suited for new models that assign different probabilities
+  to trees with the same branching times but different topologies.
+}
+\new{
+  The default option for the twin diversification model $p_T$ 
+  is the standard BD model.
+}
+\verb;pirouette;
+\new{
+   has a built-in function to use a Yule model as well.
+  Additionally, a user can specify a function to generate a twin tree
+  from any speciation model, such as, for example, a coalescent model.
+}
+
+It is then possible to use the likelihood function 
+$L_{\mathit{T}}$ for this diversification model to find 
+the parameters $\theta^{*}_{\mathit{T}}$ 
+(e.g. speciation and extinction rates, in case of a BD model) 
+that maximize this likelihood applied 
+to the true tree, conditioned on its number of tips $n_{\mathit{G}}$:
+\begin{align}
+    \max[L_{\mathit{T}}(\theta_{\mathit{T}}|\tau_{\mathit{G}}, n_{\mathit{G}})] 
+\rightarrow \theta^{*}_{\mathit{T}}.
+\end{align}
+We use $\theta^{*}_{\mathit{T}}$ to simulate a number 
+$n_{\mathit{T}} = n_{\mathit{G}}$ 
+of branching times $\Vec{t}_{\mathit{T}}$ for the twin tree 
+$\tau_{\mathit{T}}$, under the process $p_{T}$, 
+while preserving the topology. 
+We simulate the new branching times using the TESS 
+package [\cite{TESS, hohna2016tess}].
+\new{
+  For simplicity, when simulating phylogenies we assumed a sampling fraction 
+  of $100\%$. A different choice might have an effect on model performance.
+}
+
+The second step ($T_2$) of the twinning process simulates the twin alignment 
+with the same clock model, site model and mutation rate 
+used to simulate the original alignment. 
+The twin alignment can be simulated in any user-defined way.
+\verb;pirouette; supplies the option to have it simulated with
+the same mutation rate as the true alignment. By default, however,
+not only the same mutation rate is used, but also the total number of 
+\new{substitution}s
+matches the true alignment. The total number of \new{substitution}s is defined
+as the amount of different nucleotides between the (known) root sequence
+compared to the sequences at the tips.
 
@@
   stored somewhere? From what I can judge from the plots the median discrepancy 
-   is 0.012 for 500 nucs, 0.008 for 1000 nucs and 0.003 for 2000 nucs.
+  is 0.012 for 500 nucleotides, 0.008 for 1000 nucleotides 
+  and 0.003 for 2000 nucleotides.
 }

### 2020-06-26 pirouette_article.tex
Ruthlessly moved 'candidate models' to supplementary materials, 3360 words left

@@
 	I feel the new wording below is good enough. Replacing 'generative'
-	by 'baseline' add to confustion when the generative model is the
+	by 'baseline' add to confusion when the generative model is the
 	generative model. Giovanni, what do you think?
@@
 
-\subsection{Twinning}\label{subsec:twinning}
+\subsection{Controls}\label{subsec:controls}
 
 Similar to lab experiments, \verb;pirouette; allows to perform
-a control measurement, by use of a process we call twinning. 
-In this context, this control results in an error distribution
-that is the baseline error of the pipeline. The difference
-between the 'true' and 'twin' error distributions is caused only
-by the mismatch between the (known-to-be-true or assumed) 
-generative tree model and the 
-tree prior used in the actual inference.
-
-\new{
-	See the supplementary materials section \ref{subsec:twinning_mathematics}
-	for more details.
-}
-
-\subsection{Candidate models}\label{subsec:candidates}
-
-The user has to specify exactly one standard inference model,
-but may be unsure which one to pick. To account for this, the user can
-specify a set of candidate inference models. Each of these candidate inference models is run in an initial, relatively short, analysis; the candidate model with the highest 
-evidence (i.e., marginal likelihood) will then be
-used in another, longer, inference run, resulting in another error distribution.
-The evidence for an inference model is estimated by nested 
-sampling [\cite{russel2019model}], using the \verb;NS; BEAST2 package.
-
-If twinning is used, a candidate model that has the highest evidence for
-the twin alignment is also used to create another error
-distribution.
+two types of control measurements. The first type of control
+is called 'twinning', which results in an error distribution
+that is the baseline error of the inference pipeline (see
+supplementary materials, section \ref{subsec:twinning} for more details).
+The second type of control is the use of candidate models,
+which results in an error distribution
+for a generative model that is determined to be the best fit to
+the tree (see supplementary materials, 
+section \ref{subsec:candidates} for more details).
 

### 2020-06-26 pirouette_supplement.tex
Ruthlessly moved 'candidate models' to supplementary materials, 3360 words left

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The mathematics of the twinning process}
-\label{subsec:twinning_mathematics}
+\subsection{The twinning process}
+\label{subsec:twinning}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Similar to lab experiments, \verb;pirouette; allows to perform
+a control measurement, by use of a process we call twinning. 
+In this context, this control results in an error distribution
+that is the baseline error of the pipeline. The difference
+between the 'true' and 'twin' error distributions is caused only
+by the mismatch between the (known-to-be-true or assumed) 
+generative tree model and the 
+tree prior used in the actual inference.
 
@@
 compared to the sequences at the tips.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Candidate models}
+\label{subsec:candidates}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The user has to specify exactly one standard inference model,
+but may be unsure which one to pick. To account for this, the user can
+specify a set of candidate inference models. Each of these candidate inference models is run in an initial, relatively short, analysis; the candidate model with the highest 
+evidence (i.e., marginal likelihood) will then be
+used in another, longer, inference run, resulting in another error distribution.
+The evidence for an inference model is estimated by nested 
+sampling [\cite{russel2019model}], using the \verb;NS; BEAST2 package.
+
+If twinning is used, a candidate model that has the highest evidence for
+the twin alignment is also used to create another error
+distribution.
 

### 2020-06-26 pirouette_article.tex
Add feedback @rsetienne in text

@@
 	generative model. Giovanni, what do you think?
+}
+\rampal{
+	This explanation again does not resolve the confusion. 
+	You have a new model with which you generate a tree and the alignment. 
+	In the pirouette paper this is DD. You could call this the generative model. 
+	You also create a twin tree using a standard tree prior, e.g. BD. 
+	Then in inference you use a tree prior, e.g. Yule. I think what you mean 
+	is that Yule is not the one used for the twin tree, 
+	and is therefore not the generative tree model. 
+	But this is far from clear as it could also refer to the DD model. 
+	Perhaps if you replace by "generative model for the twin tree", 
+	then it is fine. Note that comparison model or baseline model 
+	are NOT synonyms.
+}
+\rampal{
+	Not synonyms for the generative tree model for the twin tree, I mean.
+}
+\richel{
+	Giovanni, could you process this feedback?
 }

### 2020-06-27 pirouette_article.tex
Start processing feedback @rsetienne
Moved section on stochasticity to supplementary materials, 3270 words

--- 2020-06-26 pirouette_article.tex
@@
 \rampal{
-	This only increases the confusion; maybe replace by baseline model everywhere
-}
-\richel{
-	I feel the new wording below is good enough. Replacing 'generative'
-	by 'baseline' add to confusion when the generative model is the
-	generative model. Giovanni, what do you think?
+	Perhaps if you replace by "generative model for the twin tree", 
+	then it is fine.
 }
@@
   closest to the actual model. In that case, the
-  'generative tree model' serves as a baseline or comparison model.
+  'generative tree model' serves as a baseline model.
+  This is is done in the example, where the Yule model is the generative model.
 }
@@
 section \ref{subsec:candidates} for more details).
-
-\subsection{Stochasticity caused by simulating phylogenies}
-
-Finally, if the goal is to evaluate BEAST2's performance 
-on a non-standard tree model, 
-one must also consider the last source of stochasticity: 
-the different phylogenies a tree model generates.
-A single phylogeny cannot be considered as fully representative of the model. 
-For this reason multiple phylogenies must be 
-considered \new{(at least 100 independent true and twin trees)}.
-If the number of considered phylogenies is high enough, 
-the comparison between the main pipeline's aggregated error distribution 
-and its twin counterpart leads to a fair evaluation 
-of the new tree model with respect to the baseline error.
+Additionally, multiple \verb;pirouette; runs are needed
+to reduce the influence of stochasticity (see supplementary materials, 
+section \ref{subsec:stochasticity} for more details).
 

### 2020-06-27 pirouette_supplement.tex
Moved section on stochasticity to supplementary materials, 3270 words

@@
 distribution.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Stochasticity caused by simulating phylogenies}
+\label{subsec:stochasticity}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The goal is to evaluate BEAST2's performance 
+on a non-standard tree model, 
+one must also consider the last source of stochasticity: 
+the different phylogenies a tree model generates.
+A single phylogeny cannot be considered as fully representative of the model. 
+For this reason multiple phylogenies must be 
+considered \new{(at least 100 independent true and twin trees)}.
+If the number of considered phylogenies is high enough, 
+the comparison between the main pipeline's aggregated error distribution 
+and its twin counterpart leads to a fair evaluation 
+of the new tree model with respect to the baseline error.
 

### 2020-06-27 pirouette_article.tex
Moved nLTT discussion to supplementary materials, 3093 words

@@
 of DNA sequences, but one can also use other heritable materials such as amino acids.
-The user must specify a root sequence, a mutation rate and a site model.
-\new{
-  The root sequence is the DNA sequence of the shared common ancestor of 
-  all species. By default it is set to four different equally-sized 
-  mononucleotide blocks (a block of length $n$ is a sequence in which the 
-  same character is repeated $n$ times), which helps interpreting the 
-  resulting alignment.
-}
-Supported nucleotide substitution models (part of a DNA
-site model) are JC, HKY, TN and GTR (see Table \ref{tab:options} for
-the meaning of these abbreviations).
-JC is the default nucleotide substitution model,
-in which the nucleotide substitution rates between all
-nucleotides are equal and constant over time.
+The user must specify a root sequence (i.e. the DNA sequence of the shared 
+common ancestor of all species), a mutation rate and a site model.
 
@@
 \verb;babette; [\cite{bilderbeek2018babette}] R package which makes use of BEAST2.
-This step usually takes up most of the pipeline's computation time.
 
@@
 which samples an unrepresentative part of parameter and tree space. 
-By default, \verb;pirouette; 
-removes the first 10\% of the posterior.
 From the remaining posterior, \verb;pirouette; 
@@
 The user can specify a function to quantify the differences between
-the true and posterior trees. By default, the package uses the nLTT 
-statistic [\cite{janzen2015approximate}], which is the absolute difference
-between the normalized lineages-through-time plots of two trees.
-The nLTT statistic is chosen, as it can operate on any two trees (regardless
-of their crown ages and number of taxa) and its results have a clear range
-from zero to one. This normalized result makes it possible to compare trees 
-from a distribution of trees from any tree model.
-\new{
-  The nLTT statistic is unsuitable, however, to distinguish
-  between a constant-rate BD model and a family of time-dependent 
-  models [\cite{louca2020extant}].
-}
+the true and posterior trees.
 
@@
 
-Using the default \verb;pirouette; settings,
-the error distribution shown in Figure \ref{fig:example_30}
-is produced.
+The error distribution shown in Figure \ref{fig:example_30}
+is produced, which uses the nLTT 
+statistic [\cite{janzen2015approximate}] to compare phylogenies (see 
+section \label{subsec:nltt} for details regarding the nLTT statistic
+and its caveats).
 

### 2020-06-27 pirouette_supplement.tex
Moved nLTT discussion to supplementary materials, 3093 words

@@
 of the new tree model with respect to the baseline error.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The nLTT statistic}
+\label{subsec:nltt}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The nLTT statistic is the absolute difference
+between the normalized lineages-through-time plots of two trees.
+The nLTT statistic is chosen, as it can operate on any two trees (regardless
+of their crown ages and number of taxa) and its results have a clear range
+from zero to one. This normalized result makes it possible to compare trees 
+from a distribution of trees from any tree model.
+\new{
+  The nLTT statistic is unsuitable, however, to distinguish
+  between a constant-rate BD model and a family of time-dependent 
+  models [\cite{louca2020extant}].
+}
 

### 2020-06-27 pirouette_article.tex
Process feedback @Giappo: shortened code, only one library call

@@
     processes underlying the species tree.
-    In practice, however, only macroevolutionary models that allow for fast computation of the prior probability are used.
-    The question is how accurate the tree estimation is when the real macroevolutionary processes are substantially different from those assumed in the tree prior. \\
+    In practice, however, only macroevolutionary models that allow 
+    for fast computation of the prior probability are used.
+    The question is how accurate the tree estimation is when 
+    the real macroevolutionary processes are substantially different 
+    from those assumed in the tree prior. \\
     \textbf{2. }
@@
     a free and open-source R package that assesses 
-    the inference error made by Bayesian phylogenetics for a given macroevolutionary diversification model. \verb;pirouette; makes use of 
+    the inference error made by Bayesian phylogenetics 
+    for a given macroevolutionary diversification model. 
+    \verb;pirouette; makes use of 
     BEAST2, but its philosophy applies to any Bayesian phylogenetic inference 
@@
 
-{\bf Keywords:} Bayesian model selection, BEAST2, computational biology, evolution, phylogenetics, R, tree prior
+{\bf Keywords:} Bayesian model selection, BEAST2, computational biology, evolution, phylogenetics, R, tree prior, babette
 
@@
 	from character data (usually nucleotide sequences) extracted from organisms
-	(usually extant, but extinction events and/or time-stamped data can also be added), 
+	(usually extant, but extinction events and/or time-stamped data 
+	can also be added), 
 	and hence in our understanding of the main drivers 
@@
 	as they represent some biologically realistic processes (e.g. 
-	viewing diversification as a branching process), while being computationally fast.
+	viewing diversification as a branching process), 
+	while being computationally fast.
 }
@@
 	For example, one can add novel diversification models 
-	by writing a BEAST2 plugin that contains the likelihood formula of a phylogeny under the novel diversification model, 
+	by writing a BEAST2 plugin that contains 
+	the likelihood formula of a phylogeny under the novel diversification model, 
 	i.e. the prior probability of a species tree.
@@
 
-\giovanni{
-  Do we need the library calls in the listing?
-}
-\richel{
-  I simplified the code and only call pirouette now. 
-  I think this is cleaner then using a scope operator.
-}
 \begin{lstlisting}[language=R]

### 2020-06-27 pirouette_article.tex
Moved 'acknowledgements', 'data accessibility' and 'author contribitions' to supplementary materials. 2967 words

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgments}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We thank the Center for Information Technology of the University 
-of Groningen for its support and for providing access to the Peregrine 
-high performance computing cluster. 
-We thank the Netherlands 
-Organization for Scientific Research (NWO) for financial support 
-through a VICI grant awarded to RSE.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Data Accessibility}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-All code is archived at 
-\url{http://github.com/richelbilderbeek/pirouette_article},
-with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Author contributions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-RJCB, GL and RSE conceived the idea for the package. 
-RJCB created, tested and revised the package.
-GL provided major contributions to the package.
-RJCB wrote the first draft of the manuscript, 
-GL and RSE contributed to revisions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Bibliography

### 2020-06-27 pirouette_supplement.tex
Moved 'acknowledgements', 'data accessibility' and 'author contribitions' to supplementary materials. 2967 words

@@
     effect of different mutation rates
+  }
+  \item{
+    subsection \ref{subsec:acknowledgments}: Acknowledgments
+  }
+  \section{
+    subsection \ref{subsec:data_accessibility}: Data accessibility
+  }
+  \section{
+    subsection \ref{subsec:author_contributions}: Author contributions
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Acknowledgments}
+\label{subsec:acknowledgments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We thank the Center for Information Technology of the University 
+of Groningen for its support and for providing access to the Peregrine 
+high performance computing cluster. 
+We thank the Netherlands 
+Organization for Scientific Research (NWO) for financial support 
+through a VICI grant awarded to RSE.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Data accessibility}
+\label{subsec:data_accessibility}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+All code is archived at 
+\url{http://github.com/richelbilderbeek/pirouette_article},
+with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Author contributions}
+\label{subsec:author_contributions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+RJCB, GL and RSE conceived the idea for the package. 
+RJCB created, tested and revised the package.
+GL provided major contributions to the package.
+RJCB wrote the first draft of the manuscript, 
+GL and RSE contributed to revisions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Bibliography

### 2020-06-28 pirouette_article.tex
Changes @Giappo

@@
 \verb;pirouette;
-new{
+\new{
 	aims to expose when assuming standard priors for the 
@@
 	The two methods therefore complement one another.
-}
-\giovanni{
-	I don't understand this sentence: "pirouette aims to identify when assuming standard priors for the 
-	species tree leads to incorrect inference if one believes more complex 
-	diversification models are operating than can be currently accommodated in 
-	inference". Maybe shortening it, it can become clearer
-}
-\richel{
-	Reworded, thanks!
-}
-However, we note that the \verb;pirouette; pipeline is not restricted 
+} However, we note that the \verb;pirouette; pipeline is not restricted 
 to exploring the effects of a new species tree model. 

### 2020-06-28 pirouette_supplement.tex
Changes @Giappo

@@
   \item{
-    subsection \ref{subsec:twinning_mathematics}: the mathematics of the
+    subsection \ref{subsec:twinning}: the
     twinning process
+  }
+  \item{
+    subsection \ref{subsec:candidates}: candidate models for the inference
+  }
+  \item{
+    subsection \ref{subsec:stochasticity}: the effects of stochasticity
+  }
+  \item{
+    subsection \ref{subsec:nltt}: the
+    nLTT statistics
   }
@@
   }
-  \section{
+  \item{
     subsection \ref{subsec:data_accessibility}: Data accessibility
   }
-  \section{
+  \item{
     subsection \ref{subsec:author_contributions}: Author contributions
@@
 % MEE style
-\bibliographystyle{pirouette_mee}
+%\bibliographystyle{pirouette_mee}
 % \bibliography{pirouette_supplement}
-\bibliography{pirouette_article}
+%\bibliography{pirouette_article}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-07-14 pirouette_article.tex

@@
 	or where diversification rates change for specific lineages 
-	and their descendants [\cite{etienne2012conceptual, rabosky2014automatic, alfaro2009nine, laudanno2020sls}].
+	and their descendants [\cite{etienne2012conceptual, rabosky2014automatic, alfaro2009nine, laudanno2020detecting}].
 	Other models treat speciation as a process that takes 
@@
 \new{
+Similarly,
+Ritchie and colleagues [\cite{Ritchie2017impact}] showed that inference was accurate when birth-death or skyline coalescent priors were used, but they simulated their trees with a Yule process only, as their focus was not so much on the diversification process but on the influence of inter- and intraspecific sampling.
+}
+
+\new{
   Another way to benchmark a diversification model, is by doing
@@
   This new method simultaneously assesses the substitution, clock and tree
-  models and thus is a hierarchical model.
-  We learn from \cite{duchene2015evaluating} that 
-  such a method may be sensitive to other aspects of the hierarchical model.
+  models [\cite{duchene2015evaluating}].
 }
@@
   \centering
-  \includegraphics[width = 0.99\textwidth]{workflow4.png}
+  \includegraphics[width = 0.97\textwidth]{workflow4.png}
   \caption{
@@
 There is also the option to generate a 'twin tree', 
-that goes through the same pipeline (see section \ref{subsec:twinning}).
+that goes through the same pipeline (see supplementary subsection \ref{subsec:twinning}).
 
@@
 that is the baseline error of the inference pipeline (see
-supplementary materials, section \ref{subsec:twinning} for more details).
+supplementary materials, subsection \ref{subsec:twinning} for more details).
 The second type of control is the use of candidate models,
@@
 statistic [\cite{janzen2015approximate}] to compare phylogenies (see 
-section \label{subsec:nltt} for details regarding the nLTT statistic
+section \ref{subsec:nltt} for details regarding the nLTT statistic
 and its caveats).
@@
 	generated with the posterior parameters!).
-}
+    By contrast,}
     \verb;pirouette;
-\new{
-	aims to expose when assuming standard priors for the 
+	\new{aims to expose when assuming standard priors for the 
 	species tree are a mis- or underparameterization. 
-	Our approach applies to simulated trees and compares 
+	Hence, our approach applies to simulated trees and compares 
 	the posterior distributions of trees generated with 
@@
 	The two methods therefore complement one another.
-} However, we note that the \verb;pirouette; pipeline is not restricted 
+}
+However, we note that the \verb;pirouette; pipeline is not restricted 
 to exploring the effects of a new species tree model. 

### 2020-07-15 pirouette_supplement.tex
Add wall clock times and number of replicates

@@
     From this alignment two posterior distributions of trees are created: 
-    one using the generative model and another one using the inference model with the highest marginal likelihood.
-For each distribution of trees, a distribution of errors, measured with the nLTT statistic, between the posterior trees and the main trees is drawn.
-From the true tree also a twin tree is created (right side of the figure) which follows the same pipeline, leading to two additional error distributions to use as baseline errors.
+    one using the generative model and another one using the inference model 
+    with the highest marginal likelihood. 
+    For each distribution of trees, a distribution of errors, 
+    measured with the nLTT statistic, 
+    between the posterior trees and the main trees is drawn. 
+    From the true tree also a twin tree is created (right side of the figure) 
+    which follows the same pipeline, 
+    leading to two additional error distributions to use as baseline errors.
   }
@@
     Aggregate error distributions, similar to Fig. \ref{fig:example_30} 
-    for the main example, but now for a collection of 20 replicate trees. 
+    for the main example, but now for a collection of 100 replicate trees. 
     For each setting (true generative, true best candidate, twin generative 
@@
     pipeline have been merged into a single distribution.
+    This took 2.7 days (wall clock time) to compute.
   }
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each true tree has 6 taxa. This took 16 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates. 
+    Here each true tree has 6 taxa. 
+    This took 2.7 days (wall clock time) to compute.
+  }
   \label{fig:example_6_taxa}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_32/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. Here each true tree has 12 taxa. This took 19 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates. 
+    Here each true tree has 12 taxa. 
+    This took 6.0 days (wall clock time) to compute.
+  }
   \label{fig:example_12_taxa}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_33/errors.png}
-    \caption{Aggregate error distributions for 20 replicates. Here each true tree has 24 taxa. This took 28 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates. 
+    Here each true tree has 24 taxa. 
+    This took 9.8 days (wall clock time) to compute.
+  }
   \label{fig:example_24_taxa}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_41/errors.png}
-  \caption{Aggregate error distributions for 5 replicates. Here each true tree has 32 taxa. This took 7 hours to compute.}
+  \caption{
+    Aggregate error distributions for 65 replicates. 
+    Here each true tree has 32 taxa. 
+    This took 8.0 days (wall clock time) to compute.
+  }
   \label{fig:example_32_taxa}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_42/errors.png}
-  \caption{Aggregate error distributions for 2 replicates. Here each true tree has 40 taxa. This took 8 hours to compute.}
+  \caption{
+    Aggregate error distributions for 5 replicates. 
+    Here each true tree has 40 taxa.
+    This took 0.83 days (wall clock time) to compute.
+  }
   \label{fig:example_40_taxa}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_19/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates. 
+    Here each each alignment has a sequence length of 500 nucleotides. 
+    This took 2.2 days (wall clock time) to compute.
+  }
   \label{fig:example_500_nucleotides}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates. 
+    Here each each alignment has a sequence length of 1000 nucleotides 
+    (as used in the main example). 
+    This took 2.7 days (wall clock time) to compute.
+  }
   \label{fig:example_1000_nucleotides}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_34/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates. 
+    Here each each alignment has a sequence length of 2000 nucleotides. 
+    This took 4.4 days (wall clock time) to compute.
+  }
   \label{fig:example_2000_nucleotides}
@@
 }
-\giovanni{I think it might be nice to have a simple lot of error medians versus number of nucleotides.
-If we want to be a bit fancier, we can think of including 0.25 and 0.75 percentiles as error bars.}
+\giovanni{
+  I think it might be nice to have a simple lot of error medians 
+  versus number of nucleotides. If we want to be a bit fancier, 
+  we can think of including 0.25 and 0.75 percentiles as error bars.
+}
 
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_22/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates. 
+    Here each true tree is generated by a Yule process. 
+    For the inference we used a Yule tree prior. 
+    This took 2.9 days (wall clock time) to compute.
+  }
   \label{fig:example_yule}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_26/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. 
+  \caption{
+    Aggregate error distributions for 100 replicates. 
     Here each true tree is generated by a BD process. 
-  For the inference we used instead a Yule tree prior. This took 
-  \richel{unknown} hours to compute.
+    For the inference we used instead a Yule tree prior. 
+    This took 2.7 days (wall clock time) to compute.
   }
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_low.png}
-  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with low likelihood.}
+  \caption{
+    Aggregate error distributions for a distribution of trees, 
+    where the true trees are DD with low likelihood.
+  }
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_mid.png}
-  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with median likelihood.}
+  \caption{
+    Aggregate error distributions for a distribution of trees, 
+    where the true trees are DD with median likelihood.
+  }
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_high.png}
-  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with high likelihood.}
+  \caption{
+    Aggregate error distributions for a distribution of trees, 
+    where the true trees are DD with high likelihood.
+  }
 \end{figure}
@@
   direction using more simulations because computational times turned to be 
-  extremely high. The entire run took 120 hours in total.
+  extremely high. 
+  The entire run took 120 hours in total.
 }
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_18/errors.png}
-  \caption{Aggregate error distributions similar to Fig.
-\ref{fig:replicate_trees}, but here the number of \new{substitution}s is not imposed to be the same between true and twin alignment.
-Instead we just use an equal mutation rate.
-This took 10 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates.
+    similar to Fig. \ref{fig:replicate_trees}, 
+    but here the number of \new{substitution}s is not imposed 
+    to be the same between true and twin alignment. 
+    Instead, an equal mutation rate is used. 
+    This took 3.3 days (wall clock time) to compute.
+  }
   \label{fig:example_random_mutations}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_35/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.25 / crown age. This took 11 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 0.25 / crown age. 
+    This took 1.8 days (wall clock time) to compute.
+  }
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_36/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.50 / crown age. This took 14 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 0.50 / crown age. 
+    This took 2.2 days (wall clock time) to compute.
+  }
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_37/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.75 / crown age. This took 16 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 0.75 / crown age. 
+    This took 2.5 days (wall clock time) to compute.
+  }
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_38/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.25 / crown age. This took 17 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 1.25 / crown age. 
+    This took 2.9 days (wall clock time) to compute.
+  }
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_39/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.50 / crown age. This took 10 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 1.50 / crown age. 
+    This took 3.0 days (wall clock time) to compute.
+  }
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_40/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 2.0 / crown age. This took 19 hours to compute.}
+  \caption{
+    Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 2.0 / crown age. 
+    This is done for 100 replicates.
+    This took 3.0 days (wall clock time) to compute.
+  }
 \end{figure}

### 2020-07-16 pirouette_supplement.tex

@@
   in the twin distribution, similar to the result of the previous 
-  subsection, subsection\ref{subsec:distribution}.
+  subsection, subsection \ref{subsec:distribution}.
   In addition to this, we cannot detect a clear trend between the number of 

### 2020-07-17 pirouette_supplement.tex
Changes @Giappo

--- 2020-06-28 pirouette_supplement.tex
@@
   }
-  \caption{
-    Full pirouette pipeline, including comparison to baseline error. 
+  \caption{Full pirouette pipeline, including comparison to baseline error. 
     The true tree (top left) is used to simulate an alignment. 
     From this alignment two posterior distributions of trees are created: 
-    one using the generative model and another one using the inference model with the highest marginal likelihood.
-For each distribution of trees, a distribution of errors, measured with the nLTT statistic, between the posterior trees and the main trees is drawn.
-From the true tree also a twin tree is created (right side of the figure) which follows the same pipeline, leading to two additional error distributions to use as baseline errors.
-  }
+    one using the generative model and another one using the inference model 
+    with the highest marginal likelihood. 
+    For each distribution of trees, a distribution of errors, 
+    measured with the nLTT statistic, 
+    between the posterior trees and the main trees is drawn. 
+    From the true tree also a twin tree is created (right side of the figure) 
+    which follows the same pipeline, 
+    leading to two additional error distributions to use as baseline errors.}
   \label{fig:example_30_full_pipeline}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
-  \caption{
-    Aggregate error distributions, similar to Fig. \ref{fig:example_30} 
-    for the main example, but now for a collection of 20 replicate trees. 
-    For each setting (true generative, true best candidate, twin generative 
-    and twin best candidate), the resulting errors from each replicate 
-    pipeline have been merged into a single distribution.
-  }
+  \caption{Aggregate error distributions, similar to Fig.
+\ref{fig:example_30} for the main example, but now for a collection of 100 replicate trees.
+For each setting (true generative, true best candidate, twin generative and twin best candidate), the resulting errors from each replicate pipeline have been merged into a single distribution.
+This took 2.7 days (wall clock time) to compute.}
   \label{fig:replicate_trees}
@@
 \begin{figure}[H]
-  \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each true tree has 6 taxa. This took 16 hours to compute.}
-  \label{fig:example_6_taxa}
-\end{figure}
-
-\begin{figure}[H]
   \includegraphics[width=0.98\textwidth]{pirouette_example_32/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. Here each true tree has 12 taxa. This took 19 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates. Here each true tree has 12 taxa. This took 6.0 days (wall clock time) to compute.}
   \label{fig:example_12_taxa}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_33/errors.png}
-    \caption{Aggregate error distributions for 20 replicates. Here each true tree has 24 taxa. This took 28 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates. Here each true tree has 24 taxa. This took 9.8 days (wall clock time) to compute.}
   \label{fig:example_24_taxa}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_41/errors.png}
-  \caption{Aggregate error distributions for 5 replicates. Here each true tree has 32 taxa. This took 7 hours to compute.}
+  \caption{Aggregate error distributions for 65 replicates. Here each true tree has 32 taxa. This took 8.0 days (wall clock time) to compute.}
   \label{fig:example_32_taxa}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_42/errors.png}
-  \caption{Aggregate error distributions for 2 replicates. Here each true tree has 40 taxa. This took 8 hours to compute.}
+  \caption{Aggregate error distributions for 5 replicates. Here each true tree has 40 taxa. This took 0.83 days (wall clock time) to compute.}
   \label{fig:example_40_taxa}
@@
 
-\new{
-  We show in figures \ref{fig:example_6_taxa}, \ref{fig:example_12_taxa}, 
+\begin{figure}[H]
+  \includegraphics[width=0.98\textwidth]{supplementary_figures/plot_error_vs_n_taxa.png}
+  \caption{Difference between median true error and median twin error for different number of taxa.}
+  \label{fig:error_vs_ntaxa}
+\end{figure}
+
+\new{
+  We show in figures \ref{fig:replicate_trees}, \ref{fig:example_12_taxa}, 
   \ref{fig:example_24_taxa}, \ref{fig:example_32_taxa} and 
   \ref{fig:example_40_taxa} what are the errors obtained when starting from 
-  phylogenies with, respectively, 6, 12, 24, 32 and 40 taxa. Again we can see 
+  phylogenies with, respectively, 6, 12, 24, 32 and 40 taxa.
+  Again we can see 
   that in each case errors tend to be greater in the true distribution than 
-  in the twin distribution, similar to the result of the previous 
-  subsection, subsection\ref{subsec:distribution}.
-  In addition to this, we cannot detect a clear trend between the number of 
-  taxa of the original phylogeny and the extent of such discrepancy. 
-  Due to long computational times we have been able to provide data only 
-  for a limited amount of replicates. 
-  We cannot exclude that a clearer trend could be observed increasing the 
-  number of replicates.
+  in the twin distribution, similar to the result of subsection \ref{subsec:distribution}.
+  Collecting all the data together we can see that errors tend to decrease as the number of taxa in the considered phylogenies increase (see Fig.
+\ref{fig:error_vs_ntaxa}).
+The data point for $40$ taxa not following the trend could be due to the limited amount of simulated trees taken in consideration due to time constrains.
 }
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_19/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates. 
+    Here each each alignment has a sequence length of 500 nucleotides. 
+    This took 2.2 days (wall clock time) to compute.}
   \label{fig:example_500_nucleotides}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
-  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates. 
+    Here each each alignment has a sequence length of 1000 nucleotides.
+    This is a replicate of Fig. \ref{fig:replicate_trees}. We put it here to facilitate the comparison with the cases with different number of nucleotides.}
   \label{fig:example_1000_nucleotides}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_34/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates. 
+    Here each each alignment has a sequence length of 2000 nucleotides. 
+    This took 4.4 days (wall clock time) to compute.}
   \label{fig:example_2000_nucleotides}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=0.98\textwidth]{supplementary_figures/plot_error_vs_sequence_length.png}
+  \caption{Difference between median true error and median twin error for different sequence lengths.}
+  \label{fig:error_vs_seqlength}
 \end{figure}
@@
   we can observe that the discrepancy between the true and twin error 
-  distributions tend to become smaller as the number of nucleotides increase. 
+  distributions tend to become smaller as the number of nucleotides increase (see also Fig. \ref{fig:error_vs_seqlength}). 
   This occurred for both the generative and best candidate cases.
@@
 }
-\giovanni{
-  Also, do we have the raw data 
-  stored somewhere? From what I can judge from the plots the median discrepancy 
-  is 0.012 for 500 nucleotides, 0.008 for 1000 nucleotides 
-  and 0.003 for 2000 nucleotides.
-}
-\richel{
-  Sure, raw data is at
-  \url{https://www.richelbilderbeek.nl/pirouette_example_19.zip},
-  \url{https://www.richelbilderbeek.nl/pirouette_example_28.zip}
-  and 
-  \url{https://www.richelbilderbeek.nl/pirouette_example_34.zip}.
-  These URLS can be found on the GitHub site.
-  What would you like to do with the raw data?
-}
-\giovanni{I think it might be nice to have a simple lot of error medians versus number of nucleotides.
-If we want to be a bit fancier, we can think of including 0.25 and 0.75 percentiles as error bars.}
 
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_22/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates. 
+    Here each true tree is generated by a Yule process. 
+    For the inference we used a Yule tree prior. 
+    This took 2.9 days (wall clock time) to compute.}
   \label{fig:example_yule}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_26/errors.png}
-  \caption{Aggregate error distributions for 10 replicates. 
+  \caption{Aggregate error distributions for 100 replicates. 
     Here each true tree is generated by a BD process. 
-  For the inference we used instead a Yule tree prior. This took 
-  \richel{unknown} hours to compute.
-}
+    For the inference we used instead a Yule tree prior. 
+    This took 2.7 days (wall clock time) to compute.}
 \end{figure}
@@
 }
-\new{
-  As we use only extant trees, it is reasonable that the method
+\new{As we use only extant trees, it is reasonable that the method
   is slightly weaker in distinguishing between the Yule and BD
   models. It is unknown what the discriminatory power would be
-  when comparing trees with extinction events.
-}
+  when comparing trees with extinction events.}
 
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_23/errors_mid.png}
-  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with median likelihood.}
+  \caption{Aggregate error distributions for a distribution of trees, 
+    where the true trees are DD with median likelihood.}
 \end{figure}
@@
 \begin{figure}[H]
-  \includegraphics[width=0.98\textwidth]{pirouette_example_18/errors.png}
-  \caption{Aggregate error distributions similar to Fig.
-\ref{fig:replicate_trees}, but here the number of \new{substitution}s is not imposed to be the same between true and twin alignment.
-Instead we just use an equal mutation rate.
-This took 10 hours to compute.}
+  %\includegraphics[width=0.98\textwidth]{pirouette_example_18/errors.png}
+  \includegraphics[width=0.94\textwidth]{pirouette_example_18/errors.png}
+  \caption{Aggregate error distributions for 100 replicates.
+    similar to Fig. \ref{fig:replicate_trees}, 
+    but here the number of \new{substitution}s is not imposed 
+    to be the same between true and twin alignment. 
+    Instead, an equal mutation rate is used. 
+    This took 3.3 days (wall clock time) to compute.}
   \label{fig:example_random_mutations}
@@
 Here, we show the same results for different mutation rates.
-
-\new{
-  We can observe that the error increases for mutation rates above
+\new{
+\iffalse
+  We can observe that the error tends to increases for mutation rates above
   `1.0 / crown age`. Below this value, the errors are the same.
@@
   as is still present in the twin distributions.
+\fi
+ The results for the different mutation rates are shown in Figs.
+\ref{fig:example_0.25_mutation_rate} (0.25 / crown age), \ref{fig:example_0.50_mutation_rate} (0.50 / crown age), \ref{fig:example_0.75_mutation_rate} (0.75 / crown age), \ref{fig:replicate_trees} (1.00 / crown age), \ref{fig:example_1.25_mutation_rate} (1.25 / crown age),
+ \ref{fig:example_1.50_mutation_rate} (1.50 / crown age) and
+ \ref{fig:example_2.00_mutation_rate} (2.00 / crown age).
+ }
+\new{
+ Fig.
+\ref{fig:error_vs_mutationrate} summarizes all the other figures showing on the y-axis, for each value of the mutation rate, the difference between the median of the true distribution and the median of the twin distribution.
+We can observe a general positive trend as the mutation rate increase, even though the value for 1.5 / crown age suggests to take this result with caution.
+It is possible, however, that a more regular trend could be observed increasing the number of simulations.
 }
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_35/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.25 / crown age. This took 11 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 0.25 / crown age. 
+    This took 1.8 days (wall clock time) to compute.}
+  \label{fig:example_0.25_mutation_rate}
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_36/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.50 / crown age. This took 14 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 0.50 / crown age. 
+    This took 2.2 days (wall clock time) to compute.}
+  \label{fig:example_0.50_mutation_rate}
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_37/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 0.75 / crown age. This took 16 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 0.75 / crown age. 
+    This took 2.5 days (wall clock time) to compute.}
+  \label{fig:example_0.75_mutation_rate}
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_38/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.25 / crown age. This took 17 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 1.25 / crown age. 
+    This took 2.9 days (wall clock time) to compute.}
+  \label{fig:example_1.25_mutation_rate}
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_39/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.50 / crown age. This took 10 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 1.50 / crown age. 
+    This took 3.0 days (wall clock time) to compute.}
+  \label{fig:example_1.50_mutation_rate}
 \end{figure}
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_40/errors.png}
-  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 2.0 / crown age. This took 19 hours to compute.}
+  \caption{Aggregate error distributions for 100 replicates,
+    for the tree distribution presented 
+    in \ref{subsec:distribution} but with a per-nucleotide mutation rate 
+    of 2.0 / crown age. 
+    This is done for 100 replicates.
+    This took 3.0 days (wall clock time) to compute.}
+  \label{fig:example_2.00_mutation_rate}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=0.98\textwidth]{supplementary_figures/plot_error_vs_mutation_rate.png}
+  \caption{Difference between median true error and median twin error for different values of mutation rate.}
+  \label{fig:error_vs_mutationrate}
 \end{figure}

### 2020-07-17 supplementary_figures/error_vs_var.R

@@ -1,0 +1,161 @@
+#' Download the data and create a plot of error vs variable
+#' @param var variable to plot on the x-axis
+#' @param datafolder folder to download the data in
+#' @param resultsfolder folder to save the results in
+#' @author Giovanni Laudanno
+#' @return a plot
+plot_error_vs_var = function(
+  var,
+  datafolder,
+  resultsfolder
+) {
+  data_filename = paste0("errors_vs_", var, ".RData")
+
+  # Download data
+  path = file.path(resultsfolder, var)
+  if (!dir.exists(path)) {dir.create(path)}
+  setwd(datafolder)
+
+  if (var == "n_taxa") {
+    examples <- c(28, 32, 33, 41, 42)
+    var_values <- c(6, 12, 24, 32, 40)
+  } else if (var == "sequence_length") {
+    examples <- c(19, 28, 34)
+    var_values <- c(500, 1000, 2000)
+  } else if (var == "mutation_rate") {
+    examples <- c(35, 36, 37, 28, 38, 39, 40)
+    var_values <- c(0.25, 0.50, 0.75, 1.00, 1.25, 1.50, 2.00)
+  } else {
+    stop("This variable cannot be used")
+  }
+  testit::assert(length(examples) == length(var_values))
+  for (example in examples) {
+    stringa = paste0("pirouette_example_", example)
+    if (!any(grepl(x = list.files(datafolder), pattern = stringa))) {
+      zipfile = file.path(datafolder, paste0(stringa, ".zip"))
+      utils::download.file(
+        url = paste0("http://www.richelbilderbeek.nl/pirouette_example_", example, ".zip"),
+        destfile = zipfile
+      )
+      utils::unzip(zipfile)
+    }
+  }
+
+  # Collect data
+  if (file.exists(data_filename)) {
+    load(file = data_filename)
+  } else {
+    ee = f = 1; seed = 314
+    maxseed = 100
+    errors = data.frame(matrix(NA, nrow = 9001 * 4 * maxseed * length(examples), ncol = 3))
+    colnames(errors) = c("value", "model", var)
+    i = 0
+    for (ee in seq_along(examples)) {
+      example = examples[ee]
+      var_value = var_values[ee]
+      cat(var, "is", var_value, "\n")
+      xx = file.path(datafolder, paste0("pirouette_example_", example), paste0("example_", example))
+      testit::assert(dir.exists(xx))
+      seeds = list.files(xx)
+      for (seed in seeds) {
[…]

### 2020-07-22 pirouette_article.tex

@@
 	BEAST has built-in priors such as the Yule [\cite{yule}] and 
-	(constant-rate) birth-death [\cite{nee1994reconstructed}] models
+	(constant-rate) birth-death (BD) [\cite{nee1994reconstructed}] models
 	as well as coalescent priors.
@@
 	Yule model [\cite{heled2015calibrated}],
-	  BD with incomplete sampling [\cite{stadler2009incomplete}],
-	  BD serial sampling [\cite{stadler2012estimating}],
-	  BD serial skyline [\cite{stadler2013birth}]
-	  Fossilized BD process [\cite{gavryushkina2014bayesian}], and
-	  BD SIR [\cite{kuhnert2014simultaneous}].
+	  the BD model with incomplete sampling [\cite{stadler2009incomplete}],
+	  the BD moel with serial sampling [\cite{stadler2012estimating}],
+	  the BD serial skyline model [\cite{stadler2013birth}],
+	  the fossilized BD process [\cite{gavryushkina2014bayesian}], and
+	  the BD SIR model [\cite{kuhnert2014simultaneous}].
 }
@@
 	depends on one or more traits 
-	[\cite{maddison2007estimating, fitzjohn2012diversitree, herrera2018detecting}].
+	[\cite{maddison2007estimating, fitzjohn2012diversitree, herrera2019detecting}].
 }
@@
 
-\new{
-  Another way to benchmark a diversification model, is by doing
-  a model comparison, in which the best model is determined
-  from a set of models. A good early example is \cite{goldman1993statistical} 
+\new{Another way to benchmark a diversification model, is by doing a model comparison, in which the best model is determined
+from a set of models. A good early example is \citet{goldman1993statistical} 
 in which Goldman compared DNA substitution models.
 A recent approach to test the impact of tree prior choice, proposed
-  by \cite{duchene2018phylodynamic}, allows to measure
+by \citet{duchene2018phylodynamic}, allows to measure
 model adequacy for phylodynamic models that are
-  mathematically described (i.e. having a known likelihood equation).
-}
+mathematically described (i.e. have a known likelihood equation).}
 
@@
 }
-\new{
-  This new method simultaneously assesses the substitution, clock and tree
-  models [\cite{duchene2015evaluating}].
-}
+\new{This new method simultaneously assesses the substitution, clock and tree models [\cite{duchene2015evaluating}].}
 The method starts with a phylogeny generated by the new model. 
@@
 as the 'generative tree model' $\mathit{p_{G}}$.
-\rampal{
-	Perhaps if you replace by "generative model for the twin tree", 
-	then it is fine.
-}
 \new{
-  A 'generative tree model', in this paper, is always a standard tree model.
-  If the actual model that generated the tree is non-standard, 
-  the 'generative tree model' is the standard tree model that is 
-  closest to the actual model. In that case, the
-  'generative tree model' serves as a baseline model.
+  A 'generative tree model', in this paper, can be either the novel diversification model for which we are testing the impact of choosing standard tree priors for, or it is the model with which we generate the twin tree that is needed for comparison (see below).
+In the latter case, we also refer to it as the actual generative tree model, and it thus serves as a baseline model.
 This is is done in the example, where the Yule model is the generative model.
@@
 \new{
-  The error we aim to quantify is not of stochastic nature.
+  The inference error we aim to quantify is not of stochastic nature.
   Stochastic errors are usually non-directional.
   We, instead, aim to expose the bias due to the mismatch between 
-  a generative model (that has either generated the phylogeny, or
-  is assumed to have do so) and the model(s) used in the actual inference.
+  a generative model (that has generated the phylogeny) and the model(s) used in the actual inference.
 }
-We define the birth-death (BD) model [\cite{nee1994reconstructed}] as
+We define the birth-death \new{(BD)} model [\cite{nee1994reconstructed}] as
 the standard tree model, as many (non-standard) tree models 
-have a parameter setting that can be reduced to this model. 
+have a parameter setting \new{such that it reduces} to this model. 
 One such example is the diversity-dependent (DD) diversification 
@@
 \new{
-  The BD model can be seen as a special case of the DD model. 
-  In fact, for
+  The BD model can be seen as a special case of the DD model, because for
 } 
@@
 when assessing a novel tree prior, 
-our example contain only one replicate,
+our example contains only one replicate,
 as the goal is to show the workings of \verb;pirouette;,
 instead of doing an extensive analysis.
-Supplementary material includes results of replicated runs under multiple settings.
+The supplementary material includes results of replicated runs under multiple settings.
 
@@
     phylogenies (4a). 
-    Each of the posteriors' trees is compared to the true phylogeny (1a) 
+    Each of the posterior trees is compared to the true phylogeny (1a) 
     using the error measure $\mathit{E}$, 
@@
     using clock model $\mathit{c_{G}}$ and site model $\mathit{s_{G}}$ 
-    imported from the generative model (the model that has
-    actually generated the phylogeny, or is assumed to have done so). 
+    used with the generative tree model (the novel tree model).
     The twin pipeline follows the procedure of the main pipeline, 
@@
     an inference model $\mathit{I}$ is chosen (which may or may not differ 
-    from the actual model that generated the tree).
+    from the model that generated the tree).
   \item The inference model and the alignment are used 
@@
 a set of standard inference models $I_{1},\dots,I_{n}$.
-For example, if the generative model is known and standard,
+For example, if the generative model is known and standard (which it is for the twin tree, see below),
 one can specify the inference model to be the same as the generative model.
@@
 to be closest to the true tree model.
-Also, if we want to run only the inference
-model that fits best to an alignment from a set of candidates,
-one can specify these inference models as 
-well (see section \ref{subsec:candidates}).
+Alternatively, if we want to run only the inference
+model that fits best to an alignment from a set of candidates \new{(regardless of whether these generated the alignments)},
+one can specify these inference models (see section \ref{subsec:candidates}).
 
@@
 
-The fourth step quantifies the inference error made. 
+The fourth step quantifies the new{impact of choosing standard models for inference, i.e.} the inference error made. 
 First the burn-in fraction is removed, i.e. the first phase of the 
@@
 
-Similar to lab experiments, \verb;pirouette; allows to perform
+\verb;pirouette; allows for
 two types of control measurements. The first type of control
@@
 supplementary materials, subsection \ref{subsec:twinning} for more details).
+\new{This the error that arises when the models used in inference are identical to the ones used in generating the alignments.}
 The second type of control is the use of candidate models,
-which results in an error distribution
+which result in an error distribution
 for a generative model that is determined to be the best fit to
@@
 section \ref{subsec:candidates} for more details).
+\new{The underlying idea is that using a substitution model in inference than used in generating the alignment may partly compensate for choosing a standard tree model instead of the generative tree model as tree prior in inference.}
 Additionally, multiple \verb;pirouette; runs are needed
@@
   \caption{
-    The inference error made 
-    for an assumed generative tree model and best candidate model,
-    compared with the error obtained for the twin tree.
-    The assumed generative tree model is the model assumed to be closest to
-    the actual generative tree model that generated the true tree.
-    The twin distributions show the baseline inference error.
+    \new{The inference error made 
+    when using the generative substitution and clock models (in this case JC69 and strict clock) and a Yule tree prior (this is the assumed generative model because the real generative model is of course unknown) and when using the best-fitting candidate model (in this case JC69, relaxed lognormal and BD), for both the true tree generated by the novel diversification model and for the twin tree.
+    The twin distributions show the baseline inference error.}
     Vertical dashed lines show the median error value per distribution.
@@
 In the upper panel of Figure \ref{fig:example_30},
-we can see that the error distributions of the assumed generative model
+we can see that the error distributions of the \new{ (assumed) generative model (i.e.
+the known generative substitution and clock models, and the tree model that is used for generating and inferring the twin tree and assumed in inference of the true tree)}
 differ strongly between the true and twin tree. 
 This difference shows the extent of the mismatch between
-the actual tree model (which is DD) and the (BD) tree prior used.
+the true tree model (which is DD) and the (Yule) tree prior used in inference.
 Because these distributions are distinctively different,
 the inference error made when using an 
-incorrect (that is, BD) tree prior on a DD tree
+incorrect (that is, Yule) tree prior on a DD tree
 is profound.
@@
 candidate model is only slightly better at inferring the true tree,
-than the assumed generative model, 
+than the (assumed) generative model, 
 thereby showing that they cannot compensate 
@@
 Figure~\ref{fig:example_30} illustrates the primary result of our pipeline: 
-it shows the error distributions for the true tree and the twin tree 
-when either the (assumed) generating model or the best candidate model is used in inference. 
+it shows the error distributions for the true tree and the twin tree \new{
+when either the generative model (for substitution and clock models these are known, for the tree model it must be assumed for the true tree and it is known for the twin tree) or the best-fitting set candidate model (i.e.
+combination of tree model, substitution model and clock model) is used in inference.}
 The clear difference between the error distributions 
@@
 However, we note that for longer DNA sequences, the assumption 
-of constant substitution rates may become less realistic 
+of the same substitution rates across the entire sequence may become less realistic (different genes may experience different substitution rates)
 and hence longer sequences may require more parameters. 
@@
   For models that make different predictions on topology, 
-  the twinning process should be modified accordingly.
+  the twinning process should be modified.
 }
@@
 	also developed a method to assess the adequacy of a tree model
-	on empirical trees. They simulate trees from the posterior distribution of 
-	the parameters and then compare this to the originally inferred tree using 
+	on empirical trees. They simulated trees from the posterior distribution of 
+	the parameters and then compared this to the originally inferred tree using 
 	tree statistics, to determine whether the assumed tree model in inference 

### 2020-07-22 pirouette_supplement.tex
Removed 'clearpage' as this results in a line of comma's like ',,,,,,,,,,,,,,,,,,,,,,,,,,,,'
Indeed, @Giappo fixed the duplicate figure, thanks! Fix #110
Fixed newpage/clearpage error by using more '\newpage'

@@
 
+\newpage
+
 \input{pirouette_example_30/example_30_314/esses_gen.latex}
 
+\newpage
+
 \input{pirouette_example_30/example_30_314/esses_best.latex}
 
+\newpage
+
 \input{pirouette_example_30/example_30_314/esses_twin_gen.latex}
 
+\newpage
+
 \input{pirouette_example_30/example_30_314/esses_twin_best.latex}
 
+\newpage
+
 \input{pirouette_example_30/example_30_314/evidence_true.latex}
 
+\newpage
+
 \input{pirouette_example_30/example_30_314/evidence_twin.latex}
 
-\clearpage
 \newpage
@@
   \includegraphics[width=0.98\textwidth]{pirouette_example_28/errors.png}
-  \caption{Aggregate error distributions, similar to Fig.
-\ref{fig:example_30} for the main example, but now for a collection of 100 replicate trees.
-For each setting (true generative, true best candidate, twin generative and twin best candidate), the resulting errors from each replicate pipeline have been merged into a single distribution.
-This took 2.7 days (wall clock time) to compute.}
+  \caption{
+    Aggregate error distributions, 
+    similar to Fig. \ref{fig:example_30} for the main example, 
+    but now for a collection of 100 replicate trees. 
+    For each setting (true generative, true best candidate, 
+    twin generative and twin best candidate), the resulting errors 
+    from each replicate pipeline have been merged into a single distribution. 
+    This took 2.7 days (wall clock time) to compute.
+  }
   \label{fig:replicate_trees}
@@
 \ref{fig:error_vs_ntaxa}).
-The data point for $40$ taxa not following the trend could be due to the limited amount of simulated trees taken in consideration due to time constrains.
+The data point for $40$ taxa not following the trend could be due to the limited amount of simulated trees taken in consideration due to time constraints.
 }
@@
 
-\clearpage
+\newpage
 

### 2020-07-23 pirouette_article.tex
Show parameters explictly, as suggested by @rsetienne

@@
 	  the BD model with incomplete sampling [\cite{stadler2009incomplete}],
-	  the BD moel with serial sampling [\cite{stadler2012estimating}],
+	  the BD model with serial sampling [\cite{stadler2012estimating}],
 	  the BD serial skyline model [\cite{stadler2013birth}],
@@
 
+# Create the inference model for the generative model explicitly
+pir_params$experiments[[1]]$inference_model$tree_prior <- create_yule_tree_prior()
+pir_params$experiments[[1]]$inference_model$site_model <- create_jc69_site_model()
+pir_params$experiments[[1]]$inference_model$clock_model <- create_strict_clock_model()
+
+# Simulate a twin tree using BD explicitly
+pir_params$twinning_params$sim_twin_tree_fun <- get_sim_bd_twin_tree_fun()
+
 # Do the runs
@@
     \new{The inference error made 
-    when using the generative substitution and clock models (in this case JC69 and strict clock) and a Yule tree prior (this is the assumed generative model because the real generative model is of course unknown) and when using the best-fitting candidate model (in this case JC69, relaxed lognormal and BD), for both the true tree generated by the novel diversification model and for the twin tree.
+    when using the generative substitution and clock models (in this case JC69 and strict clock) and a Yule tree prior (this is the assumed generative model because the real generative model is of course unknown) and when using the best-fitting candidate model (in this case JC69, relaxed log-normal and BD), for both the true tree generated by the novel diversification model and for the twin tree.
     The twin distributions show the baseline inference error.}

### 2020-07-25 pirouette_article.tex
Changes @rsetienne

@@
 
-# Create phylogeny
-phylogeny <- create_exemplary_dd_tree(
-  n_taxa = 6, 
-  crown_age = 10,
-  extinction_rate = 0.1
-)
+# Create a DD phylogeny with 6 taxa and a crown age of 10
+phylogeny <- create_exemplary_dd_tree()
 
-# Use standard pirouette setup
+# Use standard pirouette setup.
+This creates a list object with all settings for generating the alignment, the inference using BEAST2,  the twinning parameters to generate the twin tree and infer it using BEAST2, and the error measure
 pir_params <- create_std_pir_params()
-
-# Create the inference model for the generative model explicitly
-pir_params$experiments[[1]]$inference_model$tree_prior <- create_yule_tree_prior()
-pir_params$experiments[[1]]$inference_model$site_model <- create_jc69_site_model()
-pir_params$experiments[[1]]$inference_model$clock_model <- create_strict_clock_model()
-
-# Simulate a twin tree using BD explicitly
-pir_params$twinning_params$sim_twin_tree_fun <- get_sim_bd_twin_tree_fun()
 
@@
   \caption{
-    \new{The inference error made 
-    when using the generative substitution and clock models (in this case JC69 and strict clock) and a Yule tree prior (this is the assumed generative model because the real generative model is of course unknown) and when using the best-fitting candidate model (in this case JC69, relaxed log-normal and BD), for both the true tree generated by the novel diversification model and for the twin tree.
+    \new{The impact of the tree prior for the example tree in Figure 2.
+The alignment for this true tree was generated using a JC69 substitution model and strict clock model.
+For inferring the tree from this alignment in the 'generative' scenario the same substitution and clock models were used, and a Yule tree prior (this is the assumed generative model because the real generative model is of course unknown).
+In the 'best' scenario the best-fitting candidate models (JC69 substitution model, relaxed log-normal clock model and BD tree prior) were used.
+For the twin tree, the same models were chosen, except that BD was used to generate and infer the twin tree in the 'generative' scenario.
     The twin distributions show the baseline inference error.}

### 2020-07-25 pirouette_supplement.tex
Changes @rsetienne

@@
 one can download and install the package from \verb;pirouette;'s GitHub 
-repository. We do first need the \verb;mcbette; and \verb;nodeSub; packages:
+repository. We first need the \verb;mcbette;, \verb;tiebeaur; and \verb;nodeSub; packages:
 \begin{lstlisting}[
@@
   "richelbilderbeek/mcbette"
+)
+remotes::install_github(
+  "richelbilderbeek/tiebeaur"
 )

### 2020-07-25 pirouette_supplement.tex
Describe pipeline in more detail, as requested by @rsetienne

@@
 
-This subsection shows the diagnostics from the main example, which
-uses one tree.
+This subsection describes the pipeline of the main example 
+and its diagnostics in more detail.
+
+The pipeline starts at the top-left panel of figure \ref{fig:pipeline} (which 
+is identical to figure \ref{fig:dd_tree}),
+which is the 'true tree'.
+The 'true tree' is generated by the diversity-dependent (DD) 
+tree model [\citep{DDD, etienne2012diversity}],
+which is a BD model with a speciation rate that is dependent on the number of species,
+with (an arbitrarily chosen) crown age of 10 time units 
+and 6 tips for an extinction rate of 0.1. 
+The carrying-capacity is set to 6. 
+The initial speciation rate $\lambda_0$ is chosen 
+such that the expected number of species in a constant-rate BD model 
+would be equal to the number of tips, which amounts to $\lambda_0 = 0.63$.
+
+From this 'true tree', a 'true alignment' is simulated, using
+the JC69 nucleotide substitution model and a strict clock model.
+The resulting alignment is shown at the center-left
+of figure \ref{fig:pipeline}.
+
+From the 'true alignment' the generative inference model is run.
+Of course, it cannot be the actual (DD) model. Instead, the
+default BEAST2 inference model is used, which assumes a JC69 nucleotide
+substitution model, a strict clock model and a Yule tree model.
+The resulting posterior trees are shows in the
+center-left panel of figure \ref{fig:pipeline}.
+
+From this 'generative true' posterior (center-left panel in 
+figure \ref{fig:pipeline}), the difference between each of its trees is
+compared to the 'true tree' (top-left panel), using the nLTT statistic,
+resulting in the error distribution shown at the bottom-left panel
+of figure \ref{fig:pipeline}.
+
+Based on the 'true alignment' (center-left panel), the candidate model
+with the highest marginal likelihood is determined, from a set of 
+15 models. The set of models consists out of all combinations
+of all 4 nucleotides substitution models (JC69, HKY, TN93, GTR),
+all 2 clock models (strict and relaxed log-normal) and 2 birth-death
+models (Yule and Birth-Death), except for the inference model used as the
+generative model (JC69, strict, Yule). The inference model that had the
+highest evidence (as shown in Table \ref{tab:true_evidence}) was 
+the inference model with a JC69 nucleotide substitution model,
+a relaxed log-normal clock model and a Yule tree model.
+From the 'true alignment' this best candidate inference model is run.
+The resulting posterior trees are shown in the second panel
+of the third row of posteriors in figure \ref{fig:pipeline}.
+
+From this 'best true' posterior (second panel in third row of
+figure \ref{fig:pipeline}), the difference between each of its trees is
+compared to the 'true tree' (top-left panel), using the nLTT statistic,
+resulting in the second error distribution in the bottom row
+of figure \ref{fig:pipeline}.
+
+From the 'true tree' (top-left) we generated a BD twin tree (top-right).
+
+From this 'twin tree', a 'twin alignment' is simulated, using
+the JC69 nucleotide substitution model and a strict clock model.
+The resulting alignment is shown at the center-right
+of figure \ref{fig:pipeline}.
+
+From the 'twin alignment' the generative inference model is run as well.
+Also here, the default BEAST2 inference model is used, 
+which assumes a JC69 nucleotide
+substitution model, a strict clock model and a Yule tree model.
+The resulting posterior trees are shows in the
+third panel of the third row of figure \ref{fig:pipeline}.
+
+From this 'generative twin' posterior (third panel in 
+third row of figure \ref{fig:pipeline}), 
+the difference between each of its trees is
+compared to the 'twin tree' (top-right panel), using the nLTT statistic,
+resulting in the error distribution shown at the third panel
+of the bottom row of figure \ref{fig:pipeline}.
+
+Based on the 'twin alignment' (center-right panel), the candidate model
+with the highest marginal likelihood is determined, from the same 
+set of 15 candidate models.The inference model that had the
+highest evidence (as shown in Table \ref{tab:twin_evidence}) was 
+the inference model with a JC69 nucleotide substitution model,
+a strict clock model and a BD tree model (note: this indeed matches how
+the twin tree and twin tree have been simulated).
+From the 'twin alignment' this best candidate inference model is run.
+The resulting posterior trees are shown in the fourth panel
+of the third row of posteriors in figure \ref{fig:pipeline}.
+
+From this 'best twin' posterior (fourth in third row of
+figure \ref{fig:pipeline}), the difference between each of its trees is
+compared to the 'twin tree' (top-right panel), using the nLTT statistic,
+resulting in the fourth error distribution in the bottom row
+of figure \ref{fig:pipeline}.
+
 To assess if the results of the inference are meaningful one important 

### 2020-07-25 pirouette_supplement.tex
No tiebauer dependency in pirouette anymore

@@
 one can download and install the package from \verb;pirouette;'s GitHub 
-repository. We first need the \verb;mcbette;, \verb;tiebeaur; and \verb;nodeSub; packages:
+repository. We first need the \verb;mcbette; and \verb;nodeSub; packages:
 \begin{lstlisting}[
@@
   "richelbilderbeek/mcbette"
-)
-remotes::install_github(
-  "richelbilderbeek/tiebeaur"
 )

### 2020-07-26 pirouette_article.tex
Changes @rsetienne

@@
 The error distribution shown in Figure \ref{fig:example_30}
-is produced, which uses the nLTT 
-statistic [\cite{janzen2015approximate}] to compare phylogenies (see 
+is produced, which uses the nLTT statistic [\cite{janzen2015approximate}] to compare phylogenies (see 
 section \ref{subsec:nltt} for details regarding the nLTT statistic
@@
   \caption{
-    \new{The impact of the tree prior for the example tree in Figure 2.
+    \new{The impact of the tree prior for the example tree in Figure \ref{fig:dd_tree}.
 The alignment for this true tree was generated using a JC69 substitution model and strict clock model.
 For inferring the tree from this alignment in the 'generative' scenario the same substitution and clock models were used, and a Yule tree prior (this is the assumed generative model because the real generative model is of course unknown).
-In the 'best' scenario the best-fitting candidate models (JC69 substitution model, relaxed log-normal clock model and BD tree prior) were used.
-For the twin tree, the same models were chosen, except that BD was used to generate and infer the twin tree in the 'generative' scenario.
+In the 'best' scenario the best-fitting candidate models (JC69 substitution model, relaxed log-normal clock model and Yule tree prior) were used.
+For the twin tree, the same models were chosen, except that BD was used to generate and infer the twin tree.
     The twin distributions show the baseline inference error.}
@@
 we can see that the error distributions of the \new{ (assumed) generative model (i.e.
-the known generative substitution and clock models, and the tree model that is used for generating and inferring the twin tree and assumed in inference of the true tree)}
+the known generative substitution and clock models, and the tree model that is assumed in inference of the true tree, and the tree model that is used for generating and inferring the twin tree)}
 differ strongly between the true and twin tree. 
@@
 the inference error made when using an 
-incorrect (that is, Yule) tree prior on a DD tree
-is profound.
+incorrect tree prior on a DD tree is profound.
 
@@
 The candidate model that had highest evidence given the simulated alignment,
-was JC, RLN, BD (see Table \ref{tab:options} for the meaning of these 
-abbreviations). JC was indeed the site model used when simulating the
-alignments. The RLN clock model is an unexpected choice: the RLN clock
-model assumes that mutation rates differ between lineages, whereas the
-alignment was simulated with equal mutation rates for all lineages.
-The BD model is indeed closer to the DD model than Yule, as both BD and DD
-model allow for extinction, while Yule does not.
+was JC69, RLN and Yule (see Table \ref{tab:options} for the meaning of these 
+abbreviations). JC69 was indeed the site model used when simulating the
+alignments. The RLN clock model is a surprising result: it assumes that mutation rates differ between lineages, whereas the
+alignment was simulated with equal mutation rates for all lineages \new{(strict clock)}.
+\new{The Yule model is not surprising as it is the standard tree prior that is probably closest to the true DD model because it shows no pull-of-the-present (but also no slowdown).}
 
@@
 
-We showed how to use \verb;pirouette; to quantify the \new{impact} of a 
-tree prior in Bayesian phylogenetics, assuming the simplest standard 
-tree model possible.
-In principle any other standard tree model can be assumed, 
-but we chose to provide the simplest example.
+We showed how to use \verb;pirouette; to quantify the \new{impact of a 
+tree prior in Bayesian phylogenetics, assuming - for illustrative purposes - the simplest standard substitution, clock and
+tree models, but also the models that would be selected among many different standard tree priors according to the highest marginal likelihood, as this would be a likely strategy for an empiricist.
+We recommend exploring different candidate models, but note that this is computationally highly demanding.}
 
@@
 for the true tree and the twin tree suggests 
-that the choice of tree prior does matter.
+that the choice of tree prior matters.
 We note, however, that only one tree from a novel tree model
@@
 decreases for longer DNA sequences. We investigated this
-superficially and confirmed this prediction (see the supplementary materials). 
+superficially and confirmed this prediction (see the supplementary material). 
 However, we note that for longer DNA sequences, the assumption 
@@
 	if sufficient data is available the species tree prior may not be important,
-	and hence the inference may be adequate 
+	and hence inference may be adequate 
 	even though the assumed species tree prior is not. 
@@
 }
-However, we note that the \verb;pirouette; pipeline is not restricted 
+Furthermore, we note that the \verb;pirouette; pipeline is not restricted 
 to exploring the effects of a new species tree model. 
@@
 In conclusion, \verb;pirouette; can show the errors to be expected
-when the model assumed in inference is different from the 
-actual generative model.
+when the model assumed in inference is different from the actual generative model.
 The user can then judge whether or not this new model 
-should be implemented in our Bayesian phylogenetic tool.
+should be implemented in the Bayesian phylogenetic tool.
 

### 2020-07-26 pirouette_supplement.tex
Changes @rsetienne

@@
     subsection \ref{subsec:nltt}: the
-    nLTT statistics
+    nLTT statistic
   }
@@
   \item{
-    subsection \ref{subsec:n_taxa}: the effect of the number
-    of taxa
+    subsection \ref{subsec:n_taxa}: the effect of the number of taxa
   }
@@
     subsection \ref{subsec:simplest_correct_parameterization} shows the
-    effect when doing an inference in the simplest use case
+    effect when performing inference in the simplest use case
   }
@@
     subsection \ref{subsec:under_parameterization} shows the
-    effect when doing an inference with an under-parameterization
+    effect when performing inference with an under-parameterization
   }
@@
 
-Similar to lab experiments, \verb;pirouette; allows to perform
-a control measurement, by use of a process we call twinning. 
-In this context, this control results in an error distribution
+\verb;pirouette; allows to perform
+a control measurement, by use of a process we call twinning. This control results in an error distribution
 that is the baseline error of the pipeline. The difference
-between the 'true' and 'twin' error distributions is caused only
-by the mismatch between the (known-to-be-true or assumed) 
-generative tree model and the 
+between the 'true' and 'twin' error distributions is caused only by the mismatch between the true tree model and the 
 tree prior used in the actual inference.
@@
 and $T_2$, which generates a 'twin alignment' (Fig.~\ref{fig:pipeline}, 2b).
-Both twin tree and alignment will be analyzed in the same way 
-as the true tree and alignment.
+Both twin tree and alignment will be analyzed in the same way as the true tree and alignment.
 
@@
 \new{
-  We choose to preserve the original topology to 
+  We chose to preserve the original topology to 
   increase the similarity between the twin to the original tree. 
@@
   assumption about topology (all topologies are equally likely).
-  However, this might not be suited for new models that assign different probabilities
+  However, this might not be suitable for new models that assign different probabilities
   to trees with the same branching times but different topologies.
@@
 (e.g. speciation and extinction rates, in case of a BD model) 
-that maximize this likelihood applied 
-to the true tree, conditioned on its number of tips $n_{\mathit{G}}$:
+that maximize this likelihood applied to the true tree, conditioned on its number of tips $n_{\mathit{G}}$:
 \begin{align}
@@
 with the same clock model, site model and mutation rate 
-used to simulate the original alignment. 
+used to simulate the alignment on the true. 
 The twin alignment can be simulated in any user-defined way.
-\verb;pirouette; supplies the option to have it simulated with
-the same mutation rate as the true alignment. By default, however,
+\verb;pirouette; provides the option simulate it with the same mutation rate as the true alignment. By default, however,
 not only the same mutation rate is used, but also the total number of 
@@
 matches the true alignment. The total number of \new{substitution}s is defined
-as the amount of different nucleotides between the (known) root sequence
+as the number of different nucleotides between the (known) root sequence
 compared to the sequences at the tips.
@@
 If twinning is used, a candidate model that has the highest evidence for
-the twin alignment is also used to create another error
+the twin alignment is also used to create the twin error
 distribution.
@@
 \new{
-  The nLTT statistic is unsuitable, however, to distinguish
+  The nLTT statistic is not suitable, however, to distinguish
   between a constant-rate BD model and a family of time-dependent 
@@
 substitution model, a strict clock model and a Yule tree model.
-The resulting posterior trees are shows in the
+The resulting posterior trees are shown in the
 center-left panel of figure \ref{fig:pipeline}.
 
-From this 'generative true' posterior (center-left panel in 
-figure \ref{fig:pipeline}), the difference between each of its trees is
+From this 'generative true' posterior (center-left panel in figure \ref{fig:pipeline}), the difference between each of its trees is
 compared to the 'true tree' (top-left panel), using the nLTT statistic,
-resulting in the error distribution shown at the bottom-left panel
+resulting in the error distribution shown in the bottom-left panel
 of figure \ref{fig:pipeline}.
@@
 with the highest marginal likelihood is determined, from a set of 
-15 models. The set of models consists out of all combinations
+15 models. The set of models consists of all combinations
 of all 4 nucleotides substitution models (JC69, HKY, TN93, GTR),
 all 2 clock models (strict and relaxed log-normal) and 2 birth-death
-models (Yule and Birth-Death), except for the inference model used as the
-generative model (JC69, strict, Yule). The inference model that had the
+models (Yule and Birth-Death), except the inference model used as the
+generative model (JC69, strict clock, Yule). The inference model that had the
 highest evidence (as shown in Table \ref{tab:true_evidence}) was 
@@
 a relaxed log-normal clock model and a Yule tree model.
-From the 'true alignment' this best candidate inference model is run.
 The resulting posterior trees are shown in the second panel
@@
 
-From this 'best true' posterior (second panel in third row of
-figure \ref{fig:pipeline}), the difference between each of its trees is
+From this 'best true' posterior, the difference between each of its trees is
 compared to the 'true tree' (top-left panel), using the nLTT statistic,
@@
 
-From this 'twin tree', a 'twin alignment' is simulated, using
+From this 'twin tree', a 'twin alignment' was simulated, using
 the JC69 nucleotide substitution model and a strict clock model.
-The resulting alignment is shown at the center-right
+The resulting alignment is shown in the center-right panel
 of figure \ref{fig:pipeline}.
@@
 substitution model, a strict clock model and a Yule tree model.
-The resulting posterior trees are shows in the
+The resulting posterior trees are shown in the
 third panel of the third row of figure \ref{fig:pipeline}.
-
-From this 'generative twin' posterior (third panel in 
-third row of figure \ref{fig:pipeline}), 
+From this 'generative twin' posterior, 
 the difference between each of its trees is
 compared to the 'twin tree' (top-right panel), using the nLTT statistic,
-resulting in the error distribution shown at the third panel
+resulting in the error distribution shown in the third panel
 of the bottom row of figure \ref{fig:pipeline}.
@@
 a strict clock model and a BD tree model (note: this indeed matches how
-the twin tree and twin tree have been simulated).
+the twin tree and twin tree were simulated).
 From the 'twin alignment' this best candidate inference model is run.
@@
 From this 'best twin' posterior (fourth in third row of
-figure \ref{fig:pipeline}), the difference between each of its trees is
+figure \ref{fig:pipeline}), the difference between each of its trees was
 compared to the 'twin tree' (top-right panel), using the nLTT statistic,
@@
   we can observe that the discrepancy between the true and twin error 
-  distributions tend to become smaller as the number of nucleotides increase (see also Fig. \ref{fig:error_vs_seqlength}). 
+  distributions tends to become smaller as the number of nucleotides increase (see also Fig. \ref{fig:error_vs_seqlength}).
   This occurred for both the generative and best candidate cases.
@@
 \new{
-  Since the two models are very similar to each other (the BD model can be 
+  Because the two models are very similar to each other (the BD model can be 
   turned into a Yule model just by setting the extinction parameter to 
@@
   case (subsection \ref{subsec:simplest_correct_parameterization}), 
-  where a Yule tree prior has been used, 
+  where a Yule tree prior was used, 
   the distributions here exhibit a greater difference.
@@
   ones relative to the full dataset of \ref{subsec:distribution}. We can also 
-  notice that in the case of median likelihood, the twin median error appear 
-  to be lower than the true mean error. This usually is a sign that the number 
+  notice that in the case of median likelihood, the twin median error appears 
+  to be lower than the true mean error. This is usually a sign that the number 
   of replicates (in this case 10) is too low to allow us to draw precise 
-  conclusions from this test. We decided to not explore further in this 
+  conclusions from this test. We did not explore further in this 
   direction using more simulations because computational times turned to be 

### 2020-07-26 pirouette_supplement.tex
Fix figures, thanks @rsetienne

@@
 generative model (JC69, strict clock, Yule). The inference model that had the
-highest evidence (as shown in Table \ref{tab:true_evidence}) was 
+highest evidence (as shown in Table \ref{tab:evidence_true}) was 
 the inference model with a JC69 nucleotide substitution model,
@@
 set of 15 candidate models.The inference model that had the
-highest evidence (as shown in Table \ref{tab:twin_evidence}) was 
+highest evidence (as shown in Table \ref{tab:evidence_twin}) was 
 the inference model with a JC69 nucleotide substitution model,
@@
 the 4 cases shown in Fig. \ref{fig:example_30_full_pipeline}: 
-"true" pipeline with generative model (Table \ref{tab:esses_true_gen}), 
-"true" pipeline with best candidate model (Table \ref{tab:esses_true_best}), 
+"true" pipeline with generative model (Table \ref{tab:esses_gen}), 
+"true" pipeline with best candidate model (Table \ref{tab:esses_best}), 
 "twin" pipeline with generative model (Table \ref{tab:esses_twin_gen}), 
@@
 We also report the marginal likelihood (or evidence) data for model selection 
-performed both in the "true" (Table \ref{tab:true_evidence}) and 
-"twin" pipeline (Table \ref{tab:twin_evidence}).
+performed both in the "true" (Table \ref{tab:evidence_true}) and 
+"twin" pipeline (Table \ref{tab:evidence_twin}).
 
@@
 
-\input{pirouette_example_30/example_30_314/esses_gen.latex}
-
-\newpage
-
-\input{pirouette_example_30/example_30_314/esses_best.latex}
-
-\newpage
-
-\input{pirouette_example_30/example_30_314/esses_twin_gen.latex}
-
-\newpage
-
-\input{pirouette_example_30/example_30_314/esses_twin_best.latex}
-
-\newpage
-
-\input{pirouette_example_30/example_30_314/evidence_true.latex}
-
-\newpage
-
-\input{pirouette_example_30/example_30_314/evidence_twin.latex}
+\input{pirouette_example_30/example_30/esses_gen.latex}
+
+\newpage
+
+\input{pirouette_example_30/example_30/esses_best.latex}
+
+\newpage
+
+\input{pirouette_example_30/example_30/esses_twin_gen.latex}
+
+\newpage
+
+\input{pirouette_example_30/example_30/esses_twin_best.latex}
+
+\newpage
+
+\input{pirouette_example_30/example_30/evidence_true.latex}
+
+\newpage
+
+\input{pirouette_example_30/example_30/evidence_twin.latex}
 

### 2020-07-26 pirouette_example_30/example_30.R

@@ -1,0 +1,76 @@
+# Works under Linux and MacOS only
+# pirouette example 30:
+# create one exemplary DD tree, as used in the pirouette article
+library(pirouette)
+
+# Constants
+example_no <- 30
+folder_name <- file.path(paste0("example_", example_no))
+crown_age <- 10
+n_taxa <- 6
+rng_seed <- 314
+
+# Create phylogenies
+set.seed(rng_seed)
+
+# Nice, but does not match example 30
+phylogeny <- create_exemplary_dd_tree(n_taxa = n_taxa, crown_age = crown_age)
+
+# Create simulation function like example 28
+sim_dd_tree_fun <- function(crown_age) {
+  extinction_rate <- 0.1
+  n_taxa <- 6
+  n_0 <- 2 # Initial number of species at stem/crown of tree
+  diff <- (log(n_taxa) - log(n_0)) / crown_age
+  speciation_rate <- 3.0 * (diff + extinction_rate)
+  carrying_capacity <- n_taxa # clade-level
+  dd_parameters <- c(speciation_rate, extinction_rate, carrying_capacity)
+  ddmodel <- 1 # linear dependence in speciation rate with parameter K
+  dd_sim_result <- DDD::dd_sim(pars = dd_parameters, age  = crown_age, ddmodel = ddmodel)
+  phylogeny <- dd_sim_result$tes # Only extant species
+  phylogeny
+}
+sim_tree_fun <- pryr::partial(
+  sim_dd_tree_fun,
+  crown_age = crown_age
+)
+phylogeny <- sim_tree_fun()
+
+
+# Setup pirouette
+pir_params <- create_std_pir_params(folder_name = folder_name)
+if (is_on_ci()) {
+  pir_params <- shorten_pir_params(pir_params)
+}
+
+# Generative experiment
+pir_params$experiments[[1]]$inference_model$tree_prior <- create_yule_tree_prior()
+pir_params$experiments[[1]]$inference_model$site_model <- create_jc69_site_model()
+pir_params$experiments[[1]]$inference_model$clock_model <- create_strict_clock_model()
+
+# Simulate a twin tree using BD
+pir_params$twinning_params$sim_twin_tree_fun <- get_sim_bd_twin_tree_fun()
+
+check_pir_params(pir_params)
+
+
+# Do the runs
+pir_out <- pir_run(
+  phylogeny,
+  pir_params = pir_params
+)
+
+# Save data
[…]

### 2020-07-26 pirouette_example_30/scripts/rerun.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+#
+# Re-run the code locally, to re-create the data and figure.
+#
+# Usage:
+#
+#   ./scripts/rerun.sh
+#
+#SBATCH --partition=gelifes
+#SBATCH --time=2:00:00
+#SBATCH --nodes=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --ntasks=1
+#SBATCH --mem=10G
+#SBATCH --job-name=pirex30
+#SBATCH --output=example_30.log
+#
+rm -rf example_30
+rm errors.png
+time Rscript example_30.R
+zip -r pirouette_example_30.zip example_30 example_30.R scripts errors.png
+

### 2020-07-26 scripts/code.R

@@ -1,62 +1,0 @@
-remotes::install_github("richelbilderbeek/pirouette")
-library(pirouette)
-install_beast2()
-install_beast2_pkg("NS")
-phylogeny <- create_yule_tree(n_taxa = 6, crown_age = 10)
-alignment_params <- create_alignment_params(
-  sim_tral_fun = get_sim_tral_with_std_nsm_fun(
-    mutation_rate = 0.1,
-    site_model = create_jc69_site_model()
-  ),
-  root_sequence = create_blocked_dna(length = 1000)
-)
-generative_experiment <- create_experiment(
-  inference_conditions = create_inference_conditions(
-    model_type = "generative", 
-    run_if = "always"
-  ), 
-  inference_model = create_inference_model(
-    tree_prior = create_yule_tree_prior(),
-    clock_model = create_strict_clock_model(), 
-    site_model = create_jc69_site_model(),
-    mcmc = create_mcmc()
-  )
-)
-experiments <- list(generative_experiment)
-error_measure_params <- create_error_measure_params(
-  error_fun = get_nltt_error_fun(),
-  burn_in_fraction = 0.1
-)
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  error_measure_params = error_measure_params
-)
-errors <- pir_run(
-  phylogeny = phylogeny,
-  pir_params = pir_params
-)
-pir_plot(errors)
-candidate_experiments <- create_all_experiments(
-  exclude_model = generative_experiment$inference_model
-)
-experiments <- c(
-  list(generative_experiment),
-  candidate_experiments
-)
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments
-)
-phylogeny  <- ape::read.tree(
-  text = "(((A:8, B:8):1, C:9):1, ((D:8, E:8):1, F:9):1);"
-)
-twinning_params <- create_twinning_params(
-  sim_twin_tree_fun = get_sim_bd_twin_tree_fun(),
-  sim_twal_fun = get_sim_twal_with_std_nsm_fun()
-)
-pir_params <- create_pir_params(
-  alignment_params = alignment_params,
-  experiments = experiments,
-  twinning_params = twinning_params
-)

### 2020-07-26 scripts/fix_tables.R

@@ -1,0 +1,32 @@
+# Fix tables by adding the labels if these are absent
+
+
+tex_filenames <- list.files(
+  "pirouette_example_30",
+  recursive = TRUE,
+  pattern = "latex",
+  full.names = TRUE
+)
+
+testthat::expect_true(length(tex_filenames) > 2)
+
+for (tex_filename in tex_filenames) {
+  text <- readr::read_lines(tex_filename)
+
+  # Only add label when it is absent
+  if (length(stringr::str_which(text, "\\label\\{tab:.*\\}")) == 1) {
+    message("File '", tex_filename, "' already had a label")
+    next()
+  }
+
+
+  last_line <- tail(text, n = 1)
+  label <- paste0(
+    "  \\label{tab:", beautier::get_file_base_sans_ext(tex_filename),"}"
+  )
+  text[length(text)] <- label
+  text <- c(text, last_line)
+  readr::write_lines(x = text, path = tex_filename)
+  message("File '", tex_filename, "' now has a label")
+}
+

### 2020-07-28 pirouette_supplement.tex
Address more points of confusion

@@
 
-To assess if the results of the inference are meaningful one important 
-parameter is the Effective Sample Size (ESS). This quantity describes how 
-many independent trees are sampled from the posterior distributions. 
-For reliable results it is good practice to have at 
-least $ESS = 200$ (see 
-\begin{sloppypar}
-  \url{https://beast.community/ess_tutorial}).
-\end{sloppypar}
-In the following we present the ESS for the posterior distributions of 
-the 4 cases shown in Fig. \ref{fig:example_30_full_pipeline}: 
-"true" pipeline with generative model (Table \ref{tab:esses_gen}), 
-"true" pipeline with best candidate model (Table \ref{tab:esses_best}), 
-"twin" pipeline with generative model (Table \ref{tab:esses_twin_gen}), 
-"twin" pipeline with best candidate model (Table \ref{tab:esses_twin_best}). 
-We also report the marginal likelihood (or evidence) data for model selection 
-performed both in the "true" (Table \ref{tab:evidence_true}) and 
-"twin" pipeline (Table \ref{tab:evidence_twin}).
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\newpage
+To assess if the results of the inference are meaningful one important 
+parameter is the Effective Sample Size (ESS). This quantity describes how 
+many independent trees are sampled from the posterior distributions. 
+For reliable results it is good practice to have at 
+least $ESS = 200$ (see 
+\begin{sloppypar}
+  \url{https://beast.community/ess_tutorial}).
+\end{sloppypar}
+In the following we present the ESS for the posterior distributions of 
+the 4 cases shown in Fig. \ref{fig:example_30_full_pipeline}.
+
+The ESSes of the 'true' pipeline for the generative model
+are shown in Table \ref{tab:esses_gen}.
+From the estimated parameters, one can deduce that
+the JC69 nucleotide substitution model was used (no
+estimated parameter needed), a strict clock model was 
+used (again, no parameter needed to be estimated)
+and a Yule tree prior is used ('Yule model' and 'birthRate' are estimated).
+Note that although the actual true tree is created by a DD process, 
+the default and standard Yule tree model is used as the closest
+standard tree model.
 
@@
 
+The ESSes of the 'twin' pipeline for the generative model
+are shown in Table \ref{tab:esses_twin_gen}.
+Note that the generative inference model is 
+re-used (which assumes a Yule tree model) in the inference, 
+where the twin tree is actually created using a BD process,
+which is the default.
+
+\input{pirouette_example_30/example_30/esses_twin_gen.latex}
+
+\newpage
+
+The ESSes of the 'true' pipeline for the best candidate model
+are shown in Table \ref{tab:esses_best}. 
+From the names of the estimated parameters, it is clear that
+the best candidate model has had a TN93 nucleotide substitution 
+model (due to 'kappa1', 'kappa2', 'freqParameter.1', 
+'freqParameter.2', 'freqParameter.3', 'freqParameter.4')
+a strict clock (no parameter needed to be estimated) and a BD 
+model ('BirthDeath', 'BDBirthRate' and 'BDDeathRate').
+
 \input{pirouette_example_30/example_30/esses_best.latex}
@@
 
-\input{pirouette_example_30/example_30/esses_twin_gen.latex}
-
-\newpage
+The ESSes of the 'twin' pipeline for the best candidate model
+are shown in Table \ref{tab:esses_twin_best}.
+
+From the names of the estimated parameters, it is clear that
+the best candidate model for the twin tree
+is JC69 nucleotide substitution 
+model (no parameter needed to be estimated),
+a strict clock (no parameter needed to be estimated) and a BD 
+model ('BirthDate', 'BDBirthRate' and 'BDDeathRate').
+Note that this matches the actual process of how the twin tree
+and twin alignment are generated.
 
@@
 
+The marginal likelihood (or evidence) data for the model comparison
+performed in the 'true' pipeline is shown in Table \ref{tab:evidence_true}.
+The best (that is, the one with the highest model weight)
+candidate model assumes a TN93 nucleotide substitution model,
+a strict clock and a BD model.
+
 \input{pirouette_example_30/example_30/evidence_true.latex}
@@
 \newpage
+
+The marginal likelihood (or evidence) data for model comparison
+performed in the 'twin' pipeline is shown in Table \ref{tab:evidence_twin}.
+The best (that is, the one with the highest model weight)
+candidate model assumes a JC69 nucleotide substitution model,
+a strict clock and a BD model.
+Note that this matches the actual process of how the twin tree
+and twin alignment are generated.
 

### 2020-07-28 pirouette_article.tex
Fix figures, thanks @rsetienne
Addressed all feedback, thanks @rsetienne, fix #105

@@
 The candidate model that had highest evidence given the simulated alignment,
-was JC69, RLN and Yule (see Table \ref{tab:options} for the meaning of these 
-abbreviations). JC69 was indeed the site model used when simulating the
-alignments. The RLN clock model is a surprising result: it assumes that mutation rates differ between lineages, whereas the
-alignment was simulated with equal mutation rates for all lineages \new{(strict clock)}.
-\new{The Yule model is not surprising as it is the standard tree prior that is probably closest to the true DD model because it shows no pull-of-the-present (but also no slowdown).}
+was TN, strict and BD (see Table \ref{tab:options} for the meaning of these 
+abbreviations). 
+The TN site model is a surprising result: it assumes nucleotide substitutions 
+occur at different rates between translations and transversions 
+and that nucleotides are available with different frequencies (that is,
+a nucleotide may be rare or overabundant).
+The strict clock model matches the model how the alignment is simulated.
+\new{
+  The Yule model is not surprising as it is the standard tree prior 
+  that is probably closest to the true DD model because it shows 
+  no pull-of-the-present (but also no slowdown).
+}
 

### 2020-07-28 pirouette_article.tex
Process feedback. Thanks @rsetienne. Fix #111

@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_30/errors.png}
+  \includegraphics[width=0.9\textwidth]{pirouette_example_30/errors.png}
   \caption{
-    \new{The impact of the tree prior for the example tree in Figure \ref{fig:dd_tree}.
-The alignment for this true tree was generated using a JC69 substitution model and strict clock model.
-For inferring the tree from this alignment in the 'generative' scenario the same substitution and clock models were used, and a Yule tree prior (this is the assumed generative model because the real generative model is of course unknown).
-In the 'best' scenario the best-fitting candidate models (JC69 substitution model, relaxed log-normal clock model and Yule tree prior) were used.
-For the twin tree, the same models were chosen, except that BD was used to generate and infer the twin tree.
-    The twin distributions show the baseline inference error.}
+    \new{
+      The impact of the tree prior for 
+      the example tree in Figure \ref{fig:dd_tree}. 
+      The alignment for this true tree was generated using a JC substitution model 
+      and strict clock model. 
+      For inferring the tree from this alignment in the 'generative' scenario 
+      the same substitution and clock models were used, 
+      and a Yule tree prior (this is the assumed generative model,
+      because the real generative model is of course unknown). 
+      In the 'best' scenario the best-fitting candidate 
+      models (TN substitution model, strict clock model and Yule tree prior) 
+      were used. For the twin tree, the same models were chosen, 
+      except that BD was used to generate and infer the twin tree.
+      The twin distributions show the baseline inference error.
+    }
     Vertical dashed lines show the median error value per distribution.

### 2020-07-28 pirouette_supplement.tex
Process feedback. Thanks @rsetienne. Fix #111

@@
 From this 'true tree', a 'true alignment' is simulated, using
-the JC69 nucleotide substitution model and a strict clock model.
+the JC nucleotide substitution model and a strict clock model.
 The resulting alignment is shown at the center-left
@@
 Of course, it cannot be the actual (DD) model. Instead, the
-default BEAST2 inference model is used, which assumes a JC69 nucleotide
+default BEAST2 inference model is used, which assumes a JC nucleotide
 substitution model, a strict clock model and a Yule tree model.
@@
 15 models. The set of models consists of all combinations
-of all 4 nucleotides substitution models (JC69, HKY, TN93, GTR),
+of all 4 nucleotides substitution models (JC, HKY, TN, GTR),
 all 2 clock models (strict and relaxed log-normal) and 2 birth-death
 models (Yule and Birth-Death), except the inference model used as the
-generative model (JC69, strict clock, Yule). The inference model that had the
+generative model (JC, strict clock, Yule). The inference model that had the
 highest evidence (as shown in Table \ref{tab:evidence_true}) was 
-the inference model with a JC69 nucleotide substitution model,
+the inference model with a JC nucleotide substitution model,
 a relaxed log-normal clock model and a Yule tree model.
@@
 From this 'twin tree', a 'twin alignment' was simulated, using
-the JC69 nucleotide substitution model and a strict clock model.
+the JC nucleotide substitution model and a strict clock model.
 The resulting alignment is shown in the center-right panel
@@
 Also here, the default BEAST2 inference model is used, 
-which assumes a JC69 nucleotide
+which assumes a JC nucleotide
 substitution model, a strict clock model and a Yule tree model.
@@
 highest evidence (as shown in Table \ref{tab:evidence_twin}) was 
-the inference model with a JC69 nucleotide substitution model,
+the inference model with a JC nucleotide substitution model,
 a strict clock model and a BD tree model (note: this indeed matches how
@@
 From the estimated parameters, one can deduce that
-the JC69 nucleotide substitution model was used (no
+the JC nucleotide substitution model was used (no
 estimated parameter needed), a strict clock model was 
@@
 From the names of the estimated parameters, it is clear that
-the best candidate model has had a TN93 nucleotide substitution 
+the best candidate model has had a TN nucleotide substitution 
 model (due to 'kappa1', 'kappa2', 'freqParameter.1', 
@@
 the best candidate model for the twin tree
-is JC69 nucleotide substitution 
+is JC nucleotide substitution 
 model (no parameter needed to be estimated),
@@
 The best (that is, the one with the highest model weight)
-candidate model assumes a TN93 nucleotide substitution model,
+candidate model assumes a TN nucleotide substitution model,
 a strict clock and a BD model.
@@
 The best (that is, the one with the highest model weight)
-candidate model assumes a JC69 nucleotide substitution model,
+candidate model assumes a JC nucleotide substitution model,
 a strict clock and a BD model.

### 2020-07-28 scripts/create_figure_2.R
Prettier tree

@@ -5,7 +5,7 @@
 #
 #   Rscript scripts/create_figure_2.R
 
-newick_filename <- "pirouette_example_2/example_2_314/true_tree.newick"
+newick_filename <- "../pirouette_example_30/example_30/phylogeny.newick"
 
 library(testthat)
 expect_true(file.exists(newick_filename))

### 2020-07-29 pirouette_supplement.tex
Use 'TESS' in the bibliography (that is, only the package)
Cite Houhna's paper again
Add note in supmat what is needed in a thesis to have the refs refer properly
Add LaTeX comments to ensure a correct thesis
Changes @rsetienne

@@
 \verb;stringr; [\cite{stringr}],
-\verb;TESS; [\cite{TESS}],
+\verb;TESS; [\cite{TESS,hohna2016tess}],
 \verb;testit; [\cite{testit}], 
@@
 We simulate the new branching times using the TESS 
-package [\cite{TESS, hohna2016tess}].
+package [\cite{TESS}].
 \new{
@@
 the best candidate model has had a TN nucleotide substitution 
-model (due to 'kappa1', 'kappa2', 'freqParameter.1', 
+model (which can be inferred from the parameters'kappa1', 'kappa2', 'freqParameter.1', 
 'freqParameter.2', 'freqParameter.3', 'freqParameter.4')

### 2020-07-29 pirouette_article.tex
Prettier tree

@@
 \begin{figure}[H]
-  \includegraphics[width=\textwidth]{pirouette_example_30/example_30/true_tree.png}
+  \includegraphics[width=\textwidth]{figure_2.png}
   \caption{
@@
       were used. For the twin tree, the same models were chosen, 
-      except that BD was used to generate and infer the twin tree.
+      except that BD was used to generate the twin tree.
       The twin distributions show the baseline inference error.
@@
 the known generative substitution and clock models, and the tree model that is assumed in inference of the true tree, and the tree model that is used for generating and inferring the twin tree)}
-differ strongly between the true and twin tree. 
+differ substantially between the true and twin tree. 
 This difference shows the extent of the mismatch between
@@
 the inference error made when using an 
-incorrect tree prior on a DD tree is profound.
+incorrect tree prior on a DD tree is quite profound.
 
@@
 
-The candidate model that had highest evidence given the simulated alignment,
+\new{The candidate model that had highest evidence given the simulated alignment,
 was TN, strict and BD (see Table \ref{tab:options} for the meaning of these 
@@
 a nucleotide may be rare or overabundant).
-The strict clock model matches the model how the alignment is simulated.
-\new{
-  The Yule model is not surprising as it is the standard tree prior 
-  that is probably closest to the true DD model because it shows 
-  no pull-of-the-present (but also no slowdown).
+The strict clock model matches the model used to simulate the alignment.
+The BD model is perhaps somewhat surprising for the true tre because the other alteranative standard tree prior, Yule, is probably closest to the true DD model because it shows no pull-of-the-present (but also no slowdown).
 }

### 2020-07-29 pirouette_supplement.tex
Address that 'create_exemplary_dd_tree' creates trees with different numbers of tips

@@
 University of Groningen. Due to this, for example, a run of 40 taxa only
-has 6 replicates, because one run takes 4 hours. For all experiments, the intermediate results can all be downloaded 
+has few replicates, because one run takes 4 hours. For all experiments, the intermediate results can all be downloaded 
 from their respective websites, which is approximately 5 gigabyte in total.
@@
 with (an arbitrarily chosen) crown age of 10 time units 
-and 6 tips for an extinction rate of 0.1. 
+and an expected number of 6 tips for an extinction rate of 0.1. 
 The carrying-capacity is set to 6. 
@@
 would be equal to the number of tips, which amounts to $\lambda_0 = 0.63$.
+Note that in the main example, a tree was generated with 5 tips, due
+to stochasticity in the tree generation algorithm.
 
@@
 
-The main example uses 6 taxa. Here we show the same results as the main example,
+The main example uses 5 taxa. Here we show the same results as the main example,
 except for a varying number of taxa. We did so, by setting the DD model's
@@
   \ref{fig:example_40_taxa} what are the errors obtained when starting from 
-  phylogenies with, respectively, 6, 12, 24, 32 and 40 taxa.
+  phylogenies with, respectively, 5, 12, 24, 32 and 40 taxa.
   Again we can see 
@@
 \begin{sloppypar}
-  \url{https://github.com/richelbilderbeek/pirouette_example_28} (6 taxa, main example),
+  \url{https://github.com/richelbilderbeek/pirouette_example_28} (5 taxa, main example),
   \url{https://github.com/richelbilderbeek/pirouette_example_32} (12 taxa),
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% MEE style
+% Keep the next commented LaTeX commands
+%
+% These two lines must be added in a thesis, 
+%   so that the references have a proper link.
+% This is done by sed in the Makefile of github.com/richelbilderbeek/thesis
+%
 %\bibliographystyle{pirouette_mee}
-% \bibliography{pirouette_supplement}
 %\bibliography{pirouette_article}

### 2020-07-29 scripts/create_figure_2.R
5 taxa, different creation of figure 2

@@ -13,8 +13,8 @@ expect_true(file.exists(newick_filename))
 phylogeny <- ape::read.tree(file = newick_filename)
 cex <- 1.5
 edge.width <- 2
-x <- 0.3
-y <- 1.38
+x <- 0.5
+y <- 2.5
 lwd <- 2
 label.offset <- 0.25
 ape::plot.phylo(phylogeny, cex = cex, edge.width = edge.width, label.offset = label.offset)

### 2020-07-29 pirouette_article.tex
Address that 'create_exemplary_dd_tree' creates trees with different numbers of tips
Processed all feedback. Thanks @rsetienne. Fix #113

@@
 
-# Create a DD phylogeny with 6 taxa and a crown age of 10
+# Create a DD phylogeny with 5 taxa and a crown age of 10
 phylogeny <- create_exemplary_dd_tree()
@@
       and strict clock model. 
-      For inferring the tree from this alignment in the 'generative' scenario
+      For inferring the tree from this alignment in the 'generative' scenario,
       the same substitution and clock models were used, 
@@
       because the real generative model is of course unknown).
-      In the 'best' scenario the best-fitting candidate 
-      models (TN substitution model, strict clock model and Yule tree prior) 
-      were used. For the twin tree, the same models were chosen, 
-      except that BD was used to generate the twin tree.
+      For the twin tree, the same inference model is used.
+      In the 'best' scenarios, for the 'true' tree, 
+      the best-fitting candidate models on the 'true' tree
+      (JC substitution model, RLN clock model and BD tree prior) 
+      was used. 
+      In the 'best' scenarios, for the 'twin' tree, 
+      the best-fitting candidate models on the 'twin' tree
+      (JC substitution model, RLN clock model and Yule tree prior) 
+      was used. 
       The twin distributions show the baseline inference error.
@@
 In the upper panel of Figure \ref{fig:example_30},
-we can see that the error distributions of the \new{ (assumed) generative model (i.e.
-the known generative substitution and clock models, and the tree model that is assumed in inference of the true tree, and the tree model that is used for generating and inferring the twin tree)}
+we can see that the error distributions of the 
+\new{ 
+  (assumed) generative model (i.e. the known generative substitution and 
+  clock models, and the tree model that is assumed in inference of the true 
+  tree, and the tree model that is used for generating and inferring the 
+  twin tree)
+}
 differ substantially between the true and twin tree. 
@@
 Comparing the upper and lower panel of Figure \ref{fig:example_30}, 
-we can see that the best
-candidate model is only slightly better at inferring the true tree,
+we can see that 
+\new{
+  the best candidate model is slightly worse at inferring the true tree,
   than the (assumed) generative model, 
-thereby showing that they cannot compensate 
-for the true generative tree model not being among the inference models.
+  indicating that the generative inference model we selected is a good pick.
+}
 
-\new{The candidate model that had highest evidence given the simulated alignment,
-was TN, strict and BD (see Table \ref{tab:options} for the meaning of these 
+\new{
+  The candidate model that had highest evidence given the simulated alignment,
+  was JC, RLN and BD (see Table \ref{tab:options} for the meaning of these 
   abbreviations). 
-The TN site model is a surprising result: it assumes nucleotide substitutions 
-occur at different rates between translations and transversions 
-and that nucleotides are available with different frequencies (that is,
-a nucleotide may be rare or overabundant).
-The strict clock model matches the model used to simulate the alignment.
-The BD model is perhaps somewhat surprising for the true tre because the other alteranative standard tree prior, Yule, is probably closest to the true DD model because it shows no pull-of-the-present (but also no slowdown).
+  The RLN clock model is a surprising result: it assumes nucleotide substitutions 
+  occur at different rates between the taxa.
+  The JC nucleotide substitution model matches the model used to simulate the alignment.
+  The BD model is perhaps somewhat surprising for the true tree, 
+  because the other alternative standard tree prior, 
+  Yule, is probably closest to the true 
+  DD model because it shows no pull-of-the-present (but also no slowdown).
 }

### 2020-07-29 pirouette_supplement.tex
Processed all feedback. Thanks @rsetienne. Fix #113

@@
 models (Yule and Birth-Death), except the inference model used as the
-generative model (JC, strict clock, Yule). The inference model that had the
-highest evidence (as shown in Table \ref{tab:evidence_true}) was 
+generative model (JC, strict clock, Yule). 
+The inference model that had the highest evidence 
+(as shown in Table \ref{tab:evidence_true}) was 
 the inference model with a JC nucleotide substitution model,
-a relaxed log-normal clock model and a Yule tree model.
+an RLN clock model and a BD tree model.
 The resulting posterior trees are shown in the second panel
@@
 the inference model with a JC nucleotide substitution model,
-a strict clock model and a BD tree model (note: this indeed matches how
+an RLN clock model and a Yule tree model (note: this indeed matches how
 the twin tree and twin tree were simulated).
@@
 From the names of the estimated parameters, it is clear that
-the best candidate model has had a TN nucleotide substitution 
-model (which can be inferred from the parameters'kappa1', 'kappa2', 'freqParameter.1', 
-'freqParameter.2', 'freqParameter.3', 'freqParameter.4')
-a strict clock (no parameter needed to be estimated) and a BD 
-model ('BirthDeath', 'BDBirthRate' and 'BDDeathRate').
+the best candidate model has had a JC nucleotide substitution 
+model (no parameter needed to be estimated)
+an RLN clock model (which can be inferred from the parameter 'rate.mean') 
+and a BD model ('BirthDeath', 'BDBirthRate' and 'BDDeathRate').
 
@@
 model (no parameter needed to be estimated),
-a strict clock (no parameter needed to be estimated) and a BD 
-model ('BirthDate', 'BDBirthRate' and 'BDDeathRate').
-Note that this matches the actual process of how the twin tree
-and twin alignment are generated.
+an RLN clock model (which can be inferred from the parameter 'rate.mean') 
+and a Yule model ('YuleModel', 'birthRate').
+Note that there is a mismatch between the actual process of how the twin tree
+and twin alignment are generated, as the twin tree is generated by a
+BD process, and the alignment is simulated using a JC nucleotide
+substitution model and a strict clock model.
 
@@
 The best (that is, the one with the highest model weight)
-candidate model assumes a TN nucleotide substitution model,
-a strict clock and a BD model.
+candidate model assumes a JC nucleotide substitution model,
+an RLN clock and a BD tree model.
 
@@
 candidate model assumes a JC nucleotide substitution model,
-a strict clock and a BD model.
-Note that this matches the actual process of how the twin tree
-and twin alignment are generated.
+an RLN clock and a Yule tree model.
+Note that there is a mismatch between the actual process of how the twin tree
+and twin alignment are generated, as the twin tree is generated by a
+BD process, and the alignment is simulated using a JC nucleotide
+substitution model and a strict clock model.
 

### 2020-07-29 pirouette_article.tex
Reviewed changes for review

@@
       For the twin tree, the same inference model is used.
-      In the 'best' scenarios, for the 'true' tree, 
-      the best-fitting candidate models on the 'true' tree
+      In the 'best' scenario, for the true tree, 
+      the best-fitting candidate models on the true tree
       (JC substitution model, RLN clock model and BD tree prior) 
       was used. 
-      In the 'best' scenarios, for the 'twin' tree, 
-      the best-fitting candidate models on the 'twin' tree
+      In the 'best' scenario, for the twin tree, 
+      the best-fitting candidate models on the twin tree
       (JC substitution model, RLN clock model and Yule tree prior) 

### 2020-07-29 pirouette_supplement.tex
Reviewed changes for review

@@
 would be equal to the number of tips, which amounts to $\lambda_0 = 0.63$.
+\new{
   Note that in the main example, a tree was generated with 5 tips, due
   to stochasticity in the tree generation algorithm.
+}
 
@@
 the inference model with a JC nucleotide substitution model,
-an RLN clock model and a Yule tree model (note: this indeed matches how
-the twin tree and twin tree were simulated).
+an RLN clock model and a Yule tree model 
+\new{
+  (note: this mismatches how
+  the twin tree and twin tree were simulated, as the twin tree
+  is generated using a BD process, where the alignment
+  is simulated using a JC substitution model and strict clock model
+  ).
+}
 From the 'twin alignment' this best candidate inference model is run.

### 2020-07-30 pirouette_article.tex
Changes @rsetienne

@@
       the example tree in Figure \ref{fig:dd_tree}. 
-      The alignment for this true tree was generated using a JC substitution model 
-      and strict clock model. 
+      The alignment for this true tree was generated using a JC substitution model and strict clock model. 
       For inferring the tree from this alignment in the 'generative' scenario,
@@
       and a Yule tree prior (this is the assumed generative model,
-      because the real generative model is of course unknown).
-      For the twin tree, the same inference model is used.
+      because the real generative model is assumed to be unknown).
+      For the twin tree, the same inference models were used.
       In the 'best' scenario, for the true tree, 
-      the best-fitting candidate models on the true tree
-      (JC substitution model, RLN clock model and BD tree prior) 
-      was used. 
-      In the 'best' scenario, for the twin tree, 
-      the best-fitting candidate models on the twin tree
-      (JC substitution model, RLN clock model and Yule tree prior) 
-      was used. 
+      the best-fitting candidate models were JC substitution model, RLN clock model and BD tree prior, while for the twin tree, the best-fitting candidate models were JC substitution model, RLN clock model and Yule tree prior.
       The twin distributions show the baseline inference error.
@@
   than the (assumed) generative model, 
-  indicating that the generative inference model we selected is a good pick.
+  indicating that the generative inference model we selected is a good choice.
 }
@@
 The clear difference between the error distributions 
-for the true tree and the twin tree suggests 
-that the choice of tree prior matters.
+for the true tree and the twin tree suggests that the choice of tree prior matters.
 We note, however, that only one tree from a novel tree model
@@
 Like most phylogenetic experiments, the setup of \verb;pirouette;
-involves many choices. A prime example is the
-length of the simulated DNA sequence. One expects that the inference error
-decreases for longer DNA sequences. We investigated this
-superficially and confirmed this prediction (see the supplementary material). 
+involves many choices. A prime example is the length of the simulated DNA sequence. One expects that the inference error
+decreases for longer DNA sequences. We investigated this superficially and confirmed this prediction (see the supplementary material). 
 However, we note that for longer DNA sequences, the assumption 
@@
 a promising starting point, as it can compare any two trees and 
-results in an error distribution of known range, but one may also explore other statistics.
-In principle, \verb;pirouette; allows for this, but in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.
+results in an error distribution of known range, but one may also explore other statistics, for example statistics that depend on the topology of the tree,
+While \verb;pirouette; allows for this in principle, in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.
 \new{
@@
 
-In conclusion, \verb;pirouette; can show the errors to be expected
+In conclusion, \verb;pirouette; can show the errors in phylogenetic reconstruction expected
 when the model assumed in inference is different from the actual generative model.
 The user can then judge whether or not this new model 
-should be implemented in the Bayesian phylogenetic tool.
+should be implemented in a Bayesian phylogenetic tool.
 

### 2020-07-30 pirouette_supplement.tex
Changes @rsetienne

@@
 are shown in Table \ref{tab:esses_twin_gen}.
-Note that the generative inference model is 
-re-used (which assumes a Yule tree model) in the inference, 
+Note that the generative inference model is re-used (which assumes a Yule tree model) in the inference, 
 where the twin tree is actually created using a BD process,
@@
 From the names of the estimated parameters, it is clear that
-the best candidate model has had a JC nucleotide substitution 
+the best candidate model has a JC nucleotide substitution 
 model (no parameter needed to be estimated)
 an RLN clock model (which can be inferred from the parameter 'rate.mean') 
-and a BD model ('BirthDeath', 'BDBirthRate' and 'BDDeathRate').
+and a BD tree prior ('BirthDeath', 'BDBirthRate' and 'BDDeathRate').
 

### 2020-07-31 pirouette_supplement.tex
Changes @rsetienne

@@
 with the highest marginal likelihood is determined, from the same 
-set of 15 candidate models.The inference model that had the
+set of 15 candidate models. The inference model that had the
 highest evidence (as shown in Table \ref{tab:evidence_twin}) was 
@@
 \new{
-  (note: this mismatches how
-  the twin tree and twin tree were simulated, as the twin tree
-  is generated using a BD process, where the alignment
-  is simulated using a JC substitution model and strict clock model
-  ).
+  (Note that this is different from how the twin tree was generated which was with a BD process and the alignment
+  was simulated with a JC substitution model and strict clock model
+  ). However, the extinction rate used in simulating the twin tree was practically 0, thus resembling a Yule process.
 }
@@
 substitution model and a strict clock model.
+Again we note that the extinction rate used to simulate the twin tree (estimated from the true tree) was practically 0, so the BD process resembled a Yule process.
 
@@
 BD process, and the alignment is simulated using a JC nucleotide
-substitution model and a strict clock model.
+substitution model and a strict clock model. The extinction rate in simulating the BD process (estimated from the true tree) was, however, practically 0, so the BD process resembled a Yule process.
 

### 2020-08-01 pirouette_article.tex
Changes @rsetienne
Remove 'new' tags, fix #122

@@
   \textbf{3. }
-    We describe \verb;pirouette;'s usage \new{providing full examples in which 
-    we interrogate a model for its power to describe another}. \\
+    We describe \verb;pirouette;'s usage providing full examples in which 
+    we interrogate a model for its power to describe another. \\
   \textbf{4. }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\new{
 The development of new powerful Bayesian phylogenetic inference tools,
@@
 and modes of diversification.
-}
 
-\new{
 BEAST [\cite{drummond2007beast}] is a typical Bayesian phylogenetics tool, 
@@
 while being computationally fast.
-}
 
-\new{
 To allow users to extend the functionalities of BEAST
@@
 the BD SIR model [\cite{kuhnert2014simultaneous}].
-}
 
-\new{
 Many other diversification models (and their associated likelihood algorithms) 
@@
 [\cite{maddison2007estimating, fitzjohn2012diversitree, herrera2019detecting}].
-}
 
-\new{
 These are, however, not yet available as tree priors in BEAST2, 
@@
 but our methods can be used with other built-in tree priors as well.
-}
 
-\new{
 The rationale of our paper is as follows.
-}
 When a novel diversification model is introduced,
@@
 This is not necessarily problematic, when the data are very informative 
-\new{
-  or when the prior is truly uninformative
-}, 
+or when the prior is truly uninformative, 
 as this will reduce the influence of the tree prior.
-However, the assumption that tree prior choice is of low \new{impact} 
+However, the assumption that tree prior choice is of low impact 
 must first be verified.
 
-There have been multiple attempts to investigate the \new{impact} of tree
+There have been multiple attempts to investigate the impact of tree
 prior choice. For example, Sarver and colleagues, [\cite{sarver2019choice}] 
@@
 However, they only compared current diversification models to one another, 
-and thus this does not inform us on the \new{impact} of a new tree prior.
+and thus this does not inform us on the impact of a new tree prior.
 
-\new{
 Similarly,
 Ritchie and colleagues [\cite{Ritchie2017impact}] showed that inference was accurate when birth-death or skyline coalescent priors were used, but they simulated their trees with a Yule process only, as their focus was not so much on the diversification process but on the influence of inter- and intraspecific sampling.
-}
 
-\new{Another way to benchmark a diversification model, is by doing a model comparison, in which the best model is determined
+Another way to benchmark a diversification model, is by doing a model comparison, in which the best model is determined
 from a set of models. A good early example is \citet{goldman1993statistical} 
@@
 model adequacy for phylodynamic models that are
-mathematically described (i.e. have a known likelihood equation).}
+mathematically described (i.e. have a known likelihood equation).
 
-Here we introduce a method to quantify the \new{impact} 
+Here we introduce a method to quantify the impact 
 of a novel tree prior,
-\new{
 i.e., a tree model, for which we can simulate phylogenies, but not yet
 calculate their likelihoods.
-}
-\new{This new method simultaneously assesses the substitution, clock and tree models [\cite{duchene2015evaluating}].}
+This new method simultaneously assesses the substitution, clock and tree models [\cite{duchene2015evaluating}].
 The method starts with a phylogeny generated by the new model. 
@@
 
-The goal of \verb;pirouette; is to quantify the \new{impact of a new} tree prior.
+The goal of \verb;pirouette; is to quantify the impact of a new tree prior.
 It does so by measuring the inference error made 
@@
 as the 'generative tree model' $\mathit{p_{G}}$.
-\new{
 A 'generative tree model', in this paper, can be either the novel diversification model for which we are testing the impact of choosing standard tree priors for, or it is the model with which we generate the twin tree that is needed for comparison (see below).
@@
 This is is done in the example, where the Yule model is the generative model.
-}
 
-\new{
 The inference error we aim to quantify is not of stochastic nature.
@@
 a generative model (that has generated the phylogeny) and the model(s) used in the actual inference.
-}
-We define the birth-death \new{(BD)} model [\cite{nee1994reconstructed}] as
+We define the birth-death (BD) model [\cite{nee1994reconstructed}] as
 the standard tree model, as many (non-standard) tree models 
-have a parameter setting \new{such that it reduces} to this model. 
+have a parameter setting such that it reduces to this model. 
 One such example is the diversity-dependent (DD) diversification 
@@
 on the number of species and a clade-level carrying capacity.
-\new{
 The BD model can be seen as a special case of the DD model, because for
-} 
 an infinite carrying capacity, the DD model reduces to the BD model.
@@
     \verb;mutation_rate; &
-    Pace at which \new{substitutions occur} &
+    Pace at which substitutions occur &
     \verb;mutation_rate; $\in \mathbb{R}_{>0}$\\
@@
 Alternatively, if we want to run only the inference
-model that fits best to an alignment from a set of candidates \new{(regardless of whether these generated the alignments)},
+model that fits best to an alignment from a set of candidates (regardless of whether these generated the alignments),
 one can specify these inference models (see section \ref{subsec:candidates}).
@@
 that is the baseline error of the inference pipeline (see
-supplementary materials, subsection \ref{subsec:twinning} for more details).
-\new{This the error that arises when the models used in inference are identical to the ones used in generating the alignments.}
+supplementary materials, subsection \ref{subsec:twinning} for more details). This the error that arises when the models used in inference are identical to the ones used in generating the alignments.
 The second type of control is the use of candidate models,
@@
 section \ref{subsec:candidates} for more details).
-\new{The underlying idea is that using a substitution model in inference than used in generating the alignment may partly compensate for choosing a standard tree model instead of the generative tree model as tree prior in inference.}
+The underlying idea is that using a substitution model in inference than used in generating the alignment may partly compensate for choosing a standard tree model instead of the generative tree model as tree prior in inference.
 Additionally, multiple \verb;pirouette; runs are needed
@@
   \caption{
-    \new{
     The impact of the tree prior for 
@@
     The twin distributions show the baseline inference error.
-    }
     Vertical dashed lines show the median error value per distribution.
@@
 we can see that the error distributions of the 
-\new{ 
 (assumed) generative model (i.e. the known generative substitution and 
@@
 twin tree)
-}
 differ substantially between the true and twin tree. 
@@
 we can see that 
-\new{
 the best candidate model is slightly worse at inferring the true tree,
@@
 indicating that the generative inference model we selected is a good choice.
-}
 
-\new{
 The candidate model that had highest evidence given the simulated alignment,
@@
 DD model because it shows no pull-of-the-present (but also no slowdown).
-}
 
@@
 
-We showed how to use \verb;pirouette; to quantify the \new{impact of a 
+We showed how to use \verb;pirouette; to quantify the impact of a 
 tree prior in Bayesian phylogenetics, assuming - for illustrative purposes - the simplest standard substitution, clock and
 tree models, but also the models that would be selected among many different standard tree priors according to the highest marginal likelihood, as this would be a likely strategy for an empiricist.
-We recommend exploring different candidate models, but note that this is computationally highly demanding.}
+We recommend exploring different candidate models, but note that this is computationally highly demanding, particularly for large trees.
 
 Figure~\ref{fig:example_30} illustrates the primary result of our pipeline: 
-it shows the error distributions for the true tree and the twin tree \new{
+it shows the error distributions for the true tree and the twin tree
 when either the generative model (for substitution and clock models these are known, for the tree model it must be assumed for the true tree and it is known for the twin tree) or the best-fitting set candidate model (i.e.
-combination of tree model, substitution model and clock model) is used in inference.}
+combination of tree model, substitution model and clock model) is used in inference.
 The clear difference between the error distributions 
@@
 We note, however, that only one tree from a novel tree model
-is not enough to determine the \new{impact} of using an incorrect
+is not enough to determine the impact of using an incorrect
 tree prior. Instead, a distribution 
@@
 While \verb;pirouette; allows for this in principle, in our example we used a diversification model (DD) that only deviates from the Yule and BD models in the temporal branching pattern, not in the topology.
-\new{
 For models that make different predictions on topology, 
 the twinning process should be modified.
-}
-\new{
+
 As noted in the introduction, Duch\^{e}ne and 
@@
 generated with the posterior parameters!).
-    By contrast,}
+By contrast,
 \verb;pirouette;
-	\new{aims to expose when assuming standard priors for the 
+aims to expose when assuming standard priors for the 
 species tree are a mis- or underparameterization. 
@@
 The two methods therefore complement one another.
-}
+
 Furthermore, we note that the \verb;pirouette; pipeline is not restricted 

### 2020-08-01 pirouette_supplement.tex
Remove 'new' tags, fix #122

@@
     effect when the twin alignment is allowed to have a different
-    number of \new{substitution}s
+    number of substitutions
   }
@@
   \item{
-    When we do not impose the same number of \new{substitution}s 
+    When we do not impose the same number of substitutions 
     between true and twin alignment, we observe a difference in the error 
     distributions with respect to the standard case (presented in the main 
-    text) where they are forced to have the same number of \new{substitution}s.
+    text) where they are forced to have the same number of substitutions.
   }
@@
 
-\new{
 We chose to preserve the original topology to 
@@
 to trees with the same branching times but different topologies.
-}
-\new{
 The default option for the twin diversification model $p_T$ 
 is the standard BD model.
-}
-\verb;pirouette;
-\new{
-   has a built-in function to use a Yule model as well.
+\verb;pirouette; has a built-in function to use a Yule model as well.
 Additionally, a user can specify a function to generate a twin tree
 from any speciation model, such as, for example, a coalescent model.
-}
 
@@
 package [\cite{TESS}].
-\new{
 For simplicity, when simulating phylogenies we assumed a sampling fraction 
 of $100\%$. A different choice might have an effect on model performance.
-}
 
@@
 not only the same mutation rate is used, but also the total number of 
-\new{substitution}s
-matches the true alignment. The total number of \new{substitution}s is defined
+substitutions
+matches the true alignment. The total number of substitutions is defined
 as the number of different nucleotides between the (known) root sequence
@@
 For this reason multiple phylogenies must be 
-considered \new{(at least 100 independent true and twin trees)}.
+considered (at least 100 independent true and twin trees).
 If the number of considered phylogenies is high enough, 
@@
 from a distribution of trees from any tree model.
-\new{
 The nLTT statistic is not suitable, however, to distinguish
@@
 models [\cite{louca2020extant}].
-}
 
@@
 would be equal to the number of tips, which amounts to $\lambda_0 = 0.63$.
-\new{
 Note that in the main example, a tree was generated with 5 tips, due
 to stochasticity in the tree generation algorithm.
-}
 
@@
 an RLN clock model and a Yule tree model 
-\new{
 (Note that this is different from how the twin tree was generated which was with a BD process and the alignment
-  was simulated with a JC substitution model and strict clock model
-  ). However, the extinction rate used in simulating the twin tree was practically 0, thus resembling a Yule process.
-}
+was simulated with a JC substitution model and strict clock model). 
+However, the extinction rate used in simulating the twin tree was practically 0, thus resembling a Yule process.
 From the 'twin alignment' this best candidate inference model is run.
@@
 
-\new{
 The resulting error distributions are shown in Fig. \ref{fig:replicate_trees}. 
@@
 the original trees.
-}
 
@@
 
-\new{
 We show in figures \ref{fig:replicate_trees}, \ref{fig:example_12_taxa}, 
@@
 The data point for $40$ taxa not following the trend could be due to the limited amount of simulated trees taken in consideration due to time constraints.
-}
 
@@
 
-\new{
 From figures \ref{fig:example_500_nucleotides}, 
@@
 more information becomes available.
-}
 
@@
 
-\new{
 As expected the twin and true distributions in Fig. \ref{fig:example_yule} 
 are extremely similar for both the generative and the best candidate case.
-}
 
@@
 
-\new{
 Because the two models are very similar to each other (the BD model can be 
@@
 the distributions here exhibit a greater difference.
-}
-\new{As we use only extant trees, it is reasonable that the method
+
+As we use only extant trees, it is reasonable that the method
 is slightly weaker in distinguishing between the Yule and BD
 models. It is unknown what the discriminatory power would be
-  when comparing trees with extinction events.}
+when comparing trees with extinction events.
 
@@
 
-\new{
 Here the median errors are similar in the three settings and similar to 
@@
 The entire run took 120 hours in total.
-}
 
@@
 The main example uses a twin alignment that has the same number
-of \new{substitution}s (as measured from the ancestral sequence) as the true alignment. Here, we show the same results, with the difference that
+of substitutions (as measured from the ancestral sequence) as the true alignment. Here, we show the same results, with the difference that
 the twin alignment uses the same mutation rate, yet is not guaranteed
-to have the same number of \new{substitution}s.
+to have the same number of substitutions.
 
@@
     similar to Fig. \ref{fig:replicate_trees}, 
-    but here the number of \new{substitution}s is not imposed 
+    but here the number of substitutions is not imposed 
     to be the same between true and twin alignment. 
@@
 
-\new{
 Comparing figures \ref{fig:example_random_mutations} 
@@
 alignments and this is reflected in the error distributions.
-}
 
@@
 Here, we show the same results for different mutation rates.
-\new{
-\iffalse
-  We can observe that the error tends to increases for mutation rates above
-  `1.0 / crown age`. Below this value, the errors are the same.
-  For these lower mutation rates, however, the shape of the error
-  distribution changes for the baseline/generative models,
-  deviating more and more from the gamma-distribution-like shape
-  as is still present in the twin distributions.
-\fi
  The results for the different mutation rates are shown in Figs.
@@
  \ref{fig:example_2.00_mutation_rate} (2.00 / crown age).
- }
-\new{
 Fig.
@@
 It is possible, however, that a more regular trend could be observed increasing the number of simulations.
-}
 

### 2020-08-07 pirouette_supplement.tex
Add DOIs to code and data

@@
 
-All code is archived at 
-\url{http://github.com/richelbilderbeek/pirouette_article},
-with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
+The pirouette code used for the examples 
+is archived at DOI \url{https://doi.org/10.5281/zenodo.3969839}.
+The pirouette examples (including intermediate data) are archived at 
+DOI \url{https://doi.org/10.5281/zenodo.3970000}.
 

### 2020-10-09 pirouette_supplement.tex
Transferred richelbilderbeek/mcbette to ropensci/mcbette

@@
 remotes::install_github(
-  "richelbilderbeek/mcbette"
+  "ropensci/mcbette"
 )
